### 2024-09-13 papers/article/active-travel-behaviour-evolution-in-Canada/active-travel-behaviour-evolution-in-Canada.Rmd
Starting paper

---
title: A Historical Analysis of the Evolution of Active Travel Behaviour in Canada
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    correspondingauthor: true
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    organization: Big Wig University
    addressline: 1 main street
    city: Gotham
    state: State
    postcode: 123456
    country: United States
  - code: Another University
    organization: Department
    addressline: A street 29
    postcode: 2054 NX
    city: Manchester,
    country: The Netherlands
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  This is the abstract.

  It consists of two paragraphs.
keywords: 
  - keyword1
  - keyword2
journal: "An awesome journal"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: mybibfile.bib
biblio-style: elsarticle-harv # author year style for natbib - use 'elsarticle-num' or 'elsarticle-num-names' for numbered scheme
classoption: preprint, 3p, authoryear # remove authoryear is not using `elsarticle-harv`
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

```{r load-pkgs, include = FALSE}
# List of packages required for this analysis
#pkg <- c("dplyr", "ggplot2", "knitr", "bookdown", "fitdistrplus", "tidyverse", "readxl", "splitstackshape","tibble", "readr", "tabulate", "janitor", "flextable", "tidyr", "gt", "xtable", "kableExtra")

# Load packages
library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
library(dplyr) # A Grammar of Data Manipulation
library(here) # A Simpler Way to Find Your Files
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(fitdistrplus) # Help to Fit of a Parametric Distribution to Non-Censored or Censored Data
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(readxl) # Read Excel Files 
library(splitstackshape) # Stack and Reshape Datasets After Splitting Concatenated Values
library(tibble) # Simple Data Frames
library(readr) # Read Rectangular Text Data
library(tabulate) # Pretty Console Output for Tables
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data
library(flextable) # Functions for Tabular Reporting
library(tidyr) # Tidy Messy Data
library(xtable) # Export Tables to LaTeX or HTML
library(extrafont) # Tools for Using Fonts
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
library(reshape2) # Flexibly Reshape Data: A Reboot of the Reshape Package
library(RColorBrewer) # ColorBrewer Palettes
```

```{r load-data, include = FALSE}
load(paste0(here(), "/data/walking_1986.Rda"))
load(paste0(here(), "/data/walking_1992.Rda"))
load(paste0(here(), "/data/walking_1998.Rda"))
load(paste0(here(), "/data/walking_2005.Rda"))
load(paste0(here(), "/data/walking_2010.Rda"))
load(paste0(here(), "/data/walking_2015.Rda"))
load(paste0(here(), "/data/cycling_1992.Rda"))
load(paste0(here(), "/data/cycling_1998.Rda"))
load(paste0(here(), "/data/cycling_2005.Rda"))
load(paste0(here(), "/data/cycling_2010.Rda"))
load(paste0(here(), "/data/cycling_2015.Rda"))
load(paste0(here(), "/data/cycling_data.Rda"))
load(paste0(here(), "/data/cycling_data_f.Rda"))
load(paste0(here(), "/data/walking_data.Rda"))
load(paste0(here(), "/data/walking_data_f.Rda"))
load(paste0(here(), "/data/walking_1986_f.Rda"))
load(paste0(here(), "/data/walking_1992_f.Rda"))
load(paste0(here(), "/data/walking_1998_f.Rda"))
load(paste0(here(), "/data/walking_2005_f.Rda"))
load(paste0(here(), "/data/walking_2010_f.Rda"))
load(paste0(here(), "/data/walking_2015_f.Rda"))
load(paste0(here(), "/data/cycling_1992_f.Rda"))
load(paste0(here(), "/data/cycling_1998_f.Rda"))
load(paste0(here(), "/data/cycling_2005_f.Rda"))
load(paste0(here(), "/data/cycling_2010_f.Rda"))
load(paste0(here(), "/data/cycling_2015_f.Rda"))
load(paste0(here(), "/data/trip.Rda"))
```

## Introduction

The idea that travel behaviour can be influenced by the city form has raising an increase interest in urban and transportation planning. Cities intend to encourage their residents to adopt more sustainable modes of transportation, including walking, cycling, and public transit, by developing an environment with a variety of transportation alternatives and, at the same time, increasing accessibility - understood as the ease of reaching destinations and opportunities [@iacono2008access]. Because of their significant role in enhancing and promoting urban sustainability [@hino2014built; @lamiquiz2015effects], active transportation modes, such as walking and cycling, are playing a central role in urban mobility research and policy-making [@handy1993regional; @clifton2001evaluating; @frank2001built; @krizek2005perspectives; @sallis2004active; @vandenbulcke2009mapping; @wu2019measuring]. Walking and cycling accessibility are closely related, contributing together to the concept of "active accessibility" or "non-motorized accessibility", and, when considered in the urban and transportation planning process, reducing dependence on private vehicles and promoting healthier and more sustainable travel behaviour among residents. 
There are two main components when measuring accessibility: (1) the location and power of attraction of urban opportunities (trip benefit) and (2) the barrier in travel from the origin in the network to the destination (trip cost). A way for measuring the cost of travel when calculating accessibility is using impedance functions, a methods that is receiving attention from transportation planning scholars, urban geography, and sustainable development [@frank2005linking; @krizek2005perspectives; @currie2010quantifying; @iacono2010; @yang2012walking; @millward2013active; @nassir2016utility; @saghapour2017measuring; @wu2019measuring]. The impedance functions have different forms and all of them serves a tool to understand the travel behaviour and as measure of the willingness to travel a certain distance to achieve a desired destination, where a service or an opportunity is located [@taylor1975distance; @fotheringham1981spatial; @kwan1998space; @eldridge1991warped; @luoma1993threshold; @papa2012gravity; @yang2012walking; @millward2013active; @vale2017influence]. In this concept, areas with higher accessibility are those characterized by a lower impedance when traveling to desirable destinations. When talking about active accessibility, increasing the distance between two points generally implies in a decreasing in the probability of that trip being done by walking or biking [@hansen1959accessibility; @pirie1979measuring; @handy1997measuring; @geurs2001accessibility; @bhat2002development; @church2003measuring; @kwan2003recent; @geurs2004; @levinson2005access; @cascetta2013new]. However, more information about the willingness of some individuals to walk or cycle greater distance is needed, as well as more data on how distance affects the type and feasibility of the activity, destinations desirability, and the characteristics of those embarking on the trip in different contexts. Because of this, investigate the evolution and dynamics of impedance function over time becomes important, since they're easily impacted by changes in the transportation network or in urban spatial configurations [@iacono2008access; @iacono2010]. Luoma, Mikkonen, and Palomaki (1993) evidenced a decreasing in the distance decay parameter over time, attributing this trend to improvements and maturation of the transportation system [@luoma1993threshold], and  Mikkonen and Luoma (1999) elucidated the factors behind these temporal patterns in the gravity models patterns [@mikkonen1999parameters].
A range of impedance functions have been applied to describe the distribution of walking and cycling trips, wether for general or specific purposes [@iacono2008access; @iacono2010; @larsen2010beyond; @yang2012walking; @millward2013active; @vale2017influence; @li2020approach]. 
Selecting an appropriate impedance function can be challenging and results in a diverse range of cost decay functions that are employed as impedance functions in accessibility measures, including **threshold functions** (e.g., binary Step Function and multiple Step Function) and **smooth cost decay functions** (e.g., inverse-potential, log-normal, logistic, exponential square-root, and half-life function) [@de2009exponential; @reggiani2011accessibility; @osth2016new; @itf2017linking]. The variety of functions relies in how scholars approach the influence of distance, with negative exponential distance-decay functions are commonly used in assessing non-motorized accessibility, capturing the willingness of individuals to walk or cycle to destinations [@handy1997measuring; @geurs2001accessibility; @iacono2010; @vega2012using; @millward2013active; @vale2017influence; @li2020approach].

The merit of this function is due to its ability to assign decreasing influences to more remote opportunities, giving a more accurate estimate for shorter journeys [@iacono2010; @kanafani1983transportation; @fotheringham1989spatial]. However, in addition to determine the form of the impedance function, scholars also need to specify the variable used to measure impedance, which can be either time, distance, monetary cost, a combination these last variables or even a generalized cost concept.

The choice between time and distance as the impedance variable has been found to be acceptable based on previous research [@iacono2010; @hull2012accessibility; @sun2012measuring; @lowry2012using; @vasconcelos2012evaluation ]. However, when it comes to non-motorized travel modes, extracting accurate travel times from existing network models can be challenging, which limits the options and makes distance a more practical choice [@handy1997measuring; @iacono2010; @yang2012walking; @arranz2019measuring]. Furthermore, researchers specializing in active modes of transportation have faced challenges stemming from a need for more objective data concerning walking and cycling behavior. Estimating specialized impedance functions specific to non-motorized modes requires appropriate travel survey data that can capture pedestrian and cycle behavior. Often, researchers have resorted to relying on retrospective questionnaires, which assess subjective aspects such as the frequency and duration of walking and cycling activities. Notably, regional household travel surveys, including trips made by non-motorized modes, have been employed for this purpose [@iacono2010; @millward2013active].

In contrast to these localized surveys, some datasets provide a nationwide perspective, encompassing travel for various trip purposes and offering insights into details like travel episode origins, destinations, and time-based lengths. This comprehensive approach furnishes a more holistic understanding of active transportation behavior. Nevertheless, only some studies, such as Yang et al. (2012), have examined active travel behavior nationally. Addressing the previously identified gaps and precisely the challenge of using uniform decay curves, such as the negative exponential function for both cycling and walking in accessibility assessments, this study endeavors to identify the most appropriate distance decay function for various destinations and time periods.  Our research delves into the complexities of actual travel behavior, emphasizing active transportation modes. We utilize historical data from the General Social Survey (GSS) from 1986 to 2015 in Canada to calculate the impedance function for cycling and walking trips. By doing so, this study conducts a comparative analysis of travel behaviors associated with these two modes.

Moreover, we recognize that non-work travel encompasses a spectrum of trip intentions and diverse traveler behaviors, making the impedance function an essential analytical tool for examining non-work accessibility. Grengs (2015) emphasizes the importance of crafting distinct functions for each travel purpose, a principle that guides our analysis [@grengs2015nonwork]. Our investigation will encompass a variety of trip intents, ranging from commutes to homes, workplaces, or educational institutions to social visits, outdoor activities, business trips, shopping, cultural outings to libraries, museums, or theaters, dining out, and engaging in religious practices. By drawing on a nationally representative sample of Canadian residents, our research intends to bridge the empirical data gap concerning the frequency and duration of typical pedestrian and cycling trips for different purposes. In doing so, it aspires to provide a nuanced understanding of active travel behavior. Ultimately, this comprehensive analysis seeks to contribute meaningfully to the ongoing conversation on active transportation, shedding light on its role in shaping travel behavior and accessibility.

## Background

Accessibility is conceptualized as the potential to access geographically dispersed opportunities, considering the challenges associated with reaching them [@paez2012measuring]. The positioning of resources concerning users, the transportation infrastructure, and how spatial relations and distances impact the potential utilization of amenities are fundamental considerations in accessibility and mobility modeling. Typically, the effect of distance on potential use is expressed by "Impedance functions" or "Distance Decay functions" [@hansen1959accessibility; @koenig1980indicators; @fotheringham1981spatial]. These functions are integral in transportation planning, commonly incorporated into forecasting models to interpret urban travel behaviors for each mode. They are typically shaped from estimates grounded in sample data distributions that mirror fluctuations in individuals' propensity to travel different distances to reach opportunities. This importance is underscored by their prevalent use in understanding accessibility to specific locations and areas covered by different services [@hsiao1997use; @zhao2003forecasting; @iacono2010; @li2020approach]. Fundamentally, accessing opportunities is tied to the travel costs to a destination. Indeed, the main goal of the impedance function is to depict the diminishing intensity of interaction as the separation between locations augments. These functions delineate how an increase in distance or associated travel costs inversely affects potential usage; in essence, distant facilities are less likely to be used compared to those in closer proximity[@hansen1959accessibility; @koenig1980indicators; @fotheringham1981spatial; @skov2001estimation]. In addition, the impedance function captures people's tendency to reduce the number of trips they take exponentially as the distance increases. The ''distance-decay'' effect suggests that adding one unit of distance to a long trip is less significant than adding one unit to a shorter trip [@carrothers1956historical]. The essence of the impedance function, as highlighted by Carrothers (1956), suggests a nuanced understanding of travel behavior, particularly in how distance or time additions impact trips of varying lengths. This concept is further elaborated upon by the idea that the perceived friction or impedance per unit of distance does not remain constant but diminishes with increasing distance. Essentially, any additional distance or time incurred is felt more acutely for shorter trips, creating a higher friction or impedance against movement. This phenomenon is due to the disproportionately greater friction per unit of distance for shorter distances compared to longer ones. Conversely, when considering longer movements, an additional unit of distance or time becomes relatively less significant. For example, in an urban context, the friction against movement within a densely populated area for a given distance is generally higher than the same distance in a less densely developed space. This implies that the effect of an extra mile or minute is more substantial when added to a shorter trip than to a longer one. Hence, the impedance function reflects this diminishing sensitivity to distance or time increments as the length of the trip increases, aligning with the principle that the impact of additional distance or time diminishes as the total trip length extends.

Examining impedance functions concerning different transportation modes and destinations is a valuable foundation for comprehending the travel behavior attributed to each mode. The parameters of impedance functions offer insights into the spatial coverage provided by each mode of transportation. By segmenting modal trips based on their purposes, comparing the distribution of trips between various purposes for each transportation mode (work-related and non-work purposes) becomes possible. Empirically derived impedance functions offer valuable evidence that can be utilized to examine and substantiate various claims about travel behavior, thus supporting urban planning endeavors. For instance, the current interest in creating "livable" communities revolves around loosely held assumptions regarding individuals' willingness to walk and cycle to different destinations. A common belief is that people are generally willing to walk up to a quarter of a mile to access most locations [@untermann1984accommodating]. However, there remains limited information regarding whether specific individuals are open to walking or cycling longer distances and, if so, how much farther they are willing to travel. Moreover, there needs to be more evidence concerning the influence of trip characteristics, destination attractiveness, and individual characteristics on the impact of distance on walking and cycling behaviors[@geurs2006accessibility].

Since Hansen's foundational research, various categories of accessibility measures have been developed, such as active-based, infrastructure-based, individual-based, and utility-based indicators [@hansen1959accessibility; @geurs2004]. Accessibility metrics, particularly those of the gravity type or potential measures, have been extensively utilized in active modes [@miller2005place]. These measures primarily stem from the gravity model, where an impedance function weighs opportunities. In fact, these are designed by weighting opportunities within an area according to an attraction measure and then diminishing each based on an impedance measure [for example, @geertman1995gis; @handy1993regional]. Equation \ref{eq:ch03-equation-01} expresses the general representation of the accessibility equation:

\begin{equation}
A_{ik}^{pt} = \sum_{j=1}^J g(O_j^t) f(C_{ij}^p)
\label{eq:ch03-equation-01}
\end{equation}

Research demonstrates that two primary types of accessibility indicators are predominantly used in studies. According to equation \ref{eq:ch03-equation-01}, the first one revolves around **Opportunities**, which are weighted by an impedance (characterized by a relevant decreasing function of travel cost or time required to access these opportunities. The accessibility measure $A^ {pt}_{ik}$ signifies accessibility from an origin i at a specific time t to a particular destination type k tailored for an individual of type p. The function $g(O^t_{jk})$ quantifies the appeal of opportunities categorized under type k at destination j, accessible at time t. Simultaneously, $f(C^p_{ij})$ represents the impedance while traveling from origin i to destination j for a person belonging to type p. It's worth noting that the functional form $f()$ delineates an impedance decay function. Meanwhile, $C^p_{ij}$ signifies the generalized travel cost, potentially encompassing factors such as time, distance, and exertion.

Within the gravity model, the second variable concerns the cost associated with the spatial gap between a trip's start and end points (origins and destinations). Moreover, the third pivotal variable, crucial for constructing this model, involves the mathematical formulation delineating the travel impedance between origins and destinations. This "cost" can be articulated in terms of physical distance, travel duration, financial outlay, or a fusion of these elements. Among these, travel duration emerges as the chief cost measure and is the metric chosen for this particular investigation[@handy1993regional; @fotheringham1989spatial; @grengs2004measuring; @hess2005access]. In fact, when selecting a format for the impedance function, the researcher must determine which variable (time, cost, or both) will measure impedance. Historically, both these measures have been employed, with specific instances incorporating the concept of generalized cost [@handy1997measuring]. For non-motorized travel modes, however, the prevalent choice is distance, attributed to the challenges in obtaining precise travel times from existing network models designed for walking and bicycling. Previous studies indicate that adopting either time or distance as the impedance variable is appropriate [@handy1997measuring]. While the first two variables are derived from the attributes of the built environment, the impedance parameter captures aspects of human behavior, making its determination an intricate undertaking.

The second type pertains to the **Cumulative-opportunity** metrics, often referred to as **isochronic indices**. These evaluate accessibility by determining the number or proportion of opportunities available within a specified travel duration or distance from a reference point. These metrics avoid making presumptions about travel decay. They utilize a rectangular function, categorizing travel as "acceptable" within certain thresholds and "unacceptable" beyond them. One of the main complexities with these metrics is deciding on the appropriate cutoff point. This decision can be based on prevailing mobility patterns of the population, or it can mirror established norms, conventions, or the researcher's informed projections [@vickerman1974accessibility]. This metric can be interpreted as a specialized solution of Equation \ref{eq:ch03-equation-02} In this context, the impedance function is designated as 1 when $C_{ij}< x$ and 0 when $C_{ij} > x$.

\begin{equation}
C_{ij} =
\begin{cases}
  1 & \text{if } c_{ij} \le x \\
  0 & \text{if } c_{ij} > x
\end{cases}
\label{eq:ch03-equation-02}
\end{equation}

The study of travel behaviors has resulted in the development of various mathematical functions that have been carefully developed by comparing observed walking and cycling trips with predefined mathematical distributions over time. The choice of the impedance function is critical, as it is deeply intertwined with the gravitational force of travel, subsequently influencing accessibility evaluation outcomes [@breheny1978measurement; @kwan1998space; @talen1998assessing]. It should be noted that the domain of transportation research has seen an increased focus on various formulations related to accessibility. Pre-eminent research in this domain has been presented by the likes of Song (1996), Handy and Niemeier (1997), Handy and Clifton (2001), and Iacono et al. (2010)[@song1996some; @handy1997measuring; @clifton2001evaluating; @iacono2010]. While the majority of these studies focus on automobile transportation, Iacono et al. (2010), Millward et al. (2013), and Vale et al. (2017) investigate active modes of transportation, such as walking and cycling. Their findings show that the attractiveness of destinations declines significantly beyond a mile.

The choice of the most appropriate impedance function and the associated parameters can significantly vary, especially when considering different modes of travel, particularly within the realm of active transportation, as well as the specific purpose of the journey [@iacono2008access; @iacono2010; @larsen2010beyond; @millward2013active]. In the domain of impedance function modeling for travel behavior analysis, both a simple inverse power function and a negative exponential function, typically based on distance or travel time, are frequently employed. However, these conventional models have been criticized for their rapid attenuation near the origin, as highlighted in Ingram's seminal work in 1971. Ingram pointed out that these traditional functions tend to diminish more quickly than empirical data would suggest, particularly in proximity to the starting point of travel. To address this limitation, he proposed the use of a modified Gaussian function. This alternative model is characterized by a slower decline rate near the origin and a less steep fall-off at greater distances, providing a more realistic representation of travel behavior over varying distances. The modified Gaussian function holds distinct advantages over its counterparts, including a slower rate of decline in proximity to the origin and a more gradual approach to zero as distances increase, setting it apart from the rapid decay exhibited by the negative exponential and inverse power functions[@ingram1971concept]. Despite the theoretical promise of the Gaussian function, as proposed by Ingram (1971), our comprehensive review of the existing literature did not reveal any specific research applications of this function to assess accessibility within the context of active transportation. To our knowledge, only Kwan (1998) stands out as the sole researcher to have employed a Gaussian function to evaluate automobile accessibility [@kwan1998space; @vale2017influence]. This absence of utilization in the active transportation context underscores a notable research gap, leaving the opportunity for future investigations to explore the potential advantages and implications of adopting the modified Gaussian function for assessing accessibility in walking and cycling.

Researchers, from Stewart (1941) to Haggett (2001), have reached a consensus regarding the non-linear nature of inverse distance-decay functions in human geographical applications [@stewart1941inverse; @taylor1971distance; @taylor1983distance; @johnston1973frictions;@robinsonm; @haggett2001geography]. At the same time, simplistic functions may not aptly reflect the observed data on distance-interaction intensity graphs, and intricate distance-decay functions have been promoted by Taylor (1983) and Robinson (1998). Among these functions, bell-shaped curves, such as the Tanner function, March's function, and the Box-Cox function, have gained prominence, directed by multiple parameters [@richardson1969elements;@gaudry1981inverse; @taylor1983distance; @mandel1997disaggregate; @robinsonm; @tiefelsdorf2003misspecifications;@de2011modelling; @paez2012measuring; @halas2014distance; @martinez2013new; @van2001accessibility]. Mozolin et al. (2000) introduced an approach that integrates distance and employment numbers, visualized in a 3D diagrammatic representation [@mozolin2000trip]. However, while these functions provide precise fits for specific scenarios, their general application across varying contexts may be limited.

As mentioned, the choice of impedance functions plays a pivotal role in modeling accessibility, particularly when analyzing travel behavior within urban environments. Among the various mathematical forms that can represent impedance functions, the negative exponential function has emerged as the dominant choice in traditional transportation planning models and recent studies [@meyer1984urban; @gutierrez1996european; @kwan1998space; @apparicio2008comparing; @iacono2008access; @iacono2010; @larsen2010beyond; @millward2013active]. Its prevalence in the literature can be attributed to several factors, including its compatibility with travel behavior theory and its ability to effectively estimate shorter trips, particularly those associated with non-motorized modes of transportation. As Song (1996) noted in his examination of alternative accessibility measures, the negative exponential form $(e ^ {-\beta x})$ stands out as the most useful in explaining population distribution due to its gradual decline, which aligns with empirical data and accurately captures the influence of proximity on accessibility [@song1996some]. This choice of impedance function aligns with the rationale that a more gradual decline near the origin is essential for estimating shorter trips, a characteristic highly relevant to active modes of transportation, such as walking and cycling. This preference for the negative exponential form is further corroborated by the work of Kanafani (1983), who highlighted its suitability for modeling non-motorized modes, emphasizing its capacity to better estimate shorter trips compared to the power function. The ubiquity of the negative exponential form in recent studies and its adaptability to various urban contexts make it a well-suited candidate for estimating impedance functions, as indicated by a record of numerous empirical applications [@kanafani1983transportation; @fotheringham1989spatial; @de2009exponential; @iacono2010; @signorino2011gravity; @prins2014many].

Furthermore, recent investigations into impedance functions have shed light on their effectiveness in modeling walking accessibility, with particular attention to the negative exponential form. In the study conducted by Vale and Pereira (2017), which examined 20 pedestrian accessibility measures, both the modified Gaussian and exponential functions were found to be robust choices for accurately modeling walking accessibility. These findings corroborate the suitability of the negative exponential form for capturing the nuanced patterns of walking behavior within urban areas. Vale and Pereira's research introduced a novel cumulative Gaussian function that considers cumulative opportunities at close distances (e.g., 200 or 400 meters) and a modified Gaussian curve for longer distances. These functions showcase the adaptability and versatility of impedance modeling, especially in the context of active travel and pedestrian accessibility analysis [@vale2017influence]. Thus, while alternative impedance functions may be considered in specific scenarios, the dominance of the negative exponential form underscores its utility as a fundamental choice for estimating impedance functions, mainly when dealing with active modes of transportation and urban accessibility modeling. We present a detailed exposition of the most widely recognized impedance functions to further elaborate on this aspect. These include the inverse power, negative exponential, modified Gaussian, logarithmic normal distribution, cumulative opportunities (both rectangular and linear), gamma, and Log-logistic decay functions. Each model offers a unique lens through which the relationship between distance or travel time and accessibility can be understood, thereby providing a comprehensive toolkit for urban planners and researchers to analyze and predict travel patterns in various contexts.

- **Inverse Power equation**

The Inverse Power function is defined as (\ref{eq:ch03-equation-03}):

\begin{equation}
f(t_{ij}) =
\begin{cases}
  1 & \text{if } t_{ij} \le 1 \\
  t_{ij}^{-\beta} & \text{otherwise}
\end{cases}
\label{eq:ch03-equation-03}
\end{equation}

This function posits that impedance is inversely proportional to travel time $(t_{ij})$ (represents the travel time between location i (origin) and location j (destination))raised to the power of $\beta$, a parameter to be estimated. For travel times less than or equal to one unit, the impedance is set to 1, reflecting a baseline accessibility. As the travel time increases beyond one unit, the function decays according to the power of $\beta$, indicating a diminishing probability of a traveler choosing a destination as the distance increases.

- **Negative Exponential equation**

The Negative Exponential function (\ref{eq:ch03-equation-04}) is given by:

\begin{equation}
f(t_{ij}) = e^{-\beta t_{ij}}
\label{eq:ch03-equation-04}
\end{equation}

This model suggests that impedance decreases exponentially with increasing travel time $(t_{ij})$. The parameter $\beta$ represents the decay rate, with higher values indicating a faster decrease in accessibility with increasing travel time. This function is widely used due to its simplicity and ability to model the rapid drop-off in accessibility over distance.

- **Modified Gaussian equation**

The Modified Gaussian function is expressed as (\ref{eq:ch03-equation-05}):

\begin{equation}
f(t_{ij}) = e^{-\frac{t_{ij}^2}{2}}
\label{eq:ch03-equation-05}
\end{equation}

This function follows a Gaussian distribution pattern, implying that the impedance decreases in a bell-shaped curve as travel time increases, with a slower decline near the origin and a less steep drop at larger distances. It captures the gradual reduction in accessibility with distance more realistically than the exponential or power functions.

- **Logarithmic normal distribution function equation**

Presented as Equation \ref{eq:ch03-equation-06} [@wu2019measuring]:

\begin{equation}
f(c_{ij}) = \alpha \frac{1}{\sqrt{2\pi} x \sigma} e^{-\frac{(\ln t - \mu)^2}{2\sigma^2}}
\label{eq:ch03-equation-06}
\end{equation}

Here, the variable $x$ represents the distance traveled in active modes. The parameters, represented as $\alpha$, $\mu$, and $\sigma$, hold significant importance in this analysis as they undergo estimation. Together, these parameters exert control over the shape of the curve under investigation. The function accounts for impedance distribution across different travel times in a log-normal format, offering a more nuanced view of accessibility. [@wu2019measuring].

- **Cumulative Opportunities Rectangular Function**

Defined as Equation \ref{eq:ch03-equation-07}:

\begin{equation}
f(t_{ij}) =
\begin{cases}
  1 & \text{if } t_{ij} \le t \\
  0 & \text{otherwise}
\end{cases}
\label{eq:ch03-equation-07}
\end{equation}

This function assumes a binary approach to accessibility. For travel times $(t_{ij})$ below a certain threshold $t$, full accessibility (value 1) is granted. Beyond this threshold, accessibility rapidly drops to zero, representing a stark division between accessible and inaccessible areas.

- **Cumulative Opportunities Linear**

The equation of cumulative opportunities linear is as follows (Equation \ref{eq:ch03-equation-08}):

\begin{equation}
f(t_{ij}) =
\begin{cases}
  1 - \frac{t_{ij}}{\bar{t}} & \text{if } t_{ij} \le \bar{t} \\
  0 & \text{otherwise}
\end{cases}
\label{eq:ch03-equation-08}
\end{equation}

This linear function provides a gradual decrease in accessibility with an increase in travel time up to a particular average time $\bar{t}$. Beyond this average time, accessibility is considered zero. This model is beneficial for representing a more gradual decay in accessibility than the abrupt drop-off in the rectangular model.

- **Gamma distribution**

The formula of this function is presented below  (Equation \ref{eq:ch03-equation-09})

\begin{equation}
f(c_{ij}) = 
   \begin{cases}
\frac{1}{\sigma^\alpha\Gamma(\alpha)} c_{ij}^{\alpha-1} \cdot e^{-\frac{c_{ij}}{\sigma}} & \text{if } 0 \leq c_{ij} <      \infty  \text{ and } \alpha, \sigma > 0 \\
0 & \text{otherwise}
   \end{cases}
\label{eq:ch03-equation-09}
\end{equation}

The Gamma function introduces a flexible model, parameterized by $\alpha$ and $\sigma$, to represent impedance. This function can adjust its shape significantly based on the parameter values, offering a versatile tool for modeling various types of decay patterns in travel behavior.

- **Log-logistic decay functions**

The log-logistic function was displayed in Equation \ref{eq:ch03-equation-10}.

\begin{equation}
f(c_{ij}) = \frac {1}{1+ \exp(a + b\ln c_{ij})}
\label{eq:ch03-equation-10}
\end{equation}

In this context, the variable $c_{ij}$ denotes the travel time between locations $i$ and $j$, while parameters $a$ and $b$ are the subject of estimation. In this log-logistic model, the impedance is a function of the natural logarithm of travel time $\ln c_{ij}$. Parameters $a$and $b$ define the steepness of the decay curve. This function is helpful for scenarios where a logistic decay pattern is more representative of the travel behavior being modeled [@thorsen1999network].

- **Beta distribution**

The Beta distribution has probability density function (\ref{eq:ch03-equation-11})

\begin{equation}
f(c_{ij}) = 
\begin{cases}
\frac{1}{B(\alpha, \beta)} c_{ij}^{\alpha - 1}(1 - c_{ij})^{\beta - 1} & \text{if } 0 \leq c_{ij} < \infty \\
0 & \text{otherwise}
\end{cases}
\label{eq:ch03-equation-11}
\end{equation}

The Equation \ref{eq:ch03-equation-11} represents the probability density function of the Beta distribution, a continuous probability distribution with two shape parameters. In this formula, $c_{ij}$is a specific impedance or cost between a point $i$ and point $j$ which is between 0 and 1, inclusive. It represents the possible outcomes of the Beta distribution, where $c_{ij}$ is the probability of an event occurring. $\alpha$ is the first shape parameter, and $\beta$ is the second shape parameter. These parameters control the shape of the Beta distribution, where both are positive and more than 0. They can be any positive real numbers, and they determine the behavior of the probability density function, dictating where most of the distribution's "weight" is. The term $B(\alpha, \beta)$ in the denominator is the Beta function, which is a normalization constant to ensure that the total probability integrates to 1 over the interval from 0 to 1. The Beta function is related to the Gamma function and is defined for positive values of $\alpha$ and $\beta$. It ensures that the probability density function is scaled correctly. This formula gives us the likelihood of $c_{ij}$ when the distribution is defined by particular $\alpha$ and $\beta$ parameters.

- **Poisson distribution**

\begin{equation}
P(c_{ij} = k) = \frac{\lambda^k e^{-\lambda}}{k!}, \quad \text{for } k = 0, 1, 2, \ldots
\label{eq:ch03-equation-12}
\end{equation}

The Poisson distribution (\ref{eq:ch03-equation-12}) is a discrete probability distribution that expresses the likelihood of a given number of events occurring within a fixed interval of time or space, assuming these events occur with a known constant mean rate and independently of the time since the last event. It is named after the French mathematician Siméon Denis Poisson. In Equation \ref{eq:ch03-equation-12}, $P(c_{ij} =k)$ represents the probability that the impedance between point $i$ and point $j$ is exactly $k$. The impedance could refer to various factors such as travel time, cost, or distance that might impede the movement or flow between these two points.$\lambda$ is the average rate of impedance observed in the system or estimated from data. It represents the expected number of impedances that occur over a given interval. In the context of accessibility, it could be the average number of barriers or frictions encountered when traveling between different locations. $k$ is the actual observed number of impedances; for example, this could be the actual number of barriers experienced on a particular route. The use of the Poisson distribution to model impedance is based on the assumption that the events (or impedances) occur randomly and independently within a given spatial or temporal frame of reference.

-  **Weibull distribution**

The Weibull distribution with shape $\alpha$ and scale $\theta$ has a probability density function as follows:

\begin{equation}
f(c_{ij}) =
\begin{cases}
\left( \frac{\alpha}{\theta}\right) \left( \frac{c_{ij}}{\theta}\right)^{\alpha - 1} e^{-\left(\frac{c_{ij}}{\theta}\right)^\alpha} & \text{if } 0 \leq c_{ij} \\
0 & \text{otherwise}
\end{cases}
\label{eq:ch03-equation-13}
\end{equation}

The Weibull distribution (Equation \ref{eq:ch03-equation-13}) is a continuous probability distribution used to model various data types, particularly where the data shows a changing failure rate over time. This flexibility makes the Weibull distribution especially suited to reliability analysis, life data analysis, and modeling failure times. $f(c_{ij})$ represents the probability density function (pdf) of the Weibull distribution, with $c_{ij}$ being the variable of interest. here, $c_{ij}$ would represent the impedance value, such as time or cost associated with traveling between two points. $\alpha$ is the shape parameter of the Weibull distribution. It determines the form of the probability distribution function and can indicate whether the failure rate is increasing, constant, or decreasing over time. In fact,it would shape the distribution of impedance experiences, such as whether impedances are more common as trips get longer or shorter. $\theta$ is the scale parameter of the Weibull distribution. It stretches or compresses the distribution along the horizontal axis and is akin to the 'characteristic life' in reliability terminology.

Based on the above description, the inverse power, negative exponential, and modified Gaussian functions consistently diminish the influence of opportunities as travel time increases, governed by an impedance parameter $\beta$ that incorporates travel cost considerations. Originating from early gravity models of spatial interaction [@stewart1948demographic; @zipf1949human], the inverse power function exhibits a rapid decrease in opportunity influence as travel time lengthens. While power functions draw parallels with Newtonian physics, their direct applicability to human travel behavior has faced scrutiny [@sen2012gravity]. In contrast, the negative exponential function adopts a more gradual approach. Supported by robust theoretical foundations rooted in entropy maximization [@wilson1971family] and choice behavior theory [@fotheringham1989spatial], this function has emerged as a de facto standard in applied accessibility analysis due to its theoretically sound underpinnings. On the other hand, the modified Gaussian function displays a considerably gentler decline in influence near its origin and an overall slower decay rate.

## Materials and Methods

### The GSS survey

It is critical to have appropriate travel survey data that includes pedestrian and cycling activities to derive impedance functions appropriate for non-motorized modes. The optimal approach would involve a dedicated survey, intricately designed to capture these specific behaviors predominantly, or data collected from Global Positioning Systems---an option that typically has higher costs. When these specific datasets are unavailable, a broader regional household travel survey can be employed, provided it encompasses trips made by non-motorized means [@iacono2010]. In this research, we utilized data from the General Social Survey (GSS), which is administered by Statistics Canada, to delve into active travel behavior in Canada.

The GSS provides a comprehensive cross-sectional snapshot of the Canadian population through telephone surveys established in 1985. The survey coverage area includes both metropolitan and non-metropolitan regions, ensuring a diverse and representative sample of the Canadian population. Specifically, the ten provinces of Canada were divided into distinct geographic strata for sampling purposes. Many Census Metropolitan Areas (CMAs), such as St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver, were treated as separate strata. Additional strata were formed by grouping other CMAs within Quebec, Ontario, and British Columbia, and by categorizing non-CMA areas within each province into their own strata. This stratification results in a total of 27 distinct strata, providing comprehensive coverage across both densely populated urban areas and more sparsely populated rural regions.

These surveys encompass an array of socio-demographic inquiries combined with questions concentrating on specific core themes, such as health, time use, and aspects like social support and aging (Statistics Canada, 2015). One of the standout features of the GSS is its recurring "time use" cycle, which delves into the daily activities of Canadians. This cycle captures the amount of time individuals allocate to various tasks and the sequence, location, and concurrent activities, offering a holistic view of Canadians' daily lives. The questions within this cycle have been adapted and refined over the years to reflect the changing dynamics of daily life, ensuring that the data remains pertinent and contemporary.

In order to investigate the historical active travel behavior in Canada, Six GSS cycles were thoroughly considered for this study, namely those from 1986, 1992, 1998, 2005, 2010, and 2015. The 1986 cycle is notable because it was the first national random sample examining Canadian time-use patterns. Data filtering was essential given the research focus on travel behavior, particularly walking and cycling. It required an exhaustive extraction of entries relevant to these two travel modes. Each GSS Cycle is derived from two microdata sources: the Main and Episode files.  The Main file comprises questionnaire responses and associated data from participants, while the Episode files furnish detailed insights into every activity episode reported by the respondents. For this study, we employed the episode files to establish a comprehensive dataset for impedance function analysis. This dataset encompasses variables such as individual ID, start time, end time, time duration, origins and destinations of each walking and cycling trip, and weight. It should be noted that each record represents a single activity in a respondent's day, ensuring that all episodes collectively span twenty-four hours (or 1440 minutes). The weight parameter signifies the number of time-use episodes that a particular record in the Episode File represents.

Before diving into the analysis, performing preliminary steps to ensure data consistency across different time-use surveys was essential. Over the years, these surveys used diverse coding schemes for activities and contexts. Therefore, harmonizing these coding systems was crucial. The classifications of various activities have evolved from 1986 to 2015. To achieve uniformity, we synchronized the activity categories from 2005, 2010, and 2015 and similarly aligned those from 1986, 1992, and 1998. As a result, for the earlier years (1986, 1992, and 1998), trip origins and destinations were categorized as "home," "other's home," and "work or school." In the later years (2005, 2010, and 2015), these categories expanded to include "Restaurant, bar or club," "Place of worship," "Grocery store, other stores or mall," "outdoors," and "Library, museum or theater." Additionally, it is essential to note that the 1986 dataset solely comprises walking data, with no records of cycling trips for that year. This careful standardization of data across different periods was fundamental for a coherent and accurate analysis of travel behavior trends.

The decision to tap into the time series data of the GSS is motivated by the opportunity it offers to discern and analyze evolving patterns over an extended time frame. This approach spans three decades and facilitates a comprehensive longitudinal assessment of walking and cycling trends in Canada. This depth of temporal coverage is instrumental in capturing the nuances and shifts in travel behaviors over time.

### Estimating impedance function parameters

The foundation of our study rested upon the calculation of impedance functions for walking and cycling trips in each of the six years (1986, 1992, 1998, 2005, 2010, and 2015) under investigation. As mentioned before, the impedance function is a fundamental construct for examining travel behavior, encapsulating the factors influencing the ease or difficulty of traversing a particular route. Diving deeper into travel behavior variations, we acknowledged that behaviors are shaped by the mode of travel (such as walking or cycling), the trip's purpose, and the location specifics of the trip's origin and destination [@geurs2006accessibility; @iacono2008access; @iacono2010; @larsen2010beyond; @millward2013active]. This insight underscored the challenge of generalizing a single impedance function, given that different travel modes and purposes necessitate differing functions. So, in this research, we endeavored to compute the impedance function individually for each destination and mode of transportation. We employed the R programming language to initiate this process and leveraged the "fitdistrplus" package. Our approach was methodical, commencing with a comprehensive exploratory analysis. We generated skewness and kurtosis graphs, providing valuable visual representations of the distributional characteristics inherent in the travel time duration for walking and cycling trips each year.

The selection of an appropriate probability distribution was pivotal in our analysis. By closely scrutinizing the skewness and kurtosis graphs, we aimed to identify the probability distribution that most faithfully mirrored the empirical characteristics of our data. This step was crucial to ensure the subsequent calculations accurately captured the underlying travel behavior. In the quest for the optimal distribution, we systematically evaluated various probability distribution models, including but not limited to the normal, gamma, exponential, and Weibull distributions. Our choice of the most suitable distribution hinged on statistical metrics, including the Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC). These criteria facilitated a quantitative assessment of the goodness of fit for different distributional models. Additionally, we leveraged the maximum likelihood estimation (MLE) method, utilizing the Nelder-Mead optimization algorithm available within the {fitdistrplus} package [@delignette2015fitdistrplus]. It ensured precise estimation of distribution parameters, underlining the rigor of our impedance function calculation.

Following a thorough process of model selection and parameter estimation, we calculated the impedance function. This crucial component assesses the resistance or difficulties associated with walking and cycling trips for each year and destination, considering the different weights of various factors. In defining the impedance function, we identified travel time as the critical metric for measuring trip impedance (cost). This choice is based on the premise that walking and cycling usually do not incur direct monetary costs, aligning with the concepts presented in previous research [@hamidi2014applying]. This approach allows for a more precise evaluation of the impediments pedestrians and cyclists face over time and across different destinations.

## Results and discusion

### Descriptive analysis of walking and cycling trips from 1986 to 2015

The General Social Survey (GSS) data emerged as an invaluable repository in our detailed examination of active travel behavior in Canada. It provided a comprehensive view of individual preferences and behaviors about active modes of transportation, specifically walking and cycling. By focusing on the years 1986 to 2015, we sought to capture the evolution of active travel against the backdrop of crucial urban development changes, societal shifts in attitudes towards health and the environment, and modifications in transportation infrastructure and policies.

\newpage

```{r ch03-make-table-01, echo=FALSE, cache=FALSE}

# Assuming 'trip' is your initial data frame and it's already been loaded.

# Filter out trips with DURATION greater than 90
trip_filtered <- trip %>%
  filter(DURATION <= 90)

# Calculate the count of trips for each mode and year
# Summarize counts by year and mode
trip_counts <- trip_filtered %>%
  group_by(YEAR, MODE) %>%
  summarise(Count = n(), .groups = "drop") %>%
  pivot_wider(names_from = YEAR, values_from = Count)

# Calculate min, mean, and max statistics
stats_data <- trip_filtered %>%
  group_by(YEAR, MODE) %>%
  summarise(
    min = min(DURATION),
    mean = mean(DURATION),
    max = max(DURATION),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(min, mean, max), names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
final_data <- bind_rows(
  trip_counts %>% mutate(Statistic = "Count"),
  stats_data
) %>%
  mutate(MODE = factor(MODE, levels = c("walking", "cycling"))) %>%
  arrange(MODE, Statistic) %>%
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Replace NA in MODE with "walking"
final_data$MODE <- replace_na(final_data$MODE, "walking")

# Rename the 'MODE' column to 'Mode' before generating the table
final_data <- final_data %>%
  rename(Mode = MODE) %>%
  dplyr::select(Mode, Statistic, everything())

# Generate the table with 'Statistic' as the second column
styled_ch03_table_01 <- final_data %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:ch03-table-01}Descriptive Analysis of Active Transportation Modes: Walking and Cycling Statistics from 1986 to 2015",  
        align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
  kable_styling(full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 2, "Year" = ncol(final_data) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01

```

Our primary method was rooted in descriptive analysis. We parsed vast quantities of data to discern overarching patterns and trends. A salient feature emerging from this analysis was the duration of trips undertaken.Table \ref{tab:ch03-table-01} contains a detailed compilation of descriptive statistics showing the fluctuating patterns of walking and cycling trips between 1986 and 2015. The data is categorized under two principal modes of transit: walking and cycling, with each category further explored through four statistical metrics: count, maximum (max), mean, and minimum (min), each quantified in minutes. In the walking trips section, the 'count' metric represents the number of recorded trips for each year. The journey began in 1986 with 384 recorded walking trips, escalating to 1,473 by 1992, indicating a burgeoning preference for walking during this period. The peak was reached in 1998 with 1,657 trips, which increased to an unprecedented 5,517 trips by 2005, possibly reflecting societal shifts toward healthier lifestyles or environmental concerns. However, post-2005, a downward trend emerged, with counts reducing to 4,360 trips in 2010 and further to 2,796 in 2015, possibly due to changing urban landscapes or the advent of alternative transportation modes.

The 'max' time invested in walking trips remains a constant 90 minutes across the years. However, the 'mean' walking time varies: it starts at 24 minutes in 1986, drops to 17 minutes in 1992, and finally to 10 minutes in 1998. This trend reverses slightly to 12 minutes in both 2005 and 2010 and ascends to 16 minutes in 2015. This fluctuation in average walking time could reflect evolving urban designs, societal pace, or changing durations in individuals' routines. The 'min' walking time, initially stable at 1 minute, drops to nonexistence in 2005 and 2010, then rebounds to 5 minutes in 2015.

For cycling trips, data is absent for 1986. Starting in 1992, 126 trips are noted, dipping marginally to 119 in 1998, then more than doubling to 325 by 2005—indicative of a surge in cycling popularity, potentially driven by environmental campaigns or improvements in cycling infrastructure. The count declined to 214 trips in 2010 but experienced a marginal recovery to 217 trips by 2015. The 'max' cycling time experiences a gradual reduction from 90 minutes in the early years to 80 minutes by 2015, possibly reflecting advancements in bicycle technology or changes in cyclists’ behaviors. The 'mean' time initially recedes from 24 minutes in 1992 to 18 minutes in 2005, then modestly increases, culminating at 22 minutes in 2015, suggesting that the average trip becomes slightly longer, which could be due to several factors such as increased commuting distances or a rise in recreational cycling. The 'min' cycling time shows a sharp decline to 1 minute in 2005 and 2010, paralleling the walking trend, before climbing back to 5 minutes in 2015, which could indicate cycling's growing popularity as a short-distance travel alternative.

As highlighted in Table \ref{tab:ch03-table-01}, throughout the 30 years under study, the duration of walking trips was consistently lower than that of cycling trips. To quantify this discrepancy, the mean duration for walking trips was approximately 33% of the corresponding duration for cycling trips, underscoring the inherent differences in these modes of transportation. An intriguing trend was observed between the years 1986 to 2005. The average duration of walking and cycling trips decreased significantly over the last two decades. Various factors might have precipitated this trend, such as urban sprawl, increased reliance on motorized transport, or societal preferences for faster modes of transportation. However, the subsequent decade (2005-2015) witnessed a marked reversal in this trend. The average duration for walking and cycling trips stabilized and began showing signs of resurgence. It could indicate many factors: growing urbanization leading to more accessible destinations, increased awareness and initiatives promoting health and sustainability, changes in urban planning that prioritize active modes of travel and improving required infrastructure, or a combination of these elements. This resurgence suggests a renewed and possibly growing affinity towards walking and cycling as viable modes of transportation.

```{r ch03-make-table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}

# Filter for walking and cycling trips, selected years, and duration <= 90
trip_stats_data <- trip %>%
  filter(MODE %in% c("Walking", "cycling"), YEAR %in% c(1986, 1992, 1998), DURATION <= 90) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    Count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- trip_stats_data %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(Count), .groups = "drop")
trip_stats_data <- left_join(trip_stats_data, total_counts, by = c("YEAR", "MODE"))
trip_stats_data$Percentage <- round((trip_stats_data$Count / trip_stats_data$total_count) * 100, 1)

# Group by YEAR, MODE, and dest, then summarize
trip_stats_data <- trip_stats_data %>%
  group_by(YEAR, MODE, dest_label) %>%
  summarise(
    Count = sum(Count),
    min = min(min),  # Minimum of minimum durations per group
    median = median(median),  # Median of median durations per group
    max = max(max),  # Maximum of maximum durations per group
    total_count = sum(total_count),
    Percentage = mean(Percentage),
    .groups = "drop"
  )

# Pivot data for the table layout
wide_data <- trip_stats_data %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  )

# Adjust column names for clarity and consistency
wide_data <- wide_data %>%
 dplyr::select(
    dest_label, MODE,
    min_1986, median_1986, max_1986, Percentage_1986,
    min_1992, median_1992, max_1992, Percentage_1992,
    min_1998, median_1998, max_1998, Percentage_1998
  )

# Arrange rows based on MODE and then the decreasing percentage of year 1998
wide_data <- wide_data %>% 
  arrange(MODE, desc(Percentage_1998))

wide_data <- wide_data %>% 
  rename(
    P_86 = Percentage_1986,
   P_92 = Percentage_1992,
    P_98 = Percentage_1998,
   min_86 = min_1986,
   min_92 = min_1992,
   min_98 = min_1998,
   med_86 = median_1986,
   med_92 = median_1992,
   med_98 = median_1998,
   max_86 = max_1986,
   max_92 = max_1992,
   max_98 = max_1998
  )

# Create a new 'label_mode' column that labels the mode at the first occurrence
wide_data <- wide_data %>%
  group_by(MODE) %>%
  mutate(label_mode = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup() %>%
  dplyr::select(-MODE) %>%
  rename(MODE = label_mode) %>%
  relocate(MODE, .before = 2)

styled_ch03_table_02 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:ch03-table-02}Comparative Trip Stats by Mode and Destination: 1986, 1992, 1998",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "P*", "Min","Med","Max", "P", "Min","Med","Max", "P")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                font_size = 6) %>%
  column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
  column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>%
  add_footnote(c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; 'P' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination."))

styled_ch03_table_02

```

```{r ch03-make-table-03,echo=FALSE, cache=FALSE, warning=FALSE}

# Rename certain 'dest_label' values
trip <- trip %>%
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Outdoors",
    TRUE ~ dest_label  # Keep all other values as they are
  ))

# Filter for walking and cycling trips, selected years, and duration <= 90
trip_stats_data <- trip %>%
  filter(MODE %in% c("Walking", "cycling"), YEAR %in% c(2005, 2010, 2015), DURATION <= 90) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    Count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- trip_stats_data %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(Count), .groups = "drop")
trip_stats_data <- left_join(trip_stats_data, total_counts, by = c("YEAR", "MODE"))
trip_stats_data$Percentage <- round((trip_stats_data$Count / trip_stats_data$total_count) * 100, 1)

# Group by YEAR, MODE, and dest, then summarize
trip_stats_data <- trip_stats_data %>%
  group_by(YEAR, MODE, dest_label) %>%
  summarise(
    Count = sum(Count),
    min = min(min),  # Minimum of minimum durations per group
    median = median(median),  # Median of median durations per group
    max = max(max),  # Maximum of maximum durations per group
    total_count = sum(total_count),
    Percentage = mean(Percentage),
    .groups = "drop"
  )

# Pivot data for the table layout
wide_data <- trip_stats_data %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  )

# Adjust column names for clarity and consistency
wide_data <- wide_data %>%
 dplyr::select(
    dest_label, MODE,
    min_2005, median_2005, max_2005, Percentage_2005,
    min_2010, median_2010, max_2010, Percentage_2010,
    min_2015, median_2015, max_2015, Percentage_2015
  )

# Arrange rows based on MODE and then the decreasing percentage of year 2015
wide_data <- wide_data %>% 
  arrange(MODE, desc(Percentage_2015))

wide_data <- wide_data %>% 
  rename(
    P_05 = Percentage_2005,
   P_10 = Percentage_2010,
    P_15 = Percentage_2015,
   min_05 = min_2005,
   min_10 = min_2010,
   min_15 = min_2015,
   med_05 = median_2005,
   med_10 = median_2010,
   med_15 = median_2015,
   max_05 = max_2005,
   max_10 = max_2010,
   max_15 = max_2015
  )

# Create a new 'label_mode' column that labels the mode at the first occurrence
wide_data <- wide_data %>%
  group_by(MODE) %>%
  mutate(label_mode = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup() %>%
  dplyr::select(-MODE) %>%
  rename(MODE = label_mode) %>%
  relocate(MODE, .before = 2)

styled_ch03_table_03 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:ch03-table-03}Comparative Trip Statistics by Transportation Mode and Destination: 2005, 2010, and 2015",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "P*", "Min","Med","Max", "P", "Min","Med","Max", "P")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                font_size = 6) %>%
  column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
  column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>%
  add_footnote(c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; 'P' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"))

styled_ch03_table_03
```

Tables \ref{tab:ch03-table-02} and \ref{tab:ch03-table-03} thoroughly explore trip statistics by transportation mode and destination for the years 1986, 1992, 1998, 2005, 2010, and 2015. These tables detail the minimum, median, and maximum trip durations, accompanied by their respective percentages, which reflect the proportion of trips within each category. The temporal measures are presented in minutes. An analytical comparison of the tables reveals evolving patterns in transportation behaviors over time, with a particular emphasis on walking and cycling.

Table \ref{tab:ch03-table-02} reveals that for the year 1986, the dataset is limited to walking trips, as cycling data is not available, possibly indicating a lack of cycling infrastructure or lower cycling popularity at that time. Walking to home destinations represents a significant portion of trips, with a median duration of 15 minutes and a maximum of 90 minutes. Also, walking trips to one's home had a median duration of 15 minutes and a maximum of 90, accounting for over 40% of trips, suggesting walking was a reasonably common mode of travel. In contrast, trips to work or school were less frequent, at 14.6%, with similar median and maximum durations. Walking trips to one's home had a median duration of 15 minutes and a maximum of 90, accounting for over 40% of trips, suggesting walking was a reasonably common mode of travel. In contrast, trips to work or school were less frequent, at 14.6%, with similar median and maximum durations mirroring those to home, indicating that walking served as a secondary mode of commute.

By 1992, cycling data appears, indicating an introduction or increased adoption of this mode of travel. The median cycling times to home and other destinations are relatively low at 20 and 10 minutes, respectively. However, the maximum times are substantial (90 and 70 minutes), suggesting some cyclists were willing to travel longer distances. Walking trips to home still dominate, with an increased percentage of nearly 59% and the median time remaining 10 minutes. The data might suggest an increasing tendency to reside closer to workplaces or schools or improved pedestrian access. Moreover, the percentages, reflective of the trip count, showed that over half of the commutes to home were done by bicycle. The presence of cycling trips to work or school is not recorded, which might suggest limited adoption or data collection constraints for this purpose. In 1998, walking trips to one's home accounted for 61.2% of the total trips, a substantial increase from 43.5% in 1986. The median duration of these trips remained consistent at 10 minutes over the years, with a maximum duration of 90 minutes. This indicates a strong preference for walking as a mode of transport to home, possibly due to various factors such as the proximity of residential areas to city centers, increased walkability of cities, or a heightened societal focus on physical health and environmental consciousness.

In comparison, walking trips to work or school comprised 20.5% of trips in 1998, a slight increase from 14.6% in 1986. The median duration for these trips was 6.5 minutes, notably lower than for walking trips home. It could suggest that work or school destinations were, on average, closer to the individuals' starting points or that there was a selective preference for walking to work or school only when the distance was relatively short. The percentage of walking trips to 'other' destinations was 28.2%, with a median duration of 5 minutes and a maximum of 90 minutes. The 'other' category likely encompasses a variety of destinations such as shopping, leisure activities, or social visits, which may be more dispersed in location, hence the broader range in trip durations.

Moving to Table 2.3, by 2005, a variety of destinations are included, such as grocery stores and restaurants, indicating a more detailed analysis of trip purposes. Cycling to work or school sees a high median time of 15 minutes and a maximum of 75 minutes, illustrating that cycling has become a more prominent mode of transportation for longer distances, possibly due to improved cycling infrastructure or increased environmental awareness. In addition, In 2005, the destination with the maximum percentage of walking was the trip to home, holding a significant 44.3%. It suggests that walking to work or school was the most common trip purpose among pedestrians. The second most common walking trip was to work or school, with a percentage of 17.1%, followed by grocery store destinations at 12.5%. The least common walking destinations were Library, museum, and theater locations, which held a negligible percentage, possibly due to their infrequency or distance from residential areas. The destination with the highest percentage for cycling was also to home, at 48.9%. This percentage, while higher than that of walking, reflects a considerable number of individuals opting to cycle for their commutes. The second and third most popular cycling destinations were work or school (22.5%), followed by grocery stores (10.2%).

By 2010, the median durations for cycling to work or school increased to 15.0 minutes. The trend shifts slightly between 2005 to 2010. The maximum percentage for walking trips remained home, albeit decreased to 43.6%. The second most common destination for walking was work or school at 15%, followed by grocery stores at 13.3%. The least common was again library, museum, and theater locations, similar to 2005, suggesting a consistent pattern over the years.
Cycling to home reached its peak in 2010, with the highest percentage at 51.4%, overtaking walking for this destination and indicating a surge in cycling's popularity for commuting. However, the second destination remains work or school with 25.7%, and the third destination was changed to restaurants and bars, which increased to 10.3%.

In 2015, the median time for cycling trips to home and work or school decreased (15.0 and 10.0 minutes, respectively) compared to 2010, which might indicate more efficient cycling conditions or a shift in the locations of homes relative to workplaces. Walking trips maintain a steady percentage, especially to home, which could signal a consistent preference for walking as a reliable and health-conscious mode of transportation. In 2015, walking home with 46.7 was the most popular destination for walking. Also, walking to work or school increased to 17.1%, tying with cycling for the destination with the highest percentage. It suggests a balancing out of the preferences between the two modes for commuting. The third most common walking trip was to grocery stores, which increased to 12.7%, and the fourth was trips to restaurants at 8.9%. The least common walking destination was to places of worship or business, which had the smallest percentage, reflecting their specific and potentially less frequent nature.
Cycling, while experiencing a decrease in the percentage of home travel compared to 2010, remained a significant mode of transportation for that purpose. In 2015, the specific percentages for other cycling destinations allocated to work or school and grocery stores increased to 31.8 and decreased to 7.4%.

When comparing walking and cycling, it is evident that cycling has been embraced for longer-distance commutes over time, as seen in the extended median and maximum durations. While maintaining a significant share of trips, mainly to home destinations, walking tends to be favored for shorter distances. This preference may reflect walking's inherent limitations in speed and convenience over longer distances compared to cycling.
The evolution from 2005 to 2015 suggests an increasing preference for shorter, more efficient trips, potentially indicative of changing urban designs and lifestyles that favor convenience and time-saving. The data points to a population increasingly incorporating cycling into their routine for various trip purposes, reflecting broader trends towards sustainable transportation and active lifestyles. Overall, these statistics offer valuable insights into the dynamics of urban transportation, highlighting the gradual but distinct shifts in how people choose to travel within their cities. These shifts are influenced by a confluence of factors, including urban development, cultural trends, environmental considerations, and transportation policies aimed at promoting active transport modes.

```{r ch03-make-Heatmap-walking_2015, echo=FALSE, cache=FALSE, include=FALSE}

#Creating walking heatmap in 2015

# Combine all origin and destination labels from 2005, 2010, and 2015 datasets
all_origins <- unique(c(as.character(walking_2005$orig_label), 
                        as.character(walking_2010$orig_label), 
                        as.character(walking_2015$orig_label)))
all_destinations <- unique(c(as.character(walking_2005$dest_label), 
                             as.character(walking_2010$dest_label), 
                             as.character(walking_2015$dest_label)))

# Apply consistent factor levels across all datasets
walking_2015$orig_label <- factor(walking_2015$orig_label, levels = all_origins)
walking_2015$dest_label <- factor(walking_2015$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_2015$WGHT_EPI <- as.numeric(as.character(walking_2015$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_2015 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI)) # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2015 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 2015
destination_order <- c("Home", "Work or school", "Grocery store, other stores or mall", "Others' home",
                       "Outdoors", "Library, museum or theatre", "Place of worship", "Restaurant, bar or club", "Sporting events, field or other")

# Create the ggplot object and assign it to a variable
heatmap_plot_2015<- ggplot(heatmap_melted_2015, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2015$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2015$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2015$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2015") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-walking_2010, echo=FALSE, cache=FALSE, include=FALSE}
# Creating walking heatmap in 2010
# Apply consistent factor levels across all datasets
walking_2010$orig_label <- factor(walking_2010$orig_label, levels = all_origins)
walking_2010$dest_label <- factor(walking_2010$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_2010$WGHT_EPI <- as.numeric(as.character(walking_2010$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_2010 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI)) # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2010 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 2010
destination_order <- c("Home", "Work or school", "Grocery store, other stores or mall", "Others' home",
                       "Outdoors", "Library, museum or theatre", "Place of worship", "Restaurant, bar or club", "Sporting events, field or other")

# Create the ggplot object and assign it to a variable
heatmap_plot_2010 <- ggplot(heatmap_melted_2010, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2010$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2010$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2010$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2010") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-walking_2005, echo=FALSE, cache=FALSE, include=FALSE}

# Creating walking heatmap in 2005
# Apply consistent factor levels across all datasets
walking_2005$orig_label <- factor(walking_2005$orig_label, levels = all_origins)
walking_2005$dest_label <- factor(walking_2005$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_2005$WGHT_EPI <- as.numeric(as.character(walking_2005$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_2005 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI)) # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2005 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 2005
destination_order <- c("Home", "Work or school", "Grocery store, other stores or mall", "Others' home",
                       "Outdoors", "Library, museum or theatre", "Place of worship", "Restaurant, bar or club", "Sporting events, field or other")

# Create the ggplot object and assign it to a variable
heatmap_plot_2005 <- ggplot(heatmap_melted_2005, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2005$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2005$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2005$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2005") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-walking_1998, echo=FALSE, cache=FALSE, include=FALSE}

# Creating walking heatmap in 1998
# Combine all origin and destination labels from 1986, 1992, and 1998 datasets
all_origins <- unique(c(as.character(walking_1986$orig_label), 
                        as.character(walking_1992$orig_label), 
                        as.character(walking_1998$orig_label)))
all_destinations <- unique(c(as.character(walking_1986$dest_label), 
                             as.character(walking_1992$dest_label), 
                             as.character(walking_1998$dest_label)))

# Apply consistent factor levels across all datasets
walking_1998$orig_label <- factor(walking_1998$orig_label, levels = all_origins)
walking_1998$dest_label <- factor(walking_1998$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_1998$WGHT_EPI <- as.numeric(as.character(walking_1998$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_1998 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI)) # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_1998 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 1998
destination_order <- c("Home", "Work or school", "Others' home")

# Create the ggplot object and assign it to a variable
heatmap_plot_1998 <- ggplot(heatmap_melted_1998, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_1998$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_1998$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_1998$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "1998") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-walking_1992, echo=FALSE, cache=FALSE, include=FALSE}

# Creating walking heatmap in 1992
# Apply consistent factor levels across all datasets
walking_1992$orig_label <- factor(walking_1992$orig_label, levels = all_origins)
walking_1992$dest_label <- factor(walking_1992$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_1992$WGHT_EPI <- as.numeric(as.character(walking_1992$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_1992 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI, na.rm = TRUE), # Calculate weighted count for each origin-destination pair
            .groups = "drop") # Avoid automatic grouping

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_1992 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 1992
destination_order <- c("Home", "Work or school", "Others' home")

# Create the ggplot object and assign it to a variable
heatmap_plot_1992 <- ggplot(heatmap_melted_1992, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_1992$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_1992$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_1992$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "1992") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-walking_1986, echo=FALSE, cache=FALSE, include=FALSE}

# Creating walking heatmap in 1986

# Apply consistent factor levels across all datasets
walking_1986$orig_label <- factor(walking_1986$orig_label, levels = all_origins)
walking_1986$dest_label <- factor(walking_1986$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
walking_1986$WGHT_EPI <- as.numeric(as.character(walking_1986$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- walking_1986 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI)) # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_1986 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Define the order of destinations for 1986
destination_order <- c("Home", "Work or school", "Others' home")

# Create the ggplot object and assign it to a variable
heatmap_plot_1986 <- ggplot(heatmap_melted_1986, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_1986$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_1986$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_1986$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "1986") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-cycling_2015,echo=FALSE, cache=FALSE, include=FALSE }
# Creating cycling heatmap in 2015
# Combine all origin and destination labels from 2005, 2010, and 2015 datasets
all_origins <- unique(c(as.character(cycling_2005$orig_label), 
                        as.character(cycling_2010$orig_label), 
                        as.character(cycling_2015$orig_label)))
all_destinations <- unique(c(as.character(cycling_2005$dest_label), 
                             as.character(cycling_2010$dest_label), 
                             as.character(cycling_2015$dest_label)))

cycling_2015$orig_label <- factor(cycling_2015$orig_label, levels = all_origins)
cycling_2015$dest_label <- factor(cycling_2015$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
cycling_2015$WGHT_EPI <- as.numeric(as.character(cycling_2015$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- cycling_2015 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI), .groups = 'drop') # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2015 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Create the ggplot object and assign it to a variable
heatmap_plot_c_2015 <- ggplot(heatmap_melted_2015, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2015$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2015$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2015$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2015") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-cycling_2010,echo=FALSE, cache=FALSE, include=FALSE }
# Creating cycling heatmap in 2010

# Convert labels in 2010 dataset to factors with consistent levels
cycling_2010$orig_label <- factor(cycling_2010$orig_label, levels = all_origins)
cycling_2010$dest_label <- factor(cycling_2010$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
cycling_2010$WGHT_EPI <- as.numeric(as.character(cycling_2010$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- cycling_2010 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI), .groups = 'drop') # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2010 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Create the ggplot object and assign it to a variable
heatmap_plot_c_2010 <- ggplot(heatmap_melted_2010, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2010$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2010$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2010$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2010") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-cycling_2005,echo=FALSE, cache=FALSE, include=FALSE }

# Creating cycling heatmap in 2005
# Specify consistent factor levels for origins and destinations
cycling_2005$orig_label <- factor(cycling_2005$orig_label, levels = all_origins)
cycling_2005$dest_label <- factor(cycling_2005$dest_label, levels = all_destinations)

# Ensure WGHT_EPI is numeric
cycling_2005$WGHT_EPI <- as.numeric(as.character(cycling_2005$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- cycling_2005 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI), .groups = 'drop') # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_2005 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Create the ggplot object and assign it to a variable
heatmap_plot_c_2005 <- ggplot(heatmap_melted_2005, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_2005$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_2005$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_2005$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "2005") +
  theme_minimal() +
 theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-cycling_1998,echo=FALSE, cache=FALSE, include=FALSE }

# Creating cycling heatmap in 1998
# Ensure WGHT_EPI is numeric
cycling_1998$WGHT_EPI <- as.numeric(as.character(cycling_1998$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- cycling_1998 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI), .groups = 'drop') # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Get the maximum percentage value for the legend
max_percentage <- max(trip_counts$percentage, na.rm = TRUE)

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_1998 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Create the ggplot object and assign it to a variable
heatmap_plot_c_1998 <- ggplot(heatmap_melted_1998, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_1998$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_1998$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_1998$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "1998") +
  theme_minimal() +
  theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

```{r ch03-make-Heatmap-cycling_1992,echo=FALSE, cache=FALSE, include=FALSE }

# Creating cycling heatmap in 1992
# Ensure WGHT_EPI is numeric
cycling_1992$WGHT_EPI <- as.numeric(as.character(cycling_1992$WGHT_EPI))

# Calculate the count of each pair of the orig-dest trip and their weighted counts
trip_counts <- cycling_1992 %>%
  group_by(orig_label, dest_label) %>%
  summarize(count_each_trip = n(), # Count the number of trips for each unique origin-destination pair
            weighted_count = sum(WGHT_EPI), .groups = 'drop') # Calculate weighted count for each origin-destination pair

# Calculate the total weighted count for normalization
total_weighted_count <- sum(trip_counts$weighted_count)

# Calculate percentages and round to 2 decimal places
trip_counts <- trip_counts %>%
  mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

# Reshape data for heatmap
heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

# Convert the data frame to a matrix for the heatmap
heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

# Melt the matrix for use with ggplot
heatmap_melted_1992 <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

# Create the ggplot object for the heatmap with a centered and styled caption
heatmap_plot_c_1992 <- ggplot(heatmap_melted_1992, aes(x = Destination, y = Origin, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                       na.value = "gray95", 
                       limits = c(0, max(heatmap_melted_1992$value, na.rm = TRUE)), 
                       breaks = c(0, max(heatmap_melted_1992$value, na.rm = TRUE)), 
                       labels = c("0%", paste0(max(heatmap_melted_1992$value, na.rm = TRUE), "%"))) +
  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = "1992") +
  theme_minimal() +
 theme(
  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
  axis.title = element_text(size = 16),
  legend.title = element_text(size = 16, margin = margin(b = 15)), 
  legend.text = element_text(size = 16),
  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt"))
)
```

Figure 1 and Figure 2, as depicted, serve as a comprehensive visual compilation of walking and cycling trip data spanning nearly three decades, from 1986 to 2015. These heat maps employ a nuanced color gradient to translate complex data into an easily interpretable format. The spectrum of colors, from deep, rich tones to pale hues, corresponds directly to the percentage of trips along various origins and destinations, with the darkest shades representing the highest percentage between origins and destinations. Lighter shades denote pathways less traveled, offering a clear visual distinction between the frequented and less frequented corridors of pedestrian/cyclist movement.

The 2015 heat map particularly highlights the dominance of trips originating from homes. Walking trips from home to work or school represent a notable 13.2%, depicted as the darkest cell on the map. This indicates a preference for walking as a mode of transport for commuting in Canada. In contrast, the map also reveals walking trips from workplaces or schools back to homes at 13.91%, indicating a return flow in commuting patterns. Additionally, home-originating trips to other homes, accounting for 9.22%, might reflect short neighborhood walks for purposes like exercise or dog walking. Trips from home to grocery stores, other stores, or malls, making up 7.14% of the trips, suggest a shift in walking frequency for shopping purposes. Trips originating from other homes are also prominent, especially to homes (6.95%), possibly indicating social visitation patterns. Trips from home to restaurants, bars, or clubs, at 4.88%, reveal a social and recreational aspect of walking trips. However, walking trips starting outdoors are less frequent, with the highest being trips to homes at 0.51%. The lower end of the spectrum includes origins like businesses, libraries, museums, theaters, and places of worship, indicating these as less common walking paths. For example, business-to-restaurant, bar, or club trips are a mere 0.11%.

The 2010 heatmap shows the most frequent walking trips from home to work or school, constituting 9.93% of all trips and emphasizing walking in daily commutes. The return trips from work or school to home also show a high percentage at 9.34%. Notably, trips from other's homes back to home, at 8.54%, may indicate a tendency for walking during social visitations. The third and fourth most common routes are from home to shopping destinations and back, accounting for 6.51% and 8.7% of the trips, respectively. Walking trips from home to restaurants or bars make up 3.28%, highlighting walking's role in leisure activities. Less frequent trips include those to libraries, places of worship, and outdoors, indicated by lighter shades on the heatmap. In 2005, the most prevalent walking trip was from home to work or school, accounting for 13.22% of all walking trips. The data further reveals that walking trips from other's homes to homes, at 8.3%, signify a notable trend in social visitations. Shopping-related walking trips also show high frequencies. Leisure activities like walking to restaurants or bars account for 3.61%. Less common trips include those to libraries, places of worship, and outdoors.

The 1998 heatmap reveals a predominance of trips from home to other's homes, at 21.83%, emphasizing the role of pedestrian mobility in maintaining social connections. Commuting trips, such as from home to work or school, constitute 11.56% of the trips. However, trips within the work or school category are least frequent at 1.7%. In 1992, the most frequent walking trips were within the home category, at 27.85%, likely encompassing leisure, dog walking, or short errands. Social visitations also represent a significant portion, at 15.88%. Commuting trips from home to work or school made up 18.62% of the walking trips, with return trips being less frequent at 14.73%. Trips originating from other's homes show diverse figures, with 'other's home' to 'home' trips being quite prevalent at 17%. Lastly, in 1986, most walking trips occurred between homes and other's homes, with trips from other's homes to homes being most prevalent at 29.21%. It underscores the importance of walking in maintaining social ties. Commuting trips, represented at 11.56%, suggest a smaller proportion of walking trips compared to social visits. Intracategory trips from home to home, likely representing leisure walks, accounted for 6.01%. Walking trips within the work or school category were low at 1.7%, possibly due to limited walking opportunities during work or school.

Over nearly three decades, the trends in walking trips reveal significant shifts in the patterns and preferences of pedestrians. In the earlier years, such as 1986 and 1992, walking trips were predominantly characterized by social visitations, as evidenced by the high percentage of trips between 'home' and 'other's home. In 1986, these social trips constituted the majority of walking activities, with 29.21% from 'other's home' to 'home' and 21.83% in the reverse direction. By 1992, there was a notable increase in intracategory 'home' trips (27.85%), suggesting a rise in walking for leisure or short errands within the neighborhood. This shift highlights transitioning from walking predominantly for social purposes to incorporating more leisure and short-distance errands.

In contrast, in the later years, especially in 2005 and 2015, there has been a clear trend towards walking being more frequently used for commuting purposes. In 2005, the commute from 'home' to 'work or school' was the most prevalent walking trip, accounting for 13.22% of all walking trips. Similarly, in 2015, walking trips from home to work or school represented the highest percentage (13.2%), indicating an increased reliance on walking as a primary mode of transportation for commuting. This transition from social to commuting purposes reflects broader changes in urban lifestyles and possibly improvements in pedestrian infrastructure, making walking a more viable option for daily commutes. The data also shows a consistent pattern of return trips from work or school to home, further underscoring the bidirectional nature of commuting trips.
These changing patterns indicate how the role of walking in urban environments has evolved. Initially, it served primarily as a means of social connectivity and short leisurely activities. Over time, however, it has increasingly become a key component of the daily commute, reflecting changes in urban planning, societal habits, and possibly environmental consciousness. The heat maps' changing color gradients over the years visually encapsulate this evolution, marking a shift in the urban pedestrian landscape

```{r ch03-make-fig-01, include=FALSE}
# Adjust each plot to have a fixed aspect ratio
heatmap_plot_1986 <- heatmap_plot_1986 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12)   
  )

heatmap_plot_1992 <- heatmap_plot_1992 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)  
  )
heatmap_plot_1998 <- heatmap_plot_1998 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_2005 <- heatmap_plot_2005 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_2010 <- heatmap_plot_2010 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_2015 <- heatmap_plot_2015 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )

# Combine the plots with adjusted spacing
ch03_fig_01 <- cowplot::plot_grid(heatmap_plot_1986, heatmap_plot_2005, heatmap_plot_1992, heatmap_plot_2010, heatmap_plot_1998, heatmap_plot_2015, ncol = 2, align = 'hv',  rel_widths = c(1, 1), rel_heights = c(1, 1, 1))

# Save the combined plot to a file
ggsave(file = "figure/ch03_fig_01.jpg", 
       plot = ch03_fig_01,
       width = 17, 
       height = 22, 
       units = "in")

```

```{r ch03-plot-fig-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_01.jpg")
```

```{r ch03-make-fig-02, include=FALSE}

# Adjust each plot to have a fixed aspect ratio

heatmap_plot_c_1992 <- heatmap_plot_c_1992 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_c_1998 <- heatmap_plot_c_1998 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_c_2005 <- heatmap_plot_c_2005 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_c_2010 <- heatmap_plot_c_2010 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"), 
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )
heatmap_plot_c_2015 <- heatmap_plot_c_2015 +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12)   
  )

# Combine the plots with adjusted spacing
ch03_fig_02 <- cowplot::plot_grid(heatmap_plot_c_1992, heatmap_plot_c_2005,heatmap_plot_c_1998, heatmap_plot_c_2010, heatmap_plot_c_2015,ncol = 2, align = 'hv', rel_widths = c(2, 2),  rel_heights = c(2, 2, 2))

# Save the combined plot to a file
ggsave(file = "figure/ch03_fig_02.jpg", 
       plot = ch03_fig_02,
       width = 17, 
       height = 22, 
       units = "in")
```

```{r ch03-plot-fig-02, echo=FALSE, out.width="100%", fig.cap="Percentage of cycling Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_02.jpg")
```
In 2015, bicycles were clearly preferred as a primary mode of transportation, especially for commuting from 'home' to 'work or school,' accounting for 32.73% of all cycling trips. It underscores the bicycle's role in daily commutes, reflecting a commitment to environmental sustainability and fitness. The return journey from work or school to home also showed a significant volume, representing 30.38% of trips, indicating a balanced bidirectional commuting pattern. Besides commuting, cycling for errands and social visitations was notable, with trips to grocery stores and other retail destinations making up 5.41% and trips for social purposes to another's home at 3.66%. Leisure activities, including visits to restaurants, bars, or clubs, accounted for 2.06% of trips. Additionally, trips from home to home were reported at 4.58%, highlighting cycling's role in local mobility.

Commuting remained a key cycling activity in 2010, with 27.97% of trips made from 'home' to 'work or school' and 25.38% in the opposite direction, reinforcing the bicycle's importance in daily travel. Social interactions played a significant role, with 8.71% of trips from 'home' to 'other's home' integrating bicycles into the social fabric of Canadian life. Errands and dining or entertainment purposes remained consistent with previous years, at 5.01% and 2.06%, respectively. Home-to-home trips accounted for 5.97%, underscoring the versatility of cycling for various personal needs.

By 2005, the landscape of cycling trips showed a diverse usage pattern, with commuting still at 18.74% for trips from 'work or school' to 'home'. The data highlights the bicycle's vital role in Canadian daily life as a reliable mode of transport. Interestingly, trips from 'home' to 'home' constituted 7.85% of the total, indicating that a significant portion of cycling activity was dedicated to local and recreational mobility. Social and errand cycling, at 7.06% and 7.15%, respectively, along with recreational trips, at 2.07%, show the broadening scope of cycling beyond mere commuting.

In 1998, the data presented a deep color saturation on the heat map for trips from 'work or school' to 'home' and vice versa, with 26.64% and 30.75% of trips, respectively, highlighting a strong bidirectional commuting behavior. The role of bicycles in social visitation was also significant, with 13.53% of trips from 'home' to 'other's home' and 20.74% for the return journey. Home-to-home trips were 4.18%, pointing to the growing trend of using bicycles for a mix of commuting, social, and recreational purposes.The bicycle's essential role in commuting was further reinforced in 1992 when data showed that 26.13% of cycling trips were made from "home" to "work or school," with a sizable return trip percentage of 20.33%. Social visitations also featured prominently, with 20.74% of trips from 'other's home' back to 'home', indicating the bicycle's role in facilitating community connectivity. Cycling trips within the 'home' category, indicating a variety of local uses, were at 15.03%.

Over the years, the cycling trends in Canada have shown an evolving pattern, with a consistent emphasis on commuting while also demonstrating an increasing integration of bicycles into various aspects of urban life. In 2015, cycling was predominantly used for commuting, with 32.73% of trips from 'home' to 'work or school' and a nearly equal percentage for the return journey. This trend reflects a strong environmental and fitness consciousness in urban Canadian society. Comparatively, in 2010, while commuting still remained the primary use of bicycles (27.97% for going to work or school and 25.38% for returning), there was a noticeable increase in cycling for social visitations, accounting for 8.71% of trips. This shift indicates a broader adoption of cycling beyond utilitarian purposes, embedding it more deeply into the social fabric.

In earlier years, such as 2005 and 1998, the emphasis on commuting was still apparent but with a slightly different focus. For instance 2005, the most prevalent cycling route was from work or school to home, constituting 18.74% of trips. This difference in the direction of the dominant route might suggest variations in cycling habits or available infrastructure at the time. In 1998, the data showed a balance between commuting and social interactions, with high percentages for trips from work or school to home (26.64%) and from 'home' to 'other's home' (13.53%). This balance demonstrates the role of bicycles in fostering community interactions and local mobility. Going back to 1992, cycling for commuting (26.13%) and social visitations (20.74%) were also significant, highlighting the bicycle's longstanding role in facilitating professional and interpersonal connectivity.

Overall, these trends across different years illustrate a consistent reliance on bicycles for commuting in Canada, coupled with a growing appreciation of their role in social and recreational activities. The shifting percentages and patterns of use reflect changing urban lifestyles, advancements in cycling infrastructure, and an increasing recognition of the bicycle as a versatile, environmentally friendly, and health-conscious mode of transportation in Canadian cities

### Impedance function analysis

The impedance function figures presented in this section illuminate the changes in walking and cycling trip durations to various destinations across Canada over an extensive period from 1986 to 2015. These distance decay curves are pivotal to understanding impedance functions and are critical for travel behavior analysis. A fundamental idea of spatial interaction is represented by their depiction of the decreasing probability of selecting walking or cycling as a mode of transportation as the distance between the origin and the destination increases. This idea states that the likelihood of traveling between two points decreases with increasing distance. These curves also indicate the critical points at which a person's tendency to walk or cycle abruptly decreases.

Our exploration of distance decay curves for walking and cycling trips to diverse destinations utilized the General Social Survey (GSS) dataset, enabling an examination of how these curves vary across transportation modes and destination types over time. The data shed light on the frequency and characteristics of walking and cycling trips, with the years 2005, 2010, and 2015 featuring a wide range of destinations, from homes to workplaces, schools, other people's homes, grocery stores, retail outlets, malls, outdoor spaces, restaurants, bars, clubs, libraries, museums, theaters, and places of worship. This breadth of destination types signifies a marked expansion from the data of earlier years (1986, 1992, and 1998), which cataloged trips primarily to homes, other people's homes, and work or school. This evolution in data collection reflects a growing understanding of the complex nature of urban mobility and the diverse purposes that motivate walking and cycling trips, providing a comprehensive foundation for analyzing distance decay and its implications for urban planning and sustainable transportation strategies.

A methodical statistical approach was adopted to determine the appropriate distribution for modeling the distance decay curves for different destinations. The analysis began with generating Cullen and Frey graphs utilizing the fitdistrplus package in R, a graphical tool pivotal for suggesting plausible distributions by plotting the square of skewness against excess kurtosis. This step paved the way for selecting potential distributions for further assessment. Upon identifying potential distributions, an empirical evaluation was conducted to fit several candidate distributions to the data, specifically targeting distributions such as Gamma, Weibull, normal, lognormal, poisson, logistic, and Exponential, among others. Each distribution was fitted using maximum likelihood estimation (MLE), with the added complexity of weights incorporated into the fitting process. This was a crucial step as the dataset included a weight variable, which is a fundamental weighting element used for estimating the frequency of activities conducted by the Canadian population as recorded in the dataset. It aims to portray the broader behavioral patterns of the population accurately.

To objectively assess the fit of each distribution, model selection criteria such as the Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), and log-likelihood (logLik) were employed. These criteria are critical in not only capturing the goodness of fit but also penalizing the complexity of the model to prevent overfitting. AIC and BIC provide a balance between the model's accuracy and complexity, where lower values are indicative of a more economical model. The log-likelihood directly measures the probability of observing the given data under the assumed statistical model, with higher values signaling a better fit. The distribution that manifested the lowest AIC and BIC, coupled with the highest log-likelihood, was deemed the most suitable for representing the distance decay curve for each specific destination in each year. This rigorous statistical procedure ensured that the chosen distribution was the most fitting given the empirical data, reflecting both the underlying process and the observed variations. The approach guarantees that the final model is not only statistically robust but also reliable for interpreting the distance decay effect within the study context. Despite initial considerations for traditional diagnostic methods such as residual analysis and Q-Q plots, these techniques proved incompatible with the complexities of weighted data. The variable "WGHT_EPI" in the dataset weighting variable is essential for representing the Canadian population's activity frequencies at the episode level. While the fitdistrplus package accommodates weights in estimation, it does not extend this functionality to diagnostic plots, which are typically unweighted. Consequently, the reliance on information criteria like AIC and BIC, adjusted for weighted data, became instrumental in guiding our selection of the most robust and representative model for our analysis, marking a methodological stride in our statistical practice.

Following the detailed statistical analysis, we present four comprehensive tables that encapsulate AIC, BIC, and log-likelihood values of each destination by mode and year (Tables \ref{tab:ch03-table-04},\ref{tab:ch03-table-05}, \ref{tab:ch03-table-07}, and \ref{tab:ch03-table-08}). These tables summarize the fitting process for each distribution evaluated, illuminating the decision-making process for selecting the most suitable distribution for each destination category. The tables illustrate our model selection process, highlighting the distribution that emerged as the best fit for each specific context based on the lowest AIC and BIC values and the highest log-likelihood scores. The first table delineates the statistical metrics for walking trips, while the second table deals with trips made by bicycle. Each table is organized to showcase the comparison across different years, thus offering insights into the temporal dynamics of travel behavior and the evolution of distance decay characteristics. By presenting the diagnostics in this tabular format, we aim to provide a clear and accessible reference demonstrating the distributional trends and patterns discerned from the GSS dataset.

```{r ch03-make-table-04, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Filter the dataframe to only include specific destinations
desired_destinations <- c("Home", "work or school", "Other's home")
walking_data_filtered <- walking_data[walking_data$Destinations %in% desired_destinations,]

# Round the numeric columns to two decimal places
numeric_columns <- c("logLik_1986", "AIC_1986", "BIC_1986",
                     "logLik_1992", "AIC_1992", "BIC_1992", 
                     "logLik_1998", "AIC_1998", "BIC_1998")
walking_data_filtered[numeric_columns] <- round(walking_data_filtered[numeric_columns], 2)

library(dplyr)
library(tidyr)
library(kableExtra)

walking_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                ends_with("_1986") | ends_with("_1992") | ends_with("_1998"), 
                -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE,
        caption = "\\label{tab:ch03-table-04}Impedance functions and selection criteria across destinations for walking trips (1986, 1992 and 1998)") |>
  kable_styling(font_size = 12) |> 
  pack_rows("Destination: Home", 1, 9) |>
  pack_rows("Destination: Other's home", 10, 21) |>
  pack_rows("Destination: Work or school", 22, 33) |>
  row_spec(1, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(4, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(7, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(10, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(14, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(18, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(22, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(26, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(30, bold = TRUE, italic = TRUE, font_size = 10)

```

```{r ch03-make-table-05, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}

# Filter the dataframe to include the new specific destinations
new_destinations <- c("Home", "work or school", "Other's home", "Restaurant")
walking_data_filtered <- walking_data[walking_data$Destinations %in% new_destinations,]

# Round the numeric columns to two decimal places for the years 2005, 2010, and 2015
numeric_columns <- c("logLik_2005", "AIC_2005", "BIC_2005",
                     "logLik_2010", "AIC_2010", "BIC_2010",
                     "logLik_2015", "AIC_2015", "BIC_2015")
walking_data_filtered[numeric_columns] <- round(walking_data_filtered[numeric_columns], 2)

walking_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                ends_with("_2005") | ends_with("_2010") | ends_with("_2015"), -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE,
        caption = "\\label{tab:ch03-table-05}Impedance functions and selection criteria across destinations for walking trips (2005, 2010, 2015)") |>
  kable_styling(font_size = 9) |> 
  pack_rows("Destination: Home", 1, 9) |>
  pack_rows("Destination: work or school", 10, 21) |>
  pack_rows("Destination: Other's home", 22, 33) |>
  pack_rows("Destination: Outdoors ", 34, 45) |>
  row_spec(1, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(4, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(7, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(10, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(15, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(18, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(25, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(26, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(30, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(34, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(41, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(42, bold = TRUE, italic = TRUE, font_size = 10)

```

```{r ch03-make-table-06, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}

# Filter the dataframe to include the new specific destinations
new_destinations <- c("Grocery store","Outdoors ", "Library", "Place of worship")
walking_data_filtered <- walking_data[walking_data$Destinations %in% new_destinations,]

# Round the numeric columns to two decimal places for the years 2005, 2010, and 2015
numeric_columns <- c("logLik_2005", "AIC_2005", "BIC_2005",
                     "logLik_2010", "AIC_2010", "BIC_2010",
                     "logLik_2015", "AIC_2015", "BIC_2015")
walking_data_filtered[numeric_columns] <- round(walking_data_filtered[numeric_columns], 2)

walking_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                ends_with("_2005") | ends_with("_2010") | ends_with("_2015"), -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE) |>
  kable_styling(font_size = 9) |> 
  pack_rows("Destination: Grocery store", 1, 9) |>
  pack_rows("Destination: Restaurant", 10, 27) |>
  pack_rows("Destination: Library", 28, 39) |>
  pack_rows("Destination: Place of worship", 40, 54)|>
  row_spec(1, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(4, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(7, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(10, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(16, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(20, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(27, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(28, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(34, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(40, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(45, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(53, bold = TRUE, italic = TRUE, font_size = 10)

```

```{r ch03-make-table-07, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}

## Filter the dataframe to only include specific destinations
desired_destinations <- c("Home", "work or school", "Other's home")
cycling_data_filtered <- cycling_data[cycling_data$Destinations %in% desired_destinations,]

# Round the numeric columns to two decimal places
numeric_columns <- c("logLik_1992", "AIC_1992", "BIC_1992", 
                     "logLik_1998", "AIC_1998", "BIC_1998")
cycling_data_filtered[numeric_columns] <- round(cycling_data_filtered[numeric_columns], 2)

cycling_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                 ends_with("_1992") | ends_with("_1998"), -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE,
        caption = "\\label{tab:ch03-table-07}Impedance functions and selection criteria across destinations for cycling trips (1992 and 1998)") |>
  kable_styling(font_size = 11) |> 
  pack_rows("Destination: Home", 1, 6) |>
  pack_rows("Destination: Other's home", 7, 16) |>
  pack_rows("Destination: Work or school", 17, 26)|>
 row_spec(1, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(4, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(11, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(16, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(17, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(22, bold = TRUE, italic = TRUE, font_size = 10) 
```

```{r ch03-make-table-08, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}

# Filter the dataframe to only include specific destinations
desired_destinations <- c("Home" , "Other's home", "work or school" )
cycling_data_filtered <- cycling_data[cycling_data$Destinations %in% desired_destinations,]

# Round the numeric columns to two decimal places for the years 2005, 2010, and 2015
numeric_columns <- c("logLik_2005", "AIC_2005", "BIC_2005",
                     "logLik_2010", "AIC_2010", "BIC_2010",
                     "logLik_2015", "AIC_2015", "BIC_2015")
cycling_data_filtered[numeric_columns] <- round(cycling_data_filtered[numeric_columns], 2)

cycling_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                 ends_with("_2005") | ends_with("_2010") | ends_with("_2015"), -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE,
        caption = "\\label{tab:ch03-table-08}Impedance functions and selection criteria across destinations for cycling trips (2005, 2010 and 2015)") |>
  kable_styling(font_size = 11) |> 
  pack_rows("Destination: Home", 1, 9) |> 
  pack_rows("Destination: Other's home", 10, 24)|>
  pack_rows("Destination: work or school", 25, 39) |>
  row_spec(1, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(4, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(7, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(10, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(15, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(20, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(25, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(30, bold = TRUE, italic = TRUE, font_size = 10) |>
  row_spec(39, bold = TRUE, italic = TRUE, font_size = 10)

```

```{r ch03-make-table-09, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Filter the dataframe to only include specific destinations
desired_destinations <- c("Outdoors " ,  "Grocery store", "Restaurant" )
cycling_data_filtered <- cycling_data[cycling_data$Destinations %in% desired_destinations,]

# Round the numeric columns to two decimal places for the years 2005, 2010, and 2015
numeric_columns <- c("logLik_2005", "AIC_2005", "BIC_2005",
                     "logLik_2010", "AIC_2010", "BIC_2010",
                     "logLik_2015", "AIC_2015", "BIC_2015")
cycling_data_filtered[numeric_columns] <- round(cycling_data_filtered[numeric_columns], 2)

cycling_data_filtered |> 
  dplyr::select(Destinations, 
                Impedance_Function, 
                 ends_with("_2005") | ends_with("_2010") | ends_with("_2015"), -starts_with("Selected_")) |>
  pivot_longer(cols = -c(Destinations, Impedance_Function, starts_with("Selected")), 
               names_to = c("fun", "year"), 
               names_sep = "_") |>
  pivot_wider(names_from = fun, values_from = value) |>
  arrange(Destinations, year) |>
  dplyr::select(-Destinations) |>
  kable(format = "latex",
        booktabs = TRUE) |>
  kable_styling(font_size = 9) |> 
  pack_rows("Destination: Outdoors ", 1, 18) |>
  pack_rows("Destination: Grocery store", 19, 36)|>
  pack_rows("Destination: Restaurant", 37, 48) |>
  row_spec(1, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(7, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(15, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(19, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(26, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(38, bold = TRUE, italic = TRUE, font_size = 9) |>
  row_spec(45, bold = TRUE, italic = TRUE, font_size = 9) 
```

\newpage

The figures below illustrate the frequency distribution of travel times for walking and cycling to various destinations, as modeled by impedance functions over a range of years. Each graph shows a curve representing the scaled frequency of trips against the trip duration in minutes for a specific year, with the peak of the curve indicating the most common trip duration. The red dots and dashed lines highlight the mode of the distribution, which is the most frequently observed trip duration. The multiple curves within each year show the variability or sensitivity of the model under different assumptions or parameters.
From these figures, we can infer trends in travel behavior for walking and cycling over time. For instance, a shift in the peak towards the right over the years would suggest an increase in the average duration of trips, possibly due to several factors such as urban sprawl, changes in urban design, or evolving societal preferences. Conversely, a shift towards the left would suggest a decrease in trip duration, potentially indicating improved walking and cycling infrastructure, leading to shorter and possibly more convenient trips.
These figures encapsulate a wealth of information about travel behavior and serve as a visual summary of the underlying data collected for walking and cycling trips to different destinations. They provide a snapshot of the changes in travel time preferences or requirements across the studied years, which is valuable for urban planners and policymakers aiming to understand and promote sustainable modes of transportation.

```{r ch03-make-impedance_graph_walking_home ,cache=FALSE, message=FALSE, echo=FALSE, include=FALSE}

walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

home_peaks <- walking_data_f %>%
  filter(dest_label == "Home") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

home_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Home"), aes(group = YEAR, colour = "Home"), linewidth = 1.2) +
  geom_segment(data = home_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = home_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = home_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for home destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Home" = "cyan3", "Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-impedance_graph_cycling_home ,cache=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# Assuming cycling_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Home' destination for each year
home_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Home") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
home_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Home"), aes(group = YEAR, colour = "Home"), linewidth = 1.2) +
  geom_segment(data = home_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = home_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = home_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for home destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # This will be the title of the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Home" = "cyan3", "Other destinations" = "grey40", "Peaks" = "red"))
```

```{r ch03-make-fig-03, cache = FALSE, include=FALSE}
ch03_fig_03 <- cowplot::plot_grid(home_plot,home_c_plot, nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_03.jpg", 
       plot = ch03_fig_03,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-03, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to home destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_03.jpg")
```

Figure \ref{fig:ch03-plot-fig-03} illustrates the shifting patterns in walking trip behavior to home destinations over several years. The red dashed lines and points highlight the peak travel times that individuals are inclined to walk from their origins to home. A discernible trend of the peak point moving toward shorter durations indicates a diminishing willingness to engage in longer walking trips. For instance, the peak travel time in 1986 was around 10 minutes, which then steadily reduced in the following years, with the 2015 data showing the peak remaining at approximately 10 minutes. The walking trip graphs consistently show that the frequency of trips declines sharply with increased travel time, a characteristic of non-motorized transportation. The leftward shift in peak frequency points to a growing reluctance to extend walking durations, potentially due to factors such as urban expansion, increased use of private vehicles, enhancements in public transportation, or shifts in societal norms and lifestyle choices.

Conversely, Figure \ref{fig:ch03-plot-fig-03} captures the progression of cycling trip behavior for home trips in Canada from 1992 to 2015. Unlike walking trips, the cycling data reveals an elongation in the peak travel time. In 1992, the average travel time was approximately five minutes. This number has gradually increased over time, with a peak of 15 minutes in 2015. This ascending trend underscores a heightened propensity to cycle for longer periods when returning home. The motivation behind this shift could be multifaceted, encompassing enhancements in cycling infrastructure, a heightened emphasis on physical fitness and environmental sustainability, and the gradual adoption of cycling as a favored means of transportation for increasingly longer distances. Walking trip patterns significantly differ from the cycling pattern, reflecting an expanding tolerance for extended travel times. This discrepancy might result from the inherent distinctions between the two forms of transportation. Because cycling is a faster mode of transportation, it has probably benefited from better pathways and social developments that support active travel. This has allowed cyclists to cover longer distances more comfortably and easily. The distribution of cycling trip frequencies, while still demonstrating a decrease as travel time lengthens, presents a less pronounced decline compared to walking trips. This suggests a wider acceptance of longer cycling durations. The progressive shift in peak travel times over the years could be attributed to amplified investments in cycling infrastructure, such as the introduction of bike lanes and the proliferation of bike-sharing initiatives, coupled with an overall rise in cycling's popularity.

```{r ch03-make-impedance_graph_walking_work, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'work or school' destination for each year
work_w_peaks <- walking_data_f %>%
  filter(dest_label == "Work or school") %>%
  group_by(YEAR) %>%
  summarize(
    peak_w_duration = DURATION[which.max(f_t)],
    peak_w_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
work_w_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Work or school"), aes(group = YEAR, colour = "Work or school"), linewidth = 1.2) +
  geom_segment(data = work_w_peaks, aes(x = peak_w_duration, y = peak_w_ft, xend = peak_w_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = work_w_peaks, aes(x = peak_w_duration, y = peak_w_ft, colour = "Peaks"), size = 3) +
  geom_text(data = work_w_peaks, aes(x = peak_w_duration, y = 0, label = peak_w_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for work or school destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # This will be the title of the legend
  ) +
  theme_minimal(base_size = 12) +
  scale_colour_manual(values = c("Work or school" = "deeppink2", "Other destinations" = "grey40", "Peaks" = "red"),
                      labels = c("Library", "Other", "Peaks")) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Work or school" = "deeppink2","Other destinations" = "grey40",  "Peaks" = "red"))
```

```{r ch03-make-impedance_graph_cycling_work, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'work or school' destination for each year
work_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Work or school") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
work_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Work or school"), aes(group = YEAR, colour = "Work or school"), linewidth = 1.2) +
  geom_segment(data = work_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = work_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = work_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for work or school destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # This will be the title of the legend
  ) +
  theme_minimal(base_size = 12) +
  scale_colour_manual(values = c("Work or school" = "deeppink2", "Other destinations" = "grey40", "Peaks" = "red"),
                      labels = c("Library", "Other", "Peaks")) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Work or school" = "deeppink2","Other destinations" = "grey40",  "Peaks" = "red"))
```

```{r ch03-make-fig-04, cache = FALSE, include=FALSE}
ch03_fig_04 <- cowplot::plot_grid(work_w_plot,work_c_plot, nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_04.jpg", 
       plot = ch03_fig_04,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-04, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_04.jpg")
```

Figure \ref{fig:ch03-plot-fig-04} captures the evolution of walking trip behavior to work or school destinations over selected years. The graphs reveal an interesting trend: while there is some fluctuation, there is a general decrease in the peak travel time duration. For instance, in 1986, the peak travel time was approximately 10 minutes; by 2015, it had reduced slightly, suggesting a preference for shorter walking commutes. This could indicate an adjustment in urban layouts that brings residences closer to work and school locations or a shift in employment or education patterns that favors proximity. The graphs show a sharp decrease in the frequency of walking trips as the travel time increases, indicating a strong preference for shorter commutes. The decreasing peak frequency over the years could reflect improved access to alternative modes of transport or an increase in the use of private vehicles, which could reduce the need for longer walking trips.

Figure \ref{fig:ch03-plot-fig-04} presents the trends for cycling trips to work or school destinations. In contrast to walking, the peak travel times for cycling vary over the years. In 1992, the most frequent travel time was nearly 10 minutes, but by 2015, the peak shifted to around 15 minutes. This indicates a growing acceptance or necessity of cycling for slightly longer durations to reach work or school. The increase in peak travel times might be associated with several factors, including expanding city boundaries, making longer commutes more common, or improving cycling infrastructure that make longer distances more feasible and safe for cyclists. For both walking and cycling, the frequency of trips decreases markedly as the duration increases. However, cycling trips show a broader acceptance for longer durations, likely due to the increased speed and efficiency of travel by bike compared to walking. The data suggests a social shift towards embracing cycling for daily commutes, which investments in cycling-friendly city planning could support, the availability of bike-sharing programs, and initiatives to encourage active transportation for health and environmental benefits.

```{r ch03-make-impedance_graph_walking_other, cache=FALSE,include=FALSE, message=FALSE, echo=FALSE}
# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Other's home' destination for each year
other_peaks <- walking_data_f %>%
  filter(dest_label == "Other's home") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
other_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Other's home"), aes(group = YEAR, colour = "Other's home"), linewidth = 1.2) +
  geom_segment(data = other_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = other_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = other_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for other's home destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # This will be the title of the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Other's home" = "darkorchid3","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-impedance_graph_cycling_other, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}

# Assuming cycling_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Other's home' destination for each year
other_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Other's home") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
other_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Other's home"), aes(group = YEAR, colour = "Other's home"), linewidth = 1.2) +
  geom_segment(data = other_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = other_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = other_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for other's home destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Other's home" = "darkorchid3","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-fig-05, cache = FALSE, include=FALSE}
ch03_fig_05 <- cowplot::plot_grid(other_plot ,other_c_plot , nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_05.jpg", 
       plot = ch03_fig_05,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-05, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_05.jpg")
```

Figure \ref{fig:ch03-plot-fig-05} maps the trends in walking trips to another person's home across a span of years. In 1986, the peak was observed at a travel time of roughly 10 minutes, suggesting a preference for short walking trips. Furthermore, most walking trips were completed in under 50 minutes during these years, emphasizing a strong inclination towards relatively short distances.
There are noticeable fluctuations in the peak travel times from 1986 to 2015. For example, in 2005, the peak duration for walking trips to another's home contracts was less than 2 minutes, indicating a shift towards even shorter trips compared to previous years. This shift could reflect changes in urban design, such as more compact neighborhoods or increased use of other transportation modes. By 2010, the willingness to walk extended to trips under 40 minutes for all purposes, signaling a slight increase in the acceptance of longer walks. However, by 2015, the data shows a return to shorter durations for the peak times, with the most common walking trips hovering around 10 minutes again. Turning to cycling, Figure \ref{fig:ch03-plot-fig-05} reveals the evolution of cycling trip behavior to another's home. In contrast to walking, cycling trips exhibit a gradual increase in the peak travel time. From an average of approximately 5 minutes in 1992, there is a discernible upward trend, with trips averaging less than 25 minutes in 1998 and rising further to an average of less than 50 minutes by 2015. It suggests not only an increasing willingness to cycle for longer periods but also possibly a rise in the number of destinations becoming accessible via cycling, reflecting improvements in cycling infrastructure or a growing cultural trend towards embracing cycling for various trip purposes. In both walking and cycling modes, the trend towards shorter trips dominates, with a marked decrease in trip frequency as duration increases. Nonetheless, the data on cycling suggests a growing acceptance of cycling as a viable mode of transport for longer distances over the years, likely due to its efficiency and the positive influence of environmental sustainability movements that promote cycling.

```{r ch03-make-impedance_graph_walking_grocery, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming walking_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Grocery store, other stores or mall' destination for each year
grocery_peaks <- walking_data_f %>%
  filter(dest_label == "Grocery store, other stores or mall") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
grocery_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Grocery store, other stores or mall"), aes(group = YEAR, colour = "Grocery store"), linewidth = 1.2) +
  geom_segment(data = grocery_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = grocery_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = grocery_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for Grocery store, other stores or mall destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Grocery store" = "red3","Other destinations" = "grey40", "Peaks" = "red"))
```

```{r ch03-make-impedance_graph_cycling_grocery, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming cycling_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Grocery store, other stores or mall' destination for each year
grocery_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Grocery store, other stores or mall") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
grocery_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Grocery store, other stores or mall"), aes(group = YEAR, colour = "Grocery store"), linewidth = 1.2) +
  geom_segment(data = grocery_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = grocery_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = grocery_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for Grocery store, other stores or mall destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Grocery store" = "red3","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-fig-06, cache = FALSE, include=FALSE}
ch03_fig_06 <- cowplot::plot_grid(grocery_plot  ,grocery_c_plot  , nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_06.jpg", 
       plot = ch03_fig_06,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-06, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_06.jpg")
```

The distance decay curves in Figure \ref{fig:ch03-plot-fig-06} and Figure \ref{fig:ch03-plot-fig-06} reflect the frequency of walking and cycling trips to grocery stores, other stores, or malls for the years where data is available, specifically 2005, 2010, and 2015. Analyzing the distance decay curves provided in Figure \ref{fig:ch03-plot-fig-06} for walking trips, we see a clear evolution in consumer behavior over a decade. In 2005, the peak frequency of walking trips to shopping destinations was approximately 4 minutes, reflecting a preference for very close proximity to these destinations. However, by 2010, the peak sharply decreases to a travel time of around 2 minutes, indicating an even stronger preference for shopping destinations within a very short walking distance. Notably, in 2015, the data illustrates a reversal in this trend, with the peak frequency rising to around 10 minutes. This increase suggests that by 2015, individuals were more inclined to walk longer distances to reach grocery stores, other stores, or malls, which could be due to changes in urban development or a renewed interest in walking as a form of exercise and sustainable travel.

For cycling trips, shown in Figure \ref{fig:ch03-plot-fig-06}, we witness a different pattern of change. The peak frequency in 2005 shows that the most common travel time for cycling to shopping destinations was around 5 minutes, suitable for quick and convenient trips. This duration represents a mode of transport that is efficient for nearby stores but utilized less for farther distances. Moving forward to 2015, we observe a dramatic increase in the peak travel time to around 10 minutes. This significant rise indicates a marked shift towards the willingness to cycle longer to reach shopping destinations.
In summary, the data from the provided graphs point to a change in the travel patterns of consumers over the years. While walking trips for shopping initially showed a trend towards extremely short distances, there was a notable shift towards longer durations by 2015. Conversely, cycling trips have expanded in terms of acceptable travel time, with a marked increase in the duration of trips by 2015, reflecting a broader acceptance and possibly encouragement of cycling as a preferred mode of transportation for shopping trips.

```{r ch03-make-impedance_graph_walking_outdoors, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming walking_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'outdoors' destination for each year
outdoors_peaks <- walking_data_f %>%
  filter(dest_label == "Outdoors") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
outdoors_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Outdoors"), aes(group = YEAR, colour = "Outdoors"), linewidth = 1.2) +
  geom_segment(data = outdoors_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = outdoors_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = outdoors_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for outdoors destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Outdoors" = "royalblue2","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-impedance_graph_cycling_outdoors, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming cycling_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'outdoors' destination for each year
outdoors_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Outdoors") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE),
    .groups = "drop" # Dropping groups after summarization
  )

# Plot the data
outdoors_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Outdoors"), aes(group = YEAR, colour = "Outdoors"), linewidth = 1.2) +
  geom_segment(data = outdoors_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = outdoors_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = outdoors_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for outdoors destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Outdoors" = "royalblue2","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-fig-07, cache = FALSE, include=FALSE}
ch03_fig_07 <- cowplot::plot_grid(outdoors_plot,outdoors_c_plot  , nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_07.jpg", 
       plot = ch03_fig_07,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-07, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_07.jpg")
```
In Figure \ref{fig:ch03-plot-fig-07}, for walking trips to outdoor destinations, the blue line in the 2005 graph exhibits a peak at approximately 3 minutes, indicating that the most common duration of walking trips was very short. This peak suggests that outdoor destinations were typically close to the starting point or that there was a preference for quick, accessible outdoor activities. In 2010, we saw a similar pattern with the peak around the 3-minute mark, demonstrating a consistent preference for short walking trips over time. By 2015, the peak shifts to around 10 minutes. This increase in the most common trip duration may reflect a change in the location of outdoor spaces relative to residential areas.Turning to cycling (Figure \ref{fig:ch03-plot-fig-07}), the blue line in the 2005 graph peaks at about 10 minutes. This suggests that the typical cycling trip to outdoor destinations was a moderate distance away, indicating that cycling was used for slightly longer but still relatively brief excursions. Across both walking and cycling trips to outdoor destinations, the observed peaks in the blue lines indicate a general trend of short-duration trips, with a notable increase in the willingness to travel longer distances by bike in recent years.

```{r ch03-make-impedance_graph_walking_restaurant, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming walking_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Restaurant, bar or club' destination for each year
restaurant_peaks <- walking_data_f %>%
  filter(dest_label == "Restaurant, bar or club") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

restaurant_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Restaurant, bar or club"), aes(group = YEAR, colour = "Restaurant"), linewidth = 1.2) +
  geom_segment(data = restaurant_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = restaurant_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = restaurant_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for Restaurant, bar or club destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Restaurant" = "yellow2","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-impedance_graph_cycling_restaurant, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}

# Assuming cycling_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
cycling_data_f <- cycling_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Restaurant, bar or club' destination for each year
restaurant_c_peaks <- cycling_data_f %>%
  filter(dest_label == "Restaurant, bar or club") %>%
  group_by(YEAR) %>%
  summarize(
    peak_c_duration = DURATION[which.max(f_t)],
    peak_c_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
restaurant_c_plot <- ggplot(cycling_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = cycling_data_f %>% filter(dest_label == "Restaurant, bar or club"), aes(group = YEAR, colour = "Restaurant"), linewidth = 1.2) +
  geom_segment(data = restaurant_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, xend = peak_c_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = restaurant_c_peaks, aes(x = peak_c_duration, y = peak_c_ft, colour = "Peaks"), size = 3) +
  geom_text(data = restaurant_c_peaks, aes(x = peak_c_duration, y = 0, label = peak_c_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for Restaurant, bar or club destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Restaurant" = "yellow2","Other destinations" = "grey40", "Peaks" = "red"))
```

```{r ch03-make-fig-08, cache = FALSE, include=FALSE}
ch03_fig_08 <- cowplot::plot_grid(restaurant_plot,restaurant_c_plot  , nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_08.jpg", 
       plot = ch03_fig_08,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-08, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_08.jpg")
```
In the yellow line graph for 2005 (Figure \ref{fig:ch03-plot-fig-08}), there is a distinct peak at around 3 minutes, indicating that the most common duration for walking trips to these social destinations was very short. This peak suggests that such venues were often chosen within proximity to the starting point, reflecting a preference for convenience and accessibility when deciding to walk to restaurants, bars, or clubs. The 2015 graph reveals a more pronounced change, with the peak frequency occurring around 10 minutes. This suggests an increased willingness to walk further distances to these destinations compared to previous years, possibly due to changes in the locations of such venues or a shift towards healthier lifestyle choices that include more walking. In  Figure \ref{fig:ch03-plot-fig-08}, for cycling trips in 2005, the peak in the yellow line graph is at approximately 15 minutes, indicating a preference for moderately distant venues when cycling. This duration suggests that cyclists were prepared to travel further than walkers to reach restaurants, bars, or clubs, possibly due to the speed and convenience of cycling over longer distances. The 2015 graph shows a significant shift in the peak frequency to around 15 minutes, which points to a clear trend of individuals being willing to cycle for longer periods to social destinations. Across the years, the yellow line graphs for walking and cycling to restaurants, bars, or clubs show a general pattern of trips most frequent at short to moderate durations.

```{r ch03-make-impedance_graph_walking_place, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming walking_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Place of worship' destination for each year
place_peaks <- walking_data_f %>%
  filter(dest_label == "Place of worship") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
place_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Place of worship"), aes(group = YEAR, colour = "Place of worship"), linewidth = 1.2) +
  geom_segment(data = place_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = place_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = place_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for place of worship destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Place of worship" = "darkorange1","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-fig-09, cache = FALSE, include=FALSE}
ch03_fig_09 <- place_plot

ggsave(file = "figure/ch03_fig_09.jpg", 
       plot = ch03_fig_09,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-09, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations"}

knitr::include_graphics("figure/ch03_fig_09.jpg")
```

Figure \ref{fig:ch03-plot-fig-09} presents the frequency of walking trips to places of worship in 2005 and 2015. In 2005, the peak frequency occurred at approximately 5 minutes, demonstrating that the most common duration for walking trips to places of worship was short. This implies that, typically, such destinations were located within close proximity to individuals' origin points, or there was a preference for accessing nearby places of worship. By 2015, there is a noticeable shift in the peak frequency to around 10 minutes, indicating a change in the walking duration to places of worship over the decade. This increase suggests that people were willing to walk a bit longer to reach their places of worship, or it could indicate that these places have become more spread out within communities, requiring longer walk times. The absence of data for the intervening years and cycling trips leaves a gap in understanding the trend over time. However, the available data from 2005 to 2015 indicates a trend of increased walking time for visits to places of worship. This could reflect changes in urban development, with places of worship being farther from residential areas or a shift in individual preferences for walking as part of their religious or spiritual practices. It is important to note that there is no data on cycling trips to places of worship destinations, which limits our ability to analyze and compare the full spectrum of non-motorized travel to places of worship.

```{r ch03-make-impedance_graph_walking_library, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE}
# Assuming walking_data_f is already loaded and has 'YEAR', 'DURATION', 'dest_label', and 'f_t' columns

# Filter out NA values if necessary
walking_data_f <- walking_data_f %>% filter(!is.na(f_t))

# Find the peaks for 'Library, museum or theatre' destination for each year
library_peaks <- walking_data_f %>%
  filter(dest_label == "Library, museum or theatre") %>%
  group_by(YEAR) %>%
  summarize(
    peak_duration = DURATION[which.max(f_t)],
    peak_ft = max(f_t, na.rm = TRUE)
  ) %>%
  ungroup()

# Plot the data
library_plot <- ggplot(walking_data_f, aes(x = DURATION, y = f_t)) +
  geom_line(aes(group = interaction(dest_label, YEAR), colour = "Other destinations"), linewidth = 0.5) +
  facet_wrap(~YEAR, scales = "free_y") +
  geom_line(data = walking_data_f %>% filter(dest_label == "Library, museum or theatre"), aes(group = YEAR, colour = "Library"), linewidth = 1.2) +
  geom_segment(data = library_peaks, aes(x = peak_duration, y = peak_ft, xend = peak_duration, yend = 0, colour = "Peaks"), linetype = "dashed", linewidth = 0.5) +
  geom_point(data = library_peaks, aes(x = peak_duration, y = peak_ft, colour = "Peaks"), size = 3) +
  geom_text(data = library_peaks, aes(x = peak_duration, y = 0, label = peak_duration), color = "red", size = 3, hjust = 1.5, vjust = -0.5) +
  labs(
    title = "Modeling distance decay curves of walking for Library, museum or theatre destination",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    colour = "Legend"  # Title for the legend
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.margin = unit(c(1, 1, 2, 1), "lines"),
    axis.line = element_line(color = "grey40"),
    axis.text.x = element_text(color = "grey40", size = 10),
    axis.text.y = element_text(color = "grey40", size = 10),
    plot.title = element_text(hjust = 0.5, size = 14),
    strip.text = element_text(color = "grey40"),
    panel.grid.major = element_line(color = "grey85", size = 0.2),
    panel.grid.minor = element_line(color = "grey85", size = 0.1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    panel.spacing = unit(0.5, "lines"),
    aspect.ratio = 1/2
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  scale_colour_manual(values = c("Library" = "olivedrab4","Other destinations" = "grey40", "Peaks" = "red"))

```

```{r ch03-make-fig-010, cache = FALSE, include=FALSE}
ch03_fig_10 <- library_plot

ggsave(file = "figure/ch03_fig_10.jpg", 
       plot = ch03_fig_10,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-10, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations"}

knitr::include_graphics("figure/ch03_fig_10.jpg")
```

Figure \ref{fig:ch03-plot-fig-10} illustrates the walking trip frequency to libraries, museums, or theaters for 2005 and 2015, as indicated by the green line curves. In 2005, the peak frequency of walking trips to these cultural destinations was approximately 7 minutes. This suggests that such venues were commonly located within a short walking distance from the trip origin, or there was a preference for nearby cultural activities that could be quickly accessed by foot. By 2015, we see a significant shift in behavior, with the peak frequency now occurring at around 10 minutes. This change over the decade may reflect a broader willingness to walk slightly longer distances to reach libraries, museums, or theaters. It is essential to note the absence of cycling data trips, which indicates a lack of available information regarding the frequency of cycling trips to these destinations.

The distance decay curves for walking (Figure \ref{fig:ch03-plot-fig-12} and cycling (Figure \ref{fig:ch03-plot-fig-14}) to different destinations and for different years from 1986 to 2015 were analyzed. The analysis revealed that the shape of the distance decay curves varied depending on the destination and the mode. The destinations for which these functions were estimated vary across different years, modes of transportation (walking and cycling), and the available data. For instance, in 2005, 2010, and 2015, the walking destinations encompassed a wide range, including home, work or school, other people’s homes, grocery stores, other retail stores or malls, outdoor spaces, restaurants, bars or clubs, libraries, museums, theaters, and places of worship. In contrast, the destinations for 1986, 1992, and 1998 were limited to just three categories: home, other people’s homes, and work or school.

```{r ch03-make-impedance_walking_2015, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 2015 only
walking_data_2015 <- walking_data_f %>%
  filter(YEAR == 2015)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_2015 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015 with dashed lines and size 0.8
walking_plot_2015 <- ggplot(walking_data_2015, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line( size = 0.8) +  # Set lines to dashed and size to 0.8
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 2015",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  ) +
  scale_x_continuous(breaks = seq(0, max(walking_data_2015$DURATION, na.rm = TRUE), by = 10)) 
```

```{r ch03-make-impedance_walking_2010, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 2010 only
walking_data_2010 <- walking_data_f %>%
  filter(YEAR == 2010)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_2010 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2010
walking_plot_2010 <- ggplot(walking_data_2010, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 2010",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")    
  ) +
  scale_x_continuous(breaks = seq(0, max(walking_data_2010$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-impedance_walking_2005, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 2005 only
walking_data_2005 <- walking_data_f %>%
  filter(YEAR == 2005)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_2005 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015
walking_plot_2005 <- ggplot(walking_data_2005, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
   geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 2005",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  )  +
  scale_x_continuous(breaks = seq(0, max(walking_data_2005$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-impedance_walking_1998, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 1998 only
walking_data_1998 <- walking_data_f %>%
  filter(YEAR == 1998)

# Define the colors for each destination
dest_colors <- c(
 "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_1998 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015
walking_plot_1998 <- ggplot(walking_data_1998, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
   geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 1998",
    x = "Duration (minutes)",
    y = "Frequency (scaled)",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")   
  ) +
  scale_x_continuous(breaks = seq(0, max(walking_data_1998$DURATION, na.rm = TRUE), by = 10))

```

```{r ch03-make-impedance_walking_1992, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 1998 only
walking_data_1992 <- walking_data_f %>%
  filter(YEAR == 1992)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_1992 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 1992
walking_plot_1992 <- ggplot(walking_data_1992, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 1992",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")     
  ) +
  scale_x_continuous(breaks = seq(0, max(walking_data_1992$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-impedance_walking_1986, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 1986 only
walking_data_1986 <- walking_data_f %>%
  filter(YEAR == 1986)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- walking_data_1986 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 1986
walking_plot_1986 <- ggplot(walking_data_1986, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of walking for 1986",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "gray10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  ) +
  scale_x_continuous(breaks = seq(0, max(walking_data_1986$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-fig-11, cache = FALSE, include=FALSE}
ch03_fig_11 <- cowplot::plot_grid(walking_plot_1986, walking_plot_1992, walking_plot_1998, nrow = 3, align = 'v')

ggsave(file = "figure/ch03_fig_11.jpg", 
       plot = ch03_fig_11,
       width = 8.5,  
       height = 11,  
       dpi = 300,
       device = 'jpg')

ch03_fig_12 <- cowplot::plot_grid( walking_plot_2005, walking_plot_2010, walking_plot_2015, nrow = 3, align = 'v')

ggsave(file = "figure/ch03_fig_12.jpg", 
       plot = ch03_fig_12,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-11, echo=FALSE, cache=FALSE, out.width="100%", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_11.jpg")
```

```{r ch03-plot-fig-12, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking trips to various destinations", fig.align="center"}
knitr::include_graphics("figure/ch03_fig_12.jpg")
```

Between 1986 and 1998, the highest frequency of walking trips occurred within the 5-10 minute range. Subsequently, from 1998 to 2010, the peak shifted to the 0-5 minute range, only to return to the 5-10 minute range during the period from 2010 to 2015. Based on the findings presented in Figures \ref{fig:ch03-plot-fig-12}, it is evident that in 2015, the majority of walking trips had a duration of less than 50 minutes. The graphical representations of walking trip patterns exhibit similar trends for destinations such as home, work or school, other people’s homes, and grocery stores. However, when it comes to recreational destinations like outdoor locations, restaurants, bars or clubs, places of worship, and libraries, individuals tend to walk for a shorter duration, typically around
40 minutes. This observation underscores the discretionary nature of entertainment-related trips, which often do not adhere to strict time constraints within an individual’s daily activity schedule. In 2010, individuals exhibited a willingness to engage in walking trips of less than 40 minutes duration, regardless of their intended purposes. However, this trend experienced a decline in 2005, with the average walking trip duration decreasing to approximately less than 30 minutes for all purposes. Furthermore, between 1986 and 1998, there was a notable decrease in people’s willingness to engage
in walking trips to reach destinations such as their own homes, others’ homes, work, and school. Specifically, the average walking duration decreased from 70 minutes in 1986 to 40 minutes in 1998.

```{r ch03-make-impedance_cycling_2015, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 2015 only
cycling_data_2015 <- cycling_data_f %>%
  filter(YEAR == 2015)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- cycling_data_2015 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015
cycling_plot_2015 <- ggplot(cycling_data_2015, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for 2015",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  )  +
  scale_x_continuous(breaks = seq(0, max(cycling_data_2015$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-impedance_cycling_2010, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the walking_data for the year 2010 only
cycling_data_2010 <- cycling_data_f %>%
  filter(YEAR == 2010)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- cycling_data_2010 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015
cycling_plot_2010 <- ggplot(cycling_data_2010, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for 2010",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  # Set legend text properties
    legend.title = element_text(size = 10, color = "grey10"), # Set legend title properties
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  ) +
  scale_x_continuous(breaks = seq(0, max(cycling_data_2010$DURATION, na.rm = TRUE), by = 10))  
```

```{r ch03-make-impedance_cycling_2005, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the cycling_data for the year 2005 only
cycling_data_2005 <- cycling_data_f %>%
  filter(YEAR == 2005)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- cycling_data_2005 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plotting code with modified x-axis breaks and added vertical lines
cycling_plot_2005 <- ggplot(cycling_data_2005, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  scale_x_continuous(breaks = seq(0, 90, by = 10)) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for 2005",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),
    legend.title = element_text(size = 10, color = "grey10"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10") 
  ) +
  scale_x_continuous(breaks = seq(0, max(cycling_data_2005$DURATION, na.rm = TRUE), by = 10))

```

```{r ch03-make-impedance_cycling_1998, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}
# Filter the cycling_data for the year 1998 only
cycling_data_1998 <- cycling_data_f %>%
  filter(YEAR == 1998)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- cycling_data_1998 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 2015
cycling_plot_1998 <- ggplot(cycling_data_1998, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  scale_x_continuous(breaks = seq(0, 90, by = 10)) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for 1998",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA),
    axis.line = element_line(color = "grey10")  
  )+
  scale_x_continuous(breaks = seq(0, max(cycling_data_1998$DURATION, na.rm = TRUE), by = 10))
```

```{r ch03-make-impedance_cycling_1992, echo=FALSE,include=FALSE, message=FALSE, warning=FALSE}

# Filter the cycling_data for the year 1992 only
cycling_data_1992 <- cycling_data_f %>%
  filter(YEAR == 1992)

# Define the colors for each destination
dest_colors <- c(
  "Place of worship" = "darkorange1",
  "Outdoors" = "royalblue2",
  "Library, museum or theatre" = "olivedrab4",
  "Grocery store, other stores or mall" = "red3",
  "Restaurant, bar or club" = "yellow2",
  "Other's home" = "darkorchid3",
  "Work or school" = "deeppink2",
  "Home" = "cyan3"
)

# Calculate the peaks for each group
peaks <- cycling_data_1992 %>%
  group_by(dest_label) %>%
  summarise(peak_duration = DURATION[which.max(f_t)], peak_f_t = max(f_t))

# Plot the data for the year 1992
cycling_plot_1992 <- ggplot(cycling_data_1992, aes(x = DURATION, y = f_t, group = dest_label, color = dest_label)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = dest_colors) +
  geom_segment(data = peaks, aes(x = peak_duration, y = 0, xend = peak_duration, yend = peak_f_t, color = dest_label),
               linetype = "dashed", size = 0.5) +
  labs(
    title = "Modeling distance decay curves of cycling for 1992",
    x = "Duration (minutes)",
    y = "Frequency",
    color = "Destination"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, color = "grey10"),
    axis.title = element_text(color = "grey10"),
    axis.text = element_text(color = "grey10"),
    legend.text = element_text(size = 10, color = "grey10"),  
    legend.title = element_text(size = 10, color = "grey10"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", colour = NA) ,
    axis.line = element_line(color = "grey10")  
  ) +
  scale_x_continuous(breaks = seq(0, max(cycling_data_1992$DURATION, na.rm = TRUE), by = 10))

```

```{r ch03-make-fig-13, cache = FALSE, include=FALSE}
ch03_fig_13 <- cowplot::plot_grid(cycling_plot_1992,cycling_plot_1998, nrow = 2, align = 'v')

ggsave(file = "figure/ch03_fig_13.jpg", 
       plot = ch03_fig_13,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')

ch03_fig_14 <- cowplot::plot_grid(cycling_plot_2005 , cycling_plot_2010, cycling_plot_2015, nrow = 3, align = 'v')

ggsave(file = "figure/ch03_fig_14.jpg", 
       plot = ch03_fig_14,
       width = 8.5,  # Width in inches, adjust as needed for your specific layout
       height = 11,  # Height in inches, adjust as needed for your specific layout
       dpi = 300,
       device = 'jpg')
```

```{r ch03-plot-fig-13, echo=FALSE, cache=FALSE, out.width="100%", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_13.jpg")
```

```{r ch03-plot-fig-14, echo=FALSE, cache=FALSE, out.width="100%", fig.cap="Modeling distance decay curves of walking and cycling to work or school destinations", fig.align="center"}

knitr::include_graphics("figure/ch03_fig_14.jpg")
```

Apart from walking trips, cycling trips have shown variations in their choice of destinations between 1992 and 2015 (Figure \ref{fig:ch03-plot-fig-14}. However, due to the limited availability of data for cycling trips across different years, it was not feasible to calculate the impedance function for various destination categories. Consequently, we derived the impedance function for specific destinations, namely home, other people’s homes, work or school, grocery stores, restaurants or bars, and outdoor locations, for the years 2005, 2010, and 2015. For the years 1992 and 1998, the impedance function was specifically computed for destinations such as home, other people’s homes, and work or school. It’s essential to highlight that data regarding cycling trips for the year 1986 was not accessible for analysis. In 1992, individuals exhibited a notable willingness to engage in cycling trips for an approximate duration of 90 minutes when traveling to their residences. This indicates that during that period, people were more inclined to use cycling for longer commutes to their homes. In contrast, for destinations such as others’ homes and work or school,cycling was favored for shorter-distance trips to these destinations, possibly due to convenience or the availability of cycling infrastructure. By 1998, the inclination to cycle for all purposes reduced to around 70 minutes. This suggests a shift in preferences towards slightly shorter cycling trips. During this time, the willingness to cycle to reach various destinations appeared to converge, possibly indicating a more balanced use of cycling for different trip types.

Furthermore, in 2005, individuals tended to cycle for approximately 70 minutes to reach their homes and work or school, while the duration for other destinations like grocery shopping, restaurants and bars, outdoor activities, and others was roughly 40 minutes. This might imply that, in 2005, cycling was primarily preferred for commuting to homes and workplaces or schools. The shorter durations for other destinations may be attributed to the availability of alternative modes of transportation or the proximity of these destinations. However, the willingness to cycle to reach homes and work or school decreased to 60 and 50 minutes in 2010, respectively. This reduction in cycling time may indicate a shift towards shorter commutes or a preference for alternative transportation methods during this period. Meanwhile, this duration increased to 80 minutes for cycling trips to reach homes and grocery shopping, possibly indicating an emphasis on active transportation for these essential activities. Conversely, the time required to reach work and school decreased to 60 minutes, possibly reflecting changes in commuting patterns.

A significant observation from the study period of 1986 to 2015 is the predominance of the home as both the principal origin and destination for walking trips. During the initial years of the study, specifically in 1986, 1992, and 1998, the subsequent most frequented destinations were identified as the other's home and school or workplace settings, in that order of prevalence [@millward2013active]. However, a notable shift in this pattern emerged between 2005 and 2015. Despite the continued preeminence of the home as the foremost starting point for pedestrian journeys, there was a reordering in the popularity of destinations [@millward2013active;]. Educational institutions and workplaces became the second most common endpoints for these excursions, indicating changes in the societal or infrastructural dynamics influencing pedestrian movement. Concurrently, Grocery stores and malls emerged as the third most favored destination, while visits to the homes of others experienced a decline, positioning them as the fourth most common choice. Additionally, it should be noted that restaurants, bars, clubs, and outdoors have been identified as the next significant destinations for pedestrian trips during this period.

Our investigation aligns with the body of existing research, which collectively illustrates that the duration of walking trips is characteristically shorter than those undertaken by cycling or driving. This observation is supported by Iacono et al. (2010), who delineated that walking trips are generally shorter in comparison to cycling excursions. Further reinforcing this notion, Vale et al. (2017) delineated that walking is predominantly adopted for utilitarian objectives rather than leisure, implying a tendency towards shorter, purpose-driven journeys.

Likewise, Yang et al. (2012) explored the distance decay effect whereby the frequency of trips decreases as distance increases and found it more acute for walking than for cycling. This effect underscores the inherent limitations in walking distance that are not as pronounced in cycling, likely due to the physical exertion and time investment required for longer distances on foot. Expanding on the factors influencing walking trip duration and likelihood, Wu et al. (2019) highlighted the significant role of urban design elements. Their study elucidated that land use patterns, street connectivity, and the presence and quality of pedestrian infrastructure are pivotal in shaping the impedance function for walking trips, effectively determining the ease and appeal of walking as a mode of transportation.

Moreover, Millward et al. (2013) contributed to this nuanced understanding by demonstrating that the duration of walking trips is also contingent upon the nature of the destination. Their findings suggest that utilitarian destinations, such as workplaces or stores, are typically associated with shorter walking trips, reflecting a pragmatic approach to pedestrian movement.
These studies underscore a multifaceted understanding of pedestrian mobility, highlighting the influence of physical, urban, and socio-economic factors on the duration and purpose of walking trips. This body of research not only provides a comprehensive overview of pedestrian travel behavior but also offers critical insights for urban planners and policymakers aiming to promote walking as a sustainable and healthy mode of transportation.[@iacono2008access; @iacono2010; @millward2013active; @wu2019measuring; @yang2012walking].

### 2024-09-22 README.Rmd
Adding the rmd with the methodology and already-processed data

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ActiveCA

<!-- badges: start -->
<!-- badges: end -->

The goal of ActiveCA is to ...

## Installation

You can install the development version of ActiveCA from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("dias-bruno/ActiveCA")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(ActiveCA)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2024-09-22 data-raw/01-pre-processing-source-files.Rmd
Adding the rmd with the methodology and already-processed data

---
title: "Reading source files and preparing data sets"
author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# layout configuration 
library(tufte)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(htmltools.dir.version = FALSE)
rm(list=ls())
```

# Introduction

This Rmarkdown file is part of the
[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. This
package is one of the products developed by the [*Mobilizing
Justice*](https://mobilizingjustice.ca/)[^1].

[^1]: The Mobilizing Justice project is a multidisciplinary and
    multi-sector collaboration with the objective of understand and
    address transportation poverty in Canada and to improve the
    well-being of Canadians at risk of transport poverty. The Social
    Sciences and Humanities Research Council (SSRHC) has provided
    funding for the project, which was created by an unprecedented
    alliance of academics from various Canadian provinces and
    institutions, transportation firms, and nonprofit organizations

The main objective of the *ActiveCA* package is to identify the most
appropriate impedance functions for active transportation modes for
various destinations and time periods in Canada. To do achieve our
objective, we analyzed historical data from the [General Social Survey
(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
from 1986 to 2015 in Canada to calculate the impedance function for
cycling and walking trips.

In this R markdown, we read the GSS surveys and prepare the data set to
the following analyses.

# The General Social Surveys (GSS)

In Canada, Statistics Canada produces the General Social Surveys (GSS)
to collect data on social trends in order to track changes in the living
conditions of Canadians and well-being over time. This survey tracks
changes in time and it is used to understand how Canadians spend and
manage their time and what factors contribute to their happiness and
stress. The survey was created in 1985 as a series of independent,
annual, cross-sectional surveys, each covering one topic in depth.
Caregiving, families, time use, social identity, volunteering, and
victimization are among the current themes of the GSS.

The six survey themes listed above are repeated every five years. In
addition to the central topic, each cycle includes new content that
addresses emerging and policy-relevant issues, collecting detailed
socio-demographic information, such as age, gender, education, religion,
ethnicity, income, etc. Time-use surveys collect data on all human
activities and can inform a wide range of policies. In particular, three
key themes have been identified as essential for policymaking and for
which no other data source is adequate: unpaid work and non-market
production, well-being, and gender equality. Leisure time, work-life
balance, health, commuting, culture, and sports are among the other
topics covered by time-use surveys. In addition, statistics Canada has
been conducting time-use surveys at five to seven year intervals since
1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
and 2015). The time use collects information on participation of
respondents and time spent on a wide range of day-to-day activities
using a 24-hour retrospective diary. Also, information is collected on
the location of these activities (e.g., at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity.

The surveys (obtained from
[here](http://odesi2.scholarsportal.info/webview/)) consist of two
files: a main file and an episode file.

## The Main File

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, enhancing the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes. This file is aggregated and prohibits the direct association
between specific activities with particular locations or companions.

## The Episode File

The Episode File records detailed data for each activity episode
reported by respondents. Each episode entry includes the start and end
times, duration, location, and accompanying social context, informing
when and where activities occurred and with whom. The file distinguishes
itself by focusing on individual episodes rather than respondents, with
the data structured around the numerous activity instances that compose
a respondent's day. Although respondent-specific characteristics are not
included within the Episode File, it is possible to link the Main File
and the Episode File by using an identifiable variable present in both
data sets.

# Reading and processing the GSS files

In this R markdown, our aim is to generate a table in the appropriate
format to perform the subsequent analyses, after reading and processing
the GSS episode surveys.

For the purpose of our analyses, the following definitions are
important:

| **Code** | **Location**                           | **Year** | **Description**                                                |
|-------------|------------------|-------------|------------------------------|
| 300      | At home or on property                 | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 301      | At place of work or school             | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 302      | Away on business                       | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 303      | At someone else's home or property     | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 304      | In the neighbourhood                   | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 305      | Outdoors                               | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 306      | Grocery store, other stores or mall    | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 307      | Library, museum or theatre             | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 308      | Sports centre, field or arena          | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 309      | Restaurant, bar or club                | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 310      | Place of worship                       | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 311      | Medical, dental or other health clinic | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 312      | Elsewhere                              | 2015     | Walking (code = 315) and Cycling (code = 318) to this location |
| 1        | Respondent's home                      | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 2        | Work place                             | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 3        | Someone else's home                    | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 4        | Restaurant/bar                         | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 5        | Place of worship                       | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 6        | Grocery store                          | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 7        | Other store/Mall                       | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 8        | School                                 | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 9        | Outdoors away from home                | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 10       | Library                                | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 11       | Other place                            | 2010     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 1        | Respondent's home                      | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 2        | Work place                             | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 3        | Someone else's home                    | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 4        | Restaurant/bar                         | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 5        | Place of worship                       | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 6        | Grocery store                          | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 7        | Other store/Mall                       | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 8        | School                                 | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 9        | Outdoors away from home                | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 10       | Library                                | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 11       | Other place                            | 2005     | Walking (code = 14) and Cycling (code = 17) to this location   |
| 1        | Respondent's home                      | 1998     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 2        | Work place                             | 1998     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 3        | Someone else's home                    | 1998     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 4        | Other plourhood                        | 1998     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 1        | Respondent's home                      | 1992     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 2        | Work place                             | 1992     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 3        | Someone else's home                    | 1992     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 4        | Other place                            | 1992     | Walking (code = 7) and Cycling (code = 9) to this location     |
| 1        | Respondent's home                      | 1886     | Walking (code = 5) to this location                            |
| 2        | Work place                             | 1886     | Walking (code = 5) to this location                            |
| 3        | At other places                        | 1886     | Walking (code = 5) to this location                            |
| 7        | other                                  | 1886     | Walking (code = 5) to this location                            |

To start processing the data, first load the `R` packages used in this
notebook:

```{r}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(here) # A Simpler Way to Find Your Files
library(readxl) # Read Excel Files
library(tidyr) # Tidy Messy Data
library(usethis) # Automate Package and Project Setup
library(fitdistrplus) # Estimates distribution parameters
library(purrr) # Provides a complete and consistent set of tools for working with functions and vectors
```

For each survey year, the process of preparing the data is very similar
and, in the end, we will obtain a table in which each row represents an
specific episode, and each column a variable. The following variables
will be added to our final table:

-   `PUMFID`: the record identification[^2].
-   `WGHT_EPI`: Episode weight.
-   `ACTCODE`: Activity code of the episode.
-   `STARTIME`: Start time of the episode.
-   `ENDTIME`: End time of the episode.
-   `DURATION` Duration (in minutes) of the episode
-   `LOCATION`: Location/mode of transport of the episode.
-   `ORIGIN`: Place of origin of the episode.
-   `DESTINATION`: Place of destination of the episode.

[^2]: This variable serves as the link between the Main and the Episode
    files. Each case has a unique record identifier and this appears on
    every episode.

The variables mentioned above are obtained directly from the GSS episode
file. To facilitate the reading and better identify the survey of origin
of the episode, we will create the following variables:

-   `orig_label`: A label for the variable `ORIGIN`.
-   `dest_label`: A label for the variable `DESTINATION`.
-   `YEAR`: The year of the survey.
-   `MODE`: The transportation mode of the episode (`walking` or
    `cycling`).

Additionally, our analysis will focus on developing impedance functions considering only the population who lives in larger urban population centre (a Census Metropolitan Area (CMA) or Census Agglomeration (CA))^[For definitions of CMA and CA check the this [website](https://www12.statcan.gc.ca/census-recensement/2021/ref/dict/az/Definition-eng.cfm?ID=geo009).]. We decided to apply this restriction because the travel behavior among residents of the CMA and CA areas, in terms of active travel, tends to be very different from those outside these large urban population centers. The variable that informs if the respondent resides in a CMA or CA region is the variable named `LUC_RST`, present in the Main File survey. According to the Statistics Canada, `LUC_RST` = 1 indicates that the resident is from a CMA or CA region.

The file name pattern is as follows: `transportation mode`[^3] + `_` +
`year of survey` [^4]. This means that, for instance, the file
`walking_2015` refers to transportation mode table `walking` in the year
2015.

[^3]:  `walking` or `cycling`

[^4]:  whether 1986, 1992, 1998, 2005, 2010 or 2015. Note: There is no
    data on cycling for the 1986 GSS survey.

The walking and cycling episodes are identified using the specific
activity code. Also, the preceding and succeeding episodes for each
walking/cycling instance are included to determine the origin and
destination locations of the trip.

## GSS 2015

Read the Episode and Main files of the 2015 GSS survey:

```{r}
# Episode File
gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))

# Main File
gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
```

### Walking 2015

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r}
# walking 2015
# Creating data of origins and destinations of walking trip 2015
inds <- which(gss_e_2015$LOCATION == 315)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested[^5], and already creating
the origin and destination variables:

[^5]: Specifically for the 2015 GSS survey, `TUI_01` refers to the
    activity code. We'll change the name of this column to be match the
    name of the activity code variable in the others surveys.

```{r}
walking_2015 <- gss_e_2015 |>
  dplyr::slice(unlist(rows)) |>
  dplyr::select(PUMFID, WGHT_EPI,TUI_01:LOCATION) |> 
  mutate(origin = lag(LOCATION),
         destination = lead(LOCATION)) |>
  filter(LOCATION == 315)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_2015 <- walking_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_2015 <- walking_2015 |>  
    filter(destination == 300 | 
           destination == 301 | 
           destination == 302 | 
           destination == 303 |
           destination == 304 |
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the column named
`TUI_01` will be renamed to `ACTCODE` to ensure consistency with the
naming conventions used in other years.

```{r}
walking_2015 <- walking_2015 |>
  dplyr::select(PUMFID:ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 2015,
         MODE = "Walking") |>
  dplyr::rename(ACTCODE = TUI_01)
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST')]

walking_2015 <- walking_2015 |> 
  left_join(gss_m_2015, by = 'PUMFID') |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_2015 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The table above shows that the most common destination for a walking trip
that starts at home are `Home` and `Work or school`. It also shows that the second most common walking trip is from `Grocery store, other stores or mall` to `Home`. Notice, though,
that quite a few trips start *and* end at Home. What are those supposed
to be?

Check those home-to-home trips:

```{r echo = FALSE}
walking_2015 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `PUMFID` == 10041 and check the walking episode:

```{r}
gss_e_2015 |>
  filter(PUMFID == 10041) |>
  slice(unique(c(which(LOCATION == 315) - 1, 
                 which(LOCATION == 315),
                 which(LOCATION == 315) + 1))) |>
  dplyr::select(PUMFID:LOCATION)
```

The person was at home doing personal care (code `27`), then went for
walk that lasted 15 minutes, came back home and spent 180 minutes
looking for work (code `9`). So, this was a recreational/leisure
trip![^6]

[^6]: `TUI_01` is an activity code of the episode (What were you doing
    at minute). The list of main activity codes (001-095) can be found
    in an appendix of the User Guide. According to this list, `27` is
    "personal care", `7` is "Transport to or from activity" and `9` is
    "Looking for work".

Save data in R format for later use:

```{r}
usethis::use_data(walking_2015, overwrite = TRUE)
```

### Cycling 2015

Identifying the rows with cycling episodes and selecting the rows after
and before the cycling episodes to obtain the origin and location
variables:

```{r}
# Creating data of origins and destinations of cycling trip 2015
inds <- which(gss_e_2015$LOCATION == 318)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with cycling episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
cycling_2015 <- gss_e_2015[unlist(rows),] |> 
  dplyr::select(PUMFID, WGHT_EPI, TUI_01:LOCATION) |>
  mutate(origin = lag(LOCATION, order_by = PUMFID)) |> 
  mutate(destination = lead(LOCATION, order_by = PUMFID)) |>
  #group_by(PUMFID) |>
  filter(LOCATION == 318)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_2015 <- cycling_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
# change destination and origins to text column
cycling_2015 <- cycling_2015 |>  
  # Choose destinations of interest
  filter(destination == 300 | 
           destination == 301 | 
           destination == 302 | 
           destination == 303 |
           destination == 304 |
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the column named
`TUI_01` will be renamed to `ACTCODE` to ensure consistency with the
naming conventions used in other years.

```{r}
cycling_2015 <- cycling_2015 |>
  dplyr::select(PUMFID:ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 2015,
         MODE = "Cycling") |>
  dplyr::rename(ACTCODE = TUI_01)
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST')]

cycling_2015 <- cycling_2015 |> 
  left_join(gss_m_2015, by = 'PUMFID') |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```
Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
cycling_2015 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is
`Work or school`. Again, we can notice that there are quite a few trips that
start *and* end at Home. What are those supposed to be?

Check those home-to-home trips:

```{r echo = FALSE}
cycling_2015 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `PUMFID` == 16591 and check the cycling episode:

```{r}
gss_e_2015 |>
  filter(PUMFID == 16591) |>
  slice(unique(c(reduce(map(which(LOCATION == 318), ~ c(.x - 1, .x, .x + 1)), c)))) |>
   dplyr::select(PUMFID:LOCATION)
```

The individual was at home engaged in personal care activities
(`TUI_01` = `2`) before cycling for approximately 10 minutes to a grocery
store. After spending 30 minutes at the store, they cycled back home.
Upon returning, the person prepared a meal and cleaned the house for approximately 85 minutes. Subsequently, the individual embarked on
another bicycle trip to the store, which took 10 minutes, and thereafter
returned home.[^7]

[^7]: Note: `TUI_01` is an activity code of the episode (What were you
    doing in minutes?).The list of main activity codes (001-095) can be
    found in an appendix of the User Guide.

Save data in R format for later use:

```{r}
usethis::use_data(cycling_2015, overwrite = TRUE)
```

As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_2015,gss_m_2015)
```
## GSS 2010

Read the Episode and Main files of the 2010 GSS survey:

```{r}
# Episode File
gss_e_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv"))

# Main File
gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv"))
```

### Walking 2010

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r}
# **walking 2010**
# Creating data of origins and destinations of walking trip 2010
inds <- which(gss_e_2010$PLACE == 14)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
walking_2010 <- gss_e_2010[unlist(rows),] |> 
  dplyr::select(RECID,WGHT_EPI, ACTCODE:ENDTIME,DURATION,PLACE) |>
  mutate(origin = lag(PLACE, order_by = RECID)) |> 
  mutate(destination = lead(PLACE, order_by = RECID)) |>
  filter(PLACE == 14)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_2010 <- walking_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_2010 <- walking_2010 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
walking_2010 <- walking_2010 |>
  dplyr::select(RECID, WGHT_EPI,ACTCODE, STARTIME:dest_label) |>
  mutate(YEAR = 2010,
         MODE = "Walking")
walking_2010 <- walking_2010 %>% dplyr::rename(
    PUMFID = RECID,
    LOCATION = PLACE)
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST')]

walking_2010 <- walking_2010 |> 
  left_join(gss_m_2010, by = c('PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_2010 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

For the year 2010, the most common destination for a walking trip that starts at home are
`Home` to `Home`, followed by `Home` to `Work or school`. Again, we can check the home-to-home trips:

```{r echo = FALSE}
walking_2010 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 3 and check the walking episode:

```{r}

gss_e_2010 %>%
  filter(RECID == 3) %>%
  slice(unique(c(reduce(map(which(PLACE == 14), ~ c(.x - 1, .x, .x + 1)), c)))) %>%
  dplyr::select(RECID:PLACE)

```

The individual spent their time at home engaging in meal preparation
(`ACTCODE` =`430`) for a duration of 15 minutes, followed by a walk
lasting approximately 20 minutes to reach a grocery store. After
spending 15 minutes there, the person walked an additional 10 minutes to
another grocery store, where they stayed for 10 minutes before returning
home in 20 minutes. Once home, the individual engaged in socialization
for approximately 10 minutes (`ACTCODE`=`751`) before taking a 10-minute
walk to another location. At this new location, the person undertook
activities related to washing and dressing (`ACTCODE`=`400`) for around 10
minutes, then went for a 10-minute walk, returning home after that.
Subsequently, they dedicated 10 minutes to meal preparation
(`ACTCODE`=`101`), followed by a 10-minute session of food (or meal) cleanup
(`ACTCODE`=`110`). The individual then took a 15-minute walk, returned home,
and spent 5 minutes on household management tasks, which included
organizing and planning activities.[^8]

[^8]: Note: `ACTCODE` is an activity code of the episode (What were you
    doing in minute).The list of main activity codes can be found in an
    appendix of the User Guide.

Save data in R format for later use:

```{r}
usethis::use_data(walking_2010, overwrite = TRUE)
```

### Cycling 2010

Identifying the rows with cycling episodes and selecting the rows after
and before the cycling episodes to obtain the origin and location
variables:

```{r}

# **cycling 2010**
# Creating data of origins and destinations of cycling trip 2010
inds <- which(gss_e_2010$PLACE == 17)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_2010 <- gss_e_2010[unlist(rows),] |> 
  dplyr::select(RECID:PLACE) |>
  mutate(origin = lag(PLACE, order_by = RECID)) |> 
  mutate(destination = lead(PLACE, order_by = RECID)) |>
  group_by(RECID) |>
  filter(PLACE == 17)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_2010 <- cycling_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

Similarly to the previous step, the description column will also be
created for destinations.

```{r}
cycling_2010 <- cycling_2010 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
cycling_2010 <- cycling_2010 |>
  dplyr::select(RECID, WGHT_EPI,ACTCODE:ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 2010,
         MODE = "Cycling")
cycling_2010 <- cycling_2010 %>% dplyr::rename(
    PUMFID = RECID,
    LOCATION = PLACE)
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST')]

cycling_2010 <- cycling_2010 |> 
  left_join(gss_m_2010, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```
Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
cycling_2010 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is
`Work or school`. Notice, though that there are some trips that start
*and* end at Home. What are those supposed to be?

Check those home-to-home trips:

```{r echo = FALSE}
cycling_2010 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 6301 and check the walking episode:

```{r}
gss_e_2010 |>
  filter(RECID == 6301) |>
    slice(unique(c(reduce(map(which(PLACE == 17), ~ c(.x - 1, .x, .x + 1)), c)))) |>
   dplyr::select(RECID,PLACE,ACTCODE,DURATION)
```

The individual engaged in watching television at their residence for
approximately 160 minutes before cycling for about 15 minutes. Upon
returning home, they socialized for roughly 165 minutes. Following this
period of socialization at a private residence, the person embarked on
another cycling session for 15 minutes, after which they returned home
once more. Subsequently, the individual carried out various household
chores for about 10 minutes.

Save data in R format for later use:

```{r}
usethis::use_data(cycling_2010, overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_2010,gss_m_2010)
```
## GSS 2005

Read the Episode and Main files of the 2005 GSS survey:

```{r}
# Episode File
gss_e_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-e_F1.csv"))

# Main File
gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv"))
```

### Walking 2005

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r creating dataset for walking 2005 }

# **walking 2005**
# Creating data of origins and destinations of walking trip 2005
inds = which(gss_e_2005$PLACE == 14)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
walking_2005 <- gss_e_2005 [unlist(rows),] |>
  dplyr::select(RECID:PLACE) |> 
  mutate(origin = lag(PLACE, order_by = RECID)) |>
  mutate(destination = lead(PLACE, order_by = RECID)) |> 
  group_by(RECID) |> 
  filter(PLACE == 14) 
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_2005 <- walking_2005 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",               
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_2005 <- walking_2005 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
walking_2005 <- walking_2005 |>
  dplyr::select(RECID, WGHT_EPI, ACTCODE:ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 2005,
         MODE = "Walking")
walking_2005 <- walking_2005 |> dplyr::rename(
    PUMFID = RECID,
    LOCATION = PLACE) 
```

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST')]

walking_2005 <- walking_2005 |> 
  left_join(gss_m_2005, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_2005 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a walking trip that starts at home are
`work or school`, `Other's home` and `home`.

Check those home-to-home trips:

```{r echo = FALSE}
walking_2005 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 89 and check the walking episode:

```{r}
gss_e_2005 |>
  filter(RECID == 89) |>
  slice(unique(c(reduce(map(which(PLACE == 14), ~ c(.x - 1, .x, .x + 1)), c)))) |>
  dplyr::select(RECID,PLACE,ACTCODE,DURATION)
```

The individual engaged in personal grooming and dressing activities
(`ACTCODE`= 400) before embarking on a brief walk lasting approximately 5
minutes, after which they returned home. Subsequently, the person
participated in religious activities at their residence, which included
services, prayer, and Bible readings (`ACTCODE` = 640). Following these
activities, the individual undertook another walk of around 5 minutes,
returning home once more. At home, they engaged in various activities,
including socializing with friends and relatives and sleeping. Later,
the person walked for approximately 5 minutes to reach a different
location and then spent an additional 5 minutes walking to return
home.[^9]

[^9]: Note: `ACTCODE` is an activity code of the episode (What were you
    doing in minute).The list of main activity codes can be found in an
    appendix of the User Guide.

Save data in R format for later use:

```{r}
usethis::use_data(walking_2005, overwrite = TRUE)
```

### Cycling 2005

Identifying the rows with cycling episodes and selecting the rows after
and before the cycling episodes to obtain the origin and location
variables:

```{r creating dataset for cycling 2005 }

# **cycling 2005**
# Creating data of origins and destinations of cycling trip 2005
inds <- which(gss_e_2005$PLACE == 17)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with cycling episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
cycling_2005 <- gss_e_2005[unlist(rows),] |> 
  dplyr::select(RECID:PLACE) |>
  mutate(origin = lag(PLACE, order_by = RECID)) |> 
  mutate(destination = lead(PLACE, order_by = RECID)) |>
  group_by(RECID) |>
  filter(PLACE == 17)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_2005 <- cycling_2005 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

Similarly to the previous step, the description column will also be
created for destinations.

```{r}
cycling_2005 <- cycling_2005 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
cycling_2005 <- cycling_2005 |>
  dplyr::select(RECID, WGHT_EPI, ACTCODE:ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 2005,
         MODE = "Cycling")
cycling_2005 <- cycling_2005 |> dplyr::rename(
    PUMFID = RECID,
    LOCATION = PLACE) 
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST')]

cycling_2005 <- cycling_2005 |> 
  left_join(gss_m_2005, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
cycling_2005 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is
`Work or school`. Check those home-to-home trips:

```{r echo = FALSE}
cycling_2005 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 1903 and check the cycling episode:

```{r}
gss_e_2005 |>
  filter(RECID == 1903) |>
  slice(unique(c(reduce(map(which(PLACE == 17), ~ c(.x - 1, .x, .x + 1)), c)))) |>  
  dplyr::select(RECID,PLACE,ACTCODE,DURATION)
```

The individual remained at home, where they slept for approximately 420
minutes. Subsequently, they ventured out for a 15-minute cycling session
before returning home. Upon their return, they engaged in an activity
labeled as `ACTCODE` = 11 for 465 minutes, which denotes working from home
as their primary workplace. Following this work period, the person went
out for another 15-minute cycling session, then returned home once more
and spent 90 minutes watching television.

Save data in R format for later use:

```{r}
usethis::use_data(cycling_2005, overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_2005,gss_m_2005)
```
## GSS 1998

Read the Episode and Main files of the 1998 GSS survey:

```{r}
# Episode File
gss_e_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12e_F1.csv"))

# Main File
gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv"))
```

### Walking 1998

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r creating dataset for walking 1998 }

# **walking 1998**
# Creating data of origins and destinations of walking trip 1998
inds = which(gss_e_1998$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
walking_1998 <- gss_e_1998 [unlist(rows),] |>
  dplyr::select(RECID:PLACE) |> 
  mutate(origin = lag(PLACE, order_by = RECID)) |>
  mutate(destination = lead(PLACE, order_by = RECID)) |> 
  group_by(RECID) |> 
  filter(PLACE == 7) 
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_1998 <- walking_1998 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_1998 <- walking_1998 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
walking_1998 <- walking_1998 |>
  dplyr::select(RECID, WGHTEPI,ACTCODE: ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 1998,
         MODE = "Walking")
walking_1998 <- walking_1998|> dplyr::rename(
    PUMFID = RECID,
    WGHT_EPI = WGHTEPI,
    LOCATION = PLACE) 
```

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_1998 <- gss_m_1998[,c('RECID','CMAPRV')]

walking_1998 <- walking_1998 |> 
  left_join(gss_m_1998, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(CMAPRV %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_1998 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

In this case, the most common destination for a walking trip that starts at home are
`other's Home`.

Check those home-to-home trips:

```{r echo = FALSE}
walking_1998 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 2302 and check the walking episode:

```{r}
gss_e_1998 |>
  filter(RECID == 2302) |>
  slice(unique(c(reduce(map(which(PLACE == 7), ~ c(.x - 1, .x, .x + 1)), c)))) |>
  dplyr::select(RECID,PLACE,ACTCODE,DURATION)

```

The person was at home doing food cleanup (`ACTCODE` = 110), then
went for walking that lasted 30 minutes, came back home and spent 15 minutes relaxing/resting. Here have here another cas of a recreational/leisure trip!

Save data in R format for later use:

```{r}
usethis::use_data(walking_1998, overwrite = TRUE)
```

### Cycling 1998

Identifying the rows with cycling episodes and selecting the rows after
and before the cycling episodes to obtain the origin and location
variables:

```{r creating dataset for cycling 1998 }
# **cycling 1998**
# Creating data of origins and destinations of cycling trip 1998
inds <- which(gss_e_1998$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with cycling episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
cycling_1998 <- gss_e_1998[unlist(rows),] |> 
  dplyr::select(RECID:PLACE) |>
  mutate(origin = lag(PLACE, order_by = RECID)) |> 
  mutate(destination = lead(PLACE, order_by = RECID)) |>
  group_by(RECID) |>
  filter(PLACE == 9)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_1998 <- cycling_1998 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

Similarly to the previous step, the description column will also be
created for destinations.

```{r}
cycling_1998 <- cycling_1998 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
cycling_1998 <- cycling_1998 |>
  dplyr::select(RECID, WGHTEPI,ACTCODE: ENDTIME, DURATION:dest_label) |>
  mutate(YEAR = 1998,
         MODE = "Cycling")
cycling_1998 <- cycling_1998|> dplyr::rename(
    PUMFID = RECID,
    WGHT_EPI = WGHTEPI,
    LOCATION = PLACE) 
```

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_1998 <- gss_m_1998[,c('RECID','CMAPRV')]

cycling_1998 <- cycling_1998 |> 
  left_join(gss_m_1998, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
  filter(CMAPRV %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
cycling_1998 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is
`Work or school`.

Check those home-to-home trips:

```{r echo = FALSE}
cycling_1998 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See `RECID` == 4695 and check the cycling episode:

```{r}
gss_e_1998 |>
  filter(RECID == 4695) |>
  slice(unique(c(reduce(map(which(PLACE == 9), ~ c(.x - 1, .x, .x + 1)), c)))) |>  
  dplyr::select(RECID,PLACE,ACTCODE,DURATION)
```

The person spent 165 minutes talking to someone (in person or on the phone, `ARTCODE` = 950), then embarked on a bike ride for 12 minutes. After that, the person spent 108 minutes shopping at home (`ARTCODE` = 380), cycled for another 12 minutes and came back to talk to someone for another 138 minutes (`ARTCODE` = 950).

Save data in R format for later use:

```{r}
usethis::use_data(cycling_1998, overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_1998,gss_m_1998)
```
## GSS 1992

Read the Episode and Main files of the 1992 GSS survey:

```{r}
# Episode File
gss_e_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-ep_F1.csv"))

# Main File
gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv"))
```

### Walking 1992

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r creating dataset for walking 1992 }

# **walking 1992**
# Creating data of origins and destinations of walking trip 1992
inds = which(gss_e_1992$PLACE == 7)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
walking_1992 <- gss_e_1992 [unlist(rows),] |>
  dplyr::select(SEQNUM:PLACE,TIMEWGT) |> 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) |>
  mutate(destination = lead(PLACE, order_by = SEQNUM)) |> 
  group_by(SEQNUM) |> 
  filter(PLACE == 7) 
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_1992 <- walking_1992 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_1992 <- walking_1992 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
walking_1992 <- walking_1992 |>
  dplyr::select(SEQNUM,TIMEWGT,ACTCODE:dest_label) |>
  mutate(YEAR = 1992,
         MODE = "Walking")
walking_1992 <- walking_1992 |> dplyr::rename(
    PUMFID = SEQNUM,
    WGHT_EPI = TIMEWGT,
    LOCATION = PLACE)
```

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_1992 <- gss_m_1992[,c('SEQNUM','DVCMA')]

walking_1992 <- walking_1992 |> 
  left_join(gss_m_1992, by = c( 'PUMFID' = 'SEQNUM')) |>  # Merging the two tables
  filter(DVCMA %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_1992 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a walking trip that starts at home is
`Home`. Check these home-to-home trips:

```{r echo = FALSE}
walking_1992 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See SEQNUM == 1235 and check the walking episode:

```{r}
gss_e_1992 |>
  filter(SEQNUM == 1235) |>
  slice(unique(c(reduce(map(which(PLACE == 7), ~ c(.x - 1, .x, .x + 1)), c)))) |>
  dplyr::select(SEQNUM,PLACE,ACTCODE,DURATION)

```

The person was at home reading newspapers (`ACTCODE` = 940), then went for walking that lasted 30 minutes, came back home and spent 90 minutes resting (`ACTCODE` = 470).[^10]

[^10]: Note: ACTCODE is an activity code of the episode (What were you
    doing in minute).The list of main activity codes can be found in an
    appendix of the User Guide.

Save data in R format for later use:

```{r}
usethis::use_data(walking_1992, overwrite = TRUE)
```

### Cycling_1992

Identifying the rows with cycling episodes and selecting the rows after
and before the cycling episodes to obtain the origin and location
variables:

```{r creating dataset for cycling 1992 }
# **cycling 1992**
# Creating data of origins and destinations of cycling trip 1992
inds <- which(gss_e_1992$PLACE == 9)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with cycling episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
cycling_1992 <- gss_e_1992[unlist(rows),] |> 
  dplyr::select(SEQNUM:PLACE, TIMEWGT) |>
  mutate(origin = lag(PLACE, order_by = SEQNUM)) |> 
  mutate(destination = lead(PLACE, order_by = SEQNUM)) |>
  group_by(SEQNUM) |>
  filter(PLACE == 9)
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
cycling_1992 <- cycling_1992 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
cycling_1992 <- cycling_1992 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
cycling_1992 <- cycling_1992 |>
  dplyr::select(SEQNUM,TIMEWGT,ACTCODE:dest_label) |>
  mutate(YEAR = 1992,
         MODE = "Cycling")
cycling_1992 <- cycling_1992 |> dplyr::rename(
    PUMFID = SEQNUM,
    WGHT_EPI = TIMEWGT,
    LOCATION = PLACE)
```
Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_1992 <- gss_m_1992[,c('SEQNUM','DVCMA')]

cycling_1992 <- cycling_1992 |> 
  left_join(gss_m_1992, by = c( 'PUMFID' = 'SEQNUM')) |>  # Merging the two tables
  filter(DVCMA %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
cycling_1992 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

The most common destination for a cycling trip that starts at home is
`Work or school`.

Check those home-to-home trips:

```{r echo = FALSE}
cycling_1992 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See SEQNUM == 8493 and check the cycling episode:

```{r}
gss_e_1992 |>
  filter(SEQNUM == 8493) |>
  slice(unique(c(reduce(map(which(PLACE == 9), ~ c(.x - 1, .x, .x + 1)), c)))) |> 
  dplyr::select(SEQNUM,PLACE,ACTCODE,DURATION)
```

The person was at home Talking or having conversation on the phone (`ACTCODE` = 950), then went for cycling that lasted 60 minutes, came back home and spent 15 minutes washing and dressing (`ACTCODE` = 400).

Save data in R format for later use:

```{r}
usethis::use_data(cycling_1992, overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_1992,gss_m_1992)
```

## GSS 1986

Read the Episode and Main files of the 1986 GSS survey:

```{r}
# Episode File
gss_e_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-ep_F1.csv"))

# Main File
gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv"))
```

### Walking 1986

Identifying the rows with walking episodes and selecting the rows after
and before the walking episodes to obtain the origin and location
variables:

```{r creating dataset for walking 1986}
# **walking 1986**
# Creating data of origins and destinations of walking trip 1986
inds = which(gss_e_1986$PLACE == 5)
rows <- lapply(inds, function(x) (x-1):(x+1))
```

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested, and already creating the
origin and destination variables:

```{r}
walking_1986 <- gss_e_1986 [unlist(rows),] |>
  dplyr::select(SEQNUM:PLACE,FWGT_MS) |> 
  mutate(origin = lag(PLACE, order_by = SEQNUM)) |>
  mutate(destination = lead(PLACE, order_by = SEQNUM)) |> 
  group_by(SEQNUM) |> 
  filter(PLACE == 5) 
```

Now, we'll select the origins of interest and create the label for the
`origin` variable:

```{r}
walking_1986 <- walking_1986 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

Similarly to the step before, we'll select the destinations of interest
and create the label for the `destination` variable:

```{r}
walking_1986 <- walking_1986 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home"),
         dest_label = factor(dest_label))
```

Creating the `year` and `mode` columns. Additionally, the columns
`RECID`, `WGHTEPI` and `PLACE` were changed to `PUMFID`, `WGHT_EPI` and
`LOCATION` to align with the names of the variables in 2015.

```{r}
walking_1986 <- walking_1986 |>
  dplyr::select(SEQNUM,ACT_CODE:dest_label) |>
  mutate(YEAR = 1986,
         MODE = "Walking")

walking_1986 <- walking_1986 |> dplyr::rename(
    PUMFID = SEQNUM,
    STARTIME= STRTTIME,
    WGHT_EPI = FWGT_MS,
    ACTCODE = ACT_CODE,
    LOCATION = PLACE)
```

Now, let's evaluate the count of episodes between each combination of
origin and destination:

```{r}
walking_1986 |> 
  group_by(orig_label, 
           dest_label) |>
  count() |>
  pivot_wider(names_from = "dest_label", values_from = "n")
```

One of the most common destination for a walking trip that starts at
home is `HOME`. Notice, though that the most trips start *and* end at
Home. What are those supposed to be?

Check those home-to-home trips:

```{r echo = FALSE}
walking_1986 |>
  filter(orig_label == "Home" & dest_label == "Home")
```

See SEQNUM == 26 and check the walking episode:

```{r}
gss_e_1986 |>
  filter(SEQNUM == 68) |>
  slice(unique(c(reduce(map(which(PLACE == 5), ~ c(.x - 1, .x, .x + 1)), c)))) |>
  dplyr::select(SEQNUM,PLACE,ACT_CODE,DURATION)
```

The person was at home listening radio for 30 minutes (`ARTCODE` = 90), then went for walking that lasted 15 minutes, came back home and spent 90 minutes
resting/relaxing (`ARTCODE` = 47).

Save data in R format for later use:

```{r}
usethis::use_data(walking_1986, overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

```{r}
rm(gss_e_1986,gss_m_1986)
```

# Creating integrated dataset of walking and cycling trips

To finish our pre-processing methodology, we will create a dataset named
`trip` that will combine all the previous tables into a unique dataset:

```{r creating an integrated data}
# creating an integrated data frame from 1986 to 2015
gss_episodes <- rbind(walking_2015[,1:13], cycling_2015[,1:13], walking_2010[,1:13], cycling_2010[,1:13], walking_2005[,1:13], cycling_2005[,1:13], cycling_1998[,1:13], walking_1998[,1:13], walking_1992[,1:13], cycling_1992[,1:13], walking_1986[,1:13])

# Change the name of some values in the dest column in the dataset 
# Ensure that 'dest_label' is the correct column name
gss_episodes <- gss_episodes |>
  mutate(dest_label = case_when(
    dest_label == "Restaurant/bar" ~ "Restaurant, bar or club",
    dest_label == "Outdoors away from home" ~ "Outdoors",
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Outdoors",
    dest_label == "Library, museum or theatre" ~ "Library",
    TRUE ~ dest_label  # This line ensures other values remain unchanged
  ))
```

Save this data frame in R format:

```{r}
usethis::use_data(gss_episodes, overwrite = TRUE)
```

### 2024-09-22 data-raw/02-calculating-impedances.Rmd
Adding the rmd with the methodology and already-processed data

---
title: "Calculating impedances for walking in cycling"
author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# layout configuration 
library(tufte)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(htmltools.dir.version = FALSE)
rm(list=ls())
```

# Introduction

This Rmarkdown file is part of the
[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. This
package is one of the products developed by the [*Mobilizing
Justice*](https://mobilizingjustice.ca/)[^1].

[^1]: The Mobilizing Justice project is a multidisciplinary and
    multi-sector collaboration with the objective of understand and
    address transportation poverty in Canada and to improve the
    well-being of Canadians at risk of transport poverty. The Social
    Sciences and Humanities Research Council (SSRHC) has provided
    funding for the project, which was created by an unprecedented
    alliance of academics from various Canadian provinces and
    institutions, transportation firms, and nonprofit organizations

The main objective of the *ActiveCA* package is to identify the most
appropriate impedance functions for active transportation modes for
various destinations and time periods in Canada. To do achieve our
objective, we analyzed historical data from the [General Social Survey
(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
from 1986 to 2015 in Canada to calculate the impedance function for
cycling and walking trips.

In this R markdown, we'll calculate the impedance functions for walking and cycling for each GSS year (from 1986 to 2015).

# Coding

To start processing the data, first load the `R` packages used in this
notebook:

```{r}
library(dplyr)# A Grammar of Data Manipulation 
library(fitdistrplus) # Help to Fit of a Parametric Distribution to Non-Censored or Censored Data
library(scales) # Scale data column-wise in a computationally efficient way
library(here) # enable easy file referencing in project-oriented workflows
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
```

To exemplify the process of selecting the best impedance function, we will use the example of the 1986 GSS dataset, already pre-processed in the `00-Reading-Source-Files.Rmd`: 
```{r}
load(file = paste0(here(), "/data/walking_1986.Rda"))
```

First, we'll analyze the travel time variable (`DURATION`) of the 1986 dataset, using the `summary()` function.

```{r}
summary(walking_1986$DURATION)
```
The minimum duration value for walking trips is `1` minute,  while the maximum is `660` minutes. Additionally, the average and median duration for walking trips are `20.85` minutes and `10` minutes, respectively.

Now, we'll see this variable displayed in a histogram:
```{r}
ggplot(walking_1986, aes(x = DURATION )) + 
  geom_histogram(binwidth = 30) + 
  geom_vline(aes(xintercept=mean(DURATION)), color="blue", linetype="dashed", linewidth=0.5) + 
  geom_vline(aes(xintercept=median(DURATION)), color="green", linetype="dashed", linewidth=0.5)
```

A histogram is a graph that illustrates the frequency distribution of a dataset corresponding to a single variable. Histograms displays the distribution and variability of the dataset, typically organizing data into 'bins' or 'range groups,' tallying the quantity of data points within each bin.

Analyzing the histogram above, we can identify that the just a few values of duration are higher than 1 hundred minutes.  To better visualize the variable distribution, let's use an other type of graph, the boxplot graph:

```{r}
ggplot(walking_1986, aes(y = DURATION, x = "")) + 
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, max(walking_1986$DURATION, na.rm = TRUE), by = 100)) +
  xlab("")
```
We can see that just a few outliers present values higher than 100 minutes. In fact, the chunk below shows that from the 4347 instances of the dataset, 117 have values of duration higher than 100 minutes. Because of this, we'll analyze only individuals who engaged in walking for less than 100 minutes, focusing on their origins and destinations.

```{r}
length(walking_1986$DURATION)
sum(walking_1986$DURATION > 100)
```
Visualizing the new summary:
```{r}
walking_1986 <- walking_1986 %>% filter(DURATION <= 100)
summary(walking_1986$DURATION)
```

## Impedance functions

The impedance function $f(c_{ij})$ reveals important information about
the travel behavior of the population, as it represents the relationship
between the "population" at an origin and where they usually go, want to
go, or can go to reach the "opportunities" at the destinations.
Because of this, defining the impedance function is extremely important.

The decay rate of the impedance function needs to be calibrated if one
wants the accessibility estimates to be representative of the people's
travel behavior. In our case, we'll use the Census data in the
calibration process.

Soukhov and Páez [-@MJ-A2-0002] reviewed commonly used impedance
functions $f(\bullet)$ in accessibility research, explaining their
impacts on the summation of opportunities at specific travel costs
$c_{ij}$. In this study, we will explore the following functions to
determine the impedance function that best fits our data:
[exponential](https://en.wikipedia.org/wiki/Exponential_distribution),
[gamma](https://en.wikipedia.org/wiki/Gamma_distribution),
[log-normal](https://en.wikipedia.org/wiki/Log-normal_distribution),
[normal](https://en.wikipedia.org/wiki/Normal_distribution), and
[uniform](https://en.wikipedia.org/wiki/Continuous_uniform_distribution).

# Trip length distribution: exemplifying the case of the 1986 GSS

All of the impedance functions mentioned require the analyst to define
parameters. One useful technique is to create a trip length distribution
(TLD) using empirically observed origin-destination travel survey data.
A TLD reflects observed travel patterns: specifically, the probability
of an observed trip of a given travel cost occurring for the population
in a region of interest. Based on the TLD, we can select the most
appropriate theoretical PDF forms (e.g., uniform, exponential, gamma),
adjust the associated parameters, and use the calibrated theoretical PDF
to incorporate the assumptions about the population's travel behavior
into the accessibility calculation.

Soukhov and Páez [-@MJ-A2-0002] demonstrated the process of calibrating
and selecting the best distribution to represent an impedance function
based on travel flows from workers who live and work (full-time) within
the City of Hamilton. The authors used the data from the R data package
[{TTS2016R}](https://soukhova.github.io/TTS2016R/) and the R package
[{fitdistrplus}](https://cloud.r-project.org/web/packages/fitdistrplus/index.html)
to generate parameters that best fit the TLD the parameters of the
uniform, exponential, and gamma functions as closely to the TLD.

The next chunks of code of this Rmarkdown explains the process of calibrating PDF
functions for the Walking dataset of the 1986 GSS.

## Empirical travel time

Creating a data frame with the all empirical travel times:

```{r empiric-travel-times}
# Bind the empirical trip length distribution:
tld_empirical <- walking_1986[, c("DURATION", "MODE", "destination", "WGHT_EPI","dest_label")] %>% 
  mutate(MODE = factor(MODE, levels = c("Walking","Cycling")), 
      distribution = "empirical")

tld_empirical$DURATION <- as.numeric(tld_empirical$DURATION)
```

## Theoretical travel time

Now that we already have the travel times for each trip, we will try to create a impedance function that better
describes the travel pattern of the respondents in each destination. To do so,
we'll first create a function that based on the lowest Akaike
Information Criterion (AIC)[^12], selects the better function between
the distributions:
[exponential](https://en.wikipedia.org/wiki/Exponential_distribution),
[gamma](https://en.wikipedia.org/wiki/Gamma_distribution),
[log-normal](https://en.wikipedia.org/wiki/Log-normal_distribution),
[normal](https://en.wikipedia.org/wiki/Normal_distribution), and
[uniform](https://en.wikipedia.org/wiki/Continuous_uniform_distribution).

[^12]: The Akaike Information Criterion (AIC) estimates prediction error
    and model quality for a given data set.

We'll use the
[{fitdistrplus}](https://cloud.r-project.org/web/packages/fitdistrplus/index.html)
to determine the parameters that best fit the TLD. The Moment Matching Estimation (MME) fitting method and the Nelder-Mead
direct optimization algorithm are used [@mullerdutang2015].

```{r lowest-aic-function}
# Select the distribution function based on the lowest AIC value
lowest_aic <- function(values){

  min_aic <- min(values)

  if(!is.na(lnorm_$aic) & min_aic == lnorm_$aic){
    choosen_f <- lnorm_
  }

  else if  (!is.na(gamma_$aic) & min_aic == gamma_$aic){
    choosen_f <- gamma_
  }

  else if  (!is.na(unif_$aic) & min_aic == unif_$aic){
    choosen_f <- unif_
  }

  else if  (!is.na(norm_$aic) & min_aic == norm_$aic){
    choosen_f <- norm_
  }
  else if  (!is.na(exp_$aic) & min_aic == exp_$aic){
    choosen_f <- exp_
  }

  return(choosen_f)
}

# Test the distributions
test_distributions_weighted <- function(x, weights){
  gamma_ <<- fitdistrplus::fitdist(data=x, "gamma", method="mme", weights = weights)
  lnorm_ <<- fitdistrplus::fitdist(data=x, "lnorm", method="mme", weights = weights)
  norm_ <<- fitdistrplus::fitdist(data=x, "norm", method="mme", weights = weights)
  exp_ <<- fitdistrplus::fitdist(data=x, "exp", method="mme", weights = weights)
  unif_ <<- fitdistrplus::fitdist(data=x, "unif", method="mme", weights = weights)

  values <- c(lnorm_$aic, gamma_$aic, unif_$aic, norm_$aic, exp_$aic)
  values <- values[!is.na(values)]

  chosen_function <- lowest_aic(values)

  return(chosen_function)
}
```

Obtaining the theoretical impedance function for the each destination mode:
```{r}
for(destino in unique(tld_empirical$destination)){
  assign(paste0("f_1986_Walking_dest", destino), 
         test_distributions_weighted(
           x = tld_empirical[tld_empirical$destination == destino, ]$DURATION,
           weights = tld_empirical[tld_empirical$destination == destino, ]$WGHT_EPI))
}
```

Based on the minimum AIC, the best function for the destination `1` (`Home`), considering walking as the mode of transportation, is the `Log-normal distribution`:

```{r}
f_1986_Walking_dest1
```

Actually, the `Log-normal distribution` was the best fuction for the others two destination from the survey: `2` (`Work or school`) and `3` (`Other's home`).

```{r}
f_1986_Walking_dest2
```

```{r}
f_1986_Walking_dest3
```

The chunk below aims to create functions to graphically display the
theoretical values obtained in the chunks:

```{r theoretical-impedance-functions}

# For Displaying theoretical data

visualize_impedance <- function(destination_function, travel_cost, name_destination){

  x <- data.frame(t = seq(1, max(travel_cost), 1))

  # lnorm impedance 
  if(destination_function$distname == "lnorm"){
      x <- x %>%
      mutate(f = dlnorm(t,
                    meanlog = destination_function$estimate[1],
                    sdlog = destination_function$estimate[2]),
         destination = name_destination)

  # Unif impedance 
  } else if(destination_function$distname == "unif") {
    x <- x %>%
    mutate(f = dunif(t,
                     min=0,
                     max=destination_function$estimate[2]), #already scaled from 1 to 0
         destination = name_destination)

  # Exponential impedance 
  } else if(destination_function$distname == "exp") {
    x <- x %>%
    mutate(f = dexp(t,
                    rate = destination_function$estimate), #|> scales::rescale(),
         destination = name_destination)

  # Gamma impedance 
  } else if (destination_function$distname == "gamma"){  
    x <- x %>%
        mutate(f = dgamma(t,
                    shape = destination_function$estimate[1],
                    rate = destination_function$estimate[2]),
         destination = name_destination)

  # Norm impedance 
  } else if (destination_function$distname == "norm"){
        x <- x %>%
        mutate(f = dnorm(t,
                    mean = destination_function$estimate[1],
                    sd = destination_function$estimate[2]),
         destination = name_destination)
        }

  return(x)
}
```

Building the theoretical values for each destination based on the best
impedance function for each destination:

```{r theoretical-values}
# Theoretical values for bike destination 
tld_theoretical <- data.frame()

for(destino in unique(tld_empirical$destination)){

  current_function <- paste0("f_1986_Walking_dest", destino)

  assign(paste0("theo_1986_Walking_dest", destino), 
         visualize_impedance(destination_function = get(current_function),
                             travel_cost = tld_empirical[tld_empirical$destination == destino, ]$DURATION,
                             name_destination = destino))

  tld_theoretical <- rbind(tld_theoretical, get(paste0("theo_1986_Walking_dest", destino)))
  }
```

Plotting the theoretical (black lines) and the empirical travel time
distributions (bars):

```{r travel-time-plots}
par(mfrow = c(2, round(length(unique(tld_theoretical$destination))/2)))  # Arrange plots in 2 columns

# Loop through each destination
for (destination in unique(tld_empirical$destination)) {
  # Subset empirical and theoretical data for the current destination
  empirical_subset <- tld_empirical[tld_empirical$destination == destination, ]
  theoretical_subset <- tld_theoretical[tld_theoretical$destination == destination, ]

  # Create histogram for empirical data
  hist(empirical_subset$DURATION, freq = FALSE, col = "grey70", breaks = 20, main = paste(empirical_subset$dest_label[1]),
       xlab = "Travel time from home to work (min)", ylab = "Density")

  # Add theoretical curve
  lines(theoretical_subset$t, theoretical_subset$f)
}

# Reset plotting parameters
par(mfrow = c(1, 1))
```

# Calculating for all GSS cicles

```{r}
load(file = paste0(here(), "/data/gss_episodes.Rda"))
```

```{r}
gss_episodes <- gss_episodes %>%
  filter(DURATION <= 100 & DURATION > 0)

gss_episodes <- gss_episodes %>%
  group_by(MODE, YEAR, destination) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  filter(count > 1)

gss_episodes$WGHT_EPI[is.na(gss_episodes$WGHT_EPI)] <- 1
```

```{r}

for(year in unique(gss_episodes$YEAR)){

  for(destino in unique(gss_episodes$destination)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode, ]

        if(nrow(gss_subset) > 0){
          print(paste("Calculating the best function for ", 
                    mode, ", destination ",  
                    gss_subset$dest_label[1],
                    ", for the ", year, " GSS cicle"))

          assign(paste0("f_",year, "_", mode, "_dest", destino),
               test_distributions_weighted(
               x = gss_subset$DURATION,
               weights = round(gss_subset$WGHT_EPI)))
        }
    }
  }
}
```

```{r}
# To generate the theoretical data
generate_impedance <- function(destination_function, travel_cost){

  f <- 0

  # lnorm impedance 
  if(destination_function$distname == "lnorm"){
  f <- dlnorm(travel_cost,
                    meanlog = destination_function$estimate[1],
                    sdlog = destination_function$estimate[2])

   estimate_1 <- destination_function$estimate[1]
   estimate_2 <- destination_function$estimate[2]
  }

  # Unif impedance 
  else if(destination_function$distname == "unif") {
   f <- dunif(travel_cost,
                     min=0,
                     max=destination_function$estimate[2])

   estimate_1 <- 0
   estimate_2 <- destination_function$estimate[2]

  # Exponential impedance 
  } else if(destination_function$distname == "exp") {
    f <- rescale(dexp(travel_cost,
                    rate = destination_function$estimate))

   estimate_1 <- destination_function$estimate
   estimate_2 <- 0

      # Gamma impedance 
  } else if (destination_function$distname == "gamma"){  
    f <- dgamma(travel_cost,
                    shape = destination_function$estimate[1],
                    rate = destination_function$estimate[2])

   estimate_1 <- destination_function$estimate[1]
   estimate_2 <- destination_function$estimate[2]

  # Norm impedance 
  } else if (destination_function$distname == "norm"){
        f <- dnorm(travel_cost,
                    mean = destination_function$estimate[1],
                    sd = destination_function$estimate[2])

   estimate_1 <- destination_function$estimate[1]
   estimate_2 <- destination_function$estimate[2]

  }

  df_f <- data.frame("f_name" = rep(destination_function$distname, length(f)),
           "f" = f, 
           "est_1" = rep(estimate_1, length(f)),
           "est_2" = rep(estimate_2, length(f)))

  return(df_f)
}
```

```{r}
gss_episodes$f_name <- ''
gss_episodes$f <- 0
gss_episodes$est_1 <- 0
gss_episodes$est_2 <- 0
gss_episodes$loglik <- 0
gss_episodes$aic <- 0
gss_episodes$bic <- 0

for(year in unique(gss_episodes$YEAR)){

  for(destino in unique(gss_episodes$destination)){

    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){

        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            ]

        if(nrow(gss_subset) > 0){

                    print(paste("Obtaining the values for ", 
                    mode, ", destination ",  
                    gss_subset$dest_label[1],
                    ", for the ", year, " GSS cicle"))

        current_function <- paste0("f_",year, "_", mode, "_dest", destino)

        df_f <- generate_impedance(destination_function = get(current_function),
                                   travel_cost = gss_subset$DURATION)

        gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("f_name","f","est_1","est_2")] <- df_f[,c("f_name","f","est_1","est_2")]

        gss_episodes[gss_episodes$destination == destino &
                            gss_episodes$YEAR == year &
                            gss_episodes$MODE == mode ,
                            c("loglik", "aic", "bic")] <- list(get(current_function)$loglik, get(current_function)$aic, get(current_function)$bic)
        }
      }
  }
}

```

```{r}
gss_impedances <- gss_episodes[,c("PUMFID","DURATION","LOCATION",
                  "origin","destination","orig_label",
                  "dest_label","YEAR","MODE","f_name",
                  "f","est_1","est_2","loglik","aic","bic")]
```

Save this data frame in R format:

```{r}
usethis::use_data(gss_impedances, overwrite = TRUE)
```

### 2024-09-22 papers/article/active-travel-behaviour-evolution-in-Canada/active-travel-behaviour-evolution-in-Canada.Rmd
Adding the rmd with the methodology and already-processed data

@@
 
-The merit of this function is due to its ability to assign decreasing influences to more remote opportunities, giving a more accurate estimate for shorter journeys [@iacono2010; @kanafani1983transportation; @fotheringham1989spatial].
+The merit of this function is due to its ability to assign decreasing influences to more remote opportunities, giving a more accurate estimate for shorter trips [@iacono2010; @kanafani1983transportation; @fotheringham1989spatial].
 However, in addition to determine the form of the impedance function, scholars also need to specify the variable used to measure impedance, which can be either time, distance, monetary cost, a combination these last variables or even a generalized cost concept.
-
-The choice between time and distance as the impedance variable has been found to be acceptable based on previous research [@iacono2010; @hull2012accessibility; @sun2012measuring; @lowry2012using; @vasconcelos2012evaluation ].
-However, when it comes to non-motorized travel modes, extracting accurate travel times from existing network models can be challenging, which limits the options and makes distance a more practical choice [@handy1997measuring; @iacono2010; @yang2012walking; @arranz2019measuring].
-Furthermore, researchers specializing in active modes of transportation have faced challenges stemming from a need for more objective data concerning walking and cycling behavior.
-Estimating specialized impedance functions specific to non-motorized modes requires appropriate travel survey data that can capture pedestrian and cycle behavior.
-Often, researchers have resorted to relying on retrospective questionnaires, which assess subjective aspects such as the frequency and duration of walking and cycling activities.
-Notably, regional household travel surveys, including trips made by non-motorized modes, have been employed for this purpose [@iacono2010; @millward2013active].
-
-In contrast to these localized surveys, some datasets provide a nationwide perspective, encompassing travel for various trip purposes and offering insights into details like travel episode origins, destinations, and time-based lengths.
-This comprehensive approach furnishes a more holistic understanding of active transportation behavior.
-Nevertheless, only some studies, such as Yang et al.
-(2012), have examined active travel behavior nationally.
-Addressing the previously identified gaps and precisely the challenge of using uniform decay curves, such as the negative exponential function for both cycling and walking in accessibility assessments, this study endeavors to identify the most appropriate distance decay function for various destinations and time periods.
-Our research delves into the complexities of actual travel behavior, emphasizing active transportation modes.
-We utilize historical data from the General Social Survey (GSS) from 1986 to 2015 in Canada to calculate the impedance function for cycling and walking trips.
-By doing so, this study conducts a comparative analysis of travel behaviors associated with these two modes.
-
-Moreover, we recognize that non-work travel encompasses a spectrum of trip intentions and diverse traveler behaviors, making the impedance function an essential analytical tool for examining non-work accessibility.
-Grengs (2015) emphasizes the importance of crafting distinct functions for each travel purpose, a principle that guides our analysis [@grengs2015nonwork].
-Our investigation will encompass a variety of trip intents, ranging from commutes to homes, workplaces, or educational institutions to social visits, outdoor activities, business trips, shopping, cultural outings to libraries, museums, or theaters, dining out, and engaging in religious practices.
-By drawing on a nationally representative sample of Canadian residents, our research intends to bridge the empirical data gap concerning the frequency and duration of typical pedestrian and cycling trips for different purposes.
-In doing so, it aspires to provide a nuanced understanding of active travel behavior.
-Ultimately, this comprehensive analysis seeks to contribute meaningfully to the ongoing conversation on active transportation, shedding light on its role in shaping travel behavior and accessibility.
+Among these options, the choice between time and distance as the impedance has been found to be most used based on previous studies [@iacono2010; @hull2012accessibility; @sun2012measuring; @lowry2012using; @vasconcelos2012evaluation], with distance being more adopted in non-motorized applications since extracting accurate travel times from existing network models is challenging [@handy1997measuring; @iacono2010; @yang2012walking; @arranz2019measuring].
+Additionally, estimate impedance function to active transportation modes requires appropriate travel survey data that is able to capture pedestrian and cycle behaviour, resulting in researchers recurring to retrospective questionnaires to assess subjective aspects such as the frequency and duration of walking and cycling activities.
+Notably, regional household travel surveys that include trips made by non-motorized modes have been employed for this purpose [@iacono2010; @millward2013active].
+In opposition to these specific surveys, some data sets provides a nationwide perspective, including travel for different purposes and detailing the trip with valuable information, named episodes, regarding the origins, destinations, and time-based lengths.
+Besides this type of data can provides a deeper comprehension about the active transportation behaviour,  such as Yang et al.
+(2012), only few studies have examined travel behaviour nationally.
+It is aiming to address the mentioned challenges that this study focuses on identifying the most appropriate impedance functions for active transportation modes for various destinations and time periods in Canada.
+To do achieve our objective, we analyze historical data from the General Social Survey (GSS) from 1986 to 2015 in Canada to calculate the impedance function for cycling and walking trips.
+Additionally, this study perform a comparative analysis of travel behaviors associated with these two modes.
+
+We recognize that non-work travel encompasses a range of trip purposes and diverse traveler behaviors, which makes impedance functions essential analytical tools for studying non-work accessibility.
+Grengs (2015) emphasizes the importance of elaborating distinct functions for each travel purpose, a principle that guides this analysis [@grengs2015nonwork].
+Our investigation covers a variety of trip goals, ranging from commutes to homes, workplaces, or educational institutions to social visits, outdoor activities, business trips, shopping, cultural outings to libraries, museums, or theaters, dining out, and engaging in religious practices.
+Our research aims to enhance the current knowledge about active travel behaviour and provide empirical data about frequency and duration of typical pedestrian and cycling trips for different purposes, by applying the methodology on a nationally representative samples of Canadian residents.
+Lastly, this comprehensive analysis seeks to contribute to the ongoing conversation on active transportation, highlighting its role in influencing transportation plans to a more sustainable alternative.
 

### 2024-10-17 data-raw/01-preprocessing-source-files.Rmd
Adding province name and urban situation

--- 2024-09-22 data-raw/01-pre-processing-source-files.Rmd
@@
 ```
+
 Now, let's select only the cases of respondents who live in a CMA or CA:
 ```{r}
-gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST')]
+gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST','PRV')]
 
@@
   left_join(gss_m_2015, by = 'PUMFID') |>  # Merging the two tables
-  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
+  filter(LUC_RST %in% c(1,2,3)) |> 
+  mutate(Pop_centre =
+           case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "non CMA/CA"),
+         Province = 
+           case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"))
 ```
@@
 ```{r}
-gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST')]
+gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST','PRV')]
 
@@
 ```{r}
-gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST')]
+gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST','PRV')]
 
@@
   left_join(gss_m_2010, by = c('PUMFID' = 'RECID')) |>  # Merging the two tables
-  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
+  filter(LUC_RST %in% c(1,2,3)) |> 
+  mutate(Pop_centre =
+           case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "non CMA/CA"),
+         Province = 
+           case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"))
 ```
@@
   left_join(gss_m_2010, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
-  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
+  filter(LUC_RST %in% c(1,2,3)) |> 
+  mutate(Pop_centre =
+           case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "non CMA/CA"),
+         Province = 
+           case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"))
 ```
@@
 ```{r}
-gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST')]
+gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST','PRV')]
 
@@
   left_join(gss_m_2005, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
-  filter(LUC_RST == 1) # Selecting only respondents from CMA/CA
+  filter(LUC_RST %in% c(1,2,3)) |> 
+  mutate(Pop_centre =
+           case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "non CMA/CA"),
+         Province = 
+           case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"))
 ```
@@
 ```{r}
-gss_m_1998 <- gss_m_1998[,c('RECID','CMAPRV')]
+gss_m_1998 <- gss_m_1998[,c('RECID','CMAPRV','PRV')]
 
@@
   left_join(gss_m_1998, by = c( 'PUMFID' = 'RECID')) |>  # Merging the two tables
-  filter(CMAPRV %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
+  mutate(Pop_centre =
+           case_when(CMAPRV %in% c(6,8,11,13,15) ~ "CMA",
+                     CMAPRV %in% c(1,2,3,4,5,7,9,10,12,14) ~ "non CMA"),
+         Province = 
+           case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia")) |> dplyr::rename(
+    LUC_RST = CMAPRV) 
 ```
@@
 ```{r}
-gss_m_1992 <- gss_m_1992[,c('SEQNUM','DVCMA')]
+gss_m_1992 <- gss_m_1992[,c('SEQNUM','DVCMA','DVPROV')]
 
@@
   left_join(gss_m_1992, by = c( 'PUMFID' = 'SEQNUM')) |>  # Merging the two tables
-  filter(DVCMA %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
+  mutate(Pop_centre =
+           case_when(DVCMA %in% c(6,8,11,13,15) ~ "CMA",
+                     DVCMA %in% c(1,2,3,4,5,7,9,10,12,14) ~ "non CMA"),
+         Province = 
+           case_when(DVPROV == 0 ~ "Newfoundland and Labrador",
+                     DVPROV == 1 ~ "Prince Edward Island",
+                     DVPROV == 2 ~ "Nova Scotia",
+                     DVPROV == 3 ~ "New Brunswick",
+                     DVPROV == 4 ~ "Quebec",
+                     DVPROV == 5 ~ "Ontario",
+                     DVPROV == 6 ~ "Manitoba",
+                     DVPROV == 7 ~ "Saskatchewan",
+                     DVPROV == 8 ~ "Alberta",
+                     DVPROV == 9 ~ "British Columbia")) |> dplyr::rename(
+    LUC_RST = DVCMA,
+    PRV = DVPROV) 
 ```
@@
 left_join(gss_m_1992, by = c( 'PUMFID' = 'SEQNUM')) |>  # Merging the two tables
-  filter(DVCMA %in% c(6,8,11,13,15)) # Selecting only respondents from CMA/CA
+  mutate(Pop_centre =
+           case_when(DVCMA %in% c(6,8,11,13,15) ~ "CMA",
+                     DVCMA %in% c(1,2,3,4,5,7,9,10,12,14) ~ "non CMA"),
+         Province = 
+           case_when(DVPROV == 0 ~ "Newfoundland and Labrador",
+                     DVPROV == 1 ~ "Prince Edward Island",
+                     DVPROV == 2 ~ "Nova Scotia",
+                     DVPROV == 3 ~ "New Brunswick",
+                     DVPROV == 4 ~ "Quebec",
+                     DVPROV == 5 ~ "Ontario",
+                     DVPROV == 6 ~ "Manitoba",
+                     DVPROV == 7 ~ "Saskatchewan",
+                     DVPROV == 8 ~ "Alberta",
+                     DVPROV == 9 ~ "British Columbia")) |> dplyr::rename(
+    LUC_RST = DVCMA,
+    PRV = DVPROV) 
 ```
@@
 ```{r}
+gss_m_1986 <- gss_m_1986[,c('SEQNUM','PROV')]
+
 walking_1986 <- walking_1986 |>
   dplyr::select(SEQNUM,ACT_CODE:dest_label) |>
+    left_join(gss_m_1986, by = c( 'SEQNUM' = 'SEQNUM')) |>  # Merging the two tables
   mutate(YEAR = 1986,
-         MODE = "Walking")
-
-walking_1986 <- walking_1986 |> dplyr::rename(
+         MODE = "Walking",
+         LUC_RST = 99,
+         Pop_centre = "No information available",
+         Province = 
+           case_when(PROV == 0 ~ "Newfoundland and Labrador",
+                     PROV == 1 ~ "Prince Edward Island",
+                     PROV == 2 ~ "Nova Scotia",
+                     PROV == 3 ~ "New Brunswick",
+                     PROV == 4 ~ "Quebec",
+                     PROV == 5 ~ "Ontario",
+                     PROV == 6 ~ "Manitoba",
+                     PROV == 7 ~ "Saskatchewan",
+                     PROV == 8 ~ "Alberta",
+                     PROV == 9 ~ "British Columbia")) |> 
+  dplyr::rename(
+    PRV = PROV,
     PUMFID = SEQNUM,
@@
 # creating an integrated data frame from 1986 to 2015
-gss_episodes <- rbind(walking_2015[,1:13], cycling_2015[,1:13], walking_2010[,1:13], cycling_2010[,1:13], walking_2005[,1:13], cycling_2005[,1:13], cycling_1998[,1:13], walking_1998[,1:13], walking_1992[,1:13], cycling_1992[,1:13], walking_1986[,1:13])
+gss_episodes <- rbind(walking_2015, cycling_2015, walking_2010, cycling_2010, walking_2005, cycling_2005, cycling_1998, walking_1998, walking_1992, cycling_1992, walking_1986)
 

### 2024-10-17 data-raw/02-calculating-impedances.Rmd
Adding province name and urban situation

@@
 gss_episodes <- gss_episodes %>%
-  filter(DURATION <= 100 & DURATION > 0)
+  filter(DURATION <= 100 & 
+           DURATION > 0 & 
+           YEAR > 1986 & 
+           Pop_centre %in% c("CMA/CA", "CMA"))
 
@@
 gss_impedances <- gss_episodes[,c("PUMFID","DURATION","LOCATION",
-                  "origin","destination","orig_label",
+                  "origin","destination","Province","Pop_centre","orig_label",
                   "dest_label","YEAR","MODE","f_name",

### 2024-11-05 data-raw/01-preprocessing-source-files.Rmd
updating episode and impedance data

@@
 gss_episodes <- rbind(walking_2015, cycling_2015, walking_2010, cycling_2010, walking_2005, cycling_2005, cycling_1998, walking_1998, walking_1992, cycling_1992, walking_1986)
-
-# Change the name of some values in the dest column in the dataset 
-# Ensure that 'dest_label' is the correct column name
-gss_episodes <- gss_episodes |>
-  mutate(dest_label = case_when(
-    dest_label == "Restaurant/bar" ~ "Restaurant, bar or club",
-    dest_label == "Outdoors away from home" ~ "Outdoors",
-    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
-    dest_label == "Restaurant, bar or club" ~ "Restaurant",
-    dest_label == "Sports centre, field or arena" ~ "Sport area",
-    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
-    dest_label == "In the neighbourhood" ~ "Outdoors",
-    dest_label == "Library, museum or theatre" ~ "Library",
-    TRUE ~ dest_label  # This line ensures other values remain unchanged
-  ))
 ```

### 2024-11-08 Data-Paper/Manuscrip-Data-Package.tex
Update documentation

\documentclass[Royal,times,sageh]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\setcitestyle{aysep={,}}

\title{A demonstration of the \LaTeX class file for Statistical Methods
in Medical Research with Rmarkdown}

\runninghead{Uthor \emph{et al}.}

\author{A. Uthor*\affilnum{1,2}, O. Tro\affilnum{2}, O.
Vriga\affilnum{3}}

\affiliation{\affilnum{1}{Department of Incredible Research, University
A, City A, Country A}\\\affilnum{2}{Department of Applied Things,
University B, City B, Country B}\\\affilnum{3}{Very Important Stuff
Committee, Institute C, City C, Country C}}

\corrauth{Corresponding author name, This is sample corresponding
address.}

\email{\href{mailto:correspondingauthor@protonmail.com}{\nolinkurl{correspondingauthor@protonmail.com}}}

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut elit
odio. Donec fermentum tellus neque, vitae fringilla orci pretium vitae.
Fusce maximus finibus facilisis. Donec ut ullamcorper turpis. Donec ut
porta ipsum. Nullam cursus mauris a sapien ornare pulvinar. Aenean
malesuada molestie erat quis mattis. Praesent scelerisque posuere
faucibus. Praesent nunc nulla, ullamcorper ut ullamcorper sed, molestie
ut est. Donec consequat libero nisi, non semper velit vulputate et.
Quisque eleifend tincidunt ligula, bibendum finibus massa cursus eget.
Curabitur aliquet vehicula quam non pulvinar. Aliquam facilisis tortor
nec purus finibus, sit amet elementum eros sodales. Ut porta porttitor
vestibulum. Integer molestie, leo ut maximus aliquam, velit dui iaculis
nibh, eget hendrerit purus risus sit amet dolor. Sed sed tincidunt ex.
Curabitur imperdiet egestas tellus in iaculis. Maecenas ante neque,
pretium vel nisl at, lobortis lacinia neque. In gravida elit vel
volutpat imperdiet. Sed ut nulla arcu. Proin blandit interdum ex sit
amet laoreet. Phasellus efficitur, sem hendrerit mattis dapibus, nunc
tellus ornare nisi, nec eleifend enim nibh ac ipsum. Aenean tincidunt
nisl sit amet facilisis faucibus. Donec odio erat, bibendum eu imperdiet
sed, gravida luctus turpis.
\end{abstract}

\keywords{Class file; \LaTeX; SMMR; Rmarkdown;}

\maketitle

\hypertarget{the-article-header-information}{%
\section{The Article Header
Information}\label{the-article-header-information}}

YAML header:

\begin{verbatim}
output:
  rticles::sim_article:
    keep_tex: TRUE
\end{verbatim}

Configure the YAML header including the following elements:

\begin{itemize}
\item
  \texttt{title}: Title
\item
  \texttt{runninghead}: Author last names, use \emph{et al.} if there
  are three or more authors.
\item
  \texttt{author}: List of author(s) containing \texttt{name} and
  \texttt{num}
\item
  \texttt{corrauth}: Corresponding author's name and address.
\item
  \texttt{email}: Correspondence email
\item
  \texttt{abstract}: Limited to 200 words
\item
  \texttt{keywords}: Keywords for the article
\item
  \texttt{bibliography}: BibTeX \texttt{.bib} file
\item
  \texttt{bibliographystyle}: sageh or sagev
\item
  \texttt{classoption}: options of the \texttt{sagej} class
\end{itemize}

\hypertarget{remarks}{%
\subsection{Remarks}\label{remarks}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
  \texttt{bibliographystyle}
\item
  \texttt{classoption}
\item
  Keywords are separated by commas.
\end{enumerate}

\hypertarget{the-body-of-the-article}{%
\section{The Body of the Article}\label{the-body-of-the-article}}

\hypertarget{mathematics}{%
\subsection{Mathematics}\label{mathematics}}

Use mathematics in Rmarkdown as usual.

\hypertarget{figures-and-tables}{%
\subsection{Figures and Tables}\label{figures-and-tables}}

Figures are supported from R code:

\begin{verbatim}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
\end{verbatim}

\begin{figure}
\includegraphics[width=1\linewidth]{Manuscrip-Data-Package_files/figure-latex/plot-ref-1} \caption{Fancy Caption\label{fig:plot}}\label{fig:plot-ref}
\end{figure}

\ldots and can be referenced (Figure \ref{fig:plot}) by including the
\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
\texttt{fig.cap} attribute of the R chunk:
\texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.
It is a quirky hack at the moment, see
\href{https://github.com/yihui/knitr/issues/323}{here}.

Analogously, use Rmarkdown to produce tables as usual:

\begin{verbatim}
if (!require("xtable")) install.packages("xtable")
\end{verbatim}

\begin{verbatim}
## Loading required package: xtable
\end{verbatim}

\begin{verbatim}
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
\end{verbatim}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & speed & dist \\ 
  \hline
1 & 4.00 & 2.00 \\ 
  2 & 4.00 & 10.00 \\ 
  3 & 7.00 & 4.00 \\ 
  4 & 7.00 & 22.00 \\ 
  5 & 8.00 & 16.00 \\ 
  6 & 9.00 & 10.00 \\ 
   \hline
\end{tabular}
\caption{A table} 
\label{tab:table}
\end{table}

Referenced via \ref{tab:table}. You can also use the YAML option
\texttt{header-includes} to includes custom \LaTeX packages for tables
(keep in mind that \texttt{pandoc} uses \texttt{longtables} by default,
and it is hardcoded; some things may require including the package
\texttt{longtable}). E.g., using \texttt{ctable}:

\begin{verbatim}
header-includes:
- \usepackage{ctable}
\end{verbatim}

Then, just write straight-up \LaTeX code and reference is as usual
(\texttt{\textbackslash{}ref\{tab:ctable\}}):

\begin{verbatim}
\ctable[cap = {Short caption},
        caption = {A caption for this table.},
        label={tab:ctable},]
        {cc}
        {
        \tnote[$\ast$]{Footnote 1}
        \tnote[$\dagger$]{Other footnote}
        \tnote[b]{Mistakes are possible.}
        }{
        \FL
        COL 1\tmark[a] & COL 2\tmark[$\ast$]
        \ML
        6.92\tmark[$\dagger$] & 0.09781 \\
        6.93\tmark[$\dagger$] & 0.09901 \\
        97 & 2000
        \LL
}
\end{verbatim}

It is also possible to set the \texttt{YAML} option
\texttt{longtable:\ true} and use markdown tables (or the
\texttt{knitr::kable} function): \texttt{knitr::kable(head(cars))}
produces the same table as the \texttt{xtable} example presented before.

\hypertarget{cross-referencing}{%
\subsection{Cross-referencing}\label{cross-referencing}}

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using
\texttt{{[}@\textless{}name\textgreater{}{]}}, equivalent of
\texttt{\textbackslash{}ref\{\textless{}name\textgreater{}\}} and
\texttt{\textbackslash{}label\{\textless{}name\textgreater{}\}}). That
works well for citations in Rmarkdown, not so well for figures and
tables. In that case, it is possible to revert to standard
\LaTeX syntax.

\hypertarget{double-spacing}{%
\subsection{Double Spacing}\label{double-spacing}}

If you need to double space your document for submission please use the
\texttt{doublespace} option in the header.

\hypertarget{bibliography}{%
\section{Bibliography}\label{bibliography}}

Link a \texttt{.bib} document via the YAML header, and bibliography will
be printed at the very end (as usual). The default bibliography style is
provided by Wiley as in \texttt{WileyNJD-AMA.bst}, do not delete that
file.

Use the Rmarkdown equivalent of the \LaTeX citation system using
\texttt{{[}@\textless{}name\textgreater{}{]}}. Example:
\citep{Taylor1937}, \citep{Knupp1999, Kamm2000}.

To include all citation from the \texttt{.bib} file, add
\texttt{\textbackslash{}nocite\{*\}} before the end of the document.

\hypertarget{further-information}{%
\section{Further information}\label{further-information}}

All \LaTeX enviroments supported by the main template are supported here
as well; see the \texttt{.tex} sample file
\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm}{here}
for more details and example.

\bibliographystyle{sageh}
\bibliography{bibfile}

\end{document}

### 2024-11-08 data-raw/01-preprocessing-source-files.Rmd
Update documentation

@@
     latex_engine: xelatex
+bibliography: activeca.bib
+link-citations: yes
 editor_options: 
@@
 This Rmarkdown file is part of the
-[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. This
-package is one of the products developed by the [*Mobilizing
-Justice*](https://mobilizingjustice.ca/)[^1].
-
-[^1]: The Mobilizing Justice project is a multidisciplinary and
-    multi-sector collaboration with the objective of understand and
-    address transportation poverty in Canada and to improve the
-    well-being of Canadians at risk of transport poverty. The Social
-    Sciences and Humanities Research Council (SSRHC) has provided
-    funding for the project, which was created by an unprecedented
-    alliance of academics from various Canadian provinces and
-    institutions, transportation firms, and nonprofit organizations
-
-The main objective of the *ActiveCA* package is to identify the most
+[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. The main objective of the *ActiveCA* package is to identify the most
 appropriate impedance functions for active transportation modes for
-various destinations and time periods in Canada. To do achieve our
-objective, we analyzed historical data from the [General Social Survey
+various destinations and time periods in Canada. To do achieve this objective, we analyzed historical data from the [General Social Survey
 (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
@@
 the following analyses.
+
+## Suggested reading
+
+-   Soukhov, A., Páez, A. (2024). Accessibility analysis for planning
+    applications (Report No. MJ-A2-0002). Mobilizing Justice.
+    <https://github.com/soukhova/MJ-Accessibility-Blogs>
 
@@
 
-Additionally, our analysis will focus on developing impedance functions considering only the population who lives in larger urban population centre (a Census Metropolitan Area (CMA) or Census Agglomeration (CA))^[For definitions of CMA and CA check the this [website](https://www12.statcan.gc.ca/census-recensement/2021/ref/dict/az/Definition-eng.cfm?ID=geo009).].
-We decided to apply this restriction because the travel behavior among residents of the CMA and CA areas, in terms of active travel, tends to be very different from those outside these large urban population centers.
-The variable that informs if the respondent resides in a CMA or CA region is the variable named `LUC_RST`, present in the Main File survey.
-According to the Statistics Canada, `LUC_RST` = 1 indicates that the resident is from a CMA or CA region.
-
 The file name pattern is as follows: `transportation mode`[^3] + `_` +
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_2015 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 cycling_2015 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_2010 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 cycling_2010 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_2005 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 cycling_2005 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_1998 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 cycling_1998 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_1992 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 cycling_1992 |>
@@
 
-```{r echo = FALSE}
+```{r include=FALSE}
 walking_1986 |>

### 2024-11-08 data-raw/02-calculating-impedances.Rmd
Update documentation

--- 2024-09-22 data-raw/02-calculating-impedances.Rmd
@@
 ---
-title: "Calculating impedances for walking in cycling"
+title: "Calculating impedances for CMA and CA areas from the GSS surveys"
 author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"
@@
 This Rmarkdown file is part of the
-[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. This
-package is one of the products developed by the [*Mobilizing
-Justice*](https://mobilizingjustice.ca/)[^1].
-
-[^1]: The Mobilizing Justice project is a multidisciplinary and
-    multi-sector collaboration with the objective of understand and
-    address transportation poverty in Canada and to improve the
-    well-being of Canadians at risk of transport poverty. The Social
-    Sciences and Humanities Research Council (SSRHC) has provided
-    funding for the project, which was created by an unprecedented
-    alliance of academics from various Canadian provinces and
-    institutions, transportation firms, and nonprofit organizations
-
-The main objective of the *ActiveCA* package is to identify the most
+[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package.The main objective of the *ActiveCA* package is to identify the most
 appropriate impedance functions for active transportation modes for
@@
 In this R markdown, we'll calculate the impedance functions for walking and cycling for each GSS year (from 1986 to 2015).
+
+## Suggested reading
+
+-   Soukhov, A., Páez, A. (2024). Accessibility analysis for planning
+    applications (Report No. MJ-A2-0002). Mobilizing Justice.
+    <https://github.com/soukhova/MJ-Accessibility-Blogs>
 
@@
 
+Soukhov and Páez [-@MJ-A2-0002] explained the process of selecting the best impedance function in detail in a report released by the [*Mobilizing Justice* partnership](https://mobilizingjustice.ca/)[^1] using data from the city of Hamilton, Ontario.
+Subsequently, Santos and Páez (2024) adapted the methodology to create [**CommuteCA**](https://github.com/dias-bruno/CommuteCA), a package to obtain accessibility measures for any Canadian city using the 2021 Population Census as the main data source.
+Here, we will apply the methodology developed by Santos and Páez (2024) to obtain the impedance functions of the GSS surveys.
+
+[^1]: The Mobilizing Justice project is a multidisciplinary and
+    multi-sector collaboration with the objective of understand and
+    address transportation poverty in Canada and to improve the
+    well-being of Canadians at risk of transport poverty. The Social
+    Sciences and Humanities Research Council (SSRHC) has provided
+    funding for the project, which was created by an unprecedented
+    alliance of academics from various Canadian provinces and
+    institutions, transportation firms, and nonprofit organizations
+
 To exemplify the process of selecting the best impedance function, we will use the example of the 1986 GSS dataset, already pre-processed in the `00-Reading-Source-Files.Rmd`:
+
 ```{r}
@@
 Analyzing the histogram above, we can identify that the just a few values of duration are higher than 1 hundred minutes.
-To better visualize the variable distribution, let's use an other type of graph, the boxplot graph:
+To better visualize the variable distribution, let's use an other type of graph, the box plot graph:
 
@@
 ```
+After excluding trip longer than 100 minutes duration, the minimum duration value for walking trips is `1` minute (same as before),  while the maximum is `95` minutes.
+The new average duration for walking trips is `17.03` minutes and the median is `10` minutes.
 
@@
 
-The next chunks of code of this Rmarkdown explains the process of calibrating PDF
-functions for the Walking dataset of the 1986 GSS.
-
 ## Empirical travel time
 
-Creating a data frame with the all empirical travel times:
+The next chunks of code of this Rmarkdown explains the process of calibrating PDF
+functions for the Walking dataset of the 1986 GSS. Creating a data frame with the all empirical travel times:
 
@@
 direct optimization algorithm are used [@mullerdutang2015].
+
+The code below introduces two functions: `test_distributions_weighted` and `lowest_aic`.
+The first takes the time values as input and applies them to the distributions mentioned above, obtaining an AIC value for each distribution.
+In turn, `lowest_aic` compares the AIC values of each distribution and selects the function with the minimum AIC (the distribution with the best fit considering the time values provided).
 
@@
 
-Obtaining the theoretical impedance function for the each destination mode:
+The next chunk selects the best impedance function, using the `test_distributions_weighted` and `lowest_aic` functions for each destination and transportation mode (in this case, walking since the 1986 survey does not have cycling as an mode option).
+The name pattern is as follows: `f_year_mode_dest` + `destination`, which means that `f_1986_Walking_dest1` displays the best function for the year 1986, walking mode, destination 1 (Home).
+
 ```{r}
@@
 
-Actually, the `Log-normal distribution` was the best fuction for the others two destination from the survey: `2` (`Work or school`) and `3` (`Other's home`).
+Actually, the `Log-normal distribution` was the best function for the others two destination from the survey: `2` (`Work or school`) and `3` (`Other's home`).
 
@@
 
-Building the theoretical values for each destination based on the best
+Obtaining the theoretical values for each destination based on the best
 impedance function for each destination:
@@
 
+Visualizing the theoretical values (first 10 rows):
+```{r}
+tld_theoretical[1:10,]
+```
+
 Plotting the theoretical (black lines) and the empirical travel time
@@
 ```{r travel-time-plots}
-par(mfrow = c(2, round(length(unique(tld_theoretical$destination))/2)))  # Arrange plots in 2 columns
+par(mfrow = c(2, round(length(unique(tld_theoretical$destination))/2)))
 
@@
 for (destination in unique(tld_empirical$destination)) {
-  # Subset empirical and theoretical data for the current destination
   empirical_subset <- tld_empirical[tld_empirical$destination == destination, ]
@@
 
-  # Create histogram for empirical data
 hist(empirical_subset$DURATION, freq = FALSE, col = "grey70", breaks = 20, main = paste(empirical_subset$dest_label[1]),
@@
 
-  # Add theoretical curve
   lines(theoretical_subset$t, theoretical_subset$f)
@@
 
+After exemplifying the process of selecting the best impedance function for the 1986 GSS survey, we will obtain the function for all GSS cycles.
+First, we'll load the GSS episodes pre-processed in the first rmarkdown:
+
 ```{r}
@@
 
+Additionally, our analysis will focus on developing impedance functions considering only the population who lives in larger urban population centre (a Census Metropolitan Area (CMA) or Census Agglomeration (CA))^[For definitions of CMA and CA check the this [website](https://www12.statcan.gc.ca/census-recensement/2021/ref/dict/az/Definition-eng.cfm?ID=geo009).].
+We decided to apply this restriction because the travel behavior among residents of the CMA and CA areas, in terms of active travel, tends to be very different from those outside these large urban population centers.
+The variable that informs if the respondent resides in a CMA or CA region is the variable named `Pop_centre`, obtained after some processing made with he Main File survey.
+
 ```{r}
 gss_episodes <- gss_episodes %>%
-  filter(DURATION <= 100 & DURATION > 0)
+  filter(DURATION <= 100 & 
+           DURATION > 0 & 
+           YEAR > 1986 & 
+           Pop_centre %in% c("CMA/CA", "CMA"))
+```
 
+Furthermore, to calculate the function, we need a combination of `mode`, `year` and `destination` greater than 1, which means that we need at least two episodes to be able to fit a function.
+Consequently, combinations with a greater number of episodes will exhibit better functions in terms of AIC and will better represent the empirical values.
+
+```{r}
 gss_episodes <- gss_episodes %>%
@@
 
+Similar as we did before, the next chunk selects the best impedance function, using the `test_distributions_weighted` and `lowest_aic` functions for each destination, transportation mode and year.
+The name pattern is as follows: `f_year_mode_dest` + `destination`, which means that `f_1986_Walking_dest1` displays the best function for the year 1986, walking mode, destination 1 (Home).
+/000
 ```{r}
@@
 ```
+After selecting the best function for every combination of destination, transportation mode and year, the next two chunks will apply the functions in all surveys, adding the information with the best function and its parameters, AIC/BIC/loglik values, and count of episodes that were used to calculate the function:
 
@@
 ```
+Creation of a new dataset called `gss_impedances`, to keep the original episode files as they are after the pre-processing stage:
 
@@
 gss_impedances <- gss_episodes[,c("PUMFID","DURATION","LOCATION",
-                  "origin","destination","orig_label",
+                  "origin","destination","Province","Pop_centre","orig_label",
                   "dest_label","YEAR","MODE","f_name",
-                  "f","est_1","est_2","loglik","aic","bic")]
+                  "f","est_1","est_2","loglik","aic","bic","count")]
 ```

### 2024-11-08 R/ActiveCA-package.R

@@ -1,0 +1,466 @@
+#' ActiveCA: Processed data to obtain impedance functions for active transportation modes in Canada
+#'
+#' The main objective of the *ActiveCA* package is to provide processed data to identify
+#' the most appropriate impedance functions for active transportation modes considering
+#' various destinations and time periods in Canada. To do achieve this objective,
+#' we processed historical data from the [General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
+#' from 1986 to 2015 in Canada. We also calculate impedance function for cycling
+#' and walking trips for the Census Agglomerations and Census Metropolitan Areas.
+#'
+#' @docType data
+#' @name ActiveCA-package
+#' @author Bruno Dos Santos, School of Earth, Environment and Society, McMaster University \email{dossanb@@mcmaster.ca}
+#' @author Mahdis Moghadasi, School of Earth, Environment and Society, McMaster University \email{paezha@@mcmaster.ca}
+#' @author Antonio Paez, School of Earth, Environment and Society, McMaster University \email{paezha@@mcmaster.ca}
+#' @references \url{https://github.com/dias-bruno/ActiveCA}
+NULL
+
+##############################################
+##                DATA SETS                 ##
+##############################################
+
+#' Preprocessed walking episodes (1986).
+#'
+#' This dataset contains all the walking episodes from the 1986 GSS Survey.
+#' We identified the walking episodes and selected the places of origin
+#' and destination, measuring the time spent, weight and other
+#' information about the activities.
+#'
+#' @format A data frame containing 4347 rows and 17 variables; each row represents a unique episode.
+#'   \describe{
+#'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
+#'   \item{WGHT_EPI}{Episode weight.}
+#'   \item{ACTCODE}{Activity code of the episode.}
+#'   \item{STARTIME}{Start time (in minutes) of the episode.}
+#'   \item{ENDTIME}{End time (in minutes) of the episode.}
+#'   \item{DURATION}{Duration (in minutes) of the episode.}
+#'   \item{LOCATION}{Location/mode of transport of the episode.}
[…]

### 2024-11-12 Data-Paper/Manuscript-Data-Package.Rmd
Manuscript

---
title: "ActiveCA: an open data product to obtain impedance functions for active transportation modes in Canada"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This article describes and visualizes the data contained in the {ActiveCA} R data package. {ActiveCA} contains open data products designed to obtain impedance functions for active transportation modes in Canada, retrieved from the General Social Survey collections from 1986 to 2015. The package provides data tables on walking and cycling episodes, detailing the trip origins, destinations, and duration of each episode. The origin and destination categories covers a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others. Addittionally, the package includes details on the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency. {ActiveCA} enables users to calculate the average travel time for each origin-destination combination for each survey year, considering the two active transportation modes: walking and cycling. For each Census Metropolitan and Agglomeration Area, we estimated distance decay curves (impedance functions) for each origin-destination. The package will continue to expand with contributions from the authors and the broader community through requests in the future. {ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data and all open data products are detailed."

keywords: Active; mobility; walking; cycling; travel time; impedance; transportation; R 
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  warning = FALSE,
  comment = '', 
  out.width = "1\\linewidth")
```

```{r install-data-package, include=FALSE}
if (!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA",
                        build_vignettes = TRUE)
}

rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
library(ActiveCA)
library(here)
library(dplyr)
library(tidyr)
# library(fitdistrplus)
library(ggplot2)
library(kableExtra)
library(reshape2)

# library(patchwork)
# library(sf)
# library(scales)
# library(stats)
# library(ggspatial)
# library(shadowtext)
# library(tmap)
# library(ggpmisc)
```

```{r load-data, include = FALSE}
load(paste0(here(), "/data/gss_episodes.Rda"))
load(paste0(here(), "/data/gss_impedances.Rda"))

```

# Introduction

This manuscript presents the open data product {ActiveCA}. Open data products (ODPs) are the outcome of a transparent process that transforms raw data (open or not) into analysis-ready data, in which all stages of development follow open principles. ODPs differ from general open data due to their high utility, added value and open availability. The product presented in this document is an `R` data package that currently consists of processed data tables retrieved from the General Social Survey (GSS) collections from 1986 to 2015, to obtain impedance functions for active transportation modes in Canada.

To create this `R` data package, we collected the GSS surveys, cleaned them and processed them to make it ready for analysis. An `R` data package contains code, data and documentation in a standardized collection format that can be installed by `R` users via a centralized software repository, such as CRAN (Comprehensive R Archive Network) and GitHub. Although the GSS surveys are publicly sourced and managed by Statistic Canada, preparing them for analysis can be time-consuming, tedious and perhaps not even possible for those who try, due to a lack of documentation or prior knowledge.

The aim of this paper is to walk readers through the data sets and invite others to experiment in its uses and applications. {ActiveCA} is freely available on GitHub for all to install and freely use in the spirit of open and reproducible research. Although {ActiveCA} was designed to obtain impedance functions, we admit and hope that its use can be adopted in various applications that even go beyond the range of possibilities we have imagined. Not only the data, but also all the code documenting the processing methodology is available for consultation and evaluation in its repository. This package contributes to reducing the barrier to using the information contained in GSS surveys to provide data-driven decisions in transportation analysis.

# General Social Survey (GSS) collection

In Canada, Statistics Canada produces the General Social Surveys (GSS)
to collect data on social trends in order to track changes in the living
conditions of Canadians and well-being over time. This survey tracks
changes in time and it is used to understand how Canadians spend and
manage their time and what factors contribute to their happiness and
stress. The survey was created in 1985 as a series of independent,
annual, cross-sectional surveys, each covering one topic in depth.
Caregiving, families, time use, social identity, volunteering, and
victimization are among the current themes of the GSS.

The six survey themes listed above are repeated every five years. In
addition to the central topic, each cycle includes new content that
addresses emerging and policy-relevant issues, collecting detailed
socio-demographic information, such as age, gender, education, religion,
ethnicity, income, etc. Time-use surveys collect data on all human
activities and can inform a wide range of policies. In particular, three
key themes have been identified as essential for policymaking and for
which no other data source is adequate: unpaid work and non-market
production, well-being, and gender equality. Leisure time, work-life
balance, health, commuting, culture, and sports are among the other
topics covered by time-use surveys. In addition, statistics Canada has
been conducting time-use surveys at five to seven year intervals since
1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
and 2015). The time use collects information on participation of
respondents and time spent on a wide range of day-to-day activities
using a 24-hour retrospective diary. Also, information is collected on
the location of these activities (e.g., at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity.

## The Main File

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, enhancing the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes. This file is aggregated and prohibits the direct association
between specific activities with particular locations or companions.

## The Episode File

The Episode File records detailed data for each activity episode
reported by respondents. Each episode entry includes the start and end
times, duration, location, and accompanying social context, informing
when and where activities occurred and with whom. The file distinguishes
itself by focusing on individual episodes rather than respondents, with
the data structured around the numerous activity instances that compose
a respondent's day. Although respondent-specific characteristics are not
included within the Episode File, it is possible to link the Main File
and the Episode File by using an identifiable variable present in both
data sets.

# Average travel times

```{r ch03-make-table-01, echo=FALSE, cache=FALSE}

# Calculate the count of trips for each mode and year
# Summarize counts by year and mode
episodes_counts <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = YEAR, values_from = count)

# Calculate min, mean, and max statistics
stats_data <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(
    min = min(DURATION),
    mean = mean(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(min, mean, median, max), names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
final_data <- bind_rows(
  episodes_counts %>% mutate(Statistic = "count"),
  stats_data
) %>%
  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) %>%
  arrange(MODE, Statistic) %>%
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Rename the 'MODE' column to 'Mode' before generating the table
final_data <- final_data %>%
  rename(Mode = MODE) %>%
  dplyr::select(Mode, Statistic, everything())

# Generate the table with 'Statistic' as the second column
styled_ch03_table_01 <- final_data %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:ch03-table-01}Descriptive Analysis of Active Transportation Modes: Walking and Cycling Statistics from 1986 to 2015",  
        align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 2, "Year" = ncol(final_data) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01
```

```{r ch03-make-table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
# Filter for selected years
episodes_stats_before <- gss_episodes %>%
  filter(YEAR < 2005) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- episodes_stats_before %>% 
  group_by(YEAR, MODE) %>% 
  summarise(total_count = sum(count), 
            .groups = "drop")

episodes_stats_before <- left_join(episodes_stats_before, total_counts, by = c("YEAR", "MODE"))

episodes_stats_before$Percentage <- round((episodes_stats_before$count / episodes_stats_before$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_before %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  )  %>% 
  arrange(MODE, desc(Percentage_1998))

wide_data <- wide_data %>%
  group_by(MODE) %>%
  mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup() %>%
 dplyr::select(
    dest_label, MODE,
    min_1986, median_1986, max_1986, Percentage_1986,
    min_1992, median_1992, max_1992, Percentage_1992,
    min_1998, median_1998, max_1998, Percentage_1998
  )

styled_ch03_table_02 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:ch03-table-02}Comparative Trip Stats by Mode and Destination: 1986, 1992, 1998",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 6) %>%
  # column_spec(1, width = "1cm") %>%  # Set width of 'Dest' column to 1.5cm
  # column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>% 
  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_02 
```

```{r ch03-make-table-03,echo=FALSE, cache=FALSE, warning=FALSE}
# Rename certain 'dest_label' values
gss_episodes_renamed <- gss_episodes %>%
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))

# Filter for walking and cycling trips, selected years, and duration <= 90
episodes_stats_after <- gss_episodes_renamed %>%
  filter(YEAR > 2000) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
episodes_stats_after <- left_join(episodes_stats_after, total_counts, by = c("YEAR", "MODE"))
episodes_stats_after$Percentage <- round((episodes_stats_after$count / episodes_stats_after$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_after %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  ) %>%
 dplyr::select(
    dest_label, MODE,
    min_2005, median_2005, max_2005, Percentage_2005,
    min_2010, median_2010, max_2010, Percentage_2010,
    min_2015, median_2015, max_2015, Percentage_2015
  ) %>% 
  arrange(MODE, desc(Percentage_2015)) %>%
  group_by(MODE) %>%
  mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup()

styled_ch03_table_03 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:ch03-table-03}Comparative Trip Statistics by Transportation Mode and Destination: 2005, 2010, and 2015",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                font_size = 6) %>%
  column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
  column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>% 
  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_03
```

```{r ch03-make-Heatmap-walking_2015, echo=FALSE, cache=FALSE, include=FALSE}
gss_heatmap <- gss_episodes_renamed %>%
  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) %>% 
  filter(YEAR %in% c(1992,2015))

for(year in unique(gss_heatmap$YEAR)){

  for(mode in unique(gss_heatmap$MODE)){

    trip_counts <- gss_heatmap %>%
      filter(YEAR == year, MODE == mode) %>% 
      group_by(orig_label, dest_label) %>%
      summarize(count_each_trip = n(), 
                weighted_count = sum(WGHT_EPI), .groups = "drop")

      # Calculate the total weighted count for normalization
      total_weighted_count <- sum(trip_counts$weighted_count)

      # Calculate percentages and round to 2 decimal places
      trip_counts <- trip_counts %>%
        mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

      # Reshape data for heatmap
      heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

      # Convert the data frame to a matrix for the heatmap
      heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
      rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

      # Melt the matrix for use with ggplot
      heatmap_melted <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

      # Define the order of destinations for 2015
      destination_order <- c("Home", "Work or school", "Grocery store", 
                             "Others' home", "Outdoors", "Neighbourhood", 
                             "Cultural venues", "Place of worship", 
                             "Restaurant", "Sport area", "Business")

      # Heatmap for every combination of year and mode
      assign(paste0("heatmap_plot_", year, "_", mode),
               ggplot(heatmap_melted, aes(x = Destination, y = Origin, fill = value)) +
                  geom_tile(color = "white") +
                  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                                       na.value = "gray95", 
                                       limits = c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       breaks =  c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
                  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
                  theme_minimal() +
                  theme(
                  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
                  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
                  axis.title = element_text(size = 16),
                  legend.title = element_text(size = 16, margin = margin(b = 15)), 
                  legend.text = element_text(size = 16),
                  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt")))
             )

  } 
}
```

```{r make-figs, include=FALSE}
heatmap_plot_2015_Walking <- heatmap_plot_2015_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_2015_Cycling <- heatmap_plot_2015_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Cycling <- heatmap_plot_1992_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Walking <- heatmap_plot_1992_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

# Combine the plots with adjusted spacing
walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"), 
       plot = walking_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)

# Combine the plots with adjusted spacing
cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"), 
       plot = cycling_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

```{r echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"))
```

```{r  echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"))
```

# Impedance function for Canadians Metropolitan and Census Agglomerations areas

```{r echo=FALSE, cache=FALSE, warning=FALSE}
gss_impedances %>% 
filter(dest_label == "Work or school", MODE == "Walking") %>%
  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic","count") %>%
  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
  mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
  unique() %>% 
  kable(format = "latex",
        booktabs = TRUE,
        caption = "\\label{tab:ch03-table-04}Impedance functions and selection criteria for walking trips considering 'Work or school' as destination")
```

```{r}
gss_impedances %>% 
filter(dest_label == "Work or school", MODE == "Walking") %>%
  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic","count") %>%
  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
  mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
  unique()
```

# Concluding remarks

# References {#references .unnumbered}

### 2024-11-12 Data-Paper/Manuscript-Data-Package.tex
Manuscript

\documentclass[Royal,times,sageh]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\setcitestyle{aysep={,}}

\title{ActiveCA: an open data product to obtain impedance functions for
active transportation modes in Canada}

\runninghead{}

\author{Anon1\affilnum{}, Anon2\affilnum{}, Anon3\affilnum{}}

\affiliation{\affilnum{}{}}

\begin{abstract}
This article describes and visualizes the data contained in the
\{ActiveCA\} R data package. \{ActiveCA\} contains open data products
designed to obtain impedance functions for active transportation modes
in Canada, retrieved from the General Social Survey collections from
1986 to 2015. The package provides data tables on walking and cycling
episodes, detailing the trip origins, destinations, and duration of each
episode. The origin and destination categories covers a wide variety of
locations, such as home, work or school, libraries, museums,
restaurants, bars, sports centers, health clinics, place of worship, and
others. Addittionally, the package includes details on the respondent's
region characteristics, specifying wheter they live in a metropolitan
area and their province of residency. \{ActiveCA\} enables users to
calculate the average travel time for each origin-destination
combination for each survey year, considering the two active
transportation modes: walking and cycling. For each Census Metropolitan
and Agglomeration Area, we estimated distance decay curves (impedance
functions) for each origin-destination. The package will continue to
expand with contributions from the authors and the broader community
through requests in the future. \{ActiveCA\} is freely accessible for
exploration and download from the associated Github repository, where
the documentation and code involved in creating and manipulating data
and all open data products are detailed.
\end{abstract}

\keywords{Active; mobility; walking; cycling; travel time; impedance;
transportation; R}

\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

This manuscript presents the open data product \{ActiveCA\}. Open data
products (ODPs) are the outcome of a transparent process that transforms
raw data (open or not) into analysis-ready data, in which all stages of
development follow open principles. ODPs differ from general open data
due to their high utility, added value and open availability. The
product presented in this document is an \texttt{R} data package that
currently consists of processed data tables retrieved from the General
Social Survey (GSS) collections from 1986 to 2015, to obtain impedance
functions for active transportation modes in Canada.

To create this \texttt{R} data package, we collected the GSS surveys,
cleaned them and processed them to make it ready for analysis. An
\texttt{R} data package contains code, data and documentation in a
standardized collection format that can be installed by \texttt{R} users
via a centralized software repository, such as CRAN (Comprehensive R
Archive Network) and GitHub. Although the GSS surveys are publicly
sourced and managed by Statistic Canada, preparing them for analysis can
be time-consuming, tedious and perhaps not even possible for those who
try, due to a lack of documentation or prior knowledge.

The aim of this paper is to walk readers through the data sets and
invite others to experiment in its uses and applications. \{ActiveCA\}
is freely available on GitHub for all to install and freely use in the
spirit of open and reproducible research. Although \{ActiveCA\} was
designed to obtain impedance functions, we admit and hope that its use
can be adopted in various applications that even go beyond the range of
possibilities we have imagined. Not only the data, but also all the code
documenting the processing methodology is available for consultation and
evaluation in its repository. This package contributes to reducing the
barrier to using the information contained in GSS surveys to provide
data-driven decisions in transportation analysis.

\hypertarget{general-social-survey-gss-collection}{%
\section{General Social Survey (GSS)
collection}\label{general-social-survey-gss-collection}}

In Canada, Statistics Canada produces the General Social Surveys (GSS)
to collect data on social trends in order to track changes in the living
conditions of Canadians and well-being over time. This survey tracks
changes in time and it is used to understand how Canadians spend and
manage their time and what factors contribute to their happiness and
stress. The survey was created in 1985 as a series of independent,
annual, cross-sectional surveys, each covering one topic in depth.
Caregiving, families, time use, social identity, volunteering, and
victimization are among the current themes of the GSS.

The six survey themes listed above are repeated every five years. In
addition to the central topic, each cycle includes new content that
addresses emerging and policy-relevant issues, collecting detailed
socio-demographic information, such as age, gender, education, religion,
ethnicity, income, etc. Time-use surveys collect data on all human
activities and can inform a wide range of policies. In particular, three
key themes have been identified as essential for policymaking and for
which no other data source is adequate: unpaid work and non-market
production, well-being, and gender equality. Leisure time, work-life
balance, health, commuting, culture, and sports are among the other
topics covered by time-use surveys. In addition, statistics Canada has
been conducting time-use surveys at five to seven year intervals since
1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
and 2015). The time use collects information on participation of
respondents and time spent on a wide range of day-to-day activities
using a 24-hour retrospective diary. Also, information is collected on
the location of these activities (e.g., at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity.

\hypertarget{the-main-file}{%
\subsection{The Main File}\label{the-main-file}}

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, enhancing the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes. This file is aggregated and prohibits the direct association
between specific activities with particular locations or companions.

\hypertarget{the-episode-file}{%
\subsection{The Episode File}\label{the-episode-file}}

The Episode File records detailed data for each activity episode
reported by respondents. Each episode entry includes the start and end
times, duration, location, and accompanying social context, informing
when and where activities occurred and with whom. The file distinguishes
itself by focusing on individual episodes rather than respondents, with
the data structured around the numerous activity instances that compose
a respondent's day. Although respondent-specific characteristics are not
included within the Episode File, it is possible to link the Main File
and the Episode File by using an identifiable variable present in both
data sets.

\hypertarget{average-travel-times}{%
\section{Average travel times}\label{average-travel-times}}

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{>{}llcccccc}
\caption{\label{tab:ch03-make-table-01}\label{tab:ch03-table-01}Descriptive Analysis of Active Transportation Modes: Walking and Cycling Statistics from 1986 to 2015}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-8}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
\midrule
 & count & 4347 & 1500 & 1670 & 5533 & 4379 & 3251\\
\nopagebreak
 & max & 660 & 300 & 255 & 515 & 480 & 900\\
\nopagebreak
 & mean & 21 & 19 & 11 & 12 & 12 & 17\\
\nopagebreak
 & median & 10 & 10 & 5 & 10 & 8 & 10\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & min & 1 & 1 & 1 & 0 & 0 & 5\\
\cmidrule{1-8}\pagebreak[0]
 & count & NA & 135 & 119 & 333 & 236 & 245\\
\nopagebreak
 & max & NA & 240 & 90 & 180 & 153 & 120\\
\nopagebreak
 & mean & NA & 31 & 21 & 19 & 21 & 24\\
\nopagebreak
 & median & NA & 20 & 15 & 15 & 15 & 15\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & min & NA & 5 & 2 & 1 & 1 & 5\\
\bottomrule
\end{longtable}
\endgroup{}

\begingroup\fontsize{6}{8}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
\end{TableNotes}
\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
\caption{\label{tab:ch03-make-table-02}\label{tab:ch03-table-02}Comparative Trip Stats by Mode and Destination: 1986, 1992, 1998}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{1986} & \multicolumn{4}{c}{1992} & \multicolumn{4}{c}{1998} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode*}} & \multicolumn{1}{c}{\textbf{Min*}} & \multicolumn{1}{c}{\textbf{Med*}} & \multicolumn{1}{c}{\textbf{Max*}} & \multicolumn{1}{c}{\textbf{(\%)*}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
\midrule
Home &  & NA & NA & NA & NA & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
Work or school & Cycling & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
Other's home &  & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
Home &  & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
Other's home & Walking & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
\addlinespace
Work or school &  & 1 & 10 & 450 & 11.3 & 2 & 10 & 60 & 19.2 & 1 & 6.5 & 75 & 20.4\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\begingroup\fontsize{6}{8}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
\end{TableNotes}
\begin{longtable}[t]{>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}}
\caption{\label{tab:ch03-make-table-03}\label{tab:ch03-table-03}Comparative Trip Statistics by Transportation Mode and Destination: 2005, 2010, and 2015}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{2005} & \multicolumn{4}{c}{2010} & \multicolumn{4}{c}{2015} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
\multicolumn{1}{>{\centering\arraybackslash}p{2cm}}{\textbf{Destination}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Mode*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)}}\\
\midrule
Home &  & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
Work or school & Cycling & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
Grocery store &  & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
Other's home &  & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
Restaurant &  & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
\addlinespace
Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 15 & 2.9\\
Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 90 & 2.0\\
Outdoors &  & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 30.0 & 45 & 1.2\\
Cultural venues &  & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
\addlinespace
Place of worship &  & 20 & 20.0 & 20 & 0.3 & NA & NA & NA & NA & 15 & 15.0 & 15 & 0.4\\
Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
Work or school & Walking & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
Restaurant &  & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
\addlinespace
Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 45 & 3.3\\
Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
Cultural venues &  & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
\addlinespace
Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking_hm_fig}

}

\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:unnamed-chunk-1}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/cycling_hm_fig}

}

\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:unnamed-chunk-2}
\end{figure}

\hypertarget{impedance-function-for-canadians-metropolitan-and-census-agglomerations-areas}{%
\section{Impedance function for Canadians Metropolitan and Census
Agglomerations
areas}\label{impedance-function-for-canadians-metropolitan-and-census-agglomerations-areas}}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}\label{tab:ch03-table-04}Impedance functions and selection criteria for walking trips considering 'Work or school' as destination}
\centering
\begin{tabular}[t]{rlrrrrrr}
\toprule
YEAR & f\_name & est\_1 & est\_2 & loglik & aic & bic & count\\
\midrule
2015 & lnorm & 2.55 & 0.64 & -3306028 & 6612061 & 6612069 & 407\\
2010 & lnorm & 2.24 & 0.81 & -2636159 & 5272321 & 5272329 & 367\\
2010 & gamma & 1.99 & 0.18 & -1298182 & 2596367 & 2596373 & 127\\
2005 & lnorm & 2.23 & 0.71 & -1507682 & 3015369 & 3015375 & 224\\
2005 & lnorm & 2.08 & 0.84 & -2565742 & 5131489 & 5131497 & 500\\
\addlinespace
1998 & gamma & 1.23 & 0.09 & -1159374 & 2318752 & 2318758 & 109\\
1992 & lnorm & 2.38 & 0.70 & -1159698 & 2319400 & 2319406 & 113\\
\bottomrule
\end{tabular}
\end{table}

\begin{verbatim}
# A tibble: 7 x 8
   YEAR f_name est_1 est_2   loglik     aic     bic count
  <dbl> <chr>  <dbl> <dbl>    <dbl>   <dbl>   <dbl> <int>
1  2015 lnorm   2.55  0.64 -3306028 6612061 6612069   407
2  2010 lnorm   2.24  0.81 -2636159 5272321 5272329   367
3  2010 gamma   1.99  0.18 -1298182 2596367 2596373   127
4  2005 lnorm   2.23  0.71 -1507682 3015369 3015375   224
5  2005 lnorm   2.08  0.84 -2565742 5131489 5131497   500
6  1998 gamma   1.23  0.09 -1159374 2318752 2318758   109
7  1992 lnorm   2.38  0.7  -1159698 2319400 2319406   113
\end{verbatim}

\hypertarget{concluding-remarks}{%
\section{Concluding remarks}\label{concluding-remarks}}

\bibliographystyle{sageh}
\bibliography{bibfile.bib}

\end{document}

### 2024-11-12 data-raw/01-preprocessing-source-files.Rmd
Manuscript

@@
                      origin == 306 ~ "Grocery store, other stores or mall",
+                     origin == 307 ~ "Library, museum or theatre",
                      origin == 308 ~ "Sports centre, field or arena",
-                     origin == 307 ~ "Library, museum or theatre",
                      origin == 309 ~ "Restaurant, bar or club",
@@
                      origin == 3 ~ "Other's home",
-                     origin == 9 ~ "Outdoors",
-                     origin == 6 ~ "Grocery store, other stores or mall",
-                     origin == 10 ~ "Library, museum or theatre",
                      origin == 4 ~ "Restaurant, bar or club",
                      origin == 5 ~ "Place of worship",
+                     origin == 6 ~ "Grocery store, other stores or mall",
                      origin == 7 ~ "Grocery store, other stores or mall",                     
-                     origin == 8 ~ "Work or school"),
+                     origin == 8 ~ "Work or school",
+                     origin == 9 ~ "Outdoors",
+                     origin == 10 ~ "Library, museum or theatre"),
          orig_label = factor(orig_label))

### 2024-11-12 data-raw/02-calculating-impedances.Rmd
Manuscript

@@
 gss_episodes <- gss_episodes %>%
-  group_by(MODE, YEAR, destination) %>%
+  group_by(MODE, YEAR, dest_label) %>%
   mutate(count = n()) %>%
@@
 Similar as we did before, the next chunk selects the best impedance function, using the `test_distributions_weighted` and `lowest_aic` functions for each destination, transportation mode and year.
-The name pattern is as follows: `f_year_mode_dest` + `destination`, which means that `f_1986_Walking_dest1` displays the best function for the year 1986, walking mode, destination 1 (Home).
+The name pattern is as follows: `f_year_mode_dest` + `dest_label`, which means that `f_1986_Walking_destHome` displays the best function for the year 1986, walking mode, Home destination.
 /000
 ```{r}
-
 for(year in unique(gss_episodes$YEAR)){
 
-  for(destino in unique(gss_episodes$destination)){
+  for(destino in unique(gss_episodes[gss_episodes$YEAR == year, ]$dest_label)){
 
-    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){
+    for(mode in unique(gss_episodes[gss_episodes$YEAR == year & gss_episodes$dest_label == destino, ]$MODE)){
 
-        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
+        gss_subset <-  gss_episodes[gss_episodes$dest_label == destino &
                             gss_episodes$YEAR == year &
@@
                     mode, ", destination ",  
-                    gss_subset$dest_label[1],
+                    destino,
                     ", for the ", year, " GSS cicle"))
@@
 
-  for(destino in unique(gss_episodes$destination)){
+  for(destino in unique(gss_episodes[gss_episodes$YEAR == year, ]$dest_label)){
 
-    for(mode in unique(gss_episodes[gss_episodes$YEAR == year, ]$MODE)){
+    for(mode in unique(gss_episodes[gss_episodes$YEAR == year & gss_episodes$dest_label == destino, ]$MODE)){
 
-        gss_subset <-  gss_episodes[gss_episodes$destination == destino &
+        gss_subset <-  gss_episodes[gss_episodes$dest_label == destino &
                             gss_episodes$YEAR == year &
@@
 
-        gss_episodes[gss_episodes$destination == destino &
+        gss_episodes[gss_episodes$dest_label == destino &
                             gss_episodes$YEAR == year &

### 2024-11-12 R/ActiveCA-package.R

@@ -1,4 +1,4 @@
-#' ActiveCA: Processed data to obtain impedance functions for active transportation modes in Canada
+#' ActiveCA: a data set to obtain impedance functions for active transportation modes in Canada
 #'
 #' The main objective of the *ActiveCA* package is to provide processed data to identify
 #' the most appropriate impedance functions for active transportation modes considering

### 2024-11-13 Data-Paper/Manuscript-Data-Package.Rmd
Data paper

@@
 ---
-title: "ActiveCA: an open data product to obtain impedance functions for active transportation modes in Canada"
+title: "ActiveCA: an open data product on active transportation episodes in Canada"
 author:
@@
 abstract: 
-  "This article describes and visualizes the data contained in the {ActiveCA} R data package.
-{ActiveCA} contains open data products designed to obtain impedance functions for active transportation modes in Canada, retrieved from the General Social Survey collections from 1986 to 2015.
+  "This paper describes the data set of the {ActiveCA} R data package.
+{ActiveCA} contains open data products to obtain impedance functions for active transportation modes in Canada, retrieved from the General Social Survey collections from 1986 to 2015.
 The package provides data tables on walking and cycling episodes, detailing the trip origins, destinations, and duration of each episode.
 The origin and destination categories covers a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.
-Addittionally, the package includes details on the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
+Addittionally, the package details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
 {ActiveCA} enables users to calculate the average travel time for each origin-destination combination for each survey year, considering the two active transportation modes: walking and cycling.
-For each Census Metropolitan and Agglomeration Area, we estimated distance decay curves (impedance functions) for each origin-destination.
+For each Census Metropolitan and Agglomeration Area, we estimated distance-decay curves (impedance functions) for each destination, transportation mode and year.
 The package will continue to expand with contributions from the authors and the broader community through requests in the future.
@@
     keep_tex: yes
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
   echo = FALSE,
-  cache = TRUE,
+  cache = FALSE,
   warning = FALSE,
@@
 library(tidyr)
-# library(fitdistrplus)
 library(ggplot2)
@@
 library(reshape2)
-
-# library(patchwork)
-# library(sf)
-# library(scales)
-# library(stats)
-# library(ggspatial)
-# library(shadowtext)
-# library(tmap)
-# library(ggpmisc)
+library(rlang)
+library(scales)
+library(gridExtra)
+library(ggpubr)
+library(grid)
+library(stats)
 ```
@@
 load(paste0(here(), "/data/gss_impedances.Rda"))
-
 ```
@@
 
-This manuscript presents the open data product {ActiveCA}.
-Open data products (ODPs) are the outcome of a transparent process that transforms raw data (open or not) into analysis-ready data, in which all stages of development follow open principles.
-ODPs differ from general open data due to their high utility, added value and open availability.
-The product presented in this document is an `R` data package that currently consists of processed data tables retrieved from the General Social Survey (GSS) collections from 1986 to 2015, to obtain impedance functions for active transportation modes in Canada.
+This paper presents the open data product {ActiveCA}. Open data products
+(ODPs) are the outcome of a transparent process that transforms raw data
+(open or not) into analysis-ready data, in which all stages of
+development follow open principles [@arribas-bel2021]. ODPs differ from
+general open data due to their high utility, added value and open
+availability. The product presented in this document is an `R` data
+package that currently consists of processed data tables retrieved from
+the Time Use collections of the General Social Surveys (GSS) from 1986
+to 2015 [@statisticscanada2024], aimed to obtain impedance functions for
+active transportation modes in Canada.
 
-To create this `R` data package, we collected the GSS surveys, cleaned them and processed them to make it ready for analysis.
-An `R` data package contains code, data and documentation in a standardized collection format that can be installed by `R` users via a centralized software repository, such as CRAN (Comprehensive R Archive Network) and GitHub.
-Although the GSS surveys are publicly sourced and managed by Statistic Canada, preparing them for analysis can be time-consuming, tedious and perhaps not even possible for those who try, due to a lack of documentation or prior knowledge.
+To create this `R` data package, we collected, cleaned and processed the
+Time Use collections from the GSS surveys to make them ready for
+analysis. An `R` data package contains code, data and documentation in a
+standardized format that can be installed by `R` users via a centralized
+software repository, such as CRAN (Comprehensive R Archive Network) and
+GitHub. Although the GSS surveys are publicly sourced and managed by
+Statistic Canada, preparing them for analysis can be time-consuming,
+tedious and perhaps not even possible for those who try, due to a lack
+of documentation or prior knowledge.
 
-The aim of this paper is to walk readers through the data sets and invite others to experiment in its uses and applications.
-{ActiveCA} is freely available on GitHub for all to install and freely use in the spirit of open and reproducible research.
-Although {ActiveCA} was designed to obtain impedance functions, we admit and hope that its use can be adopted in various applications that even go beyond the range of possibilities we have imagined.
-Not only the data, but also all the code documenting the processing methodology is available for consultation and evaluation in its repository.
-This package contributes to reducing the barrier to using the information contained in GSS surveys to provide data-driven decisions in transportation analysis.
+The aim of this paper is to walk readers through the data sets and
+invite others to experiment in its uses and applications. {ActiveCA} is
+freely available on GitHub for all to install and freely use in the
+spirit of open and reproducible research. Although {ActiveCA} was
+designed to obtain impedance functions, we admit and hope that its use
+can be adopted in various applications that even go beyond the range of
+possibilities we have imagined. Not only the data, but also all the code
+documenting the processing methodology is available for consultation and
+evaluation in its repository. This package contributes to reducing the
+barrier to using the information contained in GSS surveys to provide
+data-driven decisions in transportation analysis.
 
@@
 
-In Canada, Statistics Canada produces the General Social Surveys (GSS)
-to collect data on social trends in order to track changes in the living
-conditions of Canadians and well-being over time. This survey tracks
-changes in time and it is used to understand how Canadians spend and
-manage their time and what factors contribute to their happiness and
-stress. The survey was created in 1985 as a series of independent,
-annual, cross-sectional surveys, each covering one topic in depth.
-Caregiving, families, time use, social identity, volunteering, and
-victimization are among the current themes of the GSS.
+Statistics Canada [-@statisticscanada2024] conducts GSS surveys to
+obtain data on social trends to track changes in Canadians' living
+conditions and well-being over time. This survey is used to understand
+how citizens spend and manage their time and what factors contribute to
+their happiness and stress. Created in 1985, the survey is part of a
+series of independent, annual, and cross-sectional surveys.
 
-The six survey themes listed above are repeated every five years. In
-addition to the central topic, each cycle includes new content that
-addresses emerging and policy-relevant issues, collecting detailed
-socio-demographic information, such as age, gender, education, religion,
-ethnicity, income, etc. Time-use surveys collect data on all human
-activities and can inform a wide range of policies. In particular, three
-key themes have been identified as essential for policymaking and for
-which no other data source is adequate: unpaid work and non-market
-production, well-being, and gender equality. Leisure time, work-life
-balance, health, commuting, culture, and sports are among the other
-topics covered by time-use surveys. In addition, statistics Canada has
-been conducting time-use surveys at five to seven year intervals since
-1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
-and 2015). The time use collects information on participation of
-respondents and time spent on a wide range of day-to-day activities
-using a 24-hour retrospective diary. Also, information is collected on
-the location of these activities (e.g., at home, at work, etc.) and, for
+In addition to the main topic, each GSS cycle includes new content that
+addresses emerging and policy-relevant issues. Every five to seven
+years, the Time Use Surveys [@statisticscanada2022] collect data on
+respondents' participation and time spent on a wide range of everyday
+activities using a 24-hour retrospective diary, with information on the
+location of these activities (e.g. at home, at work, etc.) and, for
 non-personal activities, the people who were present with the respondent
-at the time of the activity.
+at the time of the activity. In addition, time-use surveys also cover
+topics related to leisure time, work-life balance, health, commuting,
+culture and sports, and many others.
+
+The most recent time use survey was carried out in 2022 [@wray2024].
+However, the 2022 dataset has not been fully published and, because of
+this, our analysis focused on the surveys from 1986 to 2015 (1986, 1992,
+1998, 2005, 2010 and 2015). Time Use surveys are composed of two data
+sets, the main one and the episode file, explained in the following
+subsections.
 
@@
 on individual activity episodes. Additionally, this file categorizes
-activities into bigger groups and subcategories, enhancing the data's
+activities into bigger groups and subcategories, facilitating the data's
 analytical utility with additional metrics such as total transit time,
 duration spent with household members, and counts of activities and
-episodes. This file is aggregated and prohibits the direct association
-between specific activities with particular locations or companions.
+episodes.
 
@@
 The Episode File records detailed data for each activity episode
-reported by respondents. Each episode entry includes the start and end
-times, duration, location, and accompanying social context, informing
-when and where activities occurred and with whom. The file distinguishes
-itself by focusing on individual episodes rather than respondents, with
-the data structured around the numerous activity instances that compose
-a respondent's day. Although respondent-specific characteristics are not
-included within the Episode File, it is possible to link the Main File
-and the Episode File by using an identifiable variable present in both
-data sets.
+reported by respondents. The entry includes the start and end times,
+duration, location, and accompanying social context, informing when and
+where activities occurred and with whom. The file distinguishes itself
+by focusing on individual episodes rather than respondents, with the
+data structured around the numerous activity instances that compose a
+day of the respondent. Although respondent-specific characteristics are
+not included within the Episode File, it is possible to link the Main
+File and the Episode File by using an identifiable variable present in
+both data sets.
 
-# Average travel times
+# Descriptive statistics
 
-```{r ch03-make-table-01, echo=FALSE, cache=FALSE}
+For each year available from the Time Use surveys, we reviewed the
+episode files to identify cases with activities listed as walking or
+cycling and selected the activities immediately before and after the
+mobility episode. The {ActiveCA} package includes all
+`r nrow(gss_episodes)` episodes that recorded walking or cycling as a
+mode of transportation, with trip durations ranging from 0 to 900
+minutes. Among the analyses possible with this package, Table
+\ref{tab:table-01} presents descriptive statistics on walking and
+cycling trips between 1986 and 2015, including metrics such as the count
+of recorded trips (count), and measures of trip duration in minutes:
+maximum (max), mean, median, and minimum (min). The 1986 survey did not
+include bicycle trips.
+
+```{r table-01, echo=FALSE, cache=FALSE}
 
@@
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
-        caption = "\\label{tab:ch03-table-01}Descriptive Analysis of Active Transportation Modes: Walking and Cycling Statistics from 1986 to 2015",  
+        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
         align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
@@
 
-```{r ch03-make-table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
+Table \ref{tab:table-01} shows that the median values for walking trips
+range between 5 and 10 minutes, while cycling trips have a consistent
+median of 15 minutes since 1998. The table also highlights very high
+maximum values, particularly for walking trips, with recorded episodes
+exceeding 4 hours in all cases.
+
+Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
+statistics for the two modes of transportation, split by destination
+categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
+Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
+walking trips destined for `home` had the highest medians. However, by
+1998, the highest medians shifted to trips to `work or school`, a
+transition that also occurred for cycling trips between 1992 and 1998.
+Table \ref{tab:table-03} indicates that the median duration for trips to
+`home` and `work or school` remained at 10 minutes.
+
+```{r table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
 # Filter for selected years
@@
   )  %>% 
-  arrange(MODE, desc(Percentage_1998))
+  arrange(dest_label, MODE, desc(Percentage_1998))
 
@@
     align = c(rep("c", ncol(wide_data))),
-    caption = "\\label{tab:ch03-table-02}Comparative Trip Stats by Mode and Destination: 1986, 1992, 1998",
+    caption = "\\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998",
     col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
@@
   column_spec(10, border_right = TRUE) %>% 
-  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
+  footnote(general = c("* The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination"), threeparttable = T)
 
@@
 
-```{r ch03-make-table-03,echo=FALSE, cache=FALSE, warning=FALSE}
+```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
 # Rename certain 'dest_label' values
@@
   ) %>% 
-  arrange(MODE, desc(Percentage_2015)) %>%
+  arrange(dest_label, MODE, desc(Percentage_2015)) %>%
   group_by(MODE) %>%
@@
     align = c(rep("c", ncol(wide_data))),
-    caption = "\\label{tab:ch03-table-03}Comparative Trip Statistics by Transportation Mode and Destination: 2005, 2010, and 2015",
+    caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
     col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
@@
 
-```{r ch03-make-Heatmap-walking_2015, echo=FALSE, cache=FALSE, include=FALSE}
+{ActiveCA} also enables visual analysis with traditional exploratory
+data analysis techniques. Figures \ref{fig:figure-01} and
+\ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
+through heat maps. These maps use color gradients to represent the
+percentage of trips between various origins and destinations, with
+darker colors indicating higher percentages and lighter colors
+representing less frequent routes. To avoid overwhelming the reader, we
+omitted the heat maps for the other years analyzed.
+
+```{=html}
+<!---
+I`m not surte if this last phrase makes sense in english! "To avoid overwhelming the reader", "Para nao sobrecarregar o leitor" 
+-->
+```
+In 1992, walking trips with `home` as both the origin and destination
+made up the majority, accounting for about 30% of all walking trips.
+These trips often involved leisure activities, like short walks or dog
+walking. Following this, trips from `home` to `work or school` comprised
+18% of walking trips. Overall, `home` emerged as a crucial hub, either
+as an origin or destination, with only 5% of trips not involving `home.`
+By 2015, `home` remained a significant node, but new locations
+distributed the proportion of trips to areas not considered in 1992. In
+2015, the highest proportion of trips were from `home` to
+`work or school` (12%) and vice versa (11%). `home` to `home` accounted
+for 8% of trips, and `grocery stores` became a notable destination for
+those leaving `home` (6%), surpassing trips to `other's home` (4%).
+
+```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
 gss_heatmap <- gss_episodes_renamed %>%
@@
                                        labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
+                  geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
+                      aes(label = round(value)), 
+                      size = 4, color = "black") +
                   labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
@@
                   plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt")))
+
              )
@@
 # Combine the plots with adjusted spacing
-walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
+walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
 
@@
 # Combine the plots with adjusted spacing
-cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
+cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
 
@@
 
-```{r echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
+```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
 
@@
 
-```{r  echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
+For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
+this mode of transportation was first included as an activity, the
+majority of trips were from `home` to `work or school`, accounting for
+about 25% of cases. This pattern remained in 2015, with these trips
+representing 30% of the cases. However, a notable change occurred in
+`home` to `home` trips, which decreased significantly from 19% in 1992
+to 5% in 2015.
+
+```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
 
@@
 
-# Impedance function for Canadians Metropolitan and Census Agglomerations areas
+# Impedance functions for Canadians Metropolitan and Census Agglomerations areas
 
-```{r echo=FALSE, cache=FALSE, warning=FALSE}
-gss_impedances %>% 
-filter(dest_label == "Work or school", MODE == "Walking") %>%
-  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic","count") %>%
+Impedance functions reveals important information about the travel
+behavior of the population, by describing the relationship between the
+population at an origin and their likelihood or ability to travel to
+specific destinations to access opportunities [@soukhov2024].
+
+Impedance functions are commonly used in accessibility analysis to
+calculate the cost of travel between different locations [@hansen1959;
+@páez2012; @palacios2022]. However, these functions need to be
+calibrated in order to accurately represent travel behavior. One
+effective way of calibrating an impedance function is by using the trip
+length distribution (TLD) obtained from origin-destination data
+[@soukhov2024]. The TLD represents the probability that a certain
+proportion of trips will be made at a specific cost, such as travel
+time. In this data set, low travel times are associated with a higher
+proportion of trips, while high travel times are associated with a lower
+proportion of trips.
+
+{ActiveCA} provides calibrated impedance functions for Canadian
+Metropolitan and Census Agglomeration areas. For each combination of
+year, destination, and transportation mode, we fitted the most suitable
+impedance function based on empirical data from the GSS surveys.
+{ActiveCA} includes a total of
+`r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
+distance-decay functions. These were estimated using the `fitdistrplus`
+package [@mullerdutang2015], selecting the distribution with the lowest
+Akaike information criterion (AIC) among exponential, gamma, log-normal,
+normal, and uniform types.
+
+Table \ref{tab:table-04} shows the best impedance functions for walking
+trips to `Work or school` across survey years, where all distributions,
+except for a gamma function in 1998, are log-normal. Figure
+\ref{fig:figure-03} illustrates these fitted functions (red line)
+alongside histograms of empirical travel times (grey bars). As for the
+others functions, these examples enable to calculate gravity-based
+accessibility measures for active transportation modes across various
+destinations and temporal scales in Canadian urban areas.
+
+```{r include = FALSE}
+visualize_impedance <- function(df_functions, travel_cost){
+
+  x <- data.frame(t = seq(1, travel_cost, 1))
+
+  for (year in unique(df_functions$YEAR)) {
+
+    name_function <- df_functions[df_functions$YEAR == year, ]$f_name
+
+    subset <- df_functions %>% 
+      filter(YEAR == year)
+
+    # Nome da nova coluna com o valor de year
+    column_name <- paste0("year_", year)
+
+    # lnorm impedance 
+    if (name_function == "lnorm") {
+      x <- x %>%
+        mutate(!!sym(column_name) := dlnorm(t,
+                                             meanlog = subset$est_1[1],
+                                             sdlog = subset$est_2[1]))
+
+    # Gamma impedance 
+    } else if (name_function == "gamma") {
+
+      x <- x %>%
+        mutate(!!sym(column_name) := dgamma(t,
+                                             shape = subset$est_1[1],
+                                             rate = subset$est_2[1]))
+    }else if (name_function == "norm") {  
+      x <- x %>%
+        mutate(!!sym(column_name) := dnorm(t,
+                                             mean = subset$est_1[1],
+                                             sd = subset$est_2[1]))
+    }else if (name_function == "exp") {  
+      x <- x %>%
+        mutate(!!sym(column_name) := dexp(t,
+                                             exp = subset$est_1[1]))
+    }else if (name_function == "unif") {  
+      x <- x %>%
+        mutate(!!sym(column_name) := dunif(t,
+                                             min = subset$est_1[1],
+                                             max = subset$est_2[1]))
+    }
+  }
+
+  return(x)
+}
+```
+
+```{r , include=FALSE}
+
+max_time <- 100
+
+for(destino in unique(gss_impedances$orig_label)){
+
+  for(mode in unique(gss_impedances[gss_impedances$orig_label == destino, ]$MODE)){
+
+  selected_functions <-  gss_impedances %>% 
+  filter(dest_label == destino, MODE == mode) %>%
+  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic") %>%
   mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
   mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
+  unique()
+
+  if(nrow(selected_functions) > 0){
+    tld_theoretical <- visualize_impedance(df_functions = selected_functions, travel_cost = max_time)
+
+  # Create an empty list to store the plots
+  plots <- list()
+
+# Loop through each unique year
+
+  for(year in unique(gss_impedances$YEAR)){
+
+    # Subset empirical and theoretical data for the current year
+    empirical_subset <- gss_impedances %>% 
+      filter(dest_label == destino, MODE == mode, YEAR == year, DURATION <= max_time)
+
+    if(nrow(empirical_subset) > 1 ){
+
+      print(paste0(mode," - ", destino," - ", year))
+
+      theoretical_subset <- tld_theoretical %>%
+      dplyr::select(t, matches(as.character(year)))
+
+      # Create the ggplot
+      plots[[as.character(year)]] <- ggplot(empirical_subset, aes(x = DURATION)) +
+        geom_histogram(aes(y = ..density..), binwidth = 5, fill = "#dce1e3", color = "#adb1b3") +
+        geom_line(data = theoretical_subset, aes(x = t, y = theoretical_subset[[2]]), color = "#b52640", size = 0.80) +
+        labs(title = paste(year)) +
+        theme_minimal() +
+        scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
+        theme(axis.title.x = element_blank(),
+            axis.title.y = element_blank(), 
+            plot.title = element_text(size = 10, face="bold"))
+
+  }
+
+# Use gridExtra to arrange the plots in a 2x3 layout and add overall axis labels
+
+    if(length(plots) == 1){
+      combined_plot <- grid.arrange(
+      grobs = plots,
+      ncol = 1,
+      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
+      left = text_grob("Density", rot = 90)
+    )
+      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+       plot = combined_plot,
+       width = 9, 
+       height = 4, 
+       units = "in", 
+       dpi = 300)
+
+    } else if(length(plots) == 3){
+      combined_plot <- grid.arrange(
+      grobs = plots,
+      ncol = 3,
+      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
+      left = text_grob("Density", rot = 90)
+    )
+      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+       plot = combined_plot,
+       width = 9, 
+       height = 4, 
+       units = "in", 
+       dpi = 300)
+    }else {
+      combined_plot <- grid.arrange(
+      grobs = plots,
+      ncol = 3,
+      bottom = textGrob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode."), gp = gpar(fontsize = 16)),
+      left = textGrob("Density", rot = 90, gp = gpar(fontsize = 16))
+    )
+      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+       plot = combined_plot,
+       width = 14, 
+       height = 8, 
+       units = "in", 
+       dpi = 300)
+    }
+  }
+}
+}
+} 
+```
+
+```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
+mode <- "Walking" 
+destino <- "Work or school"
+
+gss_impedances %>% 
+filter(dest_label == destino, MODE == mode) %>%
+  dplyr::select("YEAR","f_name", "est_1","est_2","aic") %>%
+  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
+  mutate(across(c(aic), ~ round(., digits = 0)))  %>%
   unique() %>% 
@@
         booktabs = TRUE,
-        caption = "\\label{tab:ch03-table-04}Impedance functions and selection criteria for walking trips considering 'Work or school' as destination")
+        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
+        col.names = c("Year", "Impedance function*","Parameter 1*","Parameter 2*", "AIC")) %>% 
+    kable_styling(full_width = FALSE, 
+                latex_options=c("scale_down"),
+                font_size = 10) %>% 
+  row_spec(0, bold = TRUE, align = "c") %>% 
+  footnote(general = c("* `lnorm` refers to the log-normal distribution.
+For `lnorm` distributions, `Parameter 1` and `Parameter 2` refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
+For the `gamma` distribution, `Parameter 1` and `Parameter 2` refer to the rate and shape of the distribution, respectively.
+`AIC` means Akaike information criterion."), threeparttable = T)
 ```
 
-```{r}
-gss_impedances %>% 
-filter(dest_label == "Work or school", MODE == "Walking") %>%
-  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic","count") %>%
-  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
-  mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
-  unique()
+```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Percentage of walking Trips Categorized by Origin and Destination", mode), fig.align="center"}
+
+knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"))
 ```
@@
 
+This article introduces {ActiveCA}, an open data product in the form of
+an R data package, developed after the collection, cleaning, and
+processing of General Social Survey data ranging from 1986 to 2015. The
+package provides analysis-ready data on active transportation episodes,
+focusing on walking and cycling activities, with information on trip
+origins, destinations, and duration. Additionally, the package includes
+a series of impedance functions calibrated for various destinations,
+considering the different transportation modes and time periods.
+
+The value of {ActiveCA} lies in its transparency, accessibility, and
+open infrastructure, which facilitates the addition of complementary
+data sets in the future. R users can seamlessly explore GSS walking and
+cycling episodes along with calibrated impedance functions, with the
+option to suggest enhancements to the package as needed. This article
+adopts the structure proposed by Anastasia and Páez [-@soukhov2023],
+whose work provided essential guidance for the creation of this package.
+Similarly, we aim to contribute to the academic community by promoting
+transparent research practices that encourage replication and innovation
+in related fields. We believe that {ActiveCA} will serve as a basis for
+further research on GSS and for the integration of additional data by
+the authors or the wider open source community.
+
+# Declaration of Conflicting Interests
+
+The author(s) declared no potential conflicts of interest with respect
+to the research, authorship, and/or publication of this article.
+
+# Funding
+
+The author(s) disclosed receipt of the following financial support for
+the research, authorship, and/or publication of this article: This work
+was supported by the Social Sciences and Humanities Research Council of
+Canada (*More description about the funding source after the review
+process*).
+
+# ORCID
+
+Author 1
+
+Author 2
+
+Author 3
+
+# Data availability statement
+
+The {ActiceCA} R data package can be found and installed on Github
+(*link*).
+
 # References {#references .unnumbered}

### 2024-11-13 Data-Paper/Manuscript-Data-Package.tex
Data paper

@@
 
-\title{ActiveCA: an open data product to obtain impedance functions for
-active transportation modes in Canada}
+\title{ActiveCA: an open data product on active transportation episodes
+in Canada}
 
@@
 \begin{abstract}
-This article describes and visualizes the data contained in the
-\{ActiveCA\} R data package. \{ActiveCA\} contains open data products
-designed to obtain impedance functions for active transportation modes
-in Canada, retrieved from the General Social Survey collections from
-1986 to 2015. The package provides data tables on walking and cycling
-episodes, detailing the trip origins, destinations, and duration of each
-episode. The origin and destination categories covers a wide variety of
-locations, such as home, work or school, libraries, museums,
-restaurants, bars, sports centers, health clinics, place of worship, and
-others. Addittionally, the package includes details on the respondent's
-region characteristics, specifying wheter they live in a metropolitan
-area and their province of residency. \{ActiveCA\} enables users to
-calculate the average travel time for each origin-destination
-combination for each survey year, considering the two active
-transportation modes: walking and cycling. For each Census Metropolitan
-and Agglomeration Area, we estimated distance decay curves (impedance
-functions) for each origin-destination. The package will continue to
-expand with contributions from the authors and the broader community
-through requests in the future. \{ActiveCA\} is freely accessible for
-exploration and download from the associated Github repository, where
-the documentation and code involved in creating and manipulating data
-and all open data products are detailed.
+This paper describes the data set of the \{ActiveCA\} R data package.
+\{ActiveCA\} contains open data products to obtain impedance functions
+for active transportation modes in Canada, retrieved from the General
+Social Survey collections from 1986 to 2015. The package provides data
+tables on walking and cycling episodes, detailing the trip origins,
+destinations, and duration of each episode. The origin and destination
+categories covers a wide variety of locations, such as home, work or
+school, libraries, museums, restaurants, bars, sports centers, health
+clinics, place of worship, and others. Addittionally, the package
+details the respondent's region characteristics, specifying wheter they
+live in a metropolitan area and their province of residency.
+\{ActiveCA\} enables users to calculate the average travel time for each
+origin-destination combination for each survey year, considering the two
+active transportation modes: walking and cycling. For each Census
+Metropolitan and Agglomeration Area, we estimated distance-decay curves
+(impedance functions) for each destination, transportation mode and
+year. The package will continue to expand with contributions from the
+authors and the broader community through requests in the future.
+\{ActiveCA\} is freely accessible for exploration and download from the
+associated Github repository, where the documentation and code involved
+in creating and manipulating data and all open data products are
+detailed.
 \end{abstract}
@@
 
-This manuscript presents the open data product \{ActiveCA\}. Open data
+This paper presents the open data product \{ActiveCA\}. Open data
 products (ODPs) are the outcome of a transparent process that transforms
 raw data (open or not) into analysis-ready data, in which all stages of
-development follow open principles. ODPs differ from general open data
-due to their high utility, added value and open availability. The
-product presented in this document is an \texttt{R} data package that
-currently consists of processed data tables retrieved from the General
-Social Survey (GSS) collections from 1986 to 2015, to obtain impedance
+development follow open principles \citep{arribas-bel2021}. ODPs differ
+from general open data due to their high utility, added value and open
+availability. The product presented in this document is an \texttt{R}
+data package that currently consists of processed data tables retrieved
+from the Time Use collections of the General Social Surveys (GSS) from
+1986 to 2015 \citep{statisticscanada2024}, aimed to obtain impedance
 functions for active transportation modes in Canada.
 
-To create this \texttt{R} data package, we collected the GSS surveys,
-cleaned them and processed them to make it ready for analysis. An
-\texttt{R} data package contains code, data and documentation in a
-standardized collection format that can be installed by \texttt{R} users
-via a centralized software repository, such as CRAN (Comprehensive R
-Archive Network) and GitHub. Although the GSS surveys are publicly
-sourced and managed by Statistic Canada, preparing them for analysis can
-be time-consuming, tedious and perhaps not even possible for those who
-try, due to a lack of documentation or prior knowledge.
+To create this \texttt{R} data package, we collected, cleaned and
+processed the Time Use collections from the GSS surveys to make them
+ready for analysis. An \texttt{R} data package contains code, data and
+documentation in a standardized format that can be installed by
+\texttt{R} users via a centralized software repository, such as CRAN
+(Comprehensive R Archive Network) and GitHub. Although the GSS surveys
+are publicly sourced and managed by Statistic Canada, preparing them for
+analysis can be time-consuming, tedious and perhaps not even possible
+for those who try, due to a lack of documentation or prior knowledge.
 
@@
 
-In Canada, Statistics Canada produces the General Social Surveys (GSS)
-to collect data on social trends in order to track changes in the living
-conditions of Canadians and well-being over time. This survey tracks
-changes in time and it is used to understand how Canadians spend and
-manage their time and what factors contribute to their happiness and
-stress. The survey was created in 1985 as a series of independent,
-annual, cross-sectional surveys, each covering one topic in depth.
-Caregiving, families, time use, social identity, volunteering, and
-victimization are among the current themes of the GSS.
+Statistics Canada \citeyearpar{statisticscanada2024} conducts GSS
+surveys to obtain data on social trends to track changes in Canadians'
+living conditions and well-being over time. This survey is used to
+understand how citizens spend and manage their time and what factors
+contribute to their happiness and stress. Created in 1985, the survey is
+part of a series of independent, annual, and cross-sectional surveys.
 
-The six survey themes listed above are repeated every five years. In
-addition to the central topic, each cycle includes new content that
-addresses emerging and policy-relevant issues, collecting detailed
-socio-demographic information, such as age, gender, education, religion,
-ethnicity, income, etc. Time-use surveys collect data on all human
-activities and can inform a wide range of policies. In particular, three
-key themes have been identified as essential for policymaking and for
-which no other data source is adequate: unpaid work and non-market
-production, well-being, and gender equality. Leisure time, work-life
-balance, health, commuting, culture, and sports are among the other
-topics covered by time-use surveys. In addition, statistics Canada has
-been conducting time-use surveys at five to seven year intervals since
-1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
-and 2015). The time use collects information on participation of
-respondents and time spent on a wide range of day-to-day activities
-using a 24-hour retrospective diary. Also, information is collected on
-the location of these activities (e.g., at home, at work, etc.) and, for
+In addition to the main topic, each GSS cycle includes new content that
+addresses emerging and policy-relevant issues. Every five to seven
+years, the Time Use Surveys \citep{statisticscanada2022} collect data on
+respondents' participation and time spent on a wide range of everyday
+activities using a 24-hour retrospective diary, with information on the
+location of these activities (e.g.~at home, at work, etc.) and, for
 non-personal activities, the people who were present with the respondent
-at the time of the activity.
+at the time of the activity. In addition, time-use surveys also cover
+topics related to leisure time, work-life balance, health, commuting,
+culture and sports, and many others.
+
+The most recent time use survey was carried out in 2022
+\citep{wray2024}. However, the 2022 dataset has not been fully published
+and, because of this, our analysis focused on the surveys from 1986 to
+2015 (1986, 1992, 1998, 2005, 2010 and 2015). Time Use surveys are
+composed of two data sets, the main one and the episode file, explained
+in the following subsections.
 
@@
 The Episode File records detailed data for each activity episode
-reported by respondents. Each episode entry includes the start and end
-times, duration, location, and accompanying social context, informing
-when and where activities occurred and with whom. The file distinguishes
-itself by focusing on individual episodes rather than respondents, with
-the data structured around the numerous activity instances that compose
-a respondent's day. Although respondent-specific characteristics are not
-included within the Episode File, it is possible to link the Main File
-and the Episode File by using an identifiable variable present in both
-data sets.
+reported by respondents. The entry includes the start and end times,
+duration, location, and accompanying social context, informing when and
+where activities occurred and with whom. The file distinguishes itself
+by focusing on individual episodes rather than respondents, with the
+data structured around the numerous activity instances that compose a
+day of the respondent. Although respondent-specific characteristics are
+not included within the Episode File, it is possible to link the Main
+File and the Episode File by using an identifiable variable present in
+both data sets.
 
-\hypertarget{average-travel-times}{%
-\section{Average travel times}\label{average-travel-times}}
+\hypertarget{descriptive-statistics}{%
+\section{Descriptive statistics}\label{descriptive-statistics}}
+
+For each year available from the Time Use surveys, we reviewed the
+episode files to identify cases with activities listed as walking or
+cycling and selected the activities immediately before and after the
+mobility episode. The \{ActiveCA\} package includes all 21748 episodes
+that recorded walking or cycling as a mode of transportation, with trip
+durations ranging from 0 to 900 minutes. Among the analyses possible
+with this package, Table \ref{tab:table-01} presents descriptive
+statistics on walking and cycling trips between 1986 and 2015, including
+metrics such as the count of recorded trips (count), and measures of
+trip duration in minutes: maximum (max), mean, median, and minimum
+(min). The 1986 survey did not include bicycle trips.
 
@@
 \begin{longtable}[t]{>{}llcccccc}
-\caption{\label{tab:ch03-make-table-01}\label{tab:ch03-table-01}Descriptive Analysis of Active Transportation Modes: Walking and Cycling Statistics from 1986 to 2015}\\
+\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
 \toprule
@@
 
+Table \ref{tab:table-01} shows that the median values for walking trips
+range between 5 and 10 minutes, while cycling trips have a consistent
+median of 15 minutes since 1998. The table also highlights very high
+maximum values, particularly for walking trips, with recorded episodes
+exceeding 4 hours in all cases.
+
+Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
+statistics for the two modes of transportation, split by destination
+categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
+Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
+walking trips destined for \texttt{home} had the highest medians.
+However, by 1998, the highest medians shifted to trips to
+\texttt{work\ or\ school}, a transition that also occurred for cycling
+trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
+median duration for trips to \texttt{home} and \texttt{work\ or\ school}
+remained at 10 minutes.
+
 \begingroup\fontsize{6}{8}\selectfont
@@
 \item \textit{Note: } 
-\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
+\item * The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(\%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination
 \end{TableNotes}
 \begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
-\caption{\label{tab:ch03-make-table-02}\label{tab:ch03-table-02}Comparative Trip Stats by Mode and Destination: 1986, 1992, 1998}\\
+\caption{\label{tab:table-02}\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998}\\
 \toprule
@@
 Home &  & NA & NA & NA & NA & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
-Work or school & Cycling & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
-Other's home &  & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
 Home &  & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
+Other's home & Cycling & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
 Other's home & Walking & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
+Work or school &  & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
 \addlinespace
@@
 \item \textit{Note: } 
-\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
+\item * The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(\%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination
 \end{TableNotes}
 \begin{longtable}[t]{>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}}
-\caption{\label{tab:ch03-make-table-03}\label{tab:ch03-table-03}Comparative Trip Statistics by Transportation Mode and Destination: 2005, 2010, and 2015}\\
+\caption{\label{tab:table-03}\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015}\\
 \toprule
@@
 \midrule
+Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
+Cultural venues &  & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
+Cultural venues & Walking & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
+Grocery store & Cycling & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
+Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
+\addlinespace
+Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 90 & 2.0\\
+Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
 Home &  & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
-Work or school & Cycling & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
-Grocery store &  & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
+Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
+Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 30.0 & 45 & 1.2\\
+\addlinespace
+Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
 Other's home &  & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
-Restaurant &  & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
-\addlinespace
-Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 15 & 2.9\\
-Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 90 & 2.0\\
+Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
 Outdoors &  & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
-Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 30.0 & 45 & 1.2\\
-Cultural venues &  & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
+Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
 \addlinespace
 Place of worship &  & 20 & 20.0 & 20 & 0.3 & NA & NA & NA & NA & 15 & 15.0 & 15 & 0.4\\
-Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
-Work or school & Walking & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
-Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
+Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
+Restaurant &  & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
 Restaurant &  & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
+Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 15 & 2.9\\
 \addlinespace
-Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
 Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 45 & 3.3\\
-Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
-Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
-Cultural venues &  & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
-\addlinespace
-Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
-Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
-Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
+Work or school &  & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
+Work or school &  & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
 \bottomrule
@@
 \endgroup{}
+
+\{ActiveCA\} also enables visual analysis with traditional exploratory
+data analysis techniques. Figures \ref{fig:figure-01} and
+\ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
+through heat maps. These maps use color gradients to represent the
+percentage of trips between various origins and destinations, with
+darker colors indicating higher percentages and lighter colors
+representing less frequent routes. To avoid overwhelming the reader, we
+omitted the heat maps for the other years analyzed.
+
+In 1992, walking trips with \texttt{home} as both the origin and
+destination made up the majority, accounting for about 30\% of all
+walking trips. These trips often involved leisure activities, like short
+walks or dog walking. Following this, trips from \texttt{home} to
+\texttt{work\ or\ school} comprised 18\% of walking trips. Overall,
+\texttt{home} emerged as a crucial hub, either as an origin or
+destination, with only 5\% of trips not involving \texttt{home.} By
+2015, \texttt{home} remained a significant node, but new locations
+distributed the proportion of trips to areas not considered in 1992. In
+2015, the highest proportion of trips were from \texttt{home} to
+\texttt{work\ or\ school} (12\%) and vice versa (11\%). \texttt{home} to
+\texttt{home} accounted for 8\% of trips, and \texttt{grocery\ stores}
+became a notable destination for those leaving \texttt{home} (6\%),
+surpassing trips to \texttt{other\textquotesingle{}s\ home} (4\%).
 
@@
 
-\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:unnamed-chunk-1}
+\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:figure-01}
 \end{figure}
+
+For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
+this mode of transportation was first included as an activity, the
+majority of trips were from \texttt{home} to \texttt{work\ or\ school},
+accounting for about 25\% of cases. This pattern remained in 2015, with
+these trips representing 30\% of the cases. However, a notable change
+occurred in \texttt{home} to \texttt{home} trips, which decreased
+significantly from 19\% in 1992 to 5\% in 2015.
 
@@
 
-\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:unnamed-chunk-2}
+\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:figure-02}
 \end{figure}
 
-\hypertarget{impedance-function-for-canadians-metropolitan-and-census-agglomerations-areas}{%
-\section{Impedance function for Canadians Metropolitan and Census
+\hypertarget{impedance-functions-for-canadians-metropolitan-and-census-agglomerations-areas}{%
+\section{Impedance functions for Canadians Metropolitan and Census
 Agglomerations
-areas}\label{impedance-function-for-canadians-metropolitan-and-census-agglomerations-areas}}
+areas}\label{impedance-functions-for-canadians-metropolitan-and-census-agglomerations-areas}}
+
+Impedance functions reveals important information about the travel
+behavior of the population, by describing the relationship between the
+population at an origin and their likelihood or ability to travel to
+specific destinations to access opportunities \citep{soukhov2024}.
+
+Impedance functions are commonly used in accessibility analysis to
+calculate the cost of travel between different locations
+\citep{hansen1959, páez2012, palacios2022}. However, these functions
+need to be calibrated in order to accurately represent travel behavior.
+One effective way of calibrating an impedance function is by using the
+trip length distribution (TLD) obtained from origin-destination data
+\citep{soukhov2024}. The TLD represents the probability that a certain
+proportion of trips will be made at a specific cost, such as travel
+time. In this data set, low travel times are associated with a higher
+proportion of trips, while high travel times are associated with a lower
+proportion of trips.
+
+\{ActiveCA\} provides calibrated impedance functions for Canadian
+Metropolitan and Census Agglomeration areas. For each combination of
+year, destination, and transportation mode, we fitted the most suitable
+impedance function based on empirical data from the GSS surveys.
+\{ActiveCA\} includes a total of 64 distance-decay functions. These were
+estimated using the \texttt{fitdistrplus} package
+\citep{mullerdutang2015}, selecting the distribution with the lowest
+Akaike information criterion (AIC) among exponential, gamma, log-normal,
+normal, and uniform types.
+
+Table \ref{tab:table-04} shows the best impedance functions for walking
+trips to \texttt{Work\ or\ school} across survey years, where all
+distributions, except for a gamma function in 1998, are log-normal.
+Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
+alongside histograms of empirical travel times (grey bars). As for the
+others functions, these examples enable to calculate gravity-based
+accessibility measures for active transportation modes across various
+destinations and temporal scales in Canadian urban areas.
 
 \begin{table}
-
-\caption{\label{tab:unnamed-chunk-3}\label{tab:ch03-table-04}Impedance functions and selection criteria for walking trips considering 'Work or school' as destination}
 \centering
-\begin{tabular}[t]{rlrrrrrr}
+\caption{\label{tab:table-04}\label{table-04}Impedance functions and AIC for `Walking` trips considering `Work or school` as destination.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{10}{12}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{rlrrr}
 \toprule
-YEAR & f\_name & est\_1 & est\_2 & loglik & aic & bic & count\\
+\multicolumn{1}{c}{\textbf{Year}} & \multicolumn{1}{c}{\textbf{Impedance function*}} & \multicolumn{1}{c}{\textbf{Parameter 1*}} & \multicolumn{1}{c}{\textbf{Parameter 2*}} & \multicolumn{1}{c}{\textbf{AIC}}\\
 \midrule
-2015 & lnorm & 2.55 & 0.64 & -3306028 & 6612061 & 6612069 & 407\\
-2010 & lnorm & 2.24 & 0.81 & -2636159 & 5272321 & 5272329 & 367\\
-2010 & gamma & 1.99 & 0.18 & -1298182 & 2596367 & 2596373 & 127\\
-2005 & lnorm & 2.23 & 0.71 & -1507682 & 3015369 & 3015375 & 224\\
-2005 & lnorm & 2.08 & 0.84 & -2565742 & 5131489 & 5131497 & 500\\
-\addlinespace
-1998 & gamma & 1.23 & 0.09 & -1159374 & 2318752 & 2318758 & 109\\
-1992 & lnorm & 2.38 & 0.70 & -1159698 & 2319400 & 2319406 & 113\\
+2015 & lnorm & 2.55 & 0.64 & 6612061\\
+2010 & lnorm & 2.21 & 0.78 & 7917431\\
+2005 & lnorm & 2.13 & 0.79 & 8182691\\
+1998 & gamma & 1.23 & 0.09 & 2318752\\
+1992 & lnorm & 2.38 & 0.70 & 2319400\\
 \bottomrule
 \end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item * `lnorm` refers to the log-normal distribution.
+For `lnorm` distributions, `Parameter 1` and `Parameter 2` refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
+For the `gamma` distribution, `Parameter 1` and `Parameter 2` refer to the rate and shape of the distribution, respectively.
+`AIC` means Akaike information criterion.
+\end{tablenotes}
+\end{threeparttable}}
 \end{table}
 
-\begin{verbatim}
-# A tibble: 7 x 8
-   YEAR f_name est_1 est_2   loglik     aic     bic count
-  <dbl> <chr>  <dbl> <dbl>    <dbl>   <dbl>   <dbl> <int>
-1  2015 lnorm   2.55  0.64 -3306028 6612061 6612069   407
-2  2010 lnorm   2.24  0.81 -2636159 5272321 5272329   367
-3  2010 gamma   1.99  0.18 -1298182 2596367 2596373   127
-4  2005 lnorm   2.23  0.71 -1507682 3015369 3015375   224
-5  2005 lnorm   2.08  0.84 -2565742 5131489 5131497   500
-6  1998 gamma   1.23  0.09 -1159374 2318752 2318758   109
-7  1992 lnorm   2.38  0.7  -1159698 2319400 2319406   113
-\end{verbatim}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/impf_Work or school_Walking}
+
+}
+
+\caption{Percentage of walking Trips Categorized by Origin and DestinationWalking}\label{fig:figure-03}
+\end{figure}
 
@@
 \section{Concluding remarks}\label{concluding-remarks}}
+
+This article introduces \{ActiveCA\}, an open data product in the form
+of an R data package, developed after the collection, cleaning, and
+processing of General Social Survey data ranging from 1986 to 2015. The
+package provides analysis-ready data on active transportation episodes,
+focusing on walking and cycling activities, with information on trip
+origins, destinations, and duration. Additionally, the package includes
+a series of impedance functions calibrated for various destinations,
+considering the different transportation modes and time periods.
+
+The value of \{ActiveCA\} lies in its transparency, accessibility, and
+open infrastructure, which facilitates the addition of complementary
+data sets in the future. R users can seamlessly explore GSS walking and
+cycling episodes along with calibrated impedance functions, with the
+option to suggest enhancements to the package as needed. This article
+adopts the structure proposed by Anastasia and Páez
+\citeyearpar{soukhov2023}, whose work provided essential guidance for
+the creation of this package. Similarly, we aim to contribute to the
+academic community by promoting transparent research practices that
+encourage replication and innovation in related fields. We believe that
+\{ActiveCA\} will serve as a basis for further research on GSS and for
+the integration of additional data by the authors or the wider open
+source community.
+
+\hypertarget{declaration-of-conflicting-interests}{%
+\section{Declaration of Conflicting
+Interests}\label{declaration-of-conflicting-interests}}
+
+The author(s) declared no potential conflicts of interest with respect
+to the research, authorship, and/or publication of this article.
+
+\hypertarget{funding}{%
+\section{Funding}\label{funding}}
+
+The author(s) disclosed receipt of the following financial support for
+the research, authorship, and/or publication of this article: This work
+was supported by the Social Sciences and Humanities Research Council of
+Canada (\emph{More description about the funding source after the review
+process}).
+
+\hypertarget{orcid}{%
+\section{ORCID}\label{orcid}}
+
+Author 1
+
+Author 2
+
+Author 3
+
+\hypertarget{data-availability-statement}{%
+\section{Data availability
+statement}\label{data-availability-statement}}
+
+The \{ActiceCA\} R data package can be found and installed on Github
+(\emph{link}).
 

### 2024-11-13 README.Rmd
Data paper

@@
 
-<!-- badges: start -->
-<!-- badges: end -->
-
-The goal of ActiveCA is to ...
+{ActiveCA} is an open data product in the form of
+an R data package, developed after the collection, cleaning, and
+processing of [General Social Survey
+(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) data ranging from 1986 to 2015. The main objective of the {ActiveCA} package is to identify the most
+appropriate impedance functions for active transportation modes for
+various destinations and time periods in Canada.
+The package also provides analysis-ready data on active transportation episodes, focusing on walking and cycling activities, with information on trip origins, destinations, and duration.
 
@@
 You can install the development version of ActiveCA from [GitHub](https://github.com/) with:
-
 ``` r
@@
 ```
-
-## Example
-
-This is a basic example which shows you how to solve a common problem:
-
-```{r example}
-#library(ActiveCA)
-## basic example code
-```
-
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
-
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2024-11-13 Data-Paper/Manuscript-Data-Package.Rmd
Data paper
adjusting legend on manuscript

@@
 
-```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
+```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking trips categorized by origin and destination", fig.align="center"}
 
@@
 
-```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}
+```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of cycling trips categorized by origin and destination", fig.align="center"}
 
@@
 
-```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Percentage of walking Trips Categorized by Origin and Destination", mode), fig.align="center"}
+```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Empirical data and impedance functions fitted for walking trips with `work or school` as destination."), fig.align="center"}
 

### 2024-11-13 Data-Paper/Manuscript-Data-Package.tex
Data paper
adjusting legend on manuscript

@@
 
-\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:figure-01}
+\caption{Percentage of walking trips categorized by origin and destination}\label{fig:figure-01}
 \end{figure}
@@
 
-\caption{Percentage of walking Trips Categorized by Origin and Destination}\label{fig:figure-02}
+\caption{Percentage of cycling trips categorized by origin and destination}\label{fig:figure-02}
 \end{figure}
@@
 
-\caption{Percentage of walking Trips Categorized by Origin and DestinationWalking}\label{fig:figure-03}
+\caption{Empirical data and impedance functions fitted for walking trips with `work or school` as destination.Walking}\label{fig:figure-03}
 \end{figure}

### 2024-11-13 paper/Manuscript-Data-Package.Rmd
Revisions to package configuration and reproducibility

--- 2024-11-13 Data-Paper/Manuscript-Data-Package.Rmd
@@
   )  %>% 
-  arrange(dest_label, MODE, desc(Percentage_1998))
+  arrange(MODE, dest_label, desc(Percentage_1998))
 
@@
 
-# Filter for walking and cycling trips, selected years, and duration <= 90
 episodes_stats_after <- gss_episodes_renamed %>%
@@
   ) %>% 
-  arrange(dest_label, MODE, desc(Percentage_2015)) %>%
+  arrange(MODE, dest_label, desc(Percentage_2015)) %>%
    group_by(MODE) %>%

### 2024-11-13 R/ActiveCA-package.R

@@ -1,18 +1,8 @@
-#' ActiveCA: a data set to obtain impedance functions for active transportation modes in Canada
-#'
-#' The main objective of the *ActiveCA* package is to provide processed data to identify
-#' the most appropriate impedance functions for active transportation modes considering
-#' various destinations and time periods in Canada. To do achieve this objective,
-#' we processed historical data from the [General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
-#' from 1986 to 2015 in Canada. We also calculate impedance function for cycling
-#' and walking trips for the Census Agglomerations and Census Metropolitan Areas.
-#'
-#' @docType data
-#' @name ActiveCA-package
-#' @author Bruno Dos Santos, School of Earth, Environment and Society, McMaster University \email{dossanb@@mcmaster.ca}
-#' @author Mahdis Moghadasi, School of Earth, Environment and Society, McMaster University \email{paezha@@mcmaster.ca}
-#' @author Antonio Paez, School of Earth, Environment and Society, McMaster University \email{paezha@@mcmaster.ca}
-#' @references \url{https://github.com/dias-bruno/ActiveCA}
+#' @keywords internal
+"_PACKAGE"
+
+## usethis namespace: start
+## usethis namespace: end
 NULL
 
 ##############################################

### 2024-11-13 paper/Manuscript-Data-Package.Rmd
Data paper update

@@
 ---
-title: "ActiveCA: an open data product on active transportation episodes in Canada"
+title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
 author:
@@
 abstract: 
-  "This paper describes the data set of the {ActiveCA} R data package.
-{ActiveCA} contains open data products to obtain impedance functions for active transportation modes in Canada, retrieved from the General Social Survey collections from 1986 to 2015.
-The package provides data tables on walking and cycling episodes, detailing the trip origins, destinations, and duration of each episode.
-The origin and destination categories covers a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.
-Addittionally, the package details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
-{ActiveCA} enables users to calculate the average travel time for each origin-destination combination for each survey year, considering the two active transportation modes: walking and cycling.
-For each Census Metropolitan and Agglomeration Area, we estimated distance-decay curves (impedance functions) for each destination, transportation mode and year.
+  "This paper describes the open data product {ActiveCA}.
+{ActiveCA} is an `R` data package that contains analysis-ready data from the General Social Survey cicles regarding active travel in Canada.
+The package provides data set detailing active trip origins, destinations, and duration, covering a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.
+The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
+{ActiveCA} also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
 The package will continue to expand with contributions from the authors and the broader community through requests in the future.
-{ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data and all open data products are detailed."
+{ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed."
 
@@
 ```{r load-packages, include=FALSE, cache=FALSE}
+# Load packages
 library(ActiveCA)
@@
 ```{r load-data, include = FALSE}
+# Load data
 load(paste0(here(), "/data/gss_episodes.Rda"))
@@
 general open data due to their high utility, added value and open
-availability. The product presented in this document is an `R` data
-package that currently consists of processed data tables retrieved from
-the Time Use collections of the General Social Surveys (GSS) from 1986
-to 2015 [@statisticscanada2024], aimed to obtain impedance functions for
-active transportation modes in Canada.
+availability.
+
+{ActiveCA} is an `R` data package that provides analysis-ready data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) [@statisticscanada2024] regarding active travel.
+These data can be used to study walking and cycling in Canada for various purposes over time, and to calculate impedance functions for accessibility analysis.
+The package provides pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
 
@@
 software repository, such as CRAN (Comprehensive R Archive Network) and
-GitHub. Although the GSS surveys are publicly sourced and managed by
-Statistic Canada, preparing them for analysis can be time-consuming,
-tedious and perhaps not even possible for those who try, due to a lack
-of documentation or prior knowledge.
+GitHub. Although the GSS surveys are publicly sourced managed by
+Statistic Canada [@statisticscanada2024], preparing them for analysis can be time-consuming, tedious and perhaps not even possible for those who try, due to a lack of documentation or prior knowledge.
 
@@
 freely available on GitHub for all to install and freely use in the
-spirit of open and reproducible research. Although {ActiveCA} was
-designed to obtain impedance functions, we admit and hope that its use
-can be adopted in various applications that even go beyond the range of
-possibilities we have imagined. Not only the data, but also all the code
-documenting the processing methodology is available for consultation and
+spirit of open and reproducible research.
+In addition of enabling to obtain impedance functions, {ActiveCA} can be adopted in various applications that even go beyond the range of possibilities we have imagined.
+Not only the data, but also all the code documenting the processing methodology is available for consultation and
 evaluation in its repository. This package contributes to reducing the
@@
 
-The Main File compiles an large array of aggregated data, summarizing
+The main file compiles an large array of aggregated data, summarizing
 the answers to the questionnaire and derived variables that summarize
@@
 
-The Episode File records detailed data for each activity episode
+The episode file records detailed data for each activity episode
 reported by respondents. The entry includes the start and end times,
@@
 day of the respondent. Although respondent-specific characteristics are
-not included within the Episode File, it is possible to link the Main
-File and the Episode File by using an identifiable variable present in
+not included within the episode file, it is possible to link the main
+file and the episode file by using an identifiable variable present in
 both data sets.
 
-# Descriptive statistics
+# Methodology
 
-For each year available from the Time Use surveys, we reviewed the
+For each selected cycle of the GSS surveys, we reviewed the
 episode files to identify cases with activities listed as walking or
-cycling and selected the activities immediately before and after the
-mobility episode. The {ActiveCA} package includes all
-`r nrow(gss_episodes)` episodes that recorded walking or cycling as a
-mode of transportation, with trip durations ranging from 0 to 900
-minutes. Among the analyses possible with this package, Table
-\ref{tab:table-01} presents descriptive statistics on walking and
-cycling trips between 1986 and 2015, including metrics such as the count
+cycling, selecting the activities immediately before and after the
+mobility episode.
+After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, transportation mode, as well the province and urban classification of the respondent's residency.
+
+Additionally, for Census Agglomerations and Census Metropolitan Areas we pre-estimated impedance functions for cycling and walking trips.
+Impedance functions reveals important information about the travel behavior of the population, by describing the relationship between the population at an origin and their likelihood or ability to travel to specific destinations to access opportunities [@soukhov2024].
+
+Impedance functions are commonly used in accessibility analysis to calculate the cost of travel between different locations [@hansen1959; @páez2012; @palacios2022].
+However, these functions need to be calibrated in order to accurately represent travel behavior.
+One effective way of calibrating an impedance function is by using the trip length distribution (TLD) obtained from origin-destination data [@soukhov2024].
+The TLD represents the probability that a certain proportion of trips will be made at a specific cost, such as travel time.
+In this data set, low travel times are associated with a higher proportion of trips, while high travel times are associated with a lower proportion of trips.
+
+For each combination of year, destination, and transportation mode (walking or cycling), we fitted the most suitable impedance function based on empirical data from the GSS surveys.
+We used the `fitdistrplus` package [@mullerdutang2015] to estimate the functions, selecting the distribution with the lowest Akaike information criterion (AIC) among exponential, gamma, log-normal, normal, and uniform types.
+
+# Results
+
+## Descriptive statistics
+
+Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
+{ActiveCA} includes all these episodes ready for analysis.
+Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, including metrics such as the count
 of recorded trips (count), and measures of trip duration in minutes:
@@
 ```{r table-01, echo=FALSE, cache=FALSE}
-
 # Calculate the count of trips for each mode and year
-# Summarize counts by year and mode
 episodes_counts <- gss_episodes %>%
@@
 
-# Calculate min, mean, and max statistics
+# Calculate min, mean, and max statictis (duration)
 stats_data <- gss_episodes %>%
@@
 
-# Generate the table with 'Statistic' as the second column
+# Creating a latex table
 styled_ch03_table_01 <- final_data %>%
@@
 ```{r table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
-# Filter for selected years
+# Filter for selected years and calculate statistics of trip duration
 episodes_stats_before <- gss_episodes %>%
@@
 
-# Calculate percentages
+# Calculate percentages of trips by mode/year 
 total_counts <- episodes_stats_before %>% 
@@
 
+# Latex table
 styled_ch03_table_02 <- wide_data %>%
@@
 ```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
-# Rename certain 'dest_label' values
+# Renaming 'dest_label' and 'orig_label' values to reduce table space
 gss_episodes_renamed <- gss_episodes %>%
@@
 
+# Selecting surveys after 2000 and calculating trip duration statistics 
 episodes_stats_after <- gss_episodes_renamed %>%
@@
 
-# Calculate percentages
+# Calculate percentages of trips by mode/year
 total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
@@
 
+# Latex table
 styled_ch03_table_03 <- wide_data %>%
@@
 
-{ActiveCA} also enables visual analysis with traditional exploratory
-data analysis techniques. Figures \ref{fig:figure-01} and
+{ActiveCA} also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques. Figures \ref{fig:figure-01} and
 \ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
@@
 ```
+
 In 1992, walking trips with `home` as both the origin and destination
@@
 ```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
+
+# Creating heatmaps for the years of 1992 and 2015
 gss_heatmap <- gss_episodes_renamed %>%
@@
 
-      # Calculate the total weighted count for normalization
+      # Calculate the total weighted count
       total_weighted_count <- sum(trip_counts$weighted_count)
@@
 
-      # Define the order of destinations for 2015
+      # Define the order of destinations 
       destination_order <- c("Home", "Work or school", "Grocery store", 
@@
              ) 
-
   }
@@
 ```{r make-figs, include=FALSE}
+# Fixing the size of the heatmaps 
 heatmap_plot_2015_Walking <- heatmap_plot_2015_Walking +
@@
 # Save the combined plot to a file
-ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"), 
+ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"), 
        plot = walking_hm_fig,
@@
 # Save the combined plot to a file
-ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"), 
+ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"), 
        plot = cycling_hm_fig,
@@
 
-knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"))
 ```
@@
 
-knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"))
 ```
 
-# Impedance functions for Canadians Metropolitan and Census Agglomerations areas
+## Pre-estimated impedance functions for active travel in Canada
 
-Impedance functions reveals important information about the travel
-behavior of the population, by describing the relationship between the
-population at an origin and their likelihood or ability to travel to
-specific destinations to access opportunities [@soukhov2024].
-
-Impedance functions are commonly used in accessibility analysis to
-calculate the cost of travel between different locations [@hansen1959;
-@páez2012; @palacios2022]. However, these functions need to be
-calibrated in order to accurately represent travel behavior. One
-effective way of calibrating an impedance function is by using the trip
-length distribution (TLD) obtained from origin-destination data
-[@soukhov2024]. The TLD represents the probability that a certain
-proportion of trips will be made at a specific cost, such as travel
-time. In this data set, low travel times are associated with a higher
-proportion of trips, while high travel times are associated with a lower
-proportion of trips.
-
-{ActiveCA} provides calibrated impedance functions for Canadian
-Metropolitan and Census Agglomeration areas. For each combination of
-year, destination, and transportation mode, we fitted the most suitable
-impedance function based on empirical data from the GSS surveys.
 {ActiveCA} includes a total of
 `r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
-distance-decay functions. These were estimated using the `fitdistrplus`
-package [@mullerdutang2015], selecting the distribution with the lowest
-Akaike information criterion (AIC) among exponential, gamma, log-normal,
-normal, and uniform types.
-
-Table \ref{tab:table-04} shows the best impedance functions for walking
-trips to `Work or school` across survey years, where all distributions,
-except for a gamma function in 1998, are log-normal. Figure
+pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
+Table \ref{tab:table-04} shows the best impedance functions for walking trips to `Work or school` across survey years, where all distributions, except for a gamma function in 1998, are log-normal.
+Figure
 \ref{fig:figure-03} illustrates these fitted functions (red line)
 alongside histograms of empirical travel times (grey bars). As for the
-others functions, these examples enable to calculate gravity-based
-accessibility measures for active transportation modes across various
-destinations and temporal scales in Canadian urban areas.
+others pre-estimated functions, these examples allow to calculate gravity-based
+accessibility measures for active travel across various destinations and temporal scales in Canadian urban areas.
 
 ```{r include = FALSE}
+# Function to visualize impedance functions
 visualize_impedance <- function(df_functions, travel_cost){
@@
 
-    # Nome da nova coluna com o valor de year
+    # new colunm with the year information in the name
     column_name <- paste0("year_", year)
 
-    # lnorm impedance 
+    # lnorm distribution 
     if (name_function == "lnorm") {
@@
 
-    # Gamma impedance 
+    # Gamma distribution 
     } else if (name_function == "gamma") {
@@
                                              rate = subset$est_2[1]))
+
+    # Norm distribution
+
     }else if (name_function == "norm") {  
@@
                                              sd = subset$est_2[1]))
+
+    # Exponential distribution
+
     }else if (name_function == "exp") {  
@@
                                              exp = subset$est_1[1]))
+
+    # Uniform distribution
+
     }else if (name_function == "unif") {  
@@
 ```{r , include=FALSE}
-
+# Setting the maximum time to be considered to plot the impedance graphs
 max_time <- 100
 
+# Loop to generate all impedance plots at once 
 for(destino in unique(gss_impedances$orig_label)){
@@
 
-  # Create an empty list to store the plots
   plots <- list()
@@
 # Loop through each unique year
-
   for(year in unique(gss_impedances$YEAR)){
@@
 
-# Use gridExtra to arrange the plots in a 2x3 layout and add overall axis labels
+# Use gridExtra to arrange the plots in accordance with the number of survey years (some destination exits only in specific surveys)
 
+    # Case of only one year 
     if(length(plots) == 1){
@@
     )
-      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
        plot = combined_plot,
@@
 
+    # Destination present in 3 surveys  
     } else if(length(plots) == 3){
@@
        dpi = 300)
+
+  # Destination present in 5 surveys  
     }else {
@@
 ```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
+# selecting mode and destination to display impedance functions
 mode <- "Walking" 
@@
 
+# filtering the impedance functions in accordance with the mode/destination setted in the step before and creating a latex table 
 gss_impedances %>% 
@@
 
-knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"))
 ```
@@
 
-This article introduces {ActiveCA}, an open data product in the form of
-an R data package, developed after the collection, cleaning, and
-processing of General Social Survey data ranging from 1986 to 2015. The
-package provides analysis-ready data on active transportation episodes,
-focusing on walking and cycling activities, with information on trip
-origins, destinations, and duration. Additionally, the package includes
-a series of impedance functions calibrated for various destinations,
-considering the different transportation modes and time periods.
+This article presents {ActiveCA}, an open data product that provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in Canada.
+In the form of an `R` data package, {ActiveCA} was developed after collecting, cleaning, and processing the survey data, providing information on origins, destinations, and duration of active travel, as well other information.
+Additionally, the package includes a series of pre-estimated impedance functions for walking and cycling trips, considering various destinations and time periods, for Canadian Metropolitan and Census Agglomeration areas.
 

### 2024-11-13 paper/Manuscript-Data-Package.tex
Data paper update

--- 2024-11-13 Data-Paper/Manuscript-Data-Package.tex
@@
 
-\title{ActiveCA: an open data product on active transportation episodes
-in Canada}
+\title{ActiveCA: Time Use Data from the General Social Survey of Canada
+to Study Active Travel}
 
@@
 \begin{abstract}
-This paper describes the data set of the \{ActiveCA\} R data package.
-\{ActiveCA\} contains open data products to obtain impedance functions
-for active transportation modes in Canada, retrieved from the General
-Social Survey collections from 1986 to 2015. The package provides data
-tables on walking and cycling episodes, detailing the trip origins,
-destinations, and duration of each episode. The origin and destination
-categories covers a wide variety of locations, such as home, work or
-school, libraries, museums, restaurants, bars, sports centers, health
-clinics, place of worship, and others. Addittionally, the package
-details the respondent's region characteristics, specifying wheter they
-live in a metropolitan area and their province of residency.
-\{ActiveCA\} enables users to calculate the average travel time for each
-origin-destination combination for each survey year, considering the two
-active transportation modes: walking and cycling. For each Census
-Metropolitan and Agglomeration Area, we estimated distance-decay curves
-(impedance functions) for each destination, transportation mode and
-year. The package will continue to expand with contributions from the
-authors and the broader community through requests in the future.
-\{ActiveCA\} is freely accessible for exploration and download from the
-associated Github repository, where the documentation and code involved
-in creating and manipulating data and all open data products are
-detailed.
+This paper describes the open data product \{ActiveCA\}. \{ActiveCA\} is
+an \texttt{R} data package that contains analysis-ready data from the
+General Social Survey cicles regarding active travel in Canada. The
+package provides data set detailing active trip origins, destinations,
+and duration, covering a wide variety of locations, such as home, work
+or school, libraries, museums, restaurants, bars, sports centers, health
+clinics, place of worship, and others. The package also details the
+respondent's region characteristics, specifying wheter they live in a
+metropolitan area and their province of residency. \{ActiveCA\} also
+provides pre-estimated impedance functions for active travel for
+Canadian Metropolitan Area and Census Agglomerations, which can be used
+in accessibility analysis to calculate the cost of travel between
+different locations. The package will continue to expand with
+contributions from the authors and the broader community through
+requests in the future. \{ActiveCA\} is freely accessible for
+exploration and download from the associated Github repository, where
+the documentation and code involved in creating and manipulating data
+are detailed.
 \end{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 from general open data due to their high utility, added value and open
-availability. The product presented in this document is an \texttt{R}
-data package that currently consists of processed data tables retrieved
-from the Time Use collections of the General Social Surveys (GSS) from
-1986 to 2015 \citep{statisticscanada2024}, aimed to obtain impedance
-functions for active transportation modes in Canada.
+availability.
+
+\{ActiveCA\} is an \texttt{R} data package that provides analysis-ready
+data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992),
+12 (1998), 19 (2005), 24 (2010), and 29 (2015)
+\citep{statisticscanada2024} regarding active travel. These data can be
+used to study walking and cycling in Canada for various purposes over
+time, and to calculate impedance functions for accessibility analysis.
+The package provides pre-estimated impedance functions for cycling and
+walking trips for Census Agglomerations and Census Metropolitan Areas.
 
@@
 (Comprehensive R Archive Network) and GitHub. Although the GSS surveys
-are publicly sourced and managed by Statistic Canada, preparing them for
-analysis can be time-consuming, tedious and perhaps not even possible
-for those who try, due to a lack of documentation or prior knowledge.
+are publicly sourced managed by Statistic Canada
+\citep{statisticscanada2024}, preparing them for analysis can be
+time-consuming, tedious and perhaps not even possible for those who try,
+due to a lack of documentation or prior knowledge.
 
@@
 is freely available on GitHub for all to install and freely use in the
-spirit of open and reproducible research. Although \{ActiveCA\} was
-designed to obtain impedance functions, we admit and hope that its use
-can be adopted in various applications that even go beyond the range of
-possibilities we have imagined. Not only the data, but also all the code
-documenting the processing methodology is available for consultation and
-evaluation in its repository. This package contributes to reducing the
-barrier to using the information contained in GSS surveys to provide
-data-driven decisions in transportation analysis.
+spirit of open and reproducible research. In addition of enabling to
+obtain impedance functions, \{ActiveCA\} can be adopted in various
+applications that even go beyond the range of possibilities we have
+imagined. Not only the data, but also all the code documenting the
+processing methodology is available for consultation and evaluation in
+its repository. This package contributes to reducing the barrier to
+using the information contained in GSS surveys to provide data-driven
+decisions in transportation analysis.
 
-\hypertarget{general-social-survey-gss-collection}{%
 \section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}}
+collection}\label{general-social-survey-gss-collection}
 
@@
 
-\hypertarget{the-main-file}{%
-\subsection{The Main File}\label{the-main-file}}
+\subsection{The Main File}\label{the-main-file}
 
-The Main File compiles an large array of aggregated data, summarizing
+The main file compiles an large array of aggregated data, summarizing
 the answers to the questionnaire and derived variables that summarize
@@
 
-\hypertarget{the-episode-file}{%
-\subsection{The Episode File}\label{the-episode-file}}
+\subsection{The Episode File}\label{the-episode-file}
 
-The Episode File records detailed data for each activity episode
+The episode file records detailed data for each activity episode
 reported by respondents. The entry includes the start and end times,
@@
 day of the respondent. Although respondent-specific characteristics are
-not included within the Episode File, it is possible to link the Main
-File and the Episode File by using an identifiable variable present in
+not included within the episode file, it is possible to link the main
+file and the episode file by using an identifiable variable present in
 both data sets.
 
-\hypertarget{descriptive-statistics}{%
-\section{Descriptive statistics}\label{descriptive-statistics}}
+\section{Methodology}\label{methodology}
 
-For each year available from the Time Use surveys, we reviewed the
-episode files to identify cases with activities listed as walking or
-cycling and selected the activities immediately before and after the
-mobility episode. The \{ActiveCA\} package includes all 21748 episodes
-that recorded walking or cycling as a mode of transportation, with trip
-durations ranging from 0 to 900 minutes. Among the analyses possible
-with this package, Table \ref{tab:table-01} presents descriptive
-statistics on walking and cycling trips between 1986 and 2015, including
-metrics such as the count of recorded trips (count), and measures of
-trip duration in minutes: maximum (max), mean, median, and minimum
-(min). The 1986 survey did not include bicycle trips.
+For each selected cycle of the GSS surveys, we reviewed the episode
+files to identify cases with activities listed as walking or cycling,
+selecting the activities immediately before and after the mobility
+episode. After that, we labeled the code variables with their
+appropriate descriptions, identifying each origin and destination,
+transportation mode, as well the province and urban classification of
+the respondent's residency.
+
+Additionally, for Census Agglomerations and Census Metropolitan Areas we
+pre-estimated impedance functions for cycling and walking trips.
+Impedance functions reveals important information about the travel
+behavior of the population, by describing the relationship between the
+population at an origin and their likelihood or ability to travel to
+specific destinations to access opportunities \citep{soukhov2024}.
+
+Impedance functions are commonly used in accessibility analysis to
+calculate the cost of travel between different locations
+\citep{hansen1959, páez2012, palacios2022}. However, these functions
+need to be calibrated in order to accurately represent travel behavior.
+One effective way of calibrating an impedance function is by using the
+trip length distribution (TLD) obtained from origin-destination data
+\citep{soukhov2024}. The TLD represents the probability that a certain
+proportion of trips will be made at a specific cost, such as travel
+time. In this data set, low travel times are associated with a higher
+proportion of trips, while high travel times are associated with a lower
+proportion of trips.
+
+For each combination of year, destination, and transportation mode
+(walking or cycling), we fitted the most suitable impedance function
+based on empirical data from the GSS surveys. We used the
+\texttt{fitdistrplus} package \citep{mullerdutang2015} to estimate the
+functions, selecting the distribution with the lowest Akaike information
+criterion (AIC) among exponential, gamma, log-normal, normal, and
+uniform types.
+
+\section{Results}\label{results}
+
+\subsection{Descriptive statistics}\label{descriptive-statistics}
+
+Considering GSS Cycles analyzed, we identified 21748 episodes that
+recorded active travel episodes, with trip duration ranging from 0 to
+900 minutes, to twelve different destinations. \{ActiveCA\} includes all
+these episodes ready for analysis. Table \ref{tab:table-01} presents
+descriptive statistics on walking and cycling trips between 1986 and
+2015, including metrics such as the count of recorded trips (count), and
+measures of trip duration in minutes: maximum (max), mean, median, and
+minimum (min). The 1986 survey did not include bicycle trips.
 
@@
 Home &  & NA & NA & NA & NA & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
+Other's home & Cycling & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
+Work or school &  & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
 Home &  & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
-Other's home & Cycling & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
 Other's home & Walking & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
-Work or school &  & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
 \addlinespace
@@
 \midrule
-Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
 Cultural venues &  & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
-Cultural venues & Walking & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
 Grocery store & Cycling & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
-Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
-\addlinespace
 Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 90 & 2.0\\
-Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
 Home &  & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
-Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
 Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 30.0 & 45 & 1.2\\
 \addlinespace
-Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
 Other's home &  & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
-Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
 Outdoors &  & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
-Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
-\addlinespace
 Place of worship &  & 20 & 20.0 & 20 & 0.3 & NA & NA & NA & NA & 15 & 15.0 & 15 & 0.4\\
-Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
 Restaurant &  & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
-Restaurant &  & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
 Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 15 & 2.9\\
 \addlinespace
+Work or school &  & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
+Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
+Cultural venues & Walking & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
+Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
+Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
+\addlinespace
+Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
+Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
+Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
+Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
+Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
+\addlinespace
+Restaurant &  & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
 Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 45 & 3.3\\
-Work or school &  & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
 Work or school &  & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
@@
 
-\{ActiveCA\} also enables visual analysis with traditional exploratory
-data analysis techniques. Figures \ref{fig:figure-01} and
-\ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
-through heat maps. These maps use color gradients to represent the
-percentage of trips between various origins and destinations, with
-darker colors indicating higher percentages and lighter colors
-representing less frequent routes. To avoid overwhelming the reader, we
-omitted the heat maps for the other years analyzed.
+\{ActiveCA\} also enables visual analysis of active travel in Canada
+using traditional exploratory data analysis techniques. Figures
+\ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling
+trips from 1992 and 2015 through heat maps. These maps use color
+gradients to represent the percentage of trips between various origins
+and destinations, with darker colors indicating higher percentages and
+lighter colors representing less frequent routes. To avoid overwhelming
+the reader, we omitted the heat maps for the other years analyzed.
 
@@
 
-\hypertarget{impedance-functions-for-canadians-metropolitan-and-census-agglomerations-areas}{%
-\section{Impedance functions for Canadians Metropolitan and Census
-Agglomerations
-areas}\label{impedance-functions-for-canadians-metropolitan-and-census-agglomerations-areas}}
+\subsection{Pre-estimated impedance functions for active travel in
+Canada}\label{pre-estimated-impedance-functions-for-active-travel-in-canada}
 
-Impedance functions reveals important information about the travel
-behavior of the population, by describing the relationship between the
-population at an origin and their likelihood or ability to travel to
-specific destinations to access opportunities \citep{soukhov2024}.
-
-Impedance functions are commonly used in accessibility analysis to
-calculate the cost of travel between different locations
-\citep{hansen1959, páez2012, palacios2022}. However, these functions
-need to be calibrated in order to accurately represent travel behavior.
-One effective way of calibrating an impedance function is by using the
-trip length distribution (TLD) obtained from origin-destination data
-\citep{soukhov2024}. The TLD represents the probability that a certain
-proportion of trips will be made at a specific cost, such as travel
-time. In this data set, low travel times are associated with a higher
-proportion of trips, while high travel times are associated with a lower
-proportion of trips.
-
-\{ActiveCA\} provides calibrated impedance functions for Canadian
-Metropolitan and Census Agglomeration areas. For each combination of
-year, destination, and transportation mode, we fitted the most suitable
-impedance function based on empirical data from the GSS surveys.
-\{ActiveCA\} includes a total of 64 distance-decay functions. These were
-estimated using the \texttt{fitdistrplus} package
-\citep{mullerdutang2015}, selecting the distribution with the lowest
-Akaike information criterion (AIC) among exponential, gamma, log-normal,
-normal, and uniform types.
-
-Table \ref{tab:table-04} shows the best impedance functions for walking
-trips to \texttt{Work\ or\ school} across survey years, where all
-distributions, except for a gamma function in 1998, are log-normal.
-Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
-alongside histograms of empirical travel times (grey bars). As for the
-others functions, these examples enable to calculate gravity-based
-accessibility measures for active transportation modes across various
-destinations and temporal scales in Canadian urban areas.
+\{ActiveCA\} includes a total of 64 pre-estimated impedance functions
+for cycling and walking trips for Census Agglomerations and Census
+Metropolitan Areas. Table \ref{tab:table-04} shows the best impedance
+functions for walking trips to \texttt{Work\ or\ school} across survey
+years, where all distributions, except for a gamma function in 1998, are
+log-normal. Figure \ref{fig:figure-03} illustrates these fitted
+functions (red line) alongside histograms of empirical travel times
+(grey bars). As for the others pre-estimated functions, these examples
+allow to calculate gravity-based accessibility measures for active
+travel across various destinations and temporal scales in Canadian urban
+areas.
 
@@
 
-\caption{Empirical data and impedance functions fitted for walking trips with `work or school` as destination.Walking}\label{fig:figure-03}
+\caption{Empirical data and impedance functions fitted for walking trips with `work or school` as destination.}\label{fig:figure-03}
 \end{figure}
 
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
+\section{Concluding remarks}\label{concluding-remarks}
 
-This article introduces \{ActiveCA\}, an open data product in the form
-of an R data package, developed after the collection, cleaning, and
-processing of General Social Survey data ranging from 1986 to 2015. The
-package provides analysis-ready data on active transportation episodes,
-focusing on walking and cycling activities, with information on trip
-origins, destinations, and duration. Additionally, the package includes
-a series of impedance functions calibrated for various destinations,
-considering the different transportation modes and time periods.
+This article presents \{ActiveCA\}, an open data product that provides
+analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
+(2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in
+Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
+developed after collecting, cleaning, and processing the survey data,
+providing information on origins, destinations, and duration of active
+travel, as well other information. Additionally, the package includes a
+series of pre-estimated impedance functions for walking and cycling
+trips, considering various destinations and time periods, for Canadian
+Metropolitan and Census Agglomeration areas.
 
@@
 
-\hypertarget{declaration-of-conflicting-interests}{%
 \section{Declaration of Conflicting
-Interests}\label{declaration-of-conflicting-interests}}
+Interests}\label{declaration-of-conflicting-interests}
 
@@
 
-\hypertarget{funding}{%
-\section{Funding}\label{funding}}
+\section{Funding}\label{funding}
 
@@
 
-\hypertarget{orcid}{%
-\section{ORCID}\label{orcid}}
+\section{ORCID}\label{orcid}
 
@@
 
-\hypertarget{data-availability-statement}{%
-\section{Data availability
-statement}\label{data-availability-statement}}
+\section{Data availability statement}\label{data-availability-statement}
 

### 2024-11-17 README.Rmd
Improved the documentation and fixed an issue with the Cycle 2 (1986) episode weights, which did not correctly use the number of decimals indicated in the user's guide

--- 2024-09-22 README.Rmd
@@
 
-The goal of ActiveCA is to ...
+{ActiveCA} is an open data product in the form of an `R` data package with information about active travel in Canada.
+The data product is based on Public Use Microdata Files of Statistics Canada's [General Social Survey (GSS) program](https://www150.statcan.gc.ca/n1/pub/45-25-0001/index-eng.htm) with a focus on the [Time Use Survey](https://www150.statcan.gc.ca/n1/pub/45-25-0001/index-eng.htm#a7) cycles.
+
+To reduce the barriers to the use of the Public Use Microdata Files, this data product has extracted all walking and cycling episodes and their corresponding episode weights.
+Origins and destinations are categorized, and it is possible to investigate active travel (number of episodes and their duration) for broad categories of purposes (e.g., travel to work or school, travel to other's home).
+
+The data covers cycles Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015), spanning a period of almost thirty years.
 

### 2024-11-17 data-raw/01-preprocessing-source-files.Rmd
Revisions to package configuration and reproducibility
Improved the documentation and fixed an issue with the Cycle 2 (1986) episode weights, which did not correctly use the number of decimals indicated in the user's guide

@@
 
-For the purpose of our analyses, the following definitions are
-important:
+For the purpose of our analyses, the following definitions are important:
 
@@
 
+Label the activity ("7" is "Transport to or from activity") and drop the LOCATION because it is already coded as MODE:    
+```{r}
+walking_2015 <- walking_2015 |>
+  mutate(act_label = "Transport to or from activity") |>
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
@@
 ```{r}
-usethis::use_data(walking_2015, overwrite = TRUE)
+usethis::use_data(walking_2015, 
+                  overwrite = TRUE)
 ```
@@
 ```
+
 Now, let's select only the cases of respondents who live in a CMA or CA:
@@
 ```
+
 Now, let's evaluate the count of episodes between each combination of
@@
 
+Label the activity ("7" is "Transport to or from activity") and drop the LOCATION because it is already coded as MODE:    
+```{r}
+cycling_2015 <- cycling_2015 |>
+  mutate(act_label = "Transport to or from activity") |>
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(cycling_2015, overwrite = TRUE)
+```{r}
+usethis::use_data(cycling_2015, 
+                  overwrite = TRUE)
 ```
@@
 ```
+
 ## GSS 2010
@@
 
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+walking_2010 <- walking_2010 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
@@
 ```{r}
-usethis::use_data(walking_2010, overwrite = TRUE)
+usethis::use_data(walking_2010, 
+                  overwrite = TRUE)
 ```
@@
 
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+cycling_2010 <- cycling_2010 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(cycling_2010, overwrite = TRUE)
-```
+```{r}
+usethis::use_data(cycling_2010, 
+                  overwrite = TRUE)
+```
+
 As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data
@@
 ```
+
 ## GSS 2005
@@
 
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+walking_2005 <- walking_2005 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(walking_2005, overwrite = TRUE)
+```{r}
+usethis::use_data(walking_2005,
+                  overwrite = TRUE)
 ```
@@
 
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+cycling_2005 <- cycling_2005 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(cycling_2005, overwrite = TRUE)
+```{r}
+usethis::use_data(cycling_2005, 
+                  overwrite = TRUE)
 ```
@@
 ```
+
 ## GSS 1998
@@
 The person was at home doing food cleanup (`ACTCODE` = 110), then
-went for walking that lasted 30 minutes, came back home and spent 15 minutes relaxing/resting. Here have here another cas of a recreational/leisure trip!
+went for walking that lasted 30 minutes, came back home and spent 15 minutes relaxing/resting. Here have here another case of a recreational/leisure trip!
+
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+walking_1998 <- walking_1998 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
 
@@
 ```{r}
-usethis::use_data(walking_1998, overwrite = TRUE)
+usethis::use_data(walking_1998, 
+                  overwrite = TRUE)
 ```
@@
 
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+cycling_1998 <- cycling_1998 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
 Save data in R format for later use:
@@
 ```{r}
-usethis::use_data(cycling_1998, overwrite = TRUE)
+usethis::use_data(cycling_1998, 
+                  overwrite = TRUE)
 ```
@@
 ```
+
 ## GSS 1992
@@
 
-Save data in R format for later use:
-
-```{r}
-usethis::use_data(walking_1992, overwrite = TRUE)
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r}
+walking_1992 <- walking_1992 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
+Save data in `R` format for later use:
+```{r}
+usethis::use_data(walking_1992, 
+                  overwrite = TRUE)
 ```
@@
 
-The person was at home Talking or having conversation on the phone (`ACTCODE` = 950), then went for cycling that lasted 60 minutes, came back home and spent 15 minutes washing and dressing (`ACTCODE` = 400).
+The person was at home Talking or having conversation on the phone (`ACTCODE` = 950), then went for cycling (`ACTCODE` = 822) that lasted 60 minutes, came back home and spent 15 minutes washing and dressing (`ACTCODE` = 400).
+
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r label-activity-codes-1992-cycling}
+cycling_1992 <- cycling_1992 |>
+  mutate(act_label = case_when(ACTCODE == 2 ~ "Refused information",
+                               ACTCODE == 12 ~ "Work for pay at other job(s)", # p. 156
+                               ACTCODE == 22 ~ "Looking for work", # p. 156
+                               ACTCODE == 23 ~ "Unpaid work in a family business or farm", # p. 156
+                               ACTCODE == 30 ~ "Travel during work", # p. 156
+                               ACTCODE == 60 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 90 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 101 ~ "Meal preparation", # p. 164
+                               ACTCODE == 110 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 130 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164 
+                               ACTCODE == 171 ~ "Gardening/grounds maintenance", # p. 
+                               ACTCODE == 172 ~ "Pet care (walking, grooming, feeding)", # p. 164
+                               ACTCODE == 190 ~ "Travel to/from unpaid household work", # p. 164
+                               ACTCODE == 200 ~ "Baby care/child care (infant to 4 years old)", # p. 174
+                               ACTCODE == 210 ~ "Child care", # p. 177 
+                               ACTCODE == 240 ~ "Play with children", # p. 177
+                               ACTCODE == 291 ~ "Travel to/from personal care activities for household children", # p.177
+                               ACTCODE == 292 ~ "Travel to/from personal care activities for household adults",  # p.177
+                               ACTCODE == 302 ~ "Everyday shopping: everyday goods and products (clothing, gas, etc.)", # p. 188
+                               ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
+                               ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
+                               ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
+                               ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
+                               ACTCODE == 792 ~ "Travel to/from socializing at private residences", # p. 234
+                               ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
+                               ACTCODE == 821 ~ "Walking, hiking, jogging, running", # p. 821
+                               ACTCODE == 822 ~ "Bicycling", # p. 244
+                               ACTCODE == 831 ~ "Hobbies done mainly for pleasure (painting, sketching, photography)", # p. 244
+                               ACTCODE == 871 ~ "Pleasure drives (as the driver)", # p. 245
+                               ACTCODE == 872 ~ "Pleasure drives (as a passenger in the car)", # p. 245
+                               ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
+                               ACTCODE == 892 ~ "Travel to/from coaching activities", # p. 245
+                               ACTCODE == 893 ~ "Travel to/from hobbies or for the sale of crafts", # p. 245
+                               ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
+                               ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
 
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(cycling_1992, overwrite = TRUE)
-```
+```{r}
+usethis::use_data(cycling_1992, 
+                  overwrite = TRUE)
+```
+
 As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data
-
 ```{r}
@@
 `LOCATION` to align with the names of the variables in 2015.
+
+**IMPORTANT:** See p.
+A96 of the User's Guide for Cycle 2.
+The `WGHT_EPI` values, which give the number of people a record represents, (in this case the `FWGT_MS`) are given as a 9 digit integer with 4 decimals, like so: XXXXX.XXXX.
+We cannot just copy `FWGT_MS` to `WGHT_EPI` below, we must divide first by 10,000:
 
@@
     ACTCODE = ACT_CODE,
-    LOCATION = PLACE)
+    LOCATION = PLACE) |>
+  mutate(WGHT_EPI = WGHT_EPI/10000)
 ```
@@
   slice(unique(c(reduce(map(which(PLACE == 5), ~ c(.x - 1, .x, .x + 1)), c)))) |>
-  dplyr::select(SEQNUM,PLACE,ACT_CODE,DURATION)
-```
-
-The person was at home listening radio for 30 minutes (`ARTCODE` = 90), then went for walking that lasted 15 minutes, came back home and spent 90 minutes
-resting/relaxing (`ARTCODE` = 47).
+  dplyr::select(SEQNUM, PLACE, ACT_CODE, DURATION)
+```
+
+The person was at home listening radio for 30 minutes (`ARTCODE` = 90), then went for a walk for gardening/pet care (`ACT_CODE` = 17) that lasted 15 minutes, came back home and spent 90 minutes resting/relaxing (`ARTCODE` = 47).
+
+```{r include=FALSE}
+walking_1986 |>
+  filter(orig_label == "Home" & dest_label == "Other's home")
+```
+
+See SEQNUM == 26 and check the walking episode:
+
+```{r}
+gss_e_1986 |>
+  filter(SEQNUM == 68) |>
+  slice(unique(c(reduce(map(which(PLACE == 5), ~ c(.x - 1, .x, .x + 1)), c)))) |>
+  dplyr::select(SEQNUM, PLACE, ACT_CODE, DURATION)
+```
+
+Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+```{r label-activity-codes-1986-walking}
+walking_1986 <- walking_1986 |>
+  mutate(act_label = case_when(ACTCODE == 1 ~ "Work for Pay",
+                               ACTCODE == 2 ~ "Overtime/looking for work", # p. 156
+                               ACTCODE == 3 ~ "Travel during work", # p. 156
+                               ACTCODE == 5 ~ "Meals/snacks at work", # p. 156
+                               ACTCODE == 6 ~ "Idle time before/after work", # p. 156
+                               ACTCODE == 7 ~ "Coffee/other breaks at work", # p. 156
+                               ACTCODE == 8 ~ "Other work activity", # p. 156
+                               ACTCODE == 9 ~ "Travel to/from paid work", # p. 156
+                               ACTCODE == 10 ~ "Meal preparation", # p. 164
+                               ACTCODE == 11 ~ "Food (or meal) cleanup", # p. 164
+                               ACTCODE == 12 ~ "Indoor cleaning", # p. 164
+                               ACTCODE == 13 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164
+                               ACTCODE == 17 ~ "Gardening, pet care", # p. 164 
+                               ACTCODE == 18 ~ "Other housework", # p. 164
+                               ACTCODE == 19 ~ "Travel to/from unpaid household work", # p.164
+                               ACTCODE == 20 ~ "Baby care/child care (infant to 4 years old)",  # p.177
+                               ACTCODE == 21 ~ "Child care", # p. 177
+                               ACTCODE == 22 ~ "Helping/teaching/reprimanding", # p. 177
+                               ACTCODE == 23 ~ "Reading/talking/conversation with child", # p. 177
+                               ACTCODE == 28 ~ "Other child care (unpaid babysitting)", # p. 177
+                               ACTCODE == 29 ~ "Travel to/from personal care activities for household members", # p. 177 
+                               ACTCODE == 30 ~ "Everyday Shopping", # p. 188
+                               ACTCODE == 33 ~ "Government and financial services", # p. 188
+                               ACTCODE == 34 ~ "Adult medical and dental care (outside home)", # p. 188
+                               ACTCODE == 36 ~ "Repair services (cleaning, auto, appliance)", # p. 
+                               ACTCODE == 38 ~ "Other shopping and services", # p. 188
+                               ACTCODE == 39 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 40 ~ "Washing, dressing", # p. 202 
+                               ACTCODE == 41 ~ "Adult medical care (at home)", # p. 202
+                               ACTCODE == 43 ~ "Meals at home/snacks/coffee", # p. 202
+                               ACTCODE == 44 ~ "Meals at restaurant", # p. 202 
+                               ACTCODE == 45 ~ "Night sleep/essential sleep", # p. 202
+                               ACTCODE == 47 ~ "Relaxing, thinking, resting, smoking", # p. 202 
+                               ACTCODE == 48 ~ "Other personal care or private activities (e.g. washroom activities, sex)", # p. 202
+                               ACTCODE == 49 ~ "Travel to/from personal activities", # p. 202
+                               ACTCODE == 50 ~ "Full-time classes", # p. 208
+                               ACTCODE == 51 ~ "Other classes – part-time", # p. 208 
+                               ACTCODE == 53 ~ "Homework (course, career, self-development)", # p. 208
+                               ACTCODE == 54 ~ "Meals/snacks/coffee at school", # p. 208
+                               ACTCODE == 55 ~ "Breaks/waiting for class", # p. 208
+                               ACTCODE == 56 ~ "Leisure and special interest classes", # p. 208
+                               ACTCODE == 58 ~ "Other study", # p. 208 
+                               ACTCODE == 59 ~ "Travel to/from school education activities", # p. 208
+                               ACTCODE == 62 ~ "Child, youth, family organizations (e.g. scout leader, school volunteer)", # p. 215
+                               ACTCODE == 63 ~ "Religious meetings, organizations (e.g. choir practice, church socials)", # p. 215
+                               ACTCODE == 64 ~ "Religious services/prayer/Bible readings", # p. 215
+                               ACTCODE == 66 ~ "Volunteer work, helping", # p. 215
+                               ACTCODE == 69 ~ "Travel to/from organizational activities", # p. 216 
+                               ACTCODE == 72 ~ "Movies/films at a theatre/cinema, art films, drive-in movies", # p. 234
+                               ACTCODE == 75 ~ "Visits, entertaining friends/relatives", # p. 234
+                               ACTCODE == 76 ~ "Socializing at bars, clubs", # p. 234
+                               ACTCODE == 78 ~ "Other social gatherings", # p. 234
+                               ACTCODE == 79 ~ "Travel to/from entertainment activities", # p. 234
+                               ACTCODE == 80 ~ "Sports, physical exercise, coaching", # p. 243
+                               ACTCODE == 81 ~ "Hunt, fish, camp", # p. 244
+                               ACTCODE == 82 ~ "Walk, hike", # p. 244
+                               ACTCODE == 85 ~ "Music, theatre, dance", # p. 244
+                               ACTCODE == 86 ~ "Games, cards, arcade", # p. 244
+                               ACTCODE == 87 ~ "Pleasure drives, sightseeing", # p. 245
+                               ACTCODE == 88 ~ "Other leisure activity", # p. 245
+                               ACTCODE == 89 ~ "Travel to/from sports, hobbies", # p. 245
+                               ACTCODE == 91 ~ "Television, rented movies", # p. 266 
+                               ACTCODE == 92 ~ "Listening to CD’s, cassette tapes or records", # p. 266
+                               ACTCODE == 93 ~ "Reading books, magazines", # p. 266
+                               ACTCODE == 95 ~ "Talking, conversation with household member only (face-to-face)", # p. 266
+                               ACTCODE == 96 ~ "Letters and mail", # p. 266
+                               ACTCODE == 97 ~ "Refused information", # p. 303
+                               ACTCODE == 98 ~ "Other media or communication", # p. 266
+                               ACTCODE == 99 ~ "Travel to/from media or communication activities", # p. 266
+                               )) |> 
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
 
 Save data in R format for later use:
-
-```{r}
-usethis::use_data(walking_1986, overwrite = TRUE)
+```{r}
+usethis::use_data(walking_1986, 
+                  overwrite = TRUE)
 ```
@@
 ```{r}
-usethis::use_data(gss_episodes, overwrite = TRUE)
-```
+usethis::use_data(gss_episodes, 
+                  overwrite = TRUE)
+```

### 2024-11-17 data-raw/02-calculating-impedances.Rmd
Improved the documentation and fixed an issue with the Cycle 2 (1986) episode weights, which did not correctly use the number of decimals indicated in the user's guide

@@
 ```{r}
-load(file = paste0(here(), "/data/walking_1986.Rda"))
+load(file = paste0(here(), "/data/walking_1986.rda"))
 ```
@@
            x = tld_empirical[tld_empirical$destination == destino, ]$DURATION,
-           weights = tld_empirical[tld_empirical$destination == destino, ]$WGHT_EPI))
+           weights = round(tld_empirical[tld_empirical$destination == destino, ]$WGHT_EPI)))
 }
@@
 ```{r}
-load(file = paste0(here(), "/data/gss_episodes.Rda"))
+load(file = paste0(here(), "/data/gss_episodes.rda"))
 ```
@@
 ```{r}
-gss_impedances <- gss_episodes[,c("PUMFID","DURATION","LOCATION",
+gss_impedances <- gss_episodes[,c("PUMFID","DURATION",
                                   "origin","destination","Province","Pop_centre","orig_label",

### 2024-11-17 paper/Manuscript-Data-Package.Rmd
Improved the documentation and fixed an issue with the Cycle 2 (1986) episode weights, which did not correctly use the number of decimals indicated in the user's guide

@@
 abstract: 
-  "This paper describes the open data product {ActiveCA}.
-{ActiveCA} is an `R` data package that contains analysis-ready data from the General Social Survey cicles regarding active travel in Canada.
+  "This paper describes the open data product, {ActiveCA}, with Canadian time use data about active travel.
+{ActiveCA} is an `R` data package that contains analysis-ready data from cycles  General Social Survey cicles regarding active travel in Canada.
 The package provides data set detailing active trip origins, destinations, and duration, covering a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.

### 2024-11-17 R/ActiveCA-package.R

@@ -19,12 +19,12 @@ NULL
 #' @format A data frame containing 4347 rows and 17 variables; each row represents a unique episode.
 #'   \describe{
 #'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
-#'   \item{WGHT_EPI}{Episode weight.}
+#'   \item{WGHT_EPI}{Episode weight. In 1986 the episode weight is the same as the summary weight, and is the number of persons that a record represents (see p. 18 of Cycle 2's Documentation and User's Guide)}
 #'   \item{ACTCODE}{Activity code of the episode.}
+#'   \item{act_label}{Activity (labeled).}
 #'   \item{STARTIME}{Start time (in minutes) of the episode.}
 #'   \item{ENDTIME}{End time (in minutes) of the episode.}
 #'   \item{DURATION}{Duration (in minutes) of the episode.}
-#'   \item{LOCATION}{Location/mode of transport of the episode.}
 #'   \item{origin}{Place of origin of the episode (in code).}
 #'   \item{destination}{Place of destination of the episode (in code).}
 #'   \item{orig_label}{Place of origin of the episode (labeled).}
@@ -40,7 +40,9 @@ NULL
 #' @keywords Walking 1986
 #' @name walking_1986
 #' @usage data(walking_1986)
-#' @source "[General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) / [Odesi](http://odesi2.scholarsportal.info/webview/)"
+#' @source [General Social Survey (GSS)](https://www150.statcan.gc.ca/n1/pub/45-25-0001/index-eng.htm#a7) / [Odesi](http://odesi2.scholarsportal.info/webview/)
+#' @source Adapted from Statistics Canada, General Social Survey Cycle 2, 1986. This does not constitute an endorsement by Statistics Canada of this product.
+#' @seealso Cycle 2: Time Use, Social Mobility and Language Use. Public Use Microdata File [Documentation and User’s Guide](https://publications.gc.ca/collections/collection_2017/statcan/12-0002/CS12-0002-1989-eng.pdf)
 "walking_1986"
 
 #' Preprocessed walking episodes (1992).
@@ -53,12 +55,12 @@ NULL
[…]

### 2024-11-29 data-raw/02-calculating-impedances.Rmd
Adding the weight variable to the impedance data set.

@@
 ```{r}
-gss_impedances <- gss_episodes[,c("PUMFID","DURATION",
+gss_impedances <- gss_episodes[,c("PUMFID","WGHT_EPI","DURATION",
                                   "origin","destination","Province","Pop_centre","orig_label",

### 2024-11-29 R/ActiveCA-package.R

@@ -451,6 +451,7 @@ NULL
 #' @format A data frame containing 12113 rows and 18 variables; each row represents a unique episode.
 #'   \describe{
 #'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
+#'   \item{WGHT_EPI}{Episode weight.}
 #'   \item{DURATION}{Duration (in minutes) of the episode.}
 #'   \item{origin}{Place of origin of the episode (in code).}
 #'   \item{destination}{Place of destination of the episode (in code).}

### 2024-12-02 paper/Manuscript-Data-Package.Rmd
Paper graphs and table update.

--- 2024-11-13 paper/Manuscript-Data-Package.Rmd
@@
 abstract: 
-  "This paper describes the open data product {ActiveCA}.
-{ActiveCA} is an `R` data package that contains analysis-ready data from the General Social Survey cicles regarding active travel in Canada.
+  "This paper describes the open data product `ActiveCA`, with Canadian time use data about active travel.
+`ActiveCA` is an `R` data package that contains analysis-ready data from cycles  General Social Survey cicles regarding active travel in Canada.
 The package provides data set detailing active trip origins, destinations, and duration, covering a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.
 The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
-{ActiveCA} also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
+`ActiveCA` also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
 The package will continue to expand with contributions from the authors and the broader community through requests in the future.
-{ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed."
+`ActiveCA` is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed."
 
@@
   out.width = "1\\linewidth")
+
+opts <- options(knitr.kable.NA = "")
 ```
@@
 
+```{r rename-labels, include = FALSE}
+# Renaming 'dest_label' and 'orig_label' values to reduce table space
+gss_episodes_renamed <- gss_episodes %>%
+  mutate(dest_label = case_when(
+    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
+    dest_label == "Restaurant, bar or club" ~ "Restaurant",
+    dest_label == "Sports centre, field or arena" ~ "Sport area",
+    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
+    dest_label == "In the neighbourhood" ~ "Neighbourhood",
+    dest_label == "Library, museum or theatre" ~ "Cultural venues",
+    TRUE ~ dest_label  # Keep all other values as they are
+  ), 
+  orig_label = case_when(
+    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
+    orig_label == "Restaurant, bar or club" ~ "Restaurant",
+    orig_label == "Sports centre, field or arena" ~ "Sport area",
+    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
+    orig_label == "In the neighbourhood" ~ "Neighbourhood",
+    orig_label == "Library, museum or theatre" ~ "Cultural venues",
+    TRUE ~ orig_label))
+
+gss_impedances_renamed <- gss_impedances %>%
+  mutate(dest_label = case_when(
+    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
+    dest_label == "Restaurant, bar or club" ~ "Restaurant",
+    dest_label == "Sports centre, field or arena" ~ "Sport area",
+    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
+    dest_label == "In the neighbourhood" ~ "Neighbourhood",
+    dest_label == "Library, museum or theatre" ~ "Cultural venues",
+    TRUE ~ dest_label  # Keep all other values as they are
+  ), 
+  orig_label = case_when(
+    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
+    orig_label == "Restaurant, bar or club" ~ "Restaurant",
+    orig_label == "Sports centre, field or arena" ~ "Sport area",
+    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
+    orig_label == "In the neighbourhood" ~ "Neighbourhood",
+    orig_label == "Library, museum or theatre" ~ "Cultural venues",
+    TRUE ~ orig_label))
+```
+
 # Introduction
 
-This paper presents the open data product {ActiveCA}. Open data products
+This paper presents the open data product `ActiveCA`. Open data products
 (ODPs) are the outcome of a transparent process that transforms raw data
@@
 
-{ActiveCA} is an `R` data package that provides analysis-ready data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) [@statisticscanada2024] regarding active travel.
+`ActiveCA` is an `R` data package that provides analysis-ready data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) [@statisticscanada2024] regarding active travel.
 These data can be used to study walking and cycling in Canada for various purposes over time, and to calculate impedance functions for accessibility analysis.
@@
 The aim of this paper is to walk readers through the data sets and
-invite others to experiment in its uses and applications. {ActiveCA} is
+invite others to experiment in its uses and applications. `ActiveCA` is
 freely available on GitHub for all to install and freely use in the
 spirit of open and reproducible research.
-In addition of enabling to obtain impedance functions, {ActiveCA} can be adopted in various applications that even go beyond the range of possibilities we have imagined.
+In addition of enabling to obtain impedance functions, `ActiveCA` can be adopted in various applications that even go beyond the range of possibilities we have imagined.
 Not only the data, but also all the code documenting the processing methodology is available for consultation and
@@
 Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
-{ActiveCA} includes all these episodes ready for analysis.
+`ActiveCA` includes all these episodes ready for analysis.
 Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, including metrics such as the count
@@
   group_by(YEAR, MODE) %>%
-  summarise(count = n(), .groups = "drop") %>%
-  pivot_wider(names_from = YEAR, values_from = count)
+  summarise(Count = n(), .groups = "drop") %>%
+  pivot_wider(names_from = YEAR, values_from = Count)
 
@@
     max = max(DURATION),
+    std = sd(DURATION),
     .groups = "drop"
   ) %>%
-  pivot_longer(cols = c(min, mean, median, max), names_to = "Statistic", values_to = "Value") %>%
+  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") %>%
   pivot_wider(names_from = YEAR, values_from = Value)
@@
 final_data <- bind_rows(
-  episodes_counts %>% mutate(Statistic = "count"),
+  episodes_counts %>% mutate(Statistic = "Count"),
   stats_data
@@
 # Rename the 'MODE' column to 'Mode' before generating the table
-final_data <- final_data %>%
+final_data <- final_data%>% 
+  arrange(MODE) %>%
   rename(Mode = MODE) %>%
-  dplyr::select(Mode, Statistic, everything())
+  dplyr::select(Mode, Statistic, everything()) %>% 
+   mutate(Statistic = case_when(
+    Statistic == "max" ~ "Maximum",
+    Statistic == "min" ~ "Minimum",
+    Statistic == "mean" ~ "Mean",
+    Statistic == "median" ~ "Median",
+    Statistic == "std" ~ "Standard deviation",
+    TRUE ~ Statistic  # Keep all other values as they are
+  ))
 
@@
 wide_data <- wide_data %>%
-  group_by(MODE) %>%
-  mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
-  ungroup() %>%
  dplyr::select(
-    dest_label, MODE,
+    MODE, dest_label,
     min_1986, median_1986, max_1986, Percentage_1986,
@@
     min_1998, median_1998, max_1998, Percentage_1998
-  )
+  ) %>% 
+  arrange(MODE, dest_label)
 
@@
   column_spec(10, border_right = TRUE) %>%
-  footnote(general = c("* The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination"), threeparttable = T)
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
+  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
 
@@
 ```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
-# Renaming 'dest_label' and 'orig_label' values to reduce table space
-gss_episodes_renamed <- gss_episodes %>%
-  mutate(dest_label = case_when(
-    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
-    dest_label == "Restaurant, bar or club" ~ "Restaurant",
-    dest_label == "Sports centre, field or arena" ~ "Sport area",
-    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
-    dest_label == "In the neighbourhood" ~ "Neighbourhood",
-    dest_label == "Library, museum or theatre" ~ "Cultural venues",
-    TRUE ~ dest_label  # Keep all other values as they are
-  ), 
-  orig_label = case_when(
-    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
-    orig_label == "Restaurant, bar or club" ~ "Restaurant",
-    orig_label == "Sports centre, field or arena" ~ "Sport area",
-    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
-    orig_label == "In the neighbourhood" ~ "Neighbourhood",
-    orig_label == "Library, museum or theatre" ~ "Cultural venues",
-    TRUE ~ orig_label))
-
 # Selecting surveys after 2000 and calculating trip duration statistics 
@@
  dplyr::select(
-    dest_label, MODE,
+    MODE, dest_label,
     min_2005, median_2005, max_2005, Percentage_2005,
@@
   ) %>% 
-  arrange(MODE, dest_label, desc(Percentage_2015)) %>%
-   group_by(MODE) %>%
-   mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
-   ungroup()
+  arrange(MODE, dest_label)
 
@@
     caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
-    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
+    col.names = c("Destination", "Mode","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
   ) %>%
@@
                 font_size = 6) %>%
-  column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
-  column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
+  #column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
+  #column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
   add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
@@
   column_spec(10, border_right = TRUE) %>% 
-  footnote(general = c("* The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination"), threeparttable = T)
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
+  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
 
@@
 
-{ActiveCA} also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques. Figures \ref{fig:figure-01} and
+`ActiveCA` also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques. Figures \ref{fig:figure-01} and
 \ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
@@
                                        labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
-                  geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
-                      aes(label = round(value)), 
-                      size = 4, color = "black") +
+                  #geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
+                  #    aes(label = round(value)), 
+                  #    size = 4, color = "black") +
                   labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
@@
 
-{ActiveCA} includes a total of
+`ActiveCA` includes a total of
 `r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
@@
 Table \ref{tab:table-04} shows the best impedance functions for walking trips to `Work or school` across survey years, where all distributions, except for a gamma function in 1998, are log-normal.
+
+```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
+# selecting mode and destination to display impedance functions
+mode <- c("Walking","Cycling")
+destino <- "Work or school"
+
+# filtering the impedance functions in accordance with the mode/destination setted in the step before and creating a latex table 
+selected_functions <- gss_impedances_renamed %>%
+  filter(dest_label == destino, MODE %in% mode) %>%
+  group_by(MODE, YEAR, dest_label) %>% 
+  summarise(Function = first(f_name), 
+            est_1 = first(est_1),
+            est_2 = first(est_2),
+            AIC = first(aic),
+            f_name = first(f_name),
+            count = first(count),
+            .groups = "drop") %>% 
+  mutate(Function = case_when(
+    Function == "lnorm" ~ "Lognormal",
+    Function == "gamma" ~ "Gamma",
+    Function == "unif" ~ "Uniform",
+    TRUE ~ Function), 
+  across(c(est_1,est_2), ~ round(., digits = 2)), 
+  across(c(AIC), ~ round(., digits = 0))) %>% 
+  rename(Mode = MODE, Year = YEAR, Destination = dest_label) %>%
+  unique() %>% 
+  arrange(Mode, Year) %>%
+  select(Mode, Year, Function, est_1, est_2, AIC, count)
+
+selected_functions_tab <- selected_functions %>% 
+  kable(format = "latex",
+        booktabs = TRUE,
+        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
+        col.names = c("Mode", "Year", "Impedance function","Parameter 1*","Parameter 2*", "AIC","Count")) %>% 
+    kable_styling(full_width = FALSE, 
+                latex_options=c("scale_down"),
+                font_size = 10) %>% 
+  row_spec(0, bold = TRUE, align = "c") %>% 
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>%
+  footnote(general = c("*For lognormal distributions, 'Parameter 1' and 'Parameter 2' refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
+For gamma distribution, 'Parameter 1' and 'Parameter 2' refer to the rate and shape of the distribution, respectively.
+`AIC` means Akaike information criterion."), threeparttable = T)
+
+selected_functions_tab
+```
+
 Figure
@@
 
-```{r include = FALSE}
-# Function to visualize impedance functions
-visualize_impedance <- function(df_functions, travel_cost){
+```{r theoretical_impedance_values, include = FALSE}
+theoretical_impedance <- function(df_functions, travel_cost) {
 
-  x <- data.frame(t = seq(1, travel_cost, 1))
+  # Initialize the main data frame to store results
+  result <- data.frame()
 
-  for (year in unique(df_functions$YEAR)) {
+  # Create the 't' sequence
+  t_seq <- seq(1, travel_cost, 1)
 
-    name_function <- df_functions[df_functions$YEAR == year, ]$f_name
+  for (mode in unique(df_functions$Mode)){
+
+    for (year in unique(df_functions[df_functions$Mode == mode, ]$Year)){
+
+      for (destino in unique(df_functions[df_functions$Year == year & df_functions$Mode == mode, ]$Destination)){
+
+      name_function <- df_functions[df_functions$Year == year & 
+                                      df_functions$Mode == mode & 
+                                      df_functions$Destination == destino, ]$f_name
+
+      print(paste0("YEAR: ", year, " MODE: ", mode," Destination: ", destino))
 
       subset <- df_functions %>% 
-      filter(YEAR == year)
+        filter(Year == year & Mode == mode & Destination == destino)
 
-    # new colunm with the year information in the name
-    column_name <- paste0("year_", year)
+      # Temporary data frame for this iteration
+      temp_df <- data.frame(t = t_seq)
 
-    # lnorm distribution 
+      # Add the appropriate distribution
       if (name_function == "lnorm") {
-      x <- x %>%
-        mutate(!!sym(column_name) := dlnorm(t,
+        temp_df$f <- dlnorm(temp_df$t,
                             meanlog = subset$est_1[1],
-                                             sdlog = subset$est_2[1]))
+                            sdlog = subset$est_2[1])
 
-    # Gamma distribution 
       } else if (name_function == "gamma") {  
+        temp_df$f <- dgamma(temp_df$t,
+                            shape = subset$est_1[1],
+                            rate = subset$est_2[1])
 
-      x <- x %>%
-        mutate(!!sym(column_name) := dgamma(t,
-                                             shape = subset$est_1[1],
-                                             rate = subset$est_2[1]))
+      } else if (name_function == "norm") {  
+        temp_df$f <- dnorm(temp_df$t,
+                           mean = subset$est_1[1],
+                           sd = subset$est_2[1])
 
-    # Norm distribution
+      } else if (name_function == "exp") {  
+        temp_df$f <- dexp(temp_df$t,
+                          rate = subset$est_1[1])
 
-    }else if (name_function == "norm") {  
-      x <- x %>%
-        mutate(!!sym(column_name) := dnorm(t,
-                                             mean = subset$est_1[1],
-                                             sd = subset$est_2[1]))
+      } else if (name_function == "unif") {  
+        temp_df$f <- dunif(temp_df$t,
+                           min = subset$est_1[1],
+                           max = subset$est_2[1])
+      }
 
-    # Exponential distribution
+      # Add mode, year and destination columns
+      temp_df$MODE <- mode
+      temp_df$YEAR <- year
+      temp_df$dest_label <- destino
 
-    }else if (name_function == "exp") {  
-      x <- x %>%
-        mutate(!!sym(column_name) := dexp(t,
-                                             exp = subset$est_1[1]))
-
-    # Uniform distribution
-
-    }else if (name_function == "unif") {  
-      x <- x %>%
-        mutate(!!sym(column_name) := dunif(t,
-                                             min = subset$est_1[1],
-                                             max = subset$est_2[1]))
+      # Append to the result data frame
+      result <- rbind(result, temp_df)
     }
   }
+}
+  return(result)
+}
 
-  return(x)
-}
+max_duration <- 100
+
+selected_functions <- gss_impedances_renamed %>% 
+  group_by(MODE, YEAR, dest_label) %>% 
+  summarise(Function = first(f_name), 
+            est_1 = first(est_1),
+            est_2 = first(est_2),
+            AIC = first(aic),
+            f_name = first(f_name),
+            count = first(count),
+            .groups = "drop") %>% 
+  mutate(Function = case_when(
+    Function == "lnorm" ~ "Lognormal",
+    Function == "gamma" ~ "Gamma",
+    Function == "unif" ~ "Uniform",
+    TRUE ~ Function), 
+  across(c(est_1,est_2), ~ round(., digits = 2)), 
+  across(c(AIC), ~ round(., digits = 0))) %>% 
+  rename(Mode = MODE, Year = YEAR, Destination = dest_label)
+
+theoretical_values <- theoretical_impedance(selected_functions, travel_cost = max_duration)
 ```
 
-```{r , include=FALSE}
-# Setting the maximum time to be considered to plot the impedance graphs
-max_time <- 100
+```{r include = FALSE}
 
-# Loop to generate all impedance plots at once 
-for(destino in unique(gss_impedances$orig_label)){
+for(destino in unique(gss_impedances_renamed$dest_label)){
 
-  for(mode in unique(gss_impedances[gss_impedances$orig_label == destino, ]$MODE)){
+  medianas <- gss_impedances_renamed %>%
+    filter(dest_label == destino) %>%
+    group_by(MODE) %>%
+    summarise(mediana_duracao = median(DURATION, na.rm = TRUE), .groups = 'drop')
 
-  selected_functions <-  gss_impedances %>% 
-  filter(dest_label == destino, MODE == mode) %>%
-  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic") %>%
-  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
-  mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
-  unique()
+  maximos <- theoretical_values %>%
+    filter(dest_label == destino) %>%
+    group_by(MODE, YEAR) %>%
+    filter(f == max(f)) %>%
+    slice(which.max(t)) %>%
+    ungroup()
 
-  if(nrow(selected_functions) > 0){
-    tld_theoretical <- visualize_impedance(df_functions = selected_functions, travel_cost = max_time)
+combined_plot <- gss_impedances_renamed %>% 
+  filter(dest_label == destino) %>%
+  mutate(dest_label = factor(dest_label, levels = destination_order)) %>%
+  ggplot(aes(x= DURATION)) +
+  geom_histogram(aes(y = ..density.., colour = MODE, fill = MODE), bins = 20, alpha=0.1, position="identity") + 
+  geom_line(data = theoretical_values[theoretical_values$dest_label == destino, ], aes(x = t, y = f, colour = MODE), size = 0.8) +   
+  geom_vline(data = medianas, aes(xintercept = mediana_duracao, color = MODE), linetype = "dashed") + 
+  geom_point(data = maximos, aes(x = t, y = f, colour = MODE), size = 3) + 
+  geom_label(data = maximos, aes(x = t, y = f, label = round(t, 2), colour = MODE), 
+            vjust = 0, size = 3) +
+  theme_minimal() +
+  facet_wrap(~YEAR) + 
+  labs(x = "Trip duration (in minutes)", y = "Density", colour = "Transportation Mode", fill = "Transportation Mode") +
+  theme(legend.position="right", 
+    strip.text = element_text(face = "bold", hjust = 0))
 
-  plots <- list()
-
-# Loop through each unique year
-  for(year in unique(gss_impedances$YEAR)){
-
-    # Subset empirical and theoretical data for the current year
-    empirical_subset <- gss_impedances %>% 
-      filter(dest_label == destino, MODE == mode, YEAR == year, DURATION <= max_time)
-
-    if(nrow(empirical_subset) > 1 ){
-
-      print(paste0(mode," - ", destino," - ", year))
-
-      theoretical_subset <- tld_theoretical %>%
-      dplyr::select(t, matches(as.character(year)))
-
-      # Create the ggplot
-      plots[[as.character(year)]] <- ggplot(empirical_subset, aes(x = DURATION)) +
-        geom_histogram(aes(y = ..density..), binwidth = 5, fill = "#dce1e3", color = "#adb1b3") +
-        geom_line(data = theoretical_subset, aes(x = t, y = theoretical_subset[[2]]), color = "#b52640", size = 0.80) +
-        labs(title = paste(year)) +
-        theme_minimal() +
-        scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
-        theme(axis.title.x = element_blank(),
-            axis.title.y = element_blank(), 
-            plot.title = element_text(size = 10, face="bold"))
-
-  }
-
-# Use gridExtra to arrange the plots in accordance with the number of survey years (some destination exits only in specific surveys)
+  plots <- length(unique(gss_impedances_renamed[gss_impedances_renamed$dest_label == destino, ]$YEAR))
 
   # Case of only one year 
-    if(length(plots) == 1){
-      combined_plot <- grid.arrange(
-      grobs = plots,
-      ncol = 1,
-      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
-      left = text_grob("Density", rot = 90)
-    )
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+    if(plots == 1){
+      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
        plot = combined_plot,
-       width = 9, 
+       width = 7, 
        height = 4, 
@@
     # Destination present in 3 surveys  
-    } else if(length(plots) == 3){
-      combined_plot <- grid.arrange(
-      grobs = plots,
-      ncol = 3,
-      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
-      left = text_grob("Density", rot = 90)
-    )
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+    } else if(plots == 3){
+      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
        plot = combined_plot,
-       width = 9, 
+       width = 14, 
        height = 4, 
@@
     }else {
-      combined_plot <- grid.arrange(
-      grobs = plots,
-      ncol = 3,
-      bottom = textGrob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode."), gp = gpar(fontsize = 16)),
-      left = textGrob("Density", rot = 90, gp = gpar(fontsize = 16))
-    )
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
+      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
        plot = combined_plot,
        width = 14, 
-       height = 8, 
+       height = 9, 
        units = "in", 
@@
 }
-}
-}
-} 
 ```
 
-```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
-# selecting mode and destination to display impedance functions
-mode <- "Walking" 
-destino <- "Work or school"
+```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Empirical data and impedance functions fitted for `work or school` as destination."), fig.align="center"}
 
-# filtering the impedance functions in accordance with the mode/destination setted in the step before and creating a latex table 
-gss_impedances %>% 
-filter(dest_label == destino, MODE == mode) %>%
-  dplyr::select("YEAR","f_name", "est_1","est_2","aic") %>%
-  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
-  mutate(across(c(aic), ~ round(., digits = 0)))  %>%
-  unique() %>% 
-  kable(format = "latex",
-        booktabs = TRUE,
-        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
-        col.names = c("Year", "Impedance function*","Parameter 1*","Parameter 2*", "AIC")) %>% 
-    kable_styling(full_width = FALSE, 
-                latex_options=c("scale_down"),
-                font_size = 10) %>% 
-  row_spec(0, bold = TRUE, align = "c") %>% 
-  footnote(general = c("* `lnorm` refers to the log-normal distribution.
-For `lnorm` distributions, `Parameter 1` and `Parameter 2` refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
-For the `gamma` distribution, `Parameter 1` and `Parameter 2` refer to the rate and shape of the distribution, respectively.
-`AIC` means Akaike information criterion."), threeparttable = T)
-```
-
-```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Empirical data and impedance functions fitted for walking trips with `work or school` as destination."), fig.align="center"}
-
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/impf_Work or school.jpg"))
 ```
@@
 
-This article presents {ActiveCA}, an open data product that provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in Canada.
-In the form of an `R` data package, {ActiveCA} was developed after collecting, cleaning, and processing the survey data, providing information on origins, destinations, and duration of active travel, as well other information.
+This article presents `ActiveCA`, an open data product that provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in Canada.
+In the form of an `R` data package, `ActiveCA` was developed after collecting, cleaning, and processing the survey data, providing information on origins, destinations, and duration of active travel, as well other information.
 Additionally, the package includes a series of pre-estimated impedance functions for walking and cycling trips, considering various destinations and time periods, for Canadian Metropolitan and Census Agglomeration areas.
 
-The value of {ActiveCA} lies in its transparency, accessibility, and
+The value of `ActiveCA` lies in its transparency, accessibility, and
 open infrastructure, which facilitates the addition of complementary
@@
 transparent research practices that encourage replication and innovation
-in related fields. We believe that {ActiveCA} will serve as a basis for
+in related fields. We believe that `ActiveCA` will serve as a basis for
 further research on GSS and for the integration of additional data by

### 2024-12-02 paper/Manuscript-Data-Package.tex
Paper graphs and table update.

@@
 \begin{abstract}
-This paper describes the open data product \{ActiveCA\}. \{ActiveCA\} is
-an \texttt{R} data package that contains analysis-ready data from the
+This paper describes the open data product \texttt{ActiveCA}, with
+Canadian time use data about active travel. \texttt{ActiveCA} is an
+\texttt{R} data package that contains analysis-ready data from cycles
 General Social Survey cicles regarding active travel in Canada. The
@@
 respondent's region characteristics, specifying wheter they live in a
-metropolitan area and their province of residency. \{ActiveCA\} also
-provides pre-estimated impedance functions for active travel for
+metropolitan area and their province of residency. \texttt{ActiveCA}
+also provides pre-estimated impedance functions for active travel for
 Canadian Metropolitan Area and Census Agglomerations, which can be used
@@
 contributions from the authors and the broader community through
-requests in the future. \{ActiveCA\} is freely accessible for
+requests in the future. \texttt{ActiveCA} is freely accessible for
 exploration and download from the associated Github repository, where
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-This paper presents the open data product \{ActiveCA\}. Open data
+This paper presents the open data product \texttt{ActiveCA}. Open data
 products (ODPs) are the outcome of a transparent process that transforms
@@
 
-\{ActiveCA\} is an \texttt{R} data package that provides analysis-ready
-data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992),
-12 (1998), 19 (2005), 24 (2010), and 29 (2015)
+\texttt{ActiveCA} is an \texttt{R} data package that provides
+analysis-ready data from the General Social Survey of Canada Cycles 2
+(1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015)
 \citep{statisticscanada2024} regarding active travel. These data can be
@@
 The aim of this paper is to walk readers through the data sets and
-invite others to experiment in its uses and applications. \{ActiveCA\}
-is freely available on GitHub for all to install and freely use in the
-spirit of open and reproducible research. In addition of enabling to
-obtain impedance functions, \{ActiveCA\} can be adopted in various
-applications that even go beyond the range of possibilities we have
-imagined. Not only the data, but also all the code documenting the
-processing methodology is available for consultation and evaluation in
-its repository. This package contributes to reducing the barrier to
-using the information contained in GSS surveys to provide data-driven
-decisions in transportation analysis.
+invite others to experiment in its uses and applications.
+\texttt{ActiveCA} is freely available on GitHub for all to install and
+freely use in the spirit of open and reproducible research. In addition
+of enabling to obtain impedance functions, \texttt{ActiveCA} can be
+adopted in various applications that even go beyond the range of
+possibilities we have imagined. Not only the data, but also all the code
+documenting the processing methodology is available for consultation and
+evaluation in its repository. This package contributes to reducing the
+barrier to using the information contained in GSS surveys to provide
+data-driven decisions in transportation analysis.
 
+\hypertarget{general-social-survey-gss-collection}{%
 \section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}
+collection}\label{general-social-survey-gss-collection}}
 
@@
 
-\subsection{The Main File}\label{the-main-file}
+\hypertarget{the-main-file}{%
+\subsection{The Main File}\label{the-main-file}}
 
@@
 
-\subsection{The Episode File}\label{the-episode-file}
+\hypertarget{the-episode-file}{%
+\subsection{The Episode File}\label{the-episode-file}}
 
@@
 
-\section{Methodology}\label{methodology}
+\hypertarget{methodology}{%
+\section{Methodology}\label{methodology}}
 
@@
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-\subsection{Descriptive statistics}\label{descriptive-statistics}
+\hypertarget{descriptive-statistics}{%
+\subsection{Descriptive statistics}\label{descriptive-statistics}}
 
@@
 recorded active travel episodes, with trip duration ranging from 0 to
-900 minutes, to twelve different destinations. \{ActiveCA\} includes all
-these episodes ready for analysis. Table \ref{tab:table-01} presents
-descriptive statistics on walking and cycling trips between 1986 and
-2015, including metrics such as the count of recorded trips (count), and
-measures of trip duration in minutes: maximum (max), mean, median, and
-minimum (min). The 1986 survey did not include bicycle trips.
+900 minutes, to twelve different destinations. \texttt{ActiveCA}
+includes all these episodes ready for analysis. Table \ref{tab:table-01}
+presents descriptive statistics on walking and cycling trips between
+1986 and 2015, including metrics such as the count of recorded trips
+(count), and measures of trip duration in minutes: maximum (max), mean,
+median, and minimum (min). The 1986 survey did not include bicycle
+trips.
 
@@
 \midrule
- & count & 4347 & 1500 & 1670 & 5533 & 4379 & 3251\\
+ & Count & 4347 & 1500 & 1670 & 5533 & 4379 & 3251\\
 \nopagebreak
- & max & 660 & 300 & 255 & 515 & 480 & 900\\
+ & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
 \nopagebreak
- & mean & 21 & 19 & 11 & 12 & 12 & 17\\
+ & Mean & 21 & 19 & 11 & 12 & 12 & 17\\
 \nopagebreak
- & median & 10 & 10 & 5 & 10 & 8 & 10\\
+ & Median & 10 & 10 & 5 & 10 & 8 & 10\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & min & 1 & 1 & 1 & 0 & 0 & 5\\
+ & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
+\nopagebreak
+\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
 \cmidrule{1-8}\pagebreak[0]
- & count & NA & 135 & 119 & 333 & 236 & 245\\
+ & Count &  & 135 & 119 & 333 & 236 & 245\\
 \nopagebreak
- & max & NA & 240 & 90 & 180 & 153 & 120\\
+ & Maximum &  & 240 & 90 & 180 & 153 & 120\\
 \nopagebreak
- & mean & NA & 31 & 21 & 19 & 21 & 24\\
+ & Mean &  & 31 & 21 & 19 & 21 & 24\\
 \nopagebreak
- & median & NA & 20 & 15 & 15 & 15 & 15\\
+ & Median &  & 20 & 15 & 15 & 15 & 15\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & min & NA & 5 & 2 & 1 & 1 & 5\\
+ & Minimum &  & 5 & 2 & 1 & 1 & 5\\
+\nopagebreak
+\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
 \bottomrule
@@
 \item \textit{Note: } 
-\item * The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(\%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination
+\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
 \end{TableNotes}
@@
 \midrule
-Home &  & NA & NA & NA & NA & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
-Other's home & Cycling & NA & NA & NA & NA & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
-Work or school &  & NA & NA & NA & NA & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
-Home &  & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
-Other's home & Walking & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
-\addlinespace
-Work or school &  & 1 & 10 & 450 & 11.3 & 2 & 10 & 60 & 19.2 & 1 & 6.5 & 75 & 20.4\\
+ & Home &  &  &  &  & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
+\nopagebreak
+ & Other's home &  &  &  &  & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
+\nopagebreak
+\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
+\cmidrule{1-14}\pagebreak[0]
+ & Home & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
+\nopagebreak
+ & Other's home & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
+\nopagebreak
+\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.3 & 2 & 10 & 60 & 19.2 & 1 & 6.5 & 75 & 20.4\\
 \bottomrule
@@
 \item \textit{Note: } 
-\item * The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(\%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination
+\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
 \end{TableNotes}
-\begin{longtable}[t]{>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{}c|>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}>{\centering\arraybackslash}p{0.4cm}}
+\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
 \caption{\label{tab:table-03}\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015}\\
@@
 \cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
-\multicolumn{1}{>{\centering\arraybackslash}p{2cm}}{\textbf{Destination}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Mode*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)*}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Min}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Med}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{Max}} & \multicolumn{1}{>{\centering\arraybackslash}p{0.4cm}}{\textbf{(\%)}}\\
+\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min*}} & \multicolumn{1}{c}{\textbf{Med*}} & \multicolumn{1}{c}{\textbf{Max*}} & \multicolumn{1}{c}{\textbf{(\%)*}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
 \midrule
-Cultural venues &  & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
-Grocery store & Cycling & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
-Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 90 & 2.0\\
-Home &  & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
-Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 30.0 & 45 & 1.2\\
-\addlinespace
-Other's home &  & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
-Outdoors &  & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
-Place of worship &  & 20 & 20.0 & 20 & 0.3 & NA & NA & NA & NA & 15 & 15.0 & 15 & 0.4\\
-Restaurant &  & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
-Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 10 & 15.0 & 15 & 2.9\\
-\addlinespace
-Work or school &  & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
-Business &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 30 & 0.2\\
-Cultural venues & Walking & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
-Grocery store &  & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
-Health clinic &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 130 & 1.0\\
-\addlinespace
-Home &  & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
-Neighbourhood &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 60 & 2.1\\
-Other's home &  & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
-Outdoors &  & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
-Place of worship &  & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
-\addlinespace
-Restaurant &  & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
-Sport area &  & NA & NA & NA & NA & NA & NA & NA & NA & 5 & 10.0 & 45 & 3.3\\
-Work or school &  & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
+ & Cultural venues & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
+\nopagebreak
+ & Grocery store & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
+\nopagebreak
+ & Health clinic &  &  &  &  &  &  &  &  & 10 & 15.0 & 90 & 2.0\\
+\nopagebreak
+ & Home & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
+\nopagebreak
+ & Neighbourhood &  &  &  &  &  &  &  &  & 10 & 30.0 & 45 & 1.2\\
+\nopagebreak
+ & Other's home & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
+\nopagebreak
+ & Outdoors & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
+\nopagebreak
+ & Place of worship & 20 & 20.0 & 20 & 0.3 &  &  &  &  & 15 & 15.0 & 15 & 0.4\\
+\nopagebreak
+ & Restaurant & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
+\nopagebreak
+ & Sport area &  &  &  &  &  &  &  &  & 10 & 15.0 & 15 & 2.9\\
+\nopagebreak
+\multirow[t]{-11}{*}{\centering\arraybackslash Cycling} & Work or school & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
+\cmidrule{1-14}\pagebreak[0]
+ & Business &  &  &  &  &  &  &  &  & 5 & 10.0 & 30 & 0.2\\
+\nopagebreak
+ & Cultural venues & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
+\nopagebreak
+ & Grocery store & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
+\nopagebreak
+ & Health clinic &  &  &  &  &  &  &  &  & 5 & 10.0 & 130 & 1.0\\
+\nopagebreak
+ & Home & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
+\nopagebreak
+ & Neighbourhood &  &  &  &  &  &  &  &  & 5 & 10.0 & 60 & 2.1\\
+\nopagebreak
+ & Other's home & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
+\nopagebreak
+ & Outdoors & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
+\nopagebreak
+ & Place of worship & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
+\nopagebreak
+ & Restaurant & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
+\nopagebreak
+ & Sport area &  &  &  &  &  &  &  &  & 5 & 10.0 & 45 & 3.3\\
+\nopagebreak
+\multirow[t]{-12}{*}{\centering\arraybackslash Walking} & Work or school & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
 \bottomrule
@@
 
-\{ActiveCA\} also enables visual analysis of active travel in Canada
-using traditional exploratory data analysis techniques. Figures
+\texttt{ActiveCA} also enables visual analysis of active travel in
+Canada using traditional exploratory data analysis techniques. Figures
 \ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling
@@
 
+\hypertarget{pre-estimated-impedance-functions-for-active-travel-in-canada}{%
 \subsection{Pre-estimated impedance functions for active travel in
-Canada}\label{pre-estimated-impedance-functions-for-active-travel-in-canada}
+Canada}\label{pre-estimated-impedance-functions-for-active-travel-in-canada}}
 
-\{ActiveCA\} includes a total of 64 pre-estimated impedance functions
-for cycling and walking trips for Census Agglomerations and Census
-Metropolitan Areas. Table \ref{tab:table-04} shows the best impedance
-functions for walking trips to \texttt{Work\ or\ school} across survey
-years, where all distributions, except for a gamma function in 1998, are
-log-normal. Figure \ref{fig:figure-03} illustrates these fitted
-functions (red line) alongside histograms of empirical travel times
-(grey bars). As for the others pre-estimated functions, these examples
-allow to calculate gravity-based accessibility measures for active
-travel across various destinations and temporal scales in Canadian urban
-areas.
+\texttt{ActiveCA} includes a total of 64 pre-estimated impedance
+functions for cycling and walking trips for Census Agglomerations and
+Census Metropolitan Areas. Table \ref{tab:table-04} shows the best
+impedance functions for walking trips to \texttt{Work\ or\ school}
+across survey years, where all distributions, except for a gamma
+function in 1998, are log-normal.
 
@@
 \caption{\label{tab:table-04}\label{table-04}Impedance functions and AIC for `Walking` trips considering `Work or school` as destination.}
+\caption{\label{tab:table-04}\label{table-04}Impedance functions and AIC for `Cycling` trips considering `Work or school` as destination.}
 \centering
@@
 \begin{threeparttable}
-\begin{tabular}[t]{rlrrr}
+\begin{tabular}[t]{lrlrrrr}
 \toprule
-\multicolumn{1}{c}{\textbf{Year}} & \multicolumn{1}{c}{\textbf{Impedance function*}} & \multicolumn{1}{c}{\textbf{Parameter 1*}} & \multicolumn{1}{c}{\textbf{Parameter 2*}} & \multicolumn{1}{c}{\textbf{AIC}}\\
+\multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Year}} & \multicolumn{1}{c}{\textbf{Impedance function}} & \multicolumn{1}{c}{\textbf{Parameter 1*}} & \multicolumn{1}{c}{\textbf{Parameter 2*}} & \multicolumn{1}{c}{\textbf{AIC}} & \multicolumn{1}{c}{\textbf{Count}}\\
 \midrule
-2015 & lnorm & 2.55 & 0.64 & 6612061\\
-2010 & lnorm & 2.21 & 0.78 & 7917431\\
-2005 & lnorm & 2.13 & 0.79 & 8182691\\
-1998 & gamma & 1.23 & 0.09 & 2318752\\
-1992 & lnorm & 2.38 & 0.70 & 2319400\\
+ & 1992 & Gamma & 3.00 & 0.17 & 433582 & 19\\
+
+ & 1998 & Gamma & 3.37 & 0.10 & 481536 & 19\\
+
+ & 2005 & Lognormal & 2.93 & 0.70 & 888655 & 64\\
+
+ & 2010 & Lognormal & 2.65 & 0.77 & 1292760 & 53\\
+
+\multirow[t]{-5}{*}{\raggedright\arraybackslash Cycling} & 2015 & Lognormal & 3.03 & 0.41 & 1162876 & 63\\
+\cmidrule{1-7}
+ & 1992 & Lognormal & 2.38 & 0.70 & 2319400 & 113\\
+
+ & 1998 & Gamma & 1.23 & 0.09 & 2318752 & 109\\
+
+ & 2005 & Lognormal & 2.13 & 0.79 & 8182691 & 724\\
+
+ & 2010 & Lognormal & 2.21 & 0.78 & 7917431 & 494\\
+
+\multirow[t]{-5}{*}{\raggedright\arraybackslash Walking} & 2015 & Lognormal & 2.55 & 0.64 & 6612061 & 407\\
 \bottomrule
@@
 \item \textit{Note: } 
-\item * `lnorm` refers to the log-normal distribution.
-For `lnorm` distributions, `Parameter 1` and `Parameter 2` refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
-For the `gamma` distribution, `Parameter 1` and `Parameter 2` refer to the rate and shape of the distribution, respectively.
+\item *For lognormal distributions, 'Parameter 1' and 'Parameter 2' refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
+For gamma distribution, 'Parameter 1' and 'Parameter 2' refer to the rate and shape of the distribution, respectively.
 `AIC` means Akaike information criterion.
@@
 
+Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
+alongside histograms of empirical travel times (grey bars). As for the
+others pre-estimated functions, these examples allow to calculate
+gravity-based accessibility measures for active travel across various
+destinations and temporal scales in Canadian urban areas.
+
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/impf_Work or school_Walking}
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/impf_Work or school}
 
@@
 
-\caption{Empirical data and impedance functions fitted for walking trips with `work or school` as destination.}\label{fig:figure-03}
+\caption{Empirical data and impedance functions fitted for `work or school` as destination.}\label{fig:figure-03}
 \end{figure}
 
-\section{Concluding remarks}\label{concluding-remarks}
+\hypertarget{concluding-remarks}{%
+\section{Concluding remarks}\label{concluding-remarks}}
 
-This article presents \{ActiveCA\}, an open data product that provides
-analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
-(2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in
-Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
-developed after collecting, cleaning, and processing the survey data,
-providing information on origins, destinations, and duration of active
-travel, as well other information. Additionally, the package includes a
-series of pre-estimated impedance functions for walking and cycling
-trips, considering various destinations and time periods, for Canadian
-Metropolitan and Census Agglomeration areas.
+This article presents \texttt{ActiveCA}, an open data product that
+provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998),
+19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel
+in Canada. In the form of an \texttt{R} data package, \texttt{ActiveCA}
+was developed after collecting, cleaning, and processing the survey
+data, providing information on origins, destinations, and duration of
+active travel, as well other information. Additionally, the package
+includes a series of pre-estimated impedance functions for walking and
+cycling trips, considering various destinations and time periods, for
+Canadian Metropolitan and Census Agglomeration areas.
 
-The value of \{ActiveCA\} lies in its transparency, accessibility, and
-open infrastructure, which facilitates the addition of complementary
+The value of \texttt{ActiveCA} lies in its transparency, accessibility,
+and open infrastructure, which facilitates the addition of complementary
 data sets in the future. R users can seamlessly explore GSS walking and
@@
 encourage replication and innovation in related fields. We believe that
-\{ActiveCA\} will serve as a basis for further research on GSS and for
-the integration of additional data by the authors or the wider open
+\texttt{ActiveCA} will serve as a basis for further research on GSS and
+for the integration of additional data by the authors or the wider open
 source community.
 
+\hypertarget{declaration-of-conflicting-interests}{%
 \section{Declaration of Conflicting
-Interests}\label{declaration-of-conflicting-interests}
+Interests}\label{declaration-of-conflicting-interests}}
 
@@
 
-\section{Funding}\label{funding}
+\hypertarget{funding}{%
+\section{Funding}\label{funding}}
 
@@
 
-\section{ORCID}\label{orcid}
+\hypertarget{orcid}{%
+\section{ORCID}\label{orcid}}
 
@@
 
-\section{Data availability statement}\label{data-availability-statement}
+\hypertarget{data-availability-statement}{%
+\section{Data availability
+statement}\label{data-availability-statement}}
 

### 2024-12-04 paper/Manuscript-Data-Package.Rmd
Started editing the paper after Bruno's work

@@
 abstract: 
-  "This paper describes the open data product `ActiveCA`, with Canadian time use data about active travel.
-`ActiveCA` is an `R` data package that contains analysis-ready data from cycles  General Social Survey cicles regarding active travel in Canada.
-The package provides data set detailing active trip origins, destinations, and duration, covering a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others.
-The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
-`ActiveCA` also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
-The package will continue to expand with contributions from the authors and the broader community through requests in the future.
-`ActiveCA` is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed."
+  "This paper describes {ActiveCA}, and open data product with Canadian time use 
+data. {ActiveCA} is an `R` data package that contains analysis-ready data related 
+to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992),
+12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
+Active travel is characterized by mode, with walking being part of every cycle 
+and bicycling starting in 1992. The attributes of active trips are the types of 
+locations of origins and destinations, the duration of trips, and episode weights
+for expanding the trips to population-wide estimates. Based on the year of the 
+survey, a variety of locations are coded. In earlier cycles, these include home,
+work or school, and other's home, whereas in later cycles these are augmented with
+locations such as grocery stores, restaurants, outdoor destinations, and others.
+The geographical resolution includes the province and whether the episode was in
+an urban or rural setting."
 
-keywords: Active; mobility; walking; cycling; travel time; impedance; transportation; R 
+keywords: Active; mobility; walking; cycling; travel time; time-use; General Social Survey 
 classoption:
@@
 ---
+
+<!-- 
+FROM THE ABSTRACT:
+
+The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
+{ActiveCA} also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
+The package will continue to expand with contributions from the authors and the broader community through requests in the future.
+{ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed.
+-->
 
@@
 # Load data
-load(paste0(here(), "/data/gss_episodes.Rda"))
-load(paste0(here(), "/data/gss_impedances.Rda"))
+data("gss_episodes")
+data("gss_impedances")
 ```
@@
 
-This paper presents the open data product `ActiveCA`. Open data products
-(ODPs) are the outcome of a transparent process that transforms raw data
-(open or not) into analysis-ready data, in which all stages of
-development follow open principles [@arribas-bel2021]. ODPs differ from
-general open data due to their high utility, added value and open
-availability.
+The objective of this paper is to introduce {ActiveCA}, an open data product with
+time use data from Canada's General Social Surveys. Open data products
+(ODPs) are the outcome of a process that transforms raw data (open or not) into 
+analysis-ready data, following a transparent process in which all stages of
+development follow open principles [@arribas-bel2021]. ODPs, while still open, 
+differ from general open data in their degree of ease of access, their heightened 
+usability, and potentially the value they add to the raw data.
 
-`ActiveCA` is an `R` data package that provides analysis-ready data from the General Social Survey of Canada Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) [@statisticscanada2024] regarding active travel.
+{ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
+a period of almost 40 years. The source of these data is Canada's program of 
+General Social Surveys (GSSs). This program is designed to provide cross-sectional
+data on topics of interest to improve the well-being of Canadians. As part of this
+program, every five to seven years the survey is done on the topic of time use. 
+Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
+24 (2010), and 29 (2015) of the GSS. Time use data in these surveys is coded using
+a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
+working or at school. These surveys have proved valuable in investigations of 
+mobility and quality of life [@spinneyTransport2009], the relationship between
+active travel and transit use [@lachapelleLonger2016], and travel behavior and time
+poverty [@kimFacing2024], to name but a few examples.
+
+For {ActiveCA}, data we extracted all data needed to characterize active travel 
+in Canada, namely, episodes where the activity was identified as moving between 
+an origin and a destination, either by walking or cycling. Although Statistics 
+Canada offers Public Use Microdata Files and documentation for the GSS program 
+[see @statisticscanada2024], accessing these files, and preparing  them for analysis
+is not a straightforward matter, given their size and complexity.  The  process 
+of extracting information of interest from the source files is  time-consuming, 
+tedious, and challenging and/or prone to error due to the expertise required to 
+work with these files. To create {ActiveCA} we collected, cleaned, and processed
+the cycles of the GSS  surveys concerning time use to make them ready for analysis.
+
+{ActiveCA} is distributed as an `R` package with a number of data 
+objects and their documentation. `R` packages contain code, data, and documentation
+in a standardized format that can be installed by `R` users via a software repository,
+such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
+adroit medium to distribute analysis-ready data.
+
+Given the level of interest in active travel [e.g., @mccurdySupport2023], reducing
+the barriers to using data contained in rich, but difficult to access and use surveys,
+such as Canada's GSSs, is a worthy endeavour that can only improve data-driven 
+decisions in transportation, urban, and health policy.
+
+<!-- 
+is an `R` data package that provides analysis-ready data from the General Social Survey of Canada  [@statisticscanada2024] regarding active travel.
 These data can be used to study walking and cycling in Canada for various purposes over time, and to calculate impedance functions for accessibility analysis.
 The package provides pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
+-->
 
-To create this `R` data package, we collected, cleaned and processed the
-Time Use collections from the GSS surveys to make them ready for
-analysis. An `R` data package contains code, data and documentation in a
-standardized format that can be installed by `R` users via a centralized
-software repository, such as CRAN (Comprehensive R Archive Network) and
-GitHub. Although the GSS surveys are publicly sourced managed by
-Statistic Canada [@statisticscanada2024], preparing them for analysis can be time-consuming, tedious and perhaps not even possible for those who try, due to a lack of documentation or prior knowledge.
-
-The aim of this paper is to walk readers through the data sets and
-invite others to experiment in its uses and applications. `ActiveCA` is
-freely available on GitHub for all to install and freely use in the
-spirit of open and reproducible research.
-In addition of enabling to obtain impedance functions, `ActiveCA` can be adopted in various applications that even go beyond the range of possibilities we have imagined.
-Not only the data, but also all the code documenting the processing methodology is available for consultation and
-evaluation in its repository. This package contributes to reducing the
-barrier to using the information contained in GSS surveys to provide
-data-driven decisions in transportation analysis.
+The rest of this paper discusses the sources of data, and the process implemented
+to retrieve and package them. Then, we explain how to use the package and show
+some selected examples of analysis to whet the imagination of potential users.
+This ODP provides not only data that are easy to use, but also all the code and 
+documentation that make this a reproducible research project. In summary, {ActiveCA}
+aims to implement and inspire the best principles of open spatial sciences 
+[@paez_open_2021; @brunsdon_opening_2021].
 
@@
 
-Statistics Canada [-@statisticscanada2024] conducts GSS surveys to
-obtain data on social trends to track changes in Canadians' living
-conditions and well-being over time. This survey is used to understand
-how citizens spend and manage their time and what factors contribute to
-their happiness and stress. Created in 1985, the survey is part of a
-series of independent, annual, and cross-sectional surveys.
+Statistics Canada [-@statisticscanada2024] conducts GSS surveys to obtain data 
+on social trends to track changes in Canadians' living conditions and well-being
+over time. The series of survey on time use are used to understand how Canadian
+residents spend and manage their time, and what factors contribute to their 
+happiness and stress. The GSS program was created in 1985, and is serialized to 
+provide a collection of annual, representative cross-sectional surveys.
 
-In addition to the main topic, each GSS cycle includes new content that
-addresses emerging and policy-relevant issues. Every five to seven
-years, the Time Use Surveys [@statisticscanada2022] collect data on
-respondents' participation and time spent on a wide range of everyday
-activities using a 24-hour retrospective diary, with information on the
-location of these activities (e.g. at home, at work, etc.) and, for
-non-personal activities, the people who were present with the respondent
-at the time of the activity. In addition, time-use surveys also cover
-topics related to leisure time, work-life balance, health, commuting,
-culture and sports, and many others.
+The topics of the survey cycle every few years to cover topics that include family,
+health, social identity, and every five to seven years time use. The first Canadian
+time use survey done as part of the GSS program was conducted in 1986, and the 
+most recent was completed in 2015. These Time Use Surveys [@statisticscanada2022]
+collect data on respondents' participation and time spent on a wide range of 
+everyday activities using a 24-hour retrospective diary, with information on the 
+location of these activities (e.g. at home, at work, etc.) and, for non-personal
+activities, the people who were present with the respondent at the time of the 
+activity. In addition, time-use surveys also cover topics related to leisure time,
+work-life balance, health, commuting, culture and sports, and many others.
 
-The most recent time use survey was carried out in 2022 [@wray2024].
-However, the 2022 dataset has not been fully published and, because of
-this, our analysis focused on the surveys from 1986 to 2015 (1986, 1992,
-1998, 2005, 2010 and 2015). Time Use surveys are composed of two data
-sets, the main one and the episode file, explained in the following
-subsections.
+The Public Use Microdata are released by Statistics Canada in two files: a 
+_Main File_ and an _Episode File_. The files are linked by keys that identify 
+households, individuals, and episodes (i.e., activities) conducted by individuals.
+We discuss these files in more detail in the following section.
 
@@
 
-The main file compiles an large array of aggregated data, summarizing
-the answers to the questionnaire and derived variables that summarize
-the respondents' time use across different activities, locations, and
-social interactions. This file documents the time and duration that
-respondents allocate to each activity and location. The Main File
-provides a overview of daily routines and social dynamics, not focusing
-on individual activity episodes. Additionally, this file categorizes
-activities into bigger groups and subcategories, facilitating the data's
-analytical utility with additional metrics such as total transit time,
-duration spent with household members, and counts of activities and
-episodes.
+The main file of the time use survey compiles a large array of aggregated data, 
+summarizing the answers to the questionnaire that describe households and 
+individuals, as well as derived variables that summarize the respondents' use of 
+time use across different activities, locations, and social interactions. This 
+file documents the time and duration that respondents allocate to each activity 
+and location. The Main File provides a overview of daily routines and social 
+dynamics, not focusing on individual activity episodes. Additionally, this file 
+categorizes activities into bigger groups and subcategories, facilitating the 
+data's analytical utility with additional metrics such as total transit time, time 
+spent with household members, and counts of activities and episodes.
+
+<!-- 
+Maybe here provide an example of the structure of this table?
+-->
 
@@
 
-The episode file records detailed data for each activity episode
-reported by respondents. The entry includes the start and end times,
-duration, location, and accompanying social context, informing when and
-where activities occurred and with whom. The file distinguishes itself
-by focusing on individual episodes rather than respondents, with the
-data structured around the numerous activity instances that compose a
-day of the respondent. Although respondent-specific characteristics are
-not included within the episode file, it is possible to link the main
-file and the episode file by using an identifiable variable present in
-both data sets.
+The episode is a much bigger file that records detailed data for each activity 
+episode reported by respondents. Each episode represents a single activity and its
+duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
+entry in this file includes the start and end times of the activity, the duration,
+location, and accompanying social context, informing when and where activities 
+occurred and with whom. The focus of the Episode File is not on the characteristics
+of the respondents, but on the characteristics of the activities, and the data 
+are structured around the numerous activity instances that compose a day of the 
+respondent. Although respondent-specific characteristics are not included within
+the episode file, it is possible to link the main file and the episode file by 
+using a key present in both the Main and Episode Files.
 
-# Methodology
+<!-- 
+Maybe here provide an example of the structure of this table?
+-->
 
-For each selected cycle of the GSS surveys, we reviewed the
-episode files to identify cases with activities listed as walking or
-cycling, selecting the activities immediately before and after the
-mobility episode.
-After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, transportation mode, as well the province and urban classification of the respondent's residency.
+# Methods
+
+For each selected cycle of the GSS surveys, we reviewed the episode files to 
+identify episodes of movement that involved walking or cycling. This allowed us 
+to select the activities immediately before and after the movement episode. After
+that, we labeled the code variables with their appropriate descriptions, identifying
+each origin and destination, mode of travel, as well the province and urban 
+classification of the episode.
+
+<!--
+
+Show a graphical example of this process: how were active trips identified? 
+What was the purpose of finding the preceding and succeeding activities?
+
+-->
+
+<!--
+
+Not needed in this paper:
 
@@
 
-# Results
+-->
+
+# How to use {ActiveCA}
+
+<!--
+
+The paper does not really need to report "Results".
+There are too many potential use cases and we would rather let users find their own results depending on what they are interested in.
+Instead of results, a section on "How to use the package" is appropriate.
+There are a number of points to highlight here, for example, how to load the package, call the data sets, in particular the use of episode weights to make sure that any statistical analysis is valid.
+
+-->
 
@@
 
-`ActiveCA` also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques. Figures \ref{fig:figure-01} and
-\ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
-through heat maps. These maps use color gradients to represent the
-percentage of trips between various origins and destinations, with
-darker colors indicating higher percentages and lighter colors
-representing less frequent routes. To avoid overwhelming the reader, we
-omitted the heat maps for the other years analyzed.
+{ActiveCA} also enables visual analysis of active travel in Canada using traditional 
+exploratory data analysis techniques. Figures \ref{fig:figure-01} and \ref{fig:figure-02}
+show walking and cycling trips from 1992 and 2015 through heat maps. These maps 
+use color gradients to represent the percentage of trips between various origins
+and destinations, with darker colors indicating higher percentages and lighter 
+colors representing less frequent routes. For conciseness, we omitted the heat maps
+for the other years analyzed.
 
@@
 
+# Python integration
+
+CAN WE PROVIDE PYTHON INTEGRATION? See here:
+
+https://github.com/paezha/idealista18/tree/master/Python
+
+<!--
+
 ## Pre-estimated impedance functions for active travel in Canada
@@
 
-Figure
-\ref{fig:figure-03} illustrates these fitted functions (red line)
+Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
 alongside histograms of empirical travel times (grey bars). As for the
 others pre-estimated functions, these examples allow to calculate gravity-based
-accessibility measures for active travel across various destinations and temporal scales in Canadian urban areas.
+accessibility measures for active travel across various destinations and temporal 
+scales in Canadian urban areas.
 
@@
 
+-->
+
 # Concluding remarks
 
-This article presents `ActiveCA`, an open data product that provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in Canada.
-In the form of an `R` data package, `ActiveCA` was developed after collecting, cleaning, and processing the survey data, providing information on origins, destinations, and duration of active travel, as well other information.
-Additionally, the package includes a series of pre-estimated impedance functions for walking and cycling trips, considering various destinations and time periods, for Canadian Metropolitan and Census Agglomeration areas.
+This paper presents `ActiveCA`, an open data product that provides analysis-ready 
+data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
+of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
+{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
+providing information on origins, destinations, and duration of active travel, as
+well other information.
 
-The value of `ActiveCA` lies in its transparency, accessibility, and
-open infrastructure, which facilitates the addition of complementary
-data sets in the future. R users can seamlessly explore GSS walking and
-cycling episodes along with calibrated impedance functions, with the
-option to suggest enhancements to the package as needed. This article
-adopts the structure proposed by Anastasia and Páez [-@soukhov2023],
-whose work provided essential guidance for the creation of this package.
-Similarly, we aim to contribute to the academic community by promoting
-transparent research practices that encourage replication and innovation
-in related fields. We believe that `ActiveCA` will serve as a basis for
-further research on GSS and for the integration of additional data by
+It is important to remark that the present version of {ActiveCA} covers all Canadian
+time use surveys up to 2015. While the most recent time use survey was carried out
+in 2022 [@wray2024], the Public Use Microdata Files are currently unavailable, and
+at the moment it is estimated that they will only be published in the later part
+of 2025. The `R` package will be updated once these files become available.
+
+<!--
+
+Additionally, the package includes a series of pre-estimated impedance functions 
+for walking and cycling trips, considering various destinations and time periods, 
+for Canadian Metropolitan and Census Agglomeration areas.
+
+-->
+
+The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
+which facilitates the addition of complementary data sets in the future. `R` users
+can seamlessly explore GSS walking and cycling episodes <!-- along with calibrated impedance functions-->,
+with the option to suggest enhancements to the package as needed. This article adopts
+the structure proposed by Anastasia and Páez [-@soukhov2023], whose work provided
+essential guidance for the creation of this package. Similarly, we aim to contribute
+to the academic community by promoting transparent research practices that encourage
+replication and innovation in related fields. We believe that {ActiveCA} will serve
+as a basis for further research on GSS and for the integration of additional data by
 the authors or the wider open source community.
@@
 
-The {ActiceCA} R data package can be found and installed on Github
+The {ActiveCA} R data package can be found and installed on Github
 (*link*).

### 2024-12-04 paper/Manuscript-Data-Package.tex
Started editing the paper after Bruno's work

@@
 \begin{abstract}
-This paper describes the open data product \texttt{ActiveCA}, with
-Canadian time use data about active travel. \texttt{ActiveCA} is an
-\texttt{R} data package that contains analysis-ready data from cycles
-General Social Survey cicles regarding active travel in Canada. The
-package provides data set detailing active trip origins, destinations,
-and duration, covering a wide variety of locations, such as home, work
-or school, libraries, museums, restaurants, bars, sports centers, health
-clinics, place of worship, and others. The package also details the
-respondent's region characteristics, specifying wheter they live in a
-metropolitan area and their province of residency. \texttt{ActiveCA}
-also provides pre-estimated impedance functions for active travel for
-Canadian Metropolitan Area and Census Agglomerations, which can be used
-in accessibility analysis to calculate the cost of travel between
-different locations. The package will continue to expand with
-contributions from the authors and the broader community through
-requests in the future. \texttt{ActiveCA} is freely accessible for
-exploration and download from the associated Github repository, where
-the documentation and code involved in creating and manipulating data
-are detailed.
+This paper describes \{ActiveCA\}, and open data product with Canadian
+time use data. \{ActiveCA\} is an \texttt{R} data package that contains
+analysis-ready data related to active travel spanning almost 40 years,
+extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
+(2010), and 29 (2015) of Canada's General Social Survey. Active travel
+is characterized by mode, with walking being part of every cycle and
+bicycling starting in 1992. The attributes of active trips are the types
+of locations of origins and destinations, the duration of trips, and
+episode weights for expanding the trips to population-wide estimates.
+Based on the year of the survey, a variety of locations are coded. In
+earlier cycles, these include home, work or school, and other's home,
+whereas in later cycles these are augmented with locations such as
+grocery stores, restaurants, outdoor destinations, and others. The
+geographical resolution includes the province and whether the episode
+was in an urban or rural setting.
 \end{abstract}
 
-\keywords{Active; mobility; walking; cycling; travel time; impedance;
-transportation; R}
+\keywords{Active; mobility; walking; cycling; travel time; time-use;
+General Social Survey}
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
-This paper presents the open data product \texttt{ActiveCA}. Open data
-products (ODPs) are the outcome of a transparent process that transforms
-raw data (open or not) into analysis-ready data, in which all stages of
-development follow open principles \citep{arribas-bel2021}. ODPs differ
-from general open data due to their high utility, added value and open
-availability.
+The objective of this paper is to introduce \{ActiveCA\}, an open data
+product with time use data from Canada's General Social Surveys. Open
+data products (ODPs) are the outcome of a process that transforms raw
+data (open or not) into analysis-ready data, following a transparent
+process in which all stages of development follow open principles
+\citep{arribas-bel2021}. ODPs, while still open, differ from general
+open data in their degree of ease of access, their heightened usability,
+and potentially the value they add to the raw data.
 
-\texttt{ActiveCA} is an \texttt{R} data package that provides
-analysis-ready data from the General Social Survey of Canada Cycles 2
-(1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015)
-\citep{statisticscanada2024} regarding active travel. These data can be
-used to study walking and cycling in Canada for various purposes over
-time, and to calculate impedance functions for accessibility analysis.
-The package provides pre-estimated impedance functions for cycling and
-walking trips for Census Agglomerations and Census Metropolitan Areas.
+\{ActiveCA\} provides analysis-ready data concerning active travel in
+Canada spanning a period of almost 40 years. The source of these data is
+Canada's program of General Social Surveys (GSSs). This program is
+designed to provide cross-sectional data on topics of interest to
+improve the well-being of Canadians. As part of this program, every five
+to seven years the survey is done on the topic of time use. Concretely,
+\{ActiveCA\} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
+(2010), and 29 (2015) of the GSS. Time use data in these surveys is
+coded using a very fine grain, from time spent in chores, leisure, and
+sleeping, to time spent working or at school. These surveys have proved
+valuable in investigations of mobility and quality of life
+\citep{spinneyTransport2009}, the relationship between active travel and
+transit use \citep{lachapelleLonger2016}, and travel behavior and time
+poverty \citep{kimFacing2024}, to name but a few examples.
 
-To create this \texttt{R} data package, we collected, cleaned and
-processed the Time Use collections from the GSS surveys to make them
-ready for analysis. An \texttt{R} data package contains code, data and
-documentation in a standardized format that can be installed by
-\texttt{R} users via a centralized software repository, such as CRAN
-(Comprehensive R Archive Network) and GitHub. Although the GSS surveys
-are publicly sourced managed by Statistic Canada
-\citep{statisticscanada2024}, preparing them for analysis can be
-time-consuming, tedious and perhaps not even possible for those who try,
-due to a lack of documentation or prior knowledge.
+For \{ActiveCA\}, data we extracted all data needed to characterize
+active travel in Canada, namely, episodes where the activity was
+identified as moving between an origin and a destination, either by
+walking or cycling. Although Statistics Canada offers Public Use
+Microdata Files and documentation for the GSS program
+\citep[see][]{statisticscanada2024}, accessing these files, and
+preparing them for analysis is not a straightforward matter, given their
+size and complexity. The process of extracting information of interest
+from the source files is time-consuming, tedious, and challenging and/or
+prone to error due to the expertise required to work with these files.
+To create \{ActiveCA\} we collected, cleaned, and processed the cycles
+of the GSS surveys concerning time use to make them ready for analysis.
 
-The aim of this paper is to walk readers through the data sets and
-invite others to experiment in its uses and applications.
-\texttt{ActiveCA} is freely available on GitHub for all to install and
-freely use in the spirit of open and reproducible research. In addition
-of enabling to obtain impedance functions, \texttt{ActiveCA} can be
-adopted in various applications that even go beyond the range of
-possibilities we have imagined. Not only the data, but also all the code
-documenting the processing methodology is available for consultation and
-evaluation in its repository. This package contributes to reducing the
-barrier to using the information contained in GSS surveys to provide
-data-driven decisions in transportation analysis.
+\{ActiveCA\} is distributed as an \texttt{R} package with a number of
+data objects and their documentation. \texttt{R} packages contain code,
+data, and documentation in a standardized format that can be installed
+by \texttt{R} users via a software repository, such as CRAN
+(Comprehensive R Archive Network) or GitHub, which makes them an adroit
+medium to distribute analysis-ready data.
 
-\hypertarget{general-social-survey-gss-collection}{%
+Given the level of interest in active travel
+\citep[e.g.,][]{mccurdySupport2023}, reducing the barriers to using data
+contained in rich, but difficult to access and use surveys, such as
+Canada's GSSs, is a worthy endeavour that can only improve data-driven
+decisions in transportation, urban, and health policy.
+
+The rest of this paper discusses the sources of data, and the process
+implemented to retrieve and package them. Then, we explain how to use
+the package and show some selected examples of analysis to whet the
+imagination of potential users. This ODP provides not only data that are
+easy to use, but also all the code and documentation that make this a
+reproducible research project. In summary, \{ActiveCA\} aims to
+implement and inspire the best principles of open spatial sciences
+\citep{paez_open_2021, brunsdon_opening_2021}.
+
 \section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}}
+collection}\label{general-social-survey-gss-collection}
 
@@
 surveys to obtain data on social trends to track changes in Canadians'
-living conditions and well-being over time. This survey is used to
-understand how citizens spend and manage their time and what factors
-contribute to their happiness and stress. Created in 1985, the survey is
-part of a series of independent, annual, and cross-sectional surveys.
+living conditions and well-being over time. The series of survey on time
+use are used to understand how Canadian residents spend and manage their
+time, and what factors contribute to their happiness and stress. The GSS
+program was created in 1985, and is serialized to provide a collection
+of annual, representative cross-sectional surveys.
 
-In addition to the main topic, each GSS cycle includes new content that
-addresses emerging and policy-relevant issues. Every five to seven
-years, the Time Use Surveys \citep{statisticscanada2022} collect data on
-respondents' participation and time spent on a wide range of everyday
+The topics of the survey cycle every few years to cover topics that
+include family, health, social identity, and every five to seven years
+time use. The first Canadian time use survey done as part of the GSS
+program was conducted in 1986, and the most recent was completed in
+2015. These Time Use Surveys \citep{statisticscanada2022} collect data
+on respondents' participation and time spent on a wide range of everyday
 activities using a 24-hour retrospective diary, with information on the
@@
 
-The most recent time use survey was carried out in 2022
-\citep{wray2024}. However, the 2022 dataset has not been fully published
-and, because of this, our analysis focused on the surveys from 1986 to
-2015 (1986, 1992, 1998, 2005, 2010 and 2015). Time Use surveys are
-composed of two data sets, the main one and the episode file, explained
-in the following subsections.
+The Public Use Microdata are released by Statistics Canada in two files:
+a \emph{Main File} and an \emph{Episode File}. The files are linked by
+keys that identify households, individuals, and episodes (i.e.,
+activities) conducted by individuals. We discuss these files in more
+detail in the following section.
 
-\hypertarget{the-main-file}{%
-\subsection{The Main File}\label{the-main-file}}
+\subsection{The Main File}\label{the-main-file}
 
-The main file compiles an large array of aggregated data, summarizing
-the answers to the questionnaire and derived variables that summarize
-the respondents' time use across different activities, locations, and
-social interactions. This file documents the time and duration that
-respondents allocate to each activity and location. The Main File
-provides a overview of daily routines and social dynamics, not focusing
-on individual activity episodes. Additionally, this file categorizes
-activities into bigger groups and subcategories, facilitating the data's
-analytical utility with additional metrics such as total transit time,
-duration spent with household members, and counts of activities and
-episodes.
+The main file of the time use survey compiles a large array of
+aggregated data, summarizing the answers to the questionnaire that
+describe households and individuals, as well as derived variables that
+summarize the respondents' use of time use across different activities,
+locations, and social interactions. This file documents the time and
+duration that respondents allocate to each activity and location. The
+Main File provides a overview of daily routines and social dynamics, not
+focusing on individual activity episodes. Additionally, this file
+categorizes activities into bigger groups and subcategories,
+facilitating the data's analytical utility with additional metrics such
+as total transit time, time spent with household members, and counts of
+activities and episodes.
 
-\hypertarget{the-episode-file}{%
-\subsection{The Episode File}\label{the-episode-file}}
+\subsection{The Episode File}\label{the-episode-file}
 
-The episode file records detailed data for each activity episode
-reported by respondents. The entry includes the start and end times,
-duration, location, and accompanying social context, informing when and
-where activities occurred and with whom. The file distinguishes itself
-by focusing on individual episodes rather than respondents, with the
-data structured around the numerous activity instances that compose a
-day of the respondent. Although respondent-specific characteristics are
-not included within the episode file, it is possible to link the main
-file and the episode file by using an identifiable variable present in
-both data sets.
+The episode is a much bigger file that records detailed data for each
+activity episode reported by respondents. Each episode represents a
+single activity and its duration, and the sum of all episodes throughout
+the day adds up to 24 hours. Each entry in this file includes the start
+and end times of the activity, the duration, location, and accompanying
+social context, informing when and where activities occurred and with
+whom. The focus of the Episode File is not on the characteristics of the
+respondents, but on the characteristics of the activities, and the data
+are structured around the numerous activity instances that compose a day
+of the respondent. Although respondent-specific characteristics are not
+included within the episode file, it is possible to link the main file
+and the episode file by using a key present in both the Main and Episode
+Files.
 
-\hypertarget{methodology}{%
-\section{Methodology}\label{methodology}}
+\section{Methods}\label{methods}
 
 For each selected cycle of the GSS surveys, we reviewed the episode
-files to identify cases with activities listed as walking or cycling,
-selecting the activities immediately before and after the mobility
-episode. After that, we labeled the code variables with their
-appropriate descriptions, identifying each origin and destination,
-transportation mode, as well the province and urban classification of
-the respondent's residency.
+files to identify episodes of movement that involved walking or cycling.
+This allowed us to select the activities immediately before and after
+the movement episode. After that, we labeled the code variables with
+their appropriate descriptions, identifying each origin and destination,
+mode of travel, as well the province and urban classification of the
+episode.
 
-Additionally, for Census Agglomerations and Census Metropolitan Areas we
-pre-estimated impedance functions for cycling and walking trips.
-Impedance functions reveals important information about the travel
-behavior of the population, by describing the relationship between the
-population at an origin and their likelihood or ability to travel to
-specific destinations to access opportunities \citep{soukhov2024}.
+\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}
 
-Impedance functions are commonly used in accessibility analysis to
-calculate the cost of travel between different locations
-\citep{hansen1959, páez2012, palacios2022}. However, these functions
-need to be calibrated in order to accurately represent travel behavior.
-One effective way of calibrating an impedance function is by using the
-trip length distribution (TLD) obtained from origin-destination data
-\citep{soukhov2024}. The TLD represents the probability that a certain
-proportion of trips will be made at a specific cost, such as travel
-time. In this data set, low travel times are associated with a higher
-proportion of trips, while high travel times are associated with a lower
-proportion of trips.
-
-For each combination of year, destination, and transportation mode
-(walking or cycling), we fitted the most suitable impedance function
-based on empirical data from the GSS surveys. We used the
-\texttt{fitdistrplus} package \citep{mullerdutang2015} to estimate the
-functions, selecting the distribution with the lowest Akaike information
-criterion (AIC) among exponential, gamma, log-normal, normal, and
-uniform types.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{descriptive-statistics}{%
-\subsection{Descriptive statistics}\label{descriptive-statistics}}
+\subsection{Descriptive statistics}\label{descriptive-statistics}
 
@@
 
-\texttt{ActiveCA} also enables visual analysis of active travel in
-Canada using traditional exploratory data analysis techniques. Figures
+\{ActiveCA\} also enables visual analysis of active travel in Canada
+using traditional exploratory data analysis techniques. Figures
 \ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling
@@
 and destinations, with darker colors indicating higher percentages and
-lighter colors representing less frequent routes. To avoid overwhelming
-the reader, we omitted the heat maps for the other years analyzed.
+lighter colors representing less frequent routes. For conciseness, we
+omitted the heat maps for the other years analyzed.
 
@@
 
-\hypertarget{pre-estimated-impedance-functions-for-active-travel-in-canada}{%
-\subsection{Pre-estimated impedance functions for active travel in
-Canada}\label{pre-estimated-impedance-functions-for-active-travel-in-canada}}
+\section{Python integration}\label{python-integration}
 
-\texttt{ActiveCA} includes a total of 64 pre-estimated impedance
-functions for cycling and walking trips for Census Agglomerations and
-Census Metropolitan Areas. Table \ref{tab:table-04} shows the best
-impedance functions for walking trips to \texttt{Work\ or\ school}
-across survey years, where all distributions, except for a gamma
-function in 1998, are log-normal.
+CAN WE PROVIDE PYTHON INTEGRATION? See here:
 
-\begin{table}
-\centering
-\caption{\label{tab:table-04}\label{table-04}Impedance functions and AIC for `Walking` trips considering `Work or school` as destination.}
-\caption{\label{tab:table-04}\label{table-04}Impedance functions and AIC for `Cycling` trips considering `Work or school` as destination.}
-\centering
-\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
-\fontsize{10}{12}\selectfont
-\begin{threeparttable}
-\begin{tabular}[t]{lrlrrrr}
-\toprule
-\multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Year}} & \multicolumn{1}{c}{\textbf{Impedance function}} & \multicolumn{1}{c}{\textbf{Parameter 1*}} & \multicolumn{1}{c}{\textbf{Parameter 2*}} & \multicolumn{1}{c}{\textbf{AIC}} & \multicolumn{1}{c}{\textbf{Count}}\\
-\midrule
- & 1992 & Gamma & 3.00 & 0.17 & 433582 & 19\\
+\url{https://github.com/paezha/idealista18/tree/master/Python}
 
- & 1998 & Gamma & 3.37 & 0.10 & 481536 & 19\\
+\section{Concluding remarks}\label{concluding-remarks}
 
- & 2005 & Lognormal & 2.93 & 0.70 & 888655 & 64\\
-
- & 2010 & Lognormal & 2.65 & 0.77 & 1292760 & 53\\
-
-\multirow[t]{-5}{*}{\raggedright\arraybackslash Cycling} & 2015 & Lognormal & 3.03 & 0.41 & 1162876 & 63\\
-\cmidrule{1-7}
- & 1992 & Lognormal & 2.38 & 0.70 & 2319400 & 113\\
-
- & 1998 & Gamma & 1.23 & 0.09 & 2318752 & 109\\
-
- & 2005 & Lognormal & 2.13 & 0.79 & 8182691 & 724\\
-
- & 2010 & Lognormal & 2.21 & 0.78 & 7917431 & 494\\
-
-\multirow[t]{-5}{*}{\raggedright\arraybackslash Walking} & 2015 & Lognormal & 2.55 & 0.64 & 6612061 & 407\\
-\bottomrule
-\end{tabular}
-\begin{tablenotes}
-\item \textit{Note: } 
-\item *For lognormal distributions, 'Parameter 1' and 'Parameter 2' refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
-For gamma distribution, 'Parameter 1' and 'Parameter 2' refer to the rate and shape of the distribution, respectively.
-`AIC` means Akaike information criterion.
-\end{tablenotes}
-\end{threeparttable}}
-\end{table}
-
-Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
-alongside histograms of empirical travel times (grey bars). As for the
-others pre-estimated functions, these examples allow to calculate
-gravity-based accessibility measures for active travel across various
-destinations and temporal scales in Canadian urban areas.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/impf_Work or school}
-
-}
-
-\caption{Empirical data and impedance functions fitted for `work or school` as destination.}\label{fig:figure-03}
-\end{figure}
-
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
-
-This article presents \texttt{ActiveCA}, an open data product that
+This paper presents \texttt{ActiveCA}, an open data product that
 provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998),
 19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel
-in Canada. In the form of an \texttt{R} data package, \texttt{ActiveCA}
-was developed after collecting, cleaning, and processing the survey
-data, providing information on origins, destinations, and duration of
-active travel, as well other information. Additionally, the package
-includes a series of pre-estimated impedance functions for walking and
-cycling trips, considering various destinations and time periods, for
-Canadian Metropolitan and Census Agglomeration areas.
+in Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
+developed after collecting, cleaning, and processing the survey data,
+providing information on origins, destinations, and duration of active
+travel, as well other information.
 
-The value of \texttt{ActiveCA} lies in its transparency, accessibility,
-and open infrastructure, which facilitates the addition of complementary
-data sets in the future. R users can seamlessly explore GSS walking and
-cycling episodes along with calibrated impedance functions, with the
-option to suggest enhancements to the package as needed. This article
-adopts the structure proposed by Anastasia and Páez
-\citeyearpar{soukhov2023}, whose work provided essential guidance for
-the creation of this package. Similarly, we aim to contribute to the
-academic community by promoting transparent research practices that
-encourage replication and innovation in related fields. We believe that
-\texttt{ActiveCA} will serve as a basis for further research on GSS and
-for the integration of additional data by the authors or the wider open
-source community.
+It is important to remark that the present version of \{ActiveCA\}
+covers all Canadian time use surveys up to 2015. While the most recent
+time use survey was carried out in 2022 \citep{wray2024}, the Public Use
+Microdata Files are currently unavailable, and at the moment it is
+estimated that they will only be published in the later part of 2025.
+The \texttt{R} package will be updated once these files become
+available.
 
-\hypertarget{declaration-of-conflicting-interests}{%
+The value of \{ActiveCA\} lies in its transparency, accessibility, and
+ease of use, which facilitates the addition of complementary data sets
+in the future. \texttt{R} users can seamlessly explore GSS walking and
+cycling episodes , with the option to suggest enhancements to the
+package as needed. This article adopts the structure proposed by
+Anastasia and Páez \citeyearpar{soukhov2023}, whose work provided
+essential guidance for the creation of this package. Similarly, we aim
+to contribute to the academic community by promoting transparent
+research practices that encourage replication and innovation in related
+fields. We believe that \{ActiveCA\} will serve as a basis for further
+research on GSS and for the integration of additional data by the
+authors or the wider open source community.
+
 \section{Declaration of Conflicting
-Interests}\label{declaration-of-conflicting-interests}}
+Interests}\label{declaration-of-conflicting-interests}
 
@@
 
-\hypertarget{data-availability-statement}{%
-\section{Data availability
-statement}\label{data-availability-statement}}
+\section{Data availability statement}\label{data-availability-statement}
 
-The \{ActiceCA\} R data package can be found and installed on Github
+The \{ActiveCA\} R data package can be found and installed on Github
 (\emph{link}).

### 2024-12-04 R/ActiveCA-package.R
Revised documentation to accurately describe PUMFID

@@ -18,7 +18,7 @@ NULL
 #'
 #' @format A data frame containing 4347 rows and 17 variables; each row represents a unique episode.
 #'   \describe{
-#'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
+#'   \item{PUMFID}{Public Use Microdata File Identifier. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
 #'   \item{WGHT_EPI}{Episode weight. In 1986 the episode weight is the same as the summary weight, and is the number of persons that a record represents (see p. 18 of Cycle 2's Documentation and User's Guide)}
 #'   \item{ACTCODE}{Activity code of the episode.}
 #'   \item{act_label}{Activity (labeled).}
@@ -54,7 +54,7 @@ NULL
 #'
 #' @format A data frame containing 1500 rows and 17 variables; each row represents a unique episode.
 #'   \describe{
-#'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
+#'   \item{PUMFID}{Public Use Microdata File Identifier. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
 #'   \item{WGHT_EPI}{Episode weight. In 1992 the episode weight is the same as the summary weight, and is the number of persons that a record represents (see p. 9 of Cycle 7's Documentation and User's Guide)}
 #'   \item{ACTCODE}{Activity code of the episode.}
 #'   \item{act_label}{Activity (labeled).}
@@ -90,7 +90,7 @@ NULL
 #'
 #' @format A data frame containing 1670 rows and 17 variables; each row represents a unique episode.
 #'   \describe{
-#'   \item{PUMFID}{Uniquely identifies the episode. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
+#'   \item{PUMFID}{Public Use Microdata File Identifier. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
 #'   \item{WGHT_EPI}{Episode weight. In 1998 the episode weight indicates the number of time use episodes that a record on the Episode File represents (see p. 9 of Cycle 12's Documentation and User's Guide).}
[…]

### 2024-12-05 data-raw/01-preprocessing-episode-files.Rmd
preprocessing main files (2015)

--- 2024-11-17 data-raw/01-preprocessing-source-files.Rmd
@@
 ---
-title: "Reading source files and preparing data sets"
+title: "Reading episode files and preparing data sets"
 author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"

### 2024-12-05 data-raw/03-preprocessing-main-files.Rmd
preprocessing main files (2015)

---
title: "Reading main files and preparing data sets"
author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: activeca.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# layout configuration 
library(tufte)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(htmltools.dir.version = FALSE)
rm(list=ls())
```

# Introduction

This Rmarkdown file is part of the
[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. The main objective of the *ActiveCA* package is to identify the most
appropriate impedance functions for active transportation modes for
various destinations and time periods in Canada. To do achieve this objective, we analyzed historical data from the [General Social Survey
(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
from 1986 to 2015 in Canada to calculate the impedance function for
cycling and walking trips.

In this R markdown, we read the main files of GSS surveys and prepare the data set to
the following analyses.

# The General Social Surveys (GSS)

In Canada, Statistics Canada produces the General Social Surveys (GSS)
to collect data on social trends in order to track changes in the living
conditions of Canadians and well-being over time. This survey tracks
changes in time and it is used to understand how Canadians spend and
manage their time and what factors contribute to their happiness and
stress. The survey was created in 1985 as a series of independent,
annual, cross-sectional surveys, each covering one topic in depth.
Caregiving, families, time use, social identity, volunteering, and
victimization are among the current themes of the GSS.

The six survey themes listed above are repeated every five years. In
addition to the central topic, each cycle includes new content that
addresses emerging and policy-relevant issues, collecting detailed
socio-demographic information, such as age, gender, education, religion,
ethnicity, income, etc. Time-use surveys collect data on all human
activities and can inform a wide range of policies. In particular, three
key themes have been identified as essential for policymaking and for
which no other data source is adequate: unpaid work and non-market
production, well-being, and gender equality. Leisure time, work-life
balance, health, commuting, culture, and sports are among the other
topics covered by time-use surveys. In addition, statistics Canada has
been conducting time-use surveys at five to seven year intervals since
1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
and 2015). The time use collects information on participation of
respondents and time spent on a wide range of day-to-day activities
using a 24-hour retrospective diary. Also, information is collected on
the location of these activities (e.g., at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity.

The surveys (obtained from
[here](http://odesi2.scholarsportal.info/webview/)) consist of two
files: a main file and an episode file.

## The Main File

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, enhancing the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes. This file is aggregated and prohibits the direct association
between specific activities with particular locations or companions.

# Reading and processing the GSS files

In this R markdown, our aim is to generate a table in the appropriate
format to perform the subsequent analyses, after reading and processing
the GSS main surveys.

To start processing the data, first load the `R` packages used in this
notebook:

```{r read-libraries}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(here) # A Simpler Way to Find Your Files
library(readxl) # Read Excel Files
library(tidyr) # Tidy Messy Data
library(usethis) # Automate Package and Project Setup
library(purrr) # Provides a complete and consistent set of tools for working with functions and vectors
```

For each survey year, the process of preparing the data is very similar
and, in the end, we will obtain a table in which each row represents an
specific respondent, and each column a variable. Since every survey introduced different variables, 
we will not try to align the name columns and values between the tables.

The file name pattern is as follows: `gss_main_` +
`year of survey` [^4]. This means that, for instance, the file
`gss_main_20105` refers to the main file of the 2005 GSS survey.

## GSS 2015

Read the Main files of the 2015 GSS survey:

```{r reading-2015-main-file}
# Main File
gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
```

Select only the socioeconomic variables:

```{r filtering-variables-2015}
gss_main_2015 <- gss_m_2015 %>%
  select(PUMFID:DVTDAY)
```

Labeling socioeconomic variables:

```{r SURVMNTH}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SURVMNTH: Survey month of data collection from April 2015 to March 2016
    SURVMNTH = 
           case_when(SURVMNTH == 1 ~ "January", 
                     SURVMNTH == 2 ~ "February",
                     SURVMNTH == 3 ~ "March",
                     SURVMNTH == 4 ~ "April",
                     SURVMNTH == 5 ~ "May",
                     SURVMNTH == 6  ~ "June",
                     SURVMNTH == 7 ~ "July",
                     SURVMNTH == 8 ~ "August",
                     SURVMNTH == 9 ~ "September",
                     SURVMNTH == 10 ~ "October",
                     SURVMNTH == 11 ~ "November",
                     SURVMNTH == 12 ~ "Devember",
                     SURVMNTH == 96 ~ "Valid skip",
                     SURVMNTH == 97 ~ "Don't know",
                     SURVMNTH == 98 ~ "Refusal",
                     SURVMNTH == 99 ~ "Not stated"), 
    SURVMNTH = factor(SURVMNTH))

summary(gss_main_2015$SURVMNTH)
```

```{r AGEGR10}
values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGEGR10: Age group of respondent (groups of 10)
    AGEGR10 = 
           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
                     AGEGR10 == 2 ~ "25 to 34 years",
                     AGEGR10 == 3 ~ "35 to 44 years",
                     AGEGR10 == 4 ~ "45 to 54 years",
                     AGEGR10 == 5 ~ "55 to 64 years",
                     AGEGR10 == 6  ~ "65 to 74 years",
                     AGEGR10 == 7 ~ "75 years and over",
                     AGEGR10 == 96 ~ "Valid skip",
                     AGEGR10 == 97 ~ "Don't know",
                     AGEGR10 == 98 ~ "Refusal",
                     AGEGR10 == 99 ~ "Not stated"), 
    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$AGEGR10)
```

```{r SEX}
values_ordered <- c("Male", "Female")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SEX: Sex of respondent
    SEX = 
           case_when(SEX == 1 ~ "Male", 
                     SEX == 2 ~ "Female",
                     SEX == 96 ~ "Valid skip",
                     SEX == 97 ~ "Don't know",
                     SEX == 98 ~ "Refusal",
                     SEX == 99 ~ "Not stated"), 
    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$SEX)
```

```{r MARSTAT}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # MARSTAT: Marital status of the respondent
    MARSTAT = 
           case_when(MARSTAT == 1 ~ "Married", 
                     MARSTAT == 2 ~ "Living common-law",
                     MARSTAT == 3 ~ "Widowed",
                     MARSTAT == 4 ~ "Separated",
                     MARSTAT == 5 ~ "Divorced",
                     MARSTAT == 6  ~ "Single, never married",
                     MARSTAT == 96 ~ "Valid skip",
                     MARSTAT == 97 ~ "Don't know",
                     MARSTAT == 98 ~ "Refusal",
                     MARSTAT == 99 ~ "Not stated"), 
    MARSTAT = factor(MARSTAT))

summary(gss_main_2015$MARSTAT)
```

```{r AGEPRGRD}
values_ordered <- c("Respondent is 11 + years older", 
                    "Respondent is 6 - 10 years older",
                    "Respondent is 5 years older",
                    "Respondent is 4 years older",
                    "Respondent is 3 years older",
                    "Respondent is 2 years older",
                    "Respondent is 1 years older",
                    "Respondent and spouse/partner are the same age",
                    "Respondent is 1 years younger",
                    "Respondent is 2 years younger",
                    "Respondent is 3 years younger",
                    "Respondent is 4 years younger",
                    "Respondent is 5 years younger",
                    "Respondent is 6 - 10 years younger",
                    "Respondent is 11 + years younger",
                    "Valid skip")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGEPRGRD: Age difference between respondent and spouse/partner.
    AGEPRGRD = 
           case_when(AGEPRGRD == 1 ~ "Respondent is 11 + years older", 
                     AGEPRGRD == 2 ~ "Respondent is 6 - 10 years older",
                     AGEPRGRD == 3 ~ "Respondent is 5 years older",
                     AGEPRGRD == 4 ~ "Respondent is 4 years older",
                     AGEPRGRD == 5 ~ "Respondent is 3 years older",
                     AGEPRGRD == 6  ~ "Respondent is 2 years older",
                     AGEPRGRD == 7  ~ "Respondent is 1 years older",
                     AGEPRGRD == 8  ~ "Respondent and spouse/partner are the same age",
                     AGEPRGRD == 9  ~ "Respondent is 1 years younger",
                     AGEPRGRD == 10  ~ "Respondent is 2 years younger",
                     AGEPRGRD == 11  ~ "Respondent is 3 years younger",
                     AGEPRGRD == 12  ~ "Respondent is 4 years younger",
                     AGEPRGRD == 13  ~ "Respondent is 5 years younger",
                     AGEPRGRD == 14  ~ "Respondent is 6 - 10 years younger",
                     AGEPRGRD == 15  ~ "Respondent is 11 + years younger",
                     AGEPRGRD == 96 ~ "Valid skip",
                     AGEPRGRD == 97 ~ "Don't know",
                     AGEPRGRD == 98 ~ "Refusal",
                     AGEPRGRD == 99 ~ "Not stated"), 
    AGEPRGRD = factor(AGEPRGRD, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$AGEPRGRD)
```

```{r SEXPR}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SEXPR: Sex of respondent's spouse/partner living in the household.
    SEXPR = 
           case_when(SEXPR == 1 ~ "Male spouse/partner in the household", 
                     SEXPR == 2 ~ "Female spouse/partner in the household",
                     SEXPR == 6 ~ "Valid skip",
                     SEXPR == 7 ~ "Don't know",
                     SEXPR == 8 ~ "Refusal",
                     SEXPR == 9 ~ "Not stated"), 
    SEXPR = factor(SEXPR))

summary(gss_main_2015$SEXPR)
```

```{r PRTYPEC}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PRTYPEC: Type of partner respondent has living in the household.
    PRTYPEC = 
           case_when(PRTYPEC == 1 ~ "Opposite sex married partner in the household", 
                     PRTYPEC == 2 ~ "Opposite sex common-law partner in the household",
                     PRTYPEC == 3 ~ "Same sex married or common-law partner in the household",
                     PRTYPEC == 6 ~ "Valid skip",
                     PRTYPEC == 7 ~ "Don't know",
                     PRTYPEC == 8 ~ "Refusal",
                     PRTYPEC == 9 ~ "Not stated"), 
    PRTYPEC = factor(PRTYPEC))

summary(gss_main_2015$PRTYPEC)
```

```{r CXRFLAG}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXRFLAG:  Child(ren) of the respondent living in the household.
    CXRFLAG = 
           case_when(CXRFLAG == 1 ~ "Yes", 
                     CXRFLAG == 2 ~ "No",
                     CXRFLAG == 6 ~ "Valid skip",
                     CXRFLAG == 7 ~ "Don't know",
                     CXRFLAG == 8 ~ "Refusal",
                     CXRFLAG == 9 ~ "Not stated"), 
    CXRFLAG = factor(CXRFLAG))

summary(gss_main_2015$CXRFLAG)
```

```{r AGECXRYG}
values_ordered <- c( "Valid skip", "5 to 4", "5 to 9","10 to 14",
                     "15 to 19","20 to 24",
                     "25 years and older")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGECXRYG: Age of respondent's youngest child in household
    AGECXRYG = 
           case_when(AGECXRYG == 1 ~ "5 to 4", 
                     AGECXRYG == 2 ~ "5 to 9",
                     AGECXRYG == 3 ~ "10 to 14",
                     AGECXRYG == 4 ~ "15 to 19",
                     AGECXRYG == 5 ~ "20 to 24",
                     AGECXRYG == 6  ~ "25 years and older",
                     AGECXRYG == 96 ~ "Valid skip",
                     AGECXRYG == 97 ~ "Don't know",
                     AGECXRYG == 98 ~ "Refusal",
                     AGECXRYG == 99 ~ "Not stated"), 
    AGECXRYG = factor(AGECXRYG, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$AGECXRYG)
```
```{r CHINHSDC}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three",
                     "Four or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CHINHSDC: Respondent’s child(ren) in household - Any age/marital status
    CHINHSDC = 
           case_when(CHINHSDC == 0 ~ "None", 
                     CHINHSDC == 1 ~ "One",
                     CHINHSDC == 2 ~ "Two",
                     CHINHSDC == 3 ~ "Three",
                     CHINHSDC == 4 ~ "Four or more",
                     CHINHSDC == 6 ~ "Valid skip",
                     CHINHSDC == 7 ~ "Don't know",
                     CHINHSDC == 8 ~ "Refusal",
                     CHINHSDC == 9 ~ "Not stated"), 
    CHINHSDC = factor(CHINHSDC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CHINHSDC)
```

```{r CXR0014C}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXR0014C:  Respondent's child(ren) in household - 0 to 14 years
    CXR0014C = 
           case_when(CXR0014C == 0 ~ "None", 
                     CXR0014C == 1 ~ "One",
                     CXR0014C == 2 ~ "Two",
                     CXR0014C == 3 ~ "Three or more",
                     CXR0014C == 6 ~ "Valid skip",
                     CXR0014C == 7 ~ "Don't know",
                     CXR0014C == 8 ~ "Refusal",
                     CXR0014C == 9 ~ "Not stated"), 
    CXR0014C = factor(CXR0014C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CXR0014C)
```

```{r CHH0014C}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CHH0014C:  Child(ren) in household - 0 to 14 years
    CHH0014C = 
           case_when(CHH0014C == 0 ~ "None", 
                     CHH0014C == 1 ~ "One",
                     CHH0014C == 2 ~ "Two",
                     CHH0014C == 3 ~ "Three or more",
                     CHH0014C == 6 ~ "Valid skip",
                     CHH0014C == 7 ~ "Don't know",
                     CHH0014C == 8 ~ "Refusal",
                     CHH0014C == 9 ~ "Not stated"), 
    CHH0014C = factor(CHH0014C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CHH0014C)
```

```{r CXRTIME6}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXRTIME6:  Age group of respondent’s child(ren) in household
    CXRTIME6 = 
           case_when(CXRTIME6 == 1 ~ "No child under 19 years of age at home",
                     CXRTIME6 == 2 ~ "All children under 5 years of age",
                     CXRTIME6 == 3 ~ "All children between 5 and 12 years of age",
                     CXRTIME6 == 4 ~ "All children 13 years of age and older",
                     CXRTIME6 == 5 ~ "At least one child under 5 years of age but not all
children",
                     CXRTIME6 == 6 ~ "Other",
                     CXRTIME6 == 96 ~ "Valid skip",
                     CXRTIME6 == 97 ~ "Don't know",
                     CXRTIME6 == 98 ~ "Refusal",
                     CXRTIME6 == 99 ~ "Not stated"), 
    CXRTIME6 = factor(CXRTIME6))

summary(gss_main_2015$CXRTIME6)
```

```{r PARNUM}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PARNUM: Number of parents the respondent has in household
    PARNUM = 
           case_when(PARNUM == 0 ~ "None",
                     PARNUM == 1 ~ "One parent",
                     PARNUM == 2 ~ "Two parents",
                     PARNUM == 6 ~ "Valid skip",
                     PARNUM == 7 ~ "Don't know",
                     PARNUM == 8 ~ "Refusal",
                     PARNUM == 9 ~ "Not stated"), 
    PARNUM = factor(PARNUM))

summary(gss_main_2015$PARNUM)
```

```{r LIVARR08}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
    LIVARR08 = 
           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only",
                     LIVARR08 == 96 ~ "Valid skip",
                     LIVARR08 == 97 ~ "Don't know",
                     LIVARR08 == 98 ~ "Refusal",
                     LIVARR08 == 99 ~ "Not stated"), 
    LIVARR08 = factor(LIVARR08))

summary(gss_main_2015$LIVARR08)
```

```{r LIVARR11} 
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LIVARR11: Living arrangement of respondent’s household (11 categories)
    LIVARR11 = 
           case_when(LIVARR11 == 1 ~ "Living alone",
                     LIVARR11 == 2 ~ "Living with sponse only",
                     LIVARR11 == 3 ~ "Living with spouse and child(ren) < 25 years",
                     LIVARR11 == 4 ~ "Living with spouse and child(ren) 25 years or older",
                     LIVARR11 == 5 ~ "Living with spouse and other",
                     LIVARR11 == 6 ~ "Living without spouse but with chil(ren) < 25 years",
                     LIVARR11 == 7 ~ "Living without spouse but with chil(ren) 25 years and older",
                     LIVARR11 == 8 ~ "Living with two parents",
                     LIVARR11 == 9 ~ "Living with one parent",
                     LIVARR11 == 10 ~ "Multiple person household - Other living arrangement",
                     LIVARR11 == 11 ~ "Multiple person household - Relationships undefined",
                     LIVARR11 == 96 ~ "Valid skip",
                     LIVARR11 == 97 ~ "Don't know",
                     LIVARR11 == 98 ~ "Refusal",
                     LIVARR11 == 99 ~ "Not stated"), 
    LIVARR11 = factor(LIVARR11))

summary(gss_main_2015$LIVARR11)
```

```{r HSDSIZEC}
values_ordered <- c("One person household",
                   "Two person household",
                   "Three person household",
                   "Four person household",
                   "Five person household",
                   "Six or more person household")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # HSDSIZEC: Household size of respondent
    HSDSIZEC = 
           case_when(HSDSIZEC == 1 ~ "One person household",
                     HSDSIZEC == 2 ~ "Two person household",
                     HSDSIZEC == 3 ~ "Three person household",
                     HSDSIZEC == 4 ~ "Four person household",
                     HSDSIZEC == 5 ~ "Five person household",
                     HSDSIZEC == 6 ~ "Six or more person household",
                     HSDSIZEC == 96 ~ "Valid skip",
                     HSDSIZEC == 97 ~ "Don't know",
                     HSDSIZEC == 98 ~ "Refusal",
                     HSDSIZEC == 99 ~ "Not stated"), 
    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$HSDSIZEC)
```

```{r HSDELIGC}
values_ordered <- c("One member",
                     "Two members",
                     "Three members",
                     "Four or more members")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # HSDELIGC: Number of members in respondent’s household 15 years of age or older
    HSDELIGC = 
           case_when(HSDELIGC == 1 ~ "One member",
                     HSDELIGC == 2 ~ "Two members",
                     HSDELIGC == 3 ~ "Three members",
                     HSDELIGC == 4 ~ "Four or more members",
                     HSDELIGC == 6 ~ "Valid skip",
                     HSDELIGC == 7 ~ "Don't know",
                     HSDELIGC == 8 ~ "Refusal",
                     HSDELIGC == 9 ~ "Not stated"), 
    HSDELIGC = factor(HSDELIGC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$HSDELIGC)
```

```{r MULTIGEN}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
    MULTIGEN = 
           case_when(MULTIGEN == 1 ~ "Yes", 
                     MULTIGEN == 2 ~ "No",
                     MULTIGEN == 6 ~ "Valid skip",
                     MULTIGEN == 7 ~ "Don't know",
                     MULTIGEN == 8 ~ "Refusal",
                     MULTIGEN == 9 ~ "Not stated"), 
    MULTIGEN = factor(MULTIGEN))

summary(gss_main_2015$MULTIGEN)
```

```{r PRV}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PRV: Province of residence
    PRV = 
      case_when(PRV == 10 ~ "Newfoundland and Labrador",
                     PRV == 11 ~ "Prince Edward Island",
                     PRV == 12 ~ "Nova Scotia",
                     PRV == 13 ~ "New Brunswick",
                     PRV == 24 ~ "Quebec",
                     PRV == 35 ~ "Ontario",
                     PRV == 46 ~ "Manitoba",
                     PRV == 47 ~ "Saskatchewan",
                     PRV == 48 ~ "Alberta",
                     PRV == 59 ~ "British Columbia",
                     PRV == 96 ~ "Valid skip",
                     PRV == 97 ~ "Don't know",
                     PRV == 98 ~ "Refusal",
                     PRV == 99 ~ "Not stated"),
    PRV = factor(PRV))

summary(gss_main_2015$PRV) 
```

```{r LUC_RST}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LUC_RST:  Population centre indicator
    LUC_RST = 
      case_when(LUC_RST == 1 ~ "CMA/CA",
                     LUC_RST == 2 ~ "non CMA/CA",
                     LUC_RST == 3 ~ "Prince Edward Island",
                     LUC_RST == 6 ~ "Valid skip",
                     LUC_RST == 7 ~ "Don't know",
                     LUC_RST == 8 ~ "Refusal",
                     LUC_RST == 9 ~ "Not stated"),
    LUC_RST = factor(LUC_RST))

summary(gss_main_2015$LUC_RST)
```

```{r GTU_110}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # GTU_110: How often do you feel rushed? Would you say it is...?
    GTU_110 = 
      case_when(GTU_110 == 1 ~ "Every day",
                     GTU_110 == 2 ~ "A few times a week",
                     GTU_110 == 3 ~ "About once a week",
                     GTU_110 == 4 ~ "About once a month",
                     GTU_110 == 5 ~ "Less than once a month",
                     GTU_110 == 6 ~ "Never",
                     GTU_110 == 96 ~ "Valid skip",
                     GTU_110 == 97 ~ "Don't know",
                     GTU_110 == 98 ~ "Refusal",
                     GTU_110 == 99 ~ "Not stated"),
    GTU_110 = factor(GTU_110, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$GTU_110)
```

```{r GTU_130}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # GTU_130: How often do you feel you have time on your hands that you don’t know what to do with?
    GTU_130 = 
      case_when(GTU_130 == 1 ~ "Every day",
                     GTU_130 == 2 ~ "A few times a week",
                     GTU_130 == 3 ~ "About once a week",
                     GTU_130 == 4 ~ "About once a month",
                     GTU_130 == 5 ~ "Less than once a month",
                     GTU_130 == 6 ~ "Never",
                     GTU_130 == 96 ~ "Valid skip",
                     GTU_130 == 97 ~ "Don't know",
                     GTU_130 == 98 ~ "Refusal",
                     GTU_130 == 99 ~ "Not stated"),
    GTU_130 = factor(GTU_130, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$GTU_130)
```

```{r DVTDAY}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # DVTDAY: Diary - Reference day - Type
    DVTDAY = 
      case_when(DVTDAY == 1 ~ "Weekday",
                     DVTDAY == 2 ~ "Saturday",
                     DVTDAY == 3 ~ "Sunday",
                     DVTDAY == 96 ~ "Valid skip",
                     DVTDAY == 97 ~ "Don't know",
                     DVTDAY == 98 ~ "Refusal",
                     DVTDAY == 99 ~ "Not stated"),
    DVTDAY = factor(DVTDAY))

summary(gss_main_2015$DVTDAY)
```
Save data in R format for later use:
```{r}
usethis::use_data(cycling_2015, 
                  overwrite = TRUE)
```

## GSS 2010

Read the Main files of the 2010 GSS survey:

```{r reading-2010-main-file}
# Main File
gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv"))
```

Select only the socioeconomic variables:

```{r filtering-variables-2010}
gss_main_2010 <- gss_m_2010 %>%
  select(RECID:DVTDAY)
```

### 2024-12-05 data-raw/03-preprocessing-main-files.Rmd
preprocessing main files (2010)

@@
 ```
+
 ```{r CHINHSDC}
@@
 gss_main_2010 <- gss_m_2010 %>%
-  select(RECID:DVTDAY)
+  select(RECID:WGHT_PER, AGEGR5:DVTDAY)
 ```
+
+```{r AGEGR5-2010}
+values_ordered <- c("15 to 17 years","18 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # AGEGR5: Age group of respondent (groups of 5)
+    AGEGR5 = 
+           case_when(AGEGR5 == 1 ~ "15 to 17 years",
+                     AGEGR5 == 2 ~ "18 to 19 years",
+                     AGEGR5 == 3 ~ "20 to 24 years",
+                     AGEGR5 == 4 ~ "25 to 29 years",
+                     AGEGR5 == 5 ~ "30 to 34 years",
+                     AGEGR5 == 6  ~ "35 to 39 year",
+                     AGEGR5 == 7 ~ "40 to 44 years",
+                     AGEGR5 == 8 ~ "45 to 49 years",
+                     AGEGR5 == 9 ~ "50 to 54 years",
+                     AGEGR5 == 10 ~ "55 to 59 year",
+                     AGEGR5 == 11 ~ "60 to 64 years",
+                     AGEGR5 == 12 ~ "65 to 69 years",
+                     AGEGR5 == 13 ~ "70 to 74 years",
+                     AGEGR5 == 14 ~ "75 to 79 year",
+                     AGEGR5 == 15 ~ "80 years and over"), 
+    AGEGR5 = factor(AGEGR5, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$AGEGR5)
+```
+
+```{r AGEGR10-2010}
+values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # AGEGR10: Age group of respondent (groups of 10)
+    AGEGR10 = 
+           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
+                     AGEGR10 == 2 ~ "25 to 34 years",
+                     AGEGR10 == 3 ~ "35 to 44 years",
+                     AGEGR10 == 4 ~ "45 to 54 years",
+                     AGEGR10 == 5 ~ "55 to 64 years",
+                     AGEGR10 == 6  ~ "65 to 74 years",
+                     AGEGR10 == 7 ~ "75 years and over"), 
+    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$AGEGR10)
+```
+
+```{r SEX-2010}
+values_ordered <- c("Male", "Female")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # SEX: Sex of respondent
+    SEX = 
+           case_when(SEX == 1 ~ "Male", 
+                     SEX == 2 ~ "Female"), 
+    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$SEX)
+```
+
+```{r MARSTAT}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # MARSTAT: Marital status of the respondent
+    MARSTAT = 
+           case_when(MARSTAT == 1 ~ "Married", 
+                     MARSTAT == 2 ~ "Living common-law",
+                     MARSTAT == 3 ~ "Widowed",
+                     MARSTAT == 4 ~ "Separated",
+                     MARSTAT == 5 ~ "Divorced",
+                     MARSTAT == 6  ~ "Single, never married",
+                     MARSTAT == 8 ~ "Don't know",
+                     MARSTAT == 9 ~ "Not stated"), 
+    MARSTAT = factor(MARSTAT))
+
+summary(gss_main_2010$MARSTAT)
+```
+
+```{r AGEPRGRDIF-2010}
+values_ordered <- c("Respondent is 11 + years older", 
+                    "Respondent is 6 - 10 years older",
+                    "Respondent is 5 years older",
+                    "Respondent is 4 years older",
+                    "Respondent is 3 years older",
+                    "Respondent is 2 years older",
+                    "Respondent is 1 years older",
+                    "Respondent and spouse/partner are the same age",
+                    "Respondent is 1 years younger",
+                    "Respondent is 2 years younger",
+                    "Respondent is 3 years younger",
+                    "Respondent is 4 years younger",
+                    "Respondent is 5 years younger",
+                    "Respondent is 6 - 10 years younger",
+                    "Respondent is 11 + years younger",
+                    "No spouse/partner in household")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # AGEPRGRDIF: Age difference between respondent and spouse/partner.
+    AGEPRGRDIF = 
+           case_when(AGEPRGRDIF == 1 ~ "Respondent is 11 + years older", 
+                     AGEPRGRDIF == 2 ~ "Respondent is 6 - 10 years older",
+                     AGEPRGRDIF == 3 ~ "Respondent is 5 years older",
+                     AGEPRGRDIF == 4 ~ "Respondent is 4 years older",
+                     AGEPRGRDIF == 5 ~ "Respondent is 3 years older",
+                     AGEPRGRDIF == 6  ~ "Respondent is 2 years older",
+                     AGEPRGRDIF == 7  ~ "Respondent is 1 years older",
+                     AGEPRGRDIF == 8  ~ "Respondent and spouse/partner are the same age",
+                     AGEPRGRDIF == 9  ~ "Respondent is 1 years younger",
+                     AGEPRGRDIF == 10  ~ "Respondent is 2 years younger",
+                     AGEPRGRDIF == 11  ~ "Respondent is 3 years younger",
+                     AGEPRGRDIF == 12  ~ "Respondent is 4 years younger",
+                     AGEPRGRDIF == 13  ~ "Respondent is 5 years younger",
+                     AGEPRGRDIF == 14  ~ "Respondent is 6 - 10 years younger",
+                     AGEPRGRDIF == 15  ~ "Respondent is 11 + years younger",
+                     AGEPRGRDIF == 97 ~ "No spouse/partner in household"), 
+    AGEPRGRDIF = factor(AGEPRGRDIF, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2010$AGEPRGRDIF)
+```
+
+```{r SEXPR-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # SEXPR: Sex of respondent's spouse/partner living in the household.
+    SEXPR = 
+           case_when(SEXPR == 1 ~ "Male spouse/partner in the household", 
+                     SEXPR == 2 ~ "Female spouse/partner in the household",
+                     SEXPR == 0 ~ "Respondent has no partner in the household"), 
+    SEXPR = factor(SEXPR))
+
+summary(gss_main_2010$SEXPR)
+```
+```{r PRTYPEC-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # PRTYPEC: Type of partner respondent has living in the household.
+    PRTYPEC = 
+           case_when(PRTYPEC == 1 ~ "Opposite sex married partner in the household", 
+                     PRTYPEC == 2 ~ "Opposite sex common-law partner in the household",
+                     PRTYPEC == 3 ~ "Same sex married or common-law partner in the household",
+                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the
+household "), 
+    PRTYPEC = factor(PRTYPEC))
+
+summary(gss_main_2010$PRTYPEC)
+```
+```{r AGECHRYC}
+
+values_ordered <- c(as.character(0:24),"25 and over","Not asked - no children in the household")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # AGECHRYC: Age of respondent's youngest child in household
+           AGECHRYC = case_when(AGECHRYC == 25 ~ "25 and over",
+                     AGECHRYC == 97 ~ "Not asked - no children in the household",
+                     TRUE ~ as.character(AGECHRYC)), 
+    AGECHRYC = factor(AGECHRYC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$AGECHRYC)
+```
+
+```{r CHRFLAG-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHRFLAG:  single child(ren) of the respondent living in the household.
+    CHRFLAG = 
+           case_when(CHRFLAG == 1 ~ "Yes", 
+                     CHRFLAG == 2 ~ "No"), 
+    CHRFLAG = factor(CHRFLAG))
+
+summary(gss_main_2010$CHRFLAG)
+```
+```{r CHRINHSDC-2010}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three",
+                     "Four or more children")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHRINHSDC: Number of respondent's children living in the household (any age or marital status).
+    CHRINHSDC = 
+           case_when(CHRINHSDC == 0 ~ "None",
+                     CHRINHSDC == 1 ~ "One",
+                     CHRINHSDC == 2 ~ "Two",
+                     CHRINHSDC == 3 ~ "Three",
+                     CHRINHSDC == 4 ~ "Four or more children"), 
+    CHRINHSDC = factor(CHRINHSDC, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2010$CHRINHSDC)
+```
+
+```{r CHR0014C-2010}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHR0014C: Number of respondent's children living in the household (any age or marital status).
+    CHR0014C = 
+           case_when(CHR0014C == 0 ~ "None",
+                     CHR0014C == 1 ~ "One",
+                     CHR0014C == 2 ~ "Two",
+                     CHR0014C == 3 ~ "Three or more children"), 
+    CHR0014C = factor(CHR0014C, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2010$CHR0014C)
+```
+```{r CHRTIME6-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHRTIME6:  Age group of respondent’s child(ren) in household
+    CHRTIME6 = 
+           case_when(CHRTIME6 == 1 ~ "No child under 19 years of age at home",
+                     CHRTIME6 == 2 ~ "All children under 5 years of age",
+                     CHRTIME6 == 3 ~ "All children between 5 and 12 years of age",
+                     CHRTIME6 == 4 ~ "All children 13 years of age and older",
+                     CHRTIME6 == 5 ~ "At least one child under 5 years of age but not all children",
+                     CHRTIME6 == 6 ~ "Other"), 
+    CHRTIME6 = factor(CHRTIME6))
+
+summary(gss_main_2010$CHRTIME6)
+```
+```{r CHH0512C-2010}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHH0512C: Number of children aged from 5 to 12 years living in the respondent's household.
+    CHH0512C = 
+           case_when(CHH0512C == 0 ~ "None",
+                     CHH0512C == 1 ~ "One",
+                     CHH0512C == 2 ~ "Two",
+                     CHH0512C == 3 ~ "Three or more children"), 
+    CHH0512C = factor(CHH0512C, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2010$CHH0512C)
+```
+```{r CHH0014C-2010}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # CHH0014C: Number of children aged from 0 to 14 years living in the respondent's household.
+    CHH0014C = 
+           case_when(CHH0014C == 0 ~ "None",
+                     CHH0014C == 1 ~ "One",
+                     CHH0014C == 2 ~ "Two",
+                     CHH0014C == 3 ~ "Three or more children"), 
+    CHH0014C = factor(CHH0014C, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2010$CHH0014C)
+```
+
+```{r PARHSDC}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # PARHSDC: Type of parents the respondent has within the household.
+    PARHSDC = 
+           case_when(PARHSDC == 1 ~ "Two biological parents",
+                     PARHSDC == 2 ~ "Two parents (biological/adoptive/step)",
+                     PARHSDC == 3 ~ "Only one parent",
+                     PARHSDC == 7 ~ "Not asked - no parents in the household"), 
+    PARHSDC = factor(PARHSDC))
+
+summary(gss_main_2010$PARHSDC)
+```
+
+```{r LIVARR08}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
+    LIVARR08 = 
+           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
+                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
+                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
+                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
+                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
+                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
+                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
+                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only"), 
+    LIVARR08 = factor(LIVARR08))
+
+summary(gss_main_2010$LIVARR08)
+```
+
+```{r LIVARR12-2010} 
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # LIVARR12: Living arrangement of respondent’s household (12 categories)
+    LIVARR12 = 
+           case_when(LIVARR12 == 1 ~ "Living alone",
+                     LIVARR12 == 2 ~ "Living with sponse only",
+                     LIVARR12 == 3 ~ "Living with spouse and single child < 25 years",
+                     LIVARR12 == 4 ~ "Living with spouse and single child 25 years or older",
+                     LIVARR12 == 5 ~ "Living with spouse and non-child(ren)",
+                     LIVARR12 == 6 ~ "Living with spouse and other",
+                     LIVARR12 == 7 ~ "Living without spouse and single child  < 25 years",
+                     LIVARR12 == 8 ~ "Living without spouse and single child 25 years or older",
+                     LIVARR12 == 9 ~ "No spouse and non-single child(ren)",
+                     LIVARR12 == 10 ~ "Living with two parents",
+                     LIVARR12 == 11 ~ "Living with one parent",
+                     LIVARR12 == 12 ~ "Other living arrangement"), 
+    LIVARR12 = factor(LIVARR12))
+
+summary(gss_main_2010$LIVARR12)
+```
+
+```{r HSDSIZEC}
+values_ordered <- c("One person household",
+                   "Two person household",
+                   "Three person household",
+                   "Four person household",
+                   "Five person household",
+                   "Six or more person household")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # HSDSIZEC: Household size of respondent
+    HSDSIZEC = 
+           case_when(HSDSIZEC == 1 ~ "One person household",
+                     HSDSIZEC == 2 ~ "Two person household",
+                     HSDSIZEC == 3 ~ "Three person household",
+                     HSDSIZEC == 4 ~ "Four person household",
+                     HSDSIZEC == 5 ~ "Five person household",
+                     HSDSIZEC == 6 ~ "Six or more person household"), 
+    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$HSDSIZEC)
+```
+
+```{r HSDELIGC}
+values_ordered <- c("One member",
+                     "Two members",
+                     "Three members",
+                     "Four members", "Five household members or more")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # HSDELIGC: Number of members in respondent’s household 15 years of age or older
+    HSDELIGC = 
+           case_when(HSDELIGC == 1 ~ "One member",
+                     HSDELIGC == 2 ~ "Two members",
+                     HSDELIGC == 3 ~ "Three members",
+                     HSDELIGC == 4 ~ "Four members",
+                     HSDELIGC == 5 ~ "Five household members or more"), 
+    HSDELIGC = factor(HSDELIGC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2010$HSDELIGC)
+```
+```{r FAMTYPE-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # FAMTYPE: Respondent's type of family structure.
+    FAMTYPE = 
+           case_when(FAMTYPE == 1 ~ "Couple only", 
+                     FAMTYPE == 2 ~ "Intact family",
+                     FAMTYPE == 3 ~ "Step family with common child",
+                     FAMTYPE == 4 ~ "Step family without a common child",
+                     FAMTYPE == 5 ~ "Lone parent family",
+                     FAMTYPE == 7 ~ "No spouse/partner or children in household"), 
+    FAMTYPE = factor(FAMTYPE))
+
+summary(gss_main_2010$FAMTYPE)
+```
+
+```{r MULTIGEN}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
+    MULTIGEN = 
+           case_when(MULTIGEN == 1 ~ "Yes", 
+                     MULTIGEN == 2 ~ "No",
+                     MULTIGEN == 6 ~ "Number of generations unknown"), 
+    MULTIGEN = factor(MULTIGEN))
+
+summary(gss_main_2010$MULTIGEN)
+```
+
+```{r PRV-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # PRV: Province of residence
+    PRV = 
+      case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"),
+    PRV = factor(PRV))
+
+summary(gss_main_2010$PRV) 
+```
+```{r REGION-2010}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # REGION:  Region of residence of the respondent.
+    REGION = 
+      case_when(REGION == 1 ~ "Atlantic region",
+                     REGION == 2 ~ "Quebec",
+                     REGION == 3 ~ "Ontario",
+                     REGION == 4 ~ "Prairie region",
+                     REGION == 5 ~ "British Columbia"),
+    REGION = factor(REGION))
+
+summary(gss_main_2010$REGION)
+```
+
+```{r LUC_RST}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # LUC_RST:  Population centre indicator
+    LUC_RST = 
+      case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "Prince Edward Island"),
+    LUC_RST = factor(LUC_RST))
+
+summary(gss_main_2010$LUC_RST)
+```
+
+```{r GTU_110}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Don't know")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # GTU_Q110: How often do you feel rushed? Would you say it is...?
+    GTU_Q110 = 
+      case_when(GTU_Q110 == 1 ~ "Every day",
+                     GTU_Q110 == 2 ~ "A few times a week",
+                     GTU_Q110 == 3 ~ "About once a week",
+                     GTU_Q110 == 4 ~ "About once a month",
+                     GTU_Q110 == 5 ~ "Less than once a month",
+                     GTU_Q110 == 6 ~ "Never",
+                     GTU_Q110 == 9 ~ "Don't know"),
+    GTU_Q110 = factor(GTU_Q110, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2010$GTU_Q110)
+```
+
+```{r GTU_130}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Don't know")
+
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # GTU_Q130: How often do you feel you have time on your hands that you don’t know what to do with?
+    GTU_Q130 = 
+      case_when(GTU_Q130 == 1 ~ "Every day",
+                     GTU_Q130 == 2 ~ "A few times a week",
+                     GTU_Q130 == 3 ~ "About once a week",
+                     GTU_Q130 == 4 ~ "About once a month",
+                     GTU_Q130 == 5 ~ "Less than once a month",
+                     GTU_Q130 == 6 ~ "Never",
+                     GTU_Q130 == 9 ~ "Don't know"),
+    GTU_Q130 = factor(GTU_Q130, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2010$GTU_Q130)
+```
+
+```{r DVTDAY}
+gss_main_2010 <- gss_main_2010 %>% 
+  mutate( # DVTDAY: Diary - Reference day - Type
+    DVTDAY = 
+      case_when(DVTDAY == 1 ~ "Weekday",
+                     DVTDAY == 2 ~ "Saturday",
+                     DVTDAY == 3 ~ "Sunday"),
+    DVTDAY = factor(DVTDAY))
+
+summary(gss_main_2010$DVTDAY)
+```
+Save data in R format for later use:
+```{r}
+usethis::use_data(gss_main_2010, 
+                  overwrite = TRUE)
+```
+
+### Aqui

### 2024-12-10 data-raw/03-preprocessing-main-files.Rmd
Socioeconomic variables from the 2005 GSS survey.

@@
 ```
+
 Save data in R format for later use:
 ```{r}
-usethis::use_data(cycling_2015, 
+usethis::use_data(gss_main_2015, 
                   overwrite = TRUE)
@@
 gss_main_2010 <- gss_m_2010 %>%
-  select(RECID:WGHT_PER, AGEGR5:DVTDAY)
+  select(RECID:DVTDAY)
 ```
@@
                      PRTYPEC == 3 ~ "Same sex married or common-law partner in the household",
-                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the
-household "), 
+                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the household "), 
     PRTYPEC = factor(PRTYPEC))
@@
 ```
+
 ```{r AGECHRYC}
@@
 ```
+
 Save data in R format for later use:
@@
 
+## GSS 2005
+
+Read the Main files of the 2010 GSS survey:
+
+```{r reading-2010-main-file}
+# Main File
+gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv"))
+```
+
+Select only the socioeconomic variables:
+
+```{r filtering-variables-2010}
+gss_main_2005 <- gss_m_2005 %>%
+  select(RECID:DVTDAY)
+```
+
+```{r AGEGR5-2005}
+values_ordered <- c("15 to 17 years","18 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # AGEGR5: Age group of respondent (groups of 5)
+    AGEGR5 = 
+           case_when(AGEGR5 == 1 ~ "15 to 17 years",
+                     AGEGR5 == 2 ~ "18 to 19 years",
+                     AGEGR5 == 3 ~ "20 to 24 years",
+                     AGEGR5 == 4 ~ "25 to 29 years",
+                     AGEGR5 == 5 ~ "30 to 34 years",
+                     AGEGR5 == 6  ~ "35 to 39 year",
+                     AGEGR5 == 7 ~ "40 to 44 years",
+                     AGEGR5 == 8 ~ "45 to 49 years",
+                     AGEGR5 == 9 ~ "50 to 54 years",
+                     AGEGR5 == 10 ~ "55 to 59 year",
+                     AGEGR5 == 11 ~ "60 to 64 years",
+                     AGEGR5 == 12 ~ "65 to 69 years",
+                     AGEGR5 == 13 ~ "70 to 74 years",
+                     AGEGR5 == 14 ~ "75 to 79 year",
+                     AGEGR5 == 15 ~ "80 years and over"), 
+    AGEGR5 = factor(AGEGR5, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2005$AGEGR5)
+```
+
+```{r AGEGR10-2005}
+values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # AGEGR10: Age group of respondent (groups of 10)
+    AGEGR10 = 
+           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
+                     AGEGR10 == 2 ~ "25 to 34 years",
+                     AGEGR10 == 3 ~ "35 to 44 years",
+                     AGEGR10 == 4 ~ "45 to 54 years",
+                     AGEGR10 == 5 ~ "55 to 64 years",
+                     AGEGR10 == 6  ~ "65 to 74 years",
+                     AGEGR10 == 7 ~ "75 years and over"), 
+    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2005$AGEGR10)
+```
+
+```{r SEX-2005}
+values_ordered <- c("Male", "Female")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # SEX: Sex of respondent
+    SEX = 
+           case_when(SEX == 1 ~ "Male", 
+                     SEX == 2 ~ "Female"), 
+    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2005$SEX)
+```
+
+```{r MARSTAT}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # MARSTAT: Marital status of the respondent
+    MARSTAT = 
+           case_when(MARSTAT == 1 ~ "Married", 
+                     MARSTAT == 2 ~ "Living common-law",
+                     MARSTAT == 3 ~ "Widowed",
+                     MARSTAT == 4 ~ "Separated",
+                     MARSTAT == 5 ~ "Divorced",
+                     MARSTAT == 6  ~ "Single, never married",
+                     MARSTAT == 8 ~ "Don't know",
+                     MARSTAT == 9 ~ "Not stated"), 
+    MARSTAT = factor(MARSTAT))
+
+summary(gss_main_2005$MARSTAT)
+```
+
+```{r AGEPRGRDIF-2005}
+values_ordered <- c("Respondent is 11 + years older", 
+                    "Respondent is 6 - 10 years older",
+                    "Respondent is 5 years older",
+                    "Respondent is 4 years older",
+                    "Respondent is 3 years older",
+                    "Respondent is 2 years older",
+                    "Respondent is 1 years older",
+                    "Respondent and spouse/partner are the same age",
+                    "Respondent is 1 years younger",
+                    "Respondent is 2 years younger",
+                    "Respondent is 3 years younger",
+                    "Respondent is 4 years younger",
+                    "Respondent is 5 years younger",
+                    "Respondent is 6 - 10 years younger",
+                    "Respondent is 11 + years younger",
+                    "No spouse/partner in household")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # AGEPRGRDIF: Age difference between respondent and spouse/partner.
+    AGEPRGRDIF = 
+           case_when(AGEPRGRDIF == 1 ~ "Respondent is 11 + years older", 
+                     AGEPRGRDIF == 2 ~ "Respondent is 6 - 10 years older",
+                     AGEPRGRDIF == 3 ~ "Respondent is 5 years older",
+                     AGEPRGRDIF == 4 ~ "Respondent is 4 years older",
+                     AGEPRGRDIF == 5 ~ "Respondent is 3 years older",
+                     AGEPRGRDIF == 6  ~ "Respondent is 2 years older",
+                     AGEPRGRDIF == 7  ~ "Respondent is 1 years older",
+                     AGEPRGRDIF == 8  ~ "Respondent and spouse/partner are the same age",
+                     AGEPRGRDIF == 9  ~ "Respondent is 1 years younger",
+                     AGEPRGRDIF == 10  ~ "Respondent is 2 years younger",
+                     AGEPRGRDIF == 11  ~ "Respondent is 3 years younger",
+                     AGEPRGRDIF == 12  ~ "Respondent is 4 years younger",
+                     AGEPRGRDIF == 13  ~ "Respondent is 5 years younger",
+                     AGEPRGRDIF == 14  ~ "Respondent is 6 - 10 years younger",
+                     AGEPRGRDIF == 15  ~ "Respondent is 11 + years younger",
+                     AGEPRGRDIF == 97 ~ "No spouse/partner in household"), 
+    AGEPRGRDIF = factor(AGEPRGRDIF, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2005$AGEPRGRDIF)
+```
+
+```{r SEXPR-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # SEXPR: Sex of respondent's spouse/partner living in the household.
+    SEXPR = 
+           case_when(SEXPR == 1 ~ "Male spouse/partner in the household", 
+                     SEXPR == 2 ~ "Female spouse/partner in the household",
+                     SEXPR == 0 ~ "Respondent has no partner in the household"), 
+    SEXPR = factor(SEXPR))
+
+summary(gss_main_2005$SEXPR)
+```
+
+```{r PRTYPEC-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # PRTYPEC: Type of partner respondent has living in the household.
+    PRTYPEC = 
+           case_when(PRTYPEC == 1 ~ "Respondent has a married partner in the household", 
+                     PRTYPEC == 2 ~ "Respondent has a common-law partner in the household",
+                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the household"), 
+    PRTYPEC = factor(PRTYPEC))
+
+summary(gss_main_2005$PRTYPEC)
+```
+
+```{r AGECHRYC}
+
+values_ordered <- c(as.character(0:24),"25 and over","Not asked - no children in the household")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # AGECHRYC: Age of respondent's youngest child in household
+           AGECHRYC = case_when(AGECHRYC == 25 ~ "25 and over",
+                     AGECHRYC == 97 ~ "Not asked - no children in the household",
+                     TRUE ~ as.character(AGECHRYC)), 
+    AGECHRYC = factor(AGECHRYC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2005$AGECHRYC)
+```
+
+```{r CHRFLAG-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # CHRFLAG:  single child(ren) of the respondent living in the household.
+    CHRFLAG = 
+           case_when(CHRFLAG == 1 ~ "Yes", 
+                     CHRFLAG == 2 ~ "No"), 
+    CHRFLAG = factor(CHRFLAG))
+
+summary(gss_main_2005$CHRFLAG)
+```
+```{r CHRINHSDC-2005}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three",
+                     "Four or more children")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # CHRINHSDC: Number of respondent's children living in the household (any age or marital status).
+    CHRINHSDC = 
+           case_when(CHRINHSDC == 0 ~ "None",
+                     CHRINHSDC == 1 ~ "One",
+                     CHRINHSDC == 2 ~ "Two",
+                     CHRINHSDC == 3 ~ "Three",
+                     CHRINHSDC == 4 ~ "Four or more children"), 
+    CHRINHSDC = factor(CHRINHSDC, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2005$CHRINHSDC)
+```
+
+```{r CHR0014C-2005}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # CHR0014C: Number of respondent's children living in the household (any age or marital status).
+    CHR0014C = 
+           case_when(CHR0014C == 0 ~ "None",
+                     CHR0014C == 1 ~ "One",
+                     CHR0014C == 2 ~ "Two",
+                     CHR0014C == 3 ~ "Three or more children"), 
+    CHR0014C = factor(CHR0014C, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_2005$CHR0014C)
+```
+```{r CHRTIME6-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # CHRTIME6:  Age group of respondent’s child(ren) in household
+    CHRTIME6 = 
+           case_when(CHRTIME6 == 1 ~ "No child under 19 years of age at home",
+                     CHRTIME6 == 2 ~ "All children under 5 years of age",
+                     CHRTIME6 == 3 ~ "All children between 5 and 12 years of age",
+                     CHRTIME6 == 4 ~ "All children 13 years of age and older",
+                     CHRTIME6 == 5 ~ "At least one child under 5 years of age but not all children",
+                     CHRTIME6 == 6 ~ "Other"), 
+    CHRTIME6 = factor(CHRTIME6))
+
+summary(gss_main_2005$CHRTIME6)
+```
+
+```{r PARHSDC}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # PARHSDC: Type of parents the respondent has within the household.
+    PARHSDC = 
+           case_when(PARHSDC == 1 ~ "Two biological parents",
+                     PARHSDC == 2 ~ "Two parents (biological/adoptive/step)",
+                     PARHSDC == 3 ~ "Only one parent",
+                     PARHSDC == 7 ~ "Not asked - no parents in the household"), 
+    PARHSDC = factor(PARHSDC))
+
+summary(gss_main_2005$PARHSDC)
+```
+
+```{r LIVARR08}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
+    LIVARR08 = 
+           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
+                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
+                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
+                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
+                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
+                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
+                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
+                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only"), 
+    LIVARR08 = factor(LIVARR08))
+
+summary(gss_main_2005$LIVARR08)
+```
+
+```{r LIVARR12-2005} 
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # LIVARR12: Living arrangement of respondent’s household (12 categories)
+    LIVARR12 = 
+           case_when(LIVARR12 == 1 ~ "Living alone",
+                     LIVARR12 == 2 ~ "Living with sponse only",
+                     LIVARR12 == 3 ~ "Living with spouse and single child < 25 years",
+                     LIVARR12 == 4 ~ "Living with spouse and single child 25 years or older",
+                     LIVARR12 == 5 ~ "Living with spouse and non-child(ren)",
+                     LIVARR12 == 6 ~ "Living with spouse and other",
+                     LIVARR12 == 7 ~ "Living without spouse and single child  < 25 years",
+                     LIVARR12 == 8 ~ "Living without spouse and single child 25 years or older",
+                     LIVARR12 == 9 ~ "No spouse and non-single child(ren)",
+                     LIVARR12 == 10 ~ "Living with two parents",
+                     LIVARR12 == 11 ~ "Living with one parent",
+                     LIVARR12 == 12 ~ "Other living arrangement"), 
+    LIVARR12 = factor(LIVARR12))
+
+summary(gss_main_2005$LIVARR12)
+```
+
+```{r HSDSIZEC-2005}
+values_ordered <- c("One person household",
+                   "Two person household",
+                   "Three person household",
+                   "Four person household",
+                   "Five person household",
+                   "Six or more person household")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # HSDSIZEC: Household size of respondent
+    HSDSIZEC = 
+           case_when(HSDSIZEC == 1 ~ "One person household",
+                     HSDSIZEC == 2 ~ "Two person household",
+                     HSDSIZEC == 3 ~ "Three person household",
+                     HSDSIZEC == 4 ~ "Four person household",
+                     HSDSIZEC == 5 ~ "Five person household",
+                     HSDSIZEC == 6 ~ "Six or more person household"), 
+    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2005$HSDSIZEC)
+```
+
+```{r FAMTYPE-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # FAMTYPE: Respondent's type of family structure.
+    FAMTYPE = 
+           case_when(FAMTYPE == 1 ~ "Couple only", 
+                     FAMTYPE == 2 ~ "Intact family",
+                     FAMTYPE == 3 ~ "Step family with common child",
+                     FAMTYPE == 4 ~ "Step family without a common child",
+                     FAMTYPE == 5 ~ "Lone parent family",
+                     FAMTYPE == 7 ~ "No spouse/partner or children in household"), 
+    FAMTYPE = factor(FAMTYPE))
+
+summary(gss_main_2005$FAMTYPE)
+```
+
+```{r MULTIGEN}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
+    MULTIGEN = 
+           case_when(MULTIGEN == 1 ~ "Yes", 
+                     MULTIGEN == 2 ~ "No",
+                     MULTIGEN == 6 ~ "Number of generations unknown"), 
+    MULTIGEN = factor(MULTIGEN))
+
+summary(gss_main_2005$MULTIGEN)
+```
+
+```{r PRV-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # PRV: Province of residence
+    PRV = 
+      case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"),
+    PRV = factor(PRV))
+
+summary(gss_main_2005$PRV) 
+```
+```{r REGION-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # REGION:  Region of residence of the respondent.
+    REGION = 
+      case_when(REGION == 1 ~ "Atlantic region",
+                     REGION == 2 ~ "Quebec",
+                     REGION == 3 ~ "Ontario",
+                     REGION == 4 ~ "Prairie region",
+                     REGION == 5 ~ "British Columbia"),
+    REGION = factor(REGION))
+
+summary(gss_main_2005$REGION)
+```
+
+```{r LUC_RST}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # LUC_RST:  Population centre indicator
+    LUC_RST = 
+      case_when(LUC_RST == 1 ~ "CMA/CA",
+                     LUC_RST == 2 ~ "non CMA/CA",
+                     LUC_RST == 3 ~ "Prince Edward Island"),
+    LUC_RST = factor(LUC_RST))
+
+summary(gss_main_2005$LUC_RST)
+```
+
+```{r GTU_Q110-2005}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Don't know")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # GTU_Q110: How often do you feel rushed? Would you say it is...?
+    GTU_Q110 = 
+      case_when(GTU_Q110 == 1 ~ "Every day",
+                     GTU_Q110 == 2 ~ "A few times a week",
+                     GTU_Q110 == 3 ~ "About once a week",
+                     GTU_Q110 == 4 ~ "About once a month",
+                     GTU_Q110 == 5 ~ "Less than once a month",
+                     GTU_Q110 == 6 ~ "Never",
+                     GTU_Q110 == 9 ~ "Don't know"),
+    GTU_Q110 = factor(GTU_Q110, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2005$GTU_Q110)
+```
+```{r GTU_Q120-2005}
+values_ordered <- c("More rushed",
+                    "About the same",
+                    "Less rushed",
+                    "Don't know")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # GTU_Q120: How often do you feel you have time on your hands that you don’t know what to do with?
+    GTU_Q120 = 
+      case_when(GTU_Q120 == 1 ~ "More rushed",
+                     GTU_Q120 == 2 ~  "About the same",
+                     GTU_Q120 == 3 ~ "Less rushed",
+                     GTU_Q120 == 9 ~ "Don't know"),
+    GTU_Q120 = factor(GTU_Q120, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2005$GTU_Q120)
+```
+
+```{r GTU_130-2005}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Don't know")
+
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # GTU_Q130: How often do you feel you have time on your hands that you don’t know what to do with?
+    GTU_Q130 = 
+      case_when(GTU_Q130 == 1 ~ "Every day",
+                     GTU_Q130 == 2 ~ "A few times a week",
+                     GTU_Q130 == 3 ~ "About once a week",
+                     GTU_Q130 == 4 ~ "About once a month",
+                     GTU_Q130 == 5 ~ "Less than once a month",
+                     GTU_Q130 == 6 ~ "Never",
+                     GTU_Q130 == 9 ~ "Don't know"),
+    GTU_Q130 = factor(GTU_Q130, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2005$GTU_Q130)
+```
+```{r GTU_Q140-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # GTU_Q140:  Do you feel that the days are just too short to do all the things you want?
+    GTU_Q140 = 
+           case_when(GTU_Q140 == 1 ~ "Yes", 
+                     GTU_Q140 == 2 ~ "No",
+                     GTU_Q140 == 8 ~ "Not stated", 
+                     GTU_Q140 == 9 ~ "Don't know"), 
+    GTU_Q140 = factor(GTU_Q140))
+
+summary(gss_main_2005$GTU_Q140)
+```
+
+```{r GTU_Q150-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # GTU_Q150: Diary - Reference day - Type
+    GTU_Q150 = 
+      case_when(GTU_Q150 == 1 ~ "Time with family and friends (spouse, children, boy-friend - girl-friend)",
+                     GTU_Q150 == 2 ~ "Relaxation - personal time",
+                     GTU_Q150 == 3 ~ "Practising sports",
+                     GTU_Q150 == 4 ~ "Crafts or hobbies",
+                     GTU_Q150 == 5 ~ "Outdoor activities",
+                     GTU_Q150 == 6 ~ "Reading - writing",
+                     GTU_Q150 == 7 ~ "Studies",
+                     GTU_Q150 == 8 ~ "Work",
+                     GTU_Q150 == 9 ~ "Other",
+                     GTU_Q150 == 10 ~ "Travel",
+                     GTU_Q150 == 11 ~ "Volunteer",
+                     GTU_Q150 == 12  ~ "Nothing",
+                     GTU_Q150 == 13  ~ "Do household work",
+                     GTU_Q150 == 98 ~ "Not stated",
+                     GTU_Q150 == 99 ~ "Don't know",
+),
+    GTU_Q150 = factor(GTU_Q150))
+
+summary(gss_main_2005$GTU_Q150)
+```
+
+```{r DVTDAY}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # DVTDAY: Diary - Reference day - Type
+    DVTDAY = 
+      case_when(DVTDAY == 1 ~ "Weekday",
+                     DVTDAY == 2 ~ "Saturday",
+                     DVTDAY == 3 ~ "Sunday"),
+    DVTDAY = factor(DVTDAY))
+
+summary(gss_main_2005$DVTDAY)
+```
+Save data in R format for later use:
+```{r}
+usethis::use_data(gss_main_2005, 
+                  overwrite = TRUE)
+```
+
+## GSS 1998
+
 ### Aqui

### 2024-12-10 data-raw/03-preprocessing-main-files.Rmd
Socioeconomic variables from the 1992 GSS survey

@@
                      SURVMNTH == 11 ~ "November",
-                     SURVMNTH == 12 ~ "Devember",
+                     SURVMNTH == 12 ~ "December",
                      SURVMNTH == 96 ~ "Valid skip",
@@
 
-```{r SEX}
+```{r SEX-2015}
 values_ordered <- c("Male", "Female")
@@
 ```
+
 ```{r PRTYPEC-2010}
@@
 ```
+
 ```{r CHRINHSDC-2010}
@@
 ```
+
 ```{r CHRTIME6-2010}
@@
 ```
+
 ```{r CHH0512C-2010}
@@
 ```
+
 ```{r CHH0014C-2010}
@@
 
-```{r AGECHRYC}
+```{r AGECHRYC-2005}
 
@@
 ```
+
 ```{r CHRINHSDC-2005}
@@
 ```
+
 ```{r CHRTIME6-2005}
@@
 ```
+
 ```{r REGION-2005}
@@
 ```
+
 ```{r GTU_Q120-2005}
@@
 
+Read the Main files of the 2010 GSS survey:
+
+```{r reading-1998-main-file}
+# Main File
+gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv"))
+```
+
+Select only the socioeconomic variables:
+
+```{r filtering-variables-1998}
+gss_main_1998 <- gss_m_1998 %>%
+  select(RECID:DVTDAY)
+```
+
+```{r SURVMNTH-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # SURVMNTH: Survey month of data collection from February 1998 to January 1999
+    SURVMNTH = 
+           case_when(SURVMNTH == 1 ~ "January", 
+                     SURVMNTH == 2 ~ "February",
+                     SURVMNTH == 3 ~ "March",
+                     SURVMNTH == 4 ~ "April",
+                     SURVMNTH == 5 ~ "May",
+                     SURVMNTH == 6  ~ "June",
+                     SURVMNTH == 7 ~ "July",
+                     SURVMNTH == 8 ~ "August",
+                     SURVMNTH == 9 ~ "September",
+                     SURVMNTH == 10 ~ "October",
+                     SURVMNTH == 11 ~ "November",
+                     SURVMNTH == 12 ~ "December"), 
+    SURVMNTH = factor(SURVMNTH))
+
+summary(gss_main_1998$SURVMNTH)
+```
+```{r LANINT-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # LANINT: Language of interview.
+    LANINT = 
+           case_when(LANINT == 1 ~ "English", 
+                     LANINT == 2 ~ "French"), 
+    LANINT = factor(LANINT))
+
+summary(gss_main_1998$LANINT)
+```
+```{r AGEGR5-1998}
+values_ordered <- c("15 to 17 years","18 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # AGEGR5: Age group of respondent (groups of 5)
+    AGEGR5 = 
+           case_when(AGEGR5 == 1 ~ "15 to 17 years",
+                     AGEGR5 == 2 ~ "18 to 19 years",
+                     AGEGR5 == 3 ~ "20 to 24 years",
+                     AGEGR5 == 4 ~ "25 to 29 years",
+                     AGEGR5 == 5 ~ "30 to 34 years",
+                     AGEGR5 == 6  ~ "35 to 39 year",
+                     AGEGR5 == 7 ~ "40 to 44 years",
+                     AGEGR5 == 8 ~ "45 to 49 years",
+                     AGEGR5 == 9 ~ "50 to 54 years",
+                     AGEGR5 == 10 ~ "55 to 59 year",
+                     AGEGR5 == 11 ~ "60 to 64 years",
+                     AGEGR5 == 12 ~ "65 to 69 years",
+                     AGEGR5 == 13 ~ "70 to 74 years",
+                     AGEGR5 == 14 ~ "75 to 79 year",
+                     AGEGR5 == 15 ~ "80 years and over"), 
+    AGEGR5 = factor(AGEGR5, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$AGEGR5)
+```
+
+```{r AGEGR10-1998}
+values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 to 84 years", "85 years and over")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # AGEGR10: Age group of respondent (groups of 10)
+    AGEGR10 = 
+           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
+                     AGEGR10 == 2 ~ "25 to 34 years",
+                     AGEGR10 == 3 ~ "35 to 44 years",
+                     AGEGR10 == 4 ~ "45 to 54 years",
+                     AGEGR10 == 5 ~ "55 to 64 years",
+                     AGEGR10 == 6  ~ "65 to 74 years",
+                     AGEGR10 == 7 ~ "75 to 84 years",
+                     AGEGR10 == 8 ~ "85 years and over"), 
+    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$AGEGR10)
+```
+
+```{r SEX-1998}
+values_ordered <- c("Male", "Female")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # SEX: Sex of respondent
+    SEX = 
+           case_when(SEX == 1 ~ "Male", 
+                     SEX == 2 ~ "Female"), 
+    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$SEX)
+```
+
+```{r MARSTAT}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # MARSTAT: Marital status of the respondent
+    MARSTAT = 
+           case_when(MARSTAT == 1 ~ "Married", 
+                     MARSTAT == 2 ~ "Living common-law",
+                     MARSTAT == 3 ~ "Widowed",
+                     MARSTAT == 4 ~ "Separated",
+                     MARSTAT == 5 ~ "Divorced",
+                     MARSTAT == 6  ~ "Single, never married",
+                     MARSTAT == 8 ~ "Don't know",
+                     MARSTAT == 9 ~ "Not stated"), 
+    MARSTAT = factor(MARSTAT))
+
+summary(gss_main_1998$MARSTAT)
+```
+
+```{r AGEPRGR5-1998}
+values_ordered <- c("Not Asked","15 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # AGEPRGR5: Age group of respondent (groups of 5)
+    AGEPRGR5 = 
+           case_when(AGEPRGR5 == 97 ~ "Not Asked",
+                     AGEPRGR5 == 1 ~  "15 to 19 years",
+                     AGEPRGR5 == 2 ~ "20 to 24 years",
+                     AGEPRGR5 == 3 ~ "25 to 29 years",
+                     AGEPRGR5 == 4 ~ "30 to 34 years",
+                     AGEPRGR5 == 5  ~ "35 to 39 year",
+                     AGEPRGR5 == 6 ~ "40 to 44 years",
+                     AGEPRGR5 == 7 ~ "45 to 49 years",
+                     AGEPRGR5 == 8 ~ "50 to 54 years",
+                     AGEPRGR5 == 9 ~ "55 to 59 year",
+                     AGEPRGR5 == 10 ~ "60 to 64 years",
+                     AGEPRGR5 == 11 ~ "65 to 69 years",
+                     AGEPRGR5 == 12 ~ "70 to 74 years",
+                     AGEPRGR5 == 13 ~ "75 to 79 year",
+                     AGEPRGR5 == 14 ~ "80 years and over"), 
+    AGEPRGR5 = factor(AGEPRGR5, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$AGEPRGR5)
+```
+```{r PRTYPEC-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # PRTYPEC: Type of partner respondent has living in the household.
+    PRTYPEC = 
+           case_when(PRTYPEC == 1 ~ "Respondent has a married partner in the household", 
+                     PRTYPEC == 2 ~ "Respondent has a common-law partner in the household",
+                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the household"), 
+    PRTYPEC = factor(PRTYPEC))
+
+summary(gss_main_1998$PRTYPEC)
+```
+```{r CHRFLAG-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHRFLAG:  Child(ren) of the respondent living in the household.
+    CHRFLAG = 
+           case_when(CHRFLAG == 1 ~ "Yes", 
+                     CHRFLAG == 2 ~ "No"), 
+    CHRFLAG = factor(CHRFLAG))
+
+summary(gss_main_1998$CHRFLAG)
+```
+```{r AGECHRYC-1998}
+
+values_ordered <- c(as.character(0:24),"25 and over","Not asked - no children in the household")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # AGECHRYC: Age of respondent's youngest child in household
+           AGECHRYC = case_when(AGECHRYC == 25 ~ "25 and over",
+                     AGECHRYC == 97 ~ "Not asked - no children in the household",
+                     TRUE ~ as.character(AGECHRYC)), 
+    AGECHRYC = factor(AGECHRYC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$AGECHRYC)
+```
+
+```{r CHR0004C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR0004C: Number of respondent’s children 0 to 4 years of age living in the household.
+    CHR0004C = 
+           case_when(CHR0004C == 0 ~ "None",
+                     CHR0004C == 1 ~ "One child",
+                     CHR0004C == 2 ~ "Two or more children"), 
+    CHR0004C = factor(CHR0004C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHR0004C)
+```
+
+```{r CHR0512C-1998}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR0512C: Number of children aged from 5 to 12 years living in the respondent's household.
+    CHR0512C = 
+           case_when(CHR0512C == 0 ~ "None",
+                     CHR0512C == 1 ~ "One",
+                     CHR0512C == 2 ~ "Two",
+                     CHR0512C == 3 ~ "Three or more children"), 
+    CHR0512C = factor(CHR0512C, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_1998$CHR0512C)
+```
+
+```{r CHR1314C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR1314C: Number of respondent’s children 13 to 14 years of age living in the household.
+    CHR1314C = 
+           case_when(CHR1314C == 0 ~ "None",
+                     CHR1314C == 1 ~ "One child",
+                     CHR1314C == 2 ~ "Two or more children"), 
+    CHR1314C = factor(CHR1314C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHR1314C)
+```
+
+```{r CHR1518C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR1518C: Number of respondent’s children 15 to 18 years of age living in the household.
+    CHR1518C = 
+           case_when(CHR1518C == 0 ~ "None",
+                     CHR1518C == 1 ~ "One child",
+                     CHR1518C == 2 ~ "Two or more children"), 
+    CHR1518C = factor(CHR1518C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHR1518C)
+```
+
+```{r CHR1924C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR1924C: Number of respondent’s children 19 to 24 years of age living in the household.
+    CHR1924C = 
+           case_when(CHR1924C == 0 ~ "None",
+                     CHR1924C == 1 ~ "One child",
+                     CHR1924C == 2 ~ "Two or more children"), 
+    CHR1924C = factor(CHR1924C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHR1924C)
+```
+```{r CHR25UPC-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHR25UPC: Number of respondent’s children 25 years of age and older living in the household.
+    CHR25UPC = 
+           case_when(CHR25UPC == 0 ~ "None",
+                     CHR25UPC == 1 ~ "One child",
+                     CHR25UPC == 2 ~ "Two or more children"), 
+    CHR25UPC = factor(CHR25UPC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHR25UPC)
+```
+```{r CHH0004C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHH0004C: Number of children aged from 0 to 4 living in the respondent’s household.
+    CHH0004C = 
+           case_when(CHH0004C == 0 ~ "None",
+                     CHH0004C == 1 ~ "One child",
+                     CHH0004C == 2 ~ "Two or more children"), 
+    CHH0004C = factor(CHH0004C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHH0004C)
+```
+
+```{r CHH1314C-1998}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHH1314C: Number of children aged from 13 to 14 living in the respondent’s household.
+    CHH1314C = 
+           case_when(CHH1314C == 0 ~ "None",
+                     CHH1314C == 1 ~ "One child",
+                     CHH1314C == 2 ~ "Two or more children"), 
+    CHH1314C = factor(CHH1314C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHH1314C)
+```
+```{r CHH0512C-1998}
+values_ordered <- c("None", "One child",
+                     "Two children","Three or more children")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CHH0512C: Number of children aged from 13 to 14 living in the respondent’s household.
+    CHH0512C = 
+           case_when(CHH0512C == 0 ~ "None",
+                     CHH0512C == 1 ~ "One child",
+                     CHH0512C == 2 ~ "Two children",
+                     CHH0512C == 3 ~ "Three or more children",), 
+    CHH0512C = factor(CHH0512C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$CHH0512C)
+```
+
+```{r CHRTIME6-2005}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # CHRTIME6:  Age group of respondent’s child(ren) in household
+    CHRTIME6 = 
+           case_when(CHRTIME6 == 1 ~ "No child under 19 years of age at home",
+                     CHRTIME6 == 2 ~ "All children under 5 years of age",
+                     CHRTIME6 == 3 ~ "All children between 5 and 12 years of age",
+                     CHRTIME6 == 4 ~ "All children 13 years of age and older",
+                     CHRTIME6 == 5 ~ "At least one child under 5 years of age but not all children",
+                     CHRTIME6 == 6 ~ "Other"), 
+    CHRTIME6 = factor(CHRTIME6))
+
+summary(gss_main_2005$CHRTIME6)
+```
+
+```{r PARHSDC}
+gss_main_2005 <- gss_main_2005 %>% 
+  mutate( # PARHSDC: Type of parents the respondent has within the household.
+    PARHSDC = 
+           case_when(PARHSDC == 1 ~ "Two biological parents",
+                     PARHSDC == 2 ~ "Two parents (biological/adoptive/step)",
+                     PARHSDC == 3 ~ "Only one parent",
+                     PARHSDC == 7 ~ "Not asked - no parents in the household"), 
+    PARHSDC = factor(PARHSDC))
+
+summary(gss_main_2005$PARHSDC)
+```
+
+```{r LIVARR08-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
+    LIVARR08 = 
+           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
+                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
+                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
+                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
+                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
+                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
+                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
+                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only"), 
+    LIVARR08 = factor(LIVARR08))
+
+summary(gss_main_1998$LIVARR08)
+```
+
+```{r LIVARR12-1998} 
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # LIVARR12: Living arrangement of respondent’s household (12 categories)
+    LIVARR12 = 
+           case_when(LIVARR12 == 1 ~ "Living alone",
+                     LIVARR12 == 2 ~ "Living with sponse only",
+                     LIVARR12 == 3 ~ "Living with spouse and single child < 25 years",
+                     LIVARR12 == 4 ~ "Living with spouse and single child 25 years or older",
+                     LIVARR12 == 5 ~ "Living with spouse and non-child(ren)",
+                     LIVARR12 == 6 ~ "Living with spouse and other",
+                     LIVARR12 == 7 ~ "Living without spouse and single child  < 25 years",
+                     LIVARR12 == 8 ~ "Living without spouse and single child 25 years or older",
+                     LIVARR12 == 9 ~ "No spouse and non-single child(ren)",
+                     LIVARR12 == 10 ~ "Living with two parents",
+                     LIVARR12 == 11 ~ "Living with one parent",
+                     LIVARR12 == 12 ~ "Other living arrangement"), 
+    LIVARR12 = factor(LIVARR12))
+
+summary(gss_main_1998$LIVARR12)
+```
+
+```{r HSDSIZEC-1998}
+values_ordered <- c("One household member",
+                   "Two household members",
+                   "Three household members",
+                   "Four household members",
+                   "Five household members",
+                   "Six household members",
+                   "Seven household members", 
+                   "Eight household members or more")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # HSDSIZEC: Household size of respondent
+    HSDSIZEC = 
+           case_when(HSDSIZEC == 1 ~ "One household member",
+                     HSDSIZEC == 2 ~ "Two household members",
+                     HSDSIZEC == 3 ~ "Three household members",
+                     HSDSIZEC == 4 ~ "Four household members",
+                     HSDSIZEC == 5 ~ "Five household members",
+                     HSDSIZEC == 6 ~ "Six household members",
+                     HSDSIZEC == 7 ~ "Seven household members",
+                     HSDSIZEC == 8 ~ "Eight household members or more"), 
+    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1998$HSDSIZEC)
+```
+
+```{r FAMTYPE-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # FAMTYPE: Respondent's type of family structure.
+    FAMTYPE = 
+           case_when(FAMTYPE == 1 ~ "Couple only", 
+                     FAMTYPE == 2 ~ "Intact family",
+                     FAMTYPE == 3 ~ "Step family with common child",
+                     FAMTYPE == 4 ~ "Step family without a common child",
+                     FAMTYPE == 5 ~ "Lone parent family",
+                     FAMTYPE == 7 ~ "Not asked"), 
+    FAMTYPE = factor(FAMTYPE))
+
+summary(gss_main_1998$FAMTYPE)
+```
+
+```{r MULTIGEN-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
+    MULTIGEN = 
+           case_when(MULTIGEN == 1 ~ "Yes", 
+                     MULTIGEN == 2 ~ "No"), 
+    MULTIGEN = factor(MULTIGEN))
+
+summary(gss_main_1998$MULTIGEN)
+```
+
+```{r PRV-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # PRV: Province of residence
+    PRV = 
+      case_when(PRV == 10 ~ "Newfoundland and Labrador",
+                     PRV == 11 ~ "Prince Edward Island",
+                     PRV == 12 ~ "Nova Scotia",
+                     PRV == 13 ~ "New Brunswick",
+                     PRV == 24 ~ "Quebec",
+                     PRV == 35 ~ "Ontario",
+                     PRV == 46 ~ "Manitoba",
+                     PRV == 47 ~ "Saskatchewan",
+                     PRV == 48 ~ "Alberta",
+                     PRV == 59 ~ "British Columbia"),
+    PRV = factor(PRV))
+
+summary(gss_main_1998$PRV) 
+```
+
+```{r REGION-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # REGION:  Region of residence of the respondent.
+    REGION = 
+      case_when(REGION == 1 ~ "Atlantic region",
+                     REGION == 2 ~ "Quebec",
+                     REGION == 3 ~ "Ontario",
+                     REGION == 4 ~ "Prairie region",
+                     REGION == 5 ~ "British Columbia"),
+    REGION = factor(REGION))
+
+summary(gss_main_1998$REGION)
+```
+
+```{r CMAPRV}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # CMAPRV:  Population centre indicator
+    CMAPRV = 
+      case_when(CMAPRV %in% c(6,8,11,13,15) ~ "CMA",
+                     CMAPRV %in% c(1,2,3,4,5,7,9,10,12,14) ~ "non CMA"),
+    CMAPRV = factor(CMAPRV))
+
+summary(gss_main_1998$CMAPRV)
+```
+
+```{r A2-1998}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Not stated")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # A2: How often do you feel rushed?
+    A2 = 
+      case_when(A2 == 1 ~ "Every day",
+                     A2 == 2 ~ "A few times a week",
+                     A2 == 3 ~ "About once a week",
+                     A2 == 4 ~ "About once a month",
+                     A2 == 5 ~ "Less than once a month",
+                     A2 == 6 ~ "Never",
+                     A2 == 9 ~ "Not stated"),
+    A2 = factor(A2, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1998$A2)
+```
+
+```{r A3-1998}
+values_ordered <- c("More rushed",
+                    "About the same",
+                    "Less rushed",
+                    "Don't know","Not stated")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # A3: Compared to five years ago, do you feel...
+    A3 = 
+      case_when(A3 == 1 ~ "More rushed",
+                     A3 == 2 ~  "About the same",
+                     A3 == 3 ~ "Less rushed",
+                     A3 == 8 ~ "Don't know",
+                     A3 == 9 ~ "Not stated"),
+    A3 = factor(A3, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1998$A3)
+```
+
+```{r GTU_130-1998}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Not stated")
+
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # A4: How often do you feel you have time on your hands that you don’t know what to do with?
+    A4 = 
+      case_when(A4 == 1 ~ "Every day",
+                     A4 == 2 ~ "A few times a week",
+                     A4 == 3 ~ "About once a week",
+                     A4 == 4 ~ "About once a month",
+                     A4 == 5 ~ "Less than once a month",
+                     A4 == 6 ~ "Never",
+                     A4 == 9 ~ "Not stated"),
+    A4 = factor(A4, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1998$A4)
+```
+
+```{r A5-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # A5:  Do you feel that the days are just too short to do all the things you want?
+    A5 = 
+           case_when(A5 == 1 ~ "Yes", 
+                     A5 == 2 ~ "No",
+                     A5 == 9 ~ "Not stated", 
+                     A5 == 8  ~ "Don't know"), 
+    A5 = factor(A5))
+
+summary(gss_main_1998$A5)
+```
+
+```{r A6-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # A6: On which main activity would you choose to spend more time on
+    A6 = 
+      case_when(A6 == 1 ~ "Time with family and friends (spouse, children, boy-friend - girl-friend)",
+                     A6 == 2 ~ "Relaxation - personal time",
+                     A6 == 3 ~ "Practising sports",
+                     A6 == 4 ~ "Crafts or hobbies",
+                     A6 == 5 ~ "Outdoor activities",
+                     A6 == 6 ~ "Reading - writing",
+                     A6 == 7 ~ "Studies",
+                     A6 == 8 ~ "Work",
+                     A6 == 9 ~ "Domestic work",
+                     A6 == 10 ~ "Holiday/Travel",
+                     A6 == 11 ~ "Volunteer",
+                     A6 == 12  ~ "Other",
+                     A6 == 97  ~ "Not asked",
+                     A6 == 98 ~ "Not stated",
+                     A6 == 99 ~ "Don't know",
+),
+    A6 = factor(A6))
+
+summary(gss_main_1998$A6)
+```
+```{r DDAY}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # DDAY: Diary - Reference day - Type
+    DDAY = 
+      case_when(DDAY == 1 ~ "Sunday",
+                     DDAY == 2 ~ "Monday",
+                     DDAY == 3 ~ "Tuesday",
+                     DDAY == 4 ~ "Wednesday",
+                     DDAY == 5 ~ "Thursday",
+                     DDAY == 6 ~ "Friday",
+                     DDAY == 7 ~ "Saturday"),
+    DDAY = factor(DDAY))
+
+summary(gss_main_1998$DDAY)
+```
+
+```{r DIARYREC}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # DIARYREC: Recall of diary interview
+    DIARYREC = 
+      case_when(DIARYREC == 1 ~ "Diary completed the day following the designated day (24 hrs recall)",
+                     DIARYREC == 2 ~ "Diary completed the 2nd day following the designated day (48 hrs recall)",
+                     DIARYREC == 3 ~ "Diary completed in 2 days (24 to 48 hrs recall)"),
+    DIARYREC = factor(DIARYREC))
+
+summary(gss_main_1998$DIARYREC)
+```
+
+```{r DVTDAY}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # DVTDAY: Diary - Reference day - Type
+    DVTDAY = 
+      case_when(DVTDAY == 1 ~ "Weekday",
+                     DVTDAY == 2 ~ "Saturday",
+                     DVTDAY == 3 ~ "Sunday"),
+    DVTDAY = factor(DVTDAY))
+
+summary(gss_main_1998$DVTDAY)
+```
+
+## GSS 1992
+
+Read the Main files of the 2010 GSS survey:
+
+```{r reading-1992-main-file}
+# Main File
+gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv"))
+```
+
+Select only the socioeconomic variables:
+
+```{r filtering-variables-1998}
+gss_main_1992 <- gss_m_1992 %>%
+  select(c("SEQNUM", 
+ "FWGHT", 
+ "SURMNTH", 
+ "FORMTYPE",
+ "A4", 
+ "A5", 
+ "A6",
+ "C2", 
+ "DVAGEGR", 
+ "DVSEX", 
+ "DVMS", 
+ "DVHHSCAP",
+ "DVPROV", 
+ "DVCMA", 
+ "DVLVGRGR", 
+ "DVAGRYC", 
+ "RESC0004", 
+ "RESC0512", 
+ "RESC1318", 
+ "RESC1924", 
+ "RESC2596", 
+ "DVAGRSP"))
+```
+
+```{r C2-1992}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # C2:  Respondent has child(ren) <15?
+    C2 = 
+           case_when(C2 == 1 ~ "Yes", 
+                     C2 == 2 ~ "No"), 
+    C2 = factor(C2))
+
+summary(gss_main_1992$C2)
+```
+
+```{r SURMNTH-1992}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # SURMNTH: Survey month of data collection from April 1992 to March 2016
+    SURMNTH = 
+           case_when(SURMNTH == 1 ~ "January", 
+                     SURMNTH == 2 ~ "February",
+                     SURMNTH == 3 ~ "March",
+                     SURMNTH == 4 ~ "April",
+                     SURMNTH == 5 ~ "May",
+                     SURMNTH == 6  ~ "June",
+                     SURMNTH == 7 ~ "July",
+                     SURMNTH == 8 ~ "August",
+                     SURMNTH == 9 ~ "September",
+                     SURMNTH == 10 ~ "October",
+                     SURMNTH == 11 ~ "November",
+                     SURMNTH == 12 ~ "December"), 
+    SURMNTH = factor(SURMNTH))
+
+summary(gss_main_1992$SURMNTH)
+```
+
+```{r FORMTYPE-1992}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # FORMTYPE: Language of interview.
+    FORMTYPE = 
+           case_when(FORMTYPE == 1 ~ "English", 
+                     FORMTYPE == 2 ~ "French"), 
+    FORMTYPE = factor(FORMTYPE))
+
+summary(gss_main_1992$FORMTYPE)
+```
+
+```{r A4-1992}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Not stated")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # A4: How often do you feel rushed?
+    A4 = 
+      case_when(A4 == 1 ~ "Every day",
+                     A4 == 2 ~ "A few times a week",
+                     A4 == 3 ~ "About once a week",
+                     A4 == 4 ~ "About once a month",
+                     A4 == 5 ~ "Less than once a month",
+                     A4 == 6 ~ "Never",
+                     A4 == 9 ~ "Not stated"),
+    A4 = factor(A4, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1992$A4)
+```
+
+```{r A5-1992}
+values_ordered <- c("More rushed",
+                    "About the same",
+                    "Less rushed","Not stated")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # A5: Compared to five years ago, do you feel...
+    A5 = 
+      case_when(A5 == 1 ~ "More rushed",
+                     A5 == 2 ~  "About the same",
+                     A5 == 3 ~ "Less rushed",
+                     A5 == 9 ~ "Not stated"),
+    A5 = factor(A5, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1992$A5)
+```
+
+```{r GTU_130-1992}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never", 
+                    "Not stated")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # A6: How often do you feel you have time on your hands that you don’t know what to do with?
+    A6 = 
+      case_when(A6 == 1 ~ "Every day",
+                     A6 == 2 ~ "A few times a week",
+                     A6 == 3 ~ "About once a week",
+                     A6 == 4 ~ "About once a month",
+                     A6 == 5 ~ "Less than once a month",
+                     A6 == 6 ~ "Never",
+                     A6 == 9 ~ "Not stated"),
+    A6 = factor(A6, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_1992$A6)
+```
+```{r DVAGEGR-1992}
+values_ordered <- c("15 to 17 years","18 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVAGEGR: Age group of respondent (groups of 5)
+    DVAGEGR = 
+           case_when(DVAGEGR == 1 ~ "15 to 17 years",
+                     DVAGEGR == 2 ~ "18 to 19 years",
+                     DVAGEGR == 3 ~ "20 to 24 years",
+                     DVAGEGR == 4 ~ "25 to 29 years",
+                     DVAGEGR == 5 ~ "30 to 34 years",
+                     DVAGEGR == 6  ~ "35 to 39 year",
+                     DVAGEGR == 7 ~ "40 to 44 years",
+                     DVAGEGR == 8 ~ "45 to 49 years",
+                     DVAGEGR == 9 ~ "50 to 54 years",
+                     DVAGEGR == 10 ~ "55 to 59 year",
+                     DVAGEGR == 11 ~ "60 to 64 years",
+                     DVAGEGR == 12 ~ "65 to 69 years",
+                     DVAGEGR == 13 ~ "70 to 74 years",
+                     DVAGEGR == 14 ~ "75 to 79 year",
+                     DVAGEGR == 15 ~ "80 years and over"), 
+    DVAGEGR = factor(DVAGEGR, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$DVAGEGR)
+```
+
+```{r DVSEX-1992}
+values_ordered <- c("Male", "Female")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVSEX: DVSEX of respondent
+    DVSEX = 
+           case_when(DVSEX == 1 ~ "Male", 
+                     DVSEX == 2 ~ "Female"), 
+    DVSEX = factor(DVSEX, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$DVSEX)
+```
+
+```{r DVMS}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVMS: Marital status of the respondent
+    DVMS = 
+           case_when(DVMS == 1 ~ "Married or common-law",
+                     DVMS == 2 ~ "Widowed",
+                     DVMS == 3 ~ "Separated or Divorced",
+                     DVMS == 4  ~ "Single, never married",
+                     DVMS == 9 ~ "Not stated"), 
+    DVMS = factor(DVMS))
+
+summary(gss_main_1992$DVMS)
+```
+
+```{r DVHHSCAP-1992}
+values_ordered <- c("One household member",
+                   "Two household members",
+                   "Three household members",
+                   "Four household members",
+                   "Five household members",
+                   "Six household members",
+                   "Seven members or more")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVHHSCAP: Household size of respondent
+    DVHHSCAP = 
+           case_when(DVHHSCAP == 1 ~ "One household member",
+                     DVHHSCAP == 2 ~ "Two household members",
+                     DVHHSCAP == 3 ~ "Three household members",
+                     DVHHSCAP == 4 ~ "Four household members",
+                     DVHHSCAP == 5 ~ "Five household members",
+                     DVHHSCAP == 6 ~ "Six household members",
+                     DVHHSCAP == 7 ~ "Seven members or more"), 
+    DVHHSCAP = factor(DVHHSCAP, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$DVHHSCAP)
+```
+
+```{r DVPROV-1992}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVPROV: Province of residence
+    DVPROV = 
+      case_when(DVPROV == 0 ~ "Newfoundland and Labrador",
+                     DVPROV == 1 ~ "Prince Edward Island",
+                     DVPROV == 2 ~ "Nova Scotia",
+                     DVPROV == 3 ~ "New Brunswick",
+                     DVPROV == 4 ~ "Quebec",
+                     DVPROV == 5 ~ "Ontario",
+                     DVPROV == 6 ~ "Manitoba",
+                     DVPROV == 7 ~ "Saskatchewan",
+                     DVPROV == 8 ~ "Alberta",
+                     DVPROV == 9 ~ "British Columbia"),
+    DVPROV = factor(DVPROV))
+
+summary(gss_main_1992$DVPROV) 
+```
+
+```{r DVCMA-1992}
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVCMA:  Population centre indicator
+    DVCMA = 
+      case_when(DVCMA %in% c(6,8,11,13,15) ~ "CMA",
+                     DVCMA %in% c(1,2,3,4,5,7,9,10,12,14) ~ "non CMA"),
+    DVCMA = factor(DVCMA))
+
+summary(gss_main_1992$DVCMA)
+```
+```{r DVLVGRGR-1992} 
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVLVGRGR: Living arrangement of respondent’s household (12 categories)
+    DVLVGRGR = 
+           case_when(DVLVGRGR == 1 ~ "Living alone",
+                     DVLVGRGR == 2 ~ "Living with sponse only",
+                     DVLVGRGR == 3 ~ "Living with spouse and single child < 25 years",
+                     DVLVGRGR == 4 ~ "Living with spouse and single child 25 years or older",
+                     DVLVGRGR == 5 ~ "Living with spouse and non-child(ren)",
+                     DVLVGRGR == 6 ~ "Living with spouse and other",
+                     DVLVGRGR == 7 ~ "Living without spouse and single child  < 25 years",
+                     DVLVGRGR == 8 ~ "Living without spouse and single child 25 years or older",
+                     DVLVGRGR == 9 ~ "No spouse and non-single child(ren)",
+                     DVLVGRGR == 10 ~ "Living with two parents",
+                     DVLVGRGR == 11 ~ "Living with one parent",
+                     DVLVGRGR == 12 ~ "Other living arrangement"), 
+    DVLVGRGR = factor(DVLVGRGR))
+
+summary(gss_main_1992$DVLVGRGR)
+```
+
+```{r RESC0004C-1992}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # RESC0004: Number of respondents single children aged 0-4.
+    RESC0004 = 
+           case_when(RESC0004 == 0 ~ "None",
+                     RESC0004 == 1 ~ "One child",
+                     RESC0004 == 2 ~ "Two or more children"), 
+    RESC0004 = factor(RESC0004, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$RESC0004)
+```
+```{r RESC0512C-1992}
+values_ordered <- c("None", "One",
+                     "Two",
+                     "Three or more children")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # RESC0512: Number of children aged from 5 to 12 years living in the respondent's household.
+    RESC0512 = 
+           case_when(RESC0512 == 0 ~ "None",
+                     RESC0512 == 1 ~ "One",
+                     RESC0512 == 2 ~ "Two",
+                     RESC0512 == 3 ~ "Three or more children"), 
+    RESC0512 = factor(RESC0512, levels = values_ordered, ordered = TRUE)) 
+summary(gss_main_1992$RESC0512)
+```
+
+```{r RESC1318-1992}
+values_ordered <- c("None", "One child",
+                     "Two children", "Three or more children")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # RESC1318: Number of respondent’s children 15 to 18 years of age living in the household.
+    RESC1318 = 
+           case_when(RESC1318 == 0 ~ "None",
+                     RESC1318 == 1 ~ "One child",
+                     RESC1318 == 2 ~ "Two children", 
+                     RESC1318 == 3 ~ "Three or more children"), 
+    RESC1318 = factor(RESC1318, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$RESC1318)
+```
+
+```{r RESC1924-1992}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # RESC1924: Number of respondent’s children 19 to 24 years of age living in the household.
+    RESC1924 = 
+           case_when(RESC1924 == 0 ~ "None",
+                     RESC1924 == 1 ~ "One child",
+                     RESC1924 == 2 ~ "Two or more children"), 
+    RESC1924 = factor(RESC1924, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$RESC1924)
+```
+
+```{r RESC2596C-1992}
+values_ordered <- c("None", "One child",
+                     "Two or more children")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # RESC2596: Number of respondent’s children 25 years of age and older living in the household.
+    RESC2596 = 
+           case_when(RESC2596 == 0 ~ "None",
+                     RESC2596 == 1 ~ "One child",
+                     RESC2596 == 2 ~ "Two or more children"), 
+    RESC2596 = factor(RESC2596, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$RESC2596)
+```
+```{r DVAGRSP-1992}
+values_ordered <- c("SPOUSE NOT LIVING IN HOUSEHOLD",
+                    "NOT APPLICABLE",
+                    "NOT STATED",
+                    "15 to 24 years","25 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVAGRSP: Age group of the respondents spouse.
+    DVAGRSP = 
+           case_when(DVAGRSP == 1 ~ "15 to 24 years", 
+                     DVAGRSP == 2 ~ "25 to 34 years",
+                     DVAGRSP == 3 ~ "35 to 44 years",
+                     DVAGRSP == 4 ~ "45 to 54 years",
+                     DVAGRSP == 5 ~ "55 to 64 years",
+                     DVAGRSP == 6  ~ "65 to 74 years",
+                     DVAGRSP == 7 ~ "75 years and over",
+                     DVAGRSP == 8 ~ "SPOUSE NOT LIVING IN HOUSEHOLD", 
+                     DVAGRSP == 0 ~ "NOT APPLICABLE", 
+                     DVAGRSP == 9 ~ "NOT STATED"), 
+    DVAGRSP = factor(DVAGRSP, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$DVAGRSP)
+```
+
 ### Aqui

### 2024-12-11 data-raw/03-preprocessing-main-files.Rmd
Documentation for main GSS files

@@
 
-```{r SURVMNTH}
+```{r SURVMNTH-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r AGEGR10}
+```{r AGEGR10-2015}
 values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
@@
 
-```{r MARSTAT}
+```{r MARSTAT-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r AGEPRGRD}
+```{r AGEPRGRD-2015}
 values_ordered <- c("Respondent is 11 + years older", 
@@
 
-```{r SEXPR}
+```{r SEXPR-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r PRTYPEC}
+```{r PRTYPEC-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r CXRFLAG}
+```{r CXRFLAG-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r AGECXRYG}
+```{r AGECXRYG-2015}
 values_ordered <- c( "Valid skip", "5 to 4", "5 to 9","10 to 14",
@@
 
-```{r CHINHSDC}
+```{r CHINHSDC-2015}
 values_ordered <- c("None", 
@@
 
-```{r CXR0014C}
+```{r CXR0014C-2015}
 values_ordered <- c("None", 
@@
 
-```{r CHH0014C}
+```{r CHH0014C-2015}
 values_ordered <- c("None", 
@@
 
-```{r CXRTIME6}
+```{r CXRTIME6-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r PARNUM}
+```{r PARNUM-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r LIVARR08}
+```{r LIVARR08-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r LIVARR11} 
+```{r LIVARR11-2015} 
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r HSDSIZEC}
+```{r HSDSIZEC-2015}
 values_ordered <- c("One person household",
@@
 
-```{r HSDELIGC}
+```{r HSDELIGC-2015}
 values_ordered <- c("One member",
@@
 
-```{r MULTIGEN}
+```{r MULTIGEN-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r PRV}
+```{r PRV-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r LUC_RST}
+```{r LUC_RST-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 
-```{r GTU_110}
+```{r GTU_110-2015}
 values_ordered <- c("Every day",
@@
 
-```{r GTU_130}
+```{r GTU_130-2015}
 values_ordered <- c("Every day",
@@
 
-```{r DVTDAY}
+```{r DVTDAY-2015}
 gss_main_2015 <- gss_main_2015 %>% 
@@
 Save data in R format for later use:
-```{r}
+```{r save-data-2015}
 usethis::use_data(gss_main_2015, 
@@
 
-```{r MARSTAT}
+```{r MARSTAT-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 
-```{r AGECHRYC}
+```{r AGECHRYC-2010}
 
@@
 
-```{r PARHSDC}
+```{r PARHSDC-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 
-```{r LIVARR08}
+```{r LIVARR08-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 
-```{r HSDSIZEC}
+```{r HSDSIZEC-2010}
 values_ordered <- c("One person household",
@@
 
-```{r HSDELIGC}
+```{r HSDELIGC-2010}
 values_ordered <- c("One member",
@@
 ```
+
 ```{r FAMTYPE-2010}
@@
 
-```{r MULTIGEN}
+```{r MULTIGEN-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 ```
+
 ```{r REGION-2010}
@@
 
-```{r LUC_RST}
+```{r LUC_RST-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 
-```{r GTU_110}
+```{r GTU_110-2010}
 values_ordered <- c("Every day",
@@
 
-```{r GTU_130}
+```{r GTU_130-2010}
 values_ordered <- c("Every day",
@@
 
-```{r DVTDAY}
+```{r DVTDAY-2010}
 gss_main_2010 <- gss_main_2010 %>% 
@@
 Save data in R format for later use:
-```{r}
+```{r save-data-2010}
 usethis::use_data(gss_main_2010, 
@@
 
-Read the Main files of the 2010 GSS survey:
-
-```{r reading-2010-main-file}
+Read the Main files of the 2005 GSS survey:
+
+```{r reading-2005-main-file}
 # Main File
@@
 
-```{r filtering-variables-2010}
+```{r filtering-variables-22005015}
 gss_main_2005 <- gss_m_2005 %>%
@@
 
-```{r MARSTAT}
+```{r MARSTAT-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 
-```{r PARHSDC}
+```{r PARHSDC-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 
-```{r LIVARR08}
+```{r LIVARR08-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 
-```{r MULTIGEN}
+```{r MULTIGEN-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 
-```{r LUC_RST}
+```{r LUC_RST-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 ```
+
 ```{r GTU_Q140-2005}
@@
 
-```{r DVTDAY}
+```{r DVTDAY-2005}
 gss_main_2005 <- gss_main_2005 %>% 
@@
 ```
+
 Save data in R format for later use:
-```{r}
+```{r save-data-2005}
 usethis::use_data(gss_main_2005, 
@@
 
-Read the Main files of the 2010 GSS survey:
+Read the Main files of the 1998 GSS survey:
 
@@
 ```
+
 ```{r LANINT-1998}
@@
 ```
+
 ```{r AGEGR5-1998}
@@
 
-```{r MARSTAT}
+```{r MARSTAT-1998}
 gss_main_1998 <- gss_main_1998 %>% 
@@
 ```
+
 ```{r PRTYPEC-1998}
@@
 ```
+
 ```{r CHRFLAG-1998}
@@
 ```
+
 ```{r AGECHRYC-1998}
@@
 ```
+
 ```{r CHR25UPC-1998}
@@
 ```
+
 ```{r CHH0512C-1998}
@@
 summary(gss_main_1998$CHH0512C)
-```
-
-```{r CHRTIME6-2005}
-gss_main_2005 <- gss_main_2005 %>% 
-  mutate( # CHRTIME6:  Age group of respondent’s child(ren) in household
-    CHRTIME6 = 
-           case_when(CHRTIME6 == 1 ~ "No child under 19 years of age at home",
-                     CHRTIME6 == 2 ~ "All children under 5 years of age",
-                     CHRTIME6 == 3 ~ "All children between 5 and 12 years of age",
-                     CHRTIME6 == 4 ~ "All children 13 years of age and older",
-                     CHRTIME6 == 5 ~ "At least one child under 5 years of age but not all children",
-                     CHRTIME6 == 6 ~ "Other"), 
-    CHRTIME6 = factor(CHRTIME6))
-
-summary(gss_main_2005$CHRTIME6)
-```
-
-```{r PARHSDC}
-gss_main_2005 <- gss_main_2005 %>% 
-  mutate( # PARHSDC: Type of parents the respondent has within the household.
-    PARHSDC = 
-           case_when(PARHSDC == 1 ~ "Two biological parents",
-                     PARHSDC == 2 ~ "Two parents (biological/adoptive/step)",
-                     PARHSDC == 3 ~ "Only one parent",
-                     PARHSDC == 7 ~ "Not asked - no parents in the household"), 
-    PARHSDC = factor(PARHSDC))
-
-summary(gss_main_2005$PARHSDC)
 ```
@@
 
-```{r CMAPRV}
+```{r CMAPRV-1998}
 gss_main_1998 <- gss_main_1998 %>% 
@@
 ```
-```{r DDAY}
+
+```{r DDAY-1998}
 gss_main_1998 <- gss_main_1998 %>% 
@@
 
-```{r DIARYREC}
+```{r DIARYREC-1998}
 gss_main_1998 <- gss_main_1998 %>% 
@@
 
-```{r DVTDAY}
+```{r DVTDAY-1998}
 gss_main_1998 <- gss_main_1998 %>% 
@@
 
+```{r PARHSDC-1998}
+gss_main_1998 <- gss_main_1998 %>% 
+  mutate( # PARHSD: Type of parents the respondent has within the household.
+    PARHSD = 
+           case_when(PARHSD == 1 ~ "Two biological parents",
+                     PARHSD == 2 ~ "Biologic parents",
+                     PARHSD == 3 ~ "Biologic-parents",
+                     PARHSD == 4 ~ "Only one parent",
+                     PARHSD == 7 ~ "Not asked - no parents in the household"), 
+    PARHSD = factor(PARHSD))
+
+summary(gss_main_1998$PARHSD)
+```
+
+Save data in R format for later use:
+```{r save-data-1998}
+usethis::use_data(gss_main_1998, 
+                  overwrite = TRUE)
+```
+
 ## GSS 1992
 
-Read the Main files of the 2010 GSS survey:
+Read the Main files of the 1992 GSS survey:
 
@@
 ```
+
 ```{r DVAGEGR-1992}
@@
 
-```{r DVMS}
+```{r DVMS-1992}
 gss_main_1992 <- gss_main_1992 %>% 
@@
 ```
+
 ```{r DVAGRSP-1992}
@@
 ```
-
-### Aqui
+```{r DVAGRYC-1992}
+values_ordered <- c(as.character(0:24),"25 and over","NOT APPLICABLE", "SINGLE CHILD OLDER THAN 25")
+
+gss_main_1992 <- gss_main_1992 %>% 
+  mutate( # DVAGRYC: Age of respondent's youngest child in household
+           DVAGRYC = case_when(DVAGRYC == 25 ~ "25 and over",
+                     DVAGRYC == 97 ~ "NOT APPLICABLE",
+                     DVAGRYC == 98 ~ "SINGLE CHILD OLDER THAN 25",
+                     TRUE ~ as.character(DVAGRYC)), 
+    DVAGRYC = factor(DVAGRYC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1992$DVAGRYC)
+```
+
+Save data in R format for later use:
+```{r save-data-1992}
+usethis::use_data(gss_main_1992, 
+                  overwrite = TRUE)
+```
+
+## GSS 1986
+
+Read the Main files of the 1986 GSS survey:
+
+```{r reading-1986-main-file}
+# Main File
+gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv"))
+```
+
+Select only the socioeconomic variables:
+
+```{r filtering-variables-1986}
+gss_main_1986 <- gss_m_1986 %>%
+  select(c("SEQNUM", 
+ "FWGT_OS", 
+ "DVAGEGR", 
+ "DVSEX", 
+ "DVMARST",
+ "DVHHLDSZ",
+ "PROV", 
+ "DVCHILD"))
+```
+
+```{r C2-1986}
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # DVCHILD: NUMBER OF CHILDREN AT HOME
+    DVCHILD = 
+           case_when(DVCHILD == 1 ~ "NONE < 19", 
+                     DVCHILD == 2 ~ "ALL < 5",
+                     DVCHILD == 3~ "4 < ALL < 13",
+                     DVCHILD == 4 ~ "12 < ALL < 19",
+                     DVCHILD == 5 ~ "SOME < 5",
+                     DVCHILD == 6 ~ "OTHER"), 
+    DVCHILD = factor(DVCHILD))
+
+summary(gss_main_1986$DVCHILD)
+```
+
+```{r DVAGEGR-1986}
+values_ordered <- c("15 to 19 years","20 to 24 years",
+                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
+                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
+                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")
+
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # DVAGEGR: Age group of respondent (groups of 5)
+    DVAGEGR = 
+           case_when(DVAGEGR == 1 ~ "15 to 19 years",
+                     DVAGEGR == 2 ~ "20 to 24 years",
+                     DVAGEGR == 3 ~ "25 to 29 years",
+                     DVAGEGR == 4 ~ "30 to 34 years",
+                     DVAGEGR == 5  ~ "35 to 39 year",
+                     DVAGEGR == 6 ~ "40 to 44 years",
+                     DVAGEGR == 7 ~ "45 to 49 years",
+                     DVAGEGR == 8 ~ "50 to 54 years",
+                     DVAGEGR == 9 ~ "55 to 59 year",
+                     DVAGEGR == 10 ~ "60 to 64 years",
+                     DVAGEGR == 11 ~ "65 to 69 years",
+                     DVAGEGR == 12 ~ "70 to 74 years",
+                     DVAGEGR == 13 ~ "75 to 79 year",
+                     DVAGEGR == 1 ~ "80 years and over"), 
+    DVAGEGR = factor(DVAGEGR, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1986$DVAGEGR)
+```
+
+```{r DVSEX-1986}
+values_ordered <- c("Male", "Female")
+
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # DVSEX: DVSEX of respondent
+    DVSEX = 
+           case_when(DVSEX == 1 ~ "Male", 
+                     DVSEX == 2 ~ "Female"), 
+    DVSEX = factor(DVSEX, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1986$DVSEX)
+```
+
+```{r DVMARST-1986}
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # DVMARST: Marital status of the respondent
+    DVMARST = 
+           case_when(DVMARST == 1 ~ "Married",
+                     DVMARST == 3 ~ "Widowed",
+                     DVMARST == 4 ~ "Separated or Divorced",
+                     DVMARST == 2  ~ "Single",
+                     DVMARST == 9 ~ "Not stated"), 
+    DVMARST = factor(DVMARST))
+
+summary(gss_main_1986$DVMARST)
+```
+
+```{r DVHHLDSZ-1986}
+values_ordered <- c("One household member",
+                   "Two household members",
+                   "Three household members",
+                   "Four members or more")
+
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # DVHHLDSZ: Household size of respondent
+    DVHHLDSZ = 
+           case_when(DVHHLDSZ == 1 ~ "One household member",
+                     DVHHLDSZ == 2 ~ "Two household members",
+                     DVHHLDSZ == 3 ~ "Three household members",
+                     DVHHLDSZ == 4 ~ "Four members or more"), 
+    DVHHLDSZ = factor(DVHHLDSZ, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_1986$DVHHLDSZ)
+```
+
+```{r PROV-1986}
+gss_main_1986 <- gss_main_1986 %>% 
+  mutate( # PROV: Province of residence
+    PROV = 
+      case_when(PROV == 0 ~ "Newfoundland and Labrador",
+                     PROV == 1 ~ "Prince Edward Island",
+                     PROV == 2 ~ "Nova Scotia",
+                     PROV == 3 ~ "New Brunswick",
+                     PROV == 4 ~ "Quebec",
+                     PROV == 5 ~ "Ontario",
+                     PROV == 6 ~ "Manitoba",
+                     PROV == 7 ~ "Saskatchewan",
+                     PROV == 8 ~ "Alberta",
+                     PROV == 9 ~ "British Columbia"),
+    PROV = factor(PROV))
+
+summary(gss_main_1986$PROV) 
+```
+Save data in R format for later use:
+```{r}
+usethis::use_data(gss_main_1986, 
+                  overwrite = TRUE)
+```

### 2024-12-11 R/ActiveCA-package.R

@@ -476,3 +476,263 @@ NULL
 #' @usage data(gss_impedances)
 #' @source [General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) / [Odesi](http://odesi2.scholarsportal.info/webview/)
 "gss_impedances"
+
+#' Selection of socioeconomic variables from the 1986 GSS Time Use Survey (Cycle 2).
+#'
+#' This dataset contains some socioeconomic variables obtained
+#' from the main file of the 1986 GSS Time Use Survey (Cycle 2). We labeled
+#' variables according to the Public Use Micro Data File Documentation and User's Guide (gss-12M0002-E-1986-c-2-main).
+#'
+#' @format A data frame containing 16390 rows and 9 variables; each row represents a unique respondent.
+#'   \describe{
+#'   \item{SEQNUM}{Record identification.}
+#'   \item{FWGT_OS}{Final Weight for Main File.}
+#'   \item{DVAGEGR}{Age group of respondent.}
+#'   \item{DVSEX}{Sex of the respondent.}
+#'   \item{DVMARST}{Marital status of the respondent.}
+#'   \item{DVHHLDSZ}{Household size.}
+#'   \item{PROV}{Province of residence of the respondent.}
+#'   \item{DVCHILD}{Number of children.}
+#'}
+#' @docType data
+#' @keywords gss main respondent socio economic 1986
+#' @name gss_main_1986
+#' @usage data(gss_main_1986)
+#' @source [General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) / [Odesi](http://odesi2.scholarsportal.info/webview/)
+"gss_main_1986"
+
+#' Selection of socioeconomic variables from the 1992 GSS Time Use Survey (Cycle 7).
+#'
+#' This dataset contains some socioeconomic variables obtained
+#' from the main file of the 1992 GSS Time Use Survey (Cycle 7). We labeled
+#' variables according to the Public Use Micro Data File Documentation and User's Guide (gss-12M0007-E-1992-c-7-m).
+#'
+#' @format A data frame containing 9815 rows and 9 variables; each row represents a unique respondent.
+#'   \describe{
+#'   \item{SEQNUM}{Record identification.}
+#'   \item{FWGHT}{Final Weight for Main File.}
[…]

### 2024-12-12 paper/Manuscript-Data-Package.Rmd
Edited heading before pushing back to Bruno
Detalling how to use the package in the data paper

@@
 data("gss_impedances")
+data("gss_main_2015")
+
+gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+
+gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
 ```
@@
 
-For {ActiveCA}, data we extracted all data needed to characterize active travel 
+For {ActiveCA} and using GSS Public Use Microdata Files (PUMFs), we extracted all data needed to characterize active travel 
 in Canada, namely, episodes where the activity was identified as moving between 
@@
 decisions in transportation, urban, and health policy.
-
-<!-- 
-is an `R` data package that provides analysis-ready data from the General Social Survey of Canada  [@statisticscanada2024] regarding active travel.
-These data can be used to study walking and cycling in Canada for various purposes over time, and to calculate impedance functions for accessibility analysis.
-The package provides pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
--->
 
@@
 
-<!-- 
-Maybe here provide an example of the structure of this table?
--->
+Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29).
+Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
+
+The main file of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
+For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
+For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.
+
+As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not labeled.
+Adittionaly, the format of the tables (comma-separated values) does not allow for the specification of variable types (whether a variable is continuous or discrete), which can lead to mistakes analysts who have limited experience working with PUMFs.
+
+```{r gss-main-file-2015}
+main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  %>%
+  kable(
+    format = "latex", 
+    booktabs = TRUE, 
+    longtable = TRUE, 
+    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
+    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
+  kable_styling(full_width = FALSE, 
+                #latex_options=c("striped", "scale_down"), 
+                latex_options=c("scale_down"),
+                 font_size = 8) %>%
+  footnote(general = c("Legend: `PUMFID`: record identification.
+`WGHT_PER`:  person weight.
+`SURVMNTH`: survey month of data collection.
+`AGEGR10`: age group of the respondent.
+`SEX`: sex of the respondent.
+`MARSTAT`: marital status of the respondent."), threeparttable = T)
+
+main_2015_unprocessed
+```
 
@@
 
-<!-- 
-Maybe here provide an example of the structure of this table?
--->
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the main file structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 episode file (Cycle 29).
+Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), the episode number (`EPINO`), the activity code (`TUI_01`), the episode's duration (`DURATION`), and the episode's location (`LOCATION`).
 
-# Methods
+In total, the episode file of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables that detail the episodes.
+Similar to the main file, Statistics Canada has created codes for the discrete variables, where each value has a corresponding label.
 
-For each selected cycle of the GSS surveys, we reviewed the episode files to 
-identify episodes of movement that involved walking or cycling. This allowed us 
-to select the activities immediately before and after the movement episode. After
-that, we labeled the code variables with their appropriate descriptions, identifying
-each origin and destination, mode of travel, as well the province and urban 
-classification of the episode.
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description while sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
+The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`.
+Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where the origin and destination were both their home.
+Such trips, where the journey starts and finishes at home, are classified as recreational and leisure trips.
 
-<!--
+In sequence, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
+Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available.
+As shown, since the dataset does not label the variable's values, decoding the values for each variable can be a time-consuming and a challenging task.
 
-Show a graphical example of this process: how were active trips identified? 
-What was the purpose of finding the preceding and succeeding activities?
+```{r gss-epi-file-2015}
+main_2015_unprocessed <- gss_e_2015 |>
+  filter(PUMFID == 10041 & EPINO <= 7) %>%
+  dplyr::select(PUMFID, EPINO,WGHT_EPI,TUI_01, DURATION, LOCATION) %>%
+  kable(
+    format = "latex", 
+    booktabs = TRUE, 
+    longtable = TRUE, 
+    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") %>%
+  kable_styling(full_width = FALSE, 
+                #latex_options=c("striped", "scale_down"), 
+                latex_options=c("scale_down"),
+                 font_size = 8) %>%
+  footnote(general = c("Legend: `PUMFID`: record identification.
+`EPINO`: episode number.
+WGHT_EPI: episode's weight.
+TUI_01: activity code.
+DURATION: episode's duration.
+LOCATION: episode's location."), threeparttable = T)
 
--->
+main_2015_unprocessed
+```
 
-<!--
+# Data extraction
 
-Not needed in this paper:
+For each selected cycle of the GSS surveys, we reviewed the episode files to identify episodes of movement that involved walking or cycling.
+This allowed us to select the activities immediately before and after the movement episode.
+After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, mode of travel, time spent in the active trip, as well the province and urban classification of the episode.
 
-Additionally, for Census Agglomerations and Census Metropolitan Areas we pre-estimated impedance functions for cycling and walking trips.
-Impedance functions reveals important information about the travel behavior of the population, by describing the relationship between the population at an origin and their likelihood or ability to travel to specific destinations to access opportunities [@soukhov2024].
+The active trips (walking and cycling) were identified by the their correspondent activity code, considering the differences between the survey cycles.
+Identifying the proceeding and succeeding activities of each activity episode helps to understand the motivation of the trip and also enables to obtain the origin and destination of the trips.
 
-Impedance functions are commonly used in accessibility analysis to calculate the cost of travel between different locations [@hansen1959; @páez2012; @palacios2022].
-However, these functions need to be calibrated in order to accurately represent travel behavior.
-One effective way of calibrating an impedance function is by using the trip length distribution (TLD) obtained from origin-destination data [@soukhov2024].
-The TLD represents the probability that a certain proportion of trips will be made at a specific cost, such as travel time.
-In this data set, low travel times are associated with a higher proportion of trips, while high travel times are associated with a lower proportion of trips.
+For each GSS main files, we selected some socioeconomic variables for helping profiling the individuals engaged in active trip episodes.
+We selected valuable socioeconomic variables, such as age group of the respondent, sex of the respondent, marital status, number of children, between many others.
+After that, and similar to as we did to episode files, we labeled the code variables with their appropriate descriptions.
 
-For each combination of year, destination, and transportation mode (walking or cycling), we fitted the most suitable impedance function based on empirical data from the GSS surveys.
-We used the `fitdistrplus` package [@mullerdutang2015] to estimate the functions, selecting the distribution with the lowest Akaike information criterion (AIC) among exponential, gamma, log-normal, normal, and uniform types.
+# How to use `ActiveCA`
 
--->
+The code chunk below demonstrates how you can install the `ActiveCA` in R:
 
-# How to use {ActiveCA}
+```{r instal-ActiveCA}
+if (!require("ActiveCA", character.only = TRUE)) {
+      remotes::install_github("dias-bruno/ActiveCA")
+}
+```
 
-<!--
+After installing `ActiveCA` R package, you can load the package according to the following code:
 
-The paper does not really need to report "Results".
-There are too many potential use cases and we would rather let users find their own results depending on what they are interested in.
-Instead of results, a section on "How to use the package" is appropriate.
-There are a number of points to highlight here, for example, how to load the package, call the data sets, in particular the use of episode weights to make sure that any statistical analysis is valid.
+```{r load-ActiveCA, warning=FALSE}
+library(ActiveCA)
+```
 
--->
+By running `?ActiveCA` in `R`, you can access all the package documentation.
+The documentation includes the description of the package, information about the authors and maintainers, and visualize all data sets already processed included in the package.
+
+Figure X, display the documentation for the walking episodes of the 2015 GSS (`walking_2015.rda`).
+The documentation provides a brief description of the data set, how to load the package, also including the description of the format of the file and the description of all variables (already labeled).
+
+After restart your R session, you can access the data set by loading each of them. The chunk below displays how to load 2015 GSS episodes cycles.
+
+```{r load-ActiveCA, warning=FALSE}
+load(walking_2015)
+```
 

### 2024-12-12 paper/Manuscript-Data-Package.tex
Detalling how to use the package in the data paper

@@
 
-For \{ActiveCA\}, data we extracted all data needed to characterize
-active travel in Canada, namely, episodes where the activity was
-identified as moving between an origin and a destination, either by
-walking or cycling. Although Statistics Canada offers Public Use
-Microdata Files and documentation for the GSS program
-\citep[see][]{statisticscanada2024}, accessing these files, and
-preparing them for analysis is not a straightforward matter, given their
-size and complexity. The process of extracting information of interest
-from the source files is time-consuming, tedious, and challenging and/or
-prone to error due to the expertise required to work with these files.
-To create \{ActiveCA\} we collected, cleaned, and processed the cycles
-of the GSS surveys concerning time use to make them ready for analysis.
+For \{ActiveCA\} and using GSS Public Use Microdata Files (PUMFs), we
+extracted all data needed to characterize active travel in Canada,
+namely, episodes where the activity was identified as moving between an
+origin and a destination, either by walking or cycling. Although
+Statistics Canada offers Public Use Microdata Files and documentation
+for the GSS program \citep[see][]{statisticscanada2024}, accessing these
+files, and preparing them for analysis is not a straightforward matter,
+given their size and complexity. The process of extracting information
+of interest from the source files is time-consuming, tedious, and
+challenging and/or prone to error due to the expertise required to work
+with these files. To create \{ActiveCA\} we collected, cleaned, and
+processed the cycles of the GSS surveys concerning time use to make them
+ready for analysis.
 
@@
 
+The table \ref{tab:main-2015-unprocessed} shows the first ten lines and
+the first 6 variables of the GSS PUMF 2015 main file (Cycle 29). Each
+line of the table refers to a respondent in a survey and the columns
+refer respectively to the record identification (\texttt{PUMFID}), the
+person's weight (\texttt{WGHT\_PER}), the month the survey data was
+collected (\texttt{SURVMNTH}), the respondent's age group
+(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
+respondent's marital status (\texttt{MARSTAT}). In total, the main file
+of the 2015 GSS surveys has 17,390 respondents corresponding to
+\ensuremath{2.9766399\times 10^{7}} people and 848 variables. For the
+discrete variables, Statistic Canada has created codes for the possible
+values and each value has its own label. For example, for the variable
+\texttt{SURVMNTH}, the value \texttt{1} means \texttt{January\ 2016},
+while 2 means \texttt{February\ 2016}, 3 is \texttt{March\ 2016} and so
+on. As you can see from the example in Table
+\ref{tab:main-2015-unprocessed}, the variables are not labeled. In
+addition, the format of the tables (comma separated) does not allow you
+to specify the type of each variable (whether it is continuous or
+discrete, for example), which can cause confusion for an analyst with
+relatively little experience in handling PUMFs.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification.
+`WGHT\_PER`:  person weight.
+`SURVMNTH`: survey month of data collection.
+`AGEGR10`: age group of the respondent.
+`SEX`: sex of the respondent.
+`MARSTAT`: marital status of the respondent.
+\end{TableNotes}
+\begin{longtable}[t]{cccccc}
+\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
+\toprule
+PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
+\midrule
+10000 & 616.6740 & 7 & 5 & 1 & 5\\
+10001 & 8516.6140 & 7 & 5 & 1 & 1\\
+10002 & 371.7520 & 1 & 4 & 2 & 1\\
+10003 & 1019.3135 & 3 & 6 & 2 & 5\\
+10004 & 1916.0708 & 9 & 2 & 1 & 6\\
+\addlinespace
+10005 & 1952.2015 & 4 & 1 & 1 & 6\\
+10006 & 5761.5528 & 8 & 1 & 1 & 6\\
+10007 & 466.0426 & 6 & 5 & 2 & 3\\
+10008 & 2479.2991 & 2 & 2 & 2 & 1\\
+10009 & 1436.1641 & 8 & 6 & 1 & 3\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
+
 \subsection{The Episode File}\label{the-episode-file}
@@
 
-\section{Methods}\label{methods}
+Similarly to Table \ref{tab:main-2015-unprocessed} that displayed an
+example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
+shows the first seven episodes for the record identification number
+\texttt{10041} and some variables of the GSS PUMF 2015 episode file
+(Cycle 29). Each line of the table refers to an episode of the
+especified record identification (\texttt{PUMFID\ =\ 10041}), the
+episode's weight (\texttt{WGHT\_EPI}), the episode number
+(\texttt{EPINO}), the activity code (\texttt{TUI\_01}), the episode's
+duration (\texttt{DURATION}) and the episode's location
+(\texttt{LOCATION}). In total, the episode file of the 2015 GSS surveys
+has 274,108 records corresponding to \ensuremath{4.6183762\times 10^{8}}
+episodes and 527 variables that detail the episodes. As similar to the
+main file, Statistic Canada has created codes for the possible values of
+the discrete variables with each value has its own label. Considering
+the case displayed in Table \ref{tab:ep-2015-unprocessed}, this person
+started the diary description while sleeping at home
+(\texttt{TUI\_01\ =\ 1} and \texttt{LOCATION\ =\ 300}) for 210 minutes,
+then engaged in personal hygiene for 40 minutes
+(\texttt{TUI\_01\ =\ 2}), then spent 15 minutes getting personal care
+(which can be getting ready for school, supervising homework, reading,
+playing, reprimanding, educational, or emotional help, sinalized by the
+\texttt{TUI\_01\ =\ 27}). After this, this person had an activity travel
+episode, walking for 15 minutes (\texttt{TUI\_01\ =\ 7} and
+\texttt{LOCATION\ =\ 315}) in which the origin and destination was their
+own house. Cases like this, where the trip starts and finishes at home
+refers to recreational and leisure trips. After that, the person spent 3
+hours looking for a job (\texttt{TUI\_01\ =\ 9}), had break for a lunch
+of 15 minutes duration (\texttt{TUI\_01\ =\ 6}) and cleaned the house
+(\texttt{TUI\_01\ =\ 18}) for two hours. Table
+\ref{tab:ep-2015-unprocessed} shows only six variables from the 527
+possible cases. As you can see, since that data set is not labeled,
+figuring out the codes for every variable can be time consuming and
+difficult.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\end{TableNotes}
+\begin{longtable}[t]{cccccc}
+\caption{\label{tab:gss-epi-file-2015}\label{tab:main-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.}\\
+\toprule
+PUMFID & EPINO & WGHT\_EPI & TUI\_01 & DURATION & LOCATION\\
+\midrule
+10041 & 1 & 1353.818 & 1 & 210 & 300\\
+10041 & 2 & 1353.818 & 2 & 40 & 300\\
+10041 & 3 & 1353.818 & 27 & 15 & 300\\
+10041 & 4 & 1353.818 & 7 & 15 & 315\\
+10041 & 5 & 1353.818 & 9 & 180 & 300\\
+\addlinespace
+10041 & 6 & 1353.818 & 6 & 15 & 300\\
+10041 & 7 & 1353.818 & 18 & 120 & 300\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
+
+\section{Data extraction}\label{data-extraction}
 

### 2024-12-12 R/ActiveCA-package.R

@@ -704,14 +704,14 @@ NULL
 #'   \item{PUMFID}{Record identification.}
 #'   \item{WGHT_PER}{Person weight.}
 #'   \item{SURVMNTH}{Survey month of data collection.}
-#'   \item{AGEGR10}{Age group of the respondent (group of 10)..}
-#'   \item{SEX}{Sex of respondent..}
-#'   \item{MARSTAT}{Marital status of the respondent..}
+#'   \item{AGEGR10}{Age group of the respondent (group of 10).}
+#'   \item{SEX}{Sex of respondent.}
+#'   \item{MARSTAT}{Marital status of the respondent.}
 #'   \item{PHSDFLG}{Respondent has a spouse/partner living in the household.}
-#'   \item{AGEPRGRD}{Age difference between respondent and spouse/partner..}
-#'   \item{SEXPR}{Sex of the respondent's spouse/partner within the household..}
-#'   \item{PRTYPEC}{Sex of the respondent's spouse/partner within the household..}
-#'   \item{CXRFLAG}{Type of partner the respondent has within the household..}
+#'   \item{AGEPRGRD}{Age difference between respondent and spouse/partner.}
+#'   \item{SEXPR}{Sex of the respondent's spouse/partner within the household.}
+#'   \item{PRTYPEC}{Sex of the respondent's spouse/partner within the household.}
+#'   \item{CXRFLAG}{Type of partner the respondent has within the household.}
 #'   \item{AGECXRYG}{Age of respondent’s youngest child in household.}
 #'   \item{AGEHSDYC}{Age of youngest household member in respondent’s household.}
 #'   \item{CHINHSDC}{Respondent’s child(ren) in household - Any age/marital status.}
@@ -723,12 +723,12 @@ NULL
 #'   \item{LIVARR11}{Living arrangement of respondent's household (in group of 8).}
 #'   \item{HSDSIZEC}{Household size of respondent.}
 #'   \item{HSDELIGC}{Number of members in respondent’s household 15 years of age or older.}
-#'   \item{MULTIGEN}{Three generations or more family in respondent's household..}
-#'   \item{PRV}{Province of residence of the respondent..}
-#'   \item{LUC_RST}{Population centres indicator..}
+#'   \item{MULTIGEN}{Three generations or more family in respondent's household.}
+#'   \item{PRV}{Province of residence of the respondent.}
+#'   \item{LUC_RST}{Population centres indicator.}
[…]

### 2024-12-13 paper/Manuscript-Data-Package.Rmd
Including labelled main files in the data paper

@@
 
+knitr::knit_hooks$set(inline = function(x) {
+  if (is.numeric(x)) {
+    format(x, big.mark = ",", scientific = FALSE)
+  } else {
+    x
+  }
+})
+
 opts <- options(knitr.kable.NA = "")
@@
 library(stats)
+library(Hmisc)
 ```
@@
 
+# Original main and episode files from 2015 to display table examples
+
 gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
 
-gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
+# Processed main files
+data("gss_main_2015")
+data("gss_main_2010")
+data("gss_main_2005")
+data("gss_main_1998")
+data("gss_main_1992")
+data("gss_main_1986")
 ```
@@
 Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the main file structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 episode file (Cycle 29).
-Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), the episode number (`EPINO`), the activity code (`TUI_01`), the episode's duration (`DURATION`), and the episode's location (`LOCATION`).
+Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
 
-In total, the episode file of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables that detail the episodes.
-Similar to the main file, Statistics Canada has created codes for the discrete variables, where each value has a corresponding label.
+In total, the episode file of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
+Similar to the main file, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
 
-In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description while sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
 The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`.
-Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where the origin and destination were both their home.
-Such trips, where the journey starts and finishes at home, are classified as recreational and leisure trips.
-
-In sequence, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
+Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where both the origin and destination were their home.
+Such trips, where the journey starts and finishes at home, can be classified as recreational or leisure trips.
+Next, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
 Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available.
-As shown, since the dataset does not label the variable's values, decoding the values for each variable can be a time-consuming and a challenging task.
+As shown, since the dataset does not label the variable values, decoding them can be both time-consuming and challenging.
 
 ```{r gss-epi-file-2015}
-main_2015_unprocessed <- gss_e_2015 |>
+ep_2015_unprocessed <- gss_e_2015 |>
   filter(PUMFID == 10041 & EPINO <= 7) %>%
@@
 
-main_2015_unprocessed
+ep_2015_unprocessed
 ```
@@
 
-The active trips (walking and cycling) were identified by the their correspondent activity code, considering the differences between the survey cycles.
-Identifying the proceeding and succeeding activities of each activity episode helps to understand the motivation of the trip and also enables to obtain the origin and destination of the trips.
+The active trips (walking and cycling) were identified by their corresponding activity codes, accounting for differences between the survey cycles.
+Identifying the preceding and succeeding activities of each activity episode helps to understand the motivation for the trip and also determine the origin and destination of the trips.
 
-For each GSS main files, we selected some socioeconomic variables for helping profiling the individuals engaged in active trip episodes.
-We selected valuable socioeconomic variables, such as age group of the respondent, sex of the respondent, marital status, number of children, between many others.
-After that, and similar to as we did to episode files, we labeled the code variables with their appropriate descriptions.
+For each GSS main file, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes.
+We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others.
+After this, as we did with the episode files, we labeled the code variables with their appropriate descriptions.
 
@@
 
-The code chunk below demonstrates how you can install the `ActiveCA` in R:
+This section presents some potential applications of the `ActiveCA` R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
 
-```{r instal-ActiveCA}
-if (!require("ActiveCA", character.only = TRUE)) {
+To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the `ActiveCA` package in R:
+
+```{r instal-ActiveCA, echo=TRUE}
+if(!require("ActiveCA", character.only = TRUE)) {
       remotes::install_github("dias-bruno/ActiveCA")
@@
 
-After installing `ActiveCA` R package, you can load the package according to the following code:
+After installing the `ActiveCA` R package, you can load it using the following code:
 
-```{r load-ActiveCA, warning=FALSE}
+```{r load-ActiveCA, echo=TRUE}
 library(ActiveCA)
@@
 
-By running `?ActiveCA` in `R`, you can access all the package documentation.
-The documentation includes the description of the package, information about the authors and maintainers, and visualize all data sets already processed included in the package.
+By running `?ActiveCA` in `R`, you can access the complete package documentation.
+The documentation includes a description of the package, information about the authors and maintainers, and an overview of all pre-processed data sets included in the package.
 
-Figure X, display the documentation for the walking episodes of the 2015 GSS (`walking_2015.rda`).
-The documentation provides a brief description of the data set, how to load the package, also including the description of the format of the file and the description of all variables (already labeled).
+Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (`walking_2015.rda`).
+The documentation provides a brief description of the dataset, instructions on how to load the package, as well as details about the file format and descriptions of all variables (already labeled).
 
-After restart your R session, you can access the data set by loading each of them. The chunk below displays how to load 2015 GSS episodes cycles.
+```{r walking_documentation, echo=FALSE, out.width="100%", fig.cap=paste0("Documentation for the dataset of walking episodes from 2015."), fig.align="center"}
 
-```{r load-ActiveCA, warning=FALSE}
-load(walking_2015)
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking-2015-documentation.png"))
+```
+
+The chunk below shows how to load the walking episodes from the 2015 GSS (Cycle 29):
+
+```{r load-walking-2015-episodes, echo=TRUE}
+data(walking_2015)
+```
+
+The GSS surveys apply a probability sampling methodology, in which each episode or person selected in the sample represents several other episodes or persons not in the sample.
+The number of episodes and persons represented by a episode or person is determined by the weight or weighting factor.
+Because of this, every estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
+
+For instance, if someone wants to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weights.
+In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group.
+It uses the dplyr package to manipulate the data.
+
+The process begins by creating a dataset that sums the population by age group.
+Then, it joins the 2015 episodes with the 2015 main file.
+Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
+After this, a new dataset, `Active_percentage`, is created by merging both previous datasets.
+The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100.
+
+The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 37%, followed by those between 25 and 34 years old with around 33%.
+There is a significant drop in percentage for the following groups, with the percentage falling to between 15% and 18% for the remaining age groups.
+
+```{r percentages-act, echo=TRUE}
+# Load the dplyr library for data manipulation
+library('dplyr')
+
+# Load the episodes dataset
+data(gss_episodes)
+
+# Calculate the total population for each age group
+Total_population <- gss_main_2015 %>% 
+  group_by(AGEGR10) %>% 
+  summarise(Total_population  = sum(WGHT_PER))
+
+# Calculate the active population for each age group
+Active_population <- gss_episodes %>% 
+      filter(YEAR == 2015) %>% 
+  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
+  group_by(AGEGR10) %>% 
+  summarise(Active_population  = sum(WGHT_PER))
+
+# Combine total and active population and calculate the percentage
+Active_percentage <- Total_population %>% 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
+  mutate(Percentage = round(Active_population/Total_population  * 100, 2))
+
+# Display the final result
+Active_percentage
+```
+
+## Processed data set
+
+```{r processed-files, include=FALSE}
+df_main_processed <- data.frame(
+  'Survey' = c("Main","Main","Main","Main","Main","Main"),
+  'Year' = c(2015, 2010, 2005, 1998, 1992, 1986),
+  'Count' = c(nrow(gss_main_2015),
+              nrow(gss_main_2010),
+              nrow(gss_main_2005),
+              nrow(gss_main_1998),
+              nrow(gss_main_1992),
+              nrow(gss_main_1986)),
+  'Weighted' = c(sum(gss_main_2015$WGHT_PER),
+              sum(gss_main_2010$WGHT_PER),
+              sum(gss_main_2005$WGHT_PER),
+              sum(gss_main_1998$WGHTFIN),
+              sum(gss_main_1992$FWGHT),
+              sum(gss_main_1986$FWGT_OS/10000)))
+
+df_ep_processed <- gss_episodes %>% 
+  group_by(YEAR) %>%
+  summarise(Survey = 'Episode', 
+            Count = n(), 
+            Weighted = sum(WGHT_EPI)) %>%
+  rename(Year = YEAR) %>%
+  dplyr::select(Survey, Year, Count, Weighted) %>% 
+  arrange(-Year)
+
+df_processed <- rbind(df_main_processed,df_ep_processed)
+```
+
+Table \ref{tab:processed-obs} displays the total number of observations processed for main and episode files.
+For the main files, a total of `r sum(df_main_processed$Count)` observations were processed, referring to all observations from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
+Table X also presents the total cases of active trips episodes identified.
+In total `r sum(df_ep_processed$Count)` observations with register of active travel activity.
+Together, these observations account for `r sum(df_ep_processed$Weighted)` episodes (`r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1986).
+
+```{r table_df_processed}
+
+table_df_processed <- df_processed %>%
+  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
+        caption = "\\label{tab:processed-obs}Total number and weighted sum of observations processed.") %>%
+  kable_styling(full_width = FALSE, font_size = 10) %>%
+  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) %>%
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major")
+
+table_df_processed
+```
+
+Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
+Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 episode file (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Only active travel episodes appear in Table \ref{tab:ep-2015-processed} since the observations were filtered to select cases with walking or cycling episodes.
+For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.
+
+```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
+main_2015_processed <- gss_main_2015[1:10, 1:6]  %>%
+  kable(
+    format = "latex", 
+    booktabs = TRUE, 
+    longtable = TRUE, 
+    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
+    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
+  kable_styling(full_width = FALSE, 
+                #latex_options=c("striped", "scale_down"), 
+                latex_options=c("scale_down"),
+                 font_size = 8) %>%
+  footnote(general = c("Legend: `PUMFID`: record identification.
+`WGHT_PER`:  person weight.
+`SURVMNTH`: survey month of data collection.
+`AGEGR10`: age group of the respondent.
+`SEX`: sex of the respondent.
+`MARSTAT`: marital status of the respondent."), threeparttable = T)
+
+main_2015_processed
+
+ep_2015_unprocessed <- gss_episodes_renamed |>
+  filter(PUMFID == 10041 & YEAR == 2015) %>%
+  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) %>%
+  kable(
+    format = "latex", 
+    booktabs = TRUE, 
+    longtable = TRUE, 
+    caption = "\\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.", 
+       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) %>% 
+  kable_styling(full_width = FALSE, 
+                #latex_options=c("striped", "scale_down"), 
+                latex_options=c("scale_down"),
+                 font_size = 8) %>%
+  footnote(general = c("Legend: `PUMFID`: record identification.
+`EPINO`: episode number.
+WGHT_EPI: episode's weight.
+TUI_01: activity code.
+DURATION: episode's duration.
+LOCATION: episode's location."), threeparttable = T)
+
+ep_2015_unprocessed
 ```
@@
 `ActiveCA` includes all these episodes ready for analysis.
-Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, including metrics such as the count
-of recorded trips (count), and measures of trip duration in minutes:
-maximum (max), mean, median, and minimum (min). The 1986 survey did not
+Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes: maximum (max), mean, median, and minimum (min).
+The 1986 survey did not
 include bicycle trips.
@@
 ```{r table-01, echo=FALSE, cache=FALSE}
-# Calculate the count of trips for each mode and year
-episodes_counts <- gss_episodes %>%
+# Calculate min, mean, and max statistic (duration)
+stats_data <- gss_episodes_renamed %>%
   group_by(YEAR, MODE) %>%
-  summarise(Count = n(), .groups = "drop") %>%
-  pivot_wider(names_from = YEAR, values_from = Count)
-
-# Calculate min, mean, and max statictis (duration)
-stats_data <- gss_episodes %>%
-  group_by(YEAR, MODE) %>%
-  summarise(
+  dplyr::summarise(
     min = min(DURATION),
-    mean = mean(DURATION),
-    median = median(DURATION),
+    mean = sum(DURATION * WGHT_EPI)/sum(WGHT_EPI),
+    median =  Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
     max = max(DURATION),
@@
 # Combine counts and stats into one data frame
-final_data <- bind_rows(
-  episodes_counts %>% mutate(Statistic = "Count"),
-  stats_data
-) %>%
+stats_data <- stats_data %>%
   mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) %>%
@@
 # Rename the 'MODE' column to 'Mode' before generating the table
-final_data <- final_data%>% 
-  arrange(MODE) %>%
+stats_data <- stats_data %>%
   rename(Mode = MODE) %>%
@@
     Statistic == "std" ~ "Standard deviation",
-    TRUE ~ Statistic  # Keep all other values as they are
-  ))
+    TRUE ~ Statistic))
 
 # Creating a latex table
-styled_ch03_table_01 <- final_data %>%
+styled_ch03_table_01 <- stats_data %>%
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
         caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
-        align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
+        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) %>%
   kable_styling(full_width = FALSE, font_size = 10) %>%
   column_spec(1, bold = TRUE) %>%
-  add_header_above(c(" " = 2, "Year" = ncol(final_data) - 2)) %>%
+  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) %>%
   collapse_rows(columns = 1, valign = "top", latex_hline = "major")
@@
 
-Table \ref{tab:table-01} shows that the median values for walking trips
-range between 5 and 10 minutes, while cycling trips have a consistent
-median of 15 minutes since 1998. The table also highlights very high
-maximum values, particularly for walking trips, with recorded episodes
-exceeding 4 hours in all cases.
+Table \ref{tab:table-01} shows that, in general, the median values for walking trips is 10 minutes, except for 1998 when the median was 5 minutes.
+In the case of cycling trips, the duration fluctuated over the years, ranging from 10 to 20 minutes.
+The table also highlights very high maximum values, particularly for walking trips, with recorded episodes exceeding 4 hours in all cases.
 
@@
 transition that also occurred for cycling trips between 1992 and 1998.
-Table \ref{tab:table-03} indicates that the median duration for trips to
+Table \ref{tab:table-03} indicates that the median duration for walking trips to
 `home` and `work or school` remained at 10 minutes.
@@
   group_by(YEAR, dest_label, MODE) %>%
-  summarise(
-    count = n(),
+  dplyr::summarise(
+    count = sum(WGHT_EPI),
     min = min(DURATION),
-    median = median(DURATION),
+    median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
     max = max(DURATION),
@@
     caption = "\\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998",
-    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
+    col.names = c("Destination", "Mode","Min","Med","Max", "(%)", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
   ) %>%
@@
   collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
-  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
+  footnote(general = c("'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
 
@@
   group_by(YEAR, dest_label, MODE) %>%
-  summarise(
-    count = n(),
+   dplyr::summarise(
+    count = sum(WGHT_EPI),
     min = min(DURATION),
-    median = median(DURATION),
+    median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
     max = max(DURATION),
@@
     caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
-    col.names = c("Destination", "Mode","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
+    col.names = c("Destination", "Mode","Min","Med","Max", "(%)", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
   ) %>%
@@
 
-{ActiveCA} also enables visual analysis of active travel in Canada using traditional 
-exploratory data analysis techniques. Figures \ref{fig:figure-01} and \ref{fig:figure-02}
-show walking and cycling trips from 1992 and 2015 through heat maps. These maps 
-use color gradients to represent the percentage of trips between various origins
-and destinations, with darker colors indicating higher percentages and lighter 
-colors representing less frequent routes. For conciseness, we omitted the heat maps
-for the other years analyzed.
-
-```{=html}
-<!---
-I`m not surte if this last phrase makes sense in english! "To avoid overwhelming the reader", "Para nao sobrecarregar o leitor" 
--->
-```
+{ActiveCA} also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques.
+Figures \ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling trips from 1992 and 2015 through heat maps.
+These maps use color gradients to represent the percentage of trips between various origins and destinations, with darker colors indicating higher percentages and lighter colors representing less frequent routes.
+For conciseness, we omitted the heat maps for the other years analyzed.
 
@@
       group_by(orig_label, dest_label) %>%
-      summarize(count_each_trip = n(), 
+        dplyr::summarise(count_each_trip = n(), 
                 weighted_count = sum(WGHT_EPI), .groups = "drop")
@@
 
-# Python integration
+`ActiveCA` also enables obtaining insights from the main processed files.
+Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2015.
+According to this plot, married respondents reported the highest level of stress, with 17% of possible cases.
 
-CAN WE PROVIDE PYTHON INTEGRATION? See here:
+```{r main_stress_figure}
+main_stress_figure <- gss_main_2015 %>%
+  filter(!is.na(MARSTAT), !is.na(GTU_110)) %>%
+  group_by(MARSTAT, GTU_110) %>%
+  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") %>%
+  mutate(percentage = n / sum(n) * 100) %>%
+  ungroup() %>%
+  ggplot(aes(x = MARSTAT,
+    y = GTU_110)) +
+    geom_tile(aes(fill = percentage)) +
+    labs(x = "Marital status of the respondent",
+    y = "How often do you feel rushed?",
+    fill = "Percentage (%)") + 
+    geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
+    theme_bw() + 
+    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd")) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 
-https://github.com/paezha/idealista18/tree/master/Python
-
-<!--
-
-## Pre-estimated impedance functions for active travel in Canada
-
-`ActiveCA` includes a total of
-`r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
-pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
-Table \ref{tab:table-04} shows the best impedance functions for walking trips to `Work or school` across survey years, where all distributions, except for a gamma function in 1998, are log-normal.
-
-```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
-# selecting mode and destination to display impedance functions
-mode <- c("Walking","Cycling")
-destino <- "Work or school"
-
-# filtering the impedance functions in accordance with the mode/destination setted in the step before and creating a latex table 
-selected_functions <- gss_impedances_renamed %>%
-  filter(dest_label == destino, MODE %in% mode) %>%
-  group_by(MODE, YEAR, dest_label) %>% 
-  summarise(Function = first(f_name), 
-            est_1 = first(est_1),
-            est_2 = first(est_2),
-            AIC = first(aic),
-            f_name = first(f_name),
-            count = first(count),
-            .groups = "drop") %>% 
-  mutate(Function = case_when(
-    Function == "lnorm" ~ "Lognormal",
-    Function == "gamma" ~ "Gamma",
-    Function == "unif" ~ "Uniform",
-    TRUE ~ Function), 
-  across(c(est_1,est_2), ~ round(., digits = 2)), 
-  across(c(AIC), ~ round(., digits = 0))) %>% 
-  rename(Mode = MODE, Year = YEAR, Destination = dest_label) %>%
-  unique() %>% 
-  arrange(Mode, Year) %>%
-  select(Mode, Year, Function, est_1, est_2, AIC, count)
-
-selected_functions_tab <- selected_functions %>% 
-  kable(format = "latex",
-        booktabs = TRUE,
-        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
-        col.names = c("Mode", "Year", "Impedance function","Parameter 1*","Parameter 2*", "AIC","Count")) %>% 
-    kable_styling(full_width = FALSE, 
-                latex_options=c("scale_down"),
-                font_size = 10) %>% 
-  row_spec(0, bold = TRUE, align = "c") %>% 
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>%
-  footnote(general = c("*For lognormal distributions, 'Parameter 1' and 'Parameter 2' refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively.
-For gamma distribution, 'Parameter 1' and 'Parameter 2' refer to the rate and shape of the distribution, respectively.
-`AIC` means Akaike information criterion."), threeparttable = T)
-
-selected_functions_tab
+# Save the combined plot to a file
+ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"), 
+       plot = main_stress_figure,
+       width = 15, 
+       height = 10, 
+       units = "cm", 
+       dpi = 300)
 ```
 
-Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
-alongside histograms of empirical travel times (grey bars). As for the
-others pre-estimated functions, these examples allow to calculate gravity-based
-accessibility measures for active travel across various destinations and temporal 
-scales in Canadian urban areas.
+```{r figure-stress, echo=FALSE, out.width="100%", fig.cap=paste0("Level of stress among respondents of different marital statuses (2015)."), fig.align="center"}
 
-```{r theoretical_impedance_values, include = FALSE}
-theoretical_impedance <- function(df_functions, travel_cost) {
-
-  # Initialize the main data frame to store results
-  result <- data.frame()
-
-  # Create the 't' sequence
-  t_seq <- seq(1, travel_cost, 1)
-
-  for (mode in unique(df_functions$Mode)){
-
-    for (year in unique(df_functions[df_functions$Mode == mode, ]$Year)){
-
-      for (destino in unique(df_functions[df_functions$Year == year & df_functions$Mode == mode, ]$Destination)){
-
-      name_function <- df_functions[df_functions$Year == year & 
-                                      df_functions$Mode == mode & 
-                                      df_functions$Destination == destino, ]$f_name
-
-      print(paste0("YEAR: ", year, " MODE: ", mode," Destination: ", destino))
-
-      subset <- df_functions %>% 
-        filter(Year == year & Mode == mode & Destination == destino)
-
-      # Temporary data frame for this iteration
-      temp_df <- data.frame(t = t_seq)
-
-      # Add the appropriate distribution
-      if (name_function == "lnorm") {
-        temp_df$f <- dlnorm(temp_df$t,
-                            meanlog = subset$est_1[1],
-                            sdlog = subset$est_2[1])
-
-      } else if (name_function == "gamma") {  
-        temp_df$f <- dgamma(temp_df$t,
-                            shape = subset$est_1[1],
-                            rate = subset$est_2[1])
-
-      } else if (name_function == "norm") {  
-        temp_df$f <- dnorm(temp_df$t,
-                           mean = subset$est_1[1],
-                           sd = subset$est_2[1])
-
-      } else if (name_function == "exp") {  
-        temp_df$f <- dexp(temp_df$t,
-                          rate = subset$est_1[1])
-
-      } else if (name_function == "unif") {  
-        temp_df$f <- dunif(temp_df$t,
-                           min = subset$est_1[1],
-                           max = subset$est_2[1])
-      }
-
-      # Add mode, year and destination columns
-      temp_df$MODE <- mode
-      temp_df$YEAR <- year
-      temp_df$dest_label <- destino
-
-      # Append to the result data frame
-      result <- rbind(result, temp_df)
-    }
-  }
-}
-  return(result)
-}
-
-max_duration <- 100
-
-selected_functions <- gss_impedances_renamed %>% 
-  group_by(MODE, YEAR, dest_label) %>% 
-  summarise(Function = first(f_name), 
-            est_1 = first(est_1),
-            est_2 = first(est_2),
-            AIC = first(aic),
-            f_name = first(f_name),
-            count = first(count),
-            .groups = "drop") %>% 
-  mutate(Function = case_when(
-    Function == "lnorm" ~ "Lognormal",
-    Function == "gamma" ~ "Gamma",
-    Function == "unif" ~ "Uniform",
-    TRUE ~ Function), 
-  across(c(est_1,est_2), ~ round(., digits = 2)), 
-  across(c(AIC), ~ round(., digits = 0))) %>% 
-  rename(Mode = MODE, Year = YEAR, Destination = dest_label)
-
-theoretical_values <- theoretical_impedance(selected_functions, travel_cost = max_duration)
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"))
 ```
 
-```{r include = FALSE}
-
-for(destino in unique(gss_impedances_renamed$dest_label)){
-
-  medianas <- gss_impedances_renamed %>%
-    filter(dest_label == destino) %>%
-    group_by(MODE) %>%
-    summarise(mediana_duracao = median(DURATION, na.rm = TRUE), .groups = 'drop')
-
-  maximos <- theoretical_values %>%
-    filter(dest_label == destino) %>%
-    group_by(MODE, YEAR) %>%
-    filter(f == max(f)) %>%
-    slice(which.max(t)) %>%
-    ungroup()
-
-combined_plot <- gss_impedances_renamed %>% 
-  filter(dest_label == destino) %>%
-  mutate(dest_label = factor(dest_label, levels = destination_order)) %>%
-  ggplot(aes(x= DURATION)) +
-  geom_histogram(aes(y = ..density.., colour = MODE, fill = MODE), bins = 20, alpha=0.1, position="identity") + 
-  geom_line(data = theoretical_values[theoretical_values$dest_label == destino, ], aes(x = t, y = f, colour = MODE), size = 0.8) +   
-  geom_vline(data = medianas, aes(xintercept = mediana_duracao, color = MODE), linetype = "dashed") + 
-  geom_point(data = maximos, aes(x = t, y = f, colour = MODE), size = 3) + 
-  geom_label(data = maximos, aes(x = t, y = f, label = round(t, 2), colour = MODE), 
-            vjust = 0, size = 3) +
-  theme_minimal() +
-  facet_wrap(~YEAR) + 
-  labs(x = "Trip duration (in minutes)", y = "Density", colour = "Transportation Mode", fill = "Transportation Mode") +
-  theme(legend.position="right", 
-    strip.text = element_text(face = "bold", hjust = 0))
-
-  plots <- length(unique(gss_impedances_renamed[gss_impedances_renamed$dest_label == destino, ]$YEAR))
-
-  # Case of only one year 
-    if(plots == 1){
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
-       plot = combined_plot,
-       width = 7, 
-       height = 4, 
-       units = "in", 
-       dpi = 300)
-
-    # Destination present in 3 surveys  
-    } else if(plots == 3){
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
-       plot = combined_plot,
-       width = 14, 
-       height = 4, 
-       units = "in", 
-       dpi = 300)
-
-  # Destination present in 5 surveys  
-    }else {
-      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
-       plot = combined_plot,
-       width = 14, 
-       height = 9, 
-       units = "in", 
-       dpi = 300)
-    }
-}
-```
-
-```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Empirical data and impedance functions fitted for `work or school` as destination."), fig.align="center"}
-
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/impf_Work or school.jpg"))
-```
-
--->
+# Python integration
 
@@
 
-<!--
-
-Additionally, the package includes a series of pre-estimated impedance functions 
-for walking and cycling trips, considering various destinations and time periods, 
-for Canadian Metropolitan and Census Agglomeration areas.
-
--->
-
 The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
 which facilitates the addition of complementary data sets in the future. `R` users
-can seamlessly explore GSS walking and cycling episodes <!-- along with calibrated impedance functions-->,
+can seamlessly explore GSS walking and cycling episodes, 
 with the option to suggest enhancements to the package as needed. This article adopts

### 2024-12-13 paper/Manuscript-Data-Package.tex
Including labelled main files in the data paper

--- 2024-12-04 paper/Manuscript-Data-Package.tex
@@
 
+Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
+six variables of the GSS PUMF 2015 main file (Cycle 29). Each row in the
+table correspond to a survey respondent, while the columns refer the
+following information: record identification (\texttt{PUMFID}), the
+person's weight (\texttt{WGHT\_PER}), the month the survey data was
+collected (\texttt{SURVMNTH}), the respondent's age group
+(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
+respondent's marital status (\texttt{MARSTAT}).
+
+The main file of the 2015 GSS surveys includes a total of 17,390
+respondents, representing 29,766,399 individuals and 848 variables. For
+discrete variables, Statistics Canada has assigned specific codes to the
+possible values, with each code accompanied by a label. For instance, in
+the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
+\texttt{January\ 2016}, 2 means \texttt{February\ 2016}, 3 corresponds
+to \texttt{March\ 2016}, and so on.
+
+As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
+labeled. Adittionaly, the format of the tables (comma-separated values)
+does not allow for the specification of variable types (whether a
+variable is continuous or discrete), which can lead to mistakes analysts
+who have limited experience working with PUMFs.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification.
+`WGHT\_PER`:  person weight.
+`SURVMNTH`: survey month of data collection.
+`AGEGR10`: age group of the respondent.
+`SEX`: sex of the respondent.
+`MARSTAT`: marital status of the respondent.
+\end{TableNotes}
+\begin{longtable}[t]{cccccc}
+\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
+\toprule
+PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
+\midrule
+10000 & 616.6740 & 7 & 5 & 1 & 5\\
+10001 & 8516.6140 & 7 & 5 & 1 & 1\\
+10002 & 371.7520 & 1 & 4 & 2 & 1\\
+10003 & 1019.3135 & 3 & 6 & 2 & 5\\
+10004 & 1916.0708 & 9 & 2 & 1 & 6\\
+\addlinespace
+10005 & 1952.2015 & 4 & 1 & 1 & 6\\
+10006 & 5761.5528 & 8 & 1 & 1 & 6\\
+10007 & 466.0426 & 6 & 5 & 2 & 3\\
+10008 & 2479.2991 & 2 & 2 & 2 & 1\\
+10009 & 1436.1641 & 8 & 6 & 1 & 3\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
+
 \subsection{The Episode File}\label{the-episode-file}
@@
 
-\section{Methods}\label{methods}
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
+example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
+presents the first seven episodes for the record identification number
+\texttt{10041} and some variables from the GSS PUMF 2015 episode file
+(Cycle 29). Each row in the table corresponds to an episode associated
+with the specified record identification (\texttt{PUMFID\ =\ 10041}),
+including the episode's weight (\texttt{WGHT\_EPI}), episode number
+(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
+(\texttt{DURATION}), and episode location (\texttt{LOCATION}).
+
+In total, the episode file of the 2015 GSS surveys contains 274,108
+records, representing 461,837,622 episodes and 527 variables detailing
+the episodes. Similar to the main file, Statistics Canada has created
+codes for the discrete variables, with each value corresponding to a
+label.
+
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this
+respondent began the diary description by sleeping at home
+(\texttt{TUI\_01\ =\ 1} and \texttt{LOCATION\ =\ 300}) for 210 minutes,
+followed by 40 minutes of personal hygiene (\texttt{TUI\_01\ =\ 2}). The
+respondent then spent 15 minutes on personal care activities, such as
+getting ready for school, supervising homework, reading, playing,
+reprimanding, or providing educational or emotional support, as
+indicated by \texttt{TUI\_01\ =\ 27}. Next, they recorded a travel
+episode, walking for 15 minutes (\texttt{TUI\_01\ =\ 7} and
+\texttt{LOCATION\ =\ 315}), where both the origin and destination were
+their home. Such trips, where the journey starts and finishes at home,
+can be classified as recreational or leisure trips. Next, the respondent
+spent 3 hours searching for a job (\texttt{TUI\_01\ =\ 9}), took a
+15-minute lunch break (\texttt{TUI\_01\ =\ 6}), and then cleaned the
+house (\texttt{TUI\_01\ =\ 18}) for two hours. Table
+\ref{tab:ep-2015-unprocessed} displays only six variables out of the 527
+available. As shown, since the dataset does not label the variable
+values, decoding them can be both time-consuming and challenging.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\end{TableNotes}
+\begin{longtable}[t]{cccccc}
+\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.}\\
+\toprule
+PUMFID & EPINO & WGHT\_EPI & TUI\_01 & DURATION & LOCATION\\
+\midrule
+10041 & 1 & 1353.818 & 1 & 210 & 300\\
+10041 & 2 & 1353.818 & 2 & 40 & 300\\
+10041 & 3 & 1353.818 & 27 & 15 & 300\\
+10041 & 4 & 1353.818 & 7 & 15 & 315\\
+10041 & 5 & 1353.818 & 9 & 180 & 300\\
+\addlinespace
+10041 & 6 & 1353.818 & 6 & 15 & 300\\
+10041 & 7 & 1353.818 & 18 & 120 & 300\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
+
+\section{Data extraction}\label{data-extraction}
 
@@
 their appropriate descriptions, identifying each origin and destination,
-mode of travel, as well the province and urban classification of the
-episode.
+mode of travel, time spent in the active trip, as well the province and
+urban classification of the episode.
 
-\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}
+The active trips (walking and cycling) were identified by their
+corresponding activity codes, accounting for differences between the
+survey cycles. Identifying the preceding and succeeding activities of
+each activity episode helps to understand the motivation for the trip
+and also determine the origin and destination of the trips.
+
+For each GSS main file, we selected socioeconomic variables to help
+profile the individuals engaged in active trip episodes. We included key
+socioeconomic variables such as the respondent's age group, sex, marital
+status, and number of children, among others. After this, as we did with
+the episode files, we labeled the code variables with their appropriate
+descriptions.
+
+\section{\texorpdfstring{How to use
+\texttt{ActiveCA}}{How to use ActiveCA}}\label{how-to-use-activeca}
+
+This section presents some potential applications of the
+\texttt{ActiveCA} R package. In fact, we expect that the application of
+this package to extend beyond our pre-imagined range of uses.
+
+To start the demonstration, it is first necessary to install the
+package. The code chunk below demonstrates how to install the
+\texttt{ActiveCA} package in R:
+
+\begin{verbatim}
+if(!require("ActiveCA", character.only = TRUE)) {
+      remotes::install_github("dias-bruno/ActiveCA")
+}
+\end{verbatim}
+
+After installing the \texttt{ActiveCA} R package, you can load it using
+the following code:
+
+\begin{verbatim}
+library(ActiveCA)
+\end{verbatim}
+
+By running \texttt{?ActiveCA} in \texttt{R}, you can access the complete
+package documentation. The documentation includes a description of the
+package, information about the authors and maintainers, and an overview
+of all pre-processed data sets included in the package.
+
+Figure \ref{fig:walking_documentation} displays the documentation for
+the walking episodes from the 2015 GSS (\texttt{walking\_2015.rda}). The
+documentation provides a brief description of the dataset, instructions
+on how to load the package, as well as details about the file format and
+descriptions of all variables (already labeled).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2015-documentation}
+
+}
+
+\caption{Documentation for the dataset of walking episodes from 2015.}\label{fig:walking_documentation}
+\end{figure}
+
+The chunk below shows how to load the walking episodes from the 2015 GSS
+(Cycle 29):
+
+\begin{verbatim}
+data(walking_2015)
+\end{verbatim}
+
+The GSS surveys apply a probability sampling methodology, in which each
+episode or person selected in the sample represents several other
+episodes or persons not in the sample. The number of episodes and
+persons represented by a episode or person is determined by the weight
+or weighting factor. Because of this, every estimates of the number of
+episodes or persons need to be calculated applying the corresponding
+weighting factors.
+
+For instance, if someone wants to calculate the percentage of
+respondents from the 2015 GSS survey with active travel episodes, it is
+necessary to account for the person's weights. In 2015, the weight
+variable is represented by \texttt{WGHT\_PER.} The code chunk below
+demonstrates how to obtain the percentage of people with active travel
+episodes by age group. It uses the dplyr package to manipulate the data.
+
+The process begins by creating a dataset that sums the population by age
+group. Then, it joins the 2015 episodes with the 2015 main file. Note
+that in both operations, the code sums the person weight variable
+(\texttt{WGHT\_PER}) to obtain the correct population values. After
+this, a new dataset, \texttt{Active\_percentage}, is created by merging
+both previous datasets. The percentage of active travel episodes by age
+group is calculated by dividing the total population by the population
+with active trip episodes, then multiplying by 100.
+
+The result shows that the age group with the highest share of active
+trips is those between 15 and 20 years old, with almost 37\%, followed
+by those between 25 and 34 years old with around 33\%. There is a
+significant drop in percentage for the following groups, with the
+percentage falling to between 15\% and 18\% for the remaining age
+groups.
+
+\begin{verbatim}
+# Load the dplyr library for data manipulation
+library('dplyr')
+
+# Load the episodes dataset
+data(gss_episodes)
+
+# Calculate the total population for each age group
+Total_population <- gss_main_2015 %>% 
+  group_by(AGEGR10) %>% 
+  summarise(Total_population  = sum(WGHT_PER))
+
+# Calculate the active population for each age group
+Active_population <- gss_episodes %>% 
+      filter(YEAR == 2015) %>% 
+  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
+  group_by(AGEGR10) %>% 
+  summarise(Active_population  = sum(WGHT_PER))
+
+# Combine total and active population and calculate the percentage
+Active_percentage <- Total_population %>% 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
+  mutate(Percentage = round(Active_population/Total_population  * 100, 2))
+
+# Display the final result
+Active_percentage
+\end{verbatim}
+
+\begin{verbatim}
+# A tibble: 7 x 4
+  AGEGR10           Total_population Active_population Percentage
+  <ord>                        <dbl>             <dbl>      <dbl>
+1 15 to 24 years            4511131.          1641468.       36.4
+2 25 to 34 years            4956386.          1644033.       33.2
+3 35 to 44 years            4734506.           847770.       17.9
+4 45 to 54 years            5136125.           790667.       15.4
+5 55 to 64 years            4831306.           767064.       15.9
+6 65 to 74 years            3283969.           559025.       17.0
+7 75 years and over         2312976.           384359.       16.6
+\end{verbatim}
+
+\subsection{Processed data set}\label{processed-data-set}
+
+Table \ref{tab:processed-obs} displays the total number of observations
+processed for main and episode files. For the main files, a total of
+89,331 observations were processed, referring to all observations from
+the Time Use Surveys from 1986 to 2015, that together represents more of
+149,389,839 respondents (29,766,399 for 2015, 28,075,610 for 2010,
+26,095,819 for 2005, 24,260,137 for 1998, 21,294,313 for 1992, and for
+2005, 24,260,137 for 1998, and 21,294,313 for 1986). Table X also
+presents the total cases of active trips episodes identified. In total
+21,748 observations with register of active travel activity. Together,
+these observations account for NA episodes (0 for 2015, 0 for 2010, 0
+for 2005, 0 for 1998, 0 for 1992, and for 2005, 0 for 1998, and 0 for
+1986).
+
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}[t]{lrrr}
+\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of observations processed.}\\
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Observations} \\
+\cmidrule(l{3pt}r{3pt}){3-4}
+Survey & Year & Count & Weighted\\
+\midrule
+ & 2015 & 17390 & 29766399\\
+\nopagebreak
+ & 2010 & 15390 & 28075610\\
+\nopagebreak
+ & 2005 & 19597 & 26095819\\
+\nopagebreak
+ & 1998 & 10749 & 24260137\\
+\nopagebreak
+ & 1992 & 9815 & 21294313\\
+\nopagebreak
+\multirow[t]{-6}{*}{\raggedright\arraybackslash Main} & 1986 & 16390 & 19897562\\
+\cmidrule{1-4}\pagebreak[0]
+ & 2015 & 3496 & 6634387\\
+\nopagebreak
+ & 2010 & 4615 & 8516753\\
+\nopagebreak
+ & 2005 & 5866 & 7583838\\
+\nopagebreak
+ & 1998 & 1789 & 3606987\\
+\nopagebreak
+ & 1992 & 1635 & \\
+\nopagebreak
+\multirow[t]{-6}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
+\bottomrule
+\end{longtable}
+\endgroup{}
+
+Table \ref{tab:main-2015-processed} shows the first ten rows and first
+six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in
+\ref{tab:main-2015-unprocessed} before our processing. Table
+\ref{tab:ep-2015-processed} presents the walking episodes for the record
+identification number \texttt{10041} from the GSS PUMF 2015 episode file
+(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Only active travel episodes appear in Table \ref{tab:ep-2015-processed}
+since the observations were filtered to select cases with walking or
+cycling episodes. For both cases, Tables \ref{tab:main-2015-processed}
+and \ref{tab:ep-2015-processed} contain labeled variables, facilitating
+the interpretation of the data.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification.
+`WGHT\_PER`:  person weight.
+`SURVMNTH`: survey month of data collection.
+`AGEGR10`: age group of the respondent.
+`SEX`: sex of the respondent.
+`MARSTAT`: marital status of the respondent.
+\end{TableNotes}
+\begin{longtable}[t]{cccccc}
+\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
+\toprule
+PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
+\midrule
+10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
+10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
+10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
+10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
+10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
+\addlinespace
+10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
+10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
+10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
+10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
+10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}
+\item \textit{Note: } 
+\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\end{TableNotes}
+\begin{longtable}[t]{rrlrlll}
+\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.}\\
+\toprule
+PUMFID & WGHT\_EPI & Activity & Duration & Origin & Destination & Mode\\
+\midrule
+10041 & 1353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
+\bottomrule
+\insertTableNotes
+\end{longtable}
+\end{ThreePartTable}
+\endgroup{}
 
@@
 
-Considering GSS Cycles analyzed, we identified 21748 episodes that
+Considering GSS Cycles analyzed, we identified 21,748 episodes that
 recorded active travel episodes, with trip duration ranging from 0 to
@@
 presents descriptive statistics on walking and cycling trips between
-1986 and 2015, including metrics such as the count of recorded trips
-(count), and measures of trip duration in minutes: maximum (max), mean,
-median, and minimum (min). The 1986 survey did not include bicycle
+1986 and 2015, with measures of trip duration in minutes: maximum (max),
+mean, median, and minimum (min). The 1986 survey did not include bicycle
 trips.
@@
 \midrule
- & Count & 4347 & 1500 & 1670 & 5533 & 4379 & 3251\\
-\nopagebreak
  & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
 \nopagebreak
- & Mean & 21 & 19 & 11 & 12 & 12 & 17\\
+ & Mean & 21 &  & 12 & 12 & 13 & 18\\
 \nopagebreak
- & Median & 10 & 10 & 5 & 10 & 8 & 10\\
+ & Median & 15 & 10 & 5 & 10 & 10 & 10\\
 \nopagebreak
@@
 \nopagebreak
-\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
 \cmidrule{1-8}\pagebreak[0]
- & Count &  & 135 & 119 & 333 & 236 & 245\\
-\nopagebreak
  & Maximum &  & 240 & 90 & 180 & 153 & 120\\
 \nopagebreak
- & Mean &  & 31 & 21 & 19 & 21 & 24\\
+ & Mean &  & 28 & 24 & 20 & 19 & 25\\
 \nopagebreak
- & Median &  & 20 & 15 & 15 & 15 & 15\\
+ & Median &  & 15 & 15 & 15 & 10 & 20\\
 \nopagebreak
@@
 \nopagebreak
-\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
 \bottomrule
@@
 
-Table \ref{tab:table-01} shows that the median values for walking trips
-range between 5 and 10 minutes, while cycling trips have a consistent
-median of 15 minutes since 1998. The table also highlights very high
-maximum values, particularly for walking trips, with recorded episodes
-exceeding 4 hours in all cases.
+Table \ref{tab:table-01} shows that, in general, the median values for
+walking trips is 10 minutes, except for 1998 when the median was 5
+minutes. In the case of cycling trips, the duration fluctuated over the
+years, ranging from 10 to 20 minutes. The table also highlights very
+high maximum values, particularly for walking trips, with recorded
+episodes exceeding 4 hours in all cases.
 
@@
 trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
-median duration for trips to \texttt{home} and \texttt{work\ or\ school}
-remained at 10 minutes.
+median duration for walking trips to \texttt{home} and
+\texttt{work\ or\ school} remained at 10 minutes.
 
@@
 \item \textit{Note: } 
-\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
+\item 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
 \end{TableNotes}
@@
 \cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
-\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode*}} & \multicolumn{1}{c}{\textbf{Min*}} & \multicolumn{1}{c}{\textbf{Med*}} & \multicolumn{1}{c}{\textbf{Max*}} & \multicolumn{1}{c}{\textbf{(\%)*}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
+\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
 \midrule
- & Home &  &  &  &  & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
+ & Home &  &  &  &  & 5 & 20 & 240 & 55.7 & 2 & 15 & 90 & 51.6\\
 \nopagebreak
- & Other's home &  &  &  &  & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
+ & Other's home &  &  &  &  & 5 & 10 & 145 & 19.7 & 2 & 10 & 80 & 15.7\\
 \nopagebreak
-\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
+\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 24.7 & 5 & 25 & 75 & 32.8\\
 \cmidrule{1-14}\pagebreak[0]
- & Home & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
+ & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 &  & 1 & 5 & 255 & 51.9\\
 \nopagebreak
- & Other's home & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
+ & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 &  & 1 & 5 & 120 & 27.8\\
 \nopagebreak
-\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.3 & 2 & 10 & 60 & 19.2 & 1 & 6.5 & 75 & 20.4\\
+\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 &  & 1 & 10 & 75 & 20.3\\
 \bottomrule
@@
 \cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
-\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min*}} & \multicolumn{1}{c}{\textbf{Med*}} & \multicolumn{1}{c}{\textbf{Max*}} & \multicolumn{1}{c}{\textbf{(\%)*}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
+\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
 \midrule
- & Cultural venues & 10 & 12.5 & 15 & 0.6 & 10 & 25 & 30 & 1.3 & 15 & 15.0 & 15 & 0.8\\
+ & Cultural venues & 10 & 10 & 15 & 0.3 & 10 & 25 & 30 & 1.0 & 15 & 15 & 15 & 0.5\\
 \nopagebreak
- & Grocery store & 2 & 10.0 & 30 & 10.2 & 5 & 10 & 75 & 8.9 & 5 & 15.0 & 80 & 6.5\\
+ & Grocery store & 2 & 10 & 30 & 10.4 & 5 & 10 & 75 & 7.2 & 5 & 15 & 80 & 5.5\\
 \nopagebreak
- & Health clinic &  &  &  &  &  &  &  &  & 10 & 15.0 & 90 & 2.0\\
+ & Health clinic &  &  &  &  &  &  &  &  & 10 & 15 & 90 & 1.9\\
 \nopagebreak
- & Home & 1 & 15.0 & 180 & 48.9 & 1 & 15 & 135 & 50.4 & 5 & 20.0 & 120 & 46.9\\
+ & Home & 1 & 15 & 180 & 48.4 & 1 & 10 & 135 & 48.8 & 5 & 20 & 120 & 46.4\\
 \nopagebreak
- & Neighbourhood &  &  &  &  &  &  &  &  & 10 & 30.0 & 45 & 1.2\\
+ & Neighbourhood &  &  &  &  &  &  &  &  & 10 & 30 & 45 & 1.3\\
 \nopagebreak
- & Other's home & 1 & 15.0 & 35 & 9.0 & 5 & 10 & 45 & 9.3 & 5 & 15.0 & 40 & 5.3\\
+ & Other's home & 1 & 10 & 35 & 9.4 & 5 & 10 & 45 & 9.7 & 5 & 15 & 40 & 4.8\\
 \nopagebreak
- & Outdoors & 5 & 15.0 & 45 & 6.0 & 3 & 10 & 115 & 3.8 & 15 & 20.0 & 30 & 1.2\\
+ & Outdoors & 5 & 15 & 45 & 6.0 & 3 & 10 & 115 & 2.6 & 15 & 30 & 30 & 1.0\\
 \nopagebreak
- & Place of worship & 20 & 20.0 & 20 & 0.3 &  &  &  &  & 15 & 15.0 & 15 & 0.4\\
+ & Place of worship & 20 & 20 & 20 & 0.1 &  &  &  &  & 15 & 15 & 15 & 0.2\\
 \nopagebreak
- & Restaurant & 5 & 20.0 & 35 & 3.0 & 10 & 15 & 153 & 2.1 & 10 & 17.5 & 60 & 4.1\\
+ & Restaurant & 5 & 15 & 35 & 2.8 & 10 & 15 & 153 & 1.7 & 10 & 20 & 60 & 3.0\\
 \nopagebreak
- & Sport area &  &  &  &  &  &  &  &  & 10 & 15.0 & 15 & 2.9\\
+ & Sport area &  &  &  &  &  &  &  &  & 10 & 15 & 15 & 2.0\\
 \nopagebreak
-\multirow[t]{-11}{*}{\centering\arraybackslash Cycling} & Work or school & 1 & 15.0 & 90 & 21.9 & 1 & 15 & 100 & 24.2 & 5 & 15.0 & 120 & 28.6\\
+\multirow[t]{-11}{*}{\centering\arraybackslash Cycling} & Work or school & 1 & 15 & 90 & 22.6 & 1 & 15 & 100 & 29.0 & 5 & 20 & 120 & 33.5\\
 \cmidrule{1-14}\pagebreak[0]
- & Business &  &  &  &  &  &  &  &  & 5 & 10.0 & 30 & 0.2\\
+ & Business &  &  &  &  &  &  &  &  & 5 & 10 & 30 & 0.2\\
 \nopagebreak
- & Cultural venues & 5 & 12.5 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 10.0 & 40 & 1.5\\
+ & Cultural venues & 5 & 10 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 15 & 40 & 1.5\\
 \nopagebreak
- & Grocery store & 1 & 10.0 & 90 & 12.5 & 1 & 8 & 105 & 13.2 & 5 & 10.0 & 130 & 11.8\\
+ & Grocery store & 1 & 10 & 90 & 11.5 & 1 & 7 & 105 & 13.1 & 5 & 10 & 130 & 10.3\\
 \nopagebreak
- & Health clinic &  &  &  &  &  &  &  &  & 5 & 10.0 & 130 & 1.0\\
+ & Health clinic &  &  &  &  &  &  &  &  & 5 & 10 & 130 & 0.9\\
 \nopagebreak
- & Home & 0 & 10.0 & 515 & 44.4 & 0 & 10 & 270 & 43.6 & 5 & 10.0 & 900 & 45.3\\
+ & Home & 0 & 10 & 515 & 43.0 & 0 & 10 & 270 & 41.4 & 5 & 10 & 900 & 44.0\\
 \nopagebreak
- & Neighbourhood &  &  &  &  &  &  &  &  & 5 & 10.0 & 60 & 2.1\\
+ & Neighbourhood &  &  &  &  &  &  &  &  & 5 & 10 & 60 & 2.6\\
 \nopagebreak
- & Other's home & 1 & 5.0 & 300 & 11.7 & 0 & 5 & 140 & 11.3 & 5 & 10.0 & 120 & 7.3\\
+ & Other's home & 1 & 5 & 300 & 10.2 & 0 & 5 & 140 & 10.2 & 5 & 10 & 120 & 6.2\\
 \nopagebreak
- & Outdoors & 1 & 5.0 & 295 & 3.6 & 0 & 10 & 480 & 5.2 & 5 & 10.0 & 135 & 2.8\\
+ & Outdoors & 1 & 5 & 295 & 3.4 & 0 & 10 & 480 & 5.0 & 5 & 10 & 135 & 3.1\\
 \nopagebreak
- & Place of worship & 1 & 10.0 & 30 & 0.8 & 1 & 8 & 60 & 0.9 & 5 & 15.0 & 45 & 1.1\\
+ & Place of worship & 1 & 10 & 30 & 0.7 & 1 & 7 & 60 & 0.8 & 5 & 15 & 45 & 1.1\\
 \nopagebreak
- & Restaurant & 0 & 5.0 & 85 & 9.3 & 1 & 5 & 153 & 10.0 & 5 & 10.0 & 120 & 8.4\\
+ & Restaurant & 0 & 5 & 85 & 9.9 & 1 & 7 & 153 & 11.0 & 5 & 10 & 120 & 9.0\\
 \nopagebreak
- & Sport area &  &  &  &  &  &  &  &  & 5 & 10.0 & 45 & 3.3\\
+ & Sport area &  &  &  &  &  &  &  &  & 5 & 10 & 45 & 3.3\\
 \nopagebreak
-\multirow[t]{-12}{*}{\centering\arraybackslash Walking} & Work or school & 0 & 10.0 & 175 & 17.1 & 0 & 10 & 150 & 15.0 & 5 & 10.0 & 190 & 15.1\\
+\multirow[t]{-12}{*}{\centering\arraybackslash Walking} & Work or school & 0 & 10 & 175 & 20.7 & 0 & 10 & 150 & 17.9 & 5 & 10 & 190 & 17.8\\
 \bottomrule
@@
 
+\texttt{ActiveCA} also enables obtaining insights from the main
+processed files. Figure \ref{fig:figure-stress} present how the level of
+stress varied among respondents depending on their marital status in
+2015. According to this plot, married respondents reported the highest
+level of stress, with 17\% of possible cases.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/main_stress_figure}
+
+}
+
+\caption{Level of stress among respondents of different marital statuses (2015).}\label{fig:figure-stress}
+\end{figure}
+
 \section{Python integration}\label{python-integration}
-
-CAN WE PROVIDE PYTHON INTEGRATION? See here:
-
-\url{https://github.com/paezha/idealista18/tree/master/Python}
 
@@
 in the future. \texttt{R} users can seamlessly explore GSS walking and
-cycling episodes , with the option to suggest enhancements to the
-package as needed. This article adopts the structure proposed by
-Anastasia and Páez \citeyearpar{soukhov2023}, whose work provided
-essential guidance for the creation of this package. Similarly, we aim
-to contribute to the academic community by promoting transparent
-research practices that encourage replication and innovation in related
-fields. We believe that \{ActiveCA\} will serve as a basis for further
-research on GSS and for the integration of additional data by the
-authors or the wider open source community.
+cycling episodes, with the option to suggest enhancements to the package
+as needed. This article adopts the structure proposed by Anastasia and
+Páez \citeyearpar{soukhov2023}, whose work provided essential guidance
+for the creation of this package. Similarly, we aim to contribute to the
+academic community by promoting transparent research practices that
+encourage replication and innovation in related fields. We believe that
+\{ActiveCA\} will serve as a basis for further research on GSS and for
+the integration of additional data by the authors or the wider open
+source community.
 

### 2024-12-14 paper/Manuscript-Data-Package.Rmd
Upload new version with adjusting of the weights for the 1992 survey

@@
 abstract: 
-  "This paper describes {ActiveCA}, and open data product with Canadian time use 
-data. {ActiveCA} is an `R` data package that contains analysis-ready data related 
-to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992),
-12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
-Active travel is characterized by mode, with walking being part of every cycle 
-and bicycling starting in 1992. The attributes of active trips are the types of 
-locations of origins and destinations, the duration of trips, and episode weights
-for expanding the trips to population-wide estimates. Based on the year of the 
-survey, a variety of locations are coded. In earlier cycles, these include home,
-work or school, and other's home, whereas in later cycles these are augmented with
-locations such as grocery stores, restaurants, outdoor destinations, and others.
-The geographical resolution includes the province and whether the episode was in
-an urban or rural setting."
+  "This paper describes {ActiveCA}, and open data product with Canadian time use data.
+{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
+Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
+The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates.
+Based on the year of the survey, a variety of locations are coded.
+In earlier cycles, these include home, work or school, and other's home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others.
+The geographical resolution includes the province and whether the episode was in an urban or rural setting."
 
@@
 ---
-
-<!-- 
-FROM THE ABSTRACT:
-
-The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency.
-{ActiveCA} also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations.
-The package will continue to expand with contributions from the authors and the broader community through requests in the future.
-{ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed.
--->
 
@@
 data("gss_impedances")
-data("gss_main_2015")
 
 # Original main and episode files from 2015 to display table examples
-
 gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
@@
 
+`ActiveCA` also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames.
+This process allows users to work with and utilize the datasets available in `ActiveCA` within a Python project.
+
 # Concluding remarks

### 2025-01-17 data-raw/01-preprocessing-episode-files.Rmd
Including labelled main files in the data paper
inclusion of section for python integration and how to use this package

@@
   dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+```
+
+Cases of episodes with `NA` weight values will be updated to value 1, meaning that since there's no weighting value for this observation, this episode only represent itself:
+
+```{r fixing-NA-weights-walking_1992}
+walking_1992$WGHT_EPI[is.na(walking_1992$WGHT_EPI)] <- 1
 ```
@@
 
+Cases of episodes with `NA` weight values will be updated to value 1, meaning that since there's no weighting value for this observation, this episode only represent itself:
+
+```{r fixing-NA-weights}
+gss_episodes$WGHT_EPI[is.na(gss_episodes$WGHT_EPI)] <- 1
+```
+
 Save this data frame in R format:

### 2025-01-17 paper/Manuscript-Data-Package.Rmd
inclusion of section for python integration and how to use this package

@@
 abstract: 
-  "This paper describes {ActiveCA}, and open data product with Canadian time use data.
+  "This paper describes {ActiveCA}, an open data product with Canadian time use data.
 {ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
@@
 
-# Original main and episode files from 2015 to display table examples
+# Original main and Episode Files from 2015 to display table examples
 gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
@@
 
-# Processed main files
+# Processed Main Files
 data("gss_main_2015")
@@
 
-The main file of the time use survey compiles a large array of aggregated data, 
+The Main File of the time use survey compiles a large array of aggregated data, 
 summarizing the answers to the questionnaire that describe households and 
@@
 
-Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29).
+Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29).
 Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
 
-The main file of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
+The Main File of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
 For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
+    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
   kable_styling(full_width = FALSE, 
@@
 respondent. Although respondent-specific characteristics are not included within
-the episode file, it is possible to link the main file and the episode file by 
+the Episode File, it is possible to link the Main File and the Episode File by 
 using a key present in both the Main and Episode Files.
 
-Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the main file structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 episode file (Cycle 29).
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
 Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
 
-In total, the episode file of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
-Similar to the main file, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
+In total, the Episode File of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
+Similar to the Main File, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
 
@@
 
-For each selected cycle of the GSS surveys, we reviewed the episode files to identify episodes of movement that involved walking or cycling.
+For each selected cycle of the GSS surveys, we reviewed the Episode Files to identify episodes of movement that involved walking or cycling.
 This allowed us to select the activities immediately before and after the movement episode.
@@
 
-For each GSS main file, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes.
+For each GSS Main File, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes.
 We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others.
-After this, as we did with the episode files, we labeled the code variables with their appropriate descriptions.
+After this, as we did with the Episode Files, we labeled the code variables with their appropriate descriptions.
 
-# How to use `ActiveCA`
+# How to use {ActiveCA}
 
-This section presents some potential applications of the `ActiveCA` R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
+This section presents some potential applications of the {ActiveCA} R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
 
-To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the `ActiveCA` package in R:
+To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the {ActiveCA} package in R:
 
@@
 
-After installing the `ActiveCA` R package, you can load it using the following code:
+After installing the {ActiveCA} R package, you can load it using the following code:
 
@@
 In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group.
-It uses the dplyr package to manipulate the data.
+It uses the `dplyr`package [@dplyr2023] to manipulate the data.
 
 The process begins by creating a dataset that sums the population by age group.
-Then, it joins the 2015 episodes with the 2015 main file.
+Then, it joins the 2015 episodes with the 2015 Main File.
 Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
 After this, a new dataset, `Active_percentage`, is created by merging both previous datasets.
-The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100.
+The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100 and rounding with 2 decimal places.
 
@@
 
-```{r percentages-act, echo=TRUE}
+```{r percentages-act, echo=TRUE, out.width='80%'}
 # Load the dplyr library for data manipulation
@@
 
-# Combine total and active population and calculate the percentage
+# Combine total and active population 
+#and calculate the percentage
 Active_percentage <- Total_population %>% 
   left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
-  mutate(Percentage = round(Active_population/Total_population  * 100, 2))
+  mutate(
+    Percentage = round(Active_population/Total_population*100,2))
 
@@
 
-Table \ref{tab:processed-obs} displays the total number of observations processed for main and episode files.
-For the main files, a total of `r sum(df_main_processed$Count)` observations were processed, referring to all observations from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
-Table X also presents the total cases of active trips episodes identified.
-In total `r sum(df_ep_processed$Count)` observations with register of active travel activity.
-Together, these observations account for `r sum(df_ep_processed$Weighted)` episodes (`r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1986).
+Table \ref{tab:processed-obs} displays the total number of records processed for Main and Episode Files.
+For the Main Files, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
+It also presents the total cases of active trips episodes identified.
+In total `r sum(df_ep_processed$Count)` records with register of active travel activity.
+Together, these records account for `r sum(df_ep_processed$Weighted)` episodes.
 
@@
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
-        caption = "\\label{tab:processed-obs}Total number and weighted sum of observations processed.") %>%
+        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") %>%
   kable_styling(full_width = FALSE, font_size = 10) %>%
@@
 
-Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
-Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 episode file (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
-Only active travel episodes appear in Table \ref{tab:ep-2015-processed} since the observations were filtered to select cases with walking or cycling episodes.
+Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
+Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes.
 For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
+    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
   kable_styling(full_width = FALSE, 
@@
     longtable = TRUE, 
-    caption = "\\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.", 
+    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.", 
        col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) %>% 
@@
 Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
-`ActiveCA` includes all these episodes ready for analysis.
-Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes: maximum (max), mean, median, and minimum (min).
-The 1986 survey did not
-include bicycle trips.
+{ActiveCA} includes all these episodes ready for analysis.
+Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes.
+The 1986 survey did not include bicycle trips.
 
@@
 
-`ActiveCA` also enables obtaining insights from the main processed files.
+{ActiveCA} also enables obtaining insights from the main processed files.
 Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2015.
-According to this plot, married respondents reported the highest level of stress, with 17% of possible cases.
+According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 17% of possible cases.
 
@@
 
-`ActiveCA` also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames.
-This process allows users to work with and utilize the datasets available in `ActiveCA` within a Python project.
+{ActiveCA} also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames.
+This process allows users to work with and utilize the datasets available in {ActiveCA} within a Python project.
 
@@
 
-This paper presents `ActiveCA`, an open data product that provides analysis-ready 
+This paper presents {ActiveCA}, an open data product that provides analysis-ready 
 data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 

### 2025-01-17 paper/Manuscript-Data-Package.tex
inclusion of section for python integration and how to use this package

@@
 \begin{abstract}
-This paper describes \{ActiveCA\}, and open data product with Canadian
+This paper describes \{ActiveCA\}, an open data product with Canadian
 time use data. \{ActiveCA\} is an \texttt{R} data package that contains
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
+\hypertarget{general-social-survey-gss-collection}{%
 \section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}
+collection}\label{general-social-survey-gss-collection}}
 
@@
 
-\subsection{The Main File}\label{the-main-file}
+\hypertarget{the-main-file}{%
+\subsection{The Main File}\label{the-main-file}}
 
-The main file of the time use survey compiles a large array of
+The Main File of the time use survey compiles a large array of
 aggregated data, summarizing the answers to the questionnaire that
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
-six variables of the GSS PUMF 2015 main file (Cycle 29). Each row in the
+six variables of the GSS PUMF 2015 Main File (Cycle 29). Each row in the
 table correspond to a survey respondent, while the columns refer the
@@
 
-The main file of the 2015 GSS surveys includes a total of 17,390
+The Main File of the 2015 GSS surveys includes a total of 17,390
 respondents, representing 29,766,399 individuals and 848 variables. For
@@
 \begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
+\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.}\\
 \toprule
@@
 of the respondent. Although respondent-specific characteristics are not
-included within the episode file, it is possible to link the main file
-and the episode file by using a key present in both the Main and Episode
+included within the Episode File, it is possible to link the Main File
+and the Episode File by using a key present in both the Main and Episode
 Files.
@@
 Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
-example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
+example of the Main File structure, Table \ref{tab:ep-2015-unprocessed}
 presents the first seven episodes for the record identification number
-\texttt{10041} and some variables from the GSS PUMF 2015 episode file
+\texttt{10041} and some variables from the GSS PUMF 2015 Episode File
 (Cycle 29). Each row in the table corresponds to an episode associated
@@
 
-In total, the episode file of the 2015 GSS surveys contains 274,108
+In total, the Episode File of the 2015 GSS surveys contains 274,108
 records, representing 461,837,622 episodes and 527 variables detailing
-the episodes. Similar to the main file, Statistics Canada has created
+the episodes. Similar to the Main File, Statistics Canada has created
 codes for the discrete variables, with each value corresponding to a
@@
 
-\section{Data extraction}\label{data-extraction}
+\hypertarget{data-extraction}{%
+\section{Data extraction}\label{data-extraction}}
 
-For each selected cycle of the GSS surveys, we reviewed the episode
-files to identify episodes of movement that involved walking or cycling.
+For each selected cycle of the GSS surveys, we reviewed the Episode
+Files to identify episodes of movement that involved walking or cycling.
 This allowed us to select the activities immediately before and after
@@
 
-For each GSS main file, we selected socioeconomic variables to help
+For each GSS Main File, we selected socioeconomic variables to help
 profile the individuals engaged in active trip episodes. We included key
@@
 status, and number of children, among others. After this, as we did with
-the episode files, we labeled the code variables with their appropriate
+the Episode Files, we labeled the code variables with their appropriate
 descriptions.
 
-\section{\texorpdfstring{How to use
-\texttt{ActiveCA}}{How to use ActiveCA}}\label{how-to-use-activeca}
+\hypertarget{how-to-use-activeca}{%
+\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}}
 
-This section presents some potential applications of the
-\texttt{ActiveCA} R package. In fact, we expect that the application of
-this package to extend beyond our pre-imagined range of uses.
+This section presents some potential applications of the \{ActiveCA\} R
+package. In fact, we expect that the application of this package to
+extend beyond our pre-imagined range of uses.
 
@@
 package. The code chunk below demonstrates how to install the
-\texttt{ActiveCA} package in R:
+\{ActiveCA\} package in R:
 
@@
 
-After installing the \texttt{ActiveCA} R package, you can load it using
-the following code:
+After installing the \{ActiveCA\} R package, you can load it using the
+following code:
 
@@
 demonstrates how to obtain the percentage of people with active travel
-episodes by age group. It uses the dplyr package to manipulate the data.
+episodes by age group. It uses the \texttt{dplyr}package
+\citep{dplyr2023} to manipulate the data.
 
 The process begins by creating a dataset that sums the population by age
-group. Then, it joins the 2015 episodes with the 2015 main file. Note
+group. Then, it joins the 2015 episodes with the 2015 Main File. Note
 that in both operations, the code sums the person weight variable
@@
 group is calculated by dividing the total population by the population
-with active trip episodes, then multiplying by 100.
+with active trip episodes, then multiplying by 100 and rounding with 2
+decimal places.
 
@@
 
-\subsection{Processed data set}\label{processed-data-set}
+\hypertarget{processed-data-set}{%
+\subsection{Processed data set}\label{processed-data-set}}
 
-Table \ref{tab:processed-obs} displays the total number of observations
-processed for main and episode files. For the main files, a total of
-89,331 observations were processed, referring to all observations from
-the Time Use Surveys from 1986 to 2015, that together represents more of
+Table \ref{tab:processed-obs} displays the total number of records
+processed for Main and Episode Files. For the Main Files, a total of
+89,331 records were processed, referring to all records from the Time
+Use Surveys from 1986 to 2015, that together represents more of
 149,389,839 respondents (29,766,399 for 2015, 28,075,610 for 2010,
 26,095,819 for 2005, 24,260,137 for 1998, 21,294,313 for 1992, and for
-2005, 24,260,137 for 1998, and 21,294,313 for 1986). Table X also
-presents the total cases of active trips episodes identified. In total
-21,748 observations with register of active travel activity. Together,
-these observations account for NA episodes (0 for 2015, 0 for 2010, 0
-for 2005, 0 for 1998, 0 for 1992, and for 2005, 0 for 1998, and 0 for
-1986).
+2005, 24,260,137 for 1998, and 21,294,313 for 1986). It also presents
+the total cases of active trips episodes identified. In total 21,748
+records with register of active travel activity. Together, these records
+account for 38,275,077 episodes.
 
@@
 \begin{longtable}[t]{lrrr}
-\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of observations processed.}\\
+\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
 \toprule
@@
 \nopagebreak
- & 1992 & 1635 & \\
+ & 1992 & 1635 & 3691918\\
 \nopagebreak
@@
 Table \ref{tab:main-2015-processed} shows the first ten rows and first
-six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in
+six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in
 \ref{tab:main-2015-unprocessed} before our processing. Table
 \ref{tab:ep-2015-processed} presents the walking episodes for the record
-identification number \texttt{10041} from the GSS PUMF 2015 episode file
+identification number \texttt{10041} from the GSS PUMF 2015 Episode File
 (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
-Only active travel episodes appear in Table \ref{tab:ep-2015-processed}
-since the observations were filtered to select cases with walking or
-cycling episodes. For both cases, Tables \ref{tab:main-2015-processed}
-and \ref{tab:ep-2015-processed} contain labeled variables, facilitating
-the interpretation of the data.
+Only the unique active travel episode appear in Table
+\ref{tab:ep-2015-processed} since the records were filtered to select
+cases with walking or cycling episodes. For both cases, Tables
+\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
+labeled variables, facilitating the interpretation of the data.
 
@@
 \begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
+\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.}\\
 \toprule
@@
 \begin{longtable}[t]{rrlrlll}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.}\\
+\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.}\\
 \toprule
@@
 
-\subsection{Descriptive statistics}\label{descriptive-statistics}
+\hypertarget{descriptive-statistics}{%
+\subsection{Descriptive statistics}\label{descriptive-statistics}}
 
@@
 recorded active travel episodes, with trip duration ranging from 0 to
-900 minutes, to twelve different destinations. \texttt{ActiveCA}
-includes all these episodes ready for analysis. Table \ref{tab:table-01}
-presents descriptive statistics on walking and cycling trips between
-1986 and 2015, with measures of trip duration in minutes: maximum (max),
-mean, median, and minimum (min). The 1986 survey did not include bicycle
-trips.
+900 minutes, to twelve different destinations. \{ActiveCA\} includes all
+these episodes ready for analysis. Table \ref{tab:table-01} presents
+descriptive statistics on walking and cycling trips between 1986 and
+2015, with measures of trip duration in minutes. The 1986 survey did not
+include bicycle trips.
 
@@
 \nopagebreak
- & Mean & 21 &  & 12 & 12 & 13 & 18\\
+ & Mean & 21 & 21 & 12 & 12 & 13 & 18\\
 \nopagebreak
@@
 \cmidrule{1-14}\pagebreak[0]
- & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 &  & 1 & 5 & 255 & 51.9\\
+ & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 & 60.5 & 1 & 5 & 255 & 51.9\\
 \nopagebreak
- & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 &  & 1 & 5 & 120 & 27.8\\
+ & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 & 18.7 & 1 & 5 & 120 & 27.8\\
 \nopagebreak
-\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 &  & 1 & 10 & 75 & 20.3\\
+\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 & 20.9 & 1 & 10 & 75 & 20.3\\
 \bottomrule
@@
 
-\texttt{ActiveCA} also enables obtaining insights from the main
-processed files. Figure \ref{fig:figure-stress} present how the level of
-stress varied among respondents depending on their marital status in
-2015. According to this plot, married respondents reported the highest
-level of stress, with 17\% of possible cases.
+\{ActiveCA\} also enables obtaining insights from the main processed
+files. Figure \ref{fig:figure-stress} present how the level of stress
+varied among respondents depending on their marital status in 2015.
+According to this plot, married respondents reported the highest level
+of stress, relating to feel stressed every day, with 17\% of possible
+cases.
 
@@
 
-\section{Python integration}\label{python-integration}
+\hypertarget{python-integration}{%
+\section{Python integration}\label{python-integration}}
 
-\section{Concluding remarks}\label{concluding-remarks}
+\{ActiveCA\} also provides a Jupyter Notebook containing a Python script
+that demonstrates how to read R data files (.rda) and convert them into
+Pandas DataFrames. This process allows users to work with and utilize
+the datasets available in \{ActiveCA\} within a Python project.
 
-This paper presents \texttt{ActiveCA}, an open data product that
-provides analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998),
-19 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel
-in Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
+\hypertarget{concluding-remarks}{%
+\section{Concluding remarks}\label{concluding-remarks}}
+
+This paper presents \{ActiveCA\}, an open data product that provides
+analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
+(2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in
+Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
 developed after collecting, cleaning, and processing the survey data,
@@
 
+\hypertarget{declaration-of-conflicting-interests}{%
 \section{Declaration of Conflicting
-Interests}\label{declaration-of-conflicting-interests}
+Interests}\label{declaration-of-conflicting-interests}}
 

### 2025-01-20 data-raw/03-preprocessing-main-files.Rmd
inclusion of section for python integration and how to use this package
adding helth-related variables to the 2015 main gss dataset

@@
 
-Read the Main files of the 2015 GSS survey:
+Read the Main file of the 2015 GSS survey:
 
@@
 gss_main_2015 <- gss_m_2015 %>%
-  select(PUMFID:DVTDAY)
+  select(PUMFID:DVTDAY, SLM_01, SRH_110, SRH_115)
 ```
@@
 
+```{r SLM_01-2015}
+values_ordered <- c("10 - Very satisfied","9","8","7","6","5",
+                    "4","3","2","1","0 - Very dissatisfied",
+                    "Valid skip", 
+                    "Don’t know", 
+                    "Refusal", 
+                    "Not stated")
+
+gss_main_2015 <- gss_main_2015 %>% 
+  mutate( # SLM_01: Using a scale of 0 to 10 where 0 means "Very dissatisfied" and 10 means "Very satisfied",how do you feel about your life as a whole right now? 
+    SLM_01 = 
+      case_when(SLM_01 == 0 ~ "0 - Very dissatisfied",
+                SLM_01 == 1 ~ "1",
+                     SLM_01 == 2 ~ "2",
+                     SLM_01 == 3 ~ "3",
+                     SLM_01 == 4 ~ "4",
+                     SLM_01 == 5 ~ "5",
+                     SLM_01 == 6 ~ "6",
+                     SLM_01 == 7 ~ "7",
+                     SLM_01 == 8 ~ "8",
+                     SLM_01 == 9 ~ "9",
+                     SLM_01 == 10 ~ "10 - Very satisfied",
+                     SLM_01 == 96 ~ "Valid skip",
+                     SLM_01 == 97 ~ "Don’t know",
+                     SLM_01 == 98 ~ "Refusal",
+                     SLM_01 == 99 ~ "Not stated"),
+    SLM_01 = factor(SLM_01, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2015$SLM_01)
+```
+
+```{r SRH_110-2015}
+values_ordered <- c("Excellent","Very good","Good",
+                    "Fair","Poor",
+                    "Valid skip", 
+                    "Don’t know", 
+                    "Refusal", 
+                    "Not stated")
+
+gss_main_2015 <- gss_main_2015 %>% 
+  mutate( # SRH_110: In general, would you say your health is...? 
+    SRH_110 = 
+      case_when(SRH_110 == 1 ~ "Excellent",
+                     SRH_110 == 2 ~ "Very good",
+                     SRH_110 == 3 ~ "Good",
+                     SRH_110 == 4 ~ "Fair",
+                     SRH_110 == 5 ~ "Poor",
+                     SRH_110 == 6 ~ "Valid skip",
+                     SRH_110 == 7 ~ "Don’t know",
+                     SRH_110 == 8 ~ "Refusal",
+                     SRH_110 == 9 ~ "Not stated"),
+    SRH_110 = factor(SRH_110, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2015$SRH_110)
+```
+
+```{r SRH_115-2015}
+values_ordered <- c("Excellent","Very good","Good",
+                    "Fair","Poor",
+                    "Valid skip", 
+                    "Don’t know", 
+                    "Refusal", 
+                    "Not stated")
+
+gss_main_2015 <- gss_main_2015 %>% 
+  mutate( # SRH_115: In general, would you say your mental health is...?
+    SRH_115 = 
+      case_when(SRH_115 == 1 ~ "Excellent",
+                     SRH_115 == 2 ~ "Very good",
+                     SRH_115 == 3 ~ "Good",
+                     SRH_115 == 4 ~ "Fair",
+                     SRH_115 == 5 ~ "Poor",
+                     SRH_115 == 6 ~ "Valid skip",
+                     SRH_115 == 7 ~ "Don’t know",
+                     SRH_115 == 8 ~ "Refusal",
+                     SRH_115 == 9 ~ "Not stated"),
+    SRH_115 = factor(SRH_115, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2015$SRH_115)
+```
+
 ```{r DVTDAY-2015}
@@
 ```
+
+Labeling socioeconomic variables:
 
@@
 
+Labeling socioeconomic variables:
+
 ```{r C2-1992}
@@
 
+Labeling socioeconomic variables:
+
 ```{r C2-1986}

### 2025-01-22 paper/Manuscript-Data-Package.tex
Some cosmetic fixes to paper

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 
@@
 
-\hypertarget{general-social-survey-gss-collection}{%
 \section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}}
+collection}\label{general-social-survey-gss-collection}
 
@@
 
-\hypertarget{the-main-file}{%
-\subsection{The Main File}\label{the-main-file}}
+\subsection{The Main File}\label{the-main-file}
 
@@
 
-\hypertarget{the-episode-file}{%
-\subsection{The Episode File}\label{the-episode-file}}
+\subsection{The Episode File}\label{the-episode-file}
 
@@
 
-\hypertarget{data-extraction}{%
-\section{Data extraction}\label{data-extraction}}
+\section{Data extraction}\label{data-extraction}
 
@@
 
-\hypertarget{how-to-use-activeca}{%
-\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}}
+\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}
 
@@
 
-By running \texttt{?ActiveCA} in \texttt{R}, you can access the complete
-package documentation. The documentation includes a description of the
-package, information about the authors and maintainers, and an overview
-of all pre-processed data sets included in the package.
+The documentation of the package can be accessed by running
+\texttt{?ActiveCA} in \texttt{R}. The documentation includes a
+description of the package, information about the authors, contributors,
+and maintainers, and an overview of all pre-processed data sets included
+in the package.
 
@@
 
-The chunk below shows how to load the walking episodes from the 2015 GSS
+The code below shows how to load the walking episodes from the 2015 GSS
 (Cycle 29):
@@
 persons represented by a episode or person is determined by the weight
-or weighting factor. Because of this, every estimates of the number of
+or weighting factor. Because of this, estimates of the number of
 episodes or persons need to be calculated applying the corresponding
@@
 
-For instance, if someone wants to calculate the percentage of
-respondents from the 2015 GSS survey with active travel episodes, it is
-necessary to account for the person's weights. In 2015, the weight
-variable is represented by \texttt{WGHT\_PER.} The code chunk below
-demonstrates how to obtain the percentage of people with active travel
-episodes by age group. It uses the \texttt{dplyr}package
-\citep{dplyr2023} to manipulate the data.
+For instance, to calculate the percentage of respondents from the 2015
+GSS survey with active travel episodes, it is necessary to account for
+the person's weight. In 2015, the weight variable is represented by
+\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
+percentage of people with active travel episodes by age group. It uses
+the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.
 
@@
 
-The result shows that the age group with the highest share of active
-trips is those between 15 and 20 years old, with almost 37\%, followed
-by those between 25 and 34 years old with around 33\%. There is a
-significant drop in percentage for the following groups, with the
-percentage falling to between 15\% and 18\% for the remaining age
-groups.
-
 \begin{verbatim}
@@
 # Calculate the total population for each age group
-Total_population <- gss_main_2015 %>% 
-  group_by(AGEGR10) %>% 
+Total_population <- gss_main_2015 |> 
+  group_by(AGEGR10) |> 
   summarise(Total_population  = sum(WGHT_PER))
@@
 # Calculate the active population for each age group
-Active_population <- gss_episodes %>% 
-      filter(YEAR == 2015) %>% 
-  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
-  group_by(AGEGR10) %>% 
+Active_population <- gss_episodes |> 
+      filter(YEAR == 2015) |> 
+  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) |> 
+  group_by(AGEGR10) |> 
   summarise(Active_population  = sum(WGHT_PER))
@@
 #and calculate the percentage
-Active_percentage <- Total_population %>% 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
+Active_percentage <- Total_population |> 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
   mutate(
     Percentage = round(Active_population/Total_population*100,2))
-
-# Display the final result
-Active_percentage
 \end{verbatim}
 
-\begin{verbatim}
-# A tibble: 7 x 4
-  AGEGR10           Total_population Active_population Percentage
-  <ord>                        <dbl>             <dbl>      <dbl>
-1 15 to 24 years            4511131.          1641468.       36.4
-2 25 to 34 years            4956386.          1644033.       33.2
-3 35 to 44 years            4734506.           847770.       17.9
-4 45 to 54 years            5136125.           790667.       15.4
-5 55 to 64 years            4831306.           767064.       15.9
-6 65 to 74 years            3283969.           559025.       17.0
-7 75 years and over         2312976.           384359.       16.6
-\end{verbatim}
+The result shows that the age group with the highest share of active
+trips is those between 15 and 20 years old, with almost 37\%, followed
+by those between 25 and 34 years old with around 33\%. There is a
+significant drop in percentage for the following groups, with the
+percentage falling to between 15\% and 18\% for the remaining age
+groups.
 
-\hypertarget{processed-data-set}{%
-\subsection{Processed data set}\label{processed-data-set}}
+\begin{longtable}[]{@{}lrrr@{}}
+\toprule\noalign{}
+AGEGR10 & Total\_population & Active\_population & Percentage \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+15 to 24 years & 4511131 & 1641468.3 & 36.39 \\
+25 to 34 years & 4956386 & 1644032.8 & 33.17 \\
+35 to 44 years & 4734506 & 847769.9 & 17.91 \\
+45 to 54 years & 5136125 & 790667.3 & 15.39 \\
+55 to 64 years & 4831306 & 767064.0 & 15.88 \\
+65 to 74 years & 3283969 & 559024.6 & 17.02 \\
+75 years and over & 2312976 & 384358.6 & 16.62 \\
+\end{longtable}
+
+\subsection{Processed data set}\label{processed-data-set}
 
@@
 
-\hypertarget{descriptive-statistics}{%
-\subsection{Descriptive statistics}\label{descriptive-statistics}}
+\subsection{Descriptive statistics}\label{descriptive-statistics}
 
@@
 
-\hypertarget{python-integration}{%
-\section{Python integration}\label{python-integration}}
+\section{Python integration}\label{python-integration}
 
@@
 
-\hypertarget{concluding-remarks}{%
-\section{Concluding remarks}\label{concluding-remarks}}
+\section{Concluding remarks}\label{concluding-remarks}
 
@@
 (\emph{link}).
+
+For review purposes, the package is currently available as a tar.gz file
+that can be installed by R. The file can be obtained from this anonymous
+location:
+
+\url{https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip}
 

### 2025-01-22 data-raw/01-preprocessing-episode-files.Rmd

@@
 # Episode File
-gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
+#gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))# Original file
+gss_e_2015 <- read.csv(paste0(here(), "/data-raw/2015-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+#gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv")) # Original file
+gss_m_2015 <- read.csv(paste0(here(), "/data-raw/2015-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv"))
+#gss_e_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv")) # Original file
+gss_e_2010 <- read.csv(paste0(here(), "/data-raw/2010-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv"))
+#gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv")) # Original file
+gss_m_2010 <- read.csv(paste0(here(), "/data-raw/2010-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 ```{r}
-
 gss_e_2010 %>%
@@
 # Episode File
-gss_e_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-e_F1.csv"))
+#gss_e_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-e_F1.csv")) # Original file
+gss_e_2005 <- read.csv(paste0(here(), "/data-raw/2005-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv"))
+#gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv")) # Original file
+gss_m_2005 <- read.csv(paste0(here(), "/data-raw/2005-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12e_F1.csv"))
+# gss_e_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12e_F1.csv")) # Original file
+gss_e_1998 <- read.csv(paste0(here(), "/data-raw/1998-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv"))
+# gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv")) # Original file
+gss_m_1998 <- read.csv(paste0(here(), "/data-raw/1998-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-ep_F1.csv"))
+# gss_e_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-ep_F1.csv")) # Original file
+gss_e_1992 <- read.csv(paste0(here(), "/data-raw/1992-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv"))
+# gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv")) # Original file
+gss_m_1992 <- read.csv(paste0(here(), "/data-raw/1992-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-ep_F1.csv"))
+# gss_e_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-ep_F1.csv")) # Original file
+gss_e_1986 <- read.csv(paste0(here(), "/data-raw/1986-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv"))
+# gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv")) # Original file
+gss_m_1992 <- read.csv(paste0(here(), "/data-raw/1986-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```

### 2025-04-04 data-raw/01-preprocessing-episode-files.Rmd
Updating the package with the new 2022 data set.

--- 2025-01-17 data-raw/01-preprocessing-episode-files.Rmd
@@
 been conducting time-use surveys at five to seven year intervals since
-1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
-and 2015). The time use collects information on participation of
+1986, with the most recent being in 2022 (1986, 1992, 1998, 2005, 2010, 2015, and 2022). The time use collects information on participation of
 respondents and time spent on a wide range of day-to-day activities
@@
 library(purrr) # Provides a complete and consistent set of tools for working with functions and vectors
+library(foreign) # To read SSPS files
+library(haven) # To read SAS files
 ```
@@
 
-[^4]:  whether 1986, 1992, 1998, 2005, 2010 or 2015. Note: There is no
+[^4]:  whether 1986, 1992, 1998, 2005, 2010, 2015, or 2022. Note: There is no
     data on cycling for the 1986 GSS survey.
@@
 # Episode File
-gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
+gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))# Original file
+#gss_e_2015 <- read.csv(paste0(here(), "/data-raw/2015-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv")) # Original file
+#gss_m_2015 <- read.csv(paste0(here(), "/data-raw/2015-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv"))
+gss_e_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-ef_F1.csv")) # Original file
+#gss_e_2010 <- read.csv(paste0(here(), "/data-raw/2010-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv"))
+gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv")) # Original file
+#gss_m_2010 <- read.csv(paste0(here(), "/data-raw/2010-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-e_F1.csv"))
+gss_e_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-e_F1.csv")) # Original file
+#gss_e_2005 <- read.csv(paste0(here(), "/data-raw/2005-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv"))
+gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv")) # Original file
+#gss_m_2005 <- read.csv(paste0(here(), "/data-raw/2005-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12e_F1.csv"))
+gss_e_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12e_F1.csv")) # Original file
+#gss_e_1998 <- read.csv(paste0(here(), "/data-raw/1998-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv"))
+gss_m_1998 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1998/gss-12M0012-E-1998-c-12-m_F1.csv")) # Original file
+#gss_m_1998 <- read.csv(paste0(here(), "/data-raw/1998-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-ep_F1.csv"))
+gss_e_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-ep_F1.csv")) # Original file
+#gss_e_1992 <- read.csv(paste0(here(), "/data-raw/1992-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv"))
+gss_m_1992 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1992/gss-12M0007-E-1992-c-7-m_F1.csv")) # Original file
+#gss_m_1992 <- read.csv(paste0(here(), "/data-raw/1992-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 # Episode File
-gss_e_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-ep_F1.csv"))
+gss_e_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-ep_F1.csv")) # Original file
+#gss_e_1986 <- read.csv(paste0(here(), "/data-raw/1986-gss-e-reproducibility.csv")) # Subset for reproducibility
 
 # Main File
-gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv"))
+gss_m_1986 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-1986/gss-12M0002-E-1986-c-2-main_F1.csv")) # Original file
+#gss_m_1992 <- read.csv(paste0(here(), "/data-raw/1986-gss-m-reproducibility.csv")) # Subset for reproducibility
 ```
@@
 
+## GSS 2022
+
+Read the Episode and Main files of the 2022 GSS survey:
+
+```{r}
+# Episode File
+gss_e_2022_spss <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_episode_2022_pumf_v1.sav"), to.data.frame = TRUE)# Original file
+
+gss_e_2022_sas <- read_sas(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_episode_2022_pumf_v1.sas7bdat")) # Original file
+
+#gss_e_2022 <- read.csv(paste0(here(), "/data-raw/2022-gss-e-reproducibility.csv")) # Subset for reproducibility
+
+# Main File
+gss_m_2022 <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_main_2022_pumf_v1.sav"), to.data.frame = TRUE) # Original file
+
+#gss_m_2022 <- read.csv(paste0(here(), "/data-raw/2022-gss-m-reproducibility.csv")) # Subset for reproducibility
+```
+
+### Walking 2022
+
+Identifying the rows with walking episodes and selecting the rows after
+and before the walking episodes to obtain the origin and location
+variables:
+
+```{r}
+# walking 2022
+# Creating data of origins and destinations of walking trip 2022
+inds_spss <- which(gss_e_2022_spss$LOCATION == 'Travel - Walk')
+rows_spss <- lapply(inds_spss, function(x) (x-1):(x+1))
+
+inds_sas <- which(gss_e_2022_sas$LOCATION == 3315)
+rows_sas <- lapply(inds_sas, function(x) (x-1):(x+1))
+```
+
+Now, we'll select all the rows with walking episodes in the GSS survey,
+selecting the columns that we are interested[^5], and already creating
+the origin and destination variables:
+
+[^5]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
+    activity code. We'll change the name of this column to be match the
+    name of the activity code variable in the others surveys.
+
+```{r}
+walking_2022_spss <- gss_e_2022_spss |>
+  dplyr::slice(unlist(rows_spss)) |>
+  dplyr::select(PUMFID:ACTIVITY, DURATION:ENDTIME) |> 
+  mutate(origin = lag(LOCATION),
+         destination = lead(LOCATION)) |>
+  filter(LOCATION == 'Travel - Walk') %>% 
+  mutate(PUMFID = as.factor(PUMFID), 
+         INSTANCE = as.factor(INSTANCE)) %>% 
+  dplyr::select(-c(LOCATION)) %>% 
+  dplyr::rename(act_label = ACTIVITY)
+
+walking_2022_sas <- gss_e_2022_sas |>
+  dplyr::slice(unlist(rows_sas)) |>
+  dplyr::select(PUMFID, INSTANCE, LOCATION, ACTIVITY) |> 
+  mutate(origin = lag(LOCATION),
+         destination = lead(LOCATION)) |>
+  dplyr::filter(LOCATION == 3315) %>% 
+  mutate(PUMFID = as.factor(PUMFID), 
+         INSTANCE = as.factor(INSTANCE)) %>% 
+  dplyr::rename(ACTCODE = ACTIVITY)
+```
+
+Now, we'll select the origins of interest and create the label for the
+`origin` variable:
+
+```{r}
+walking_2022_spss <- walking_2022_spss |>
+  filter(origin == 'At home or on property' | 
+           origin == 'At place of work or school' | 
+           origin == 'Grocery store, other stores or mall'| 
+           origin == 'In the neighbourhood'| 
+           origin == 'Sports centre, field or arena'| 
+           origin == 'Outdoors'| 
+           origin == 'Restaurant, bar or club'| 
+           origin == 'At someone else\'s home or property'| 
+           origin == 'Medical, dental or other health clinic'| 
+           origin == 'Library, museum or theatre'| 
+           origin == 'Place of worship' |
+           origin == 'Away at business') |>
+  mutate(orig_label =
+           case_when(origin == 'At home or on property' ~ "Home", 
+                     origin == 'At place of work or school' ~ "Work or school",
+                     origin == 'Away at business' ~ "Business",
+                     origin == 'At someone else\'s home or property' ~ "Other's home",
+                     TRUE ~ origin),
+         orig_label = factor(orig_label))
+```
+
+Similarly to the step before, we'll select the destinations of interest
+and create the label for the `destination` variable:
+
+```{r}
+walking_2022_spss <- walking_2022_spss |>  
+  filter(destination == 'At home or on property' | 
+           destination == 'At place of work or school' | 
+           destination == 'Grocery store, other stores or mall'| 
+           destination == 'In the neighbourhood'| 
+           destination == 'Sports centre, field or arena'| 
+           destination == 'Outdoors'| 
+           destination == 'Restaurant, bar or club'| 
+           destination == 'At someone else\'s home or property'| 
+           destination == 'Medical, dental or other health clinic'| 
+           destination == 'Library, museum or theatre'| 
+           destination == 'Place of worship' |
+           destination == 'Away at business') |>
+  mutate(dest_label =
+           case_when(destination == 'At home or on property' ~ "Home", 
+                     destination == 'At place of work or school' ~ "Work or school",
+                     destination == 'Away at business' ~ "Business",
+                     destination == 'At someone else\'s home or property' ~ "Other's home",
+                     TRUE ~ destination), 
+         dest_label = factor(dest_label))
+```
+
+Creating the `year` and `mode` columns. Additionally, the column named
+`TUI_01` will be renamed to `ACTCODE` to ensure consistency with the
+naming conventions used in other years.
+
+```{r}
+walking_2022 <- walking_2022_spss %>% 
+   dplyr::select(-c(origin,destination)) %>% 
+   left_join(walking_2022_sas, by = c('PUMFID', 'INSTANCE'))
+
+walking_2022 <- walking_2022 |>
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION,
+                origin, destination, orig_label, dest_label) |>
+  mutate(YEAR = 2022,
+         MODE = "Walking") 
+```
+
+Now, let's select only the cases of respondents who live in a CMA or CA:
+```{r}
+gss_m_2022 <- gss_m_2022 %>%
+  dplyr::select(PUMFID,LUC_RST,PRV) %>%
+  dplyr::rename(LUC = LUC_RST, 
+                Province = PRV) %>% 
+  mutate(PRV = 
+           case_when(Province == "Newfoundland and Labroador" ~ 10,
+                     Province == "Prince Edward Island" ~ 11,
+                     Province == "Nova Scotia" ~ 12,
+                     Province == "New Brunswick" ~ 13,
+                     Province == "Quebec" ~ 24,
+                     Province == "Ontario" ~ 35,
+                     Province == "Manitoba" ~ 46,
+                     Province == "Saskatchewan" ~ 47,
+                     Province == "Alberta" ~ 48,
+                     Province == "British Columbia" ~ 59),
+         Pop_centre =
+           case_when(LUC == 'Larger urban population centres (CMA/CA)' ~ "CMA/CA",
+                     LUC == 'Prince Edward Island' ~ "non CMA/CA",
+                     LUC == 'Rural areas/small population centres (non-CMA/CA)' ~ "non CMA/CA"),
+         LUC_RST =
+           case_when(LUC == 'Larger urban population centres (CMA/CA)' ~ 1,
+                     LUC == 'Prince Edward Island' ~ 3,
+                     LUC == 'Rural areas/small population centres (non-CMA/CA)' ~ 2))
+
+walking_2022 <- walking_2022 |> 
+  left_join(gss_m_2022, by = 'PUMFID') |>  # Merging the two tables
+  dplyr::select(names(walking_2015))
+
+for (col in names(walking_2015)) {
+  target_class <- class(walking_2015[[col]])
+  walking_2022[[col]] <- match.fun(paste0("as.", target_class[1]))(walking_2022[[col]])
+}
+```
+
+Now, let's evaluate the count of episodes between each combination of
+origin and destination:
+
+```{r}
+walking_2022 |> 
+  group_by(orig_label, 
+           dest_label) |>
+  count() |>
+  pivot_wider(names_from = "dest_label", values_from = "n")
+```
+
+The table above shows that the most common destination for a walking trip
+that starts at home are `Home` and `Work or school`. It also shows that the second most common walking trip is from `Grocery store, other stores or mall` to `Home`. Notice, though,
+that quite a few trips start *and* end at Home. What are those supposed
+to be?
+
+Check those home-to-home trips:
+
+```{r include=FALSE}
+walking_2022 |>
+  filter(orig_label == "Home" & dest_label == "Home")
+```
+
+See `PUMFID` == 1205 and check the walking episode:
+
+```{r}
+gss_e_2022_spss |>
+  filter(PUMFID == '1205') |>
+  slice(unique(c(which(LOCATION == 'Travel - Walk') - 1, 
+                 which(LOCATION == 'Travel - Walk'),
+                 which(LOCATION == 'Travel - Walk') + 1))) |>
+  dplyr::select(PUMFID, INSTANCE, ACTIVITY, LOCATION, DURATION)
+```
+
+The person was at home doing regular household tasks, then went for
+walk that lasted 20 minutes, came back home and returned to the household tasks for more 180 minutes.
+
+Save data in R format for later use:
+
+```{r}
+usethis::use_data(walking_2022, 
+                  overwrite = TRUE)
+```
+
+### Cycling 2022
+
+Identifying the rows with cycling episodes and selecting the rows after
+and before the cycling episodes to obtain the origin and location
+variables:
+
+```{r}
+# cycling 2022
+# Creating data of origins and destinations of cycling trip 2022
+inds_spss <- which(gss_e_2022_spss$LOCATION == 'Travel - Bicycle')
+rows_spss <- lapply(inds_spss, function(x) (x-1):(x+1))
+
+inds_sas <- which(gss_e_2022_sas$LOCATION == 3318)
+rows_sas <- lapply(inds_sas, function(x) (x-1):(x+1))
+```
+
+Now, we'll select all the rows with cycling episodes in the GSS survey,
+selecting the columns that we are interested[^5], and already creating
+the origin and destination variables:
+
+[^5]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
+    activity code. We'll change the name of this column to be match the
+    name of the activity code variable in the others surveys.
+
+```{r}
+cycling_2022_spss <- gss_e_2022_spss |>
+  dplyr::slice(unlist(rows_spss)) |>
+  dplyr::select(PUMFID:ACTIVITY, DURATION:ENDTIME) |> 
+  mutate(origin = lag(LOCATION),
+         destination = lead(LOCATION)) |>
+  filter(LOCATION == 'Travel - Bicycle') %>% 
+  mutate(PUMFID = as.factor(PUMFID), 
+         INSTANCE = as.factor(INSTANCE)) %>% 
+  dplyr::select(-c(LOCATION)) %>% 
+  dplyr::rename(act_label = ACTIVITY)
+
+cycling_2022_sas <- gss_e_2022_sas |>
+  dplyr::slice(unlist(rows_sas)) |>
+  dplyr::select(PUMFID, INSTANCE, LOCATION, ACTIVITY) |> 
+  mutate(origin = lag(LOCATION),
+         destination = lead(LOCATION)) |>
+  dplyr::filter(LOCATION == 3318) %>% 
+  mutate(PUMFID = as.factor(PUMFID), 
+         INSTANCE = as.factor(INSTANCE)) %>% 
+  dplyr::rename(ACTCODE = ACTIVITY)
+```
+
+Now, we'll select the origins of interest and create the label for the
+`origin` variable:
+
+```{r}
+cycling_2022_spss <- cycling_2022_spss |>
+  filter(origin == 'At home or on property' | 
+           origin == 'At place of work or school' | 
+           origin == 'Grocery store, other stores or mall'| 
+           origin == 'In the neighbourhood'| 
+           origin == 'Sports centre, field or arena'| 
+           origin == 'Outdoors'| 
+           origin == 'Restaurant, bar or club'| 
+           origin == 'At someone else\'s home or property'| 
+           origin == 'Medical, dental or other health clinic'| 
+           origin == 'Library, museum or theatre'| 
+           origin == 'Place of worship' |
+           origin == 'Away at business') |>
+  mutate(orig_label =
+           case_when(origin == 'At home or on property' ~ "Home", 
+                     origin == 'At place of work or school' ~ "Work or school",
+                     origin == 'Away at business' ~ "Business",
+                     origin == 'At someone else\'s home or property' ~ "Other's home",
+                     TRUE ~ origin),
+         orig_label = factor(orig_label))
+```
+
+Similarly to the step before, we'll select the destinations of interest
+and create the label for the `destination` variable:
+
+```{r}
+cycling_2022_spss <- cycling_2022_spss |>  
+  filter(destination == 'At home or on property' | 
+           destination == 'At place of work or school' | 
+           destination == 'Grocery store, other stores or mall'| 
+           destination == 'In the neighbourhood'| 
+           destination == 'Sports centre, field or arena'| 
+           destination == 'Outdoors'| 
+           destination == 'Restaurant, bar or club'| 
+           destination == 'At someone else\'s home or property'| 
+           destination == 'Medical, dental or other health clinic'| 
+           destination == 'Library, museum or theatre'| 
+           destination == 'Place of worship' |
+           destination == 'Away at business') |>
+  mutate(dest_label =
+           case_when(destination == 'At home or on property' ~ "Home", 
+                     destination == 'At place of work or school' ~ "Work or school",
+                     destination == 'Away at business' ~ "Business",
+                     destination == 'At someone else\'s home or property' ~ "Other's home",
+                     TRUE ~ destination), 
+         dest_label = factor(dest_label))
+```
+
+Creating the `year` and `mode` columns. Additionally, the column named
+`TUI_01` will be renamed to `ACTCODE` to ensure consistency with the
+naming conventions used in other years.
+
+```{r}
+cycling_2022 <- cycling_2022_spss %>% 
+   dplyr::select(-c(origin,destination)) %>% 
+   left_join(cycling_2022_sas, by = c('PUMFID', 'INSTANCE'))
+
+cycling_2022 <- cycling_2022 |>
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION,
+                origin, destination, orig_label, dest_label) |>
+  mutate(YEAR = 2022,
+         MODE = "Cycling") 
+```
+
+Now, let's select only the cases of respondents who live in a CMA or CA:
+```{r}
+cycling_2022 <- cycling_2022 |> 
+  left_join(gss_m_2022, by = 'PUMFID') |>  # Merging the two tables
+  dplyr::select(names(walking_2015))
+
+for (col in names(walking_2015)) {
+  target_class <- class(walking_2015[[col]])
+  cycling_2022[[col]] <- match.fun(paste0("as.", target_class[1]))(cycling_2022[[col]])
+}
+```
+
+Now, let's evaluate the count of episodes between each combination of
+origin and destination:
+
+```{r}
+cycling_2022 |> 
+  group_by(orig_label, 
+           dest_label) |>
+  count() |>
+  pivot_wider(names_from = "dest_label", values_from = "n")
+```
+
+The table above shows that the most common destination for a cycling trip
+that starts at home are `Home` and `Work or school`. It also shows that the second most common cycling trip is from `Grocery store, other stores or mall` to `Home`. Notice, though,
+that quite a few trips start *and* end at Home. What are those supposed
+to be?
+
+Check those home-to-home trips:
+
+```{r include=FALSE}
+cycling_2022 |>
+  filter(orig_label == "Home" & dest_label == "Home")
+```
+
+See `PUMFID` == 5002 and check the cycling episode:
+
+```{r}
+gss_e_2022_spss |>
+  filter(PUMFID == '5002') |>
+  slice(unique(c(which(LOCATION == 'Travel - Bicycle') - 1, 
+                 which(LOCATION == 'Travel - Bicycle'),
+                 which(LOCATION == 'Travel - Bicycle') + 1))) |>
+  dplyr::select(PUMFID, INSTANCE, ACTIVITY, LOCATION, DURATION)
+```
+
+The person was at home doing `Culture, sports events, hobbies, leisure or outdoor activity`, then went for cycling trip that lasted 60 minutes, came back home and spent 120 minutes
+`reading, whatching TV or listening music`. So, this was a recreational/leisure
+trip!
+
+Save data in R format for later use:
+
+```{r}
+usethis::use_data(cycling_2022, 
+                  overwrite = TRUE)
+```
+
 # Creating integrated dataset of walking and cycling trips
@@
 # creating an integrated data frame from 1986 to 2015
-gss_episodes <- rbind(walking_2015, cycling_2015, walking_2010, cycling_2010, walking_2005, cycling_2005, cycling_1998, walking_1998, walking_1992, cycling_1992, walking_1986)
+gss_episodes <- rbind(walking_2022, cycling_2022, walking_2015, cycling_2015, walking_2010, cycling_2010, walking_2005, cycling_2005, cycling_1998, walking_1998, walking_1992, cycling_1992, walking_1986)
 ```

### 2025-04-04 data-raw/02-calculating-impedances.Rmd
Updating the package with the new 2022 data set.

@@
 
-In this R markdown, we'll calculate the impedance functions for walking and cycling for each GSS year (from 1986 to 2015).
+In this R markdown, we'll calculate the impedance functions for walking and cycling for each GSS year (from 1986 to 2022).
 
@@
 The name pattern is as follows: `f_year_mode_dest` + `dest_label`, which means that `f_1986_Walking_destHome` displays the best function for the year 1986, walking mode, Home destination.
-/000
+
 ```{r}
@@
 ```
+
 After selecting the best function for every combination of destination, transportation mode and year, the next two chunks will apply the functions in all surveys, adding the information with the best function and its parameters, AIC/BIC/loglik values, and count of episodes that were used to calculate the function:

### 2025-04-04 data-raw/03-preprocessing-main-files.Rmd
Updating the package with the new 2022 data set.

@@
 library(purrr) # Provides a complete and consistent set of tools for working with functions and vectors
+library(foreign) # Read SPSS files
 ```
@@
 `gss_main_20105` refers to the main file of the 2005 GSS survey.
+
+## GSS 2022
+
+```{r reading-2022-main-file}
+# Main File
+gss_m_2022 <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_main_2022_pumf_v1.sav"), to.data.frame = TRUE) # Original file
+```
+
+The main data set for 2022 is already labeled and only needs a selection of socioeconomic and demographic variables:
+
+```{r filtering-variables-2022}
+gss_main_2022 <- gss_m_2022 %>%
+  dplyr::select("PUMFID", "WGHT_PER","SURVMNTH", "PRV", "LUC_RST", "REGION",  "HSDSIZEC","AGEGR10", "GENDER2",
+"MARSTAT", "PHSDFLG", "AGEPRC",  "GENPR2",  "COUPGEN2","CHH0017C","CHH0004C","CHH0514C","CHH1517C","CXRFLAG",
+"PARNUM",  "SENFLAG", "MULTIGEN","DDEV_FL", "DDEX_FL", "DFLEX_FL","DHEAR_FL","DLRN_FL", "DMEM_FL", "DMENT_FL",
+ "DMOB_FL", "DPAIN_FL","DUNK_FL", "DVIS_FL", "DDIS_FL", "DTYPER",  "DCLASS",  "TCS_110", "TCS_120", "TCS_130",
+ "TCS_140", "TCS_150", "TCS_160", "TCS_170", "TCS_180", "TCS_190", "TCS_200",  "GEN_01",  "GEN_02",  "SRS_10",  "MSS_130", "ACT7DAYC","MRW_05C", "MRW_D40A",
+"MRW_D40B","WET_120",  "WHW_230", "WHWD140G","TLWK_01A","TLWK_06","WLYD170G",
+"ATT_120", "ATT_130", "ATT_131",
+"ATT_150C","ATT_160", "ATT_170",
+"CTW_140A","CTW_140B","CTW_140C","CTW_140D",
+"CTW_140E","CTW140GR","CTW_140I","CTW_150G","CTW_160", "CTW_170", "CTW_180", "CTW_190", "CTW_210", "CTW_215",
+"ED_05",  "MAP_110C","MAP_D40A","DOR_110C","IM_01A",  "YRARRI_C","BPP_01A", "BPR_20", "ABM_01A",
+"AIP_01A", "VISMINFL","VISPRFL", "REL_02",  "RELIGFLG","INC_C",  
+"FAMINC_C","GTU_110", "GTU_130",  "DDAY",  "DVTDAY")
+```
+
+Save data in R format for later use:
+```{r save-data-2022}
+usethis::use_data(gss_main_2022, 
+                  overwrite = TRUE)
+```
 

### 2025-04-04 README.Rmd
inclusion of section for python integration and how to use this package

@@
 Origins and destinations are categorized, and it is possible to investigate active travel (number of episodes and their duration) for broad categories of purposes (e.g., travel to work or school, travel to other's home).
+{ActiveCA} also include some socioeconomic variables from the GSS Main Files to facilitate profiling the survey's respondents.
 
-The data covers cycles Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015), spanning a period of almost thirty years.
+The data covers cycles Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 2022, spanning a period of more than thirty five years.
 

### 2025-04-04 data-raw/01-preprocessing-episode-files.Rmd

@@
   left_join(gss_m_2022, by = 'PUMFID') |>  # Merging the two tables
-  dplyr::select(names(walking_2015))
+  dplyr::select(names(walking_2015)) %>% 
+  mutate(PUMFID = as.numeric(as.character(PUMFID)),
+         WGHT_EPI = as.numeric(as.character(WGHT_EPI)))
 

### 2025-04-04 data-raw/02-calculating-impedances.Rmd

@@
 (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
-from 1986 to 2015 in Canada to calculate the impedance function for
+from 1986 to 2022 in Canada to calculate the impedance function for
 cycling and walking trips.

### 2025-04-04 data-raw/03-preprocessing-main-files.Rmd

@@
 "AIP_01A", "VISMINFL","VISPRFL", "REL_02",  "RELIGFLG","INC_C",  
-"FAMINC_C","GTU_110", "GTU_130",  "DDAY",  "DVTDAY")
+"FAMINC_C","GTU_110", "GTU_130",  "DDAY",  "DVTDAY") %>% 
+  mutate(PUMFID = as.numeric(as.character(PUMFID)),
+         WGHT_PER = as.numeric(as.character(WGHT_PER)))
 ```

### 2025-05-13 data-raw/01-preprocessing-episode-files.Rmd
Fixing 2022 data set columns format

--- 2025-04-04 data-raw/01-preprocessing-episode-files.Rmd
@@
 
-for (col in names(walking_2015)) {
-  target_class <- class(walking_2015[[col]])
-  walking_2022[[col]] <- match.fun(paste0("as.", target_class[1]))(walking_2022[[col]])
+numeric_columns <- c("WGHT_EPI")
+
+int_columns <- c("PUMFID", "ACTCODE", "STARTIME", "ENDTIME", "DURATION",
+                     "origin", "destination", "LUC_RST",  "PRV")
+
+for (col in numeric_columns) {
+  walking_2022[[col]] <- as.numeric(as.character(walking_2022[[col]]))
+}
+
+for (col in int_columns) {
+  walking_2022[[col]] <- as.integer(as.character(walking_2022[[col]]))
 }
@@
 
-for (col in names(walking_2015)) {
-  target_class <- class(walking_2015[[col]])
-  cycling_2022[[col]] <- match.fun(paste0("as.", target_class[1]))(cycling_2022[[col]])
+for (col in numeric_columns) {
+  cycling_2022[[col]] <- as.numeric(as.character(cycling_2022[[col]]))
+}
+
+for (col in int_columns) {
+  cycling_2022[[col]] <- as.integer(as.character(cycling_2022[[col]]))
 }

### 2025-05-15 data-raw/03-preprocessing-main-files.Rmd
Updating data set variables to fix some columns that were wrongly classified as ordered

@@
                      SEX == 99 ~ "Not stated"), 
-    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))
+    SEX = factor(SEX, levels = values_ordered, ordered = FALSE))
 
@@
                      SEX == 2 ~ "Female"), 
-    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))
+    SEX = factor(SEX, levels = values_ordered, ordered = FALSE))
 
@@
                      DVSEX == 2 ~ "Female"), 
-    DVSEX = factor(DVSEX, levels = values_ordered, ordered = TRUE))
+    DVSEX = factor(DVSEX, levels = values_ordered, ordered = FALSE))
 

### 2025-05-15 R/ActiveCA-package.R

@@ -800,10 +800,10 @@ NULL
 #'   \item{LUC_RST}{Population centres indicator.}
 #'   \item{GTU_110}{How often do you feel rushed? Would you say it is:.}
 #'   \item{GTU_130}{How often do you feel you have time on your hands that you Don't know what to do with? Would you say it is:.}
+#'   \item{DVTDAY}{Type of day of the designated day of interview.}
 #'   \item{SLM_01}{Using a scale of 0 to 10 where 0 means "Very dissatisfied" and 10 means "Very satisfied", how do you feel about your life as a whole right now?}
 #'   \item{SRH_110}{In general, would you say your health is...?}
-#'   \item{SRH_150}{In general, would you say your mental health is...?}
-#'   \item{DVTDAY}{Type of day of the designated day of interview.}
+#'   \item{SRH_115}{In general, would you say your mental health is...?}
 #'}
 #' @docType data
 #' @keywords gss main respondent socio economic 2015
@@ -823,13 +823,11 @@ NULL
 #'   \item{PUMFID}{Record identification.}
 #'   \item{WGHT_PER}{Person weight.}
 #'   \item{SURVMNTH}{Survey month of data collection.}
-#'   \item{AGEGR10}{Age group of the respondent (group of 10).}
-#'   \item{SEX}{Sex of respondent.}
 #'   \item{PRV}{Province of residence of the respondent.}
 #'   \item{LUC_RST}{Population centres indicator.}
 #'   \item{REGION}{Region of residence of the respondent.}
 #'   \item{HSDSIZEC}{Household size of respondent.}
-#'   \item{AGEPRGRD}{Age group of respondent (groups of 10).}
+#'   \item{AGEGR10}{Age group of the respondent (group of 10).}
 #'   \item{GENDER2}{Gender (2) after distribution of the non-binary persons (Men+ or Women+).}
 #'   \item{MARSTAT}{Marital status of the respondent.}
 #'   \item{PHSDFLG}{Respondent has a spouse/partner living in the household.}
@@ -840,12 +838,20 @@ NULL
 #'   \item{CHH0004C}{Child(ren) in household - 0 to 4 years (flag).}
 #'   \item{CHH0514C}{Child(ren) in household - 5 to 14 years (flag).}
 #'   \item{CHH1517C}{Child(ren) in household - 15-17 years (flag).}
[…]

### 2025-05-29 paper/Manuscript-Data-Package-1Round-Review.Rmd
Editions for the first round of revision for the data paper.

--- 2025-01-17 paper/Manuscript-Data-Package.Rmd
@@
   "This paper describes {ActiveCA}, an open data product with Canadian time use data.
-{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
+{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS).
 Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
@@
 # Original main and Episode Files from 2015 to display table examples
-gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
-gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
+gss_m_2015 <- read.csv(paste0(here(),
+                              "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+gss_e_2015 <- read.csv(paste0(here(), 
+                              "/data-raw/source-files/Time-use-2015/gss-e.csv"))
 
 # Processed Main Files
+data("gss_main_2022")
 data("gss_main_2015")
@@
 # Renaming 'dest_label' and 'orig_label' values to reduce table space
-gss_episodes_renamed <- gss_episodes %>%
+gss_episodes_renamed <- gss_episodes |>
   mutate(dest_label = case_when(
@@
 
-gss_impedances_renamed <- gss_impedances %>%
+gss_impedances_renamed <- gss_impedances |>
   mutate(dest_label = case_when(
@@
 
-The objective of this paper is to introduce {ActiveCA}, an open data product with
-time use data from Canada's General Social Surveys. Open data products
+The objective of this paper is to introduce {ActiveCA}, an open data product with data from all Time Use Survey (TUS) cycles of the General Social Surveys (GSSs). Open data products
 (ODPs) are the outcome of a process that transforms raw data (open or not) into 
@@
 {ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
-a period of almost 40 years. The source of these data is Canada's program of 
-General Social Surveys (GSSs). This program is designed to provide cross-sectional
-data on topics of interest to improve the well-being of Canadians. As part of this
+a period of almost 40 years. The source of these data is TUS from the GSS.
+This program is designed to provide cross-sectional data on topics of interest to improve the well-being of Canadians. As part of this 
 program, every five to seven years the survey is done on the topic of time use. 
 Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
-24 (2010), and 29 (2015) of the GSS. Time use data in these surveys is coded using
+24 (2010), 29 (2015), and 34 (2022) of the GSS. Time use data in these surveys is coded using
 a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
@@
 
-For {ActiveCA} and using GSS Public Use Microdata Files (PUMFs), we extracted all data needed to characterize active travel 
+For {ActiveCA} and using TUS - GSS Public Use Microdata Files (PUMFs), we extracted all data needed to characterize active travel 
 in Canada, namely, episodes where the activity was identified as moving between 
@@
 
-# General Social Survey (GSS) collection
+# The Time Use Survey (TUS) collection
 
@@
 The topics of the survey cycle every few years to cover topics that include family,
-health, social identity, and every five to seven years time use. The first Canadian
-time use survey done as part of the GSS program was conducted in 1986, and the 
-most recent was completed in 2015. These Time Use Surveys [@statisticscanada2022]
+health, social identity, and every five to seven years time use (TUS). The first Canadian
+TUS done as part of the GSS program was conducted in 1986, and the 
+most recent was completed in 2022. These Time Use Surveys [@statisticscanada2022]
 collect data on respondents' participation and time spent on a wide range of 
@@
 work-life balance, health, commuting, culture and sports, and many others.
+
+TUS allows researchers to identify the origin and destination of trips, travel times and modes of transport used, providing a valuable dataset for analyzing active travel behavior.
+It also provides the empirical basis for tools used in transportation analysis, such as the development of impedance functions for accessibility analysis, a measure of the ease with which people can reach destinations and opportunities [@hansen1959; @paez2012].
+The Canadian TUS is unique at the national level in collecting detailed information on travel behaviour.
+Its consistent application across survey cycles enables the identification of long-term trends, with some questions present in the questionnaires since the first application of the survey.
+
+Most respondents to the 2022 TUS completed it online, reflecting Statistics Canada's effort to adapt to technological changes and growing time demands by offering greater flexibility and convenience [-@statisticscanada2022].
+While such methodological changes may affect data comparability over time, it is not possible to determine whether observed differences result from actual population changes or shifts in data collection methods.
+Despite rigorous efforts to ensure data quality, the use of electronic questionnaires may have influenced estimates.
+Statistics Canada assessed the impact of collection mode on a limited set of key questions, constrained by sample size.
+Importantly, none of the variables used in this research are flagged in the 2022 PUMF User Guide as unsuitable for trend analysis.
+
+Until 2022, Statistics Canada employed a telephone-based sampling frame, which was replaced by a dwelling-based frame in the most recent cycle.
+Each survey cycle spans a 12-month period, typically from July to the following July.
+The target population includes all Canadians aged 15 and over, excluding residents of the Yukon, Northwest Territories, and Nunavut, full-time institutional residents, and individuals living on Indigenous reserves.
+
+The survey encompasses both rural and urban areas, including metropolitan and non-metropolitan regions, to ensure a diverse and representative sample.
+For sampling, the ten provinces were divided into geographic strata.
+Several Census Metropolitan Areas (CMAs) - such as St.
+John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver - were treated as separate strata.
+Additional strata grouped other CMAs within Quebec, Ontario, and British Columbia, as well as non-CMA areas within each province.
 
@@
 As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not labeled.
-Adittionaly, the format of the tables (comma-separated values) does not allow for the specification of variable types (whether a variable is continuous or discrete), which can lead to mistakes analysts who have limited experience working with PUMFs.
+Additionally, the format of the tables (comma-separated values) does not allow for the specification of variable types (whether a variable is continuous or discrete), which can lead to mistakes analysts who have limited experience working with PUMFs.
 
 ```{r gss-main-file-2015}
-main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  %>%
+main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  |>
   kable(
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
+    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") |>
   kable_styling(full_width = FALSE, 
@@
                 latex_options=c("scale_down"),
-                 font_size = 8) %>%
+                 font_size = 8) |>
   footnote(general = c("Legend: `PUMFID`: record identification.
@@
 ep_2015_unprocessed <- gss_e_2015 |>
-  filter(PUMFID == 10041 & EPINO <= 7) %>%
-  dplyr::select(PUMFID, EPINO,WGHT_EPI,TUI_01, DURATION, LOCATION) %>%
+  filter(PUMFID == 10041 & EPINO <= 7) |>
+  dplyr::select(PUMFID, EPINO,WGHT_EPI,TUI_01, DURATION, LOCATION) |>
   kable(
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") %>%
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") |>
   kable_styling(full_width = FALSE, 
@@
 
-By running `?ActiveCA` in `R`, you can access the complete package documentation.
-The documentation includes a description of the package, information about the authors and maintainers, and an overview of all pre-processed data sets included in the package.
+The documentation of the package can be accessed by running `?ActiveCA` in `R`.
+The documentation includes a description of the package, information about the authors, contributors, and maintainers, and an overview of all pre-processed data sets included in the package.
 
-Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (`walking_2015.rda`).
+Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2022 GSS (`walking_2022.rda`).
 The documentation provides a brief description of the dataset, instructions on how to load the package, as well as details about the file format and descriptions of all variables (already labeled).
 
-```{r walking_documentation, echo=FALSE, out.width="100%", fig.cap=paste0("Documentation for the dataset of walking episodes from 2015."), fig.align="center"}
+```{r walking_documentation, echo=FALSE, out.width="100%", fig.cap=paste0("Documentation for the dataset of walking episodes from 2022."), fig.align="center"}
 
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking-2015-documentation.png"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking-2022-documentation.png"))
 ```
 
-The chunk below shows how to load the walking episodes from the 2015 GSS (Cycle 29):
+The code below shows how to load the walking episodes from the 2022 GSS (Cycle 29):
 
-```{r load-walking-2015-episodes, echo=TRUE}
-data(walking_2015)
+```{r load-walking-2022-episodes, echo=TRUE}
+data(walking_2022)
 ```
@@
 The number of episodes and persons represented by a episode or person is determined by the weight or weighting factor.
-Because of this, every estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
+Because of this, estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
 
-For instance, if someone wants to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weights.
-In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group.
-It uses the `dplyr`package [@dplyr2023] to manipulate the data.
+For instance, to calculate the percentage of respondents from the 2022 GSS survey with active travel episodes, it is necessary to account for the person's weight.
+In 2022, the weight variable is represented by `WGHT_PER.` The code below demonstrates how to obtain the percentage of people with active travel episodes by age group.
+It uses the `dplyr` package [@dplyr2023] to manipulate the data.
 
 The process begins by creating a dataset that sums the population by age group.
-Then, it joins the 2015 episodes with the 2015 Main File.
+Then, it joins the 2022 episodes with the 2022 Main File.
 Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
@@
 
-The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 37%, followed by those between 25 and 34 years old with around 33%.
-There is a significant drop in percentage for the following groups, with the percentage falling to between 15% and 18% for the remaining age groups.
-
-```{r percentages-act, echo=TRUE, out.width='80%'}
+```
 # Load the dplyr library for data manipulation
@@
 # Calculate the total population for each age group
-Total_population <- gss_main_2015 %>% 
-  group_by(AGEGR10) %>% 
+Total_population <- gss_main_2022 |> 
+  group_by(AGEGR10) |> 
   summarise(Total_population  = sum(WGHT_PER))
@@
 # Calculate the active population for each age group
-Active_population <- gss_episodes %>% 
-      filter(YEAR == 2015) %>% 
-  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
-  group_by(AGEGR10) %>% 
+Active_population <- gss_episodes |> 
+      filter(YEAR == 2022) |> 
+  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
+  group_by(AGEGR10) |> 
   summarise(Active_population  = sum(WGHT_PER))
@@
 #and calculate the percentage
-Active_percentage <- Total_population %>% 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
+Active_percentage <- Total_population |> 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
+  mutate(
+    Percentage = round(Active_population/Total_population*100,2))
+```
+The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 44%, followed by those between 25 and 34 years old with around 19%.
+There is a significant drop in percentage for the following groups, with the percentage falling to between 12% and 15% for the remaining age groups.
+
+```{r percentages-act, echo=FALSE, out.width='80%'}
+# Load the dplyr library for data manipulation
+library('dplyr')
+
+# Load the episodes dataset
+data(gss_episodes)
+
+# Calculate the total population for each age group
+Total_population <- gss_main_2022 |> 
+  group_by(AGEGR10) |> 
+  summarise(Total_population  = sum(WGHT_PER))
+
+# Calculate the active population for each age group
+Active_population <- gss_episodes |> 
+      filter(YEAR == 2022) |> 
+  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
+  group_by(AGEGR10) |> 
+  summarise(Active_population  = sum(WGHT_PER))
+
+# Combine total and active population 
+#and calculate the percentage
+Active_percentage <- Total_population |> 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
   mutate(
@@
 # Display the final result
-Active_percentage
+Active_percentage |>
+  kable()
 ```
@@
 df_main_processed <- data.frame(
-  'Survey' = c("Main","Main","Main","Main","Main","Main"),
-  'Year' = c(2015, 2010, 2005, 1998, 1992, 1986),
-  'Count' = c(nrow(gss_main_2015),
+  'Survey' = c("Main", "Main","Main","Main","Main","Main","Main"),
+  'Year' = c(2022, 2015, 2010, 2005, 1998, 1992, 1986),
+  'Count' = c(nrow(gss_main_2022),
+              nrow(gss_main_2015),
               nrow(gss_main_2010),
@@
               nrow(gss_main_1986)),
-  'Weighted' = c(sum(gss_main_2015$WGHT_PER),
+  'Weighted' = c(sum(gss_main_2022$WGHT_PER),
+              sum(gss_main_2015$WGHT_PER),
               sum(gss_main_2010$WGHT_PER),
@@
 
-df_ep_processed <- gss_episodes %>% 
-  group_by(YEAR) %>%
+df_ep_processed <- gss_episodes |> 
+  group_by(YEAR) |>
   summarise(Survey = 'Episode', 
             Count = n(), 
-            Weighted = sum(WGHT_EPI)) %>%
-  rename(Year = YEAR) %>%
-  dplyr::select(Survey, Year, Count, Weighted) %>% 
+            Weighted = sum(WGHT_EPI)) |>
+  rename(Year = YEAR) |>
+  dplyr::select(Survey, Year, Count, Weighted) |> 
   arrange(-Year)
@@
 df_processed <- rbind(df_main_processed,df_ep_processed)
+
+# For table
+
+df_ep_processed_table <- gss_episodes |> 
+  group_by(YEAR) |>
+  summarise(Survey_ep = 'Episode', 
+            Count_ep = n(), 
+            Weighted_ep = sum(WGHT_EPI)) |>
+  rename(Year_ep = YEAR) |>
+  dplyr::select(Survey_ep, Year_ep, Count_ep, Weighted_ep) |> 
+  arrange(-Year_ep)
+
+df_processed_table <- cbind(df_main_processed,df_ep_processed_table)
 ```
@@
 Table \ref{tab:processed-obs} displays the total number of records processed for Main and Episode Files.
-For the Main Files, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
+For the Main Files, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2022, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2022',]$Weighted)` for 2022, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
 It also presents the total cases of active trips episodes identified.
@@
 
-table_df_processed <- df_processed %>%
+table_df_processed <- df_processed_table |> 
+  select(-c(Survey_ep, Survey, Year_ep)) %>% 
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
-        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") %>%
-  kable_styling(full_width = FALSE, font_size = 10) %>%
-  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) %>%
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major")
+        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
+  kable_styling(full_width = FALSE, font_size = 8) |>
+  add_header_above(c(" " = 1, "Main" = 2, "Episode" = 2 )) %>%
+  column_spec(3, border_right = TRUE)
 
@@
 ```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
-main_2015_processed <- gss_main_2015[1:10, 1:6]  %>%
+main_2015_processed <- gss_main_2015[1:10, 1:6]  |>
   kable(
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
+    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.") |>
   kable_styling(full_width = FALSE, 
@@
 ep_2015_unprocessed <- gss_episodes_renamed |>
-  filter(PUMFID == 10041 & YEAR == 2015) %>%
-  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) %>%
+  filter(PUMFID == 10041 & YEAR == 2015) |>
+  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) |>
   kable(
@@
     caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.", 
-       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) %>% 
+       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) |> 
   kable_styling(full_width = FALSE, 
@@
 
-Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
+Considering all TUS analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
 {ActiveCA} includes all these episodes ready for analysis.
-Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes.
+Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2022, with measures of trip duration in minutes.
 The 1986 survey did not include bicycle trips.
@@
 # Calculate min, mean, and max statistic (duration)
-stats_data <- gss_episodes_renamed %>%
-  group_by(YEAR, MODE) %>%
+stats_data <- gss_episodes_renamed |>
+  group_by(YEAR, MODE) |>
   dplyr::summarise(
@@
     .groups = "drop"
-  ) %>%
-  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") %>%
+  ) |>
+  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") |>
   pivot_wider(names_from = YEAR, values_from = Value)
@@
 # Combine counts and stats into one data frame
-stats_data <- stats_data %>%
-  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) %>%
-  arrange(MODE, Statistic) %>%
+stats_data <- stats_data |>
+  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) |>
+  arrange(MODE, Statistic) |>
   mutate(across(where(is.numeric), ~ round(.x, 0)))
@@
 # Rename the 'MODE' column to 'Mode' before generating the table
-stats_data <- stats_data %>%
-  rename(Mode = MODE) %>%
-  dplyr::select(Mode, Statistic, everything()) %>% 
+stats_data <- stats_data |>
+  rename(Mode = MODE) |>
+  dplyr::select(Mode, Statistic, everything()) |> 
    mutate(Statistic = case_when(
@@
 # Creating a latex table
-styled_ch03_table_01 <- stats_data %>%
+styled_ch03_table_01 <- stats_data |>
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
         caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
-        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) %>%
-  kable_styling(full_width = FALSE, font_size = 10) %>%
-  column_spec(1, bold = TRUE) %>%
-  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) %>%
+        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) |>
+  kable_styling(full_width = FALSE, font_size = 8) |>
+  column_spec(1, bold = TRUE) |>
+  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) |>
   collapse_rows(columns = 1, valign = "top", latex_hline = "major")
@@
 
-Table \ref{tab:table-01} shows that, in general, the median values for walking trips is 10 minutes, except for 1998 when the median was 5 minutes.
-In the case of cycling trips, the duration fluctuated over the years, ranging from 10 to 20 minutes.
+Table \ref{tab:table-01} shows that, until 2022 the median values for walking trips were 10 minutes, increasing to 15 minutes in the last survey.
+In the case of cycling trips, the duration fluctuated over the years, ranging from 10 to 30 minutes.
 The table also highlights very high maximum values, particularly for walking trips, with recorded episodes exceeding 4 hours in all cases.
 
-Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
-statistics for the two modes of transportation, split by destination
-categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
-Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
-walking trips destined for `home` had the highest medians. However, by
-1998, the highest medians shifted to trips to `work or school`, a
-transition that also occurred for cycling trips between 1992 and 1998.
-Table \ref{tab:table-03} indicates that the median duration for walking trips to
-`home` and `work or school` remained at 10 minutes.
-
-```{r table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
-# Filter for selected years and calculate statistics of trip duration
-episodes_stats_before <- gss_episodes %>%
-  filter(YEAR < 2005) %>%
-  group_by(YEAR, dest_label, MODE) %>%
-  dplyr::summarise(
-    count = sum(WGHT_EPI),
-    min = min(DURATION),
-    median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
-    max = max(DURATION),
-    .groups = "drop"
-  )
-
-# Calculate percentages of trips by mode/year 
-total_counts <- episodes_stats_before %>% 
-  group_by(YEAR, MODE) %>% 
-  summarise(total_count = sum(count), 
-            .groups = "drop")
-
-episodes_stats_before <- left_join(episodes_stats_before, total_counts, by = c("YEAR", "MODE"))
-
-episodes_stats_before$Percentage <- round((episodes_stats_before$count / episodes_stats_before$total_count) * 100, 1)
-
-# Pivot data for the table layout
-wide_data <- episodes_stats_before %>%
-  pivot_wider(
-    id_cols = c(dest_label, MODE),
-    names_from = YEAR,
-    values_from = c(min, median, max, Percentage)
-  )  %>% 
-  arrange(MODE, dest_label, desc(Percentage_1998))
-
-wide_data <- wide_data %>%
- dplyr::select(
-    MODE, dest_label,
-    min_1986, median_1986, max_1986, Percentage_1986,
-    min_1992, median_1992, max_1992, Percentage_1992,
-    min_1998, median_1998, max_1998, Percentage_1998
-  ) %>% 
-  arrange(MODE, dest_label)
-
-# Latex table
-styled_ch03_table_02 <- wide_data %>%
-  kable(
-    format = "latex", 
-    booktabs = TRUE, 
-    longtable = TRUE, 
-    align = c(rep("c", ncol(wide_data))),
-    caption = "\\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998",
-    col.names = c("Destination", "Mode","Min","Med","Max", "(%)", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
-  ) %>%
-  kable_styling(full_width = FALSE, 
-                #latex_options=c("striped", "scale_down"), 
-                latex_options=c("scale_down"),
-                 font_size = 6) %>%
-  # column_spec(1, width = "1cm") %>%  # Set width of 'Dest' column to 1.5cm
-  # column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
-  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) %>%
-  row_spec(0, bold = TRUE, align = "c") %>%
-  column_spec(6, border_right = TRUE) %>%
-  column_spec(10, border_right = TRUE) %>%
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
-  footnote(general = c("'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
-
-styled_ch03_table_02 
-```
-
-```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
-# Selecting surveys after 2000 and calculating trip duration statistics 
-episodes_stats_after <- gss_episodes_renamed %>%
-  filter(YEAR > 2000) %>%
-  group_by(YEAR, dest_label, MODE) %>%
-   dplyr::summarise(
-    count = sum(WGHT_EPI),
-    min = min(DURATION),
-    median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
-    max = max(DURATION),
-    .groups = "drop"
-  )
-
-# Calculate percentages of trips by mode/year
-total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
-episodes_stats_after <- left_join(episodes_stats_after, total_counts, by = c("YEAR", "MODE"))
-episodes_stats_after$Percentage <- round((episodes_stats_after$count / episodes_stats_after$total_count) * 100, 1)
-
-# Pivot data for the table layout
-wide_data <- episodes_stats_after %>%
-  pivot_wider(
-    id_cols = c(dest_label, MODE),
-    names_from = YEAR,
-    values_from = c(min, median, max, Percentage)
-  ) %>%
- dplyr::select(
-    MODE, dest_label,
-    min_2005, median_2005, max_2005, Percentage_2005,
-    min_2010, median_2010, max_2010, Percentage_2010,
-    min_2015, median_2015, max_2015, Percentage_2015
-  ) %>% 
-  arrange(MODE, dest_label)
-
-# Latex table
-styled_ch03_table_03 <- wide_data %>%
-  kable(
-    format = "latex", 
-    booktabs = TRUE, 
-    longtable = TRUE, 
-    align = c(rep("c", ncol(wide_data))),
-    caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
-    col.names = c("Destination", "Mode","Min","Med","Max", "(%)", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
-  ) %>%
-  kable_styling(full_width = FALSE, 
-                #latex_options=c("striped", "scale_down"), 
-                latex_options=c("scale_down"),
-                font_size = 6) %>%
-  #column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
-  #column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
-  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
-  row_spec(0, bold = TRUE, align = "c") %>%
-  column_spec(6, border_right = TRUE) %>%
-  column_spec(10, border_right = TRUE) %>% 
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
-  footnote(general = c("'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
-
-styled_ch03_table_03
-```
-
 {ActiveCA} also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques.
-Figures \ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling trips from 1992 and 2015 through heat maps.
-These maps use color gradients to represent the percentage of trips between various origins and destinations, with darker colors indicating higher percentages and lighter colors representing less frequent routes.
+Figure \ref{fig:figure-01} shows walking trips from 2022 through heat maps.
+This graph uses color gradients to represent the percentage of trips between various origins and destinations, with darker colors indicating higher percentages and lighter colors representing less frequent routes.
 For conciseness, we omitted the heat maps for the other years analyzed.
 
-In 1992, walking trips with `home` as both the origin and destination
-made up the majority, accounting for about 30% of all walking trips.
-These trips often involved leisure activities, like short walks or dog
-walking. Following this, trips from `home` to `work or school` comprised
-18% of walking trips. Overall, `home` emerged as a crucial hub, either
-as an origin or destination, with only 5% of trips not involving `home.`
-By 2015, `home` remained a significant node, but new locations
-distributed the proportion of trips to areas not considered in 1992. In
-2015, the highest proportion of trips were from `home` to
-`work or school` (12%) and vice versa (11%). `home` to `home` accounted
-for 8% of trips, and `grocery stores` became a notable destination for
-those leaving `home` (6%), surpassing trips to `other's home` (4%).
+In 2022, `home` served as a central hub for most trips, with fewer than 10% of journeys not involving it as either a starting point or destination.
+The most common trip types were from `home` to `work or school` (17%) and the reverse, from `work or school` to `home` (13%).
+Notably, 7% of trips began and ended at `home`, often reflecting leisure activities such as short walks or dog walking.
+`Grocery stores` were also a key destination, comprising 10% of trips departing from `home.`
 
 ```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
+# Walking heat map
+walking_hm_fig <- gss_impedances_renamed %>%
+  filter(YEAR == 2022, MODE == 'Walking') %>%
+  group_by(orig_label, dest_label, YEAR) %>%
+  dplyr::summarise(n = sum(WGHT_EPI), .groups = "drop") %>%
+  group_by(YEAR) %>%
+  mutate(percentage = n / sum(n) * 100) %>%
+  ungroup() %>%
+  ggplot(aes(x = dest_label,
+    y = orig_label)) +
+    geom_tile(aes(fill = percentage)) +
+    labs(x = "Destination",
+    y = "Origin",
+    fill = "Percentage (%)") + 
+    #geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
+    theme_bw() + 
+    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd")) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 
-# Creating heatmaps for the years of 1992 and 2015
-gss_heatmap <- gss_episodes_renamed %>%
-  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) %>% 
-  filter(YEAR %in% c(1992,2015))
-
-for(year in unique(gss_heatmap$YEAR)){
-
-  for(mode in unique(gss_heatmap$MODE)){
-
-    trip_counts <- gss_heatmap %>%
-      filter(YEAR == year, MODE == mode) %>% 
-      group_by(orig_label, dest_label) %>%
-        dplyr::summarise(count_each_trip = n(), 
-                weighted_count = sum(WGHT_EPI), .groups = "drop")
-
-      # Calculate the total weighted count
-      total_weighted_count <- sum(trip_counts$weighted_count)
-
-      # Calculate percentages and round to 2 decimal places
-      trip_counts <- trip_counts %>%
-        mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))
-
-      # Reshape data for heatmap
-      heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")
-
-      # Convert the data frame to a matrix for the heatmap
-      heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
-      rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column
-
-      # Melt the matrix for use with ggplot
-      heatmap_melted <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))
-
-      # Define the order of destinations 
-      destination_order <- c("Home", "Work or school", "Grocery store", 
-                             "Others' home", "Outdoors", "Neighbourhood", 
-                             "Cultural venues", "Place of worship", 
-                             "Restaurant", "Sport area", "Business")
-
-      # Heatmap for every combination of year and mode
-      assign(paste0("heatmap_plot_", year, "_", mode),
-               ggplot(heatmap_melted, aes(x = Destination, y = Origin, fill = value)) +
-                  geom_tile(color = "white") +
-                  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
-                                       na.value = "gray95", 
-                                       limits = c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
-                                       breaks =  c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
-                                       labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
-                  #geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
-                  #    aes(label = round(value)), 
-                  #    size = 4, color = "black") +
-                  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
-                  theme_minimal() +
-                  theme(
-                  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
-                  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
-                  axis.title = element_text(size = 16),
-                  legend.title = element_text(size = 16, margin = margin(b = 15)), 
-                  legend.text = element_text(size = 16),
-                  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt")))
-
-             ) 
-  }
-} 
-```
-
-```{r make-figs, include=FALSE}
-# Fixing the size of the heatmaps 
-heatmap_plot_2015_Walking <- heatmap_plot_2015_Walking +
-  coord_fixed(ratio = 1) +
-  theme(
-    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
-    axis.text.x = element_text(size = 14),  
-    axis.text.y = element_text(size = 14))
-
-heatmap_plot_2015_Cycling <- heatmap_plot_2015_Cycling +
-  coord_fixed(ratio = 1) +
-  theme(
-    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
-    axis.text.x = element_text(size = 14),  
-    axis.text.y = element_text(size = 14))
-
-heatmap_plot_1992_Cycling <- heatmap_plot_1992_Cycling +
-  coord_fixed(ratio = 1) +
-  theme(
-    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
-    axis.text.x = element_text(size = 14),  
-    axis.text.y = element_text(size = 14))
-
-heatmap_plot_1992_Walking <- heatmap_plot_1992_Walking +
-  coord_fixed(ratio = 1) +
-  theme(
-    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
-    axis.text.x = element_text(size = 14),  
-    axis.text.y = element_text(size = 14))
-
-# Combine the plots with adjusted spacing
-walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
-
-# Save the combined plot to a file
-ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"), 
+# Saving figure
+ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"), 
        plot = walking_hm_fig,
-       width = 17, 
-       height = 8, 
-       units = "in", 
-       dpi = 300)
-
-# Combine the plots with adjusted spacing
-cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))
-
-# Save the combined plot to a file
-ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"), 
-       plot = cycling_hm_fig,
-       width = 17, 
-       height = 8, 
-       units = "in", 
+       width = 11, 
+       height = 7, 
+       units = "cm", 
        dpi = 300)
@@
 
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"))
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"))
 ```
 
-For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
-this mode of transportation was first included as an activity, the
-majority of trips were from `home` to `work or school`, accounting for
-about 25% of cases. This pattern remained in 2015, with these trips
-representing 30% of the cases. However, a notable change occurred in
-`home` to `home` trips, which decreased significantly from 19% in 1992
-to 5% in 2015.
+The {ActiveCA} dataset also includes information on the type of population centre in which respondents reside - specifically, whether they live in a CMA, a Census Agglomeration (CA), or outside these areas - as well as the respondent’s province.
+This information is important, as patterns of active travel often differ between metropolitan and non-metropolitan populations.
+For example, Table \ref{tab:cma-durations} presents the median walking durations by population centre type and province for 2022.
+Overall, respondents living in CMA/CA areas tend to report higher median walking durations compared to those living outside these centres.
+The most pronounced difference is observed in Nova Scotia: metropolitan residents reported a median walking duration of 30 minutes, whereas non-metropolitan residents reported a median of only 5 minutes.
 
-```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of cycling trips categorized by origin and destination", fig.align="center"}
+```{r cma-durations, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
+median_duration_cma <- gss_episodes_renamed %>%
+  filter(YEAR == 2022, 
+         MODE == 'Walking') %>%
+  group_by(Province, MODE, Pop_centre) %>%
+  summarise(
+    Median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
+    .groups = "drop") %>%
+  select(-MODE) %>% 
+  pivot_wider(
+    names_from = Pop_centre,
+    values_from = Median) %>% 
+  mutate(Province = as.character(Province)) %>%
+  arrange(Province)
 
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"))
+# Latex table
+median_duration_table <- median_duration_cma |>
+  kable(
+    format = "latex", 
+    booktabs = TRUE, 
+    caption = "\\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.") |>
+  kable_styling(full_width = FALSE, 
+                #latex_options=c("striped", "scale_down"), 
+                latex_options=c("scale_down"),
+                 font_size = 8) |>
+  add_header_above(c(" " = 1, "Population centre type" = 2)) |> 
+  footnote(general = c("'CMA' denotes Census Metropolitan Area and 'CA' denotes Census agglomeration."), threeparttable = T)
+
+median_duration_table 
 ```
@@
 
-```{r main_stress_figure}
-main_stress_figure <- gss_main_2015 %>%
-  filter(!is.na(MARSTAT), !is.na(GTU_110)) %>%
-  group_by(MARSTAT, GTU_110) %>%
-  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") %>%
-  mutate(percentage = n / sum(n) * 100) %>%
-  ungroup() %>%
+```{r main_stress_figure, include = FALSE}
+main_stress_figure <- gss_main_2015 |>
+  filter(!is.na(MARSTAT), !is.na(GTU_110)) |>
+  group_by(MARSTAT, GTU_110) |>
+  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") |>
+  mutate(percentage = n / sum(n) * 100) |>
+  ungroup() |>
   ggplot(aes(x = MARSTAT,
@@
     fill = "Percentage (%)") + 
-    geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
+    #geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
     theme_bw() + 
-    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd")) +
+    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlGn")) +
     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
@@
        plot = main_stress_figure,
-       width = 15, 
-       height = 10, 
+       width = 12, 
+       height = 8, 
        units = "cm", 
@@
 This paper presents {ActiveCA}, an open data product that provides analysis-ready 
-data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
-of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
+data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) 
+of TUS GSSs on active travel in Canada. In the form of an `R` data package, 
 {ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
@@
 
-It is important to remark that the present version of {ActiveCA} covers all Canadian
-time use surveys up to 2015. While the most recent time use survey was carried out
-in 2022 [@wray2024], the Public Use Microdata Files are currently unavailable, and
-at the moment it is estimated that they will only be published in the later part
-of 2025. The `R` package will be updated once these files become available.
+Although we did not select non-AT episodes, the process for obtaining them is very similar to that used for selecting AT episodes.
+Researchers interested in non-AT modes can use our framework to guide their methodology, making the small but necessary adjustments.
+We focused exclusively on AT episodes because the ActiveCA package is part of a larger project aimed at analyzing the historical evolution of active travel behaviour in Canada.
 
@@
 replication and innovation in related fields. We believe that {ActiveCA} will serve
-as a basis for further research on GSS and for the integration of additional data by
+as a basis for further research on TUS GSS and for the integration of additional data by
 the authors or the wider open source community.
@@
 
+For review purposes, the package is currently available as a tar.gz file that can 
+be installed by R. The file can be obtained from this anonymous location:
+
+https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip
+
 # References {#references .unnumbered}

### 2025-05-29 paper/Manuscript-Data-Package-1Round-Review.tex
Editions for the first round of revision for the data paper.

--- 2025-01-22 paper/Manuscript-Data-Package.tex
@@
 extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
-(2010), and 29 (2015) of Canada's General Social Survey. Active travel
-is characterized by mode, with walking being part of every cycle and
-bicycling starting in 1992. The attributes of active trips are the types
-of locations of origins and destinations, the duration of trips, and
-episode weights for expanding the trips to population-wide estimates.
-Based on the year of the survey, a variety of locations are coded. In
-earlier cycles, these include home, work or school, and other's home,
-whereas in later cycles these are augmented with locations such as
-grocery stores, restaurants, outdoor destinations, and others. The
-geographical resolution includes the province and whether the episode
-was in an urban or rural setting.
+(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles
+from the General Social Survey (GSS). Active travel is characterized by
+mode, with walking being part of every cycle and bicycling starting in
+1992. The attributes of active trips are the types of locations of
+origins and destinations, the duration of trips, and episode weights for
+expanding the trips to population-wide estimates. Based on the year of
+the survey, a variety of locations are coded. In earlier cycles, these
+include home, work or school, and other's home, whereas in later cycles
+these are augmented with locations such as grocery stores, restaurants,
+outdoor destinations, and others. The geographical resolution includes
+the province and whether the episode was in an urban or rural setting.
 \end{abstract}
@@
 The objective of this paper is to introduce \{ActiveCA\}, an open data
-product with time use data from Canada's General Social Surveys. Open
-data products (ODPs) are the outcome of a process that transforms raw
-data (open or not) into analysis-ready data, following a transparent
-process in which all stages of development follow open principles
-\citep{arribas-bel2021}. ODPs, while still open, differ from general
-open data in their degree of ease of access, their heightened usability,
-and potentially the value they add to the raw data.
+product with data from all Time Use Survey (TUS) cycles of the General
+Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
+process that transforms raw data (open or not) into analysis-ready data,
+following a transparent process in which all stages of development
+follow open principles \citep{arribas-bel2021}. ODPs, while still open,
+differ from general open data in their degree of ease of access, their
+heightened usability, and potentially the value they add to the raw
+data.
 
@@
 Canada spanning a period of almost 40 years. The source of these data is
-Canada's program of General Social Surveys (GSSs). This program is
-designed to provide cross-sectional data on topics of interest to
-improve the well-being of Canadians. As part of this program, every five
-to seven years the survey is done on the topic of time use. Concretely,
-\{ActiveCA\} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
-(2010), and 29 (2015) of the GSS. Time use data in these surveys is
-coded using a very fine grain, from time spent in chores, leisure, and
-sleeping, to time spent working or at school. These surveys have proved
-valuable in investigations of mobility and quality of life
+TUS from the GSS. This program is designed to provide cross-sectional
+data on topics of interest to improve the well-being of Canadians. As
+part of this program, every five to seven years the survey is done on
+the topic of time use. Concretely, \{ActiveCA\} covers Cycles 2 (1986),
+7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of
+the GSS. Time use data in these surveys is coded using a very fine
+grain, from time spent in chores, leisure, and sleeping, to time spent
+working or at school. These surveys have proved valuable in
+investigations of mobility and quality of life
 \citep{spinneyTransport2009}, the relationship between active travel and
@@
 
-For \{ActiveCA\} and using GSS Public Use Microdata Files (PUMFs), we
-extracted all data needed to characterize active travel in Canada,
+For \{ActiveCA\} and using TUS - GSS Public Use Microdata Files (PUMFs),
+we extracted all data needed to characterize active travel in Canada,
 namely, episodes where the activity was identified as moving between an
@@
 
-\section{General Social Survey (GSS)
-collection}\label{general-social-survey-gss-collection}
+\section{The Time Use Survey (TUS)
+collection}\label{the-time-use-survey-tus-collection}
 
@@
 include family, health, social identity, and every five to seven years
-time use. The first Canadian time use survey done as part of the GSS
-program was conducted in 1986, and the most recent was completed in
-2015. These Time Use Surveys \citep{statisticscanada2022} collect data
-on respondents' participation and time spent on a wide range of everyday
+time use (TUS). The first Canadian TUS done as part of the GSS program
+was conducted in 1986, and the most recent was completed in 2022. These
+Time Use Surveys \citep{statisticscanada2022} collect data on
+respondents' participation and time spent on a wide range of everyday
 activities using a 24-hour retrospective diary, with information on the
@@
 culture and sports, and many others.
+
+TUS allows researchers to identify the origin and destination of trips,
+travel times and modes of transport used, providing a valuable dataset
+for analyzing active travel behavior. It also provides the empirical
+basis for tools used in transportation analysis, such as the development
+of impedance functions for accessibility analysis, a measure of the ease
+with which people can reach destinations and opportunities
+\citep{hansen1959, paez2012}. The Canadian TUS is unique at the national
+level in collecting detailed information on travel behaviour. Its
+consistent application across survey cycles enables the identification
+of long-term trends, with some questions present in the questionnaires
+since the first application of the survey.
+
+Most respondents to the 2022 TUS completed it online, reflecting
+Statistics Canada's effort to adapt to technological changes and growing
+time demands by offering greater flexibility and convenience
+\citeyearpar{statisticscanada2022}. While such methodological changes
+may affect data comparability over time, it is not possible to determine
+whether observed differences result from actual population changes or
+shifts in data collection methods. Despite rigorous efforts to ensure
+data quality, the use of electronic questionnaires may have influenced
+estimates. Statistics Canada assessed the impact of collection mode on a
+limited set of key questions, constrained by sample size. Importantly,
+none of the variables used in this research are flagged in the 2022 PUMF
+User Guide as unsuitable for trend analysis.
+
+Until 2022, Statistics Canada employed a telephone-based sampling frame,
+which was replaced by a dwelling-based frame in the most recent cycle.
+Each survey cycle spans a 12-month period, typically from July to the
+following July. The target population includes all Canadians aged 15 and
+over, excluding residents of the Yukon, Northwest Territories, and
+Nunavut, full-time institutional residents, and individuals living on
+Indigenous reserves.
+
+The survey encompasses both rural and urban areas, including
+metropolitan and non-metropolitan regions, to ensure a diverse and
+representative sample. For sampling, the ten provinces were divided into
+geographic strata. Several Census Metropolitan Areas (CMAs) - such as
+St.~John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
+Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver
+- were treated as separate strata. Additional strata grouped other CMAs
+within Quebec, Ontario, and British Columbia, as well as non-CMA areas
+within each province.
 
@@
 As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
-labeled. Adittionaly, the format of the tables (comma-separated values)
+labeled. Additionally, the format of the tables (comma-separated values)
 does not allow for the specification of variable types (whether a
@@
 Figure \ref{fig:walking_documentation} displays the documentation for
-the walking episodes from the 2015 GSS (\texttt{walking\_2015.rda}). The
+the walking episodes from the 2022 GSS (\texttt{walking\_2022.rda}). The
 documentation provides a brief description of the dataset, instructions
@@
 
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2015-documentation}
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2022-documentation}
 
@@
 
-\caption{Documentation for the dataset of walking episodes from 2015.}\label{fig:walking_documentation}
+\caption{Documentation for the dataset of walking episodes from 2022.}\label{fig:walking_documentation}
 \end{figure}
 
-The code below shows how to load the walking episodes from the 2015 GSS
+The code below shows how to load the walking episodes from the 2022 GSS
 (Cycle 29):
@@
 \begin{verbatim}
-data(walking_2015)
+data(walking_2022)
 \end{verbatim}
@@
 
-For instance, to calculate the percentage of respondents from the 2015
+For instance, to calculate the percentage of respondents from the 2022
 GSS survey with active travel episodes, it is necessary to account for
-the person's weight. In 2015, the weight variable is represented by
+the person's weight. In 2022, the weight variable is represented by
 \texttt{WGHT\_PER.} The code below demonstrates how to obtain the
@@
 The process begins by creating a dataset that sums the population by age
-group. Then, it joins the 2015 episodes with the 2015 Main File. Note
+group. Then, it joins the 2022 episodes with the 2022 Main File. Note
 that in both operations, the code sums the person weight variable
@@
 # Calculate the total population for each age group
-Total_population <- gss_main_2015 |> 
+Total_population <- gss_main_2022 |> 
   group_by(AGEGR10) |> 
@@
 Active_population <- gss_episodes |> 
-      filter(YEAR == 2015) |> 
-  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) |> 
+      filter(YEAR == 2022) |> 
+  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
   group_by(AGEGR10) |> 
@@
 The result shows that the age group with the highest share of active
-trips is those between 15 and 20 years old, with almost 37\%, followed
-by those between 25 and 34 years old with around 33\%. There is a
+trips is those between 15 and 20 years old, with almost 44\%, followed
+by those between 25 and 34 years old with around 19\%. There is a
 significant drop in percentage for the following groups, with the
-percentage falling to between 15\% and 18\% for the remaining age
+percentage falling to between 12\% and 15\% for the remaining age
 groups.
@@
 \endlastfoot
-15 to 24 years & 4511131 & 1641468.3 & 36.39 \\
-25 to 34 years & 4956386 & 1644032.8 & 33.17 \\
-35 to 44 years & 4734506 & 847769.9 & 17.91 \\
-45 to 54 years & 5136125 & 790667.3 & 15.39 \\
-55 to 64 years & 4831306 & 767064.0 & 15.88 \\
-65 to 74 years & 3283969 & 559024.6 & 17.02 \\
-75 years and over & 2312976 & 384358.6 & 16.62 \\
+15 to 24 years & 4537139 & 1984658.8 & 43.74 \\
+25 to 34 years & 5364748 & 1048653.8 & 19.55 \\
+35 to 44 years & 5227988 & 669688.8 & 12.81 \\
+45 to 54 years & 4720990 & 605162.1 & 12.82 \\
+55 to 64 years & 5143935 & 690336.5 & 13.42 \\
+65 to 74 years & 4160941 & 619732.8 & 14.89 \\
+75 years and over & 2981061 & 422799.5 & 14.18 \\
 \end{longtable}
@@
 processed for Main and Episode Files. For the Main Files, a total of
-89,331 records were processed, referring to all records from the Time
-Use Surveys from 1986 to 2015, that together represents more of
-149,389,839 respondents (29,766,399 for 2015, 28,075,610 for 2010,
-26,095,819 for 2005, 24,260,137 for 1998, 21,294,313 for 1992, and for
-2005, 24,260,137 for 1998, and 21,294,313 for 1986). It also presents
-the total cases of active trips episodes identified. In total 21,748
-records with register of active travel activity. Together, these records
-account for 38,275,077 episodes.
+101,667 records were processed, referring to all records from the Time
+Use Surveys from 1986 to 2022, that together represents more of
+181,526,641 respondents (32,136,802 for 2022, 29,766,399 for 2015,
+28,075,610 for 2010, 26,095,819 for 2005, 24,260,137 for 1998,
+21,294,313 for 1992, and for 2005, 24,260,137 for 1998, and 21,294,313
+for 1986). It also presents the total cases of active trips episodes
+identified. In total 23,513 records with register of active travel
+activity. Together, these records account for 44,316,110 episodes.
 
-\begingroup\fontsize{10}{12}\selectfont
+\begingroup\fontsize{8}{10}\selectfont
 
-\begin{longtable}[t]{lrrr}
+\begin{longtable}[t]{rr>{}r|rr}
 \caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Observations} \\
-\cmidrule(l{3pt}r{3pt}){3-4}
-Survey & Year & Count & Weighted\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Main} & \multicolumn{2}{c}{Episode} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+Year & Count & Weighted & Count\_ep & Weighted\_ep\\
 \midrule
- & 2015 & 17390 & 29766399\\
-\nopagebreak
- & 2010 & 15390 & 28075610\\
-\nopagebreak
- & 2005 & 19597 & 26095819\\
-\nopagebreak
- & 1998 & 10749 & 24260137\\
-\nopagebreak
- & 1992 & 9815 & 21294313\\
-\nopagebreak
-\multirow[t]{-6}{*}{\raggedright\arraybackslash Main} & 1986 & 16390 & 19897562\\
-\cmidrule{1-4}\pagebreak[0]
- & 2015 & 3496 & 6634387\\
-\nopagebreak
- & 2010 & 4615 & 8516753\\
-\nopagebreak
- & 2005 & 5866 & 7583838\\
-\nopagebreak
- & 1998 & 1789 & 3606987\\
-\nopagebreak
- & 1992 & 1635 & 3691918\\
-\nopagebreak
-\multirow[t]{-6}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
+2022 & 12336 & 32136802 & 1765 & 6041032\\
+2015 & 17390 & 29766399 & 3496 & 6634387\\
+2010 & 15390 & 28075610 & 4615 & 8516753\\
+2005 & 19597 & 26095819 & 5866 & 7583838\\
+1998 & 10749 & 24260137 & 1789 & 3606987\\
+\addlinespace
+1992 & 9815 & 21294313 & 1635 & 3691918\\
+1986 & 16390 & 19897562 & 4347 & 8241196\\
 \bottomrule
@@
 \begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.}\\
+\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}\\
 \toprule
@@
 
-Considering GSS Cycles analyzed, we identified 21,748 episodes that
+Considering all TUS analyzed, we identified 23,513 episodes that
 recorded active travel episodes, with trip duration ranging from 0 to
@@
 descriptive statistics on walking and cycling trips between 1986 and
-2015, with measures of trip duration in minutes. The 1986 survey did not
+2022, with measures of trip duration in minutes. The 1986 survey did not
 include bicycle trips.
 
-\begingroup\fontsize{10}{12}\selectfont
+\begingroup\fontsize{8}{10}\selectfont
 
-\begin{longtable}[t]{>{}llcccccc}
+\begin{longtable}[t]{>{}llccccccc}
 \caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
-\cmidrule(l{3pt}r{3pt}){3-8}
-Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
+\multicolumn{2}{c}{ } & \multicolumn{7}{c}{Year} \\
+\cmidrule(l{3pt}r{3pt}){3-9}
+Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015 & 2022\\
 \midrule
- & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
+ & Maximum & 660 & 300 & 255 & 515 & 480 & 900 & 480\\
 \nopagebreak
- & Mean & 21 & 21 & 12 & 12 & 13 & 18\\
+ & Mean & 21 & 21 & 12 & 12 & 13 & 18 & 19\\
 \nopagebreak
- & Median & 15 & 10 & 5 & 10 & 10 & 10\\
+ & Median & 15 & 10 & 5 & 10 & 10 & 10 & 15\\
 \nopagebreak
- & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
+ & Minimum & 1 & 1 & 1 & 0 & 0 & 5 & 5\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
-\cmidrule{1-8}\pagebreak[0]
- & Maximum &  & 240 & 90 & 180 & 153 & 120\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27 & 24\\
+\cmidrule{1-9}\pagebreak[0]
+ & Maximum &  & 240 & 90 & 180 & 153 & 120 & 150\\
 \nopagebreak
- & Mean &  & 28 & 24 & 20 & 19 & 25\\
+ & Mean &  & 28 & 24 & 20 & 19 & 25 & 40\\
 \nopagebreak
- & Median &  & 15 & 15 & 15 & 10 & 20\\
+ & Median &  & 15 & 15 & 15 & 10 & 20 & 30\\
 \nopagebreak
- & Minimum &  & 5 & 2 & 1 & 1 & 5\\
+ & Minimum &  & 5 & 2 & 1 & 1 & 5 & 5\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20 & 20\\
 \bottomrule
@@
 
-Table \ref{tab:table-01} shows that, in general, the median values for
-walking trips is 10 minutes, except for 1998 when the median was 5
-minutes. In the case of cycling trips, the duration fluctuated over the
-years, ranging from 10 to 20 minutes. The table also highlights very
+Table \ref{tab:table-01} shows that, until 2022 the median values for
+walking trips were 10 minutes, increasing to 15 minutes in the last
+survey. In the case of cycling trips, the duration fluctuated over the
+years, ranging from 10 to 30 minutes. The table also highlights very
 high maximum values, particularly for walking trips, with recorded
@@
 
-Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
-statistics for the two modes of transportation, split by destination
-categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
-Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
-walking trips destined for \texttt{home} had the highest medians.
-However, by 1998, the highest medians shifted to trips to
-\texttt{work\ or\ school}, a transition that also occurred for cycling
-trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
-median duration for walking trips to \texttt{home} and
-\texttt{work\ or\ school} remained at 10 minutes.
+\{ActiveCA\} also enables visual analysis of active travel in Canada
+using traditional exploratory data analysis techniques. Figure
+\ref{fig:figure-01} shows walking trips from 2022 through heat maps.
+This graph uses color gradients to represent the percentage of trips
+between various origins and destinations, with darker colors indicating
+higher percentages and lighter colors representing less frequent routes.
+For conciseness, we omitted the heat maps for the other years analyzed.
 
-\begingroup\fontsize{6}{8}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
-\end{TableNotes}
-\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
-\caption{\label{tab:table-02}\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998}\\
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{4}{c}{1986} & \multicolumn{4}{c}{1992} & \multicolumn{4}{c}{1998} \\
-\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
-\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
-\midrule
- & Home &  &  &  &  & 5 & 20 & 240 & 55.7 & 2 & 15 & 90 & 51.6\\
-\nopagebreak
- & Other's home &  &  &  &  & 5 & 10 & 145 & 19.7 & 2 & 10 & 80 & 15.7\\
-\nopagebreak
-\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 24.7 & 5 & 25 & 75 & 32.8\\
-\cmidrule{1-14}\pagebreak[0]
- & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 & 60.5 & 1 & 5 & 255 & 51.9\\
-\nopagebreak
- & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 & 18.7 & 1 & 5 & 120 & 27.8\\
-\nopagebreak
-\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 & 20.9 & 1 & 10 & 75 & 20.3\\
-\bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
-
-\begingroup\fontsize{6}{8}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
-\end{TableNotes}
-\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
-\caption{\label{tab:table-03}\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015}\\
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{4}{c}{2005} & \multicolumn{4}{c}{2010} & \multicolumn{4}{c}{2015} \\
-\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
-\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
-\midrule
- & Cultural venues & 10 & 10 & 15 & 0.3 & 10 & 25 & 30 & 1.0 & 15 & 15 & 15 & 0.5\\
-\nopagebreak
- & Grocery store & 2 & 10 & 30 & 10.4 & 5 & 10 & 75 & 7.2 & 5 & 15 & 80 & 5.5\\
-\nopagebreak
- & Health clinic &  &  &  &  &  &  &  &  & 10 & 15 & 90 & 1.9\\
-\nopagebreak
- & Home & 1 & 15 & 180 & 48.4 & 1 & 10 & 135 & 48.8 & 5 & 20 & 120 & 46.4\\
-\nopagebreak
- & Neighbourhood &  &  &  &  &  &  &  &  & 10 & 30 & 45 & 1.3\\
-\nopagebreak
- & Other's home & 1 & 10 & 35 & 9.4 & 5 & 10 & 45 & 9.7 & 5 & 15 & 40 & 4.8\\
-\nopagebreak
- & Outdoors & 5 & 15 & 45 & 6.0 & 3 & 10 & 115 & 2.6 & 15 & 30 & 30 & 1.0\\
-\nopagebreak
- & Place of worship & 20 & 20 & 20 & 0.1 &  &  &  &  & 15 & 15 & 15 & 0.2\\
-\nopagebreak
- & Restaurant & 5 & 15 & 35 & 2.8 & 10 & 15 & 153 & 1.7 & 10 & 20 & 60 & 3.0\\
-\nopagebreak
- & Sport area &  &  &  &  &  &  &  &  & 10 & 15 & 15 & 2.0\\
-\nopagebreak
-\multirow[t]{-11}{*}{\centering\arraybackslash Cycling} & Work or school & 1 & 15 & 90 & 22.6 & 1 & 15 & 100 & 29.0 & 5 & 20 & 120 & 33.5\\
-\cmidrule{1-14}\pagebreak[0]
- & Business &  &  &  &  &  &  &  &  & 5 & 10 & 30 & 0.2\\
-\nopagebreak
- & Cultural venues & 5 & 10 & 40 & 0.6 & 2 & 10 & 40 & 0.7 & 5 & 15 & 40 & 1.5\\
-\nopagebreak
- & Grocery store & 1 & 10 & 90 & 11.5 & 1 & 7 & 105 & 13.1 & 5 & 10 & 130 & 10.3\\
-\nopagebreak
- & Health clinic &  &  &  &  &  &  &  &  & 5 & 10 & 130 & 0.9\\
-\nopagebreak
- & Home & 0 & 10 & 515 & 43.0 & 0 & 10 & 270 & 41.4 & 5 & 10 & 900 & 44.0\\
-\nopagebreak
- & Neighbourhood &  &  &  &  &  &  &  &  & 5 & 10 & 60 & 2.6\\
-\nopagebreak
- & Other's home & 1 & 5 & 300 & 10.2 & 0 & 5 & 140 & 10.2 & 5 & 10 & 120 & 6.2\\
-\nopagebreak
- & Outdoors & 1 & 5 & 295 & 3.4 & 0 & 10 & 480 & 5.0 & 5 & 10 & 135 & 3.1\\
-\nopagebreak
- & Place of worship & 1 & 10 & 30 & 0.7 & 1 & 7 & 60 & 0.8 & 5 & 15 & 45 & 1.1\\
-\nopagebreak
- & Restaurant & 0 & 5 & 85 & 9.9 & 1 & 7 & 153 & 11.0 & 5 & 10 & 120 & 9.0\\
-\nopagebreak
- & Sport area &  &  &  &  &  &  &  &  & 5 & 10 & 45 & 3.3\\
-\nopagebreak
-\multirow[t]{-12}{*}{\centering\arraybackslash Walking} & Work or school & 0 & 10 & 175 & 20.7 & 0 & 10 & 150 & 17.9 & 5 & 10 & 190 & 17.8\\
-\bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
-
-\{ActiveCA\} also enables visual analysis of active travel in Canada
-using traditional exploratory data analysis techniques. Figures
-\ref{fig:figure-01} and \ref{fig:figure-02} show walking and cycling
-trips from 1992 and 2015 through heat maps. These maps use color
-gradients to represent the percentage of trips between various origins
-and destinations, with darker colors indicating higher percentages and
-lighter colors representing less frequent routes. For conciseness, we
-omitted the heat maps for the other years analyzed.
-
-In 1992, walking trips with \texttt{home} as both the origin and
-destination made up the majority, accounting for about 30\% of all
-walking trips. These trips often involved leisure activities, like short
-walks or dog walking. Following this, trips from \texttt{home} to
-\texttt{work\ or\ school} comprised 18\% of walking trips. Overall,
-\texttt{home} emerged as a crucial hub, either as an origin or
-destination, with only 5\% of trips not involving \texttt{home.} By
-2015, \texttt{home} remained a significant node, but new locations
-distributed the proportion of trips to areas not considered in 1992. In
-2015, the highest proportion of trips were from \texttt{home} to
-\texttt{work\ or\ school} (12\%) and vice versa (11\%). \texttt{home} to
-\texttt{home} accounted for 8\% of trips, and \texttt{grocery\ stores}
-became a notable destination for those leaving \texttt{home} (6\%),
-surpassing trips to \texttt{other\textquotesingle{}s\ home} (4\%).
+In 2022, \texttt{home} served as a central hub for most trips, with
+fewer than 10\% of journeys not involving it as either a starting point
+or destination. The most common trip types were from \texttt{home} to
+\texttt{work\ or\ school} (17\%) and the reverse, from
+\texttt{work\ or\ school} to \texttt{home} (13\%). Notably, 7\% of trips
+began and ended at \texttt{home}, often reflecting leisure activities
+such as short walks or dog walking. \texttt{Grocery\ stores} were also a
+key destination, comprising 10\% of trips departing from \texttt{home.}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking_hm_fig}
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking_hm_2022}
 
@@
 
-For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
-this mode of transportation was first included as an activity, the
-majority of trips were from \texttt{home} to \texttt{work\ or\ school},
-accounting for about 25\% of cases. This pattern remained in 2015, with
-these trips representing 30\% of the cases. However, a notable change
-occurred in \texttt{home} to \texttt{home} trips, which decreased
-significantly from 19\% in 1992 to 5\% in 2015.
+The \{ActiveCA\} dataset also includes information on the type of
+population centre in which respondents reside - specifically, whether
+they live in a CMA, a Census Agglomeration (CA), or outside these areas
+- as well as the respondent's province. This information is important,
+as patterns of active travel often differ between metropolitan and
+non-metropolitan populations. For example, Table \ref{tab:cma-durations}
+presents the median walking durations by population centre type and
+province for 2022. Overall, respondents living in CMA/CA areas tend to
+report higher median walking durations compared to those living outside
+these centres. The most pronounced difference is observed in Nova
+Scotia: metropolitan residents reported a median walking duration of 30
+minutes, whereas non-metropolitan residents reported a median of only 5
+minutes.
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/cycling_hm_fig}
-
-}
-
-\caption{Percentage of cycling trips categorized by origin and destination}\label{fig:figure-02}
-\end{figure}
+\begin{table}
+\centering
+\caption{\label{tab:cma-durations}\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{8}{10}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{lrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Population centre type} \\
+\cmidrule(l{3pt}r{3pt}){2-3}
+Province & CMA/CA & non CMA/CA\\
+\midrule
+Alberta & 15 & 5\\
+British Columbia & 15 & 10\\
+Manitoba & 10 & 10\\
+New Brunswick & 5 & 5\\
+Newfoundland and Labroador & 20 & 10\\
+\addlinespace
+Nova Scotia & 30 & 5\\
+Ontario & 15 & 15\\
+Prince Edward Island &  & 5\\
+Quebec & 15 & 15\\
+Saskatchewan & 10 & 5\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item 'CMA' denotes Census Metropolitan Area and 'CA' denotes Census agglomeration.
+\end{tablenotes}
+\end{threeparttable}}
+\end{table}
 
@@
 analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
-(2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in
-Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
+(2005), 24 (2010), 29 (2015), and 34 (2022) of TUS GSSs on active travel
+in Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
 developed after collecting, cleaning, and processing the survey data,
@@
 
-It is important to remark that the present version of \{ActiveCA\}
-covers all Canadian time use surveys up to 2015. While the most recent
-time use survey was carried out in 2022 \citep{wray2024}, the Public Use
-Microdata Files are currently unavailable, and at the moment it is
-estimated that they will only be published in the later part of 2025.
-The \texttt{R} package will be updated once these files become
-available.
+Although we did not select non-AT episodes, the process for obtaining
+them is very similar to that used for selecting AT episodes. Researchers
+interested in non-AT modes can use our framework to guide their
+methodology, making the small but necessary adjustments. We focused
+exclusively on AT episodes because the ActiveCA package is part of a
+larger project aimed at analyzing the historical evolution of active
+travel behaviour in Canada.
 
@@
 encourage replication and innovation in related fields. We believe that
-\{ActiveCA\} will serve as a basis for further research on GSS and for
-the integration of additional data by the authors or the wider open
+\{ActiveCA\} will serve as a basis for further research on TUS GSS and
+for the integration of additional data by the authors or the wider open
 source community.

### 2025-05-30 paper/Manuscript-Data-Package-1Round-Review.Rmd
Changing the data paper due to the revision process.

@@
 
-# How to use {ActiveCA}
+# {ActiveCA} data sets
 
-This section presents some potential applications of the {ActiveCA} R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
+This section presents some potential applications of the {ActiveCA} R package.
+In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
+The installation instruction and also some examples of application of the {ActiveCA} R package are available in the `vignettes`, available in the Github repository.
 
-To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the {ActiveCA} package in R:
-
-```{r instal-ActiveCA, echo=TRUE}
-if(!require("ActiveCA", character.only = TRUE)) {
-      remotes::install_github("dias-bruno/ActiveCA")
-}
-```
-
-After installing the {ActiveCA} R package, you can load it using the following code:
-
-```{r load-ActiveCA, echo=TRUE}
-library(ActiveCA)
-```
-
-The documentation of the package can be accessed by running `?ActiveCA` in `R`.
-The documentation includes a description of the package, information about the authors, contributors, and maintainers, and an overview of all pre-processed data sets included in the package.
-
-Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2022 GSS (`walking_2022.rda`).
-The documentation provides a brief description of the dataset, instructions on how to load the package, as well as details about the file format and descriptions of all variables (already labeled).
-
-```{r walking_documentation, echo=FALSE, out.width="100%", fig.cap=paste0("Documentation for the dataset of walking episodes from 2022."), fig.align="center"}
-
-knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking-2022-documentation.png"))
-```
-
-The code below shows how to load the walking episodes from the 2022 GSS (Cycle 29):
-
-```{r load-walking-2022-episodes, echo=TRUE}
-data(walking_2022)
-```
-
-The GSS surveys apply a probability sampling methodology, in which each episode or person selected in the sample represents several other episodes or persons not in the sample.
-The number of episodes and persons represented by a episode or person is determined by the weight or weighting factor.
-Because of this, estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
-
-For instance, to calculate the percentage of respondents from the 2022 GSS survey with active travel episodes, it is necessary to account for the person's weight.
-In 2022, the weight variable is represented by `WGHT_PER.` The code below demonstrates how to obtain the percentage of people with active travel episodes by age group.
-It uses the `dplyr` package [@dplyr2023] to manipulate the data.
-
-The process begins by creating a dataset that sums the population by age group.
-Then, it joins the 2022 episodes with the 2022 Main File.
-Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
-After this, a new dataset, `Active_percentage`, is created by merging both previous datasets.
-The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100 and rounding with 2 decimal places.
-
-```
-# Load the dplyr library for data manipulation
-library('dplyr')
-
-# Load the episodes dataset
-data(gss_episodes)
-
-# Calculate the total population for each age group
-Total_population <- gss_main_2022 |> 
-  group_by(AGEGR10) |> 
-  summarise(Total_population  = sum(WGHT_PER))
-
-# Calculate the active population for each age group
-Active_population <- gss_episodes |> 
-      filter(YEAR == 2022) |> 
-  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
-  group_by(AGEGR10) |> 
-  summarise(Active_population  = sum(WGHT_PER))
-
-# Combine total and active population 
-#and calculate the percentage
-Active_percentage <- Total_population |> 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
-  mutate(
-    Percentage = round(Active_population/Total_population*100,2))
-```
-The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 44%, followed by those between 25 and 34 years old with around 19%.
-There is a significant drop in percentage for the following groups, with the percentage falling to between 12% and 15% for the remaining age groups.
-
-```{r percentages-act, echo=FALSE, out.width='80%'}
-# Load the dplyr library for data manipulation
-library('dplyr')
-
-# Load the episodes dataset
-data(gss_episodes)
-
-# Calculate the total population for each age group
-Total_population <- gss_main_2022 |> 
-  group_by(AGEGR10) |> 
-  summarise(Total_population  = sum(WGHT_PER))
-
-# Calculate the active population for each age group
-Active_population <- gss_episodes |> 
-      filter(YEAR == 2022) |> 
-  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
-  group_by(AGEGR10) |> 
-  summarise(Active_population  = sum(WGHT_PER))
-
-# Combine total and active population 
-#and calculate the percentage
-Active_percentage <- Total_population |> 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
-  mutate(
-    Percentage = round(Active_population/Total_population*100,2))
-
-# Display the final result
-Active_percentage |>
-  kable()
-```
-
-## Processed data set
+## Active episodes
 
@@
 {ActiveCA} also enables obtaining insights from the main processed files.
-Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2015.
-According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 17% of possible cases.
+Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2022.
+According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 15% of possible.
 
 ```{r main_stress_figure, include = FALSE}
-main_stress_figure <- gss_main_2015 |>
+main_stress_figure <- gss_main_2022 |>
   filter(!is.na(MARSTAT), !is.na(GTU_110)) |>
@@
   ungroup() |>
-  ggplot(aes(x = MARSTAT,
-    y = GTU_110)) +
+  ggplot(aes(x = GTU_110,
+    y = MARSTAT)) +
     geom_tile(aes(fill = percentage)) +
@@
     fill = "Percentage (%)") + 
-    #geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
+    geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
     theme_bw() + 
@@
        plot = main_stress_figure,
-       width = 12, 
-       height = 8, 
+       width = 15, 
+       height = 10, 
        units = "cm", 

### 2025-05-30 paper/Manuscript-Data-Package-1Round-Review.tex
Changing the data paper due to the revision process.

@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% From pandoc table feature
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 
@@
 
-\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}
+\section{\{ActiveCA\} data sets}\label{activeca-data-sets}
 
@@
 package. In fact, we expect that the application of this package to
-extend beyond our pre-imagined range of uses.
+extend beyond our pre-imagined range of uses. The installation
+instruction and also some examples of application of the \{ActiveCA\} R
+package are available in the \texttt{vignettes}, available in the Github
+repository.
 
-To start the demonstration, it is first necessary to install the
-package. The code chunk below demonstrates how to install the
-\{ActiveCA\} package in R:
-
-\begin{verbatim}
-if(!require("ActiveCA", character.only = TRUE)) {
-      remotes::install_github("dias-bruno/ActiveCA")
-}
-\end{verbatim}
-
-After installing the \{ActiveCA\} R package, you can load it using the
-following code:
-
-\begin{verbatim}
-library(ActiveCA)
-\end{verbatim}
-
-The documentation of the package can be accessed by running
-\texttt{?ActiveCA} in \texttt{R}. The documentation includes a
-description of the package, information about the authors, contributors,
-and maintainers, and an overview of all pre-processed data sets included
-in the package.
-
-Figure \ref{fig:walking_documentation} displays the documentation for
-the walking episodes from the 2022 GSS (\texttt{walking\_2022.rda}). The
-documentation provides a brief description of the dataset, instructions
-on how to load the package, as well as details about the file format and
-descriptions of all variables (already labeled).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2022-documentation}
-
-}
-
-\caption{Documentation for the dataset of walking episodes from 2022.}\label{fig:walking_documentation}
-\end{figure}
-
-The code below shows how to load the walking episodes from the 2022 GSS
-(Cycle 29):
-
-\begin{verbatim}
-data(walking_2022)
-\end{verbatim}
-
-The GSS surveys apply a probability sampling methodology, in which each
-episode or person selected in the sample represents several other
-episodes or persons not in the sample. The number of episodes and
-persons represented by a episode or person is determined by the weight
-or weighting factor. Because of this, estimates of the number of
-episodes or persons need to be calculated applying the corresponding
-weighting factors.
-
-For instance, to calculate the percentage of respondents from the 2022
-GSS survey with active travel episodes, it is necessary to account for
-the person's weight. In 2022, the weight variable is represented by
-\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
-percentage of people with active travel episodes by age group. It uses
-the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.
-
-The process begins by creating a dataset that sums the population by age
-group. Then, it joins the 2022 episodes with the 2022 Main File. Note
-that in both operations, the code sums the person weight variable
-(\texttt{WGHT\_PER}) to obtain the correct population values. After
-this, a new dataset, \texttt{Active\_percentage}, is created by merging
-both previous datasets. The percentage of active travel episodes by age
-group is calculated by dividing the total population by the population
-with active trip episodes, then multiplying by 100 and rounding with 2
-decimal places.
-
-\begin{verbatim}
-# Load the dplyr library for data manipulation
-library('dplyr')
-
-# Load the episodes dataset
-data(gss_episodes)
-
-# Calculate the total population for each age group
-Total_population <- gss_main_2022 |> 
-  group_by(AGEGR10) |> 
-  summarise(Total_population  = sum(WGHT_PER))
-
-# Calculate the active population for each age group
-Active_population <- gss_episodes |> 
-      filter(YEAR == 2022) |> 
-  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
-  group_by(AGEGR10) |> 
-  summarise(Active_population  = sum(WGHT_PER))
-
-# Combine total and active population 
-#and calculate the percentage
-Active_percentage <- Total_population |> 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
-  mutate(
-    Percentage = round(Active_population/Total_population*100,2))
-\end{verbatim}
-
-The result shows that the age group with the highest share of active
-trips is those between 15 and 20 years old, with almost 44\%, followed
-by those between 25 and 34 years old with around 19\%. There is a
-significant drop in percentage for the following groups, with the
-percentage falling to between 12\% and 15\% for the remaining age
-groups.
-
-\begin{longtable}[]{@{}lrrr@{}}
-\toprule\noalign{}
-AGEGR10 & Total\_population & Active\_population & Percentage \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-15 to 24 years & 4537139 & 1984658.8 & 43.74 \\
-25 to 34 years & 5364748 & 1048653.8 & 19.55 \\
-35 to 44 years & 5227988 & 669688.8 & 12.81 \\
-45 to 54 years & 4720990 & 605162.1 & 12.82 \\
-55 to 64 years & 5143935 & 690336.5 & 13.42 \\
-65 to 74 years & 4160941 & 619732.8 & 14.89 \\
-75 years and over & 2981061 & 422799.5 & 14.18 \\
-\end{longtable}
-
-\subsection{Processed data set}\label{processed-data-set}
+\subsection{Active episodes}\label{active-episodes}
 

### 2025-05-30 paper/Manuscript-Data-Package-1Round-Review.Rmd
Adding process diagram to the paper.

@@
 
-# Data extraction
+# Data process
 
-For each selected cycle of the GSS surveys, we reviewed the Episode Files to identify episodes of movement that involved walking or cycling.
+Figure \ref{fig:process-figure} presents the processes applied to the Main and Episodes files for each survey year to obtain the ActiveCA datasets.
+For each selected TUS cycle, we reviewed the Episode Files to identify episodes of movement that involved walking or cycling.
 This allowed us to select the activities immediately before and after the movement episode.
-After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, mode of travel, time spent in the active trip, as well the province and urban classification of the episode.
-
 The active trips (walking and cycling) were identified by their corresponding activity codes, accounting for differences between the survey cycles.
 Identifying the preceding and succeeding activities of each activity episode helps to understand the motivation for the trip and also determine the origin and destination of the trips.
+After that, we labeled and factored the code variables with their appropriate descriptions, classifying each origin and destination, mode of travel, and time spent in the active trip.
 
@@
 We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others.
-After this, as we did with the Episode Files, we labeled the code variables with their appropriate descriptions.
+After this, as we did with the Episode Files, we labeled and factored the code variables with their appropriate descriptions.
+
+We joined the main files to episodes files to , as well the province and urban classification of the episode.
+
+```{r process-figure, echo=FALSE, out.width="100%", fig.cap="Processes applied to the main and episode files to obtain the ActiveCA datasets.", fig.align="center"}
+
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/RPackages - ActiveCA.jpeg"))
+```
 

### 2025-05-31 paper/Manuscript-Data-Package-1Round-Review.Rmd
Paper revision.

--- 2025-05-30 paper/Manuscript-Data-Package-1Round-Review.Rmd
@@
 
-# Original main and Episode Files from 2015 to display table examples
+# Original main and _Episode files_ from 2015 to display table examples
 gss_m_2015 <- read.csv(paste0(here(),
@@
 
-# Processed Main Files
+# Processed _Main files_
 data("gss_main_2022")
@@
 {ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
-a period of almost 40 years. The source of these data is TUS from the GSS.
-This program is designed to provide cross-sectional data on topics of interest to improve the well-being of Canadians. As part of this 
-program, every five to seven years the survey is done on the topic of time use. 
+a period of almost 40 years, obtained from the TUS cycles of the Canadian GSS.
+The GSS program is designed to provide cross-sectional data on topics of interest to improve the well-being of Canadians. As part of this 
+program, every five to seven years the survey is done on the topic of time use (TUS). 
 Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
-24 (2010), 29 (2015), and 34 (2022) of the GSS. Time use data in these surveys is coded using
-a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
+24 (2010), 29 (2015), and 34 (2022) of the TUS. Time use data in these surveys is coded using a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
 working or at school. These surveys have proved valuable in investigations of 
@@
 
-For {ActiveCA} and using TUS - GSS Public Use Microdata Files (PUMFs), we extracted all data needed to characterize active travel 
-in Canada, namely, episodes where the activity was identified as moving between 
-an origin and a destination, either by walking or cycling. Although Statistics 
-Canada offers Public Use Microdata Files and documentation for the GSS program 
-[see @statisticscanada2024], accessing these files, and preparing  them for analysis
-is not a straightforward matter, given their size and complexity.  The  process 
+Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS), we extracted all data necessary to characterize active travel in Canada - specifically, episodes in which the activity involved moving between an origin and a destination by walking or cycling.
+Although Statistics Canada provides PUMFs and accompanying documentation for the GSS program [see @statisticscanada2024], accessing and preparing these files for analysis is not a straightforward task due to their size and complexity.
+The  process
 of extracting information of interest from the source files is  time-consuming, 
 tedious, and challenging and/or prone to error due to the expertise required to 
-work with these files. To create {ActiveCA} we collected, cleaned, and processed
-the cycles of the GSS  surveys concerning time use to make them ready for analysis.
+work with these files. To create {ActiveCA} we selected, labelled, and processed
+the TUS cycles to make them ready for analysis.
 
@@
 the barriers to using data contained in rich, but difficult to access and use surveys,
-such as Canada's GSSs, is a worthy endeavour that can only improve data-driven 
+such as TUS, is a worthy endeavour that can only improve data-driven 
 decisions in transportation, urban, and health policy.
@@
 The topics of the survey cycle every few years to cover topics that include family,
-health, social identity, and every five to seven years time use (TUS). The first Canadian
+health, social identity, and every five to seven years time use. The first Canadian
 TUS done as part of the GSS program was conducted in 1986, and the 
@@
 TUS allows researchers to identify the origin and destination of trips, travel times and modes of transport used, providing a valuable dataset for analyzing active travel behavior.
-It also provides the empirical basis for tools used in transportation analysis, such as the development of impedance functions for accessibility analysis, a measure of the ease with which people can reach destinations and opportunities [@hansen1959; @paez2012].
+It also provides the empirical basis for tools used in transportation analysis, such as the development of impedance functions for accessibility analysis, a measure of the ease with which people can reach destinations and opportunities [@hansen1959].
 The Canadian TUS is unique at the national level in collecting detailed information on travel behaviour.
@@
 
-Most respondents to the 2022 TUS completed it online, reflecting Statistics Canada's effort to adapt to technological changes and growing time demands by offering greater flexibility and convenience [-@statisticscanada2022].
+Most respondents to the 2022 TUS completed it online, reflecting Statistics Canada's effort to adapt to technological changes and growing time demands by offering greater flexibility and convenience [@statisticscanada2022].
 While such methodological changes may affect data comparability over time, it is not possible to determine whether observed differences result from actual population changes or shifts in data collection methods.
@@
 Statistics Canada assessed the impact of collection mode on a limited set of key questions, constrained by sample size.
-Importantly, none of the variables used in this research are flagged in the 2022 PUMF User Guide as unsuitable for trend analysis.
+Importantly, none of the variables used in this research are `R` package in the 2022 PUMF User Guide as unsuitable for trend analysis.
 
@@
 The Public Use Microdata are released by Statistics Canada in two files: a 
-_Main File_ and an _Episode File_. The files are linked by keys that identify 
+_Main file_ and an _Episode file_. The files are linked by keys that identify 
 households, individuals, and episodes (i.e., activities) conducted by individuals.
@@
 
-## The Main File
+## The _Main file_
 
-The Main File of the time use survey compiles a large array of aggregated data, 
+The _Main file_ of the time use survey compiles a large array of aggregated data, 
 summarizing the answers to the questionnaire that describe households and 
@@
 file documents the time and duration that respondents allocate to each activity 
-and location. The Main File provides a overview of daily routines and social 
+and location. The _Main file_ provides a overview of daily routines and social 
 dynamics, not focusing on individual activity episodes. Additionally, this file 
@@
 
-Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29).
+Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the TUS PUMF 2015 _Main file_ (Cycle 29).
 Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
 
-The Main File of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
+The _Main file_ of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
 For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") |>
+    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.") |>
   kable_styling(full_width = FALSE, 
@@
 
-## The Episode File
+## The _Episode file_
 
@@
 location, and accompanying social context, informing when and where activities 
-occurred and with whom. The focus of the Episode File is not on the characteristics
+occurred and with whom. The focus of the _Episode file_ is not on the characteristics
 of the respondents, but on the characteristics of the activities, and the data 
@@
 respondent. Although respondent-specific characteristics are not included within
-the Episode File, it is possible to link the Main File and the Episode File by 
-using a key present in both the Main and Episode Files.
+the _Episode file_, it is possible to link the _Main file_ and the _Episode file_ by 
+using a key present in both the _Main_ and _Episode files_.
 
-Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the _Main file_ structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the TUS PUMF 2015 _Episode file_ (Cycle 29).
 Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
 
-In total, the Episode File of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
-Similar to the Main File, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
+In total, the _Episode file_ of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
+Similar to the _Main file_, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
 
@@
 
-# Data extraction
+# Data process
 
-For each selected cycle of the GSS surveys, we reviewed the Episode Files to identify episodes of movement that involved walking or cycling.
-This allowed us to select the activities immediately before and after the movement episode.
-After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, mode of travel, time spent in the active trip, as well the province and urban classification of the episode.
+Figure \ref{fig:process-figure} presents a diagram illustrating the processes applied to the _Main_ and _Episode files_ to create the {ActiveCA} datasets.
+For each TUS cycle, we reviewed the _Episode files_ to identify movement episodes involving walking or cycling.
+This allowed us to also select the activities immediately before and after the movement episode, helping to infer the trip’s purpose and determine its origin and destination.
+Active trips were identified by their corresponding activity codes, accounting for variations across survey cycles.
 
-The active trips (walking and cycling) were identified by their corresponding activity codes, accounting for differences between the survey cycles.
-Identifying the preceding and succeeding activities of each activity episode helps to understand the motivation for the trip and also determine the origin and destination of the trips.
+Next, we labeled the coded variables with their appropriate descriptions, classifying each origin and destination, mode of travel, and time spent in the active trip.
+Categorical variables were converted into factor variables, and ordinal variables were defined as ordered factors.
 
-For each GSS Main File, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes.
-We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others.
-After this, as we did with the Episode Files, we labeled the code variables with their appropriate descriptions.
+For the _Main files_, we selected socioeconomic variables to help profile individuals engaged in active travel episodes.
+These included key indicators such as age group, sex, marital status, and number of children, among others.
+As with the _Episode files_, we labeled and factored the coded variables.
+Using the appropriate identifiers, we joined the Episode and _Main files_ to determine the province and the type of population centre - whether it is a CMA, a Census Agglomeration (CA), or a non-CMA/CA area.
+
+As output, the {ActiveCA} package provides processed datasets of walking and cycling episodes by year, a unified dataset that contains all walking and cycling episodes across TUS cycles, and processed _Main files_ (not unified, as socioeconomic and demographic variables and their categories differ across TUS cycles).
+
+```{r process-figure, echo=FALSE, out.width="100%", fig.cap="Diagram with the processes applied to the main (blue arrows) and episode files (green arrows) to obtain the ActiveCA datasets.", fig.align="center"}
+
+knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/RPackages - ActiveCA.jpeg"))
+```
 
@@
 
-This section presents some potential applications of the {ActiveCA} R package.
+This section presents some potential applications of the {ActiveCA} `R` package.
 In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
-The installation instruction and also some examples of application of the {ActiveCA} R package are available in the `vignettes`, available in the Github repository.
+The installation instruction and also some examples of application of the {ActiveCA} `R` package are available in the `vignettes`, available in the Github repository.
 
@@
 
-Table \ref{tab:processed-obs} displays the total number of records processed for Main and Episode Files.
-For the Main Files, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2022, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2022',]$Weighted)` for 2022, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986).
+Table \ref{tab:processed-obs} displays the total number of records processed for Main and _Episode files_.
+For the _Main files_, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2022, that together represents more of `r sum(df_main_processed$Weighted)` respondents.
 It also presents the total cases of active trips episodes identified.
@@
 
-Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
-Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
+Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the TUS PUMF 2015 _Episode file_ (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
 Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes.
@@
 
-{ActiveCA} also enables visual analysis of active travel in Canada using traditional exploratory data analysis techniques.
+{ActiveCA} also enables visual analysis of active travel in Canada using exploratory data analysis techniques.
 Figure \ref{fig:figure-01} shows walking trips from 2022 through heat maps.
@@
 
-The {ActiveCA} dataset also includes information on the type of population centre in which respondents reside - specifically, whether they live in a CMA, a Census Agglomeration (CA), or outside these areas - as well as the respondent’s province.
+The {ActiveCA} dataset also includes information on the type of population centre in which respondents reside - specifically, whether they live in a CMA, a CA, or outside these areas - as well as the respondent’s province.
 This information is important, as patterns of active travel often differ between metropolitan and non-metropolitan populations.
@@
 
-{ActiveCA} also enables obtaining insights from the main processed files.
+The package also enables obtaining insights from the main processed files.
 Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2022.
-According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 15% of possible.
+According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 15% of possible cases.
 
@@
 Researchers interested in non-AT modes can use our framework to guide their methodology, making the small but necessary adjustments.
-We focused exclusively on AT episodes because the ActiveCA package is part of a larger project aimed at analyzing the historical evolution of active travel behaviour in Canada.
+We focused exclusively on AT episodes because the {ActiveCA} package is part of a larger project aimed at analyzing the historical evolution of active travel behaviour in Canada.
 
@@
 which facilitates the addition of complementary data sets in the future. `R` users
-can seamlessly explore GSS walking and cycling episodes, 
+can seamlessly explore TUS walking and cycling episodes, 
 with the option to suggest enhancements to the package as needed. This article adopts
@@
 replication and innovation in related fields. We believe that {ActiveCA} will serve
-as a basis for further research on TUS GSS and for the integration of additional data by
+as a basis for further research on TUS and for the integration of additional data by
 the authors or the wider open source community.

### 2025-05-31 paper/Manuscript-Data-Package-1Round-Review.tex
Paper revision.

@@
 \{ActiveCA\} provides analysis-ready data concerning active travel in
-Canada spanning a period of almost 40 years. The source of these data is
-TUS from the GSS. This program is designed to provide cross-sectional
-data on topics of interest to improve the well-being of Canadians. As
-part of this program, every five to seven years the survey is done on
-the topic of time use. Concretely, \{ActiveCA\} covers Cycles 2 (1986),
-7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of
-the GSS. Time use data in these surveys is coded using a very fine
-grain, from time spent in chores, leisure, and sleeping, to time spent
-working or at school. These surveys have proved valuable in
+Canada spanning a period of almost 40 years, obtained from the TUS
+cycles of the Canadian GSS. The GSS program is designed to provide
+cross-sectional data on topics of interest to improve the well-being of
+Canadians. As part of this program, every five to seven years the survey
+is done on the topic of time use (TUS). Concretely, \{ActiveCA\} covers
+Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
+and 34 (2022) of the TUS. Time use data in these surveys is coded using
+a very fine grain, from time spent in chores, leisure, and sleeping, to
+time spent working or at school. These surveys have proved valuable in
 investigations of mobility and quality of life
@@
 
-For \{ActiveCA\} and using TUS - GSS Public Use Microdata Files (PUMFs),
-we extracted all data needed to characterize active travel in Canada,
-namely, episodes where the activity was identified as moving between an
-origin and a destination, either by walking or cycling. Although
-Statistics Canada offers Public Use Microdata Files and documentation
-for the GSS program \citep[see][]{statisticscanada2024}, accessing these
-files, and preparing them for analysis is not a straightforward matter,
-given their size and complexity. The process of extracting information
-of interest from the source files is time-consuming, tedious, and
-challenging and/or prone to error due to the expertise required to work
-with these files. To create \{ActiveCA\} we collected, cleaned, and
-processed the cycles of the GSS surveys concerning time use to make them
-ready for analysis.
+Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
+we extracted all data necessary to characterize active travel in Canada
+- specifically, episodes in which the activity involved moving between
+an origin and a destination by walking or cycling. Although Statistics
+Canada provides PUMFs and accompanying documentation for the GSS program
+\citep[see][]{statisticscanada2024}, accessing and preparing these files
+for analysis is not a straightforward task due to their size and
+complexity. The process of extracting information of interest from the
+source files is time-consuming, tedious, and challenging and/or prone to
+error due to the expertise required to work with these files. To create
+\{ActiveCA\} we selected, labelled, and processed the TUS cycles to make
+them ready for analysis.
 
@@
 \citep[e.g.,][]{mccurdySupport2023}, reducing the barriers to using data
-contained in rich, but difficult to access and use surveys, such as
-Canada's GSSs, is a worthy endeavour that can only improve data-driven
-decisions in transportation, urban, and health policy.
+contained in rich, but difficult to access and use surveys, such as TUS,
+is a worthy endeavour that can only improve data-driven decisions in
+transportation, urban, and health policy.
 
@@
 include family, health, social identity, and every five to seven years
-time use (TUS). The first Canadian TUS done as part of the GSS program
-was conducted in 1986, and the most recent was completed in 2022. These
-Time Use Surveys \citep{statisticscanada2022} collect data on
-respondents' participation and time spent on a wide range of everyday
-activities using a 24-hour retrospective diary, with information on the
-location of these activities (e.g.~at home, at work, etc.) and, for
-non-personal activities, the people who were present with the respondent
-at the time of the activity. In addition, time-use surveys also cover
-topics related to leisure time, work-life balance, health, commuting,
-culture and sports, and many others.
+time use. The first Canadian TUS done as part of the GSS program was
+conducted in 1986, and the most recent was completed in 2022. These Time
+Use Surveys \citep{statisticscanada2022} collect data on respondents'
+participation and time spent on a wide range of everyday activities
+using a 24-hour retrospective diary, with information on the location of
+these activities (e.g.~at home, at work, etc.) and, for non-personal
+activities, the people who were present with the respondent at the time
+of the activity. In addition, time-use surveys also cover topics related
+to leisure time, work-life balance, health, commuting, culture and
+sports, and many others.
 
@@
 with which people can reach destinations and opportunities
-\citep{hansen1959, paez2012}. The Canadian TUS is unique at the national
-level in collecting detailed information on travel behaviour. Its
-consistent application across survey cycles enables the identification
-of long-term trends, with some questions present in the questionnaires
-since the first application of the survey.
+\citep{hansen1959}. The Canadian TUS is unique at the national level in
+collecting detailed information on travel behaviour. Its consistent
+application across survey cycles enables the identification of long-term
+trends, with some questions present in the questionnaires since the
+first application of the survey.
 
@@
 time demands by offering greater flexibility and convenience
-\citeyearpar{statisticscanada2022}. While such methodological changes
-may affect data comparability over time, it is not possible to determine
+\citep{statisticscanada2022}. While such methodological changes may
+affect data comparability over time, it is not possible to determine
 whether observed differences result from actual population changes or
@@
 limited set of key questions, constrained by sample size. Importantly,
-none of the variables used in this research are flagged in the 2022 PUMF
-User Guide as unsuitable for trend analysis.
+none of the variables used in this research are \texttt{R} package in
+the 2022 PUMF User Guide as unsuitable for trend analysis.
 
@@
 The Public Use Microdata are released by Statistics Canada in two files:
-a \emph{Main File} and an \emph{Episode File}. The files are linked by
+a \emph{Main file} and an \emph{Episode file}. The files are linked by
 keys that identify households, individuals, and episodes (i.e.,
@@
 
-\subsection{The Main File}\label{the-main-file}
+\subsection{\texorpdfstring{The \emph{Main
+file}}{The Main file}}\label{the-main-file}
 
-The Main File of the time use survey compiles a large array of
+The \emph{Main file} of the time use survey compiles a large array of
 aggregated data, summarizing the answers to the questionnaire that
@@
 duration that respondents allocate to each activity and location. The
-Main File provides a overview of daily routines and social dynamics, not
-focusing on individual activity episodes. Additionally, this file
-categorizes activities into bigger groups and subcategories,
+\emph{Main file} provides a overview of daily routines and social
+dynamics, not focusing on individual activity episodes. Additionally,
+this file categorizes activities into bigger groups and subcategories,
 facilitating the data's analytical utility with additional metrics such
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
-six variables of the GSS PUMF 2015 Main File (Cycle 29). Each row in the
-table correspond to a survey respondent, while the columns refer the
-following information: record identification (\texttt{PUMFID}), the
+six variables of the TUS PUMF 2015 \emph{Main file} (Cycle 29). Each row
+in the table correspond to a survey respondent, while the columns refer
+the following information: record identification (\texttt{PUMFID}), the
 person's weight (\texttt{WGHT\_PER}), the month the survey data was
@@
 
-The Main File of the 2015 GSS surveys includes a total of 17,390
+The \emph{Main file} of the 2015 GSS surveys includes a total of 17,390
 respondents, representing 29,766,399 individuals and 848 variables. For
@@
 \begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.}\\
+\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.}\\
 \toprule
@@
 
-\subsection{The Episode File}\label{the-episode-file}
+\subsection{\texorpdfstring{The \emph{Episode
+file}}{The Episode file}}\label{the-episode-file}
 
@@
 social context, informing when and where activities occurred and with
-whom. The focus of the Episode File is not on the characteristics of the
-respondents, but on the characteristics of the activities, and the data
-are structured around the numerous activity instances that compose a day
-of the respondent. Although respondent-specific characteristics are not
-included within the Episode File, it is possible to link the Main File
-and the Episode File by using a key present in both the Main and Episode
-Files.
+whom. The focus of the \emph{Episode file} is not on the characteristics
+of the respondents, but on the characteristics of the activities, and
+the data are structured around the numerous activity instances that
+compose a day of the respondent. Although respondent-specific
+characteristics are not included within the \emph{Episode file}, it is
+possible to link the \emph{Main file} and the \emph{Episode file} by
+using a key present in both the \emph{Main} and \emph{Episode files}.
 
 Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
-example of the Main File structure, Table \ref{tab:ep-2015-unprocessed}
-presents the first seven episodes for the record identification number
-\texttt{10041} and some variables from the GSS PUMF 2015 Episode File
-(Cycle 29). Each row in the table corresponds to an episode associated
-with the specified record identification (\texttt{PUMFID\ =\ 10041}),
-including the episode's weight (\texttt{WGHT\_EPI}), episode number
-(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
-(\texttt{DURATION}), and episode location (\texttt{LOCATION}).
+example of the \emph{Main file} structure, Table
+\ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
+record identification number \texttt{10041} and some variables from the
+TUS PUMF 2015 \emph{Episode file} (Cycle 29). Each row in the table
+corresponds to an episode associated with the specified record
+identification (\texttt{PUMFID\ =\ 10041}), including the episode's
+weight (\texttt{WGHT\_EPI}), episode number (\texttt{EPINO}), activity
+code (\texttt{TUI\_01}), episode duration (\texttt{DURATION}), and
+episode location (\texttt{LOCATION}).
 
-In total, the Episode File of the 2015 GSS surveys contains 274,108
-records, representing 461,837,622 episodes and 527 variables detailing
-the episodes. Similar to the Main File, Statistics Canada has created
-codes for the discrete variables, with each value corresponding to a
-label.
+In total, the \emph{Episode file} of the 2015 GSS surveys contains
+274,108 records, representing 461,837,622 episodes and 527 variables
+detailing the episodes. Similar to the \emph{Main file}, Statistics
+Canada has created codes for the discrete variables, with each value
+corresponding to a label.
 
@@
 
-\section{Data extraction}\label{data-extraction}
+\section{Data process}\label{data-process}
 
-For each selected cycle of the GSS surveys, we reviewed the Episode
-Files to identify episodes of movement that involved walking or cycling.
-This allowed us to select the activities immediately before and after
-the movement episode. After that, we labeled the code variables with
-their appropriate descriptions, identifying each origin and destination,
-mode of travel, time spent in the active trip, as well the province and
-urban classification of the episode.
+Figure \ref{fig:process-figure} presents a diagram illustrating the
+processes applied to the \emph{Main} and \emph{Episode files} to create
+the \{ActiveCA\} datasets. For each TUS cycle, we reviewed the
+\emph{Episode files} to identify movement episodes involving walking or
+cycling. This allowed us to also select the activities immediately
+before and after the movement episode, helping to infer the trip's
+purpose and determine its origin and destination. Active trips were
+identified by their corresponding activity codes, accounting for
+variations across survey cycles.
 
-The active trips (walking and cycling) were identified by their
-corresponding activity codes, accounting for differences between the
-survey cycles. Identifying the preceding and succeeding activities of
-each activity episode helps to understand the motivation for the trip
-and also determine the origin and destination of the trips.
+Next, we labeled the coded variables with their appropriate
+descriptions, classifying each origin and destination, mode of travel,
+and time spent in the active trip. Categorical variables were converted
+into factor variables, and ordinal variables were defined as ordered
+factors.
 
-For each GSS Main File, we selected socioeconomic variables to help
-profile the individuals engaged in active trip episodes. We included key
-socioeconomic variables such as the respondent's age group, sex, marital
-status, and number of children, among others. After this, as we did with
-the Episode Files, we labeled the code variables with their appropriate
-descriptions.
+For the \emph{Main files}, we selected socioeconomic variables to help
+profile individuals engaged in active travel episodes. These included
+key indicators such as age group, sex, marital status, and number of
+children, among others. As with the \emph{Episode files}, we labeled and
+factored the coded variables. Using the appropriate identifiers, we
+joined the Episode and \emph{Main files} to determine the province and
+the type of population centre - whether it is a CMA, a Census
+Agglomeration (CA), or a non-CMA/CA area.
+
+As output, the \{ActiveCA\} package provides processed datasets of
+walking and cycling episodes by year, a unified dataset that contains
+all walking and cycling episodes across TUS cycles, and processed
+\emph{Main files} (not unified, as socioeconomic and demographic
+variables and their categories differ across TUS cycles).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/RPackages - ActiveCA}
+
+}
+
+\caption{Diagram with the processes applied to the main (blue arrows) and episode files (green arrows) to obtain the ActiveCA datasets.}\label{fig:process-figure}
+\end{figure}
 
@@
 
-This section presents some potential applications of the \{ActiveCA\} R
-package. In fact, we expect that the application of this package to
-extend beyond our pre-imagined range of uses. The installation
-instruction and also some examples of application of the \{ActiveCA\} R
-package are available in the \texttt{vignettes}, available in the Github
-repository.
+This section presents some potential applications of the \{ActiveCA\}
+\texttt{R} package. In fact, we expect that the application of this
+package to extend beyond our pre-imagined range of uses. The
+installation instruction and also some examples of application of the
+\{ActiveCA\} \texttt{R} package are available in the \texttt{vignettes},
+available in the Github repository.
 
@@
 Table \ref{tab:processed-obs} displays the total number of records
-processed for Main and Episode Files. For the Main Files, a total of
-101,667 records were processed, referring to all records from the Time
-Use Surveys from 1986 to 2022, that together represents more of
-181,526,641 respondents (32,136,802 for 2022, 29,766,399 for 2015,
-28,075,610 for 2010, 26,095,819 for 2005, 24,260,137 for 1998,
-21,294,313 for 1992, and for 2005, 24,260,137 for 1998, and 21,294,313
-for 1986). It also presents the total cases of active trips episodes
-identified. In total 23,513 records with register of active travel
-activity. Together, these records account for 44,316,110 episodes.
+processed for Main and \emph{Episode files}. For the \emph{Main files},
+a total of 101,667 records were processed, referring to all records from
+the Time Use Surveys from 1986 to 2022, that together represents more of
+181,526,641 respondents. It also presents the total cases of active
+trips episodes identified. In total 23,513 records with register of
+active travel activity. Together, these records account for 44,316,110
+episodes.
 
@@
 Table \ref{tab:main-2015-processed} shows the first ten rows and first
-six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in
+six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
 \ref{tab:main-2015-unprocessed} before our processing. Table
 \ref{tab:ep-2015-processed} presents the walking episodes for the record
-identification number \texttt{10041} from the GSS PUMF 2015 Episode File
-(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
-Only the unique active travel episode appear in Table
-\ref{tab:ep-2015-processed} since the records were filtered to select
-cases with walking or cycling episodes. For both cases, Tables
-\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
-labeled variables, facilitating the interpretation of the data.
+identification number \texttt{10041} from the TUS PUMF 2015
+\emph{Episode file} (Cycle 29), previously displayed in Table
+\ref{tab:ep-2015-unprocessed}. Only the unique active travel episode
+appear in Table \ref{tab:ep-2015-processed} since the records were
+filtered to select cases with walking or cycling episodes. For both
+cases, Tables \ref{tab:main-2015-processed} and
+\ref{tab:ep-2015-processed} contain labeled variables, facilitating the
+interpretation of the data.
 
@@
 \{ActiveCA\} also enables visual analysis of active travel in Canada
-using traditional exploratory data analysis techniques. Figure
-\ref{fig:figure-01} shows walking trips from 2022 through heat maps.
-This graph uses color gradients to represent the percentage of trips
-between various origins and destinations, with darker colors indicating
-higher percentages and lighter colors representing less frequent routes.
-For conciseness, we omitted the heat maps for the other years analyzed.
+using exploratory data analysis techniques. Figure \ref{fig:figure-01}
+shows walking trips from 2022 through heat maps. This graph uses color
+gradients to represent the percentage of trips between various origins
+and destinations, with darker colors indicating higher percentages and
+lighter colors representing less frequent routes. For conciseness, we
+omitted the heat maps for the other years analyzed.
 
@@
 population centre in which respondents reside - specifically, whether
-they live in a CMA, a Census Agglomeration (CA), or outside these areas
-- as well as the respondent's province. This information is important,
-as patterns of active travel often differ between metropolitan and
-non-metropolitan populations. For example, Table \ref{tab:cma-durations}
-presents the median walking durations by population centre type and
-province for 2022. Overall, respondents living in CMA/CA areas tend to
-report higher median walking durations compared to those living outside
-these centres. The most pronounced difference is observed in Nova
-Scotia: metropolitan residents reported a median walking duration of 30
-minutes, whereas non-metropolitan residents reported a median of only 5
-minutes.
+they live in a CMA, a CA, or outside these areas - as well as the
+respondent's province. This information is important, as patterns of
+active travel often differ between metropolitan and non-metropolitan
+populations. For example, Table \ref{tab:cma-durations} presents the
+median walking durations by population centre type and province for
+2022. Overall, respondents living in CMA/CA areas tend to report higher
+median walking durations compared to those living outside these centres.
+The most pronounced difference is observed in Nova Scotia: metropolitan
+residents reported a median walking duration of 30 minutes, whereas
+non-metropolitan residents reported a median of only 5 minutes.
 
@@
 
-\{ActiveCA\} also enables obtaining insights from the main processed
+The package also enables obtaining insights from the main processed
 files. Figure \ref{fig:figure-stress} present how the level of stress
-varied among respondents depending on their marital status in 2015.
+varied among respondents depending on their marital status in 2022.
 According to this plot, married respondents reported the highest level
-of stress, relating to feel stressed every day, with 17\% of possible
+of stress, relating to feel stressed every day, with 15\% of possible
 cases.
@@
 methodology, making the small but necessary adjustments. We focused
-exclusively on AT episodes because the ActiveCA package is part of a
+exclusively on AT episodes because the \{ActiveCA\} package is part of a
 larger project aimed at analyzing the historical evolution of active
@@
 ease of use, which facilitates the addition of complementary data sets
-in the future. \texttt{R} users can seamlessly explore GSS walking and
+in the future. \texttt{R} users can seamlessly explore TUS walking and
 cycling episodes, with the option to suggest enhancements to the package
@@
 encourage replication and innovation in related fields. We believe that
-\{ActiveCA\} will serve as a basis for further research on TUS GSS and
-for the integration of additional data by the authors or the wider open
+\{ActiveCA\} will serve as a basis for further research on TUS and for
+the integration of additional data by the authors or the wider open
 source community.

### 2025-06-01 paper/Manuscript-Data-Package-1Round-Review.Rmd
Create a document with all manuscript changes and add answer letters for the reviewers.

@@
     affiliation: CODE
-    footnote: 1
   - name: "Anon2"
@@
   "This paper describes {ActiveCA}, an open data product with Canadian time use data.
-{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS).
+{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS).
 Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
@@
 
-The objective of this paper is to introduce {ActiveCA}, an open data product with data from all Time Use Survey (TUS) cycles of the General Social Surveys (GSSs). Open data products
-(ODPs) are the outcome of a process that transforms raw data (open or not) into 
-analysis-ready data, following a transparent process in which all stages of
-development follow open principles [@arribas-bel2021]. ODPs, while still open, 
-differ from general open data in their degree of ease of access, their heightened 
-usability, and potentially the value they add to the raw data.
+The objective of this paper is to introduce {ActiveCA}, an open data
+product with data from all Time Use Survey (TUS) cycles of the General
+Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
+process that transforms raw data (open or not) into analysis-ready data,
+following a transparent process in which all stages of development
+follow open principles [@arribas-bel2021]. ODPs, while still open,
+differ from general open data in their degree of ease of access, their
+heightened usability, and potentially the value they add to the raw
+data.
 
-{ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
-a period of almost 40 years, obtained from the TUS cycles of the Canadian GSS.
-The GSS program is designed to provide cross-sectional data on topics of interest to improve the well-being of Canadians. As part of this 
-program, every five to seven years the survey is done on the topic of time use (TUS). 
-Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
-24 (2010), 29 (2015), and 34 (2022) of the TUS. Time use data in these surveys is coded using a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
-working or at school. These surveys have proved valuable in investigations of 
-mobility and quality of life [@spinneyTransport2009], the relationship between
-active travel and transit use [@lachapelleLonger2016], and travel behavior and time
-poverty [@kimFacing2024], to name but a few examples.
+{ActiveCA} provides analysis-ready data concerning active travel in
+Canada spanning a period of almost 40 years, obtained from the TUS
+cycles of the Canadian GSS. The GSS program is designed to provide
+cross-sectional data on topics of interest to improve the well-being of
+Canadians. As part of this program, every five to seven years the survey
+is done on the topic of time use (TUS). Concretely, {ActiveCA} covers
+Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
+and 34 (2022) of the TUS. Time use data in these surveys is coded using
+a very fine grain, from time spent in chores, leisure, and sleeping, to
+time spent working or at school. These surveys have proved valuable in
+investigations of mobility and quality of life [@spinneyTransport2009],
+the relationship between active travel and transit use
+[@lachapelleLonger2016], and travel behavior and time poverty
+[@kimFacing2024], to name but a few examples.
 
-Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS), we extracted all data necessary to characterize active travel in Canada - specifically, episodes in which the activity involved moving between an origin and a destination by walking or cycling.
-Although Statistics Canada provides PUMFs and accompanying documentation for the GSS program [see @statisticscanada2024], accessing and preparing these files for analysis is not a straightforward task due to their size and complexity.
-The  process
-of extracting information of interest from the source files is  time-consuming, 
-tedious, and challenging and/or prone to error due to the expertise required to 
-work with these files. To create {ActiveCA} we selected, labelled, and processed
-the TUS cycles to make them ready for analysis.
+Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
+we extracted all data necessary to characterize active travel in
+Canada - specifically, episodes in which the activity involved moving
+between an origin and a destination by walking or cycling. Although
+Statistics Canada provides PUMFs and accompanying documentation for the
+GSS program [see @statisticscanada2024], accessing and preparing these
+files for analysis is not a straightforward task due to their size and
+complexity. The process of extracting information of interest from the
+source files is time-consuming, tedious, and challenging and/or prone to
+error due to the expertise required to work with these files. To create
+{ActiveCA} we selected, labelled, and processed the TUS cycles to make
+them ready for analysis.
 
-{ActiveCA} is distributed as an `R` package with a number of data 
-objects and their documentation. `R` packages contain code, data, and documentation
-in a standardized format that can be installed by `R` users via a software repository,
-such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
-adroit medium to distribute analysis-ready data.
+{ActiveCA} is distributed as an `R` package with a number of data objects
+and their documentation. `R` packages contain code, data, and
+documentation in a standardized format that can be installed by R users
+via a software repository, such as CRAN (Comprehensive R Archive
+Network) or GitHub, which makes them an adroit medium to distribute
+analysis-ready data.
 
-Given the level of interest in active travel [e.g., @mccurdySupport2023], reducing
-the barriers to using data contained in rich, but difficult to access and use surveys,
-such as TUS, is a worthy endeavour that can only improve data-driven 
-decisions in transportation, urban, and health policy.
+Given the level of interest in active travel [e.g.,
+@mccurdySupport2023], reducing the barriers to using data contained in
+rich, but difficult to access and use surveys, such as TUS, is a worthy
+endeavour that can only improve data-driven decisions in transportation,
+urban, and health policy.
 
-The rest of this paper discusses the sources of data, and the process implemented
-to retrieve and package them. Then, we explain how to use the package and show
-some selected examples of analysis to whet the imagination of potential users.
-This ODP provides not only data that are easy to use, but also all the code and 
-documentation that make this a reproducible research project. In summary, {ActiveCA}
-aims to implement and inspire the best principles of open spatial sciences 
+The rest of this paper discusses the sources of data, and the process
+implemented to retrieve and package them. Then, we explain how to use
+the package and show some selected examples of analysis to whet the
+imagination of potential users. This ODP provides not only data that are
+easy to use, but also all the code and documentation that make this a
+reproducible research project. In summary, {ActiveCA} aims to implement
+and inspire the best principles of open spatial sciences
 [@paez_open_2021; @brunsdon_opening_2021].
@@
 
-Statistics Canada [-@statisticscanada2024] conducts GSS surveys to obtain data 
-on social trends to track changes in Canadians' living conditions and well-being
-over time. The series of survey on time use are used to understand how Canadian
-residents spend and manage their time, and what factors contribute to their 
-happiness and stress. The GSS program was created in 1985, and is serialized to 
-provide a collection of annual, representative cross-sectional surveys.
+Statistics Canada [-@statisticscanada2024] conducts GSS surveys to
+obtain data on social trends to track changes in Canadians' living
+conditions and well-being over time. The series of survey on time use
+are used to understand how Canadian residents spend and manage their
+time, and what factors contribute to their happiness and stress. The GSS
+program was created in 1985, and is serialized to provide a collection
+of annual, representative cross-sectional surveys.
 
-The topics of the survey cycle every few years to cover topics that include family,
-health, social identity, and every five to seven years time use. The first Canadian
-TUS done as part of the GSS program was conducted in 1986, and the 
-most recent was completed in 2022. These Time Use Surveys [@statisticscanada2022]
-collect data on respondents' participation and time spent on a wide range of 
-everyday activities using a 24-hour retrospective diary, with information on the 
-location of these activities (e.g. at home, at work, etc.) and, for non-personal
-activities, the people who were present with the respondent at the time of the 
-activity. In addition, time-use surveys also cover topics related to leisure time,
-work-life balance, health, commuting, culture and sports, and many others.
+The topics of the survey cycle every few years to cover topics that
+include family, health, social identity, and every five to seven years
+time use. The first Canadian TUS done as part of the GSS program was
+conducted in 1986, and the most recent was completed in 2022. These Time
+Use Surveys [@statisticscanada2022] collect data on respondents'
+participation and time spent on a wide range of everyday activities
+using a 24-hour retrospective diary, with information on the location of
+these activities (e.g. at home, at work, etc.) and, for non-personal
+activities, the people who were present with the respondent at the time
+of the activity. In addition, time-use surveys also cover topics related
+to leisure time, work-life balance, health, commuting, culture and
+sports, and many others.
 
-TUS allows researchers to identify the origin and destination of trips, travel times and modes of transport used, providing a valuable dataset for analyzing active travel behavior.
-It also provides the empirical basis for tools used in transportation analysis, such as the development of impedance functions for accessibility analysis, a measure of the ease with which people can reach destinations and opportunities [@hansen1959].
-The Canadian TUS is unique at the national level in collecting detailed information on travel behaviour.
-Its consistent application across survey cycles enables the identification of long-term trends, with some questions present in the questionnaires since the first application of the survey.
+TUS allows researchers to identify the origin and destination of trips,
+travel times and modes of transport used, providing a valuable dataset
+for analyzing active travel behavior. It also provides the empirical
+basis for tools used in transportation analysis, such as the development
+of impedance functions for accessibility analysis, a measure of the ease
+with which people can reach destinations and opportunities
+[@hansen1959]. The Canadian TUS is unique at the national level in
+collecting detailed information on travel behaviour. Its consistent
+application across survey cycles enables the identification of long-term
+trends, with some questions present in the questionnaires since the
+first application of the survey.
 
-Most respondents to the 2022 TUS completed it online, reflecting Statistics Canada's effort to adapt to technological changes and growing time demands by offering greater flexibility and convenience [@statisticscanada2022].
-While such methodological changes may affect data comparability over time, it is not possible to determine whether observed differences result from actual population changes or shifts in data collection methods.
-Despite rigorous efforts to ensure data quality, the use of electronic questionnaires may have influenced estimates.
-Statistics Canada assessed the impact of collection mode on a limited set of key questions, constrained by sample size.
-Importantly, none of the variables used in this research are `R` package in the 2022 PUMF User Guide as unsuitable for trend analysis.
+Most respondents to the 2022 TUS completed it online, reflecting
+Statistics Canada's effort to adapt to technological changes and growing
+time demands by offering greater flexibility and convenience
+[@statisticscanada2022]. While such methodological changes may affect
+data comparability over time, it is not possible to determine whether
+observed differences result from actual population changes or shifts in
+data collection methods. Despite rigorous efforts to ensure data
+quality, the use of electronic questionnaires may have influenced
+estimates. Statistics Canada assessed the impact of collection mode on a
+limited set of key questions, constrained by sample size. Importantly,
+none of the variables used in this research are `R` package in the 2022
+PUMF User Guide as unsuitable for trend analysis.
 
-Until 2022, Statistics Canada employed a telephone-based sampling frame, which was replaced by a dwelling-based frame in the most recent cycle.
-Each survey cycle spans a 12-month period, typically from July to the following July.
-The target population includes all Canadians aged 15 and over, excluding residents of the Yukon, Northwest Territories, and Nunavut, full-time institutional residents, and individuals living on Indigenous reserves.
+Until 2022, Statistics Canada employed a telephone-based sampling frame,
+which was replaced by a dwelling-based frame in the most recent cycle.
+Each survey cycle spans a 12-month period, typically from July to the
+following July. The target population includes all Canadians aged 15 and
+over, excluding residents of the Yukon, Northwest Territories, and
+Nunavut, full-time institutional residents, and individuals living on
+Indigenous reserves.
 
-The survey encompasses both rural and urban areas, including metropolitan and non-metropolitan regions, to ensure a diverse and representative sample.
-For sampling, the ten provinces were divided into geographic strata.
-Several Census Metropolitan Areas (CMAs) - such as St.
-John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver - were treated as separate strata.
-Additional strata grouped other CMAs within Quebec, Ontario, and British Columbia, as well as non-CMA areas within each province.
+The survey encompasses both rural and urban areas, including
+metropolitan and non-metropolitan regions, to ensure a diverse and
+representative sample. For sampling, the ten provinces were divided into
+geographic strata. Several Census Metropolitan Areas (CMAs) - such as
+St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
+Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and
+Vancouver - were treated as separate strata. Additional strata grouped
+other CMAs within Quebec, Ontario, and British Columbia, as well as
+non-CMA areas within each province.
 
-The Public Use Microdata are released by Statistics Canada in two files: a 
-_Main file_ and an _Episode file_. The files are linked by keys that identify 
-households, individuals, and episodes (i.e., activities) conducted by individuals.
-We discuss these files in more detail in the following section.
+The Public Use Microdata are released by Statistics Canada in two files:
+a *Main file* and an *Episode file*. The files are linked by keys that
+identify households, individuals, and episodes (i.e., activities)
+conducted by individuals. We discuss these files in more detail in the
+following section.
 
-## The _Main file_
+## The Main file
 
-The _Main file_ of the time use survey compiles a large array of aggregated data, 
-summarizing the answers to the questionnaire that describe households and 
-individuals, as well as derived variables that summarize the respondents' use of 
-time use across different activities, locations, and social interactions. This 
-file documents the time and duration that respondents allocate to each activity 
-and location. The _Main file_ provides a overview of daily routines and social 
-dynamics, not focusing on individual activity episodes. Additionally, this file 
-categorizes activities into bigger groups and subcategories, facilitating the 
-data's analytical utility with additional metrics such as total transit time, time 
-spent with household members, and counts of activities and episodes.
+The *Main file* of the time use survey compiles a large array of
+aggregated data, summarizing the answers to the questionnaire that
+describe households and individuals, as well as derived variables that
+summarize the respondents' use of time use across different activities,
+locations, and social interactions. This file documents the time and
+duration that respondents allocate to each activity and location. The
+*Main file* provides a overview of daily routines and social dynamics,
+not focusing on individual activity episodes. Additionally, this file
+categorizes activities into bigger groups and subcategories,
+facilitating the data's analytical utility with additional metrics such
+as total transit time, time spent with household members, and counts of
+activities and episodes.
 
-Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the TUS PUMF 2015 _Main file_ (Cycle 29).
-Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
+Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
+six variables of the TUS PUMF 2015 *Main file* (Cycle 29). Each row in
+the table correspond to a survey respondent, while the columns refer the
+following information: record identification (`PUMFID`), the person's
+weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`),
+the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and
+the respondent's marital status (`MARSTAT`).
 
-The _Main file_ of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.
-For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
-For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.
+The *Main file* of the 2015 GSS surveys includes a total of 17,390
+respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and
+848 variables. For discrete variables, Statistics Canada has assigned
+specific codes to the possible values, with each code accompanied by a
+label. For instance, in the case of the variable `SURVMNTH`, a value of `1`
+means `January 2016`, `2` means `February 2016`, `3` corresponds to `March 2016`,
+and so on.
 
-As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not labeled.
-Additionally, the format of the tables (comma-separated values) does not allow for the specification of variable types (whether a variable is continuous or discrete), which can lead to mistakes analysts who have limited experience working with PUMFs.
+As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
+labeled. Additionally, the format of the tables (comma-separated values)
+does not allow for the specification of variable types (whether a
+variable is continuous or discrete), which can lead to mistakes analysts
+who have limited experience working with PUMFs.
 
@@
                  font_size = 8) |>
-  footnote(general = c("Legend: `PUMFID`: record identification.
-`WGHT_PER`:  person weight.
-`SURVMNTH`: survey month of data collection.
-`AGEGR10`: age group of the respondent.
-`SEX`: sex of the respondent.
-`MARSTAT`: marital status of the respondent."), threeparttable = T)
+  footnote(general = c("Legend: PUMFID: record identification.
+WGHT_PER:  person weight.
+SURVMNTH: survey month of data collection.
+AGEGR10: age group of the respondent.
+SEX: sex of the respondent.
+MARSTAT: marital status of the respondent."), threeparttable = T)
 
@@
 
-## The _Episode file_
+## The Episode file
 
-The episode is a much bigger file that records detailed data for each activity 
-episode reported by respondents. Each episode represents a single activity and its
-duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
-entry in this file includes the start and end times of the activity, the duration,
-location, and accompanying social context, informing when and where activities 
-occurred and with whom. The focus of the _Episode file_ is not on the characteristics
-of the respondents, but on the characteristics of the activities, and the data 
-are structured around the numerous activity instances that compose a day of the 
-respondent. Although respondent-specific characteristics are not included within
-the _Episode file_, it is possible to link the _Main file_ and the _Episode file_ by 
-using a key present in both the _Main_ and _Episode files_.
+The episode is a much bigger file that records detailed data for each
+activity episode reported by respondents. Each episode represents a
+single activity and its duration, and the sum of all episodes throughout
+the day adds up to 24 hours. Each entry in this file includes the start
+and end times of the activity, the duration, location, and accompanying
+social context, informing when and where activities occurred and with
+whom. The focus of the *Episode file* is not on the characteristics of
+the respondents, but on the characteristics of the activities, and the
+data are structured around the numerous activity instances that compose
+a day of the respondent. Although respondent-specific characteristics
+are not included within the *Episode file*, it is possible to link the
+*Main file* and the *Episode file* by using a key present in both the
+*Main* and *Episode files*.
 
-Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the _Main file_ structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the TUS PUMF 2015 _Episode file_ (Cycle 29).
-Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
+example of the *Main file* structure, Table
+\ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
+record identification number 10041 and some variables from the TUS PUMF
+2015 *Episode file* (Cycle 29). Each row in the table corresponds to an
+episode associated with the specified record identification (`PUMFID` =
+`10041`), including the episode's weight (`WGHT_EPI`), episode number
+(`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and
+episode location (`LOCATION`).
 
-In total, the _Episode file_ of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes.
-Similar to the _Main file_, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.
+In total, the *Episode file* of the 2015 GSS surveys contains 274,108
+records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527
+variables detailing the episodes. Similar to the *Main file*, Statistics
+Canada has created codes for the discrete variables, with each value
+corresponding to a label.
 
-In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
-The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`.
-Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where both the origin and destination were their home.
-Such trips, where the journey starts and finishes at home, can be classified as recreational or leisure trips.
-Next, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
-Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available.
-As shown, since the dataset does not label the variable values, decoding them can be both time-consuming and challenging.
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this
+respondent began the diary description by sleeping at home (`TUI_01 = 1`
+and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal
+hygiene (`TUI_01 = 2`). The respondent then spent 15 minutes on personal
+care activities, such as getting ready for school, supervising homework,
+reading, playing, reprimanding, or providing educational or emotional
+support, as indicated by `TUI_01 = 27.` Next, they recorded a travel
+episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where
+both the origin and destination were their home. Such trips, where the
+journey starts and finishes at home, can be classified as recreational
+or leisure trips. Next, the respondent spent 3 hours searching for a job
+(`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then
+cleaned the house (`TUI_01 = 18`) for two hours. Table
+\ref{tab:ep-2015-unprocessed} displays only six variables out of the 527
+available. As shown, since the dataset does not label the variable
+values, decoding them can be both time-consuming and challenging.
 
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") |>
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.") |>
   kable_styling(full_width = FALSE, 
@@
                  font_size = 8) |>
-  footnote(general = c("Legend: `PUMFID`: record identification.
-`EPINO`: episode number.
+  footnote(general = c("Legend: PUMFID: record identification.
+EPINO: episode number.
 WGHT_EPI: episode's weight.
@@
 
-Figure \ref{fig:process-figure} presents a diagram illustrating the processes applied to the _Main_ and _Episode files_ to create the {ActiveCA} datasets.
-For each TUS cycle, we reviewed the _Episode files_ to identify movement episodes involving walking or cycling.
-This allowed us to also select the activities immediately before and after the movement episode, helping to infer the trip’s purpose and determine its origin and destination.
-Active trips were identified by their corresponding activity codes, accounting for variations across survey cycles.
+Figure \ref{fig:process-figure} presents a diagram illustrating the
+processes applied to the *Main* and *Episode files* to create the
+{ActiveCA} datasets. For each TUS cycle, we reviewed the *Episode files*
+to identify movement episodes involving walking or cycling. This allowed
+us to also select the activities immediately before and after the
+movement episode, helping to infer the trip’s purpose and determine its
+origin and destination. Active trips were identified by their
+corresponding activity codes, accounting for variations across survey
+cycles.
 
-Next, we labeled the coded variables with their appropriate descriptions, classifying each origin and destination, mode of travel, and time spent in the active trip.
-Categorical variables were converted into factor variables, and ordinal variables were defined as ordered factors.
+Next, we labeled the coded variables with their appropriate
+descriptions, classifying each origin and destination, mode of travel,
+and time spent in the active trip. Categorical variables were converted
+into factor variables, and ordinal variables were defined as ordered
+factors.
 
-For the _Main files_, we selected socioeconomic variables to help profile individuals engaged in active travel episodes.
-These included key indicators such as age group, sex, marital status, and number of children, among others.
-As with the _Episode files_, we labeled and factored the coded variables.
-Using the appropriate identifiers, we joined the Episode and _Main files_ to determine the province and the type of population centre - whether it is a CMA, a Census Agglomeration (CA), or a non-CMA/CA area.
+For the *Main files*, we selected socioeconomic variables to help
+profile individuals engaged in active travel episodes. These included
+key indicators such as age group, sex, marital status, and number of
+children, among others. As with the *Episode files*, we labeled and
+factored the coded variables. Using the appropriate identifiers, we
+joined the Episode and *Main files* to determine the province and the
+type of population centre - whether it is a CMA, a Census Agglomeration
+(CA), or a non-CMA/CA area.
 
-As output, the {ActiveCA} package provides processed datasets of walking and cycling episodes by year, a unified dataset that contains all walking and cycling episodes across TUS cycles, and processed _Main files_ (not unified, as socioeconomic and demographic variables and their categories differ across TUS cycles).
+As output, the {ActiveCA} package provides processed datasets of walking
+and cycling episodes by year, a unified dataset that contains all
+walking and cycling episodes across TUS cycles, and processed main
+files (not unified, as socioeconomic and demographic variables and
+their categories differ across TUS cycles).
 
@@
 
-Table \ref{tab:processed-obs} displays the total number of records processed for Main and _Episode files_.
-For the _Main files_, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2022, that together represents more of `r sum(df_main_processed$Weighted)` respondents.
-It also presents the total cases of active trips episodes identified.
-In total `r sum(df_ep_processed$Count)` records with register of active travel activity.
-Together, these records account for `r sum(df_ep_processed$Weighted)` episodes.
+Table \ref{tab:processed-obs} displays the total number of records
+processed for Main and *Episode files*. For the *Main files*, a total of
+`r sum(df_main_processed$Count)` records were processed, referring to
+all records from the Time Use Surveys from 1986 to 2022, that together
+represents more of `r sum(df_main_processed$Weighted)` respondents. It
+also presents the total cases of active trips episodes identified. In
+total `r sum(df_ep_processed$Count)` records with register of active
+travel activity. Together, these records account for
+`r sum(df_ep_processed$Weighted)` episodes.
 
@@
 
-Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
-Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the TUS PUMF 2015 _Episode file_ (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
-Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes.
-For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.
+Table \ref{tab:main-2015-processed} shows the first ten rows and first
+six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
+\ref{tab:main-2015-unprocessed} before our processing. Table
+\ref{tab:ep-2015-processed} presents the walking episodes for the record
+identification number `10041` from the TUS PUMF 2015 *Episode file* (Cycle
+29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only
+the unique active travel episode appear in Table
+\ref{tab:ep-2015-processed} since the records were filtered to select
+cases with walking or cycling episodes. For both cases, Tables
+\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
+labeled variables, facilitating the interpretation of the data.
 
@@
     longtable = TRUE, 
-    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.", 
+    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.", 
        col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) |> 
@@
 
-Considering all TUS analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations.
-{ActiveCA} includes all these episodes ready for analysis.
-Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2022, with measures of trip duration in minutes.
-The 1986 survey did not include bicycle trips.
+Considering all TUS analyzed, we identified `r nrow(gss_episodes)`
+episodes that recorded active travel episodes, with trip duration
+ranging from 0 to 900 minutes, to twelve different destinations.
+{ActiveCA} includes all these episodes ready for analysis. Table
+\ref{tab:table-01} presents descriptive statistics on walking and
+cycling trips between 1986 and 2022, with measures of trip duration in
+minutes. The 1986 survey did not include bicycle trips.
 
@@
 
-Table \ref{tab:table-01} shows that, until 2022 the median values for walking trips were 10 minutes, increasing to 15 minutes in the last survey.
-In the case of cycling trips, the duration fluctuated over the years, ranging from 10 to 30 minutes.
-The table also highlights very high maximum values, particularly for walking trips, with recorded episodes exceeding 4 hours in all cases.
+Table \ref{tab:table-01} shows that, until 2022 the median values for
+walking trips were 10 minutes, increasing to 15 minutes in the last
+survey. In the case of cycling trips, the duration fluctuated over the
+years, ranging from 10 to 30 minutes. The table also highlights very
+high maximum values, particularly for walking trips, with recorded
+episodes exceeding 4 hours in all cases.
 
-{ActiveCA} also enables visual analysis of active travel in Canada using exploratory data analysis techniques.
-Figure \ref{fig:figure-01} shows walking trips from 2022 through heat maps.
-This graph uses color gradients to represent the percentage of trips between various origins and destinations, with darker colors indicating higher percentages and lighter colors representing less frequent routes.
-For conciseness, we omitted the heat maps for the other years analyzed.
+{ActiveCA} also enables visual analysis of active travel in Canada using
+exploratory data analysis techniques. Figure \ref{fig:figure-01} shows
+walking trips from 2022 through heat maps. This graph uses color
+gradients to represent the percentage of trips between various origins
+and destinations, with darker colors indicating higher percentages and
+lighter colors representing less frequent routes. For conciseness, we
+omitted the heat maps for the other years analyzed.
 
-In 2022, `home` served as a central hub for most trips, with fewer than 10% of journeys not involving it as either a starting point or destination.
-The most common trip types were from `home` to `work or school` (17%) and the reverse, from `work or school` to `home` (13%).
-Notably, 7% of trips began and ended at `home`, often reflecting leisure activities such as short walks or dog walking.
-`Grocery stores` were also a key destination, comprising 10% of trips departing from `home.`
+In 2022, `home` served as a central hub for most trips, with fewer than
+10% of journeys not involving it as either a starting point or
+destination. The most common trip types were from `home` to `work or school`
+(17%) and the reverse, from `work or school` to `home` (13%). Notably, 7% of
+trips began and ended at `home`, often reflecting leisure activities such
+as short walks or dog walking. `Grocery stores` were also a key
+destination, comprising 10% of trips departing from `home.`
 
@@
 
-The {ActiveCA} dataset also includes information on the type of population centre in which respondents reside - specifically, whether they live in a CMA, a CA, or outside these areas - as well as the respondent’s province.
-This information is important, as patterns of active travel often differ between metropolitan and non-metropolitan populations.
-For example, Table \ref{tab:cma-durations} presents the median walking durations by population centre type and province for 2022.
-Overall, respondents living in CMA/CA areas tend to report higher median walking durations compared to those living outside these centres.
-The most pronounced difference is observed in Nova Scotia: metropolitan residents reported a median walking duration of 30 minutes, whereas non-metropolitan residents reported a median of only 5 minutes.
+The {ActiveCA} dataset also includes information on the type of
+population centre in which respondents reside - specifically, whether
+they live in a CMA, a CA, or outside these areas - as well as the
+respondent’s province. This information is important, as patterns of
+active travel often differ between metropolitan and non-metropolitan
+populations. For example, Table \ref{tab:cma-durations} presents the
+median walking durations by population centre type and province for
+2022. Overall, respondents living in CMA/CA areas tend to report higher
+median walking durations compared to those living outside these centres.
+The most pronounced difference is observed in Nova Scotia: metropolitan
+residents reported a median walking duration of 30 minutes, whereas
+non-metropolitan residents reported a median of only 5 minutes.
 
@@
   add_header_above(c(" " = 1, "Population centre type" = 2)) |> 
-  footnote(general = c("'CMA' denotes Census Metropolitan Area and 'CA' denotes Census agglomeration."), threeparttable = T)
+  footnote(general = c("CMA denotes Census Metropolitan Area and CA denotes Census agglomeration."), threeparttable = T)
 
@@
 
-The package also enables obtaining insights from the main processed files.
-Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2022.
-According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 15% of possible cases.
+The package also enables obtaining insights from the main processed
+files. Figure \ref{fig:figure-stress} present how the level of stress
+varied among respondents depending on their marital status in 2022.
+According to this plot, married respondents reported the highest level
+of stress, relating to feel stressed every day, with 15% of possible
+cases.
 
@@
 
-{ActiveCA} also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames.
-This process allows users to work with and utilize the datasets available in {ActiveCA} within a Python project.
+{ActiveCA} also provides a Jupyter Notebook containing a `Python` script
+that demonstrates how to read `R` data files (.rda) and convert them into
+Pandas DataFrames. This process allows users to work with and utilize
+the datasets available in {ActiveCA} within a `Python` project.
 
@@
 
-This paper presents {ActiveCA}, an open data product that provides analysis-ready 
-data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) 
-of TUS GSSs on active travel in Canada. In the form of an `R` data package, 
-{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
-providing information on origins, destinations, and duration of active travel, as
+This paper presents {ActiveCA}, an open data product that provides
+analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
+(2005), 24 (2010), 29 (2015), and 34 (2022) of TUS GSSs on active travel
+in Canada. In the form of an `R` data package, {ActiveCA} was developed
+after collecting, cleaning, and processing the survey data, providing
+information on origins, destinations, and duration of active travel, as
 well other information.
 
-Although we did not select non-AT episodes, the process for obtaining them is very similar to that used for selecting AT episodes.
-Researchers interested in non-AT modes can use our framework to guide their methodology, making the small but necessary adjustments.
-We focused exclusively on AT episodes because the {ActiveCA} package is part of a larger project aimed at analyzing the historical evolution of active travel behaviour in Canada.
+Although we did not select non-AT episodes, the process for obtaining
+them is very similar to that used for selecting AT episodes. Researchers
+interested in non-AT modes can use our framework to guide their
+methodology, making the small but necessary adjustments. We focused
+exclusively on AT episodes because the {ActiveCA} package is part of a
+larger project aimed at analyzing the historical evolution of active
+travel behaviour in Canada.
 
-The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
-which facilitates the addition of complementary data sets in the future. `R` users
-can seamlessly explore TUS walking and cycling episodes, 
-with the option to suggest enhancements to the package as needed. This article adopts
-the structure proposed by Anastasia and Páez [-@soukhov2023], whose work provided
-essential guidance for the creation of this package. Similarly, we aim to contribute
-to the academic community by promoting transparent research practices that encourage
-replication and innovation in related fields. We believe that {ActiveCA} will serve
-as a basis for further research on TUS and for the integration of additional data by
-the authors or the wider open source community.
+The value of {ActiveCA} lies in its transparency, accessibility, and
+ease of use, which facilitates the addition of complementary data sets
+in the future. R users can seamlessly explore TUS walking and cycling
+episodes, with the option to suggest enhancements to the package as
+needed. This article adopts the structure proposed by Anastasia and Páez
+[-@soukhov2023], whose work provided essential guidance for the creation
+of this package. Similarly, we aim to contribute to the academic
+community by promoting transparent research practices that encourage
+replication and innovation in related fields. We believe that {ActiveCA}
+will serve as a basis for further research on TUS and for the
+integration of additional data by the authors or the wider open source
+community.
 
@@
 
-For review purposes, the package is currently available as a tar.gz file that can 
-be installed by R. The file can be obtained from this anonymous location:
+For review purposes, the package is currently available as a tar.gz file
+that can be installed by R. The file can be obtained from this anonymous
+location:
 
-https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip
+<https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip>
 

### 2025-06-01 paper/Manuscript-Data-Package-1Round-Review.tex
Create a document with all manuscript changes and add answer letters for the reviewers.

@@
 This paper describes \{ActiveCA\}, an open data product with Canadian
-time use data. \{ActiveCA\} is an \texttt{R} data package that contains
+time use data. \{ActiveCA\} is an R data package that contains
 analysis-ready data related to active travel spanning almost 40 years,
@@
 data, and documentation in a standardized format that can be installed
-by \texttt{R} users via a software repository, such as CRAN
-(Comprehensive R Archive Network) or GitHub, which makes them an adroit
-medium to distribute analysis-ready data.
+by R users via a software repository, such as CRAN (Comprehensive R
+Archive Network) or GitHub, which makes them an adroit medium to
+distribute analysis-ready data.
 
@@
 
-\subsection{\texorpdfstring{The \emph{Main
-file}}{The Main file}}\label{the-main-file}
+\subsection{The Main file}\label{the-main-file}
 
@@
 the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
-\texttt{January\ 2016}, 2 means \texttt{February\ 2016}, 3 corresponds
-to \texttt{March\ 2016}, and so on.
+\texttt{January\ 2016}, \texttt{2} means \texttt{February\ 2016},
+\texttt{3} corresponds to \texttt{March\ 2016}, and so on.
 
@@
 \item \textit{Note: } 
-\item Legend: `PUMFID`: record identification.
-`WGHT\_PER`:  person weight.
-`SURVMNTH`: survey month of data collection.
-`AGEGR10`: age group of the respondent.
-`SEX`: sex of the respondent.
-`MARSTAT`: marital status of the respondent.
+\item Legend: PUMFID: record identification.
+WGHT\_PER:  person weight.
+SURVMNTH: survey month of data collection.
+AGEGR10: age group of the respondent.
+SEX: sex of the respondent.
+MARSTAT: marital status of the respondent.
 \end{TableNotes}
@@
 
-\subsection{\texorpdfstring{The \emph{Episode
-file}}{The Episode file}}\label{the-episode-file}
+\subsection{The Episode file}\label{the-episode-file}
 
@@
 \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
-record identification number \texttt{10041} and some variables from the
-TUS PUMF 2015 \emph{Episode file} (Cycle 29). Each row in the table
-corresponds to an episode associated with the specified record
-identification (\texttt{PUMFID\ =\ 10041}), including the episode's
-weight (\texttt{WGHT\_EPI}), episode number (\texttt{EPINO}), activity
-code (\texttt{TUI\_01}), episode duration (\texttt{DURATION}), and
-episode location (\texttt{LOCATION}).
+record identification number 10041 and some variables from the TUS PUMF
+2015 \emph{Episode file} (Cycle 29). Each row in the table corresponds
+to an episode associated with the specified record identification
+(\texttt{PUMFID} = \texttt{10041}), including the episode's weight
+(\texttt{WGHT\_EPI}), episode number (\texttt{EPINO}), activity code
+(\texttt{TUI\_01}), episode duration (\texttt{DURATION}), and episode
+location (\texttt{LOCATION}).
 
@@
 reprimanding, or providing educational or emotional support, as
-indicated by \texttt{TUI\_01\ =\ 27}. Next, they recorded a travel
+indicated by \texttt{TUI\_01\ =\ 27.} Next, they recorded a travel
 episode, walking for 15 minutes (\texttt{TUI\_01\ =\ 7} and
@@
 \item \textit{Note: } 
-\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
 \end{TableNotes}
 \begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.}\\
+\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.}\\
 \toprule
@@
 walking and cycling episodes by year, a unified dataset that contains
-all walking and cycling episodes across TUS cycles, and processed
-\emph{Main files} (not unified, as socioeconomic and demographic
-variables and their categories differ across TUS cycles).
+all walking and cycling episodes across TUS cycles, and processed main
+files (not unified, as socioeconomic and demographic variables and their
+categories differ across TUS cycles).
 
@@
 \item \textit{Note: } 
-\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
 \end{TableNotes}
 \begin{longtable}[t]{rrlrlll}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.}\\
+\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.}\\
 \toprule
@@
 \item \textit{Note: } 
-\item 'CMA' denotes Census Metropolitan Area and 'CA' denotes Census agglomeration.
+\item CMA denotes Census Metropolitan Area and CA denotes Census agglomeration.
 \end{tablenotes}
@@
 
-\{ActiveCA\} also provides a Jupyter Notebook containing a Python script
-that demonstrates how to read R data files (.rda) and convert them into
-Pandas DataFrames. This process allows users to work with and utilize
-the datasets available in \{ActiveCA\} within a Python project.
+\{ActiveCA\} also provides a Jupyter Notebook containing a
+\texttt{Python} script that demonstrates how to read \texttt{R} data
+files (.rda) and convert them into Pandas DataFrames. This process
+allows users to work with and utilize the datasets available in
+\{ActiveCA\} within a \texttt{Python} project.
 
@@
 ease of use, which facilitates the addition of complementary data sets
-in the future. \texttt{R} users can seamlessly explore TUS walking and
-cycling episodes, with the option to suggest enhancements to the package
-as needed. This article adopts the structure proposed by Anastasia and
-Páez \citeyearpar{soukhov2023}, whose work provided essential guidance
-for the creation of this package. Similarly, we aim to contribute to the
+in the future. R users can seamlessly explore TUS walking and cycling
+episodes, with the option to suggest enhancements to the package as
+needed. This article adopts the structure proposed by Anastasia and Páez
+\citeyearpar{soukhov2023}, whose work provided essential guidance for
+the creation of this package. Similarly, we aim to contribute to the
 academic community by promoting transparent research practices that

### 2025-06-01 paper/Manuscript-Data-Package2.Rmd
Create a document with all manuscript changes and add answer letters for the reviewers.

--- 2025-01-17 paper/Manuscript-Data-Package.Rmd
@@
   "This paper describes {ActiveCA}, an open data product with Canadian time use data.
-{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
+{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
 Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
@@
 # Original main and Episode Files from 2015 to display table examples
-gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
-gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))
+gss_m_2015 <- read.csv(paste0(here(),
+                              "/data-raw/source-files/Time-use-2015/gss-m.csv"))
+gss_e_2015 <- read.csv(paste0(here(), 
+                              "/data-raw/source-files/Time-use-2015/gss-e.csv"))
 
@@
 
-{ActiveCA} is distributed as an `R` package with a number of data 
-objects and their documentation. `R` packages contain code, data, and documentation
-in a standardized format that can be installed by `R` users via a software repository,
+{ActiveCA} is distributed as an R package with a number of data 
+objects and their documentation. R packages contain code, data, and documentation
+in a standardized format that can be installed by R users via a software repository,
 such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29).
-Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
+Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (PUMFID), the person's weight (WGHT_PER), the month the survey data was collected (SURVMNTH), the respondent's age group (AGEGR10), the respondent's sex (SEX), and the respondent's marital status (MARSTAT).
 
@@
 For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
-For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.
+For instance, in the case of the variable SURVMNTH, a value of 1 means January 2016, 2 means February 2016, 3 corresponds to March 2016, and so on.
 
@@
 ```{r gss-main-file-2015}
-main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  %>%
+main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  |>
   kable(
@@
 
-Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
-Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number 10041 and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
+Each row in the table corresponds to an episode associated with the specified record identification (PUMFID = 10041), including the episode's weight (WGHT_EPI), episode number (EPINO), activity code (TUI_01), episode duration (DURATION), and episode location (LOCATION).
 
@@
 
-In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
-The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`.
-Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where both the origin and destination were their home.
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (TUI_01 = 1 and LOCATION = 300) for 210 minutes, followed by 40 minutes of personal hygiene (TUI_01 = 2).
+The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by TUI_01 = 27.
+Next, they recorded a travel episode, walking for 15 minutes (TUI_01 = 7 and LOCATION = 315), where both the origin and destination were their home.
 Such trips, where the journey starts and finishes at home, can be classified as recreational or leisure trips.
-Next, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
+Next, the respondent spent 3 hours searching for a job (TUI_01 = 9), took a 15-minute lunch break (TUI_01 = 6), and then cleaned the house (TUI_01 = 18) for two hours.
 Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") %>%
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.") |>
   kable_styling(full_width = FALSE, 
@@
                 latex_options=c("scale_down"),
-                 font_size = 8) %>%
-  footnote(general = c("Legend: `PUMFID`: record identification.
-`EPINO`: episode number.
+                 font_size = 8) |>
+  footnote(general = c("Legend: PUMFID: record identification.
+EPINO: episode number.
 WGHT_EPI: episode's weight.
@@
 
-By running `?ActiveCA` in `R`, you can access the complete package documentation.
-The documentation includes a description of the package, information about the authors and maintainers, and an overview of all pre-processed data sets included in the package.
+The documentation of the package can be accessed by running ?ActiveCA in R.
+The documentation includes a description of the package, information about the authors, contributors, and maintainers, and an overview of all pre-processed data sets included in the package.
 
-Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (`walking_2015.rda`).
+Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (walking_2015.rda).
 The documentation provides a brief description of the dataset, instructions on how to load the package, as well as details about the file format and descriptions of all variables (already labeled).
@@
 
-The chunk below shows how to load the walking episodes from the 2015 GSS (Cycle 29):
+The code below shows how to load the walking episodes from the 2015 GSS (Cycle 29):
 
@@
 The number of episodes and persons represented by a episode or person is determined by the weight or weighting factor.
-Because of this, every estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
+Because of this, estimates of the number of episodes or persons need to be calculated applying the corresponding weighting factors.
 
-For instance, if someone wants to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weights.
-In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group.
-It uses the `dplyr`package [@dplyr2023] to manipulate the data.
+For instance, to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weight.
+In 2015, the weight variable is represented by WGHT_PER.
+The code below demonstrates how to obtain the percentage of people with active travel episodes by age group.
+It uses the dplyr package [@dplyr2023] to manipulate the data.
 
@@
 Then, it joins the 2015 episodes with the 2015 Main File.
-Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
-After this, a new dataset, `Active_percentage`, is created by merging both previous datasets.
+Note that in both operations, the code sums the person weight variable (WGHT_PER) to obtain the correct population values.
+After this, a new dataset, Active_percentage, is created by merging both previous datasets.
 The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100 and rounding with 2 decimal places.
 
-The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 37%, followed by those between 25 and 34 years old with around 33%.
-There is a significant drop in percentage for the following groups, with the percentage falling to between 15% and 18% for the remaining age groups.
-
-```{r percentages-act, echo=TRUE, out.width='80%'}
+```
 # Load the dplyr library for data manipulation
@@
 #and calculate the percentage
-Active_percentage <- Total_population %>% 
-  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
+Active_percentage <- Total_population |> 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
+  mutate(
+    Percentage = round(Active_population/Total_population*100,2))
+```
+The result shows that the age group with the highest share of active trips is those between 15 and 20 years old, with almost 37%, followed by those between 25 and 34 years old with around 33%.
+There is a significant drop in percentage for the following groups, with the percentage falling to between 15% and 18% for the remaining age groups.
+
+```{r percentages-act, echo=FALSE, out.width='80%'}
+# Load the dplyr library for data manipulation
+library('dplyr')
+
+# Load the episodes dataset
+data(gss_episodes)
+
+# Calculate the total population for each age group
+Total_population <- gss_main_2015 |> 
+  group_by(AGEGR10) |> 
+  summarise(Total_population  = sum(WGHT_PER))
+
+# Calculate the active population for each age group
+Active_population <- gss_episodes |> 
+      filter(YEAR == 2015) |> 
+  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) |> 
+  group_by(AGEGR10) |> 
+  summarise(Active_population  = sum(WGHT_PER))
+
+# Combine total and active population 
+#and calculate the percentage
+Active_percentage <- Total_population |> 
+  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) |>
   mutate(
@@
 
-table_df_processed <- df_processed %>%
+table_df_processed <- df_processed |>
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
-        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") %>%
-  kable_styling(full_width = FALSE, font_size = 10) %>%
-  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) %>%
+        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
+  kable_styling(full_width = FALSE, font_size = 10) |>
+  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) |>
   collapse_rows(columns = 1, valign = "top", latex_hline = "major")
@@
 Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
-Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number 10041 from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
 Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
+    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") |>
   kable_styling(full_width = FALSE, 
@@
 # Creating a latex table
-styled_ch03_table_01 <- stats_data %>%
+styled_ch03_table_01 <- stats_data |>
   kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
         caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
-        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) %>%
-  kable_styling(full_width = FALSE, font_size = 10) %>%
-  column_spec(1, bold = TRUE) %>%
-  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) %>%
+        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) |>
+  kable_styling(full_width = FALSE, font_size = 10) |>
+  column_spec(1, bold = TRUE) |>
+  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) |>
   collapse_rows(columns = 1, valign = "top", latex_hline = "major")
@@
 Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
-walking trips destined for `home` had the highest medians. However, by
-1998, the highest medians shifted to trips to `work or school`, a
+walking trips destined for home had the highest medians. However, by
+1998, the highest medians shifted to trips to work or school, a
 transition that also occurred for cycling trips between 1992 and 1998.
 Table \ref{tab:table-03} indicates that the median duration for walking trips to
-`home` and `work or school` remained at 10 minutes.
+home and work or school remained at 10 minutes.
 
@@
 # Filter for selected years and calculate statistics of trip duration
-episodes_stats_before <- gss_episodes %>%
-  filter(YEAR < 2005) %>%
-  group_by(YEAR, dest_label, MODE) %>%
+episodes_stats_before <- gss_episodes |>
+  filter(YEAR < 2005) |>
+  group_by(YEAR, dest_label, MODE) |>
   dplyr::summarise(
@@
 # Calculate percentages of trips by mode/year 
-total_counts <- episodes_stats_before %>% 
-  group_by(YEAR, MODE) %>% 
+total_counts <- episodes_stats_before |> 
+  group_by(YEAR, MODE) |> 
   summarise(total_count = sum(count), 
@@
 # Pivot data for the table layout
-wide_data <- episodes_stats_before %>%
+wide_data <- episodes_stats_before |>
   pivot_wider(
@@
     values_from = c(min, median, max, Percentage)
-  )  %>% 
+  )  |> 
   arrange(MODE, dest_label, desc(Percentage_1998))
 
-wide_data <- wide_data %>%
+wide_data <- wide_data |>
  dplyr::select(
@@
     min_1998, median_1998, max_1998, Percentage_1998
-  ) %>% 
+  ) |> 
   arrange(MODE, dest_label)
@@
 # Latex table
-styled_ch03_table_02 <- wide_data %>%
+styled_ch03_table_02 <- wide_data |>
   kable(
@@
     col.names = c("Destination", "Mode","Min","Med","Max", "(%)", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
-  ) %>%
+  ) |>
   kable_styling(full_width = FALSE, 
@@
                 latex_options=c("scale_down"),
-                 font_size = 6) %>%
-  # column_spec(1, width = "1cm") %>%  # Set width of 'Dest' column to 1.5cm
-  # column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
-  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) %>%
-  row_spec(0, bold = TRUE, align = "c") %>%
-  column_spec(6, border_right = TRUE) %>%
-  column_spec(10, border_right = TRUE) %>%
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
+                 font_size = 6) |>
+  # column_spec(1, width = "1cm") |>  # Set width of 'Dest' column to 1.5cm
+  # column_spec(2:14, width = "0.5cm") |>  # Set width of remaining columns to 0.7cm
+  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) |>
+  row_spec(0, bold = TRUE, align = "c") |>
+  column_spec(6, border_right = TRUE) |>
+  column_spec(10, border_right = TRUE) |>
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major") |> 
   footnote(general = c("'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
@@
 # Selecting surveys after 2000 and calculating trip duration statistics 
-episodes_stats_after <- gss_episodes_renamed %>%
-  filter(YEAR > 2000) %>%
-  group_by(YEAR, dest_label, MODE) %>%
+episodes_stats_after <- gss_episodes_renamed |>
+  filter(YEAR > 2000) |>
+  group_by(YEAR, dest_label, MODE) |>
    dplyr::summarise(
@@
 # Calculate percentages of trips by mode/year
-total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
+total_counts <- episodes_stats_after |> group_by(YEAR, MODE) |> summarise(total_count = sum(count), .groups = "drop")
 episodes_stats_after <- left_join(episodes_stats_after, total_counts, by = c("YEAR", "MODE"))
@@
 # Pivot data for the table layout
-wide_data <- episodes_stats_after %>%
+wide_data <- episodes_stats_after |>
   pivot_wider(
@@
     values_from = c(min, median, max, Percentage)
-  ) %>%
+  ) |>
  dplyr::select(
@@
     min_2015, median_2015, max_2015, Percentage_2015
-  ) %>% 
+  ) |> 
   arrange(MODE, dest_label)
@@
 # Latex table
-styled_ch03_table_03 <- wide_data %>%
+styled_ch03_table_03 <- wide_data |>
   kable(
@@
                 latex_options=c("scale_down"),
-                font_size = 6) %>%
-  #column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
-  #column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
-  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
-  row_spec(0, bold = TRUE, align = "c") %>%
-  column_spec(6, border_right = TRUE) %>%
-  column_spec(10, border_right = TRUE) %>% 
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
+                font_size = 6) |>
+  #column_spec(1, width = "2cm") |>  # Set width of 'Dest' column to 1.5cm
+  #column_spec(2:14, width = "0.4cm") |>  # Set width of remaining columns to 0.7cm
+  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) |>
+  row_spec(0, bold = TRUE, align = "c") |>
+  column_spec(6, border_right = TRUE) |>
+  column_spec(10, border_right = TRUE) |> 
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major") |> 
   footnote(general = c("'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)
@@
 
-In 1992, walking trips with `home` as both the origin and destination
+In 1992, walking trips with home as both the origin and destination
 made up the majority, accounting for about 30% of all walking trips.
 These trips often involved leisure activities, like short walks or dog
-walking. Following this, trips from `home` to `work or school` comprised
-18% of walking trips. Overall, `home` emerged as a crucial hub, either
-as an origin or destination, with only 5% of trips not involving `home.`
-By 2015, `home` remained a significant node, but new locations
+walking. Following this, trips from home to work or school comprised
+18% of walking trips. Overall, home emerged as a crucial hub, either
+as an origin or destination, with only 5% of trips not involving home.
+By 2015, home remained a significant node, but new locations
 distributed the proportion of trips to areas not considered in 1992. In
-2015, the highest proportion of trips were from `home` to
-`work or school` (12%) and vice versa (11%). `home` to `home` accounted
-for 8% of trips, and `grocery stores` became a notable destination for
-those leaving `home` (6%), surpassing trips to `other's home` (4%).
+2015, the highest proportion of trips were from home to
+work or school (12%) and vice versa (11%). home to home accounted
+for 8% of trips, and grocery stores became a notable destination for
+those leaving home (6%), surpassing trips to other's home (4%).
 
@@
 # Creating heatmaps for the years of 1992 and 2015
-gss_heatmap <- gss_episodes_renamed %>%
-  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) %>% 
+gss_heatmap <- gss_episodes_renamed |>
+  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) |> 
   filter(YEAR %in% c(1992,2015))
@@
 
-    trip_counts <- gss_heatmap %>%
-      filter(YEAR == year, MODE == mode) %>% 
-      group_by(orig_label, dest_label) %>%
+    trip_counts <- gss_heatmap |>
+      filter(YEAR == year, MODE == mode) |> 
+      group_by(orig_label, dest_label) |>
         dplyr::summarise(count_each_trip = n(), 
@@
       # Calculate percentages and round to 2 decimal places
-      trip_counts <- trip_counts %>%
+      trip_counts <- trip_counts |>
         mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))
@@
 this mode of transportation was first included as an activity, the
-majority of trips were from `home` to `work or school`, accounting for
+majority of trips were from home to work or school, accounting for
 about 25% of cases. This pattern remained in 2015, with these trips
 representing 30% of the cases. However, a notable change occurred in
-`home` to `home` trips, which decreased significantly from 19% in 1992
+home to home trips, which decreased significantly from 19% in 1992
 to 5% in 2015.
@@
 ```{r main_stress_figure}
-main_stress_figure <- gss_main_2015 %>%
-  filter(!is.na(MARSTAT), !is.na(GTU_110)) %>%
-  group_by(MARSTAT, GTU_110) %>%
-  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") %>%
-  mutate(percentage = n / sum(n) * 100) %>%
-  ungroup() %>%
+main_stress_figure <- gss_main_2015 |>
+  filter(!is.na(MARSTAT), !is.na(GTU_110)) |>
+  group_by(MARSTAT, GTU_110) |>
+  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") |>
+  mutate(percentage = n / sum(n) * 100) |>
+  ungroup() |>
   ggplot(aes(x = MARSTAT,
@@
 data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
-of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
+of the GSS surveys on active travel in Canada. In the form of an R data package, 
 {ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
@@
 at the moment it is estimated that they will only be published in the later part
-of 2025. The `R` package will be updated once these files become available.
+of 2025. The R package will be updated once these files become available.
 
 The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
-which facilitates the addition of complementary data sets in the future. `R` users
+which facilitates the addition of complementary data sets in the future. R users
 can seamlessly explore GSS walking and cycling episodes, 

### 2025-06-01 paper/Manuscript-Data-Package2.tex
Create a document with all manuscript changes and add answer letters for the reviewers.

--- 2025-01-22 paper/Manuscript-Data-Package.tex
@@
 
-\{ActiveCA\} is distributed as an \texttt{R} package with a number of
-data objects and their documentation. \texttt{R} packages contain code,
-data, and documentation in a standardized format that can be installed
-by \texttt{R} users via a software repository, such as CRAN
-(Comprehensive R Archive Network) or GitHub, which makes them an adroit
-medium to distribute analysis-ready data.
+\{ActiveCA\} is distributed as an R package with a number of data
+objects and their documentation. R packages contain code, data, and
+documentation in a standardized format that can be installed by R users
+via a software repository, such as CRAN (Comprehensive R Archive
+Network) or GitHub, which makes them an adroit medium to distribute
+analysis-ready data.
 
@@
 table correspond to a survey respondent, while the columns refer the
-following information: record identification (\texttt{PUMFID}), the
-person's weight (\texttt{WGHT\_PER}), the month the survey data was
-collected (\texttt{SURVMNTH}), the respondent's age group
-(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
-respondent's marital status (\texttt{MARSTAT}).
+following information: record identification (PUMFID), the person's
+weight (WGHT\_PER), the month the survey data was collected (SURVMNTH),
+the respondent's age group (AGEGR10), the respondent's sex (SEX), and
+the respondent's marital status (MARSTAT).
 
@@
 possible values, with each code accompanied by a label. For instance, in
-the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
-\texttt{January\ 2016}, 2 means \texttt{February\ 2016}, 3 corresponds
-to \texttt{March\ 2016}, and so on.
+the case of the variable SURVMNTH, a value of 1 means January 2016, 2
+means February 2016, 3 corresponds to March 2016, and so on.
 
@@
 presents the first seven episodes for the record identification number
-\texttt{10041} and some variables from the GSS PUMF 2015 Episode File
-(Cycle 29). Each row in the table corresponds to an episode associated
-with the specified record identification (\texttt{PUMFID\ =\ 10041}),
-including the episode's weight (\texttt{WGHT\_EPI}), episode number
-(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
-(\texttt{DURATION}), and episode location (\texttt{LOCATION}).
+10041 and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
+Each row in the table corresponds to an episode associated with the
+specified record identification (PUMFID = 10041), including the
+episode's weight (WGHT\_EPI), episode number (EPINO), activity code
+(TUI\_01), episode duration (DURATION), and episode location (LOCATION).
 
@@
 In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this
-respondent began the diary description by sleeping at home
-(\texttt{TUI\_01\ =\ 1} and \texttt{LOCATION\ =\ 300}) for 210 minutes,
-followed by 40 minutes of personal hygiene (\texttt{TUI\_01\ =\ 2}). The
-respondent then spent 15 minutes on personal care activities, such as
-getting ready for school, supervising homework, reading, playing,
-reprimanding, or providing educational or emotional support, as
-indicated by \texttt{TUI\_01\ =\ 27}. Next, they recorded a travel
-episode, walking for 15 minutes (\texttt{TUI\_01\ =\ 7} and
-\texttt{LOCATION\ =\ 315}), where both the origin and destination were
-their home. Such trips, where the journey starts and finishes at home,
-can be classified as recreational or leisure trips. Next, the respondent
-spent 3 hours searching for a job (\texttt{TUI\_01\ =\ 9}), took a
-15-minute lunch break (\texttt{TUI\_01\ =\ 6}), and then cleaned the
-house (\texttt{TUI\_01\ =\ 18}) for two hours. Table
+respondent began the diary description by sleeping at home (TUI\_01 = 1
+and LOCATION = 300) for 210 minutes, followed by 40 minutes of personal
+hygiene (TUI\_01 = 2). The respondent then spent 15 minutes on personal
+care activities, such as getting ready for school, supervising homework,
+reading, playing, reprimanding, or providing educational or emotional
+support, as indicated by TUI\_01 = 27. Next, they recorded a travel
+episode, walking for 15 minutes (TUI\_01 = 7 and LOCATION = 315), where
+both the origin and destination were their home. Such trips, where the
+journey starts and finishes at home, can be classified as recreational
+or leisure trips. Next, the respondent spent 3 hours searching for a job
+(TUI\_01 = 9), took a 15-minute lunch break (TUI\_01 = 6), and then
+cleaned the house (TUI\_01 = 18) for two hours. Table
 \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527
@@
 
-The documentation of the package can be accessed by running
-\texttt{?ActiveCA} in \texttt{R}. The documentation includes a
-description of the package, information about the authors, contributors,
-and maintainers, and an overview of all pre-processed data sets included
-in the package.
+The documentation of the package can be accessed by running ?ActiveCA in
+R. The documentation includes a description of the package, information
+about the authors, contributors, and maintainers, and an overview of all
+pre-processed data sets included in the package.
 
 Figure \ref{fig:walking_documentation} displays the documentation for
-the walking episodes from the 2015 GSS (\texttt{walking\_2015.rda}). The
+the walking episodes from the 2015 GSS (walking\_2015.rda). The
 documentation provides a brief description of the dataset, instructions
@@
 the person's weight. In 2015, the weight variable is represented by
-\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
-percentage of people with active travel episodes by age group. It uses
-the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.
+WGHT\_PER. The code below demonstrates how to obtain the percentage of
+people with active travel episodes by age group. It uses the dplyr
+package \citep{dplyr2023} to manipulate the data.
 
@@
 that in both operations, the code sums the person weight variable
-(\texttt{WGHT\_PER}) to obtain the correct population values. After
-this, a new dataset, \texttt{Active\_percentage}, is created by merging
-both previous datasets. The percentage of active travel episodes by age
-group is calculated by dividing the total population by the population
-with active trip episodes, then multiplying by 100 and rounding with 2
+(WGHT\_PER) to obtain the correct population values. After this, a new
+dataset, Active\_percentage, is created by merging both previous
+datasets. The percentage of active travel episodes by age group is
+calculated by dividing the total population by the population with
+active trip episodes, then multiplying by 100 and rounding with 2
 decimal places.
@@
 2005, 24,260,137 for 1998, and 21,294,313 for 1986). It also presents
-the total cases of active trips episodes identified. In total 21,748
+the total cases of active trips episodes identified. In total 23,513
 records with register of active travel activity. Together, these records
-account for 38,275,077 episodes.
+account for 44,316,110 episodes.
 
@@
 \cmidrule{1-4}\pagebreak[0]
+ & 2022 & 1765 & 6041032\\
+\nopagebreak
  & 2015 & 3496 & 6634387\\
@@
 \nopagebreak
-\multirow[t]{-6}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
+\multirow[t]{-7}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
 \bottomrule
@@
 \ref{tab:ep-2015-processed} presents the walking episodes for the record
-identification number \texttt{10041} from the GSS PUMF 2015 Episode File
-(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
-Only the unique active travel episode appear in Table
+identification number 10041 from the GSS PUMF 2015 Episode File (Cycle
+29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only
+the unique active travel episode appear in Table
 \ref{tab:ep-2015-processed} since the records were filtered to select
@@
 
-Considering GSS Cycles analyzed, we identified 21,748 episodes that
+Considering GSS Cycles analyzed, we identified 23,513 episodes that
 recorded active travel episodes, with trip duration ranging from 0 to
@@
 
-\begin{longtable}[t]{>{}llcccccc}
+\begin{longtable}[t]{>{}llccccccc}
 \caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
-\cmidrule(l{3pt}r{3pt}){3-8}
-Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
+\multicolumn{2}{c}{ } & \multicolumn{7}{c}{Year} \\
+\cmidrule(l{3pt}r{3pt}){3-9}
+Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015 & 2022\\
 \midrule
- & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
+ & Maximum & 660 & 300 & 255 & 515 & 480 & 900 & 480\\
 \nopagebreak
- & Mean & 21 & 21 & 12 & 12 & 13 & 18\\
+ & Mean & 21 & 21 & 12 & 12 & 13 & 18 & 19\\
 \nopagebreak
- & Median & 15 & 10 & 5 & 10 & 10 & 10\\
+ & Median & 15 & 10 & 5 & 10 & 10 & 10 & 15\\
 \nopagebreak
- & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
+ & Minimum & 1 & 1 & 1 & 0 & 0 & 5 & 5\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
-\cmidrule{1-8}\pagebreak[0]
- & Maximum &  & 240 & 90 & 180 & 153 & 120\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27 & 24\\
+\cmidrule{1-9}\pagebreak[0]
+ & Maximum &  & 240 & 90 & 180 & 153 & 120 & 150\\
 \nopagebreak
- & Mean &  & 28 & 24 & 20 & 19 & 25\\
+ & Mean &  & 28 & 24 & 20 & 19 & 25 & 40\\
 \nopagebreak
- & Median &  & 15 & 15 & 15 & 10 & 20\\
+ & Median &  & 15 & 15 & 15 & 10 & 20 & 30\\
 \nopagebreak
- & Minimum &  & 5 & 2 & 1 & 1 & 5\\
+ & Minimum &  & 5 & 2 & 1 & 1 & 5 & 5\\
 \nopagebreak
-\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
+\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20 & 20\\
 \bottomrule
@@
 Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
-walking trips destined for \texttt{home} had the highest medians.
-However, by 1998, the highest medians shifted to trips to
-\texttt{work\ or\ school}, a transition that also occurred for cycling
-trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
-median duration for walking trips to \texttt{home} and
-\texttt{work\ or\ school} remained at 10 minutes.
+walking trips destined for home had the highest medians. However, by
+1998, the highest medians shifted to trips to work or school, a
+transition that also occurred for cycling trips between 1992 and 1998.
+Table \ref{tab:table-03} indicates that the median duration for walking
+trips to home and work or school remained at 10 minutes.
 
@@
 
-In 1992, walking trips with \texttt{home} as both the origin and
-destination made up the majority, accounting for about 30\% of all
-walking trips. These trips often involved leisure activities, like short
-walks or dog walking. Following this, trips from \texttt{home} to
-\texttt{work\ or\ school} comprised 18\% of walking trips. Overall,
-\texttt{home} emerged as a crucial hub, either as an origin or
-destination, with only 5\% of trips not involving \texttt{home.} By
-2015, \texttt{home} remained a significant node, but new locations
-distributed the proportion of trips to areas not considered in 1992. In
-2015, the highest proportion of trips were from \texttt{home} to
-\texttt{work\ or\ school} (12\%) and vice versa (11\%). \texttt{home} to
-\texttt{home} accounted for 8\% of trips, and \texttt{grocery\ stores}
-became a notable destination for those leaving \texttt{home} (6\%),
-surpassing trips to \texttt{other\textquotesingle{}s\ home} (4\%).
+In 1992, walking trips with home as both the origin and destination made
+up the majority, accounting for about 30\% of all walking trips. These
+trips often involved leisure activities, like short walks or dog
+walking. Following this, trips from home to work or school comprised
+18\% of walking trips. Overall, home emerged as a crucial hub, either as
+an origin or destination, with only 5\% of trips not involving home. By
+2015, home remained a significant node, but new locations distributed
+the proportion of trips to areas not considered in 1992. In 2015, the
+highest proportion of trips were from home to work or school (12\%) and
+vice versa (11\%). home to home accounted for 8\% of trips, and grocery
+stores became a notable destination for those leaving home (6\%),
+surpassing trips to other's home (4\%).
 
@@
 this mode of transportation was first included as an activity, the
-majority of trips were from \texttt{home} to \texttt{work\ or\ school},
-accounting for about 25\% of cases. This pattern remained in 2015, with
-these trips representing 30\% of the cases. However, a notable change
-occurred in \texttt{home} to \texttt{home} trips, which decreased
-significantly from 19\% in 1992 to 5\% in 2015.
+majority of trips were from home to work or school, accounting for about
+25\% of cases. This pattern remained in 2015, with these trips
+representing 30\% of the cases. However, a notable change occurred in
+home to home trips, which decreased significantly from 19\% in 1992 to
+5\% in 2015.
 
@@
 (2005), 24 (2010), and 29 (2015) of the GSS surveys on active travel in
-Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
-developed after collecting, cleaning, and processing the survey data,
-providing information on origins, destinations, and duration of active
-travel, as well other information.
+Canada. In the form of an R data package, \{ActiveCA\} was developed
+after collecting, cleaning, and processing the survey data, providing
+information on origins, destinations, and duration of active travel, as
+well other information.
 
@@
 estimated that they will only be published in the later part of 2025.
-The \texttt{R} package will be updated once these files become
-available.
+The R package will be updated once these files become available.
 
@@
 ease of use, which facilitates the addition of complementary data sets
-in the future. \texttt{R} users can seamlessly explore GSS walking and
-cycling episodes, with the option to suggest enhancements to the package
-as needed. This article adopts the structure proposed by Anastasia and
-Páez \citeyearpar{soukhov2023}, whose work provided essential guidance
-for the creation of this package. Similarly, we aim to contribute to the
+in the future. R users can seamlessly explore GSS walking and cycling
+episodes, with the option to suggest enhancements to the package as
+needed. This article adopts the structure proposed by Anastasia and Páez
+\citeyearpar{soukhov2023}, whose work provided essential guidance for
+the creation of this package. Similarly, we aim to contribute to the
 academic community by promoting transparent research practices that

### 2025-06-01 paper/older-format-version/Manuscript-Data-Package.Rmd
Create a document with all manuscript changes and add answer letters for the reviewers.

--- 2025-06-01 paper/Manuscript-Data-Package2.Rmd
@@
   "This paper describes {ActiveCA}, an open data product with Canadian time use data.
-{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
+{ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
 Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
@@
 
-{ActiveCA} is distributed as an R package with a number of data 
-objects and their documentation. R packages contain code, data, and documentation
-in a standardized format that can be installed by R users via a software repository,
+{ActiveCA} is distributed as an `R` package with a number of data 
+objects and their documentation. `R` packages contain code, data, and documentation
+in a standardized format that can be installed by `R` users via a software repository,
 such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29).
-Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (PUMFID), the person's weight (WGHT_PER), the month the survey data was collected (SURVMNTH), the respondent's age group (AGEGR10), the respondent's sex (SEX), and the respondent's marital status (MARSTAT).
+Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).
 
@@
 For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label.
-For instance, in the case of the variable SURVMNTH, a value of 1 means January 2016, 2 means February 2016, 3 corresponds to March 2016, and so on.
+For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.
 
@@
 
-Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number 10041 and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
-Each row in the table corresponds to an episode associated with the specified record identification (PUMFID = 10041), including the episode's weight (WGHT_EPI), episode number (EPINO), activity code (TUI_01), episode duration (DURATION), and episode location (LOCATION).
+Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 Episode File (Cycle 29).
+Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).
 
@@
 
-In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (TUI_01 = 1 and LOCATION = 300) for 210 minutes, followed by 40 minutes of personal hygiene (TUI_01 = 2).
-The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by TUI_01 = 27.
-Next, they recorded a travel episode, walking for 15 minutes (TUI_01 = 7 and LOCATION = 315), where both the origin and destination were their home.
+In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`).
+The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`.
+Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where both the origin and destination were their home.
 Such trips, where the journey starts and finishes at home, can be classified as recreational or leisure trips.
-Next, the respondent spent 3 hours searching for a job (TUI_01 = 9), took a 15-minute lunch break (TUI_01 = 6), and then cleaned the house (TUI_01 = 18) for two hours.
+Next, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours.
 Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available.
@@
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.") |>
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") |>
   kable_styling(full_width = FALSE, 
@@
                  font_size = 8) |>
-  footnote(general = c("Legend: PUMFID: record identification.
-EPINO: episode number.
+  footnote(general = c("Legend: `PUMFID`: record identification.
+`EPINO`: episode number.
 WGHT_EPI: episode's weight.
@@
 
-The documentation of the package can be accessed by running ?ActiveCA in R.
+The documentation of the package can be accessed by running `?ActiveCA` in `R`.
 The documentation includes a description of the package, information about the authors, contributors, and maintainers, and an overview of all pre-processed data sets included in the package.
 
-Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (walking_2015.rda).
+Figure \ref{fig:walking_documentation} displays the documentation for the walking episodes from the 2015 GSS (`walking_2015.rda`).
 The documentation provides a brief description of the dataset, instructions on how to load the package, as well as details about the file format and descriptions of all variables (already labeled).
@@
 For instance, to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weight.
-In 2015, the weight variable is represented by WGHT_PER.
-The code below demonstrates how to obtain the percentage of people with active travel episodes by age group.
-It uses the dplyr package [@dplyr2023] to manipulate the data.
+In 2015, the weight variable is represented by `WGHT_PER.` The code below demonstrates how to obtain the percentage of people with active travel episodes by age group.
+It uses the `dplyr` package [@dplyr2023] to manipulate the data.
 
@@
 Then, it joins the 2015 episodes with the 2015 Main File.
-Note that in both operations, the code sums the person weight variable (WGHT_PER) to obtain the correct population values.
-After this, a new dataset, Active_percentage, is created by merging both previous datasets.
+Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values.
+After this, a new dataset, `Active_percentage`, is created by merging both previous datasets.
 The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100 and rounding with 2 decimal places.
@@
 Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing.
-Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number 10041 from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
 Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes.
@@
 Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
-walking trips destined for home had the highest medians. However, by
-1998, the highest medians shifted to trips to work or school, a
+walking trips destined for `home` had the highest medians. However, by
+1998, the highest medians shifted to trips to `work or school`, a
 transition that also occurred for cycling trips between 1992 and 1998.
 Table \ref{tab:table-03} indicates that the median duration for walking trips to
-home and work or school remained at 10 minutes.
+`home` and `work or school` remained at 10 minutes.
 
@@
 
-In 1992, walking trips with home as both the origin and destination
+In 1992, walking trips with `home` as both the origin and destination
 made up the majority, accounting for about 30% of all walking trips.
 These trips often involved leisure activities, like short walks or dog
-walking. Following this, trips from home to work or school comprised
-18% of walking trips. Overall, home emerged as a crucial hub, either
-as an origin or destination, with only 5% of trips not involving home.
-By 2015, home remained a significant node, but new locations
+walking. Following this, trips from `home` to `work or school` comprised
+18% of walking trips. Overall, `home` emerged as a crucial hub, either
+as an origin or destination, with only 5% of trips not involving `home.`
+By 2015, `home` remained a significant node, but new locations
 distributed the proportion of trips to areas not considered in 1992. In
-2015, the highest proportion of trips were from home to
-work or school (12%) and vice versa (11%). home to home accounted
-for 8% of trips, and grocery stores became a notable destination for
-those leaving home (6%), surpassing trips to other's home (4%).
+2015, the highest proportion of trips were from `home` to
+`work or school` (12%) and vice versa (11%). `home` to `home` accounted
+for 8% of trips, and `grocery stores` became a notable destination for
+those leaving `home` (6%), surpassing trips to `other's home` (4%).
 
@@
 this mode of transportation was first included as an activity, the
-majority of trips were from home to work or school, accounting for
+majority of trips were from `home` to `work or school`, accounting for
 about 25% of cases. This pattern remained in 2015, with these trips
 representing 30% of the cases. However, a notable change occurred in
-home to home trips, which decreased significantly from 19% in 1992
+`home` to `home` trips, which decreased significantly from 19% in 1992
 to 5% in 2015.
@@
 data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
-of the GSS surveys on active travel in Canada. In the form of an R data package, 
+of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
 {ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
@@
 at the moment it is estimated that they will only be published in the later part
-of 2025. The R package will be updated once these files become available.
+of 2025. The `R` package will be updated once these files become available.
 
 The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
-which facilitates the addition of complementary data sets in the future. R users
+which facilitates the addition of complementary data sets in the future. `R` users
 can seamlessly explore GSS walking and cycling episodes, 

### 2025-06-01 paper/older-format-version/Manuscript-Data-Package.tex
Create a document with all manuscript changes and add answer letters for the reviewers.

--- 2025-01-22 paper/Manuscript-Data-Package.tex
@@
-\documentclass[Royal,times,sageh]{sagej}
-
-\usepackage{moreverb,url,natbib, multirow, tabularx}
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% From pandoc table feature
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  Royal,
+  times]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs,array}
@@
 \makesavenoteenv{longtable}
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\begin{document}
-
-\setcitestyle{aysep={,}}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel},
+  pdfkeywords={Active; mobility; walking; cycling; travel time;
+time-use; General Social Survey},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 to Study Active Travel}
+\author{true \and true \and true}
+\date{}
 
-\runninghead{}
-
-\author{Anon1\affilnum{}, Anon2\affilnum{}, Anon3\affilnum{}}
-
-\affiliation{\affilnum{}{}}
-
+\begin{document}
+\maketitle
 \begin{abstract}
@@
 
-\keywords{Active; mobility; walking; cycling; travel time; time-use;
-General Social Survey}
-
-\maketitle
-
 \section{Introduction}\label{introduction}
@@
 process in which all stages of development follow open principles
-\citep{arribas-bel2021}. ODPs, while still open, differ from general
+(Arribas-Bel et al. 2021). ODPs, while still open, differ from general
 open data in their degree of ease of access, their heightened usability,
@@
 sleeping, to time spent working or at school. These surveys have proved
-valuable in investigations of mobility and quality of life
-\citep{spinneyTransport2009}, the relationship between active travel and
-transit use \citep{lachapelleLonger2016}, and travel behavior and time
-poverty \citep{kimFacing2024}, to name but a few examples.
+valuable in investigations of mobility and quality of life (Spinney,
+Scott, and Newbold 2009), the relationship between active travel and
+transit use (Lachapelle and Pinto 2016), and travel behavior and time
+poverty (Kim et al. 2024), to name but a few examples.
 
@@
 Statistics Canada offers Public Use Microdata Files and documentation
-for the GSS program \citep[see][]{statisticscanada2024}, accessing these
-files, and preparing them for analysis is not a straightforward matter,
-given their size and complexity. The process of extracting information
-of interest from the source files is time-consuming, tedious, and
-challenging and/or prone to error due to the expertise required to work
-with these files. To create \{ActiveCA\} we collected, cleaned, and
-processed the cycles of the GSS surveys concerning time use to make them
-ready for analysis.
+for the GSS program (see Canada 2024), accessing these files, and
+preparing them for analysis is not a straightforward matter, given their
+size and complexity. The process of extracting information of interest
+from the source files is time-consuming, tedious, and challenging and/or
+prone to error due to the expertise required to work with these files.
+To create \{ActiveCA\} we collected, cleaned, and processed the cycles
+of the GSS surveys concerning time use to make them ready for analysis.
 
@@
 
-Given the level of interest in active travel
-\citep[e.g.,][]{mccurdySupport2023}, reducing the barriers to using data
-contained in rich, but difficult to access and use surveys, such as
-Canada's GSSs, is a worthy endeavour that can only improve data-driven
-decisions in transportation, urban, and health policy.
+Given the level of interest in active travel (e.g., McCurdy et al.
+2023), reducing the barriers to using data contained in rich, but
+difficult to access and use surveys, such as Canada's GSSs, is a worthy
+endeavour that can only improve data-driven decisions in transportation,
+urban, and health policy.
 
@@
 reproducible research project. In summary, \{ActiveCA\} aims to
-implement and inspire the best principles of open spatial sciences
-\citep{paez_open_2021, brunsdon_opening_2021}.
+implement and inspire the best principles of open spatial sciences (Páez
+2021; Brunsdon and Comber 2021).
 
@@
 
-Statistics Canada \citeyearpar{statisticscanada2024} conducts GSS
-surveys to obtain data on social trends to track changes in Canadians'
-living conditions and well-being over time. The series of survey on time
-use are used to understand how Canadian residents spend and manage their
-time, and what factors contribute to their happiness and stress. The GSS
-program was created in 1985, and is serialized to provide a collection
-of annual, representative cross-sectional surveys.
+Statistics Canada (2024) conducts GSS surveys to obtain data on social
+trends to track changes in Canadians' living conditions and well-being
+over time. The series of survey on time use are used to understand how
+Canadian residents spend and manage their time, and what factors
+contribute to their happiness and stress. The GSS program was created in
+1985, and is serialized to provide a collection of annual,
+representative cross-sectional surveys.
 
@@
 program was conducted in 1986, and the most recent was completed in
-2015. These Time Use Surveys \citep{statisticscanada2022} collect data
-on respondents' participation and time spent on a wide range of everyday
-activities using a 24-hour retrospective diary, with information on the
-location of these activities (e.g.~at home, at work, etc.) and, for
-non-personal activities, the people who were present with the respondent
-at the time of the activity. In addition, time-use surveys also cover
-topics related to leisure time, work-life balance, health, commuting,
-culture and sports, and many others.
+2015. These Time Use Surveys (Canada 2022) collect data on respondents'
+participation and time spent on a wide range of everyday activities
+using a 24-hour retrospective diary, with information on the location of
+these activities (e.g.~at home, at work, etc.) and, for non-personal
+activities, the people who were present with the respondent at the time
+of the activity. In addition, time-use surveys also cover topics related
+to leisure time, work-life balance, health, commuting, culture and
+sports, and many others.
 
@@
 
-\begin{verbatim}
-if(!require("ActiveCA", character.only = TRUE)) {
-      remotes::install_github("dias-bruno/ActiveCA")
-}
-\end{verbatim}
+\begin{Shaded}
+\begin{Highlighting}[]
+\ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{require}\NormalTok{(}\StringTok{"ActiveCA"}\NormalTok{, }\AttributeTok{character.only =} \ConstantTok{TRUE}\NormalTok{)) \{}
+\NormalTok{      remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"dias{-}bruno/ActiveCA"}\NormalTok{)}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{verbatim}
-library(ActiveCA)
-\end{verbatim}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(ActiveCA)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{verbatim}
-data(walking_2015)
-\end{verbatim}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{data}\NormalTok{(walking\_2015)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 percentage of people with active travel episodes by age group. It uses
-the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.
+the \texttt{dplyr} package (Wickham et al. 2023) to manipulate the data.
 
@@
 covers all Canadian time use surveys up to 2015. While the most recent
-time use survey was carried out in 2022 \citep{wray2024}, the Public Use
+time use survey was carried out in 2022 (Wray 2024), the Public Use
 Microdata Files are currently unavailable, and at the moment it is
@@
 as needed. This article adopts the structure proposed by Anastasia and
-Páez \citeyearpar{soukhov2023}, whose work provided essential guidance
-for the creation of this package. Similarly, we aim to contribute to the
-academic community by promoting transparent research practices that
-encourage replication and innovation in related fields. We believe that
-\{ActiveCA\} will serve as a basis for further research on GSS and for
-the integration of additional data by the authors or the wider open
-source community.
+Páez (2023), whose work provided essential guidance for the creation of
+this package. Similarly, we aim to contribute to the academic community
+by promoting transparent research practices that encourage replication
+and innovation in related fields. We believe that \{ActiveCA\} will
+serve as a basis for further research on GSS and for the integration of
+additional data by the authors or the wider open source community.
 
@@
 
-\bibliographystyle{sageh}
-\bibliography{bibfile.bib}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-arribas-bel2021}
+Arribas-Bel, Dani, Mark Green, Francisco Rowe, and Alex Singleton. 2021.
+{``Open Data Products-A Framework for Creating Valuable Analysis Ready
+Data.''} \emph{Journal of Geographical Systems} 23 (4): 497--514.
+\url{https://doi.org/10.1007/s10109-021-00363-5}.
+
+\bibitem[\citeproctext]{ref-brunsdon_opening_2021}
+Brunsdon, Chris, and Alexis Comber. 2021. {``Opening Practice:
+Supporting Reproducibility and Critical Spatial Data Science.''}
+\emph{Journal of Geographical Systems} 23 (4): 477--96.
+\url{https://doi.org/10.1007/s10109-020-00334-2}.
+
+\bibitem[\citeproctext]{ref-statisticscanada2022}
+Canada, Statistics. 2022. {``Time Use Survey.''}
+\url{https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&amp;SDDS=4503}.
+
+\bibitem[\citeproctext]{ref-statisticscanada2024}
+---------. 2024. {``Statistics Canada: Canada's National Statistical
+Agency.''} \url{https://www.statcan.gc.ca/en/start}.
+
+\bibitem[\citeproctext]{ref-kimFacing2024}
+Kim, Sang-O, Matthew Palm, Soojung Han, and Nicholas J. Klein. 2024.
+{``Facing a Time Crunch: {Time} Poverty and Travel Behaviour in
+{Canada}.''} \emph{Transportation Research Part D: Transport and
+Environment} 126 (January): 104028.
+\url{https://doi.org/10.1016/j.trd.2023.104028}.
+
+\bibitem[\citeproctext]{ref-lachapelleLonger2016}
+Lachapelle, Ugo, and Diogo Gianini Pinto. 2016. {``Longer or More
+Frequent Walks: {Examining} the Relationship Between Transit Use and
+Active Transportation in {Canada}.''} \emph{Journal of Transport \&
+Health}, Special {Issue}: {Public} {Transport} and {Health}, 3 (2):
+173--80. \url{https://doi.org/10.1016/j.jth.2016.02.005}.
+
+\bibitem[\citeproctext]{ref-mccurdySupport2023}
+McCurdy, Ashley, Guy Faulkner, Christine Cameron, Christa
+Costas-Bradstreet, and John C. Spence. 2023. {``Support for {Active}
+{Transport} {Policy} {Initiatives} {Among} {Canadian} {Adults}: {The}
+{Canadian} {National} {Active} {Transportation} {Survey}.''}
+\emph{Active Travel Studies} 3 (2).
+\url{https://doi.org/10.16997/ats.1450}.
+
+\bibitem[\citeproctext]{ref-paez_open_2021}
+Páez, Antonio. 2021. {``Open Spatial Sciences: An Introduction.''}
+\emph{Journal of Geographical Systems} 23 (4): 467--76.
+\url{https://doi.org/10.1007/s10109-021-00364-4}.
+
+\bibitem[\citeproctext]{ref-soukhov2023}
+Soukhov, Anastasia, and Antonio Páez. 2023. {``TTS2016R: A Data Set to
+Study Population and Employment Patterns from the 2016 Transportation
+Tomorrow Survey in the Greater Golden Horseshoe Area, Ontario,
+Canada.''} \emph{Environment and Planning B: Urban Analytics and City
+Science} 50 (2): 556--63.
+\url{https://doi.org/10.1177/23998083221146781}.
+
+\bibitem[\citeproctext]{ref-spinneyTransport2009}
+Spinney, J. E. L., D. M. Scott, and K. B. Newbold. 2009. {``Transport
+Mobility Benefits and Quality of Life: {A} Time-Use Perspective of
+Elderly {Canadians}.''} \emph{Transport Policy} 16 (1): 1--11.
+\url{https://doi.org/10.1016/j.tranpol.2009.01.002}.
+
+\bibitem[\citeproctext]{ref-dplyr2023}
+Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
+Vaughan. 2023. \emph{Dplyr: A Grammar of Data Manipulation}.
+\url{https://dplyr.tidyverse.org}.
+
+\bibitem[\citeproctext]{ref-wray2024}
+Wray, Dana. 2024. {``Telework, Time Use, and Well-Being: Evidence from
+the 2022 Time Use Survey.''}
+\url{https://www150.statcan.gc.ca/n1/daily-quotidien/240605/dq240605a-eng.htm}.
+
+\end{CSLReferences}
 

### 2025-06-01 data-raw/03-preprocessing-main-files.Rmd
Updating package documentation and 2022 TUS GSS variables.

@@
 gss_m_2022 <- read.spss(paste0(here(), "/data-raw/source-files/Time-use-2022/gss_tu_main_2022_pumf_v1.sav"), to.data.frame = TRUE) # Original file
+
+gss_m_2022_b <- gss_m_2022
+gss_m_2022 <- gss_m_2022_b
 ```
@@
  "DMOB_FL", "DPAIN_FL","DUNK_FL", "DVIS_FL", "DDIS_FL", "DTYPER",  "DCLASS",  "TCS_110", "TCS_120", "TCS_130",
- "TCS_140", "TCS_150", "TCS_160", "TCS_170", "TCS_180", "TCS_190", "TCS_200",  "GEN_01",  "GEN_02",  "SRS_10",  "MSS_130", "ACT7DAYC","MRW_05C", "MRW_D40A",
+ "TCS_140", "TCS_150", "TCS_160", "TCS_170", "TCS_180", "TCS_190", "TCS_200",  "TIMECR", "GEN_01",  "GEN_02",  "SRS_10",  "MSS_130", "ACT7DAYC","MRW_05C", "MRW_D40A",
 "MRW_D40B","WET_120",  "WHW_230", "WHWD140G","TLWK_01A","TLWK_06","WLYD170G",
@@
          WGHT_PER = as.numeric(as.character(WGHT_PER)))
+```
+
+```{r LUC_RST-2022}
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate( # LUC_RST:  Population centre indicator
+    LUC_RST = 
+      case_when(LUC_RST == "Larger urban population centres (CMA/CA)" ~ "CMA/CA",
+                     LUC_RST == "Rural areas/small population centres (non-CMA/CA)" ~ "non CMA/CA",
+                     LUC_RST == "Prince Edward Island" ~ "Prince Edward Island"),
+    LUC_RST = factor(LUC_RST))
+
+summary(gss_main_2022$LUC_RST)
+```
+```{r HSDSIZEC-2022}
+values_ordered <- c("One person", 
+           "Two persons",
+           "Three persons",
+           "Four persons",
+           "Five persons or more")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# HSDSIZEC:  Household size of respondent (capped)
+    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$HSDSIZEC)
+```
+```{r AGEGR10-2022}
+values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# AGEGR10:  Age group of respondent (groups of 10)
+    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE), 
+    AGEPRC = factor(AGEPRC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$AGEGR10)
+```
+```{r AGEPRC-2022}
+values_ordered <- c("15 to 34 years","35 to 44 years",
+                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# AGEPRC:  Age group of respondent’s spouse/partner (collapsed)
+    AGEPRC = factor(AGEPRC, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$AGEPRC)
+```
+
+```{r CHH0017C-2022}
+summary(gss_main_2022$CHH0017C)
+
+values_ordered <- c("No children","One child",
+                    "Two children", "Three children or more")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# CHH0017C:  Number of child(ren) in household - 0 to 17 years (capped)
+    CHH0017C = factor(CHH0017C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$CHH0017C)
+```
+```{r CHH0004C-2022}
+summary(gss_main_2022$CHH0004C)
+
+values_ordered <- c("No children","One or more children")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# CHH0004C:  Number of child(ren) in household - 0 to 4 years (capped)
+    CHH0004C = factor(CHH0004C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$CHH0004C)
+```
+
+```{r CHH0514C-2022}
+summary(gss_main_2022$CHH0514C)
+
+values_ordered <- c("No children","One or more children")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# CHH0514C:  Child(ren) in household - 5 to 14 years (flag)
+    CHH0514C = factor(CHH0514C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$CHH0514C)
+```
+
+```{r CHH1517C-2022}
+summary(gss_main_2022$CHH1517C)
+
+values_ordered <- c("No children","One or more children")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# CHH1517C:  Child(ren) in household - 15 to 17 years (flag)
+    CHH1517C = factor(CHH1517C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$CHH1517C)
+```
+
+```{r PARNUM-2022}
+summary(gss_main_2022$PARNUM)
+
+values_ordered <- c("None","One parent","Two parents")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# PARNUM: Number of parents the respondent has in household
+    PARNUM = factor(PARNUM, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$PARNUM)
+```
+```{r DTYPER-2022}
+summary(gss_main_2022$DTYPER)
+
+values_ordered <- c("Does not have a disability",
+                    "Has one disability type",
+                    "Has two or three disability types",
+                    "Has more than three disability types")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# DTYPER: Disability type counter - Grouped
+    DTYPER = factor(DTYPER, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$DTYPER)
+```
+
+```{r DCLASS-2022}
+summary(gss_main_2022$DCLASS)
+
+values_ordered <- c("Does not have a disability",
+                    "Unknown severity",
+                    "Mild",
+                    "Moderate",
+                    "Severe",
+                    "Very severe")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# DCLASS: Global (disability) severity class
+    DCLASS = factor(DCLASS, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$DCLASS)
+```
+
+```{r TIMECR-2022}
+summary(gss_main_2022$TIMECR)
+
+values_ordered <- c("Zero Yes codes",
+                    "One Yes code",
+                    "Two Yes codes",
+                    "Three Yes codes",
+                    "Four Yes codes",
+                    "Five Yes codes",
+                    "Six Yes codes",
+                    "Seven Yes codes",
+                    "Eight Yes codes",
+                    "Nine Yes codes",
+                    "Ten Yes codes")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# TIMECR: Time crunch indicator
+    TIMECR = factor(TIMECR, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$TIMECR)
+```
+
+```{r GEN_01-2022}
+summary(gss_main_2022$GEN_01)
+
+values_ordered <- c("Poor",
+                    "Fair",
+                    "Good",
+                    "Very good",
+                    "Excellent")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# GEN_01: Self rated health - In general, how is your health?
+    GEN_01 = factor(GEN_01, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$GEN_01)
+```
+
+```{r GEN_02-2022}
+summary(gss_main_2022$GEN_02)
+
+values_ordered <- c("Poor",
+                    "Fair",
+                    "Good",
+                    "Very good",
+                    "Excellent")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# GEN_02: Self rated mental health - In general, how is your mental health?
+    GEN_02 = factor(GEN_02, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$GEN_02)
+```
+
+```{r SRS_10-2022}
+summary(gss_main_2022$SRS_10)
+
+values_ordered <- c("Not at all stressful",
+                    "Not very stressful",
+                    "A bit stressful",
+                    "Quite a bit stressful",
+                    "Extremely stressful")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# SRS_10: Self Rated Stress - Amount - Thinking about the amount of stress in your life, how would you describe most of your days?
+    SRS_10 = factor(SRS_10, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$SRS_10)
+```
+
+```{r TLWK_06-2022}
+summary(gss_main_2022$TLWK_06)
+
+values_ordered <- c("1 - Not important at all",
+                    "2",
+                    "3",
+                    "4",
+                    "5 - Very important")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# TLWK_06: Telework aspects - Avoid the commute. How important is the opportunity not to commute every day to your workplace?
+    TLWK_06 = factor(TLWK_06, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$TLWK_06)
+```
+
+```{r WLYD170G-2022}
+summary(gss_main_2022$WLYD170G)
+
+values_ordered <- c("Respondent works from home",
+                    "0 to 4 kilometers",
+                    "5 to 9 kilometers",
+                    "10 to 14 kilometers",
+                    "15 to 19 kilometers",
+                    "20 to 24 kilometers",
+                    "25 to 29 kilometers",
+                    "30 to 39 kilometers",
+                    "40 to 49 kilometers",
+                    "50 kilometers or more")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# WLYD170G: Distance from residence to workplace (grouped)
+    WLYD170G = factor(WLYD170G, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$WLYD170G)
+```
+
+```{r ATT_120-2022}
+summary(gss_main_2022$ATT_120)
+
+values_ordered <- c("All the time",
+                    "Some of the time",
+                    "Rarely",
+                    "Never")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# ATT_120: Vehicle at respondent’s disposal. How often do you have a vehicle at your disposal?
+    ATT_120 = factor(ATT_120, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$ATT_120)
+```
+
+```{r ATT_130-2022}
+summary(gss_main_2022$ATT_130)
+
+values_ordered <- c("All the time",
+                    "Some of the time",
+                    "Rarely",
+                    "Never")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# ATT_130: Need someone else to drive - Frequency. How often do you need someone to drive you somewhere for an appointment, errand or something else?
+    ATT_130 = factor(ATT_130, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$ATT_130)
+```
+
+```{r ATT_131-2022}
+summary(gss_main_2022$ATT_131)
+
+values_ordered <- c("All the time",
+                    "Some of the time",
+                    "Rarely",
+                    "Never")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# ATT_131: Someone else can drive when needed - Frequency. How often can someone drive you when you need it?
+    ATT_131 = factor(ATT_131, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$ATT_131)
+```
+```{r ATT_170-2022}
+summary(gss_main_2022$ATT_170)
+
+values_ordered <- c("Not in past 12 months",
+                    "A few times a year",
+                    "A few times a month",
+                    "Several times a week",
+                    "Every day")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# ATT_170: Use of public transportation - Past 12 months.
+    ATT_170 = factor(ATT_170, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$ATT_170)
+```
+
+```{r ED_05-2022}
+summary(gss_main_2022$ED_05)
+
+values_ordered <- c("Less than high school diploma or its equivalent",
+                    "High school diploma or a high school equivalency certificate",
+                    "Trades certificate or diploma",
+                    "College, CEGEP or other non-university certificate or diplom",
+                    "University certificate or diploma below the bachelor's level",
+                    "Bachelor's degree",
+                    "University certificate, diploma, or degree above the bachelo")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# ED_05: Educational attainment - Highest degree (7 categories)
+    ED_05 = factor(ED_05, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$ED_05)
+```
+
+```{r REL_02-2022}
+summary(gss_main_2022$REL_02)
+
+values_ordered <- c("Not at all",
+                    "Once or twice a year",
+                    "At least three times a year",
+                    "At least once a month",
+                    "At least once a week")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# REL_02: Frequency of religious participation - Past 12 months.
+    REL_02 = factor(REL_02, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$REL_02)
+```
+```{r INC_C-2022}
+summary(gss_main_2022$INC_C)
+
+values_ordered <- c("Less than $25,000",
+                    "$25,000 to $49,999",
+                    "$50,000 to $74,999",
+                    "$75,000 to $99,999",
+                    "$100,000 and over")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# INC_C: Income - Total (before tax) (grouped).
+    INC_C = factor(INC_C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$INC_C)
+```
+
+```{r FAMINC_C-2022}
+summary(gss_main_2022$FAMINC_C)
+
+values_ordered <- c("Less than $50,000",
+                    "$50,000 to $74,999",
+                    "$75,000 to $99,999",
+                    "$100,000 to $124,999",
+                    "$125,000 and over")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate(# FAMINC_C: Family Income - Total (before tax) (grouped).
+    FAMINC_C = factor(FAMINC_C, levels = values_ordered, ordered = TRUE))
+
+summary(gss_main_2022$FAMINC_C)
+```
+```{r GTU_110-2022}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate( # GTU_110: How often do you feel rushed? Would you say it is...?
+    GTU_110 = factor(GTU_110, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2022$GTU_110)
+```
+
+```{r GTU_130-2022}
+values_ordered <- c("Every day",
+                    "A few times a week",
+                    "About once a week",
+                    "About once a month",
+                    "Less than once a month",
+                    "Never")
+
+gss_main_2022 <- gss_main_2022 %>% 
+  mutate( # GTU_130: How often do you feel you have time on your hands that you don’t know what to do with?
+    GTU_130 = factor(GTU_130, levels = rev(values_ordered), ordered = TRUE))
+
+summary(gss_main_2022$GTU_130)
 ```

### 2025-06-01 R/ActiveCA-package.R

@@ -191,7 +191,7 @@ NULL
 
 #' Preprocessed walking episodes (2015).
 #'
-#' This dataset contains all the walking episodes from the 2010 GSS Survey.
+#' This dataset contains all the walking episodes from the 2015 GSS Survey.
 #' We identified the walking episodes and selected the places of origin
 #' and destination, measuring the time spent, weight and other
 #' information about the activities.
@@ -477,14 +477,14 @@ NULL
 #' @seealso 2022 Time Use. Public Use Microdata File [Documentation and User’s Guide](https://www23.statcan.gc.ca/imdb/p3Instr.pl?Function=assembleInstr&lang=en&Item_Id=1395477)
 "cycling_2022"
 
-#' Preprocessed walking and cycling episodes (from 1986 to 2015).
+#' Preprocessed walking and cycling episodes (from 1986 to 2022).
 #'
-#' This dataset contains all the walking and cycling episodes from the 1986 to 2015 GSS Surveys.
+#' This dataset contains all the walking and cycling episodes from the 1986 to 2022 GSS Surveys.
 #' We identified the cycling episodes and selected the places of origin
 #' and destination, measuring the time spent, weight and other
 #' information about the activities.
 #'
-#' @format A data frame containing 21748 rows and 17 variables; each row represents a unique episode.
+#' @format A data frame containing 23513 rows and 17 variables; each row represents a unique episode.
 #'   \describe{
 #'   \item{PUMFID}{Public Use Microdata File Identifier. This variable serves as the link between the Main and the Episode files. Each case has a unique record identifier and this appears on every episode}
 #'   \item{WGHT_EPI}{Episode weight.}
@@ -511,16 +511,16 @@ NULL
 #' @source [General Social Survey (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) / [Odesi](http://odesi2.scholarsportal.info/webview/)
 "gss_episodes"
 
-#' Impedance functions for Census Agglomerations and Census Metropolitan Areas (from 1986 to 2015).
+#' Impedance functions for Census Agglomerations and Census Metropolitan Areas (from 1986 to 2022).
 #'
 #' This dataset contains all impedance functions for walking
-#' and cycling transportation, obtained from the 1986 to 2015
+#' and cycling transportation, obtained from the 1986 to 2022
 #' GSS Surveys, for Census Metropolitan Areas and Census Agglomerations.
[…]

### 2025-08-14 paper/EPB-Final-Submission.Rmd
Final version for EPB publication

--- 2025-06-01 paper/Manuscript-Data-Package-1Round-Review.Rmd
@@
 title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
+runninghead: Dos Santos \emph{et al}.
 author:
-  - name: "Anon1"
-    email: "Anon1@anon.org"
-    affiliation: CODE
-  - name: "Anon2"
-    email: "Anon2@anon.org"
-    affiliation: CODE
-  - name: "Anon3"
-    email: "Anon3@anon.org"
-    affiliation: CODE
+- name: Bruno Dias dos Santos*
+  num: 1
+- name: Mahdis Moghadasi
+  num: 1 
+- name: Antonio Páez
+  num: 1
 address:
-  - code: CODE
-    address: "Some school"
-footnote:
-  - code: 1
-    text: "Corresponding Author"
+- num: 1
+  org: McMaster University
+corrauth: "Bruno Dias dos Santos - School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
+email: dossanb@mcmaster.ca
 abstract: 
   "This paper describes {ActiveCA}, an open data product with Canadian time use data.
-{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS).
-Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
+{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) from the General Social Survey (GSS).
+Active travel episodes are characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
 The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates.
@@
 The geographical resolution includes the province and whether the episode was in an urban or rural setting."
-
-keywords: Active; mobility; walking; cycling; travel time; time-use; General Social Survey 
+link-citations: yes
+linkcolor: blue
+keywords: Active; mobility; walking; cycling; travel time; time-use; 
 classoption:
@@
 Canadians. As part of this program, every five to seven years the survey
-is done on the topic of time use (TUS). Concretely, {ActiveCA} covers
+is done on the topic of time use. Concretely, {ActiveCA} covers
 Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
@@
 
-Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
+Using Public Use Microdata Files (PUMFs) from the TUS,
 we extracted all data necessary to characterize active travel in
@@
 
-{ActiveCA} is distributed as an `R` package with a number of data objects
-and their documentation. `R` packages contain code, data, and
+{ActiveCA} is distributed as an R package with a number of data objects
+and their documentation. R packages contain code, data, and
 documentation in a standardized format that can be installed by R users
@@
 The rest of this paper discusses the sources of data, and the process
-implemented to retrieve and package them. Then, we explain how to use
-the package and show some selected examples of analysis to whet the
+implemented to retrieve and package them. Then, we show some selected examples of analysis to whet the
 imagination of potential users. This ODP provides not only data that are
@@
 limited set of key questions, constrained by sample size. Importantly,
-none of the variables used in this research are `R` package in the 2022
+none of the variables used in this research are in the 2022
 PUMF User Guide as unsuitable for trend analysis.
@@
 representative sample. For sampling, the ten provinces were divided into
-geographic strata. Several Census Metropolitan Areas (CMAs) - such as
+geographic stratas. Several Census Metropolitan Areas (CMAs) - such as
 St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
@@
 The Public Use Microdata are released by Statistics Canada in two files:
-a *Main file* and an *Episode file*. The files are linked by keys that
+a main file and an episode file. The files are linked by keys that
 identify households, individuals, and episodes (i.e., activities)
@@
 
-The *Main file* of the time use survey compiles a large array of
+The main file of the time use survey compiles a large array of
 aggregated data, summarizing the answers to the questionnaire that
@@
 duration that respondents allocate to each activity and location. The
-*Main file* provides a overview of daily routines and social dynamics,
+main file provides a overview of daily routines and social dynamics,
 not focusing on individual activity episodes. Additionally, this file
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
-six variables of the TUS PUMF 2015 *Main file* (Cycle 29). Each row in
+six variables of the TUS PUMF 2015 main file (Cycle 29). Each row in
 the table correspond to a survey respondent, while the columns refer the
@@
 
-The *Main file* of the 2015 GSS surveys includes a total of 17,390
+The main file of the 2015 GSS surveys includes a total of 17,390
 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and
@@
     booktabs = TRUE, 
-    longtable = TRUE, 
+    longtable = FALSE, 
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
@@
 social context, informing when and where activities occurred and with
-whom. The focus of the *Episode file* is not on the characteristics of
+whom. The focus of the episode file is not on the characteristics of
 the respondents, but on the characteristics of the activities, and the
@@
 a day of the respondent. Although respondent-specific characteristics
-are not included within the *Episode file*, it is possible to link the
-*Main file* and the *Episode file* by using a key present in both the
-*Main* and *Episode files*.
+are not included within the episode file, it is possible to link the
+main file and the episode file by using a key present in both the
+main and episodes files.
 
 Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
-example of the *Main file* structure, Table
+example of the main file structure, Table
 \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
 record identification number 10041 and some variables from the TUS PUMF
-2015 *Episode file* (Cycle 29). Each row in the table corresponds to an
+2015 episode file (Cycle 29). Each row in the table corresponds to an
 episode associated with the specified record identification (`PUMFID` =
@@
 
-In total, the *Episode file* of the 2015 GSS surveys contains 274,108
+In total, the episode file of the 2015 GSS surveys contains 274,108
 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527
-variables detailing the episodes. Similar to the *Main file*, Statistics
+variables detailing the episodes. Similar to the main file, Statistics
 Canada has created codes for the discrete variables, with each value
@@
     booktabs = TRUE, 
-    longtable = TRUE, 
+    longtable = FALSE, 
     align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
-    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.") |>
+    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the  first seven episodes of the record number 10041.") |>
   kable_styling(full_width = FALSE, 
@@
 Figure \ref{fig:process-figure} presents a diagram illustrating the
-processes applied to the *Main* and *Episode files* to create the
-{ActiveCA} datasets. For each TUS cycle, we reviewed the *Episode files*
+processes applied to the main and episode files to create the
+{ActiveCA} datasets. For each TUS cycle, we reviewed the episode files
 to identify movement episodes involving walking or cycling. This allowed
@@
 
-For the *Main files*, we selected socioeconomic variables to help
+For the main files, we selected socioeconomic variables to help
 profile individuals engaged in active travel episodes. These included
 key indicators such as age group, sex, marital status, and number of
-children, among others. As with the *Episode files*, we labeled and
+children, among others. As with the episode files, we labeled and
 factored the coded variables. Using the appropriate identifiers, we
-joined the Episode and *Main files* to determine the province and the
+joined the episode and main files to determine the province and the
 type of population centre - whether it is a CMA, a Census Agglomeration
@@
 
-This section presents some potential applications of the {ActiveCA} `R` package.
+This section presents some potential applications of the {ActiveCA} R package.
 In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.
-The installation instruction and also some examples of application of the {ActiveCA} `R` package are available in the `vignettes`, available in the Github repository.
+The installation instruction and also some examples of application of the {ActiveCA} R package are available in the vignettes, available in the Github repository.
 
@@
 Table \ref{tab:processed-obs} displays the total number of records
-processed for Main and *Episode files*. For the *Main files*, a total of
+processed for Main and episode files. For the main files, a total of
 `r sum(df_main_processed$Count)` records were processed, referring to
@@
   select(-c(Survey_ep, Survey, Year_ep)) %>% 
-  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
+  kable(format = "latex", booktabs = TRUE, longtable = FALSE, 
         caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
@@
 \ref{tab:ep-2015-processed} presents the walking episodes for the record
-identification number `10041` from the TUS PUMF 2015 *Episode file* (Cycle
+identification number `10041` from the TUS PUMF 2015 episode file (Cycle
 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only
-the unique active travel episode appear in Table
+the unique active travel episode appears in Table
 \ref{tab:ep-2015-processed} since the records were filtered to select
@@
     booktabs = TRUE, 
-    longtable = TRUE, 
+    longtable = FALSE, 
     caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.", 
@@
 styled_ch03_table_01 <- stats_data |>
-  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
-        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
+  kable(format = "latex", booktabs = TRUE, longtable = FALSE, 
+        caption = "\\label{tab:table-01}Descriptive statistics of the duration (in minutes) of episodes with active transport records.",  
         align = c('l', 'l', rep('c', ncol(stats_data) - 2))) |>
@@
 
-In 2022, `home` served as a central hub for most trips, with fewer than
+In 2022, home location served as a central hub for most trips, with fewer than
 10% of journeys not involving it as either a starting point or
-destination. The most common trip types were from `home` to `work or school`
-(17%) and the reverse, from `work or school` to `home` (13%). Notably, 7% of
-trips began and ended at `home`, often reflecting leisure activities such
-as short walks or dog walking. `Grocery stores` were also a key
-destination, comprising 10% of trips departing from `home.`
+destination. The most common trip types were from home to work or school
+(17%) and the reverse, from work or school to home (13%). Notably, 7% of
+trips began and ended at home, often reflecting leisure activities such
+as short walks or dog walking. Grocery stores were also a key
+destination, comprising 10% of trips departing from home.
 
@@
     booktabs = TRUE, 
-    caption = "\\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.") |>
+    caption = "\\label{tab:cma-durations}Differences in walking duration (in minutes) between provinces and population centre type.") |>
   kable_styling(full_width = FALSE, 
@@
 
-{ActiveCA} also provides a Jupyter Notebook containing a `Python` script
-that demonstrates how to read `R` data files (.rda) and convert them into
-Pandas DataFrames. This process allows users to work with and utilize
-the datasets available in {ActiveCA} within a `Python` project.
+{ActiveCA} also provides a Jupyter Notebook containing a Python script
+that demonstrates how to read R data files (.rda) and convert them into
+Pandas Data Frames. This process allows users to work with and utilize
+the datasets available in {ActiveCA} within a Python project.
 
@@
 (2005), 24 (2010), 29 (2015), and 34 (2022) of TUS GSSs on active travel
-in Canada. In the form of an `R` data package, {ActiveCA} was developed
+in Canada. In the form of an R data package, {ActiveCA} was developed
 after collecting, cleaning, and processing the survey data, providing
@@
 
-The author(s) declared no potential conflicts of interest with respect
+The authors declared no potential conflicts of interest with respect
 to the research, authorship, and/or publication of this article.
+
+# Acknowledgments
+
+The authors thank the anonymous reviewers for their insightful comments.
 
@@
 
-The author(s) disclosed receipt of the following financial support for
-the research, authorship, and/or publication of this article: This work
-was supported by the Social Sciences and Humanities Research Council of
-Canada (*More description about the funding source after the review
-process*).
+The work has been supported by funding from the Social Sciences and Humanities Research Council’s Partnership Grant: Mobilizing justice: towards evidence-based transportation equity policy.
 
@@
 
-Author 1
+Bruno Dias dos Santos - https://orcid.org/0000-0002-6748-2038
 
-Author 2
+Mahdis Moghadasi - https://orcid.org/0000-0003-1009-9369
 
-Author 3
+Antonio Páez - https://orcid.org/0000-0001-6912-9919
 
@@
 
-The {ActiveCA} R data package can be found and installed on Github
-(*link*).
-
-For review purposes, the package is currently available as a tar.gz file
-that can be installed by R. The file can be obtained from this anonymous
-location:
-
-<https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip>
+The {ActiveCA} R data package can be found and installed on [Github](https://github.com/dias-bruno/ActiveCA).
 

### 2025-08-14 paper/EPB-Final-Submission.tex
Final version for EPB publication

--- 2025-06-01 paper/Manuscript-Data-Package-1Round-Review.tex
@@
 
-\runninghead{}
+\runninghead{Dos Santos \emph{et al}.}
 
-\author{Anon1\affilnum{}, Anon2\affilnum{}, Anon3\affilnum{}}
+\author{Bruno Dias dos Santos*\affilnum{1}, Mahdis
+Moghadasi\affilnum{1}, Antonio Páez\affilnum{1}}
 
-\affiliation{\affilnum{}{}}
+\affiliation{\affilnum{1}{McMaster University}}
+
+\corrauth{Bruno Dias dos Santos}
+
+\email{\href{mailto:dossanb@mcmaster.ca}{\nolinkurl{dossanb@mcmaster.ca}}}
 
@@
 analysis-ready data related to active travel spanning almost 40 years,
-extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
-(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles
-from the General Social Survey (GSS). Active travel is characterized by
+extracted from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
+(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) from the
+General Social Survey (GSS). Active travel episodes are characterized by
 mode, with walking being part of every cycle and bicycling starting in
@@
 
-\keywords{Active; mobility; walking; cycling; travel time; time-use;
-General Social Survey}
+\keywords{Active; mobility; walking; cycling; travel time; time-use;}
 
@@
 Canadians. As part of this program, every five to seven years the survey
-is done on the topic of time use (TUS). Concretely, \{ActiveCA\} covers
-Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
-and 34 (2022) of the TUS. Time use data in these surveys is coded using
-a very fine grain, from time spent in chores, leisure, and sleeping, to
-time spent working or at school. These surveys have proved valuable in
+is done on the topic of time use. Concretely, \{ActiveCA\} covers Cycles
+2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34
+(2022) of the TUS. Time use data in these surveys is coded using a very
+fine grain, from time spent in chores, leisure, and sleeping, to time
+spent working or at school. These surveys have proved valuable in
 investigations of mobility and quality of life
@@
 
-Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
-we extracted all data necessary to characterize active travel in Canada
-- specifically, episodes in which the activity involved moving between
-an origin and a destination by walking or cycling. Although Statistics
-Canada provides PUMFs and accompanying documentation for the GSS program
+Using Public Use Microdata Files (PUMFs) from the TUS, we extracted all
+data necessary to characterize active travel in Canada - specifically,
+episodes in which the activity involved moving between an origin and a
+destination by walking or cycling. Although Statistics Canada provides
+PUMFs and accompanying documentation for the GSS program
 \citep[see][]{statisticscanada2024}, accessing and preparing these files
@@
 
-\{ActiveCA\} is distributed as an \texttt{R} package with a number of
-data objects and their documentation. \texttt{R} packages contain code,
-data, and documentation in a standardized format that can be installed
-by R users via a software repository, such as CRAN (Comprehensive R
-Archive Network) or GitHub, which makes them an adroit medium to
-distribute analysis-ready data.
+\{ActiveCA\} is distributed as an R package with a number of data
+objects and their documentation. R packages contain code, data, and
+documentation in a standardized format that can be installed by R users
+via a software repository, such as CRAN (Comprehensive R Archive
+Network) or GitHub, which makes them an adroit medium to distribute
+analysis-ready data.
 
@@
 The rest of this paper discusses the sources of data, and the process
-implemented to retrieve and package them. Then, we explain how to use
-the package and show some selected examples of analysis to whet the
-imagination of potential users. This ODP provides not only data that are
-easy to use, but also all the code and documentation that make this a
-reproducible research project. In summary, \{ActiveCA\} aims to
-implement and inspire the best principles of open spatial sciences
-\citep{paez_open_2021, brunsdon_opening_2021}.
+implemented to retrieve and package them. Then, we show some selected
+examples of analysis to whet the imagination of potential users. This
+ODP provides not only data that are easy to use, but also all the code
+and documentation that make this a reproducible research project. In
+summary, \{ActiveCA\} aims to implement and inspire the best principles
+of open spatial sciences \citep{paez_open_2021, brunsdon_opening_2021}.
 
@@
 limited set of key questions, constrained by sample size. Importantly,
-none of the variables used in this research are \texttt{R} package in
-the 2022 PUMF User Guide as unsuitable for trend analysis.
+none of the variables used in this research are in the 2022 PUMF User
+Guide as unsuitable for trend analysis.
 
@@
 representative sample. For sampling, the ten provinces were divided into
-geographic strata. Several Census Metropolitan Areas (CMAs) - such as
+geographic stratas. Several Census Metropolitan Areas (CMAs) - such as
 St.~John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
@@
 The Public Use Microdata are released by Statistics Canada in two files:
-a \emph{Main file} and an \emph{Episode file}. The files are linked by
-keys that identify households, individuals, and episodes (i.e.,
-activities) conducted by individuals. We discuss these files in more
-detail in the following section.
+a main file and an episode file. The files are linked by keys that
+identify households, individuals, and episodes (i.e., activities)
+conducted by individuals. We discuss these files in more detail in the
+following section.
 
@@
 
-The \emph{Main file} of the time use survey compiles a large array of
+The main file of the time use survey compiles a large array of
 aggregated data, summarizing the answers to the questionnaire that
@@
 duration that respondents allocate to each activity and location. The
-\emph{Main file} provides a overview of daily routines and social
-dynamics, not focusing on individual activity episodes. Additionally,
-this file categorizes activities into bigger groups and subcategories,
+main file provides a overview of daily routines and social dynamics, not
+focusing on individual activity episodes. Additionally, this file
+categorizes activities into bigger groups and subcategories,
 facilitating the data's analytical utility with additional metrics such
@@
 Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
-six variables of the TUS PUMF 2015 \emph{Main file} (Cycle 29). Each row
-in the table correspond to a survey respondent, while the columns refer
-the following information: record identification (\texttt{PUMFID}), the
+six variables of the TUS PUMF 2015 main file (Cycle 29). Each row in the
+table correspond to a survey respondent, while the columns refer the
+following information: record identification (\texttt{PUMFID}), the
 person's weight (\texttt{WGHT\_PER}), the month the survey data was
@@
 
-The \emph{Main file} of the 2015 GSS surveys includes a total of 17,390
+The main file of the 2015 GSS surveys includes a total of 17,390
 respondents, representing 29,766,399 individuals and 848 variables. For
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item Legend: PUMFID: record identification.
-WGHT\_PER:  person weight.
-SURVMNTH: survey month of data collection.
-AGEGR10: age group of the respondent.
-SEX: sex of the respondent.
-MARSTAT: marital status of the respondent.
-\end{TableNotes}
-\begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.}\\
+\begin{table}
+\centering
+\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{8}{10}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{cccccc}
 \toprule
@@
 \bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
+\end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item Legend: PUMFID: record identification.
+WGHT\_PER: person weight.
+SURVMNTH: survey month of data collection.
+AGEGR10: age group of the respondent.
+SEX: sex of the respondent.
+MARSTAT: marital status of the respondent.
+\end{tablenotes}
+\end{threeparttable}}
+\end{table}
 
@@
 social context, informing when and where activities occurred and with
-whom. The focus of the \emph{Episode file} is not on the characteristics
-of the respondents, but on the characteristics of the activities, and
-the data are structured around the numerous activity instances that
-compose a day of the respondent. Although respondent-specific
-characteristics are not included within the \emph{Episode file}, it is
-possible to link the \emph{Main file} and the \emph{Episode file} by
-using a key present in both the \emph{Main} and \emph{Episode files}.
+whom. The focus of the episode file is not on the characteristics of the
+respondents, but on the characteristics of the activities, and the data
+are structured around the numerous activity instances that compose a day
+of the respondent. Although respondent-specific characteristics are not
+included within the episode file, it is possible to link the main file
+and the episode file by using a key present in both the main and
+episodes files.
 
 Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
-example of the \emph{Main file} structure, Table
-\ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
-record identification number 10041 and some variables from the TUS PUMF
-2015 \emph{Episode file} (Cycle 29). Each row in the table corresponds
-to an episode associated with the specified record identification
-(\texttt{PUMFID} = \texttt{10041}), including the episode's weight
-(\texttt{WGHT\_EPI}), episode number (\texttt{EPINO}), activity code
-(\texttt{TUI\_01}), episode duration (\texttt{DURATION}), and episode
-location (\texttt{LOCATION}).
+example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
+presents the first seven episodes for the record identification number
+10041 and some variables from the TUS PUMF 2015 episode file (Cycle 29).
+Each row in the table corresponds to an episode associated with the
+specified record identification (\texttt{PUMFID} = \texttt{10041}),
+including the episode's weight (\texttt{WGHT\_EPI}), episode number
+(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
+(\texttt{DURATION}), and episode location (\texttt{LOCATION}).
 
-In total, the \emph{Episode file} of the 2015 GSS surveys contains
-274,108 records, representing 461,837,622 episodes and 527 variables
-detailing the episodes. Similar to the \emph{Main file}, Statistics
-Canada has created codes for the discrete variables, with each value
-corresponding to a label.
+In total, the episode file of the 2015 GSS surveys contains 274,108
+records, representing 461,837,622 episodes and 527 variables detailing
+the episodes. Similar to the main file, Statistics Canada has created
+codes for the discrete variables, with each value corresponding to a
+label.
 
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
-\end{TableNotes}
-\begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.}\\
+\begin{table}
+\centering
+\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the  first seven episodes of the record number 10041.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{8}{10}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{cccccc}
 \toprule
@@
 \bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
+\end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
+\end{tablenotes}
+\end{threeparttable}}
+\end{table}
 
@@
 Figure \ref{fig:process-figure} presents a diagram illustrating the
-processes applied to the \emph{Main} and \emph{Episode files} to create
-the \{ActiveCA\} datasets. For each TUS cycle, we reviewed the
-\emph{Episode files} to identify movement episodes involving walking or
-cycling. This allowed us to also select the activities immediately
-before and after the movement episode, helping to infer the trip's
-purpose and determine its origin and destination. Active trips were
-identified by their corresponding activity codes, accounting for
-variations across survey cycles.
+processes applied to the main and episode files to create the
+\{ActiveCA\} datasets. For each TUS cycle, we reviewed the episode files
+to identify movement episodes involving walking or cycling. This allowed
+us to also select the activities immediately before and after the
+movement episode, helping to infer the trip's purpose and determine its
+origin and destination. Active trips were identified by their
+corresponding activity codes, accounting for variations across survey
+cycles.
 
@@
 
-For the \emph{Main files}, we selected socioeconomic variables to help
-profile individuals engaged in active travel episodes. These included
-key indicators such as age group, sex, marital status, and number of
-children, among others. As with the \emph{Episode files}, we labeled and
+For the main files, we selected socioeconomic variables to help profile
+individuals engaged in active travel episodes. These included key
+indicators such as age group, sex, marital status, and number of
+children, among others. As with the episode files, we labeled and
 factored the coded variables. Using the appropriate identifiers, we
-joined the Episode and \emph{Main files} to determine the province and
-the type of population centre - whether it is a CMA, a Census
-Agglomeration (CA), or a non-CMA/CA area.
+joined the episode and main files to determine the province and the type
+of population centre - whether it is a CMA, a Census Agglomeration (CA),
+or a non-CMA/CA area.
 
@@
 
-This section presents some potential applications of the \{ActiveCA\}
-\texttt{R} package. In fact, we expect that the application of this
-package to extend beyond our pre-imagined range of uses. The
-installation instruction and also some examples of application of the
-\{ActiveCA\} \texttt{R} package are available in the \texttt{vignettes},
-available in the Github repository.
+This section presents some potential applications of the \{ActiveCA\} R
+package. In fact, we expect that the application of this package to
+extend beyond our pre-imagined range of uses. The installation
+instruction and also some examples of application of the \{ActiveCA\} R
+package are available in the vignettes, available in the Github
+repository.
 
@@
 Table \ref{tab:processed-obs} displays the total number of records
-processed for Main and \emph{Episode files}. For the \emph{Main files},
-a total of 101,667 records were processed, referring to all records from
-the Time Use Surveys from 1986 to 2022, that together represents more of
+processed for Main and episode files. For the main files, a total of
+101,667 records were processed, referring to all records from the Time
+Use Surveys from 1986 to 2022, that together represents more of
 181,526,641 respondents. It also presents the total cases of active
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{longtable}[t]{rr>{}r|rr}
-\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
+\begin{table}
+\centering
+\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{rr>{}r|rr}
 \toprule
@@
 \bottomrule
-\end{longtable}
-\endgroup{}
+\end{tabular}
+\end{table}
 
@@
 \ref{tab:ep-2015-processed} presents the walking episodes for the record
-identification number \texttt{10041} from the TUS PUMF 2015
-\emph{Episode file} (Cycle 29), previously displayed in Table
-\ref{tab:ep-2015-unprocessed}. Only the unique active travel episode
-appear in Table \ref{tab:ep-2015-processed} since the records were
-filtered to select cases with walking or cycling episodes. For both
-cases, Tables \ref{tab:main-2015-processed} and
-\ref{tab:ep-2015-processed} contain labeled variables, facilitating the
-interpretation of the data.
+identification number \texttt{10041} from the TUS PUMF 2015 episode file
+(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
+Only the unique active travel episode appears in Table
+\ref{tab:ep-2015-processed} since the records were filtered to select
+cases with walking or cycling episodes. For both cases, Tables
+\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
+labeled variables, facilitating the interpretation of the data.
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item Legend: PUMFID: record identification.
-WGHT\_PER:  person weight.
-SURVMNTH: survey month of data collection.
-AGEGR10: age group of the respondent.
-SEX: sex of the respondent.
-MARSTAT: marital status of the respondent.
-\end{TableNotes}
-\begin{longtable}[t]{cccccc}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}\\
+\begin{table}
+\centering
+\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{8}{10}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{cccccc}
 \toprule
@@
 \bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
+\end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item Legend: PUMFID: record identification.
+WGHT\_PER:  person weight.
+SURVMNTH: survey month of data collection.
+AGEGR10: age group of the respondent.
+SEX: sex of the respondent.
+MARSTAT: marital status of the respondent.
+\end{tablenotes}
+\end{threeparttable}}
+\end{table}
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}
-\item \textit{Note: } 
-\item Legend: PUMFID: record identification. EPINO: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
-\end{TableNotes}
-\begin{longtable}[t]{rrlrlll}
-\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.}\\
+\begin{table}
+\centering
+\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.}
+\centering
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
+\fontsize{8}{10}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{rrlrlll}
 \toprule
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{longtable}[t]{>{}llccccccc}
-\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
+\begin{table}
+\centering
+\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics of the duration (in minutes) of episodes with active transport records.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{>{}llccccccc}
 \toprule
@@
  & Maximum & 660 & 300 & 255 & 515 & 480 & 900 & 480\\
-\nopagebreak
+
  & Mean & 21 & 21 & 12 & 12 & 13 & 18 & 19\\
-\nopagebreak
+
  & Median & 15 & 10 & 5 & 10 & 10 & 10 & 15\\
-\nopagebreak
+
  & Minimum & 1 & 1 & 1 & 0 & 0 & 5 & 5\\
-\nopagebreak
+
 \multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27 & 24\\
-\cmidrule{1-9}\pagebreak[0]
+\cmidrule{1-9}
  & Maximum &  & 240 & 90 & 180 & 153 & 120 & 150\\
-\nopagebreak
+
  & Mean &  & 28 & 24 & 20 & 19 & 25 & 40\\
-\nopagebreak
+
  & Median &  & 15 & 15 & 15 & 10 & 20 & 30\\
-\nopagebreak
+
  & Minimum &  & 5 & 2 & 1 & 1 & 5 & 5\\
-\nopagebreak
+
 \multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20 & 20\\
 \bottomrule
-\end{longtable}
-\endgroup{}
+\end{tabular}
+\end{table}
 
@@
 
-In 2022, \texttt{home} served as a central hub for most trips, with
+In 2022, home location served as a central hub for most trips, with
 fewer than 10\% of journeys not involving it as either a starting point
-or destination. The most common trip types were from \texttt{home} to
-\texttt{work\ or\ school} (17\%) and the reverse, from
-\texttt{work\ or\ school} to \texttt{home} (13\%). Notably, 7\% of trips
-began and ended at \texttt{home}, often reflecting leisure activities
-such as short walks or dog walking. \texttt{Grocery\ stores} were also a
-key destination, comprising 10\% of trips departing from \texttt{home.}
+or destination. The most common trip types were from home to work or
+school (17\%) and the reverse, from work or school to home (13\%).
+Notably, 7\% of trips began and ended at home, often reflecting leisure
+activities such as short walks or dog walking. Grocery stores were also
+a key destination, comprising 10\% of trips departing from home.
 
@@
 \centering
-\caption{\label{tab:cma-durations}\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.}
+\caption{\label{tab:cma-durations}\label{tab:cma-durations}Differences in walking duration (in minutes) between provinces and population centre type.}
 \centering
@@
 
-\{ActiveCA\} also provides a Jupyter Notebook containing a
-\texttt{Python} script that demonstrates how to read \texttt{R} data
-files (.rda) and convert them into Pandas DataFrames. This process
-allows users to work with and utilize the datasets available in
-\{ActiveCA\} within a \texttt{Python} project.
+\{ActiveCA\} also provides a Jupyter Notebook containing a Python script
+that demonstrates how to read R data files (.rda) and convert them into
+Pandas Data Frames. This process allows users to work with and utilize
+the datasets available in \{ActiveCA\} within a Python project.
 
@@
 (2005), 24 (2010), 29 (2015), and 34 (2022) of TUS GSSs on active travel
-in Canada. In the form of an \texttt{R} data package, \{ActiveCA\} was
-developed after collecting, cleaning, and processing the survey data,
-providing information on origins, destinations, and duration of active
-travel, as well other information.
+in Canada. In the form of an R data package, \{ActiveCA\} was developed
+after collecting, cleaning, and processing the survey data, providing
+information on origins, destinations, and duration of active travel, as
+well other information.
 
@@
 
-The author(s) declared no potential conflicts of interest with respect
-to the research, authorship, and/or publication of this article.
+The authors declared no potential conflicts of interest with respect to
+the research, authorship, and/or publication of this article.
+
+\section{Acknowledgments}\label{acknowledgments}
+
+The authors thank the anonymous reviewers for their insightful comments.
 
@@
 
-The author(s) disclosed receipt of the following financial support for
-the research, authorship, and/or publication of this article: This work
-was supported by the Social Sciences and Humanities Research Council of
-Canada (\emph{More description about the funding source after the review
-process}).
+The work has been supported by funding from the Social Sciences and
+Humanities Research Council's Partnership Grant: Mobilizing justice:
+towards evidence-based transportation equity policy.
 
@@
 
-Author 1
+Bruno Dias dos Santos - \url{https://orcid.org/0000-0002-6748-2038}
 
-Author 2
+Mahdis Moghadasi - \url{https://orcid.org/0000-0003-1009-9369}
 
-Author 3
+Antonio Páez - \url{https://orcid.org/0000-0001-6912-9919}
 
@@
 
-The \{ActiveCA\} R data package can be found and installed on Github
-(\emph{link}).
-
-For review purposes, the package is currently available as a tar.gz file
-that can be installed by R. The file can be obtained from this anonymous
-location:
-
-\url{https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip}
+The \{ActiveCA\} R data package can be found and installed on
+\href{https://github.com/dias-bruno/ActiveCA}{Github}.
 

### 2025-08-14 paper/EPB-Final-Submission.Rmd
adjusting tables for publication

@@
 The objective of this paper is to introduce {ActiveCA}, an open data
-product with data from all Time Use Survey (TUS) cycles of the General
+product with data from all Time Use Survey (TUS) cycles of the Canadian General
 Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
@@
 
-Using Public Use Microdata Files (PUMFs) from the TUS,
+Using PUMFs from the TUS,
 we extracted all data necessary to characterize active travel in
@@
 endeavour that can only improve data-driven decisions in transportation,
-urban, and health policy.
-
-The rest of this paper discusses the sources of data, and the process
+urban, and health policy. The rest of this paper discusses the sources of data, and the process
 implemented to retrieve and package them. Then, we show some selected examples of analysis to whet the
@@
 obtain data on social trends to track changes in Canadians' living
-conditions and well-being over time. The series of survey on time use
-are used to understand how Canadian residents spend and manage their
+conditions and well-being over time. TUS are used to understand how Canadian residents spend and manage their
 time, and what factors contribute to their happiness and stress. The GSS
@@
 time use. The first Canadian TUS done as part of the GSS program was
-conducted in 1986, and the most recent was completed in 2022. These Time
-Use Surveys [@statisticscanada2022] collect data on respondents'
+conducted in 1986, and the most recent was completed in 2022. These surveys [@statisticscanada2022] collect data on respondents'
 participation and time spent on a wide range of everyday activities
@@
 
-The Public Use Microdata are released by Statistics Canada in two files:
+The Public Use Microdata Files (PUMFs) are released by Statistics Canada in two files:
 a main file and an episode file. The files are linked by keys that
@@
 
-The main file of the time use survey compiles a large array of
+The main file of the TUS compiles a large array of
 aggregated data, summarizing the answers to the questionnaire that
@@
     format = "latex", 
+    format.args = list(big.mark = ",", decimal.mark = "."), 
     booktabs = TRUE, 
@@
 Table \ref{tab:processed-obs} displays the total number of records
-processed for Main and episode files. For the main files, a total of
+processed for main and episode files. For the main files, a total of
 `r sum(df_main_processed$Count)` records were processed, referring to
-all records from the Time Use Surveys from 1986 to 2022, that together
+all records from the TUSs from 1986 to 2022, that together
 represents more of `r sum(df_main_processed$Weighted)` respondents. It
@@
   select(-c(Survey_ep, Survey, Year_ep)) %>% 
-  kable(format = "latex", booktabs = TRUE, longtable = FALSE, 
+  kable(format = "latex", 
+    format.args = list(big.mark = ",", decimal.mark = "."),
+    booktabs = TRUE, longtable = FALSE, 
         caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
@@
 Table \ref{tab:main-2015-processed} shows the first ten rows and first
-six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
+six variables of the TUS PUMF 2015 main file (Cycle 29), displayed in
 \ref{tab:main-2015-unprocessed} before our processing. Table
@@
     format = "latex", 
+    format.args = list(big.mark = ",", decimal.mark = "."),
     booktabs = TRUE, 
@@
 styled_ch03_table_01 <- stats_data |>
-  kable(format = "latex", booktabs = TRUE, longtable = FALSE, 
+  kable(format = "latex", 
+    format.args = list(big.mark = ",", decimal.mark = "."),
+    booktabs = TRUE, longtable = FALSE, 
         caption = "\\label{tab:table-01}Descriptive statistics of the duration (in minutes) of episodes with active transport records.",  
@@
 
-Bruno Dias dos Santos - https://orcid.org/0000-0002-6748-2038
-
-Mahdis Moghadasi - https://orcid.org/0000-0003-1009-9369
-
-Antonio Páez - https://orcid.org/0000-0001-6912-9919
+- Bruno Dias dos Santos - https://orcid.org/0000-0002-6748-2038 
+- Mahdis Moghadasi - https://orcid.org/0000-0003-1009-9369 
+- Antonio Páez - https://orcid.org/0000-0001-6912-9919
 

### 2025-08-14 paper/EPB-Final-Submission.tex
adjusting tables for publication

@@
 
-\corrauth{Bruno Dias dos Santos}
+\corrauth{Bruno Dias dos Santos - School of Earth, Environment and
+Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1
+Canada.}
 
@@
 The objective of this paper is to introduce \{ActiveCA\}, an open data
-product with data from all Time Use Survey (TUS) cycles of the General
-Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
-process that transforms raw data (open or not) into analysis-ready data,
-following a transparent process in which all stages of development
+product with data from all Time Use Survey (TUS) cycles of the Canadian
+General Social Surveys (GSSs). Open data products (ODPs) are the outcome
+of a process that transforms raw data (open or not) into analysis-ready
+data, following a transparent process in which all stages of development
 follow open principles \citep{arribas-bel2021}. ODPs, while still open,
@@
 
-Using Public Use Microdata Files (PUMFs) from the TUS, we extracted all
-data necessary to characterize active travel in Canada - specifically,
-episodes in which the activity involved moving between an origin and a
-destination by walking or cycling. Although Statistics Canada provides
-PUMFs and accompanying documentation for the GSS program
+Using PUMFs from the TUS, we extracted all data necessary to
+characterize active travel in Canada - specifically, episodes in which
+the activity involved moving between an origin and a destination by
+walking or cycling. Although Statistics Canada provides PUMFs and
+accompanying documentation for the GSS program
 \citep[see][]{statisticscanada2024}, accessing and preparing these files
@@
 is a worthy endeavour that can only improve data-driven decisions in
-transportation, urban, and health policy.
-
-The rest of this paper discusses the sources of data, and the process
-implemented to retrieve and package them. Then, we show some selected
-examples of analysis to whet the imagination of potential users. This
-ODP provides not only data that are easy to use, but also all the code
-and documentation that make this a reproducible research project. In
-summary, \{ActiveCA\} aims to implement and inspire the best principles
-of open spatial sciences \citep{paez_open_2021, brunsdon_opening_2021}.
+transportation, urban, and health policy. The rest of this paper
+discusses the sources of data, and the process implemented to retrieve
+and package them. Then, we show some selected examples of analysis to
+whet the imagination of potential users. This ODP provides not only data
+that are easy to use, but also all the code and documentation that make
+this a reproducible research project. In summary, \{ActiveCA\} aims to
+implement and inspire the best principles of open spatial sciences
+\citep{paez_open_2021, brunsdon_opening_2021}.
 
@@
 surveys to obtain data on social trends to track changes in Canadians'
-living conditions and well-being over time. The series of survey on time
-use are used to understand how Canadian residents spend and manage their
-time, and what factors contribute to their happiness and stress. The GSS
-program was created in 1985, and is serialized to provide a collection
-of annual, representative cross-sectional surveys.
+living conditions and well-being over time. TUS are used to understand
+how Canadian residents spend and manage their time, and what factors
+contribute to their happiness and stress. The GSS program was created in
+1985, and is serialized to provide a collection of annual,
+representative cross-sectional surveys.
 
@@
 time use. The first Canadian TUS done as part of the GSS program was
-conducted in 1986, and the most recent was completed in 2022. These Time
-Use Surveys \citep{statisticscanada2022} collect data on respondents'
+conducted in 1986, and the most recent was completed in 2022. These
+surveys \citep{statisticscanada2022} collect data on respondents'
 participation and time spent on a wide range of everyday activities
@@
 
-The Public Use Microdata are released by Statistics Canada in two files:
-a main file and an episode file. The files are linked by keys that
-identify households, individuals, and episodes (i.e., activities)
-conducted by individuals. We discuss these files in more detail in the
-following section.
+The Public Use Microdata Files (PUMFs) are released by Statistics Canada
+in two files: a main file and an episode file. The files are linked by
+keys that identify households, individuals, and episodes (i.e.,
+activities) conducted by individuals. We discuss these files in more
+detail in the following section.
 
@@
 
-The main file of the time use survey compiles a large array of
-aggregated data, summarizing the answers to the questionnaire that
-describe households and individuals, as well as derived variables that
-summarize the respondents' use of time use across different activities,
-locations, and social interactions. This file documents the time and
-duration that respondents allocate to each activity and location. The
-main file provides a overview of daily routines and social dynamics, not
-focusing on individual activity episodes. Additionally, this file
-categorizes activities into bigger groups and subcategories,
-facilitating the data's analytical utility with additional metrics such
-as total transit time, time spent with household members, and counts of
-activities and episodes.
+The main file of the TUS compiles a large array of aggregated data,
+summarizing the answers to the questionnaire that describe households
+and individuals, as well as derived variables that summarize the
+respondents' use of time use across different activities, locations, and
+social interactions. This file documents the time and duration that
+respondents allocate to each activity and location. The main file
+provides a overview of daily routines and social dynamics, not focusing
+on individual activity episodes. Additionally, this file categorizes
+activities into bigger groups and subcategories, facilitating the data's
+analytical utility with additional metrics such as total transit time,
+time spent with household members, and counts of activities and
+episodes.
 
@@
 \midrule
-10000 & 616.6740 & 7 & 5 & 1 & 5\\
-10001 & 8516.6140 & 7 & 5 & 1 & 1\\
-10002 & 371.7520 & 1 & 4 & 2 & 1\\
-10003 & 1019.3135 & 3 & 6 & 2 & 5\\
-10004 & 1916.0708 & 9 & 2 & 1 & 6\\
+10,000 & 616.6740 & 7 & 5 & 1 & 5\\
+10,001 & 8,516.6140 & 7 & 5 & 1 & 1\\
+10,002 & 371.7520 & 1 & 4 & 2 & 1\\
+10,003 & 1,019.3135 & 3 & 6 & 2 & 5\\
+10,004 & 1,916.0708 & 9 & 2 & 1 & 6\\
 \addlinespace
-10005 & 1952.2015 & 4 & 1 & 1 & 6\\
-10006 & 5761.5528 & 8 & 1 & 1 & 6\\
-10007 & 466.0426 & 6 & 5 & 2 & 3\\
-10008 & 2479.2991 & 2 & 2 & 2 & 1\\
-10009 & 1436.1641 & 8 & 6 & 1 & 3\\
+10,005 & 1,952.2015 & 4 & 1 & 1 & 6\\
+10,006 & 5,761.5528 & 8 & 1 & 1 & 6\\
+10,007 & 466.0426 & 6 & 5 & 2 & 3\\
+10,008 & 2,479.2991 & 2 & 2 & 2 & 1\\
+10,009 & 1,436.1641 & 8 & 6 & 1 & 3\\
 \bottomrule
@@
 \midrule
-10041 & 1 & 1353.818 & 1 & 210 & 300\\
-10041 & 2 & 1353.818 & 2 & 40 & 300\\
-10041 & 3 & 1353.818 & 27 & 15 & 300\\
-10041 & 4 & 1353.818 & 7 & 15 & 315\\
-10041 & 5 & 1353.818 & 9 & 180 & 300\\
+10,041 & 1 & 1,353.818 & 1 & 210 & 300\\
+10,041 & 2 & 1,353.818 & 2 & 40 & 300\\
+10,041 & 3 & 1,353.818 & 27 & 15 & 300\\
+10,041 & 4 & 1,353.818 & 7 & 15 & 315\\
+10,041 & 5 & 1,353.818 & 9 & 180 & 300\\
 \addlinespace
-10041 & 6 & 1353.818 & 6 & 15 & 300\\
-10041 & 7 & 1353.818 & 18 & 120 & 300\\
+10,041 & 6 & 1,353.818 & 6 & 15 & 300\\
+10,041 & 7 & 1,353.818 & 18 & 120 & 300\\
 \bottomrule
@@
 Table \ref{tab:processed-obs} displays the total number of records
-processed for Main and episode files. For the main files, a total of
-101,667 records were processed, referring to all records from the Time
-Use Surveys from 1986 to 2022, that together represents more of
-181,526,641 respondents. It also presents the total cases of active
-trips episodes identified. In total 23,513 records with register of
-active travel activity. Together, these records account for 44,316,110
-episodes.
+processed for main and episode files. For the main files, a total of
+101,667 records were processed, referring to all records from the TUSs
+from 1986 to 2022, that together represents more of 181,526,641
+respondents. It also presents the total cases of active trips episodes
+identified. In total 23,513 records with register of active travel
+activity. Together, these records account for 44,316,110 episodes.
 
@@
 \midrule
-2022 & 12336 & 32136802 & 1765 & 6041032\\
-2015 & 17390 & 29766399 & 3496 & 6634387\\
-2010 & 15390 & 28075610 & 4615 & 8516753\\
-2005 & 19597 & 26095819 & 5866 & 7583838\\
-1998 & 10749 & 24260137 & 1789 & 3606987\\
+2,022 & 12,336 & 32,136,802 & 1,765 & 6,041,032\\
+2,015 & 17,390 & 29,766,399 & 3,496 & 6,634,387\\
+2,010 & 15,390 & 28,075,610 & 4,615 & 8,516,753\\
+2,005 & 19,597 & 26,095,819 & 5,866 & 7,583,838\\
+1,998 & 10,749 & 24,260,137 & 1,789 & 3,606,987\\
 \addlinespace
-1992 & 9815 & 21294313 & 1635 & 3691918\\
-1986 & 16390 & 19897562 & 4347 & 8241196\\
+1,992 & 9,815 & 21,294,313 & 1,635 & 3,691,918\\
+1,986 & 16,390 & 19,897,562 & 4,347 & 8,241,196\\
 \bottomrule
@@
 \midrule
-10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
-10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
-10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
-10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
-10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
+10,000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
+10,001 & 8,516.6140 & July & 55 to 64 years & Male & Married\\
+10,002 & 371.7520 & January & 45 to 54 years & Female & Married\\
+10,003 & 1,019.3135 & March & 65 to 74 years & Female & Divorced\\
+10,004 & 1,916.0708 & September & 25 to 34 years & Male & Single, never married\\
 \addlinespace
-10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
-10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
-10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
-10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
-10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
+10,005 & 1,952.2015 & April & 15 to 24 years & Male & Single, never married\\
+10,006 & 5,761.5528 & August & 15 to 24 years & Male & Single, never married\\
+10,007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
+10,008 & 2,479.2991 & February & 25 to 34 years & Female & Married\\
+10,009 & 1,436.1641 & August & 65 to 74 years & Male & Widowed\\
 \bottomrule
@@
 \midrule
-10041 & 1353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
+10,041 & 1,353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
 \bottomrule
@@
 
+\begin{itemize}
+\tightlist
+\item
   Bruno Dias dos Santos - \url{https://orcid.org/0000-0002-6748-2038}
-
+\item
   Mahdis Moghadasi - \url{https://orcid.org/0000-0003-1009-9369}
-
+\item
   Antonio Páez - \url{https://orcid.org/0000-0001-6912-9919}
+\end{itemize}
 

### 2025-11-06 data-raw/01-preprocessing-episode-files.Rmd
Adjusting origins/destinations to include travel options (such as walking to take the bus or car)

@@
 This Rmarkdown file is part of the
-[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. The main objective of the *ActiveCA* package is to identify the most
-appropriate impedance functions for active transportation modes for
-various destinations and time periods in Canada. To do achieve this objective, we analyzed historical data from the [General Social Survey
+[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. The main
+objective of the *ActiveCA* package is to identify the most appropriate
+impedance functions for active transportation modes for various
+destinations and time periods in Canada. To do achieve this objective,
+we analyzed historical data from the [General Social Survey
 (GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
@@
 been conducting time-use surveys at five to seven year intervals since
-1986, with the most recent being in 2022 (1986, 1992, 1998, 2005, 2010, 2015, and 2022). The time use collects information on participation of
+1986, with the most recent being in 2022 (1986, 1992, 1998, 2005, 2010,
+2015, and 2022). The time use collects information on participation of
 respondents and time spent on a wide range of day-to-day activities
@@
 
-For the purpose of our analyses, the following definitions are important:
+For the purpose of our analyses, the following definitions are
+important:
 
 | **Code** | **Location** | **Year** | **Description** |
-|-------------|------------------|-------------|------------------------------|
+|----------------|----------------|----------------|--------------------------|
 | 300 | At home or on property | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
@@
 
--   `PUMFID`: the record identification[^2].
+-   `PUMFID`: the record identification[^1].
 -   `WGHT_EPI`: Episode weight.
@@
 
-[^2]: This variable serves as the link between the Main and the Episode
+[^1]: This variable serves as the link between the Main and the Episode
     files. Each case has a unique record identifier and this appears on
@@
 
-The file name pattern is as follows: `transportation mode`[^3] + `_` +
-`year of survey` [^4]. This means that, for instance, the file
+The file name pattern is as follows: `transportation mode`[^2] + `_` +
+`year of survey` [^3]. This means that, for instance, the file
 `walking_2015` refers to transportation mode table `walking` in the year
@@
 
-[^3]:  `walking` or `cycling`
-
-[^4]:  whether 1986, 1992, 1998, 2005, 2010, 2015, or 2022. Note: There is no
-    data on cycling for the 1986 GSS survey.
+[^2]: `walking` or `cycling`
+
+[^3]: whether 1986, 1992, 1998, 2005, 2010, 2015, or 2022. Note: There
+    is no data on cycling for the 1986 GSS survey.
 
@@
 Now, we'll select all the rows with walking episodes in the GSS survey,
-selecting the columns that we are interested[^5], and already creating
+selecting the columns that we are interested[^4], and already creating
 the origin and destination variables:
 
-[^5]: Specifically for the 2015 GSS survey, `TUI_01` refers to the
+[^4]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
     activity code. We'll change the name of this column to be match the
@@
 walking_2015 <- walking_2015 |>
-  filter(origin == 300 | 
-           origin == 301 | 
-           origin == 302 | 
-           origin == 303 | 
-           origin == 304 |
-           origin == 305 | 
-           origin == 306 | 
-           origin == 307 |
-           origin == 308 |
-           origin == 309 | 
-           origin == 310 |
-           origin == 311) |>
+  filter(origin <= 321) |>
   mutate(orig_label =
@@
                      origin == 310 ~ "Place of worship",
-                     origin == 311 ~ "Medical, dental or other health clinic"),
+                     origin == 311 ~ "Medical, dental or other health clinic",
+                     origin == 312 ~ "Elsewhere",
+                     origin %in% c(313,314,319) ~ "Travel - Motorized",
+                     origin %in% c(317,320,321) ~ "Travel - Other",
+                     origin == 316 ~ "Travel - Transit",
+                     origin == 315 ~ "Travel - Walk",
+                     origin == 318 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 walking_2015 <- walking_2015 |>  
-    filter(destination == 300 | 
-           destination == 301 | 
-           destination == 302 | 
-           destination == 303 |
-           destination == 304 |
-           destination == 305 |
-           destination == 306 | 
-           destination == 307 |
-           destination == 308 |
-           destination == 309 | 
-           destination == 310 |
-           destination == 311)  |>
-  # Label the destinations
+  filter(destination <= 321) |>
   mutate(dest_label =
@@
                      destination == 310 ~ "Place of worship",
-                     destination == 311 ~ "Medical, dental or other health clinic"),
+                     destination == 311 ~ "Medical, dental or other health clinic",
+                     destination == 312 ~ "Elsewhere",
+                     destination %in% c(313,314,319) ~ "Travel - Motorized",
+                     destination %in% c(317,320,321) ~ "Travel - Other",
+                     destination == 316 ~ "Travel - Transit",
+                     destination == 315 ~ "Travel - Walk",
+                     destination == 318 ~ "Travel - Bycicle"),
          dest_label = factor(dest_label))
@@
 Now, let's select only the cases of respondents who live in a CMA or CA:
+
 ```{r}
@@
 
-The table above shows that the most common destination for a walking trip
-that starts at home are `Home` and `Work or school`. It also shows that the second most common walking trip is from `Grocery store, other stores or mall` to `Home`. Notice, though,
-that quite a few trips start *and* end at Home. What are those supposed
-to be?
+The table above shows that the most common destination for a walking
+trip that starts at home are `Home` and `Work or school`. It also shows
+that the second most common walking trip is from
+`Grocery store, other stores or mall` to `Home`. Notice, though, that
+quite a few trips start *and* end at Home. What are those supposed to
+be?
 
@@
 looking for work (code `9`). So, this was a recreational/leisure
-trip![^6]
-
-[^6]: `TUI_01` is an activity code of the episode (What were you doing
+trip![^5]
+
+[^5]: `TUI_01` is an activity code of the episode (What were you doing
     at minute). The list of main activity codes (001-095) can be found
@@
 
-Label the activity ("7" is "Transport to or from activity") and drop the LOCATION because it is already coded as MODE:    
+Label the activity ("7" is "Transport to or from activity") and drop the
+LOCATION because it is already coded as MODE:
+
 ```{r}
@@
 cycling_2015 <- cycling_2015 |>
-  filter(origin == 300 | 
-           origin == 301 | 
-           origin == 302 | 
-           origin == 303 | 
-           origin == 304 |
-           origin == 305 | 
-           origin == 306 | 
-           origin == 307 |
-           origin == 308 |
-           origin == 309 | 
-           origin == 310 |
-           origin == 311) |>
+  filter(origin <= 321) |>
   mutate(orig_label =
@@
                      origin == 306 ~ "Grocery store, other stores or mall",
+                     origin == 307 ~ "Library, museum or theatre",
                      origin == 308 ~ "Sports centre, field or arena",
-                     origin == 307 ~ "Library, museum or theatre",
                      origin == 309 ~ "Restaurant, bar or club",
                      origin == 310 ~ "Place of worship",
-                     origin == 311 ~ "Medical, dental or other health clinic"),
+                     origin == 311 ~ "Medical, dental or other health clinic",
+                     origin == 312 ~ "Elsewhere",
+                     origin %in% c(313,314,319) ~ "Travel - Motorized",
+                     origin %in% c(317,320,321) ~ "Travel - Other",
+                     origin == 316 ~ "Travel - Transit",
+                     origin == 315 ~ "Travel - Walk",
+                     origin == 318 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 cycling_2015 <- cycling_2015 |>  
-  # Choose destinations of interest
-  filter(destination == 300 | 
-           destination == 301 | 
-           destination == 302 | 
-           destination == 303 |
-           destination == 304 |
-           destination == 305 |
-           destination == 306 | 
-           destination == 307 |
-           destination == 308 |
-           destination == 309 | 
-           destination == 310 |
-           destination == 311)  |>
-  # Label the destinations
+  filter(destination <= 321) |>
   mutate(dest_label =
@@
 The most common destination for a cycling trip that starts at home is
-`Work or school`. Again, we can notice that there are quite a few trips that
-start *and* end at Home. What are those supposed to be?
+`Work or school`. Again, we can notice that there are quite a few trips
+that start *and* end at Home. What are those supposed to be?
 
@@
 
-The individual was at home engaged in personal care activities
-(`TUI_01` = `2`) before cycling for approximately 10 minutes to a grocery
-store. After spending 30 minutes at the store, they cycled back home.
-Upon returning, the person prepared a meal and cleaned the house for approximately 85 minutes. Subsequently, the individual embarked on
+The individual was at home engaged in personal care activities (`TUI_01`
+= `2`) before cycling for approximately 10 minutes to a grocery store.
+After spending 30 minutes at the store, they cycled back home. Upon
+returning, the person prepared a meal and cleaned the house for
+approximately 85 minutes. Subsequently, the individual embarked on
 another bicycle trip to the store, which took 10 minutes, and thereafter
-returned home.[^7]
-
-[^7]: Note: `TUI_01` is an activity code of the episode (What were you
+returned home.[^6]
+
+[^6]: Note: `TUI_01` is an activity code of the episode (What were you
     doing in minutes?).The list of main activity codes (001-095) can be
@@
 Save data in R format for later use:
+
 ```{r}
@@
 
-As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data
+As we have already processed this year's data, we will clear the Episode
+and Main files for this year from the computer's memory before uploading
+any more data
 
@@
 walking_2010 <- walking_2010 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 | 
-           origin == 4 | 
-           origin == 5 |
-           origin == 6 | 
-           origin == 7 | 
-           origin == 8 |
-           origin == 9 |
-           origin == 10 ) |>
+  filter(origin <= 21 ) |>
   mutate(orig_label =
            case_when(origin == 1 ~ "Home", 
-                     origin == 2 ~ "Work or school",
+                     origin %in% c(2,8) ~ "Work or school",
                      origin == 3 ~ "Other's home",
+                     origin == 9 ~ "Outdoors",
+                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
+                     origin == 10 ~ "Library, museum or theatre",
                      origin == 4 ~ "Restaurant, bar or club",
                      origin == 5 ~ "Place of worship",
-                     origin == 6 ~ "Grocery store, other stores or mall",
-                     origin == 7 ~ "Grocery store, other stores or mall",                     
-                     origin == 8 ~ "Work or school",
-                     origin == 9 ~ "Outdoors",
-                     origin == 10 ~ "Library, museum or theatre"),
+                     origin == 11 ~ "Elsewhere",
+                     origin %in% c(12,13,19) ~ "Travel - Motorized",
+                     origin %in% c(18,20,21) ~ "Travel - Other",
+                     origin %in% c(15,16) ~ "Travel - Transit",
+                     origin == 14 ~ "Travel - Walk",
+                     origin == 17 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 walking_2010 <- walking_2010 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 | 
-           destination == 4 | 
-           destination == 5 |
-           destination == 6 | 
-           destination == 7 | 
-           destination == 8 |
-           destination == 9 |
-           destination == 10 ) |>
+  filter(destination <= 21 ) |>
   mutate(dest_label =
            case_when(destination == 1 ~ "Home", 
-                     destination == 2 ~ "Work or school",
+                     destination %in% c(2,8) ~ "Work or school",
                      destination == 3 ~ "Other's home",
                      destination == 9 ~ "Outdoors",
-                     destination == 6 ~ "Grocery store, other stores or mall",
+                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                      destination == 10 ~ "Library, museum or theatre",
@@
                      destination == 5 ~ "Place of worship",
-                     destination == 7 ~ "Grocery store, other stores or mall",                     
-                     destination == 8 ~ "Work or school"),
+                     destination == 11 ~ "Elsewhere",
+                     destination %in% c(12,13,19) ~ "Travel - Motorized",
+                     destination %in% c(18,20,21) ~ "Travel - Other",
+                     destination %in% c(15,16) ~ "Travel - Transit",
+                     destination == 14 ~ "Travel - Walk",
+                     destination == 17 ~ "Travel - Bycicle"),
          dest_label = factor(dest_label))
@@
 
-For the year 2010, the most common destination for a walking trip that starts at home are
-`Home` to `Home`, followed by `Home` to `Work or school`. Again, we can check the home-to-home trips:
+For the year 2010, the most common destination for a walking trip that
+starts at home are `Home` to `Home`, followed by `Home` to
+`Work or school`. Again, we can check the home-to-home trips:
 
@@
 walk to another location. At this new location, the person undertook
-activities related to washing and dressing (`ACTCODE`=`400`) for around 10
-minutes, then went for a 10-minute walk, returning home after that.
+activities related to washing and dressing (`ACTCODE`=`400`) for around
+10 minutes, then went for a 10-minute walk, returning home after that.
 Subsequently, they dedicated 10 minutes to meal preparation
-(`ACTCODE`=`101`), followed by a 10-minute session of food (or meal) cleanup
-(`ACTCODE`=`110`). The individual then took a 15-minute walk, returned home,
-and spent 5 minutes on household management tasks, which included
-organizing and planning activities.[^8]
-
-[^8]: Note: `ACTCODE` is an activity code of the episode (What were you
+(`ACTCODE`=`101`), followed by a 10-minute session of food (or meal)
+cleanup (`ACTCODE`=`110`). The individual then took a 15-minute walk,
+returned home, and spent 5 minutes on household management tasks, which
+included organizing and planning activities.[^7]
+
+[^7]: Note: `ACTCODE` is an activity code of the episode (What were you
     doing in minute).The list of main activity codes can be found in an
@@
 
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r}
@@
 cycling_2010 <- cycling_2010 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 | 
-           origin == 4 | 
-           origin == 5 |
-           origin == 6 | 
-           origin == 7 | 
-           origin == 8 |
-           origin == 9 |
-           origin == 10 ) |>
+  filter(origin <= 21 ) |>
   mutate(orig_label =
            case_when(origin == 1 ~ "Home", 
-                     origin == 2 ~ "Work or school",
+                     origin %in% c(2,8) ~ "Work or school",
                      origin == 3 ~ "Other's home",
                      origin == 9 ~ "Outdoors",
-                     origin == 6 ~ "Grocery store, other stores or mall",
+                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                      origin == 10 ~ "Library, museum or theatre",
@@
                      origin == 5 ~ "Place of worship",
-                     origin == 7 ~ "Grocery store, other stores or mall",                     
-                     origin == 8 ~ "Work or school"),
+                     origin == 11 ~ "Elsewhere",
+                     origin %in% c(12,13,19) ~ "Travel - Motorized",
+                     origin %in% c(18,20,21) ~ "Travel - Other",
+                     origin %in% c(15,16) ~ "Travel - Transit",
+                     origin == 14 ~ "Travel - Walk",
+                     origin == 17 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 cycling_2010 <- cycling_2010 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 | 
-           destination == 4 | 
-           destination == 5 |
-           destination == 6 | 
-           destination == 7 | 
-           destination == 8 |
-           destination == 9 |
-           destination == 10 ) |>
+  filter(destination <= 21 ) |>
   mutate(dest_label =
            case_when(destination == 1 ~ "Home", 
-                     destination == 2 ~ "Work or school",
+                     destination %in% c(2,8) ~ "Work or school",
                      destination == 3 ~ "Other's home",
                      destination == 9 ~ "Outdoors",
-                     destination == 6 ~ "Grocery store, other stores or mall",
+                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                      destination == 10 ~ "Library, museum or theatre",
@@
 walking_2005 <- walking_2005 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 | 
-           origin == 4 | 
-           origin == 5 |
-           origin == 6 | 
-           origin == 7 | 
-           origin == 8 |
-           origin == 9 |
-           origin == 10 ) |>
+  filter(origin <= 21 ) |>
   mutate(orig_label =
            case_when(origin == 1 ~ "Home", 
-                     origin == 2 ~ "Work or school",
+                     origin %in% c(2,8) ~ "Work or school",
                      origin == 3 ~ "Other's home",
                      origin == 9 ~ "Outdoors",
-                     origin == 6 ~ "Grocery store, other stores or mall",
+                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                      origin == 10 ~ "Library, museum or theatre",
@@
                      origin == 5 ~ "Place of worship",
-                     origin == 7 ~ "Grocery store, other stores or mall",               
-                     origin == 8 ~ "Work or school"),
+                     origin == 11 ~ "Elsewhere",
+                     origin %in% c(12,13,19) ~ "Travel - Motorized",
+                     origin %in% c(18,20,21) ~ "Travel - Other",
+                     origin %in% c(15,16) ~ "Travel - Transit",
+                     origin == 14 ~ "Travel - Walk",
+                     origin == 17 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 walking_2005 <- walking_2005 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 | 
-           destination == 4 | 
-           destination == 5 |
-           destination == 6 | 
-           destination == 7 | 
-           destination == 8 |
-           destination == 9 |
-           destination == 10 ) |>
+  filter(destination <= 21 ) |>
   mutate(dest_label =
            case_when(destination == 1 ~ "Home", 
-                     destination == 2 ~ "Work or school",
+                     destination %in% c(2,8) ~ "Work or school",
                      destination == 3 ~ "Other's home",
                      destination == 9 ~ "Outdoors",
-                     destination == 6 ~ "Grocery store, other stores or mall",
+                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                      destination == 10 ~ "Library, museum or theatre",
@@
 The individual engaged in personal grooming and dressing activities
-(`ACTCODE`= 400) before embarking on a brief walk lasting approximately 5
-minutes, after which they returned home. Subsequently, the person
+(`ACTCODE`= 400) before embarking on a brief walk lasting approximately
+5 minutes, after which they returned home. Subsequently, the person
 participated in religious activities at their residence, which included
@@
 location and then spent an additional 5 minutes walking to return
-home.[^9]
-
-[^9]: Note: `ACTCODE` is an activity code of the episode (What were you
+home.[^8]
+
+[^8]: Note: `ACTCODE` is an activity code of the episode (What were you
     doing in minute).The list of main activity codes can be found in an
@@
 cycling_2005 <- cycling_2005 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 | 
-           origin == 4 | 
-           origin == 5 |
-           origin == 6 | 
-           origin == 7 | 
-           origin == 8 |
-           origin == 9 |
-           origin == 10 ) |>
+  filter(origin <= 21 ) |>
   mutate(orig_label =
            case_when(origin == 1 ~ "Home", 
-                     origin == 2 ~ "Work or school",
+                     origin %in% c(2,8) ~ "Work or school",
                      origin == 3 ~ "Other's home",
                      origin == 9 ~ "Outdoors",
-                     origin == 6 ~ "Grocery store, other stores or mall",
+                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                      origin == 10 ~ "Library, museum or theatre",
@@
 cycling_2005 <- cycling_2005 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 | 
-           destination == 4 | 
-           destination == 5 |
-           destination == 6 | 
-           destination == 7 | 
-           destination == 8 |
-           destination == 9 |
-           destination == 10 ) |>
+  filter(destination <= 21 ) |>
   mutate(dest_label =
            case_when(destination == 1 ~ "Home", 
-                     destination == 2 ~ "Work or school",
+                     destination %in% c(2,8) ~ "Work or school",
                      destination == 3 ~ "Other's home",
                      destination == 9 ~ "Outdoors",
-                     destination == 6 ~ "Grocery store, other stores or mall",
+                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                      destination == 10 ~ "Library, museum or theatre",
@@
 before returning home. Upon their return, they engaged in an activity
-labeled as `ACTCODE` = 11 for 465 minutes, which denotes working from home
-as their primary workplace. Following this work period, the person went
-out for another 15-minute cycling session, then returned home once more
-and spent 90 minutes watching television.
-
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+labeled as `ACTCODE` = 11 for 465 minutes, which denotes working from
+home as their primary workplace. Following this work period, the person
+went out for another 15-minute cycling session, then returned home once
+more and spent 90 minutes watching television.
+
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r}
@@
 ```
-As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data
+
+As we have already processed this year's data, we will clear the Episode
+and Main files for this year from the computer's memory before uploading
+any more data
 
@@
 walking_1998 <- walking_1998 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 ) |>
+  filter(origin <= 10 ) |>
   mutate(orig_label =
@@
                      origin == 2 ~ "Work or school",
-                     origin == 3 ~ "Other's home"),
+                     origin == 3 ~ "Other's home",
+                     origin == 4 ~ "Elsewhere",
+                     origin %in% c(5,6) ~ "Travel - Motorized",
+                     origin == 10 ~ "Travel - Other",
+                     origin == 8  ~ "Travel - Transit",
+                     origin == 7 ~ "Travel - Walk",
+                     origin == 9 ~ "Travel - Bycicle"),
          orig_label = factor(orig_label))
@@
 walking_1998 <- walking_1998 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 ) |>
+  filter(destination <= 10 ) |>
   mutate(dest_label =
@@
                      destination == 2 ~ "Work or school",
-                     destination == 3 ~ "Other's home"),
+                     destination == 3 ~ "Other's home",
+                     destination == 4 ~ "Elsewhere",
+                     destination %in% c(5,6) ~ "Travel - Motorized",
+                     destination == 10 ~ "Travel - Other",
+                     destination == 8  ~ "Travel - Transit",
+                     destination == 7 ~ "Travel - Walk",
+                     destination == 9 ~ "Travel - Bycicle"),
          dest_label = factor(dest_label))
@@
 
-In this case, the most common destination for a walking trip that starts at home are
-`other's Home`.
+In this case, the most common destination for a walking trip that starts
+at home are `other's Home`.
 
@@
 
-The person was at home doing food cleanup (`ACTCODE` = 110), then
-went for walking that lasted 30 minutes, came back home and spent 15 minutes relaxing/resting. Here have here another case of a recreational/leisure trip!
-
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+The person was at home doing food cleanup (`ACTCODE` = 110), then went
+for walking that lasted 30 minutes, came back home and spent 15 minutes
+relaxing/resting. Here have here another case of a recreational/leisure
+trip!
+
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r}
@@
 cycling_1998 <- cycling_1998 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 ) |>
+  filter(origin <= 10 ) |>
   mutate(orig_label =
@@
 cycling_1998 <- cycling_1998 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 ) |>
+  filter(destination <= 10 ) |>
   mutate(dest_label =
@@
 
-The person spent 165 minutes talking to someone (in person or on the phone, `ARTCODE` = 950), then embarked on a bike ride for 12 minutes.
-After that, the person spent 108 minutes shopping at home (`ARTCODE` = 380), cycled for another 12 minutes and came back to talk to someone for another 138 minutes (`ARTCODE` = 950).
-
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+The person spent 165 minutes talking to someone (in person or on the
+phone, `ARTCODE` = 950), then embarked on a bike ride for 12 minutes.
+After that, the person spent 108 minutes shopping at home (`ARTCODE` =
+380), cycled for another 12 minutes and came back to talk to someone for
+another 138 minutes (`ARTCODE` = 950).
+
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r}
@@
 walking_1992 <- walking_1992 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3) |>
+  filter(origin <= 10 ) |>
   mutate(orig_label =
@@
 walking_1992 <- walking_1992 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 ) |>
+  filter(destination <= 10 ) |>
   mutate(dest_label =
@@
 
-The person was at home reading newspapers (`ACTCODE` = 940), then went for walking that lasted 30 minutes, came back home and spent 90 minutes resting (`ACTCODE` = 470).[^10]
-
-[^10]: Note: ACTCODE is an activity code of the episode (What were you
+The person was at home reading newspapers (`ACTCODE` = 940), then went
+for walking that lasted 30 minutes, came back home and spent 90 minutes
+resting (`ACTCODE` = 470).[^9]
+
+[^9]: Note: ACTCODE is an activity code of the episode (What were you
     doing in minute).The list of main activity codes can be found in an
@@
                                ACTCODE == 303 ~ "Everyday shopping: Take-out food", # p. 188
+                               ACTCODE == 331 ~ "Financial Services (e.g. banking)", # p. 188
+                               ACTCODE == 332 ~ "Government Services (e.g. UIC)", # p. 188
                                ACTCODE == 370 ~ "Waiting for purchases or services", # p. 188
+                               ACTCODE == 380 ~ "Other Shopping and Services", # p. 188
                                ACTCODE == 390 ~ "Travel to/from shopping or obtaining services", # p. 188
+                               ACTCODE == 400 ~ "Washing, Dressing", # p. 202
+                               ACTCODE == 470 ~ "Relaxing, Thinking, Resting", # p. 202
+                               ACTCODE == 480 ~ "Other Personal Care or Private Activities", # p. 202
                                ACTCODE == 491 ~ "Travel to/from restaurant meals", # p. 202
                                ACTCODE == 492 ~ "Travel to/from other personal care activities", # p. 202
+                               ACTCODE == 550 ~ "Breaks/Waiting for Class", # p. 208
+                               ACTCODE == 580 ~ "Other Study", # p. 208
                                ACTCODE == 590 ~ "Travel to/from school education activities", # p. 208
                                ACTCODE == 610 ~ "Political, civic activity (e.g. voting, jury duty, city council, donating blood)", # p. 215
+                               ACTCODE == 640 ~ "Religious ServicesfPrayerfRead Bible", # p. 215
                                ACTCODE == 673 ~ "Unpaid babysitting", # p. 215
                                ACTCODE == 674 ~ "Transportation assistance to someone other than a household member", # p. 215
+                               ACTCODE == 678 ~ "Other unpaid work ", # p. 215
+                               ACTCODE == 680 ~ "Other Organizational, Voluntary and Religious Activity", # p. 215
                                ACTCODE == 691 ~ "Travel to/from civic and voluntary activity", # p. 216
                                ACTCODE == 692 ~ "Travel to/from religious services", # p. 216
+                               ACTCODE == 753 ~ "Socializing w/friends (non-private residence)", # p. 234
+                               ACTCODE == 780 ~ "Other Social Gatherings (Weddings, Wakes)", # p. 234
                                ACTCODE == 791 ~ "Travel to/from attending sports, movies or other entertainment events or visit sites", # p. 234
@@
                                ACTCODE == 793 ~ "Travel to/from other socializing (to bars, hospitals, weddings)", # p. 234
+                               ACTCODE == 814 ~ "Camping", # p. 244
                                ACTCODE == 816 ~ "Other outdoor activities/excursions (picnic, car rally, bird watching)", # p. 244
@@
                                ACTCODE == 894 ~ "Travel to/from other leisure activities", # p. 245
+                               ACTCODE == 950 ~ "Talking, Conversation, Phone", # p. 266
                                ACTCODE == 961 ~ "Reading personal mail (including flyers and advertisements)", # p. 266
@@
 
-Cases of episodes with `NA` weight values will be updated to value 1, meaning that since there's no weighting value for this observation, this episode only represent itself:
+Cases of episodes with `NA` weight values will be updated to value 1,
+meaning that since there's no weighting value for this observation, this
+episode only represent itself:
 
@@
 Save data in `R` format for later use:
+
 ```{r}
@@
 cycling_1992 <- cycling_1992 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3 ) |>
+  filter(origin <= 10 ) |>
   mutate(orig_label =
@@
 cycling_1992 <- cycling_1992 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 ) |>
+  filter(destination <= 10 ) |>
   mutate(dest_label =
@@
 
-The person was at home Talking or having conversation on the phone (`ACTCODE` = 950), then went for cycling (`ACTCODE` = 822) that lasted 60 minutes, came back home and spent 15 minutes washing and dressing (`ACTCODE` = 400).
-
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+The person was at home Talking or having conversation on the phone
+(`ACTCODE` = 950), then went for cycling (`ACTCODE` = 822) that lasted
+60 minutes, came back home and spent 15 minutes washing and dressing
+(`ACTCODE` = 400).
+
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r label-activity-codes-1992-cycling}
@@
                                ACTCODE == 873 ~ "Other pleasure drives (e.g. on a tour bus)", # p. 245
+                               ACTCODE == 880 ~ "Other Sport or Active Leisure", # p. 245
                                ACTCODE == 891 ~ "Travel to/from participation in active sport/outdoor activities", # p. 245
@@
 
-As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data
+As we have already processed this year's data, we will clear the Episode
+and Main files for this year from the computer's memory before uploading
+any more data
+
 ```{r}
@@
 walking_1986 <- walking_1986 |>
-  filter(origin == 1 | 
-           origin == 2 | 
-           origin == 3) |>
+  filter(origin < 8 ) |>
   mutate(orig_label =
@@
                      origin == 2 ~ "Work or school",
-                     origin == 3 ~ "Other's home"),
+                     origin == 3 ~ "Elsewhere",
+                     origin == 4 ~ "Travel - Motorized",
+                     origin == 7 ~ "Travel - Other",
+                     origin == 6 ~ "Travel - Transit",
+                     origin == 5 ~ "Travel - Walk"),
          orig_label = factor(orig_label))
@@
 walking_1986 <- walking_1986 |>
-  filter(destination == 1 | 
-           destination == 2 | 
-           destination == 3 ) |>
+  filter(destination < 8 ) |>
   mutate(dest_label =
@@
                      destination == 2 ~ "Work or school",
-                     destination == 3 ~ "Other's home"),
+                     destination == 3 ~ "Elsewhere",
+                     destination == 4 ~ "Travel - Motorized",
+                     destination == 7 ~ "Travel - Other",
+                     destination == 6 ~ "Travel - Transit",
+                     destination == 5 ~ "Travel - Walk"),
          dest_label = factor(dest_label))
@@
 
-**IMPORTANT:** See p.
-A96 of the User's Guide for Cycle 2.
-The `WGHT_EPI` values, which give the number of people a record represents, (in this case the `FWGT_MS`) are given as a 9 digit integer with 4 decimals, like so: XXXXX.XXXX.
-We cannot just copy `FWGT_MS` to `WGHT_EPI` below, we must divide first by 10,000:
+**IMPORTANT:** See p. A96 of the User's Guide for Cycle 2. The
+`WGHT_EPI` values, which give the number of people a record represents,
+(in this case the `FWGT_MS`) are given as a 9 digit integer with 4
+decimals, like so: XXXXX.XXXX. We cannot just copy `FWGT_MS` to
+`WGHT_EPI` below, we must divide first by 10,000:
 
@@
 
-The person was at home listening radio for 30 minutes (`ARTCODE` = 90), then went for a walk for gardening/pet care (`ACT_CODE` = 17) that lasted 15 minutes, came back home and spent 90 minutes resting/relaxing (`ARTCODE` = 47).
+The person was at home listening radio for 30 minutes (`ARTCODE` = 90),
+then went for a walk for gardening/pet care (`ACT_CODE` = 17) that
+lasted 15 minutes, came back home and spent 90 minutes resting/relaxing
+(`ARTCODE` = 47).
 
@@
 
-Label the activity (see the activity codes in the 2010 User's Guide) and drop the LOCATION because it is already coded as MODE:
+Label the activity (see the activity codes in the 2010 User's Guide) and
+drop the LOCATION because it is already coded as MODE:
+
 ```{r label-activity-codes-1986-walking}
@@
                                ACTCODE == 3 ~ "Travel during work", # p. 156
+                               ACTCODE == 4 ~ "Waiting/Delays at Work", # p. 156
                                ACTCODE == 5 ~ "Meals/snacks at work", # p. 156
@@
                                ACTCODE == 12 ~ "Indoor cleaning", # p. 164
+                               ACTCODE == 14 ~ "Laundry, Ironing, Folding", # p. 164
                                ACTCODE == 13 ~ "Outdoor cleaning (garbage, snow removal, garage)", # p. 164
@@
                                ACTCODE == 23 ~ "Reading/talking/conversation with child", # p. 177
+                               ACTCODE == 24 ~ "Play with Children", # p. 177
                                ACTCODE == 28 ~ "Other child care (unpaid babysitting)", # p. 177
@@
                                ACTCODE == 30 ~ "Everyday Shopping", # p. 188
+                               ACTCODE == 32 ~ "Personal Care Services", # p. 188
                                ACTCODE == 33 ~ "Government and financial services", # p. 188
@@
                                ACTCODE == 93 ~ "Reading books, magazines", # p. 266
+                               ACTCODE == 94 ~ "Reading Newspapers", # p. 266
                                ACTCODE == 95 ~ "Talking, conversation with household member only (face-to-face)", # p. 266
@@
 Now, we'll select all the rows with walking episodes in the GSS survey,
-selecting the columns that we are interested[^5], and already creating
+selecting the columns that we are interested[^10], and already creating
 the origin and destination variables:
 
-[^5]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
+[^10]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
     activity code. We'll change the name of this column to be match the
@@
 walking_2022_spss <- walking_2022_spss |>
-  filter(origin == 'At home or on property' | 
-           origin == 'At place of work or school' | 
-           origin == 'Grocery store, other stores or mall'| 
-           origin == 'In the neighbourhood'| 
-           origin == 'Sports centre, field or arena'| 
-           origin == 'Outdoors'| 
-           origin == 'Restaurant, bar or club'| 
-           origin == 'At someone else\'s home or property'| 
-           origin == 'Medical, dental or other health clinic'| 
-           origin == 'Library, museum or theatre'| 
-           origin == 'Place of worship' |
-           origin == 'Away at business') |>
+  filter(!is.na(origin)) |>
     mutate(orig_label =
@@
                      origin == 'At someone else\'s home or property' ~ "Other's home",
+                     origin %in% c('Travel - Car, truck or van,  - as passenger', 
+                                   'Travel - Car, truck or van,  - as driver',
+                                   'Travel - Motorcycle, scooter or moped',
+                                   'Travel - Taxi, Limousine service') ~ "Travel - Motorized", 
+                     origin %in% c('Travel - Airplane', 
+                                   'Travel - Other',
+                                   'Travel - not stated') ~ "Travel - Other", 
+                     origin == 'Travel - Public transit (bus, streetcar, subway, light-rail)' ~ "Travel - Transit",
                      TRUE ~ origin),
@@
 walking_2022_spss <- walking_2022_spss |>
-  filter(destination == 'At home or on property' | 
-           destination == 'At place of work or school' | 
-           destination == 'Grocery store, other stores or mall'| 
-           destination == 'In the neighbourhood'| 
-           destination == 'Sports centre, field or arena'| 
-           destination == 'Outdoors'| 
-           destination == 'Restaurant, bar or club'| 
-           destination == 'At someone else\'s home or property'| 
-           destination == 'Medical, dental or other health clinic'| 
-           destination == 'Library, museum or theatre'| 
-           destination == 'Place of worship' |
-           destination == 'Away at business') |>
+  filter(!is.na(destination)) |>
     mutate(dest_label =
@@
                      destination == 'At someone else\'s home or property' ~ "Other's home",
+                     destination %in% c('Travel - Car, truck or van,  - as passenger', 
+                                   'Travel - Car, truck or van,  - as driver',
+                                   'Travel - Motorcycle, scooter or moped',
+                                   'Travel - Taxi, Limousine service') ~ "Travel - Motorized", 
+                     origin %in% c('Travel - Airplane', 
+                                   'Travel - Other',
+                                   'Travel - not stated') ~ "Travel - Other", 
+                     destination == 'Travel - Public transit (bus, streetcar, subway, light-rail)' ~ "Travel - Transit",
                      TRUE ~ destination),
@@
 
-The person was at home doing regular household tasks, then went for
-walk that lasted 20 minutes, came back home and returned to the household tasks for more 180 minutes.
+The person was at home doing regular household tasks, then went for walk
+that lasted 20 minutes, came back home and returned to the household
+tasks for more 180 minutes.
 
@@
 Now, we'll select all the rows with cycling episodes in the GSS survey,
-selecting the columns that we are interested[^5], and already creating
+selecting the columns that we are interested[^11], and already creating
 the origin and destination variables:
 
-[^5]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
+[^11]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
     activity code. We'll change the name of this column to be match the
@@
 cycling_2022_spss <- cycling_2022_spss |>
-  filter(origin == 'At home or on property' | 
-           origin == 'At place of work or school' | 
-           origin == 'Grocery store, other stores or mall'| 
-           origin == 'In the neighbourhood'| 
-           origin == 'Sports centre, field or arena'| 
-           origin == 'Outdoors'| 
-           origin == 'Restaurant, bar or club'| 
-           origin == 'At someone else\'s home or property'| 
-           origin == 'Medical, dental or other health clinic'| 
-           origin == 'Library, museum or theatre'| 
-           origin == 'Place of worship' |
-           origin == 'Away at business') |>
+  filter(!is.na(origin)) |>
     mutate(orig_label =
@@
 cycling_2022_spss <- cycling_2022_spss |>
-  filter(destination == 'At home or on property' | 
-           destination == 'At place of work or school' | 
-           destination == 'Grocery store, other stores or mall'| 
-           destination == 'In the neighbourhood'| 
-           destination == 'Sports centre, field or arena'| 
-           destination == 'Outdoors'| 
-           destination == 'Restaurant, bar or club'| 
-           destination == 'At someone else\'s home or property'| 
-           destination == 'Medical, dental or other health clinic'| 
-           destination == 'Library, museum or theatre'| 
-           destination == 'Place of worship' |
-           destination == 'Away at business') |>
+  filter(!is.na(destination)) |>
     mutate(dest_label =
@@
 
-The table above shows that the most common destination for a cycling trip
-that starts at home are `Home` and `Work or school`. It also shows that the second most common cycling trip is from `Grocery store, other stores or mall` to `Home`. Notice, though,
-that quite a few trips start *and* end at Home. What are those supposed
-to be?
+The table above shows that the most common destination for a cycling
+trip that starts at home are `Home` and `Work or school`. It also shows
+that the second most common cycling trip is from
+`Grocery store, other stores or mall` to `Home`. Notice, though, that
+quite a few trips start *and* end at Home. What are those supposed to
+be?
 
@@
 
-The person was at home doing `Culture, sports events, hobbies, leisure or outdoor activity`, then went for cycling trip that lasted 60 minutes, came back home and spent 120 minutes
-`reading, whatching TV or listening music`. So, this was a recreational/leisure
-trip!
+The person was at home doing
+`Culture, sports events, hobbies, leisure or outdoor activity`, then
+went for cycling trip that lasted 60 minutes, came back home and spent
+120 minutes `reading, whatching TV or listening music`. So, this was a
+recreational/leisure trip!
 

### 2025-11-07 README.Rmd
Final version for EPB publication
Updating dataset to remove duplicated instances. Updating documentation to set new version.

--- 2024-11-13 README.Rmd
@@
 
-# ActiveCA
+# ActiveCA: Time use data from the general social survey of Canada to study active travel
 
-{ActiveCA} is an open data product in the form of
-an R data package, developed after the collection, cleaning, and
-processing of [General Social Survey
-(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) data ranging from 1986 to 2015. The main objective of the {ActiveCA} package is to identify the most
-appropriate impedance functions for active transportation modes for
-various destinations and time periods in Canada.
-The package also provides analysis-ready data on active transportation episodes, focusing on walking and cycling activities, with information on trip origins, destinations, and duration.
+<!-- badges: start -->
+<!-- badges: end -->
+
+-    Bruno Santos, School of Earth, Environment and Society, McMaster University
+-    Mahdis Mogadhasi, School of Earth, Environment and Society, McMaster University
+-    Antonio Paez, School of Earth, Environment and Society, McMaster University
+
+This paper describes {ActiveCA}, an open data product with Canadian time use data.
+{ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) from the General Social Survey (GSS).
+Active travel episodes are characterized by mode, with walking being part of every cycle and bicycling starting in 1992.
+The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates.
+Based on the year of the survey, a variety of locations are coded.
+In earlier cycles, these include home, work or school, and other’s home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others.
+The geographical resolution includes the province and whether the episode was in an urban or rural setting.
+
+https://doi-org.libaccess.lib.mcmaster.ca/10.1177/23998083251374724
 
@@
 
-You can install the development version of ActiveCA from [GitHub](https://github.com/) with:
+You can install the development version of ActiveCA from [GitHub](https://github.com/dias-bruno/ActiveCA) with:
+
 ``` r

### 2025-11-07 data-raw/01-preprocessing-episode-files.Rmd
Updating dataset to remove duplicated instances. Updating documentation to set new version.

@@
 
-| **Code** | **Location** | **Year** | **Description** |
-|----------------|----------------|----------------|--------------------------|
-| 300 | At home or on property | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 301 | At place of work or school | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 302 | Away on business | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 303 | At someone else's home or property | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 304 | In the neighbourhood | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 305 | Outdoors | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 306 | Grocery store, other stores or mall | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 307 | Library, museum or theatre | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 308 | Sports centre, field or arena | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 309 | Restaurant, bar or club | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 310 | Place of worship | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 311 | Medical, dental or other health clinic | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 312 | Elsewhere | 2015 | Walking (code = 315) and Cycling (code = 318) to this location |
-| 1 | Respondent's home | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 2 | Work place | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 3 | Someone else's home | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 4 | Restaurant/bar | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 5 | Place of worship | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 6 | Grocery store | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 7 | Other store/Mall | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 8 | School | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 9 | Outdoors away from home | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 10 | Library | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 11 | Other place | 2010 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 1 | Respondent's home | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 2 | Work place | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 3 | Someone else's home | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 4 | Restaurant/bar | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 5 | Place of worship | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 6 | Grocery store | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 7 | Other store/Mall | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 8 | School | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 9 | Outdoors away from home | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 10 | Library | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 11 | Other place | 2005 | Walking (code = 14) and Cycling (code = 17) to this location |
-| 1 | Respondent's home | 1998 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 2 | Work place | 1998 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 3 | Someone else's home | 1998 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 4 | Other plourhood | 1998 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 1 | Respondent's home | 1992 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 2 | Work place | 1992 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 3 | Someone else's home | 1992 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 4 | Other place | 1992 | Walking (code = 7) and Cycling (code = 9) to this location |
-| 1 | Respondent's home | 1886 | Walking (code = 5) to this location |
-| 2 | Work place | 1886 | Walking (code = 5) to this location |
-| 3 | At other places | 1886 | Walking (code = 5) to this location |
-| 7 | other | 1886 | Walking (code = 5) to this location |
+| Origin/Destination                     | LOCATION (2022)        | LOCATION (2015) | PLACE (2010) | PLACE (2005) | PLACE (1998) | PLACE (1998) | LOCATION (1986) |
+| -------------------------------------- | ---------------------- | --------------- | ------------ | ------------ | ------------ | ------------ | --------------- |
+| Home                                   | 3300                   | 300             | 1            | 1            | 1            | 1            | 1               |
+| Work or school                         | 3301                   | 301             | 2, 8         | 2, 8         | 2            | 2            | 2               |
+| Business                               | 3302                   | 302             |              |              |              |              |                 |
+| Other's home                           | 3303                   | 303             | 3            | 3            | 3            | 3            |                 |
+| In the neighbourhood                   | 3304                   | 304             |              |              |              |              |                 |
+| Outdoors                               | 3305                   | 305             | 9            | 9            |              |              |                 |
+| Grocery store, other stores or mall    | 3306                   | 306             | 6, 7         | 6, 7         |              |              |                 |
+| Library, museum or theatre             | 3307                   | 307             | 10           | 10           |              |              |                 |
+| Sports centre, field or arena          | 3308                   | 308             |              |              |              |              |                 |
+| Restaurant, bar or club                | 3309                   | 309             | 4            | 4            |              |              |                 |
+| Place of worship                       | 3310                   | 310             | 5            | 5            |              |              |                 |
+| Medical, dental or other health clinic | 3311                   | 311             |              |              |              |              |                 |
+| Elsewhere                              | 3312                   | 312             | 11           | 11           | 4            | 4            | 3               |
+| Travel - Motorized                     | 3313, 3314, 3319, 3320 | 313, 314, 319   | 12, 13, 19   | 12, 13, 19   | 5, 6         | 5, 6         | 4               |
+| Travel - Public transit                | 3316                   | 316             | 15, 16       | 15, 16       | 8            | 8            | 6               |
+| Travel - Walk                          | 3315                   | 315             | 14           | 14           | 7            | 7            | 5               |
+| Travel - Bicycle                       | 3318                   | 318             | 17           | 17           | 9            | 9            |                 |
+| Travel - Other                         |  3317, 3323, 3399      | 317, 320, 321   | 18, 20, 21   | 18, 20, 21   | 10           | 10           | 7               |
 
@@
 walking_2015 <- walking_2015 |>
-  filter(origin <= 321) |>
+  filter(!origin %in% c(315, 318, 317,320,321)) |>
   mutate(orig_label =
@@
                      origin %in% c(313,314,319) ~ "Travel - Motorized",
-                     origin %in% c(317,320,321) ~ "Travel - Other",
-                     origin == 316 ~ "Travel - Transit",
-                     origin == 315 ~ "Travel - Walk",
-                     origin == 318 ~ "Travel - Bycicle"),
-         orig_label = factor(orig_label))
+                     origin == 316 ~ "Travel - Transit"),
+         orig_label = factor(orig_label)) %>%   filter(!is.na(orig_label))
 ```
@@
 walking_2015 <- walking_2015 |>  
-  filter(destination <= 321) |>
+  filter(!destination %in% c(315, 318, 317,320,321)) |>
   mutate(dest_label =
@@
                      destination %in% c(313,314,319) ~ "Travel - Motorized",
-                     destination %in% c(317,320,321) ~ "Travel - Other",
-                     destination == 316 ~ "Travel - Transit",
-                     destination == 315 ~ "Travel - Walk",
-                     destination == 318 ~ "Travel - Bycicle"),
-         dest_label = factor(dest_label))
+                     destination == 316 ~ "Travel - Transit"),
+         dest_label = factor(dest_label)) %>%   filter(!is.na(dest_label))
 ```
@@
   mutate(act_label = "Transport to or from activity") |>
-  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province) %>% 
+  distinct()
 ```
@@
 cycling_2015 <- cycling_2015 |>
-  filter(origin <= 321) |>
+  filter(!origin %in% c(315, 318, 317,320,321)) |>
   mutate(orig_label =
@@
 cycling_2015 <- cycling_2015 |>  
-  filter(destination <= 321) |>
+  filter(!destination %in% c(315, 318, 317,320,321)) |>
   mutate(dest_label =
@@
   mutate(act_label = "Transport to or from activity") |>
-  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province) %>%    distinct()
 ```
@@
 walking_2010 <- walking_2010 |>
-  filter(origin <= 21 ) |>
+  filter(!origin %in% c(18, 20, 21, 14, 17)) |>
   mutate(orig_label =
@@
                      origin %in% c(12,13,19) ~ "Travel - Motorized",
-                     origin %in% c(18,20,21) ~ "Travel - Other",
-                     origin %in% c(15,16) ~ "Travel - Transit",
-                     origin == 14 ~ "Travel - Walk",
-                     origin == 17 ~ "Travel - Bycicle"),
-         orig_label = factor(orig_label))
+                     origin %in% c(15,16) ~ "Travel - Transit"),
+         orig_label = factor(orig_label)) %>%   filter(!is.na(orig_label))
 ```
@@
 walking_2010 <- walking_2010 |>
-  filter(destination <= 21 ) |>
+  filter(!destination %in% c(18, 20, 21, 14, 17)) |>
   mutate(dest_label =
@@
                      destination %in% c(12,13,19) ~ "Travel - Motorized",
-                     destination %in% c(18,20,21) ~ "Travel - Other",
-                     destination %in% c(15,16) ~ "Travel - Transit",
-                     destination == 14 ~ "Travel - Walk",
-                     destination == 17 ~ "Travel - Bycicle"),
-         dest_label = factor(dest_label))
+                     destination %in% c(15,16) ~ "Travel - Transit"),
+         dest_label = factor(dest_label)) %>%   filter(!is.na(dest_label))
 ```
@@
                                ACTCODE == 990 ~ "Travel to/from media or communication activities")) |> # p. 266 
-  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province) %>%    distinct()
 ```
@@
 cycling_2010 <- cycling_2010 |>
-  filter(origin <= 21 ) |>
+  filter(!origin %in% c(18, 20, 21, 14, 17)) |>
   mutate(orig_label =
@@
 cycling_2010 <- cycling_2010 |>
-  filter(destination <= 21 ) |>
+  filter(!destination %in% c(18, 20, 21, 14, 17)) |>
   mutate(dest_label =
@@
 walking_2005 <- walking_2005 |>
-  filter(origin <= 21 ) |>
+  filter(!origin %in% c(18, 20, 21, 14, 17)) |>
   mutate(orig_label =
@@
 walking_2005 <- walking_2005 |>
-  filter(destination <= 21 ) |>
+  filter(!destination %in% c(18, 20, 21, 14, 17)) |>
   mutate(dest_label =
@@
 cycling_2005 <- cycling_2005 |>
-  filter(origin <= 21 ) |>
+  filter(!origin %in% c(18, 20, 21, 14, 17)) |>
   mutate(orig_label =
@@
 cycling_2005 <- cycling_2005 |>
-  filter(destination <= 21 ) |>
+  filter(!destination %in% c(18, 20, 21, 14, 17)) |>
   mutate(dest_label =
@@
 walking_1998 <- walking_1998 |>
-  filter(origin <= 10 ) |>
+  filter(!origin %in% c(7, 9, 10)) |>
   mutate(orig_label =
@@
                      origin %in% c(5,6) ~ "Travel - Motorized",
-                     origin == 10 ~ "Travel - Other",
-                     origin == 8  ~ "Travel - Transit",
-                     origin == 7 ~ "Travel - Walk",
-                     origin == 9 ~ "Travel - Bycicle"),
-         orig_label = factor(orig_label))
+                     origin == 8  ~ "Travel - Transit"),
+         orig_label = factor(orig_label)) %>%   filter(!is.na(orig_label))
 ```
@@
 walking_1998 <- walking_1998 |>
-  filter(destination <= 10 ) |>
+  filter(!destination %in% c(7, 9, 10)) |>
   mutate(dest_label =
@@
                      destination %in% c(5,6) ~ "Travel - Motorized",
-                     destination == 10 ~ "Travel - Other",
-                     destination == 8  ~ "Travel - Transit",
-                     destination == 7 ~ "Travel - Walk",
-                     destination == 9 ~ "Travel - Bycicle"),
-         dest_label = factor(dest_label))
+                     destination == 8  ~ "Travel - Transit"),
+         dest_label = factor(dest_label)) %>%   filter(!is.na(dest_label))
 ```
@@
 cycling_1998 <- cycling_1998 |>
-  filter(origin <= 10 ) |>
+  filter(!origin %in% c(7, 9, 10)) |>
   mutate(orig_label =
@@
 cycling_1998 <- cycling_1998 |>
-  filter(destination <= 10 ) |>
+  filter(!destination %in% c(7, 9, 10)) |>
   mutate(dest_label =
@@
 walking_1992 <- walking_1992 |>
-  filter(origin <= 10 ) |>
+  filter(!origin %in% c(7, 9, 10)) |>
   mutate(orig_label =
@@
 walking_1992 <- walking_1992 |>
-  filter(destination <= 10 ) |>
+  filter(!destination %in% c(7, 9, 10)) |>
   mutate(dest_label =
@@
 cycling_1992 <- cycling_1992 |>
-  filter(origin <= 10 ) |>
+  filter(!origin %in% c(7, 9, 10)) |>
   mutate(orig_label =
@@
 cycling_1992 <- cycling_1992 |>
-  filter(destination <= 10 ) |>
+  filter(!destination %in% c(7, 9, 10)) |>
   mutate(dest_label =
@@
 walking_1986 <- walking_1986 |>
-  filter(origin < 8 ) |>
+  filter(!origin %in% c(5,7)) |>
   mutate(orig_label =
@@
                      origin == 4 ~ "Travel - Motorized",
-                     origin == 7 ~ "Travel - Other",
-                     origin == 6 ~ "Travel - Transit",
-                     origin == 5 ~ "Travel - Walk"),
-         orig_label = factor(orig_label))
+                     origin == 6 ~ "Travel - Transit"),
+         orig_label = factor(orig_label)) %>%   filter(!is.na(orig_label))
 ```
@@
 walking_1986 <- walking_1986 |>
-  filter(destination < 8 ) |>
+  filter(!origin %in% c(5,7)) |>
   mutate(dest_label =
@@
                      destination == 4 ~ "Travel - Motorized",
-                     destination == 7 ~ "Travel - Other",
-                     destination == 6 ~ "Travel - Transit",
-                     destination == 5 ~ "Travel - Walk"),
-         dest_label = factor(dest_label))
+                     destination == 6 ~ "Travel - Transit"),
+         dest_label = factor(dest_label)) %>%   filter(!is.na(dest_label))
 ```
@@
                                )) |> 
-  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
+  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province) %>%    distinct()
 ```
@@
   filter(!is.na(origin)) |>
+  filter(!origin %in% c('Travel - Airplane', 
+                                   'Travel - Other',
+                                   'Travel - not stated', 
+                        'Travel - Walk', 'Travel - Bicycle')) %>% 
   mutate(orig_label =
@@
                                    'Travel - Taxi, Limousine service') ~ "Travel - Motorized", 
-                     origin %in% c('Travel - Airplane', 
-                                   'Travel - Other',
-                                   'Travel - not stated') ~ "Travel - Other", 
                      origin == 'Travel - Public transit (bus, streetcar, subway, light-rail)' ~ "Travel - Transit",
                      TRUE ~ origin),
-         orig_label = factor(orig_label))
+         orig_label = factor(orig_label)) %>%   filter(!is.na(orig_label))
 ```
@@
   filter(!is.na(destination)) |>
+  filter(!destination %in% c('Travel - Airplane', 
+                                   'Travel - Other',
+                                   'Travel - not stated', 
+                        'Travel - Walk', 'Travel - Bicycle')) %>% 
     mutate(dest_label =
@@
                                    'Travel - Taxi, Limousine service') ~ "Travel - Motorized", 
-                     origin %in% c('Travel - Airplane', 
-                                   'Travel - Other',
-                                   'Travel - not stated') ~ "Travel - Other", 
                      destination == 'Travel - Public transit (bus, streetcar, subway, light-rail)' ~ "Travel - Transit",
                      TRUE ~ destination),
-         dest_label = factor(dest_label))
+         dest_label = factor(dest_label)) %>%   filter(!is.na(dest_label))
 ```
